### 2013-09-27 WriteUp/conj-prior.tex
Added current main write-up files

%%%
%%% Conjugate Priors
%%%

\section{Conjugate Priors}

In Sindhu's notes it's assumed that the agent has a belief $\q$, and moves the market state from initial state $\q_{init}$ to a convex combination $\tilde{\q} = \lambda \q + (1-\lambda)\q_{init}$. This note tries to justify this based on Bayesian updating, rather than budget limitations. (But it doesn't quite succeed.)

\medskip\noindent
Let $p(x;\theta)$ denote a probability density drawn from an exponential family with sufficient statistic $\phi : \mX \rightarrow \bR^d$, where $\theta$ is the natural parameter:
$$
p(x;\theta) = \exp \left[ \la \theta, \phi(x) \ra \right],
$$
and
$$
g(\theta) = \log \int_{\mX} \exp\la\phi(x),\theta\ra d\!x.
$$
Recall that $\grad g(\theta) = \Exp[\phi(x)]$ and $\grad^2 g(\theta) = \Var[\phi(x)]$. The family of conjugate priors is also an exponential family and takes the form
$$
p(\theta; n,\nu) = \exp\left[ \la n\nu,\theta \ra - ng(\theta) - h(\nu,n) \right].
$$
Here the feature map is $\psi(\theta) = (\theta, -g(\theta))$, the natural parameter is $(n\nu, n)$ where $n \in \bR$ and $\nu \in \bR^d$. The normalizer $h(\nu, n)$ is convex in $(n\nu,n)$. It is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$. The justification for this is that
$$
\Exp_{\theta} \left[ \Exp_x \left[ \phi(x) | \theta \right]\right] 
= \Exp_{\theta} \left[\grad g(\theta)\right]
= \nu.
$$
(The proof is straightforward but not obvious. I have a reference for this but it's impossible to read---it's a mathematical statistics paper.)

Suppose we draw a sample $X = (x_1,\ldots,x_m)$ of size $m$, and denote the empirical mean by $\mu[X] = \sum_{i=1}^m \phi(x_i)$. The posterior distribution is then
$$
p(\theta|X) \propto p(X|\theta)p(\theta|n,\nu) \propto \exp\left[ \la\mu[X]+n\nu,\theta\ra - (m+n)g(\theta) \right],
$$
and so the posterior mean is 
\begin{equation} \label{posterior-mean}
\frac{m\mu[X] + n\nu}{m+n}.
\end{equation}
Thus the posterior mean is a convex combination of the prior and posterior means, and their relative weights depend on the phantom and empirical sample sizes.

In the context of prediction markets, one could imagine an agent who uses the current market estimate as its prior, and draws a empirical sample of size $m$. Its belief then takes the form~(\ref{posterior-mean}), where $n$ depends on the importance the agent places on the market estimate (perhaps based on how long the market has been running). However, note that the \emph{mean parameter} becomes a convex combination of market state and empirical belief, and this does not translate to the \emph{natural parameter}, which is what we would have liked. The new natural parameter is
$$
\grad g^{-1} \left(\frac{m\mu[X] + n\nu}{m+n}\right) = \grad g^* \left(\frac{m\mu[X] + n\nu}{m+n}\right).
$$
where $g^*$ is the convex conjugate of $g$.

### 2013-09-27 WriteUp/mm.tex
Added current main write-up files

\documentclass[11pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}

\usepackage{parskip}

\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\mX}{\mathcal{X}}
\newcommand{\bR}{\mathbf{R}}
\newcommand{\grad}{\nabla}
\newcommand{\Exp}{\mathbf{E}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\q}{\mathbf{q}}
\newcommand{\mM}{\ensuremath{\mathcal{M}}}
\newcommand{\hmu}{\ensuremath{\hat{\mu}}}
\newcommand{\mP}{\ensuremath{\mathcal{P}}}
\newcommand{\mY}{\ensuremath{\mathcal{Y}}}
\newcommand{\mS}{\ensuremath{\mathcal{S}}}
\newcommand{\ds}{\displaystyle}
\newcommand{\mF}{\mathcal{F}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\htheta}{\hat{\theta}}

\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\newcounter{ExampleCount}
\setcounter{ExampleCount}{1}
\newenvironment{example}
{% This is the begin code
\medskip\noindent
{\bf Example \arabic{ExampleCount}.}
\stepcounter{ExampleCount} 
}
{% This is the end code
\medskip
}

\title{Maximum Entropy, Scoring Rules, and Prediction Markets}

\begin{document}
\maketitle

%%%
%%%
%%%

\section{Scoring Rule}

* Discrete or continuous outcome space.

* Want to elicit statistics of the distribution (not full distribution).

* Single-agent setting: scoring rules for expectations of random variables.

* Can also elicit one-to-one-functions of those expectations (e.g., mean and variance from first and second uncentered moments). Need to understand possible technical subtleties: see ``functional invariance'' property of maximum likelihood estimate.

* Strictly proper scoring rule implicitly defined via solution of a maximum entropy problem.

* Take log of maxent distribution with mean parameters given by the agent. Generalizes the usual log scoring rule for complete probability distributions.

* Although it's a generalization of the log scoring rule, may not look anything like log scoring rule: e.g.\ normal, exponential, beta distributions.

* Informational advantage: the agent doesn't need to know full probability density to be incentivized to reveal the statistics truthfully. Only needs to know the statistics.

* Does not have to agree that underlying density if from an exponential family.

* Technical subtlety (need to understand better): agent must agree on the \emph{support} of the distribution. This is related to the ``base measure'' of the exponential family.

* Important aspect of this approach: the primitive is now the \emph{uncertainty function} (e.g., negative entropy) defined over the space of probability densities.

* Other uncertainty functions can be used, they just have to be convex (is that all?). But the point is that they're defined over the space of densities, and then the convex function over the space of statistics is derived from that primitive via an optimization problem.

* It would be nice to show how many properties of the market follow from properties of the uncertainty function used.

* Do all cost functions arise in this way? From an optimization problem having objective function an uncertainty function over probability densities? This is straightforward for proper scoring rules but not as much for \emph{strictly} proper scoring rules.

* Connection between log scoring rule and maximum likelihood. This can give scoring rules for other kinds of parameters besides expectations? (Bounds on the support for the uniform distribution, for example). But it assumes that the agent's belief distribution takes a specific form (e.g., exponential, uniform, etc.) The maxent approach works for statistics of any underlying distribution, but is limited to expectations.

* Can we broaden the maxent scoring rule approach beyond expectations? Can we develop a scoring rule for the median with this approach?

* More generally, rather than just the maximum likelihood estimator, we can look at $m$-estimators to construct these kinds of scoring rules.

* Need to understand the consistency proofs of these kinds of estimators well.

%%%
%%%
%%%

\section{Prediction Market}

* Re-interpretation of the exponential family elements: sufficient statistic becomes payoff function, natural parameter is share vector, normalizer becomes cost function.

* Go over desirable properties of a price/cost function: monotonicity, path independence, etc.

* Important technical detail when looking at statistics: the domain of the cost function. For instance with first and second uncentered moments as statistics, can't have negative oustanding shares for the second security. How odd is this for practice?

* Insight: LMSR prediction market is estimating the natural parameter of an exponential family.

* Sequence of traders seeking to maximize profit can be viewed as MLE estimation of natural parameters.

* With budgets, adversarial traders can cause limited damage (in what formal sense again?)

* Budgets of informative traders grow in expectation over time

* This assumes a market that occurs over rounds, i.e., there is not just a single ultimate outcome (like a Presidential election) but several outcomes over time so that securities pay off.

* Need to think more about the model of bidder behavior: Bayesian updating? Budget constrained? Risk neutral or risk averse?

* Connection to mirror descent: the update with Bayesian traders is the mirror descent update. Frongillo et al. also observed this but in their model they have budget-constrained bidders with log utility, so there are some differences. Need to understand this better.

* What is the advantage of the mirror descent perspective/connection? Can we talk about the regret of the market and its convergence rate with Bayesian traders? How does it compare to the optimal rate (usually something like $\sqrt{T}$)?

* Note: The kind of aggregation performed by Bayesian update bidders is also very reminiscent of \emph{exponential smoothing} (not related to ``exponential'' families, at least not directly as far as I'm aware). It's used for forecasting from time series data and has some nice properties.

* Many of the results for prediction market bidders currently assume that they believe the underlying distribution to be an exponential family. It would be nice to do away with that as much as possible to try to match the mild informational requirements from the scoring rule section.

%%%
%%%
%%%

\section{Duality}

* The duality between scoring and market scoring rules (i.e., cost function prediction markets) is understood from Hanson's work.

* Exponential families perspective make the duality in even more concrete: share vector and agent belief live in dual spaces. Scoring rules work on mean parametrization and prediction markets work on natural parametrization.

%%%
%%%
%%%

\appendix

\include{maxent-market}
\include{sindhu-outline}
\include{conj-prior}
\include{maxlike-scoring}

\end{document}

### 2013-10-01 WriteUp/mm.tex
Added bullets related to Sindhu's write-up

@@
 
+* Show how to recover Savage representation of scoring rule from the primitives.
+Cost function is composition of adjoint of expectation map (linear map) and conjugate of the information function.
+Savage convex function is then the conjugate of the cost function.
+
 %%%
@@
 
+* Share vector is maximum likelihood estimate of the natural parameter of distribution.
+
+* Adversarial and strategic traders.
+
+* Budgeted prediction market suffers limited damage under concerted attacks by adversarial traders.
+
+* Informative trader will eventually be unconstrained and will be able to carry out unrestricted trade.
+
+* Arbitrage-free property: cost of making a ``sure bet'' is exactly the payoff of that bet.
+Generalizes the fact that sum of all prices must equal 1 when there is a security for each outcome (in discrete case).
+
+* Do we really want to bound sufficient statistics?
+
+* Interpretation function is nice, what is the intuition there?
+
+* Do we want to assume that the sufficient statistics are \emph{minimal}?
+What are advantages and disadvantages of this from a market standpoint?
+By definition, there are no arbitrage opportunities with minimal statistics, independent of the cost function.
+(There are no sure bets.)
+
+* If natural parameter space and sufficient statistics are bounded, then market has bounded loss. Are there weaker conditions?
+
+* Need to justify how traders choose to move share vector (equivalently, prices) given empirical beliefs. Several elements: empirical belief, market prior, budget, risk sensitivity.
+
+* The harm on the log loss of the market prediction that can be caused by adversarial traders is at most their combined budgets.
+It should be straightforward to generalize this to other losses (market uses a different cost function than LMSR).
+
+* Previous result is also interesting because it means log loss and budget should have the same units.
+
+* For Theorem 4, perhaps there is cleaner form in terms of just the natural/mean parameters and not the entire distributions.
+Could look at convex function defined on parameters via maximum entropy approach.
+
+* Perhaps introduce equivalence of uncertainty (entropy) function, Bregman divergence, and scoring rule from Dawid work.
+
+* Net profit of a trader equals the reduction in KL divergence with respect to the true distribution.
+
+* Can 5.4 proof work if trader has beliefs just over statistics and not entire distribution?
+
+* Theorem 5 makes no assumptions about optimality of belief updating procedure.
+Can we say something with respect to true distribution if the traders are doing correct Bayesian updates?
+What exactly does it mean for belief formation to be optimal?
+
+* Still don't understand the attack described in the conclusion.
+
+* Would be very nice for weak informational requirements of scoring rule setting could be carried over to prediction market setting (so don't have to assume beliefs are drawn from an exponential family).
+
+* Can something be said about the form of the agent utilities and how it interacts with cost function of the market.
+For instance, there are nice results for LMSR with Kelly betters (log utility), in a sense these two are aligned.
+
+* Is it interesting to have risk-averse agents, perhaps with just mean-variance utilities?
+
+* Can many of the nice properties of a scoring rule/market follow from properties of the underlying uncertainty function?
+E.g., conditional independence properties of LMSR that might relate to Hammersley-Clifford theorem.
+
 %%%

### 2013-12-16 WriteUp/maxentpredmkts.tex
rough draft EC paper

% Adjusted to add \usepackage[numbers]{natbib}
% \bibliographystyle{acmsmall} and \documentclass[prodmode,acmec]{ec-acmsmall}
% Jan 5, 2013 - David Parkes
%
% v2-acmsmall-sample.tex, dated March 6 2012
% This is a sample file for ACM small trim journals
%
% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
% (c) 2010 Association for Computing Machinery (ACM)
%
% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
% Users can also go through the FAQs available on the journal's submission webpage.
%
% Steps to compile: latex, bibtex, latex latex
%
% For tracking purposes => this is v1.3 - March 2012

\documentclass[prodmode,acmec]{ec-acmsmall} % Aptara syntax

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

% Metadata Information
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2013}
\acmMonth{6}
\usepackage[numbers]{natbib}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsthm}
\usepackage{parskip}
\usepackage{booktabs}

%new commands
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\mX}{\mathcal{X}}
\newcommand{\bR}{\mathbf{R}}
\newcommand{\grad}{\nabla}
\newcommand{\Exp}{\mathbf{E}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\q}{\mathbf{q}}
\newcommand{\mM}{\ensuremath{\mathcal{M}}}
\newcommand{\hmu}{\ensuremath{\hat{\mu}}}
\newcommand{\mP}{\ensuremath{\mathcal{P}}}
\newcommand{\mY}{\ensuremath{\mathcal{Y}}}
\newcommand{\mS}{\ensuremath{\mathcal{S}}}
\newcommand{\ds}{\displaystyle}
\newcommand{\mF}{\mathcal{F}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\htheta}{\hat{\theta}}
%---------------------------------------------------------------------------
\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{con}[thm]{Conjecture}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{ex}[thm]{Example}
\newtheorem{qn}[thm]{Question}
\newcommand{\defn}{\bigbreak\noindent{\bf Definition. }}
\newcommand{\undefn}{\bigbreak}
\newcommand{\rmk}{\bigbreak\noindent{\bf Remark. }}
\newcommand{\unrmk}{\bigbreak}
%\newcommand{\proof}{\noindent{\bf Proof. }}
\newcommand{\unproof}{\hfill$\Box$\bigbreak}
\newcommand{\defeq}{\stackrel{\mathrm{def}}{=}}
\newcommand{\holds}[1]{\stackrel{?}{#1}}
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\ie} {{\it i.e. }}
\newcommand{\eg} {{\it e.g. }}
\newcommand{\etal} {{\it et al. }}
\newcommand{\E}{\mathbf{E}}
\newcommand{\betavec}{\pmb{\beta}}
\newcommand{\qvec}{\mathbf{q}}
\newcommand{\lpf}{\mathbf{\psi}} %logpartition function
\newcommand{\family}{\mathcal F} 
\newcommand{\suff}{\pmb{\phi}} % suff stats
\newcommand{\muvec}{\pmb{\mu}}

% Document starts

\begin{document}
% Page heads
\markboth{Abernethy et al.}{Maximum Entropy Prediction Markets}
% Title portion
\title{Maximum Entropy Prediction Markets}
\author{JACOB ABERNETHY
\affil{University of Michigan, Ann Arbor}
SINDHU KUTTY
\affil{University of Michigan, Ann Arbor}
S\'{E}BASTIEN LAHAIE
\affil{Microsoft Research, New York City}
RAHUL SAMI
\affil{University of Michigan, Ann Arbor}}
%%\author{GANG ZHOU
%%\affil{College of William and Mary}
%%YAFENG WU
%%\affil{University of Virginia}
%%TING YAN
%%\affil{Eaton Innovation Center}
%%TIAN HE
%%\affil{University of Minnesota}
%%CHENGDU HUANG
%%\affil{Google}
%%JOHN A. STANKOVIC
%%\affil{University of Virginia}
%%TAREK F. ABDELZAHER
%%\affil{University of Illinois at Urbana-Champaign}}
% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{abstract}
In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets. We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data. Similarly,  the market state can be interpreted as a distribution over the outcome space. In particular, we show that a proper scoring rule can be derived from maximum entropy distributions. This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces. In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility. We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function. These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders. In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market. Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
\end{abstract}

\category{J.4}{Social and Behavioral Sciences}{Economics}
\category{I.2.6} {Artificial Intelligence}{Learning}

\terms{Prediction Market Design, Machine Learning Algorithms}

\keywords{Exponential Families, Bayesian Learning, Maximum Entropy Distributions, generalized LMSR}

\acmformat{Jacob Abernethy, Sindhu Kutty, S\'{e}bastien Lahaie, and Rahul Sami, 2014. Maximum Entropy Prediction Markets}
%%Gang Zhou, Yafeng Wu, Ting Yan, Tian He, Chengdu Huang, John A. Stankovic, and Tarek F. Abdelzaher, 2010. A multifrequency MAC specially designed for  wireless sensor network applications.

% At a minimum you need to supply the author names, year and a title.
% IMPORTANT:
% Full first names whenever they are known, surname last, followed by a period.
% In the case of two authors, 'and' is placed between them.
% In the case of three or more authors, the serial comma is used, that is, all author names
% except the last one but including the penultimate author's name are followed by a comma,
% and then 'and' is placed before the final author's name.
% If only first and middle initials are known, then each initial
% is followed by a period and they are separated by a space.
% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.

\begin{bottomstuff}
%%This work is supported by the National Science Foundation, under
%%grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.
%%
%%Author's addresses: G. Zhou, Computer Science Department,
%%College of William and Mary; Y. Wu  {and} J. A. Stankovic,
%%Computer Science Department, University of Virginia; T. Yan,
%%Eaton Innovation Center; T. He, Computer Science Department,
%%University of Minnesota; C. Huang, Google; T. F. Abdelzaher,
%%(Current address) NASA Ames Research Center, Moffett Field, California 94035.
\end{bottomstuff}

\maketitle

\section{Introduction}
We have dual goals in this paper. On the one hand, we highlight a structural similarity between prediction markets and exponential families. We see this as the syntax of our prediction market mechanism. On the other, this formulation has rich semantics as well: it allows for analysis of market behavior under various environments. For instance, we analyze market behavior with budget limited traders, traders that are exponential-utility-maximizers.
\section{Notation and Definitions}
convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesi
\section{Maximum Entropy Market Making}

The purpose of a prediction market is to elicit and aggregate the beliefs (i.e., subjective probabilities) of agents over a space of \emph{outcomes}. In a single-agent setting, a scoring rule is used to elicit the agent's beliefs. In a multi-agent setting, an information market is used to aggregate the agent's beliefs. Hanson~\cite{} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both so that they can perform well in both thin and thick markets. In this work we derive a wide-ranging generalization of the logarithmic market scoring rule. Using a maximum entropy approach, we explain how a market scoring rule can be developed for outcome spaces both discrete and continuous, and for generic properties of the underlying distribution (e.g., mean and variance), rather than just the probabilities of individual outcomes.

%%%
%%%
%%%

\subsection{The Model}
\label{model}

Let $\mX$ be the outcome space, which may be discrete or continuous. Let $\mP$ denote the set of probability distributions over the outcome space. We represent a probability distribution as a density $p$ absolutely continuous with respect to some base measure $\nu$ (e.g., counting for discrete outcomes, or Lebesgue for continuous outcomes).% We will discuss the interpretation of the base measure in a later section.

There is an unknown distribution $p$ over outcomes. We are interested in estimating the expected value of different outcome \emph{statistics} under this distribution by aggregating the beliefs of agents. A statistic is a real-valued function $\phi_s : \mX \rightarrow \bR$ over outcomes, independent of $p$; here $s$ belongs to a finite index set $\mS$ of size $d$. The collection of functions $\phi = (\phi_s,\, s \in \mS)$ can be viewed as a vector-valued statistic mapping outcomes into $\bR^d$. Formally, the aim is to estimate
%
\begin{equation} \label{constraint}
\Exp_p[\phi(x)] = \mu.
\end{equation}
%
We stress that we are only seek to elicit $\mu \in \bR^d$, not the full distribution $p$. As an example, suppose that $\mX = \bR$. If we were interested in eliciting the first two uncentered moments of $p$, we would include $\phi_1(x) = x$ and $\phi_2(x) = x^2$ as statistics. Note that the variance cannot be directly obtained through any single statistic: it corresponds to the expectation of $(x - \Exp_p[x])$, which depends on $p$, violating our definition of a statistic. Of course, the variance can be indirectly obtained by eliciting the first two moments separately.\footnote{This is related to the notion of \emph{elicitation complexity} of statistical properties introduced by Lambert et al.~\cite{Lambert}, but we do not pursue this connection here.}

There are two approaches to eliciting an agent's estimated $\mu$ in~(\ref{constraint}). The first approach is to incentivize the agent to directly report $\mu$ via a \emph{scoring rule}. The second approach is to form an \emph{information market} by issuing securities for each statistic $s \in \mS$, with payoff depending on the realized value of the statistic. (Such securities are known as contingent claims.) The amount of shares of each security acquired by the agent indirectly reveals it estimate $\mu$. In a market scoring rule~\cite{Hanson}, these two approaches are dual to each other in a formal sense.

%%%
%%%
%%%

\subsection{Scoring Rule}

We first consider how to directly elicit $\mu$ via a scoring rule. 
%Note that the usual definition of a scoring rule presumes that a full probability distribution is elicited, rather than just some of its properties (e.g., mean and variance), so below we will adapt the definition to our setting. 
The space of feasible reports from an agent corresponds to
%
\begin{equation} \label{marginal}
\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}.
\end{equation}
%
A scoring rule is a function $S : \mM \times \mX \rightarrow \bR$ that rewards an agent with $S(\mu,x)$ based on how its report $\mu$ agreed with the eventual outcome $x$. A scoring rule is \emph{proper} if for all $\mu \in \mM$, and $p \in \mP$ such that $\Exp_p[\phi(x)] = \mu$, we have
%
\begin{equation} \label{scoring}
\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)]
\end{equation}
%
for all $\hmu \neq \mu$; the rule is \emph{strictly proper} if the inequality is strict. Note that in the literature, scoring rules are defined over entire probability distributions, not just properties of those distributions. Our definition imposes a minimum of informational requirements on agents, because an agent does not need to assess the full distribution $p$ that might have lead to its estimates $\mu$ in order to realize that truthful reporting is an optimal strategy.

Our scoring rule is implicitly defined via the solution of a mathematical program. The program computes the maximum entropy distribution consistent with the agent's reported beliefs $\mu$.
%
\begin{eqnarray}
\max_{p \geq 0} & \ds - \int_{x \in \mX} p(x) \log p(x) \nu(dx) & \label{obj} \\
\mbox{s.t.} & \ds \int_{x \in \mX} \phi(x) p(x) \nu(dx) = \mu & \label{mean} \\
 & \ds \int_{x \in \mX} p(x) \nu(dx) = 1 & \label{normalize}
\end{eqnarray}
%
The non-negativity constraints together with~(\ref{normalize}) ensure that the solution is a probability distribution. Constraints~(\ref{mean}), which correspond to $d$ separate constraints, ensure consistency with the agent's report. The objective~(\ref{obj}) is the entropy of the solution $p$ with respect to the measure $\nu$. The program has a convex objective and linear constraints. %For continuous outcome spaces we will see in the next section that the optimal solution can still be cleanly characterized. 
%
\begin{theorem} \label{thm:scoring}
Let $p(x;\mu)$ be the optimal solution to the maximum entropy program, given report $\mu \in \bR^d$. The scoring rule defined by
%
\begin{equation} \label{scoring}
S(\mu,x) = a_x + b\log p(x;\mu),
\end{equation}
%
where $b,a_x \in \bR$ and $b > 0$, is strictly proper.
\end{theorem}
%
Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$

Note that the maximum entropy distribution is an exponential family distribution. 
 Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.

 We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
   Or 
\begin{equation}\label{eq:proper}
\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
\end{equation}

   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
   Rearranging the terms,
   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
   which is the first order condition on the convexity of $\psi(\cdot)$.
 \unproof

 \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters. That is,
   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
   \unrmk

\noindent
This rule represents a generalization of the logarithmic scoring rule for probability distributions to any properties of distributions that can be captured as expectations of statistics.  As explained in the next section, the solution to the maximum entropy program takes the form of an \emph{exponential family} distribution. For many common properties of interest, these distributions are familiar; for example, the maximum entropy distribution with a given mean is an exponential distribution, and the maximum entropy distribution with a given mean and variance is a normal distribution~\cite{}. We provide here two examples in more depth.

%%
%%

\paragraph{Multinomial Distribution}

As a first example, consider the problem of estimating the individual probabilities of a finite set of outcomes. We have $\mX = \{1,\ldots,k\}$. The relevant statistics indicate which outcome actually occurs, so the index set is $\mS = \mX$. Statistic $\phi_x(x')$ is 1 if $x' = x$ and 0 otherwise. Observe that if $p$ is the distribution over outcomes, then $\Exp_p[\phi(x)] = p$. A feasible report from an agent is any non-negative $\pi = (\pi_1,\ldots,\pi_k)$ such that $\sum_{i=1}^k \pi_i = 1$. Given a report of $\mu = \pi$, the unique solution to the maximum entropy program is $p(x; \pi) = \pi$, and~(\ref{scoring}) corresponds to the classic logarithmic scoring rule of $S(\pi,x) = \log \pi_x$. \hfill{$\Box$}

\paragraph{Normal Distribution}

As a second example, suppose the outcome space is $\mX = \bR$ and that we are interested in estimating the mean $\mu$ and variance $\sigma^2$ of the underlying distribution. We choose the first and second (uncentered) moments, $\kappa_1 = \mu$ and  $\kappa_2 = \mu^2 + \sigma^2$ as our statistics.  It is well-known that the maximum entropy distribution with a given mean and variance is the normal distribution, so the solution to the maximum entropy program (using Lebesgue as the base measure $\nu$) is
%
\[ p(x; \kappa) = \frac{1}{\sqrt{2\pi\sigma^2}} \exp\left[ -\frac{(x-\mu)^2}{2\sigma^2} \right]. \]
%
Applying Theorem~\ref{thm:scoring} and choosing appropriate constants, we find that the following scoring rule is strictly proper in this context. %
\[ S(\kappa, x) = -\frac{(x-\mu)^2}{\sigma^2} - \log \sigma^2. \]
%
Note that the fact that this scoring rule is proper for the mean amounts to the well-known fact that reporting the mean is a Bayes act under quadratic loss~\cite{someone}. \hfill{$\Box$}

\medskip\noindent
A brief observation on parametrizations is in order. Note that the maximum entropy distribution here is parametrized by the uncentered moments $\kappa = (\kappa_1,\kappa_2)$ rather than $(\mu,\sigma^2)$ because the variance is not a valid statistic. However, it is clear that the agent does not have to directly report $\kappa$: the market maker itself can do the translation from reported mean and variance to uncentered moments. This is relevant because the $\kappa$ parametrization may be unintuitive for an agent, especially given the constraints $\kappa_2 - \kappa_1^2 \geq 0$ that must be enforced. In the $(\mu, \sigma^2)$ parametrization each parameter is unrestricted, and the parameters have intuitive interpretations.

Our approach so far has been to identify relevant properties of the unknown distribution, and elicit them via a scoring rule computed through a maximum entropy program. However, it is the case that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}. Therefore, we see the following alternative approach. The market designer can begin by considering the outcome space $\mX$, and assessing what family of distributions over $\mX$ the unknown $p$ might come from. If this family is an exponential family, then Theorem~\ref{thm:scoring} immediately gives a proper scoring rule to elicit the parameters of the distribution. Many of the most familiar distributions are exponential families, including the normal, exponential, gamma, beta, binomial, Poisson, Weibull, and Dirichlet. To illustrate, we consider two examples in more depth.

%%
%%

\paragraph{Exponential}
Suppose the designer would like an estimate of the rate at which servers fail in a data center. An outcome is the time between failures, so $\mX = [0,+\infty)$. The usual distribution for failure times is the exponential distribution $p(x;\lambda) = \lambda e^{-\lambda x}$, which is parametrized by the rate $\lambda$ and has mean $\mu = 1/\lambda$. The exponential distribution is the maximum entropy distribution given a fixed mean $\mu$. From Theorem~\ref{thm:scoring} we obtain the following proper scoring rule for this setting:
%
\[ S(\mu, x) = \log \lambda - \lambda x. \]
%
It is simple to check using basic calculus that an agent maximizes the expected score by reporting $\lambda = 1/\hmu$, where here $\hmu$ is its estimate of the mean. In this case, it is unclear a priori which of the mean or rate parametrizations is most intuitive, but the maximum entropy approach easily allows for either. \hfill{$\Box$}

%%
%%

\paragraph{Beta}
Suppose the designer would like an estimate of the click-through rate of an online advertisement. The outcome space here is $\mX = (0,1)$. A typical prior over a probability like this is the beta distribution, which has two parameters $\alpha, \beta > 0$ and density function $p(x; \alpha, \beta) = x^{\alpha-1}(1-x)^{\beta-1}/B(\alpha,\beta)$, where $B$ is the beta function. By theorem~\ref{thm:scoring} this leads to the proper scoring rule
%
\[ S((\alpha, \beta), x) = (\alpha-1)\log x + (\beta-1)\log (1-x) - \log B(\alpha,\beta). \]
%
Although this parametrization is standard for the beta, it is perhaps more intuitive to use the parametrization $\mu = \alpha/(\alpha+\beta) \in (0,1)$ and $n = \alpha+\beta > 0$. Here $\mu$ is the mean of the distribution---corresponding to the agent's actual estimate of the click-through rate---and $n$ is a measure of the agent's confidence in the estimate. \hfill{$\Box$}

%%%
%%% 
%%%

\subsection{Information Market}

We next consider how to indirectly elicit $\mu$ by setting up a market of contingent claim securities. Under this approach, the elements of the index set $\mS$ are interpreted not as statistics but as securities. The payoff from one share of security $s$ when outcome $x$ occurs is given by the mapping $\phi_s$. Thus if the vector of shares held by the agent is $\theta \in \bR^d$, where entry $\theta_s$ corresponds the number of shares of security $s$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \theta, \phi(x) \ra$. As a concrete example, recall our multinomial distribution example from the previous section, where $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise. This means that security $x \in \mS$ pays 1 dollar if outcome $x \in \mX$ occurs, and nothing otherwise. (Such securities are known as Arrow-Debreu securities.) In our normal distribution example, the statistic for the mean had the mapping $\phi_1(x) = x$. Therefore a share of the corresponding security has a payoff that is linear in the outcome. (Such securities amount to futures contracts.)

To extract useful information from such a securities market, we set a pricing scheme and examine the number of shares the agent chooses to acquire. Let $\Omega$ be the set of all portfolios (i.e., vectors of shares) that the agent can feasibly hold; we will see how this domain is determined in an instant. Each security $s$ has an associated price function $c_s : \Omega \rightarrow \bR$, which gives the marginal price $c_s(\theta)$ of security $s$ when the agent holds portfolio $\theta$. (Note that the marginal price depends on the entire portfolio $\theta$, not just the number of shares $\theta_s$.) A risk-neutral agent will choose to acquire shares up to the point where, for each share, expected payoff equals marginal price. Formally, if the agent acquires portfolio $\theta$, then for each $s \in \mS$ we must have
%
\begin{equation} \label{market}
\Exp_p[\phi_s(x)] = c_s(\theta).
\end{equation}
%
In this way, by its choice of $\theta$, the agent reveals that its belief is $\mu = c(\theta)$. There are several important properties that the price function $c$ should have. To simplify the agent's portfolio acquisition process, it should not be the case that the total cost of acquiring $\theta$ depends on the order in which shares are bought. This means that there should exist a \emph{cost function} $C : \Omega \rightarrow \bR$ such that $c = \nabla C$. The gradient $\nabla C$ must be onto $\mM$ to ensure that~(\ref{market}) always has a solution---this is the most important but technical condition to realize. To ensure that the solution is unique, it would also be convenient if the cost function were strictly convex. 
%
\begin{proposition}
The following cost function is monotone, convex, and onto $\mM$: 
%
\begin{equation} \label{cost}
C(\theta) = \log \int_{x \in \mX} \exp \left[\la \theta, \phi(x) \ra\right] \nu(dx).
\end{equation}
%
\end{proposition}
%
\noindent
Observe that in the context of our earlier multinomial example, cost function~(\ref{cost}) is exactly the cost function for Hanson's logarithmic market scoring rule. Our approach here provides a generalization of this market scoring rule to markets with contingent claim securities with arbitrary payoffs, designed to elicit specific properties of distributions, beyond just the probabilities of different outcomes.

Because an agent would never select a portfolio with infinite cost, the effective domain of $C$ is $\Omega = \{\theta \:|\: C(\theta) < +\infty \}$. In the context of the multinomial distribution example, $\Omega = \bR^d$, so that shares of each security can always be bought or sold short. In the context of the normal distribution example the effective domain is $\Omega = \{(\theta_1,\theta_2) \in \bR^2 \:|\: \theta_2 > 0 \}$ if we re-define the second statistic to be $\phi_2(x) = -x^2$; this means that the security has a negative payoff for each share, and consequently the agent must be compensated to acquire such shares. Evaluating~(\ref{cost}), the cost function takes the form $C(\theta) = \frac{\theta_1^2}{4\theta_2} - \frac{1}{2}\log(2\theta_2)$.

Now, in our previous elicitation approach that used a scoring rule, the process of computing the scoring rule also provided a complete distribution $p(x;\mu)$ over outcomes, which could be used to infer other properties beyond just the agent beliefs $\mu$. In the current market-based approach, the agent's chosen portfolio $\theta$ can also form the basis of a distribution over outcomes. Consider the following distribution, represented as a density with respect to a base measure $\nu$:
%
\begin{equation} \label{expfam}
p(x;\theta) = \exp\left[ \la \theta, \phi(x) \ra - C(\theta) \right].
\end{equation}
%
Meaning, the probability that a subset $X \subseteq \mX$ of the outcomes occurs is $\int_{x \in X} p(x;\theta)\nu(dx)$. Observe that by definition~(\ref{cost}), this probability density indeed integrates to 1 over $\mX$, and it is clear that the density is non-negative as required.

In the statistical literature a distribution that takes the form~(\ref{expfam}) is known as an \emph{exponential family}. The mapping $\phi$ is known as the \emph{sufficient statistic}, $\theta$ is the \emph{natural parameter}, and $C$ is the log-partition or \emph{cumulant} function. %This connection between information markets and exponential families allows us to draw on a vast statistical literature to understand the market's properties, and a growing computer science literature to apply efficient algorithms to compute costs and perform inference.

%%%
%%%
%%%

\subsection{Duality}

There is a well-known duality between the maximum entropy approach and exponential families, which translates into a duality between the scoring rule and information market just developed. The duality implies that the approach leads to a \emph{market scoring rule}, applicable to both thin and thick multi-agent settings.

It is known that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}. To see this, let $\theta(\mu) \in \bR^d$ be the Lagrange multiplier corresponding to constraints~(\ref{mean}) when solving the maximum entropy program given the agent report $\mu$. (Our choice of notation is deliberately suggestive.) Let $A(\mu)$ be the Lagrange multiplier corresponding to~(\ref{normalize}). From the first order necessary conditions for optimality, we find that
%
\[ p(x;\mu) = \exp \left[ \la \theta(\mu), \phi(x) \ra - A(\mu) - 1 \right], \]
%
so that the solution is indeed in exponential family form. Conversely, given a cumulant $C$ from an exponential family and a parameter $\theta$, we see that $p(x;\theta)$ defined according to~(\ref{expfam}) is the maximum entropy distribution under constraints $\Exp_p[\phi(x)] = \mu$ where $\mu = \nabla C(\theta)$. We obtain the following.
%
\begin{proposition}
Let $C$ be the cumulant (i.e, cost function) for the exponential family corresponding to $\phi$, and let $\theta(\mu)$ be the optimal Lagrange multiplier for the mean constraints given an agent report of $\mu \in \mM$. Then the following scoring rule is equivalent to~(\ref{scoring}):
%
\begin{equation} \label{expscore}
S(\mu, x) = \la \theta(\mu), \phi(x) \ra - C(\theta(\mu)).
\end{equation}
%
\end{proposition}
%
\noindent
The scoring rule~(\ref{expscore}) decomposes neatly into payoff and cost functions. The first term $\la \theta(\mu), \phi(x) \ra$ defines the agent's outcome-contingent reward, while the second term~$C(\theta(\mu))$ is the cost of acquiring a portfolio $\theta(\mu)$. We see here the duality between the approach of directly reporting $\mu$, or acquiring shares $\theta$: an agent reporting beliefs $\mu$ under scoring rule~(\ref{scoring}) would choose to acquire shares $\theta(\mu)$ in the information market with cost function~(\ref{cost}). Since~(\ref{scoring}) is a proper scoring rule, the information market with cost function~(\ref{cost}) is based on a proper market scoring rule.

\section{The Exponential Family Market Mechanism: A Maximum Likelihood Approach}
In this section, we will set up a prediction market that aggregates beliefs from traders where the outcome is a continuous random variable. In particular, we assume that the outcome is drawn from an exponential family distribution. Each of the traders has access to a series of points drawn from this distribution. In other words, the traders have access an empirical mean of the sufficient statistics of the exponential family. Every trader has infinite budget so that the current market price after a trader has traded in the prediction market, exactly reflects his beliefs.

%We will now provide a more formal description of the problem statement.
%\subsection{Problem Statement}
%\begin{description}
%\item[Setup: ] Learning proceeds in rounds $j=1,2,\ldots$ where at the end of each round the algorithm may be provided with $x_{j}$  drawn from a set $\mathcal{X}$ according to some distribution $\mathcal{D}_{j}$, where $\mathcal{D}_{j}$ belongs to an exponential family defined by $k$ sufficient statistics.  In each round, some sequence of $n$ traders provide their estimates of the empirical means of each of the $k$ sufficient statistics drawn from $\mathcal{D}_{j}$.
%\item[Define: ] A learning algorithm that learns this distribution by providing a maximum likelihood estimate of its unknown natural parameters based on the empirical means.  
%\item[Proposed solution: ] Simulate a prediction market defined by a cost function $C$, securities $s_{0},s_{1},s_{2},\ldots, s_k$ and their payoffs $\phi_i(x)$ for $i=0,1,2,\ldots,k$ for the LMSR. Define a correspondence between the state of the prediction market and the output of the learning algorithm.
%\end{description}

For $x_{i}\sim P_{\betavec}$, recall that the likelihood function for independently drawn data $x_1,\ldots,x_n$ is given by $\prod_{i=1}^{n} P_{\betavec}(x_i)$. The maximum likelihood estimate of the natural parameters $\betavec$ is the value of the natural parameters that maximizes the likelihood function.

We will now set up a prediction market with log market scoring rule (LMSR) and infinite budget traders such that the market state represents the maximum likelihood estimate (MLE) of the natural parameters of an exponential family distribution. %Given  under some interpretation of what information the participating agents have, and how they bet.

For a given exponential family distribution, the prediction market is defined as follows:
\begin{description}
\item[Traders] The prediction market will simulate a trader $i$ corresponding to expert $i$. This trader processes all information samples available to her directly or inferred from previous trades, and  forms a belief distribution such that the believed means of the sufficient statistics matches the empirical means of the sufficient statistics from the information samples. The trader trades in the market to maximize her expected payoff under her believed distribution.
\item[Securities and their payoff]
For each $i=1,2,\ldots,k$, we define a security $s_{i}$ with payoff $\phi_{i}(x)$ where $x$ is the ultimate outcome and $\Phi()$ defines the vector of sufficient statistics of the exponential family distribution according to which $x$ is drawn. We define an additional security $s_{0}$ with payoff $\phi_{0}(x):=a-\sum_{i=1}^{k}\phi_{i}(x)$ where $a$ is an appropriately chosen constant dependent on the range of $\Phi$ so that the payoff of $s_{0}$ is non-negative. 
\end{description}

We also note that since we want non-negative payoffs, we restrict the exponential families under consideration so that the sufficient statistics are lower bounded. If we have a constant lower bound on the sufficient statistics, then without loss of generality we can add a constant to each sufficient statistic without changing the exponential family in any way, as the constants will be absorbed in the log-partition function.

\paragraph{Arbitrage-free Property}
We can easily show that the prediction market we have defined is arbitrage-free: there is no sequence of trades that guarantees the trader a profit under all conditions. First, note that the incremental cost and payoff function are additive over multiple trades, so the net profit of a sequence of trades depends only on the initial and final market position, and is independent of the actual path along which the trade takes place.

Now, consider any trade (or sequence of trades) that moves the market from an initial state of $\qvec$ to a final state of $\qvec + \qvec'$. The cost of this trade is $C(\qvec + \qvec') - C(\qvec)$. The cost function of our market is the log-partition function of an exponential family, and thus, $C(.)$ is a strictly convex function~\cite{WainJordan08}. Thus, we have the inequality:
\[
   C(\qvec + \qvec') - C(\qvec)  > \qvec' \nabla C|_\qvec = \qvec' \mu,
 \]
 where $\mu$ is the mean sufficient statistics vector for the distribution with parameters $\qvec$.
The payoff due to this trade, with outcome $x$, is $\qvec'.\phi(x)$. 
Now, we observe that under the distribution with parameters $\qvec$, the expected payoff is $\qvec' E[\Phi(x)] = \qvec' \mu$. Thus, under this distribution over outcomes $x$, the expected profit is strictly less than $0$. This is only possible if the
realized profit is less than $0$ for at least one outcome $x$. As this statement is true for every $\qvec$ and $\qvec'$, and the profits are path-independent, the market is arbitrage-free.

%Let $\qvec^*=(q^{*}_{0},q^{*}_{1},\ldots,q^{*}_{k})$ be the number of shares of each security held by the traders. First we note that under the assumption of perfectly rational, risk-neutral traders with infinite budget and  beliefs as indicated above, the gradient of the cost function at this point is
 %$$\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}=\mu_{i}$$
%where $\mu_{i}$ is believed to be the expected payoff of the $i^{th}$ security by the last trader who traded in this market. We show below that the instantaneous price of security $i$, $\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$ at any $\qvec=\qvec^{*}$, can be written as $\left(\frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$. This means that the instantaneous price of the $i^{th}$ security only depends on the difference between the number of held securities of $i$ and 0. This ensures that if there exists an instantaneous price that reflects the belief of a trader at any vector $\qvec^{*}$, there must exist a $\qvec'$  that also reflects his beliefs and with the desirable property that $\forall i=0,1,2,\ldots,k\ q_{i}'\geq 0$. This additional property ensures that no trader will be required to short sell securities to move the market state to  reflect his beliefs.

\paragraph{Market State and Natural Parameters}
We define the interpretation function 
\begin{eqnarray}\label{eqn:int}
I(\qvec)&=&(q_{1}-q_{0},q_{2}-q_{0},\ldots,q_{k}-q_{0})\nonumber\\
&=&(\beta_{1}, \beta_{2},\ldots, \beta_{k})=\betavec
\end{eqnarray}
This allows us to interpret the state of the market in terms of a prediction on the natural parameters of the distribution.

\begin{theorem}
The natural parameter vector corresponding to the interpretation of the market state given by Equation \ref{eqn:int} is the vector of their maximum likelihood estimates. Further, this interpretation is unique.
\end{theorem}
\begin{proof}
We observe that for $i=1,\ldots,n$,
\begin{eqnarray*}
\mu_{i}&=&\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial\log\int \exp(\qvec^{T}\Phi(x))dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial\log\int e^{q_{0}\phi_{0}(x)+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial\log\int e^{q_{0}(a-\sum_{i=1}^{n}\phi_{i}(x))+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial\log\int e^{q_{0}a} e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial\log (e^{q_{0}a} \int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial (q_{0}a)}{\partial q_{i}} \frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial \lpf(\betavec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\frac{\partial \beta_{i}}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\right)_{\rm \betavec=\betavec^{*}}
\end{eqnarray*}
Thus, this choice of parameters satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$. We will now argue that this vector is unique.

If the exponential family is represented so that there is a unique parameter vector associated with each distribution, the representation is said to be minimal. The Bernoulli, Gaussian, and Poisson distributions all have minimal representations. Now, for an exponential family whose representation is minimal, the gradient is an injection \cite[Prop. 3.2]{WainJordan08}. Thus, there is a unique parameter vector $\betavec$ that satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$, for each $i\in\{1,\ldots,n\}$. Thus, if the $\mu_{i}$'s correspond to the empirical means, since our choice of $\betavec$ satisfies the equality, it must also be the vector corresponding to the MLE.
\end{proof}

Thus, we have shown that this prediction market (along with its strategic traders) aggregates trader information in a way that the market state can be interpreted as a predictive distribution (the maximum likelihood distribution) over an infinite outcome space.

It is worth noting here that if we can bound the parameter space and the sufficient statistics of the distribution, the market maker defined here has bounded worst case loss: The probability density, and hence log loss, would then be bounded for all market states that could be reached and all outcomes. In fact, for many practical applications, a distribution with unbounded parameter space is an approximation for a true space which itself may be bounded. In this case, worst case loss would thus be bounded.

\subsection{The Exponential Family Market Mechanism in Adversarial Markets}
In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.

In this section, we consider the the prediction market setup with traders that may be either informative or malicious. The malicious traders may want to inject faulty information into the market. The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.

We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market. Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.

\subsubsection{Budget-limited Aggregation}
Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices. We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market. This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader. Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.

%There is distribution under which expected profit is $\leq 0$. Cost function is convex.

%For any $\qvec$ there is a
In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
Let $\alpha$ be the budget of a trader in the market. Suppose with infinite budget, the trader would have moved the market state from $\qvec_{init}$ to $\qvec$, where $\qvec$ represents his true beliefs. Suppose further that $\alpha<C(\qvec)-C(\qvec_{init})$. In this case, we want to budget-limit the trader's influence on the market state.

We define the budget-limited final market state as $\tilde{\qvec}$. 
Here, we consider a specific functional form of $\tilde{\qvec}$: $$\tilde{\qvec}=\lambda\qvec+(1-\lambda)\qvec_{init}$$  where $$\lambda=\min\Big(1, \frac{\alpha}{C(\qvec) - C(\qvec_{init})}\Big)$$ We first show that this trade is feasible given the trader's budget:
\begin{theorem}
Let the current market state be given by the vector $\qvec_{init}$ and $\tilde{\qvec}=\lambda\qvec+(1-\lambda)\qvec_{init}$. For $\lambda=\min\Big(1, \frac{\alpha}{C(\qvec) - C(\qvec_{init})}\Big)$, the cost to the trader to move the market state from $\qvec_{init}$ to $\tilde{\qvec}$ is at most his budget $\alpha$.
\end{theorem}
\begin{proof}
From the convexity of $C$, we have
 $$C(\tilde{\qvec})\leq(1-\lambda)C(\qvec_{init})+\lambda C(\qvec)$$

Now 
\begin{eqnarray*}
C(\tilde{\qvec}) - C(\qvec_{init}) &\leq& (1-\lambda)C(\qvec_{init})\\&&+\lambda C(\qvec) - C(\qvec_{init})\\
&=& \lambda\ (C(\qvec) - C(\qvec_{init}))
\end{eqnarray*}

%So when $\alpha<C(\qvec)-C(\qvec_{init})$, we pick $\lambda$ so that 
%$$\lambda\leq \frac{\alpha}{(C(\qvec) - C(\qvec_{init}))}$$
%
%where we want $\lambda$ such that
 Thus, $C(\tilde{\qvec}) - C(\qvec_{init}) \leq \alpha$. 
 \end{proof}

 We note that moving to $\tilde{\qvec}$ as defined may not be the optimal trade for a rational trader maximizing her expected profit. In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\tilde{\qvec}$. Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\qvec$.
\subsubsection{Damage Bound}
Given this restriction on the initial budget of traders, we can now show that the total loss that can be induced by malicious traders is bounded. We define the loss function for $\qvec$ shares held as:
$$L(\qvec,x)=-\log(p_{\betavec}(x))=\log\frac{\int e^{\qvec^{T}\Phi(x)}\ dx}{e^{\qvec^{T}\Phi(x)}}$$
with the correspondence between $\betavec$ and $\qvec$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.

%{\bf Setup:} The market evolves in rounds $t=1,\ldots,T$. The end of a time segment occurs either when each expert has made his report or when the actual value of the random variable is received as input. Each expert $i$ reports the means of the sufficient statistics as available to him at time $t$. These are combined with the previously reported means to yield $\mu_{0}^{t}(x),\mu_{1}^{t}(x),\ldots,\mu_{k}^{t}(x)$ and an associated weight $w_{i}^{t}$ which can be interpreted as the number of points from which these values have been determined. These sufficient statistics correspond to a particular value of $B$ and hence of $\qvec$, the number of shares of each security that are held by traders. Since trader $i$ has finite budget $\alpha_{i}^{t}$ at time $t$, the current market state is moved to $\qvec'$ instead, so that cost of moving it to $\qvec'$ is within his budget.

%For an initial positive budget of $\alpha_{i}^{0}$, we perform updates to the budgets of each agent as
%$$\alpha_{i}^{t}:=\alpha_{i}^{t-1}+\lambda_{i}^{t}(L(\tilde{q}_{i-1}^{t},x^{t})-L({q}_{i}^{t},x^{t}))$$ where we define
%$$\lambda_{i}^{t}:=\min\Big(1, \frac{\alpha_{i}^{t-1}}{C(\qvec_{i}) - C(\tilde{\qvec}_{i-1})}\Big)$$

Suppose the prediction market runs over multiple rounds $t$. Let $\qvec_{0}^{t}$ be the initial number of shares of each security that are held. Let $\tilde{\qvec}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports. Let us assume that at this point we receive the value of the random variable $x^{t}$.

%Let the loss function be $L(\cdot,\cdot)$, where the two arguments are the vector representing the number of shares held of each security and the actual value of the random variable.

%We want to show:
%$$L({\qvec}_{0}^{t},x^{t})-L(\tilde{\qvec}_{k}^{t},x^{t})+\sum_{i}\alpha_{i}^{t-1}\leq \sum_{i}\alpha_{i}^{t}$$
%That is, the loss due to incorporating the budget-limited information of the traders, is not more than the increase in their budgets.
Over multiple instances of the prediction market, we can track the change in budget of each trader. The change in budget for trader $i$ is
\begin{eqnarray*}
\alpha_{i}^{t}-\alpha_{i}^{t-1}&=&C(\tilde{\qvec}_{i-1}^{t})-C(\tilde{\qvec}_{i}^{t})-(\tilde{\qvec}_{i-1}^{t}-\tilde{\qvec}_{i}^{t})^{T}\Phi(x^{t})\\
&=&L(\tilde{\qvec}_{i-1}^{t},x^{t})-L(\tilde{\qvec}_{i}^{t},x^{t})
\end{eqnarray*}

Define the myopic impact of a trader $i$ in segment $t$ as
$$\Delta_{i}^{t}:=L(\tilde{\qvec}_{i-1}^{t},x^{t})-L(\tilde{\qvec}_{i}^{t},x^{t})$$
Thus, the myopic impact captures incremental gain due to the trader in a round. Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.

The total myopic impact due to all $k$ active traders is given by
$$\Delta^{t}=L(\qvec_{0}^{t},x^{t})-L(\tilde{\qvec}_{k}^{t},x^{t})$$
Thus $-\Delta^{t}$ captures the incremental loss due to the predictive probability after aggregation of all $k$ traders.

%The myopic impact of a trader $i$ in segment $t$, is given by
%$$\Delta_{i}^{t}:=L(\tilde{B}_{i-1}^{t},x^{t})-L(\tilde{B}_{i}^{t},x^{t})$$
%where $\tilde{B}_{i}$ is the vector of sufficient statistics corresponding to the the budget-limited state of the market as induced by trader $i$.
%The total myopic impact due to all $k$ active traders is thus given by
%$$\Delta^{t}=L(B_{0}^{t},x^{t})-L(\tilde{B}_{k}^{t},x^{t})$$

%If the loss function is convex in the first argument, then we have:
%\begin{eqnarray*}
%\Delta_{i}^{t}&=&L(\tilde{B}_{i-1}^{t},x^{t})-L(\tilde{B}_{i}^{t},x^{t})\\
%&=&L(\tilde{q}_{i-1}^{t},x^{t})-L(\tilde{q}_{i}^{t},x^{t})\\
%&=&L(\tilde{q}_{i-1}^{t},x^{t})-L(\lambda_{i}{q}_{i}^{t}+(1-\lambda_{i})\tilde{q}_{i-1}^{t},x^{t})\\
%&\geq&L(\tilde{q}_{i-1}^{t},x^{t})-\lambda_{i}L({q}_{i}^{t},x^{t})-(1-\lambda_{i})L(\tilde{q}_{i-1}^{t},x^{t})\\
%&=&\lambda_{i}(L(\tilde{q}_{i-1}^{t},x^{t})-L({q}_{i}^{t},x^{t}))\\
%&=& \alpha_{i}^{t} - \alpha_{i}^{t-1}
%\end{eqnarray*}
%Thus, the myopic damage caused by trader $i$ at round $t$ is upper-bounded by the change in his budget in that round.

%If we pick a loss function that is also bounded in $[0,1]$, then we see that the budget never falls below zero. 
\begin{theorem}
A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
\end{theorem}
\begin{proof}
Consider the myopic impact of a single trader $i$ after participating in the market $T$ times. Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$

Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
\end{proof}

This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders.

In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation. We provide an information-theoretic justification for this claim in Section \ref{sec:infth}. The intuition behind this claim is that his prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.

\subsubsection{An Information-Theoretic Interpretation}\label{sec:infth}
 We observe a useful alternative view of the market 
scoring rule prediction market for exponential family learning. We connect the cost, payoff and profit function to information-theoretic 
quantities associated with the exponential family.

The following result has been previously pointed out by Amari \cite{Amari-KL}. 
\begin{theorem}\label{lem:profit_decomposition}{\bf (Profit Decomposition)}: Consider an exponential 
family $\family$ of distributions over some set of statistics $\suff(x)$, with natural parameters $\betavec$. Let $\pi, \rho \in \family$ be any two probability distributions in the family. We use 
$\betavec_\pi$ to denote the natural parameters of $\pi$ and $\muvec_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
we can define $\betavec_\rho$ and  $\muvec_\rho$. We abuse notation slightly and let $\lpf(\rho)$ indicate the log partition function of $\rho$ which technically depends on its natural parameters.
Let $H(\pi)$ denote the entropy of the distribution $\pi$, and 
 $K(\pi||\rho)$ denote the KL-divergence of $\rho$ relative to $\pi$.
Then, the following equality holds:  
\begin{equation}
 K(\pi||\rho) + H(\pi) = \lpf(\rho) - \betavec_\rho \cdot \muvec_\pi
\label{eq:profit_decomposition}
\end{equation}
\end{theorem}
\begin{proof}
\begin{eqnarray*}
 K(\pi||\rho) + H(\pi) &=& \int_x \pi(x)\log{\frac{\pi(x)}{\rho(x)}}dx\\
 &&\quad - \int_x \pi(x)\log{\pi(x)}dx \\
&=& - \int_x \pi(x)\log{\rho(x)}dx \\
&=& - \int_x \pi(x)\left[ \betavec_\rho \cdot \suff(x) - \lpf(\rho)\right]dx\\
&=& - \betavec_\rho \cdot \int_x \pi(x) \suff(x) dx\\
&&\quad + \lpf(\rho) \int_x \pi(x) dx \\
&=&  \lpf(\rho) - \betavec_\rho \cdot \muvec_\pi
\end{eqnarray*}
\end{proof}

Equation~\ref{eq:profit_decomposition} gives us an alternative view of the
market scoring rule construction. Assume that $\pi$ is the true distribution, and consider two arbitrary
distributions $\rho_1, \rho_2 \in \family$. Note that $\lpf(\rho)$  is independent of $\pi$, and $\betavec_\rho \cdot \muvec_\pi$ is linear in the probabilities $\pi(x)$. If we want to measure loss by the KL-divergence, we can do so (in expectation) by setting a cost function that captures the first term, and defining security quantities ($\betavec$) and payoffs ($\suff$) to capture the second term.  In particular, in a market with cost function $\lpf$, if the
market price initially implies a distribution $\rho_1$, and a trader moves
the market to price than implies a distribution $\rho_2$, then the cost she
incurs is $\lpf(\rho_2) - \lpf(\rho_1)$. The number of securities bought to make this trade is given by the vector ($\betavec_{\rho_2} - \betavec_{\rho_1}$), and the expected payoff of the securities are given by $\muvec_\pi$. Thus, by equation \ref{eq:profit_decomposition}, the {\em net profit} of the trader is equal to $K(\pi||\rho_1) - K(\pi||\rho_2)$, {\it i.e.}, the reduction in KL-divergence with respect to the true distribution.  We note one useful property of this construction: For a fixed vector $\betavec$ of purchased securities, the cost is independent of the outcome 
(and outcome distribution $\pi$), while the payoff is independent of the initial market state in which these securities were purchased.

%Thus, we have shown that there exist deep connections between learning the natural parameters of an exponential family distribution and cost function based prediction markets. In Chapter \ref{nonmyopic} we exploit this connection and propose a bounded regret learning algorithm in a Bayesian model under less-than-ideal conditions: where the experts may be adversarial and strategic.

\subsubsection{Budget of Informative Traders}\label{sec:inf}
Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.

%For this, we need the following simple extension of Jensen's inequality that follows directly from the definition of convexity of a function.
%\begin{lemma}\label{lem:convex}
%For a strictly convex function $f$ with unique minimum $x_{min}$, $x\neq x_{min}$ in the domain of $f$, and $\lambda \in (0,1]$,
%$$f(\lambda x_{min}+ (1-\lambda)x)< f(x)$$
%\end{lemma}
%\begin{proof}
%\begin{eqnarray*}
%f(\lambda x_{min}+ (1-\lambda)x) &<& \lambda f(x_{min})+ (1-\lambda)f(x)\\
%&=& \lambda(f(x_{min})-f(x)) +f(x)\\
%&\leq& f(x)
%\end{eqnarray*}
%\end{proof}

We now characterize the expected change in budget for an informative trader. The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
\begin{theorem}\label{thm:growth}
 Suppose that each informative trader gets a random sample of data, 
resulting in a sequence of trader beliefs $\qvec_i$, and hence a sequence of budget-limited market positions
$\tilde{\qvec}_i$, before a final outcome $x$. Then, the expectation, over trader $i$'s belief distribution $\qvec_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\tilde{\qvec}_{i-1}$. 
\end {theorem}
\begin{proof}
Trader $i$'s believed distribution is the distribution parametrized by $\qvec_i$.
Therefore, the expected profit, over possible outcome values $x$, for a trader $i$ in a given round is given by
\begin{eqnarray*}
&& E[C(\tilde{\qvec}_{i-1})-C(\tilde{\qvec}_{i})-(\tilde{\qvec}_{i-1}-\tilde{\qvec}_{i})\Phi(x)]\\
&=& K(\qvec_{i}||\tilde{\qvec}_{i-1})] + H(\qvec_{i}) - [K(\qvec_{i}||\tilde{\qvec}_{i}) + H(\pi)]]\\
&=& K(\qvec_{i}||\tilde{\qvec}_{i-1}) - K(\qvec_{i}||\tilde{\qvec}_{i})
\end{eqnarray*}
We recall that $\tilde{\qvec}_{i}$ can be expanded as:
$$\tilde{\qvec_{i}} = \lambda \qvec_{i} + (1- \lambda) \tilde{\qvec}_{i-1}$$ where $\lambda$ is strictly greater than $0$, but no more than $1$. It is a standard result from information theory that $K(.,.)$ is a convex function, and therefore, we have:
\[
K(\qvec_{i}||\tilde{\qvec}_{i-1}) \leq \lambda K(\qvec_{i}||\qvec_{i}) + (1-\lambda) K(\qvec_{i}||\tilde{\qvec}_{i-1})
\]
The first term on the right hand side is zero, and thus, we get:
\[
  K(\qvec_{i}||\tilde{\qvec}_{i-1}) - K(\qvec_{i}||\tilde{\qvec}_{i}) \geq \lambda K(\qvec_{i}||\tilde{\qvec}_{i-1})
\]
%Thus, it is sufficient to prove that E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t}] >= E_{\qvec}[K(\pi||\tilde{\qvec}_{i}^{t}].
 \end{proof}
%\begin{theorem}
%The expected budget of an informative trader increases when he participates in the market. Here, expectation is taken with respect to the true distribution from which the outcome is drawn.
%We show that the budget of an informative trader grows in expectation:\end{theorem}
%\begin{proof}
%Let $\pi$ be the true distribution from which the outcome $x^{t}$ is drawn. Let $K(\pi||\qvec)$ be the relative entropy between the true distribution $\pi$ and the predictive distribution determined by the securities vector $\qvec$. Then the expected change in budget for a trader $i$ in round $t$ is given by
%\begin{eqnarray*}
%E_{x,\qvec}[\alpha_{i}^{t}&-&\alpha_{i}^{t-1}]\\
%&=&E_{x,\qvec}[C(\tilde{\qvec}_{i-1}^{t})-C(\tilde{\qvec}_{i}^{t})-(\tilde{\qvec}_{i-1}^{t}-\tilde{\qvec}_{i}^{t})^{T}\Phi(x^{t})]\\
%&=& E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t})] + H(\pi) - E[K(\pi||\tilde{\qvec}_{i}^{t}) + H(\pi)]]
%\end{eqnarray*}
%Thus, it is sufficient to prove that E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t}] >= E_{\qvec}[K(\pi||\tilde{\qvec}_{i}^{t}].

%In the case that

%It is well known that minimizing the relative entropy of a predictive distribution with respect to the true distribution is equivalent to maximizing the likelihood function. % (see, for instance, \cite{whom?}). 
%In other words, the MLE of the natural parameters of an exponential family distribution minimizes the relative entropy of a predictive distribution with respect to the true distribution.

%Further, for a true distribution $\pi$ and predictive distribution determined by $\qvec$, $K(\pi||\qvec) + H(\pi)$ is a convex function of $\qvec$. In fact, it is strictly convex when the representation is minimal. Thus, it has a unique minimum. %Further, $K(\pi||\rho) + H(\pi)$ is non-negative.

%Suppose the current market state is represented by $\tilde{\qvec}_{i-1}$ and $\qvec_{i}$ represents the belief of trader $i$. Since the trader is budget limited, he moves the market state to $\tilde{\qvec}_{i}$ instead where $\tilde{\qvec}_{i}=  \lambda \qvec_{i}+ (1-\lambda)\tilde{\qvec}_{i-1}$. For an informative trader $i$, $\qvec_{i}$ represents the MLE of the natural parameters of the distribution. Thus, $K(\pi||\qvec) + H(\pi)$ has a unique minimum at $\qvec_{i}$.

%Thus,   from Lemma \ref{lem:convex}, $E[K(\pi||\tilde{\qvec}_{i-1}^{t})+ H(\pi)] > E[K(\pi||\tilde{\qvec}_{i}^{t})+ H(\pi)]$ and trader $i$'s expected budget increases.
%\end{proof}

For continuous distributions with a density, the probability that a trader with private information will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs. 
This result suggests that, eventually, every informative trader will have the ability to influence the market state in accordance with his beliefs, without being budget limited.

We note one important aspect of Theorem~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution. This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure; depending on the distribution family and the prior distribution over parameter values, maximum likelihood estimation might not optimize the true expected score. If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
%To summarize our results for this in a market with budget-limited traders, the total loss induced by malicious traders is limited, while the budget of  informative traders grows.
\section{The MaxEnt Market Mechanism: A Bayesian View}
\subsubsection{Conjugate Priors}

%In the previous section it's assumed that the agent has a belief $\q$, and moves the market state from initial state $\q_{init}$ to a convex combination $\tilde{\q} = \lambda \q + (1-\lambda)\q_{init}$. This note tries to justify this based on Bayesian updating, rather than budget limitations. (But it doesn't quite succeed.)

\medskip\noindent
Let $p(x;\theta)$ denote a probability density drawn from an exponential family with sufficient statistic $\phi : \mX \rightarrow \bR^d$, where $\theta$ is the natural parameter:
$$
p(x;\theta) = \exp \left[ \la \theta, \phi(x) \ra \right],
$$
and
$$
g(\theta) = \log \int_{\mX} \exp\la\phi(x),\theta\ra d\!x.
$$
Recall that $\grad g(\theta) = \Exp[\phi(x)]$ and $\grad^2 g(\theta) = \Var[\phi(x)]$. The family of conjugate priors is also an exponential family and takes the form
$$
p(\theta; n,\nu) = \exp\left[ \la n\nu,\theta \ra - ng(\theta) - h(\nu,n) \right].
$$
Here the feature map is $\psi(\theta) = (\theta, -g(\theta))$, the natural parameter is $(n\nu, n)$ where $n \in \bR$ and $\nu \in \bR^d$. The normalizer $h(\nu, n)$ is convex in $(n\nu,n)$. It is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$. The justification for this is that
$$
\Exp_{\theta} \left[ \Exp_x \left[ \phi(x) | \theta \right]\right] 
= \Exp_{\theta} \left[\grad g(\theta)\right]
= \nu.
$$
%(The proof is straightforward but not obvious. I have a reference for this but it's impossible to read---it's a mathematical statistics paper.)

Suppose we draw a sample $X = (x_1,\ldots,x_m)$ of size $m$, and denote the empirical mean by $\mu[X] = \sum_{i=1}^m \phi(x_i)$. The posterior distribution is then
$$
p(\theta|X) \propto p(X|\theta)p(\theta|n,\nu) \propto \exp\left[ \la\mu[X]+n\nu,\theta\ra - (m+n)g(\theta) \right],
$$
and so the posterior mean is 
\begin{equation} \label{posterior-mean}
\frac{m\mu[X] + n\nu}{m+n}.
\end{equation}
Thus the posterior mean is a convex combination of the prior and posterior means, and their relative weights depend on the phantom and empirical sample sizes.

In the context of prediction markets, one could imagine an agent who uses the current market estimate as its prior, and draws a empirical sample of size $m$. Its belief then takes the form~(\ref{posterior-mean}), where $n$ depends on the importance the agent places on the market estimate (perhaps based on how long the market has been running). However, note that the \emph{mean parameter} becomes a convex combination of market state and empirical belief, and this does not translate to the \emph{natural parameter}, which is what we would have liked. The new natural parameter is
$$
\grad g^{-1} \left(\frac{m\mu[X] + n\nu}{m+n}\right) = \grad g^* \left(\frac{m\mu[X] + n\nu}{m+n}\right).
$$
where $g^*$ is the convex conjugate of $g$.

\subsubsection{Exponential Utility}

Assume the agent's belief distribution $p$ belongs to an exponential family, so it takes the form
%
\[
p(x;\theta) = \exp[\theta x - T(\theta)]
\]
%
where $\theta$ is the natural parameter and $T$ is the log-partition function. (I'm assuming that the sufficient statistic is $\phi(x) = x$ just for simplicity.) Assume also that the agent has an exponential utility for money $w$:
%
\[
U(w) = -\frac{1}{a} \exp(-aw).
\]
%
Here $a$ is the coefficient of risk aversion (higher means more risk averse, and the utility function is more concave).

%
\begin{proposition}
An agent with exponential family belief with natural parameter $\htheta$, and exponential utility with coefficient $a$, makes a trade that moves the current market share vector $\theta$ to the convex combination $\frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$ assuming an LMSR cost function.
\end{proposition}
%
%
\begin{proof}
Let $\delta$ be the vector of shares the agent trades. The payoff given eventual outcome $x$ is then $\delta x - C(\delta + \theta) + C(\theta)$. The utility for this payoff is as follows (recall that $\htheta$ is the agent's believed natural parameter).
%
\[ U( \delta x - C(\delta + \theta) + C(\theta) ) = -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)).
\]
%
Taking the expected utility, we obtain
%
\begin{eqnarray*}
&   & \Exp\left[ U( \delta x - C(\delta + \theta) + C(\theta) ) \right] \\
& = & \int_{\mX} -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)) \exp[\htheta x - T(\htheta)]\, dx \\
& = & -\frac{1}{a} \int_{\mX} \exp[(\htheta -a \delta) x + a C(\delta + \theta) - a C(\theta)) - T(\htheta)]\, dx \\
& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)] \int_{\mX} \exp[(\htheta -a \delta) x - T(\htheta -a \delta)]\, dx \\
& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)] \int_{\mX} p(x; \htheta -a \delta)\, dx \\
& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)]\\
& = & U\left(- C(\delta + \theta) + C(\theta)) - \frac{1}{a} T(\htheta - a\delta) + \frac{1}{a} T(\htheta)\right)
\end{eqnarray*}
%
The second-last equality follows from the fact that $\int_{\mX} p(x; \htheta -a \delta)\, dx = 1$. Since utility $U$ is monotone increasing, it is maximized by maximizing its argument, which is a concave function of $\delta$ by convexity of $C$ and $T$. The optimality condition for the argument is
%
\begin{equation} \label{eq:optim}
\grad C(\delta^* + \theta) = \grad T(\htheta - a\delta^*)
\end{equation}
%
Now if the market maker is using LMSR, then $C$ is the log-partition function of the corresponding exponential family and $C = T$. Then~(\ref{eq:optim}) can be solved by equating the arguments. This leads to $\delta^* = (\htheta - \theta) / (1+a)$, which moves the share vector to $\theta + \delta^* = \frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$.
\end{proof}
%

In the statement of the result I use the term ``LMSR cost function'' somewhat loosely, because we are not necessarily dealing with a market over exhaustive, mutually exclusive outcomes. What is meant is the cost function that arises by taking the dual to entropy of the maxent distribution with given mean parameter $\mu$. As we've discussed this seems like the right generalization of LMSR to arbitrary mean parameter spaces.

Note that as $a \rightarrow 0$, we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$. As $a$ grows (agent grows more risk averse) the agent makes smaller and smaller trades that keep it closer to the current estimate $\theta$.

So there is a kind of congruence between exponential family beliefs and exponential utility (as the names would suggest). 
\subsection{Properties of the Maximum Entropy Market}
Consider the dual of the cost function, $C^{*}(\mu)$ defined as
$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
Thus, the mean parameters are the dual variables to the natural parameters.

\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
\proof
To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
\begin{eqnarray*}
-H(p) &=& \int_{x}p(x)\log p(x)\\
&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
\end{eqnarray*}
\unproof
\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
\rmk The LMSR is essentially a special case applied to a multinomial distribution. The LMSR is known to have bounded ($\log n$) market maker loss. Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can. 
\unrmk

{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
\proof
Note that, for an exponential family distribution
\begin{eqnarray*}
H(p) &=& -\int_{x}p(x)\log p(x)dx\\
&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
\end{eqnarray*}
If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
\unproof

%{\cor The exponential family market maker has unbounded worst case loss.}
%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
\paragraph{Loss of the market maker}
First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.

Recall the definition of the conjugate dual:
$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
So we have 
$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)

Let $q_{0}$ be the initial and $q_{f}$ the final market state. Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
\begin{eqnarray*}
\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
\end{eqnarray*}

\subsubsection{Example: Gaussian Markets}
{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution. In this case,  $\beta$ is a $2$-dimensional  vector. Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$. So 
\begin{eqnarray*}
\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
\end{eqnarray*}
Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
{\ex
We will now derive the expression for the differential entropy for the normal distribution. Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
\begin{eqnarray*}
H(p) &=& -\int_{x}p(x)\log p(x)dx\\
&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
\end{eqnarray*}
Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution. Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}

{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
Let's now work out the loss for the Gaussian market maker. First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have 
\begin{eqnarray*}
D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
\end{eqnarray*}

So the Gaussian market maker loss is 
%parallels to quadratic market maker loss? 
\begin{eqnarray*}
D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
\end{eqnarray*}¥
Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk

\section{Connection to Mirror Descent?}

\section{Conclusions}

\end{document}

### 2014-01-18 WriteUp/rep_trades.tex
note on repeated trades

\documentclass[12pt]{article}

\usepackage{amsfonts}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[oldstylenums]{kpfonts}
\usepackage[ruled]{algorithm2e}

\usepackage{hyperref}
\usepackage{enumerate}
\usepackage{fancyhdr}

\textwidth6.5in

\setlength{\topmargin}{0in} \addtolength{\topmargin}{-\headheight}
\addtolength{\topmargin}{-\headsep}
\textheight8.5in
\setlength{\oddsidemargin}{0in}

\oddsidemargin  0.0in \evensidemargin 0.0in \parindent0em

\pagestyle{fancy}\lhead{} \rhead{\bf \thepage}
\chead{{\Large\textsc{Exponential Family Prediction Markets}}} \lfoot{} \rfoot{} \cfoot{}
\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{con}[thm]{Conjecture}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{ex}[thm]{Example}
\newtheorem{qn}[thm]{Question}
\newcommand{\defn}{\bigbreak\noindent{\bf Definition. }}
\newcommand{\undefn}{\bigbreak}
\newcommand{\rmk}{\bigbreak\noindent{\bf Remark. }}
\newcommand{\unrmk}{\bigbreak}
%\newcommand{\proof}{\noindent{\bf Proof. }}
\newcommand{\unproof}{\hfill$\Box$\bigbreak}
\newcommand{\defeq}{\stackrel{\mathrm{def}}{=}}
\newcommand{\holds}[1]{\stackrel{?}{#1}}
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\ie} {{\it i.e. }}
\newcommand{\eg} {{\it e.g. }}
\newcommand{\etal} {{\it et al. }}
\newcommand{\E}{\mathbb{E}}

\begin{document}
\section*{Reinterpreting repeated trades in a market}
Let us suppose that traders in this market have exponential utility \[
U(w) = -\frac{1}{a} \exp(-aw).
\]
%
Here $a$ is the coefficient of risk aversion (higher means more risk averse, and the utility function is more concave).

Let $\delta_{1}^{*}$ be the optimal vector of shares the agent decides to trade on first entering the market with exponential family belief parametrized by natural parameter $\hat{\theta}$. Thus, his belief distribution is given by the pdf $$\exp\{\hat{\theta} \phi(x) - T(\hat{\theta})\}]$$ where $T(\hat{\theta}) =\int_{\mathcal{X}}\exp\{\hat{\theta} \phi(x)\}\ dx$ is the log partition function and $\phi(x)$ are the sufficient statistics.
On a subsequent entry into this market when the market state is $\theta'$, his optimal purchase $\delta_{2}^{*}$ is given by the solution of
$$\arg\max_{\delta_{2}} \E_{x\sim \hat{\theta}} U[\mathrm{payoff}]$$
Here, the payoff given eventual outcome $x$ is  $[\delta_{1}^{*}+\delta_{2}^{*} ]\phi(x) - C(\delta_{1}^{*} + \theta) + C(\theta)- C(\delta_{2}^{*} + \theta') + C(\theta')$. Here $\theta' = \theta+\delta_{1}^{*}+\delta'$ where $\delta'$ captures the movement of the share vector by other traders in the market. 
The expected utility for this payoff is as follows.

\begin{eqnarray*}
&   & \arg\max_{\delta_{2}}\E\left[ (\delta_{1}^{*}+\delta_{2} )\phi(x) - C(\delta_{1}^{*} + \theta) + C(\theta)- C(\delta_{2} + \theta') + C(\theta') \right] \\
& = & \arg\max_{\delta_{2}}\int_{\mathcal{X}} -\frac{1}{a} [\exp(-a (\delta_{1}^{*}+\delta_{2} )\phi(x) \\
&  & + a C(\delta_{1}^{*} + \theta) -a C(\theta)+a C(\delta_{2} + \theta') -a C(\theta')) \exp\{\hat{\theta} \phi(x) - T(\hat{\theta})\}]\, dx \\
& = & \arg\max_{\delta_{2}}\int_{\mathcal{X}} -\frac{1}{a} [\exp\{-a \delta_{2} \phi(x)+a C(\delta_{2} + \theta') -a C(\theta') \\
&  & + a C(\delta_{1}^{*} + \theta) -a C(\theta)\} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})\}]\, dx \\
& = & \arg\max_{\delta_{2}}\int_{\mathcal{X}} U[N(\delta_{2},\theta')] \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}]\, dx 
\end{eqnarray*}

where we have written $-\frac{1}{a} [\exp\{-a \delta_{2} \phi(x)+a C(\delta_{2} + \theta') -a C(\theta')]$ as $U[N(\delta_{2},\theta')]$ where $N(\delta_{2},\theta')$ is the net payoff when $\delta_{2}$ shares are purchased when the current market state is $\theta'$. This is equivalent to maximizing expected utility of $N(\delta_{2},\theta')$ so long as 
$$\int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}]\, dx = c\times \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)-T(\hat{\theta}-a \delta_{1}^{*})\} \, dx$$

In other words we want the following statement to hold.
\begin{eqnarray*}
\exists c: c\times\exp\{- T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}&=&
T(\hat{\theta}-a \delta_{1}^{*})\\
&=& \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)\}\ dx
\end{eqnarray*}

where $c$ is a constant with respect to $\delta_{2}$ and $x$ (but could potentially depend on $\hat{\theta}$, $a$ and $\delta_{1}^{*}$). This is true because the right hand side of the equation is a function of only $\hat{\theta}$, $a$ and $\delta_{1}^{*}$.

Thus, optimizing a trade in a market with prior exposure $\delta_{1}^{*}$ and belief parameter $\hat{\theta}$ is equivalent to optimizing with a belief parameter $\hat{\theta}-a \delta_{1}^{*}$. Thus trades in the exponential family market can be understood as updates on the traders' beliefs.
\end{document}

### 2014-01-24 WriteUp/rep_trades.tex
additional notes on repeated trades

@@
 \newcommand{\E}{\mathbb{E}}
+\newcommand{\hmu}{\hat{\mu}}
 
 \begin{document}
-\section*{Reinterpreting repeated trades in a market}
+
+\section{Reinterpreting repeated trades in a market}
 Let us suppose that traders in this market have exponential utility \[
@@
 The expected utility for this payoff is as follows.
-
+%%
+%\[ U( \delta x - C(\delta + \theta) + C(\theta) ) = -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)).
+%\]
+%%
 \begin{eqnarray*}
@@
 
-where we have written $-\frac{1}{a} [\exp\{-a \delta_{2} \phi(x)+a C(\delta_{2} + \theta') -a C(\theta')]$ as $U[N(\delta_{2},\theta')]$ where $N(\delta_{2},\theta')$ is the net payoff when $\delta_{2}$ shares are purchased when the current market state is $\theta'$.
-This is equivalent to maximizing expected utility of $N(\delta_{2},\theta')$ so long as
+Here the first equality follows from the fact that we are taking expectation over the traders belief parameter $\hat{\theta}$ and the second equality follows simply from rearranging the factors of $\phi(x)$.
+And lastly we have written $-\frac{1}{a} [\exp\{-a \delta_{2} \phi(x)+a C(\delta_{2} + \theta') -a C(\theta')]$ as $U[N(\delta_{2},\theta')]$ where $N(\delta_{2},\theta')$ is the net payoff when $\delta_{2}$ shares are purchased when the current market state is $\theta'$.
+
+This is equivalent to maximizing expected utility of $N(\delta_{2},\theta')$, where expectation is taken with respect to an exponential family distribution over $\mathcal{X}$ parametrized by $\hat{\theta}-a \delta_{1}^{*}$, so long as
 $$\int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}]\, dx = c\times \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)-T(\hat{\theta}-a \delta_{1}^{*})\} \, dx$$
 
+%$$\int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)-\psi(\hat{\theta}-a \delta_{1}^{*})\} \, dx =1$$
+%That is
+%$$\psi(\hat{\theta}-a \delta_{1}^{*})= \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)\}\ dx$$
 In other words we want the following statement to hold.
@@
 This is true because the right hand side of the equation is a function of only $\hat{\theta}$, $a$ and $\delta_{1}^{*}$.
+Let $\Theta\defeq\hat{\theta}-a \delta_{1}^{*}  $ be the effective belief.
+Thus we have that the trader chooses his share vector as follows.
+\begin{eqnarray*}
+&&  \arg\max_{\delta_{2}}\int_{\mathcal{X}} U[N(\delta_{2},\theta')] \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}\, dx \\
+& = &\arg\max_{\delta_{2}}\int_{\mathcal{X}} U[N(\delta_{2},\theta')] \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta}-a \delta_{1}^{*})\}]\, dx\\
+& = & \arg\max_{\delta_{2}} \int_{\mathcal{X}} -\frac{1}{a} \exp\{-a \delta_{2} \phi(x)+a C(\delta_{2} + \theta') -a C(\theta')\} \exp\{\Theta\cdot \phi(x) - T(\Theta)\}\, dx\\
+& = & \arg\max_{\delta_{2}} U\left[- C(\delta_{2} + \theta') + C(\theta') - \frac{1}{a} T(\Theta - a\delta_{2}) + \frac{1}{a} T(\Theta)\right]\\
+\end{eqnarray*}
+This follows from the argument detailed in in exp-util and simple substitution.
+Further,  the maximizer is $\delta_{2}^* = (\Theta - \theta') / (1+a)$, which moves the share vector to $\theta' + \delta_{2}^* = \frac{1}{1+a} \Theta + \frac{a}{1+a} \theta'$ which is a convex combination of the effective belief and the current market state.
 
-Thus, optimizing a trade in a market with prior exposure $\delta_{1}^{*}$ and belief parameter $\hat{\theta}$ is equivalent to optimizing with a belief parameter $\hat{\theta}-a \delta_{1}^{*}$.
-Thus trades in the exponential family market can be understood as updates on the traders' beliefs.
+In other words, an exponential utility maximizing trader who has belief $\hat{\theta}$ with prior exposure $\delta$ in a market  will behave identically as an exponential utility maximizing trader with belief $\hat{\theta}-a \delta$ and no prior exposure in the market.
+Here $a$ is the utility parameter.
+This means that financial exposure can be equivalently understood as changing the privately held beliefs.
+
+Note that as the trader becomes more risk averse, the  belief is updated more aggressively.
+As $a \rightarrow 0$, the trader becomes more risk neutral and his effective belief is closer to his true belief.
+
+\subsection{Equilibrium in a market with multiple traders}
+For each trader $i$, 
+\begin{equation}\label{eq:eqbm}
+\beta_{i}-a_{i}\delta_{i}=\theta+\sum_{j=1}^{n}\delta_{j}
+\end{equation}¥
+Rewriting, we have for each trader $i$, 
+$$\frac{\beta_{i}}{a_{i}}-\delta_{i}=\frac{1}{a_{i}}(\theta+\sum_{j=1}^{n}\delta_{j})$$
+Thus,
+$$\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)-\sum_{i=1}^{n}\delta_{i}=\frac{1}{\sum_{i=1}^{n}a_{i}}\left(\theta+\sum_{j=1}^{n}\delta_{j}\right)$$
+And
+$$\sum_{j=1}^{n}\delta_{j}=
+\frac{\sum_{i=1}^{n}a_{i}\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)-\theta}{1+\sum_{i=1}^{n}a_{i}}$$
+Substituting in Equation \ref{eq:eqbm} we have the following expression for the final market state.
+$$\frac{\sum_{i=1}^{n}a_{i}}{1+\sum_{i=1}^{n}a_{i}}\left[\sum_{i=1}^{n}\left(a_{i}^{-1}\beta_{i}\right)+\theta\right]$$
 \end{document}

### 2014-01-28 WriteUp/rep_trades.tex
added notes on equilibrium; fixed proof

@@
 As $a \rightarrow 0$, the trader becomes more risk neutral and his effective belief is closer to his true belief.
+%
+%----------------------------------------------------------------------------------------
+%	BIBLIOGRAPHY
+%----------------------------------------------------------------------------------------
 
 \subsection{Equilibrium in a market with multiple traders}
-For each trader $i$, 
+We have shown that every exponential-utility maximizing trader picks the share vector $\delta_{i}$ so as to maximize the utility of $- C(\delta_{i} + \theta) + C(\theta) - \frac{1}{a_{i}} T(\beta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\beta_{i})$ where $\theta$ is the initial market state, $a_{i}$ the utility parameter and $\beta_{i}$ the belief parameter of  trader $i$.
+To maximize utility, it suffices to maximize the argument, assuming that $C()=T()=\psi()$.
+
+Now consider a market with multiple exponential-utility maximizing traders.
+We wish to compute the equilibrium state.
+Let us assume that equilibrium is reached at a market state $\theta+\sum_{i}\delta_{i}$ where $\theta$ is the initial market state before trading begins and $\delta_{i}$ is each trader's purchase.
+
+Recall the following result from game theory.
+\begin{thm}\label{thm:potl}
+Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}\defeq \delta_{1},\ldots,\delta_{i},\ldots,\delta_{n}$.
+If there exists a potential function $\phi(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=\phi(\vec{\delta})-\phi(\vec{\delta}_{-i},\delta_{i}')$$
+then when $\phi(\vec{\delta})$ is maximized, $\vec{\delta}$ is an equilibrium.
+\end{thm}¥
+
+Let $$\phi(\vec{\delta})\defeq  \psi(\theta+\sum_{i}\delta_{i})+\sum_{i}\frac{1}{a_{i}}\psi(\beta_{i}-a_{i}\delta_{i})$$
+Now $U_{i}(\vec{\delta})=- \psi(\theta+\sum_{j}\delta_{j}) + \psi(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} \psi(\beta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} \psi(\beta_{i})$.
+Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $\phi(\vec{\delta})$ for each $\delta_{i}$.
+\begin{eqnarray*}
+\nabla_{\delta_{i}}\phi(\vec{\delta})&=&\nabla\psi(\theta+\sum_{j=1}^{n}\delta_{i})-\nabla\psi(\beta_{i}-a_{i}\delta_{i})\\
+&=&0
+\end{eqnarray*}¥
+This can be achieved by equating the arguments. That is, for each trader $i$, 
 \begin{equation}\label{eq:eqbm}
@@
 Thus,
-$$\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)-\sum_{i=1}^{n}\delta_{i}=\frac{1}{\sum_{i=1}^{n}a_{i}}\left(\theta+\sum_{j=1}^{n}\delta_{j}\right)$$
+$$\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)-\sum_{i=1}^{n}\delta_{i}=\left(\theta+\sum_{j=1}^{n}\delta_{j}\right)\sum_{i=1}^{n}\frac{1}{a_{i}}$$
 And
 $$\sum_{j=1}^{n}\delta_{j}=
-\frac{\sum_{i=1}^{n}a_{i}\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)-\theta}{1+\sum_{i=1}^{n}a_{i}}$$
+\frac{\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)-\theta\sum_{i=1}^{n}\left(\frac{1}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
 Substituting in Equation \ref{eq:eqbm} we have the following expression for the final market state.
-$$\frac{\sum_{i=1}^{n}a_{i}}{1+\sum_{i=1}^{n}a_{i}}\left[\sum_{i=1}^{n}\left(a_{i}^{-1}\beta_{i}\right)+\theta\right]$$
+$$\theta+\sum_{j=1}^{n}\delta_{j}=
+\frac{\theta+\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
 \end{document}

### 2014-01-30 WriteUp/maxentpredmkts.tex
added notes on equilibrium; fixed proof
merged writeup files

@@
 \section{Introduction}
+%------------------------------------------------------------
+%Outline
+%------------------------------------------------------------
+%- exp family review
+%- maxent scoring rule
+%	- generalized lmsr
+%	- does modularity (a la Hanson) still hold?
+%- prediction market setup
+%	- MLE with risk neutrality
+%- conjugate prior market
+%	- also on hyperparam?
+%- exp-util
+%- equilibrium for repeated trades
+%- budgets and adversarial markets
+
 We have dual goals in this paper.
@@
 \section{Notation and Definitions}
-convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesi
+convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesian updates, conjugate prior.
+\section{Maximum Likelihood Scoring Rules}
+
+%%%
+%%%
+%%%
+
+\subsection{Preliminaries}
+
+We consider a state space $\mX$ which may be discrete or continuous, and a measure $\nu$ over sets of states.
+We will typically take the state space to be  $\bR$ or $\bR^d$ together with the Lebesgue measure (perhaps restricted to some subset of states), or some finite set together with the counting measure.\footnote{We will not need to make reference to the underlying $\sigma$-algebra, but to complete the specification of the measure space we take it to be the collection of Borel sets in the case of $\bR$ and $\bR^d$, and the power set in the case of a discrete set.} A \emph{probability density} is a $\nu$-measurable function $p : \mX \rightarrow \bR_+$ that integrates to 1 over $\mX$.
+
+Let $\mP$ denote the set of all probability densities over the given measure space.
+We are interested in eliciting information about some unknown density $p \in \mP$ which corresponds to the beliefs of an expert.
+The expert may be familiar with $p$ in its entirety, or only with certain properties of $p$.
+Given a set of densities $\mD \subseteq \mP$, a \emph{property}  is a real (possibly vector-valued) function $\Gamma : \mD \rightarrow \bR^k$.
+With a slight abuse of terminology we call elements of its image $\Theta = \Gamma(\mD)$ \emph{properties} and we say that $p$ has property $\theta$ when $\Gamma(p) = \theta$.
+If $\Gamma$ is one-to-one, we instead call $\Gamma$ a \emph{parametrization} and refer to the elements of its image $\Theta$ as \emph{parameters}.
+
+To be concrete, our outcome space could for instance be $\bR$ with $\nu$ the Lebesgue measure restricted to $[0,+\infty)$.
+The set $\mP$ then consists of all probability densities with support contained in the non-negative reals, and $\mD \subset \mP$ might consist of the densities of exponential distributions.
+An example of a property over $\mP$ is the mapping from a density to its mean.\footnote{The property may not be defined for all $p \in \mP$.
+In this case we implicitly restrict it to the subset of $\mP$ for which it is defined whenever its domain $\mD$ is not made explicit.} When restricted to the exponential densities $\mD$, this mapping is a parametrization.
+In much of this work, as in this example, the purpose of the measure $\nu$ is to implicitly encode bounds on the support of possible densities.
+The reason we use this approach, rather than explicitly stating bounds, is that it allows for clearer connections with maximum entropy estimation later on.
+
+Given the set of relevant densities $\mD \subseteq \mP$, which contains the expert's belief $p$, and a property $\Gamma$ over this set, the expert is asked to report $\theta = \Gamma(p)$.
+The expert is rewarded according to a \emph{scoring rule} $S : \Theta \times \mX \rightarrow \bR \cup \{-\infty\}$ which pays it $S(\htheta,x)$ according to how well its report $\htheta \in \Theta$ agrees with the eventual outcome $x \in \mX$.
+The idea is to design the scoring rule so that the expert is incentivized to report its true belief, setting $\htheta = \theta$.
+%
+\begin{definition}
+Given a property $\Gamma : \mD \rightarrow \bR^k$, a scoring rule $S$ is  \emph{proper} over domain $\mD$ if for each $\theta \in \Theta$ and $p \in \mD$ with property $\theta$, we have
+%
+\begin{equation} \label{eq-proper}
+\Exp_p[S(\theta,x)] > \Exp_p[S(\htheta,x)]
+\end{equation}
+%
+for all $\htheta \neq \theta$. If the domain is $\mD = \mP$, the set of all possible densities, then we say the scoring rule is \emph{strongly proper}.
+\end{definition}
+%
+
+\begin{itemize}
+\item Note that the inequality is strict, in the literature these rules are called \emph{strictly} proper, but we only consider strictly proper rules here so don't make this qualification.
+\item Lambert et al.'s rules are actually strongly proper, however we consider scoring rules over vector valued properties.
+Lambert et al only develop these by adding together different scoring rules for each property.
+We go further in our constructions and develop rules that are not necessarily separable.
+\item It is useful to consider rules that are not strongly proper because we might have constructions for these for properties where no strongly proper rule is known.
+\item It is important to understand the informational requirements placed on the expert.
+Under a proper scoring rule, the expert does not need to be aware of the entire density $p$ to be incentivized to report its property $\theta$, since $\theta$ is an optimal report no matter what the density might be.
+The agent only needs to agree that the density indeed lies in $\mD$.
+\item For instance, if we are considering densities with support on $[0,+\infty)$, and $\mD$ is the set of lognormal densities, then a proper scoring rule over $\mD$ for the mean does not require the agent to know anything beyond the mean (e.g., it does not need to know the variance).
+\item A strongly proper scoring rule goes further; with such a rule we do not require the expert to believe the density is drawn from some subset $\mD$, only that it agrees with the support implied by the base measure $\nu$.
+Therefore, a strongly proper scoring rule for the mean here would elicit the mean no matter what kind of density over the positive reals might apply (e.g., exponential, Pareto, lognormal).
+\item In practical scenarios this seems like an important extension, because while it might be clear which states are possible, it might be difficult to ensure the expert agrees that the probability density for states falls within some restricted family $\mD$.
+\end{itemize}
+
+%%%
+%%%
+%%%
+
+\subsection{Maximum Likelihood}
+
+%
+\begin{theorem} \label{maxlike}
+The logarithmic scoring rule defined by
+%
+\[ S(\theta, x) = \log p(x;\theta) \]
+%
+is weakly proper for $\mF$ over $\Theta$ if and only if the maximum likelihood estimate for $\theta$ is consistent.
+Furthermore, if $\mM$ is an alternative parameter space such that there exists a bijection $t : \mM \rightarrow \Theta$, then $\bar{S}(\mu, x) = S(t(\mu), x)$ is weakly proper for $\mF$ over $\mM$.
+\end{theorem}
+%
+
+%
+\begin{example}
+Suppose $X$ is supported on $[0,+\infty)$ and follows an exponential distribution with density $f(x;\lambda) = \lambda e^{-\lambda x}$ parametrized by a rate $\lambda > 0$.
+The mean $\mu$ is related to the rate via the one-to-one mapping $\mu = 1/\lambda$, so the density can alternatively be parametrized by the mean.
+According to Theorem~\ref{maxlike} the following is a weakly proper scoring rule for the mean of an exponential distribution:
+%
+\begin{equation} \label{exp-scoring}
+S(\mu, x) = -\frac{x}{\mu} - \log \mu.
+\end{equation}
+%
+To verify this, suppose the agent believes $X$ follows an exponential distribution with rate $\lambda_0$, equivalently mean $\mu_0 = 1/\lambda_0$.
+By the one-to-one relation between rate and mean, choosing $\mu$ to maximize the expected score~(\ref{exp-scoring}) is equivalent to choosing $\lambda$ to maximize $\Exp{-\lambda x + \log \lambda} = -\lambda \mu_0 + \log \lambda$.
+The latter is strictly concave in $\lambda$, and using basic calculus its unique global maximum is $\lambda^* = 1/\mu_0$, which therefore leads the agent to report $\mu = \mu_0 = 1/\lambda^*$ under scoring rule~(\ref{exp-scoring}).
+\end{example}
+%
+
+%
+\begin{example}
+Suppose that $X$ is supported on $[1,+\infty)$ and follows a Pareto distribution with density $f(x;\alpha) = \alpha/x^{\alpha+1}$ parametrized by an index $\alpha > 0$.
+The mean $\mu$ is related to the index via the one-to-one mapping $\mu = \frac{\alpha}{\alpha-1}$, so the density can alternatively be parametrized by the mean.
+Theorem~\ref{maxlike} gives the following weakly proper scoring rule for the mean of a Pareto distribution with support on $[1,+\infty)$:
+%
+\begin{equation} \label{pareto-scoring}
+S(\mu,x) = \log \frac{\mu}{\mu-1} - \left(\frac{\mu}{\mu-1} + 1\right) \log x.
+\end{equation}
+%
+To verify this, suppose the agent believes $X$ follows a Pareto distribution with index $\alpha_0$, equivalently mean $\mu_0 = \frac{\alpha_0}{\alpha_0-1}$.
+By the one-to-one relation between the index and mean, choosing $\mu$ to maximize the expected score~(\ref{pareto-scoring}) is equivalent to choosing $\alpha$ to maximize $\Exp{\log \alpha - (\alpha+1)\log x} = \log \alpha - (\alpha+1)\Exp{\log x}$.
+The latter is strictly concave in $\alpha$, and using basic calculus its unique global maximum is $\alpha^* = 1/\Exp{\log X} = \alpha_0$, where the last equality follows from the fact that $\log X$ has an exponential distribution with rate $\alpha$ when $X$ is Pareto with rate $\alpha$.
+Therefore~(\ref{pareto-scoring}) leads the agent to report $\mu = \mu_0 = \frac{\alpha^*}{\alpha^* - 1}$.
+\end{example}
+%
+
+\begin{itemize}
+\item Explain that the scoring rule from Example 2 does not elicit the mean when the agent's distribution is exponential (with change of variable so that it lies in $[1,+\infty)$. 
+\item Observe that scoring rule from Example does elicit the mean when the agent's distribution is exponential, Pareto, or in fact any other distribution with support on $[0,+\infty)$ such as the lognormal.
+\item This occurs because the scoring rule is linear in $x$, so its expectation is the same for any distribution with mean $\mu$. 
+\item Another feature to note is that under an appropriate parametrization (rate instead of mean), the scoring rule (1) is convex in the parameter being elicited. 
+\item These two observations taken together lay the groundwork for the characterization of strongly proper scoring rules for distribution statistics. 
+\item Before we proceed should note that the arguments behind Theorem~\ref{maxlike} apply not just to maximum likelihood but to any \emph{m-estimator}, which are notions from robust statistics.
+Explain what an m-estimator looks like and how there are general consistency results for those as well.
+\item The point of these alternative estimators is to be more robust to outliers in the sample, and so these kinds of estimators could translate into scoring rule more robust to inaccuracies in the agent's assessment of the underlying probability distribution or its parameters.
+(Just a hunch.)
+\end{itemize}
 \section{Maximum Entropy Market Making}
@@
 \rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
-
-\section{Connection to Mirror Descent?}
-
+\section{Generalized Log Scoring Rule}
+Suppose that an agent holds a belief in the expected value of the moments of a random variable $x$.
+The maximum entropy distribution that is consistent with these beliefs is an exponential family distribution.
+ \defn For an outcome $x$ with predicted probability $p$, the generalized log scoring rule is defined as $$S(p,x)\defeq a+b\log p(x)$$ where $a$ and $b$ are constants that we will set to $0$ and $1$ respectively in these notes.\undefn
+
+\begin{thm} The generalized log scoring rule for the maximum entropy distribution is proper.
+\end{thm}
+\proof
+Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$
+
+Note that the maximum entropy distribution is an exponential family distribution. 
+ Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.
+
+ We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
+  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
+   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
+   Or 
+\begin{equation}\label{eq:proper}
+\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
+\end{equation}
+
+   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
+   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
+   Rearranging the terms,
+   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
+   which is the first order condition on the convexity of $\psi(\cdot)$.
+ \unproof
+
+ \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters.
+That is,
+   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
+   \unrmk
+
+ \subsection{Exponential Smoothing}
+One way to model the exponential family prediction market might be through exponential smoothing.
+\begin{itemize}
+\item A \emph{time series} is a set of data points sampled at periodic instances.
+\item \emph{Smoothing} is the process of creating an approximation function based on time series that while capturing  patterns in the data is relatively insensitive to noise.
+\item Natural first step is to use moving average. But average on how many samples (say $k$)? What to do until the first $k$ samples have been received?
+\item
+Next attempt might be to use weighted moving average. Advantage is that you can give higher weight to more recent terms. But still same disadvantage as simple moving average technique.
+\item
+\emph{Exponential Smoothing} attempts to remove this disadvantage.
+\end{itemize}
+
+Exponential smoothing is defined iteratively for $\alpha\in(0,1)$ as
+\begin{eqnarray*}
+s_{1} &=& x_{0}\\
+s_{t}&=&\alpha x_{t-1}+(1-\alpha)s_{t-1}
+\end{eqnarray*}
+
+Smaller values of $\alpha$ means greater smoothing e.g., $\alpha=0$ gives a constant function ($=x_{0}$).
+Note that the selection of $x_{0}$ is important -- and increases in importance as $\alpha$.
+Some advantages over moving average: all data points matter (decreasing importance with time) and computationally only last data point need be stored.
+
+\section{Bayesian Traders}
+Suppose data is drawn from an exponential family distribution.
+Further suppose, following a Bayesian model, that each trader has some prior belief distribution over the parameters of the data distribution.
+Assuming a conjugate prior, this prior distribution is also an exponential family distribution.
+
+The data distribution is given by $\exp\{\beta\cdot\phi(x)-\psi(\beta)\}$ where $\psi()$ is the log partition function and $\phi()$ are the sufficient statistics.
+Then the conjugate prior parametrized by $b_{0}=(n\nu, n)$ is given by $\exp\{b_{0}\cdot(\beta,\psi(\beta))-\Psi(b_{0})\}$where $\Psi()$ is the corresponding log partition function.
+The posterior distribution is given by $b=(n\nu+m\hmu, n+m)$ where $\hmu$ is the empirical mean of the sufficient statistics of the $m$ data points drawn from the data distribution, $\hmu=\sum_{i=1}^{m}\phi(x_{i})$.
+We also have $\E_{\beta\sim b_{0}}\E_{x\sim \beta}[\phi(x)]=\nu$.
+
+\subsection{Linear Utility}
+
+Suppose the trader wishes to maximize his expected payoff.
+Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by $$\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]$$
+But, $\E_{\beta\sim b}\E_{x\sim \beta}[\phi(x)]=\frac{n\nu+m\hmu}{ n+m}$.
+To obtain the maximum, we set the gradient to $0$.
+Thus, we have $$\nabla C(\delta + \theta)=\frac{n\nu+m\hmu}{ n+m}$$ For an exponential family prediction market, the final market state is given by
+$$\nabla C^{*}\left(\frac{n\nu+m\hmu}{ n+m}\right)$$
+
+\subsection{Exponential Utility}
+Now suppose the trader has exponential utility, given by $$U(w) = -\frac{1}{a} \exp(-aw)$$ where $a$ is the coefficient of risk aversion.
+Suppose the trader wishes to maximize his expected utility.
+Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by
+\begin{eqnarray*}
+&&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}U\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]\\
+&=&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}-\frac{1}{a}\exp\left[ -a\delta\phi(x) +a C(\delta + \theta) -a C(\theta)\right]
+\end{eqnarray*}¥
+
+\begin{eqnarray*}
+\E_{\beta\sim b}\E_{x\sim \beta}\exp\left[-a\delta\phi(x)\right]&=&\int_{\mathcal{B}}\int_{\mathcal{X}}\exp[-a\delta\phi(x)]\, \exp\{\beta\cdot\phi(x)-\psi(\beta)\}\, dx\\
+&&\qquad\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\, d\beta\\
+&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
+&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\int_{\mathcal{X}}\exp[(\beta-a\delta)\cdot\phi(x)-\psi(\beta-a\delta)]\, dx\\
+&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
+&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\\
+\end{eqnarray*}¥
+
+\section{Reinterpreting repeated trades in a market}
+Let us suppose that traders in this market have exponential utility \[
+U(w) = -\frac{1}{a} \exp(-aw).
+\]
+%
+Here $a$ is the coefficient of risk aversion (higher means more risk averse, and the utility function is more concave).
+
+Let $\delta_{1}^{*}$ be the optimal vector of shares the agent decides to trade on first entering the market with exponential family belief parametrized by natural parameter $\hat{\theta}$.
+Thus, his belief distribution is given by the pdf $$\exp\{\hat{\theta} \phi(x) - T(\hat{\theta})\}]$$ where $T(\hat{\theta}) =\int_{\mathcal{X}}\exp\{\hat{\theta} \phi(x)\}\ dx$ is the log partition function and $\phi(x)$ are the sufficient statistics.
+On a subsequent entry into this market when the market state is $\theta'$, his optimal purchase $\delta_{2}^{*}$ is given by the solution of
+$$\arg\max_{\delta_{2}} \E_{x\sim \hat{\theta}} U[\mathrm{payoff}]$$
+Here, the payoff given eventual outcome $x$ is  $[\delta_{1}^{*}+\delta_{2}^{*} ]\phi(x) - C(\delta_{1}^{*} + \theta) + C(\theta)- C(\delta_{2}^{*} + \theta') + C(\theta')$.
+Here $\theta' = \theta+\delta_{1}^{*}+\delta'$ where $\delta'$ captures the movement of the share vector by other traders in the market.
+The expected utility for this payoff is as follows.
+%%
+%\[ U( \delta x - C(\delta + \theta) + C(\theta) ) = -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)).
+%\]
+%%
+\begin{eqnarray*}
+&   & \arg\max_{\delta_{2}}\E\left[ (\delta_{1}^{*}+\delta_{2} )\phi(x) - C(\delta_{1}^{*} + \theta) + C(\theta)- C(\delta_{2} + \theta') + C(\theta') \right] \\
+& = & \arg\max_{\delta_{2}}\int_{\mathcal{X}} -\frac{1}{a} [\exp(-a (\delta_{1}^{*}+\delta_{2} )\phi(x) \\
+&  & + a C(\delta_{1}^{*} + \theta) -a C(\theta)+a C(\delta_{2} + \theta') -a C(\theta')) \exp\{\hat{\theta} \phi(x) - T(\hat{\theta})\}]\, dx \\
+& = & \arg\max_{\delta_{2}}\int_{\mathcal{X}} -\frac{1}{a} [\exp\{-a \delta_{2} \phi(x)+a C(\delta_{2} + \theta') -a C(\theta') \\
+&  & + a C(\delta_{1}^{*} + \theta) -a C(\theta)\} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})\}]\, dx \\
+& = & \arg\max_{\delta_{2}}\int_{\mathcal{X}} U[N(\delta_{2},\theta')] \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}]\, dx 
+\end{eqnarray*}
+
+Here the first equality follows from the fact that we are taking expectation over the traders belief parameter $\hat{\theta}$ and the second equality follows simply from rearranging the factors of $\phi(x)$.
+And lastly we have written $-\frac{1}{a} [\exp\{-a \delta_{2} \phi(x)+a C(\delta_{2} + \theta') -a C(\theta')]$ as $U[N(\delta_{2},\theta')]$ where $N(\delta_{2},\theta')$ is the net payoff when $\delta_{2}$ shares are purchased when the current market state is $\theta'$.
+
+This is equivalent to maximizing expected utility of $N(\delta_{2},\theta')$, where expectation is taken with respect to an exponential family distribution over $\mathcal{X}$ parametrized by $\hat{\theta}-a \delta_{1}^{*}$, so long as
+$$\int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}]\, dx = c\times \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)-T(\hat{\theta}-a \delta_{1}^{*})\} \, dx$$
+
+%$$\int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)-\psi(\hat{\theta}-a \delta_{1}^{*})\} \, dx =1$$
+%That is
+%$$\psi(\hat{\theta}-a \delta_{1}^{*})= \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)\}\ dx$$
+In other words we want the following statement to hold.
+\begin{eqnarray*}
+\exists c: c\times\exp\{- T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}&=&
+T(\hat{\theta}-a \delta_{1}^{*})\\
+&=& \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)\}\ dx
+\end{eqnarray*}
+
+where $c$ is a constant with respect to $\delta_{2}$ and $x$ (but could potentially depend on $\hat{\theta}$, $a$ and $\delta_{1}^{*}$).
+This is true because the right hand side of the equation is a function of only $\hat{\theta}$, $a$ and $\delta_{1}^{*}$.
+Let $\Theta\defeq\hat{\theta}-a \delta_{1}^{*}  $ be the effective belief.
+Thus we have that the trader chooses his share vector as follows.
+\begin{eqnarray*}
+&&  \arg\max_{\delta_{2}}\int_{\mathcal{X}} U[N(\delta_{2},\theta')] \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}\, dx \\
+& = &\arg\max_{\delta_{2}}\int_{\mathcal{X}} U[N(\delta_{2},\theta')] \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta}-a \delta_{1}^{*})\}]\, dx\\
+& = & \arg\max_{\delta_{2}} \int_{\mathcal{X}} -\frac{1}{a} \exp\{-a \delta_{2} \phi(x)+a C(\delta_{2} + \theta') -a C(\theta')\} \exp\{\Theta\cdot \phi(x) - T(\Theta)\}\, dx\\
+& = & \arg\max_{\delta_{2}} U\left[- C(\delta_{2} + \theta') + C(\theta') - \frac{1}{a} T(\Theta - a\delta_{2}) + \frac{1}{a} T(\Theta)\right]\\
+\end{eqnarray*}
+This follows from the argument detailed in in exp-util and simple substitution.
+Further,  the maximizer is $\delta_{2}^* = (\Theta - \theta') / (1+a)$, which moves the share vector to $\theta' + \delta_{2}^* = \frac{1}{1+a} \Theta + \frac{a}{1+a} \theta'$ which is a convex combination of the effective belief and the current market state.
+
+In other words, an exponential utility maximizing trader who has belief $\hat{\theta}$ with prior exposure $\delta$ in a market  will behave identically as an exponential utility maximizing trader with belief $\hat{\theta}-a \delta$ and no prior exposure in the market.
+Here $a$ is the utility parameter.
+This means that financial exposure can be equivalently understood as changing the privately held beliefs.
+
+Note that as the trader becomes more risk averse, the  belief is updated more aggressively.
+As $a \rightarrow 0$, the trader becomes more risk neutral and his effective belief is closer to his true belief.
+%
+%----------------------------------------------------------------------------------------
+%	BIBLIOGRAPHY
+%----------------------------------------------------------------------------------------
+
+\subsection{Equilibrium in a market with multiple traders}
+We have shown that every exponential-utility maximizing trader picks the share vector $\delta_{i}$ so as to maximize the utility of $- C(\delta_{i} + \theta) + C(\theta) - \frac{1}{a_{i}} T(\beta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\beta_{i})$ where $\theta$ is the initial market state, $a_{i}$ the utility parameter and $\beta_{i}$ the belief parameter of  trader $i$.
+To maximize utility, it suffices to maximize the argument, assuming that $C()=T()=\psi()$.
+
+Now consider a market with multiple exponential-utility maximizing traders.
+We wish to compute the equilibrium state.
+Let us assume that equilibrium is reached at a market state $\theta+\sum_{i}\delta_{i}$ where $\theta$ is the initial market state before trading begins and $\delta_{i}$ is each trader's purchase.
+
+%Thus, each trader picks $\delta_{i}$ as follows.
+%\begin{eqnarray*}
+%&&\arg\max_{\delta_{i}}\left(- \psi(\theta+\sum_{i}\delta_{i}) + \psi(\theta) - \frac{1}{a_{i}} \psi(\beta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} \psi(\beta_{i})\right)\\
+%&\iff&\arg\min_{\delta_{i}}\left( \psi(\theta+\sum_{i}\delta_{i})+\sum_{i}\frac{1}{a_{i}}\psi(\beta_{i}-a_{i}\delta_{i})\right)
+%\end{eqnarray*}¥
+Recall the following result from game theory.
+\begin{thm}\label{thm:potl}
+Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}\defeq \delta_{1},\ldots,\delta_{i},\ldots,\delta_{n}$.
+If there exists a potential function $\phi(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=\phi(\vec{\delta})-\phi(\vec{\delta}_{-i},\delta_{i}')$$
+then when $\phi(\vec{\delta})$ is maximized, $\vec{\delta}$ is an equilibrium.
+\end{thm}¥
+
+Let $$\phi(\vec{\delta})\defeq  \psi(\theta+\sum_{i}\delta_{i})+\sum_{i}\frac{1}{a_{i}}\psi(\beta_{i}-a_{i}\delta_{i})$$
+Now $U_{i}(\vec{\delta})=- \psi(\theta+\sum_{j}\delta_{j}) + \psi(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} \psi(\beta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} \psi(\beta_{i})$.
+Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $\phi(\vec{\delta})$ for each $\delta_{i}$.
+\begin{eqnarray*}
+\nabla_{\delta_{i}}\phi(\vec{\delta})&=&\nabla\psi(\theta+\sum_{j=1}^{n}\delta_{i})-\nabla\psi(\beta_{i}-a_{i}\delta_{i})\\
+&=&0
+\end{eqnarray*}¥
+This can be achieved by equating the arguments. That is, for each trader $i$, 
+\begin{equation}\label{eq:eqbm}
+\beta_{i}-a_{i}\delta_{i}=\theta+\sum_{j=1}^{n}\delta_{j}
+\end{equation}¥
+Rewriting, we have for each trader $i$, 
+$$\frac{\beta_{i}}{a_{i}}-\delta_{i}=\frac{1}{a_{i}}(\theta+\sum_{j=1}^{n}\delta_{j})$$
+Thus,
+$$\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)-\sum_{i=1}^{n}\delta_{i}=\left(\theta+\sum_{j=1}^{n}\delta_{j}\right)\sum_{i=1}^{n}\frac{1}{a_{i}}$$
+And
+$$\sum_{j=1}^{n}\delta_{j}=
+\frac{\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)-\theta\sum_{i=1}^{n}\left(\frac{1}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
+Substituting in Equation \ref{eq:eqbm} we have the following expression for the final market state.
+$$\theta+\sum_{j=1}^{n}\delta_{j}=
+\frac{\theta+\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
+Let us modify the cost function so that it includes a liquidity parameter:
+$$C_{\lambda}(q)\defeq\lambda \psi(q/\lambda)$$
+
+\subsubsection{Lower Bound on Expected Payoff for Exponential Utility Traders}
+Let $\theta$ be the current market state.
+We have shown that an exponential utility trader with belief distribution  parametrized by $\beta$ will move the market state to $\theta'=\frac{1}{1+a} \beta + \frac{a}{1+a} \theta$.
+Therefore, the trader's expected net payoff is given by
+\begin{eqnarray*}
+&& \E_{x\sim P_{\beta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
+&=& \psi(\theta)-\theta\E_{x\sim P_{\beta}}[\phi(x)]-(\psi(\theta')-\theta' \E_{x\sim P_{\beta}}[\phi(x)])\\
+&=& \psi(\theta)-\theta\nabla\psi(\beta)-(\psi(\theta')-\theta' \nabla\psi(\beta))\\
+&=& \psi(\theta)-\psi(\beta)-\nabla\psi(\beta)(\theta-\beta)-(\psi(\theta')-\psi(\beta)- \nabla\psi(\beta)(\theta'-\beta))\\
+&=& D_{\psi}(\theta,\beta)-D_{\psi}(\theta',\beta)\\
+&\geq&\frac{1}{a}D_{\psi}(\theta',\beta)\geq0
+\end{eqnarray*}
+The second to last inequality holds since $D_{\psi}(\theta',\beta)$ is convex in $\theta'$ and we have:
+\begin{eqnarray*}
+D_{\psi}(\theta',\beta)&=&D_{\psi}\left(\frac{1}{1+a} \beta + \frac{a}{1+a} \theta,\beta\right)\\
+&\leq&\frac{1}{1+a}D_{\psi}( \beta ,\beta)+ \frac{a}{1+a} D_{\psi}( \theta ,\beta)\\
+&=&\frac{a}{1+a} D_{\psi}( \theta ,\beta)
+%\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
+\end{eqnarray*}¥
+Thus, a trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{\psi}(\theta',\beta)$.
 \section{Conclusions}

### 2014-01-31 WriteUp/Original/ExpFly.tex
tidied up main folder to just have submission doc

--- 2013-12-16 WriteUp/maxentpredmkts.tex
@@
-% Adjusted to add \usepackage[numbers]{natbib}
-% \bibliographystyle{acmsmall} and \documentclass[prodmode,acmec]{ec-acmsmall}
-% Jan 5, 2013 - David Parkes
-%
-% v2-acmsmall-sample.tex, dated March 6 2012
-% This is a sample file for ACM small trim journals
-%
-% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
-% (c) 2010 Association for Computing Machinery (ACM)
-%
-% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
-% Users can also go through the FAQs available on the journal's submission webpage.
-%
-% Steps to compile: latex, bibtex, latex latex
-%
-% For tracking purposes => this is v1.3 - March 2012
+\documentclass{article}
+%\usepackage{mystyle}
 
-\documentclass[prodmode,acmec]{ec-acmsmall} % Aptara syntax
-
-% Package to generate and customize Algorithm as per ACM style
-\usepackage[ruled]{algorithm2e}
-\renewcommand{\algorithmcfname}{ALGORITHM}
-\SetAlFnt{\small}
-\SetAlCapFnt{\small}
-\SetAlCapNameFnt{\small}
-\SetAlCapHSkip{0pt}
-\IncMargin{-\parindent}
-
-% Metadata Information
-\acmVolume{9}
-\acmNumber{4}
-\acmArticle{39}
-\acmYear{2013}
-\acmMonth{6}
-\usepackage[numbers]{natbib}
 \usepackage{amsmath}
 \usepackage{amssymb}
-%\usepackage{amsthm}
-\usepackage{parskip}
-\usepackage{booktabs}
+\usepackage{times}
+\usepackage{graphics}
+\usepackage{enumerate}
+\usepackage{amstext}
 
-%new commands
-\newcommand{\la}{\langle}
-\newcommand{\ra}{\rangle}
-\newcommand{\mX}{\mathcal{X}}
-\newcommand{\bR}{\mathbf{R}}
-\newcommand{\grad}{\nabla}
-\newcommand{\Exp}{\mathbf{E}}
-\newcommand{\Var}{\mathrm{Var}}
-\newcommand{\q}{\mathbf{q}}
-\newcommand{\mM}{\ensuremath{\mathcal{M}}}
-\newcommand{\hmu}{\ensuremath{\hat{\mu}}}
-\newcommand{\mP}{\ensuremath{\mathcal{P}}}
-\newcommand{\mY}{\ensuremath{\mathcal{Y}}}
-\newcommand{\mS}{\ensuremath{\mathcal{S}}}
-\newcommand{\ds}{\displaystyle}
-\newcommand{\mF}{\mathcal{F}}
-\newcommand{\mD}{\mathcal{D}}
-\newcommand{\htheta}{\hat{\theta}}
-%---------------------------------------------------------------------------
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{prop}[thm]{Proposition}
-\newtheorem{lem}[thm]{Lemma}
-\newtheorem{con}[thm]{Conjecture}
-\newtheorem{cor}[thm]{Corollary}
-\newtheorem{ex}[thm]{Example}
-\newtheorem{qn}[thm]{Question}
-\newcommand{\defn}{\bigbreak\noindent{\bf Definition. }}
-\newcommand{\undefn}{\bigbreak}
-\newcommand{\rmk}{\bigbreak\noindent{\bf Remark. }}
-\newcommand{\unrmk}{\bigbreak}
-%\newcommand{\proof}{\noindent{\bf Proof. }}
-\newcommand{\unproof}{\hfill$\Box$\bigbreak}
+\newcommand{\xvec}{\mathbf{x}} %outcome values %skk changed to xvec from betavec
+\newcommand{\family}{\mathcal F} 
+\newcommand{\betavec}{\pmb{\beta}}
+\newcommand{\muvec}{\pmb{\mu}}
+\newcommand{\lpf}{\mathbf{\psi}} %logpartition function
+\newcommand{\suff}{\pmb{\phi}} % suff stats
+\newcommand{\eps}{\mbox{$\epsilon$}}
+\newcommand{\phantomset}{\mathbb{H}}
+\newcommand{\pexp}{\overline{e}}
+\newcommand{\aexp}{\tilde{e}}
+
+\newcommand{\hyperfamily}{\mathcal{F}^{*}}
+\newcommand{\hyperP}{P^{*}}
+\newcommand{\hyperb}{\mathbf{b}}
+\newcommand{\hyperm}{\mathbf{m}}
+\newcommand{\hyperbeta}{\pmb{\beta}^{*}}
+\newcommand{\hyperlpf}{\mathbf{\psi}^{*}}
+\newcommand{\zvec}{\mathbf{z}} %% \hyperbeta is decomposed as (r\zvec,r)
+\newcommand{\pvec}{\mathbf{p}}
+%XXX agents and data
+\newcommand{\honestset}{\mathcal{H}}
+\newcommand{\attackset}{\overline{\mathcal{H}}}
+\newcommand{\xseq}{\mathbf{x}}
+\newcommand{\att}{\mathcal{A}}
+\newcommand{\nullatt}{\mathcal{A}_\phi}
+\newcommand{\xt}{\tilde{\mathbf{x}}}
+%notation: \xt_\att(\xseq) is attacked seq
+%\newcommand{\bt}{\tilde{\mathbf{b}}} %attack vector in conj prior section
+\newcommand{\Pt}{\tilde{P}} %attack hyperdistribution in conj prior section
+%XXX algorithms and predictions
+\newcommand{\pred}{Q}
+%learning algorithm: Z
+%omniscient/optimal learning algorithm: Z_O
+%XXX losses and gains, informativeness, regret
+% L(\pred_k, \xvec_k) : loss from pred when true outcome is x
+% L(\pred_k, \betavec_k) : loss in revealed parameter model
+% G_Z(u_i, \xt, \xvec_k) :
+% G_Z (u_i, \xt): expected loss, under prior conditioned by \xseq
+\newcommand{\info}{h} %% use \info_{ik}
+\newcommand{\reg}{\mbox{Reg}} % use \reg(Z)
+\newcommand{\ureg}{\mbox{UR}} % unaccounted regret, used later;%XXX  WTM/ILS notation
+\newcommand{\Inf}[2]{y_{{#1}{#2}}} %{Y_{#1}^{(#2)}}
+\newcommand{\Infvec}{\mathbf{y}}
+\newcommand{\Gain}[2]{G_{{#1}{#2}}} %{G_{#1}^{(#2)}}
+\newcommand{\Rep}[2]{r_{{#1}{#2}}}
+\newcommand{\Iik}{\Inf{i}{k}}
+\newcommand{\Gik}{\Gain{i}{k}}
+\newcommand{\Ghik}{\hat{G}_{ik}} %estimate of unnormalized gain
+\newcommand{\rik}{\Rep{i}{k}}
+%\newcommand{\xik}{x_{ik}}
+\newcommand{\yb}{\overline{y}}
+%\newcommand{\ubar}{{\overline{u}}}
+
+% G(\yvec, u_i, \xt, \xvec_k)
+% G(y_i, u_i, \xt, \xvec_k)
+% B(r): budget associated with reputation r
+
+%XXX assorted notation
+\newcommand{\var}{\mbox{Var}}
+\newcommand{\IL}{\mbox{IL}_i} %% in its use, always refers to user i
+\newcommand{\rinit}{\Rep{i}{1}} % in its use in IL proof, always refers to user i
+
+\newcommand{\ddG}[1]{\frac{d#1}{dG}}
+\newcommand{\sdG}[1]{\frac{d^2 #1}{dG^2}}
+\newcommand{\ddg}[1]{\frac{d#1}{dg}}
+\newcommand{\sdg}[1]{\frac{d^2 #1}{dg^2}}
+\newcommand{\ddu}{\frac{d}{du}}
+\newcommand{\sdu}{\frac{d^2}{du^2}}
+
+\newcommand{\hk}{\info_{ik}} %shorthand used in info loss section
+\newcommand{\htt}{h_t}
+\newcommand{\Hk}{H_{ik}}
+\newcommand{\Ht}{H_t}
+\newcommand{\Gk}{G_k}
+\newcommand{\gk}{g_k}
+\newcommand{\gt}{g_t}
+\newcommand{\kmin}{\overline{k}}
+\newcommand{\Hmin}{\overline{H}}
+\newcommand{\HM}{H_M}
+
+\newcommand{\GFe}{\mbox{$\mathrm{GF}_\epsilon$}}
+\newcommand{\EGFe}{\mbox{$\mathrm{EGF}_\epsilon$}}
+\newcommand{\pihat}{\hat{\pi}}
+\newcommand{\qvec}{\mathbf{q}}
+\newcommand{\uvec}{\mathbf{u}}
+\newcommand{\wvec}{\mathbf{w}}
+\newcommand{\qlim}{\tilde{q}}
+\newcommand{\qbar}{\overline{q}}
+\newcommand{\Gt}{\~{G}}
+\newcommand{\seq}{\sigma}
+\newcommand{\EG}{\mbox{EG}}
+\newcommand{\avec}{\mathbf{a}}
+\newcommand{\xbar}{{\overline{x}}}
+\newcommand{\yhat}{\hat{y}}
+\newcommand{\what}{\hat{w}}
+\newcommand{\gik}{g_{ik}}
+\newcommand{\xtilde}{\tilde{x}}
+
+%margin comments in red
+\newcommand{\com}[1]{\marginpar{\textcolor{red}{#1}}}
+
+%\newcommand{\xvec}{\mathbf{x}}
+\newcommand{\yvec}{\mathbf{y}}
+\newcommand{\Gauss}{\mathcal{N}}
+\newcommand{\Xvec}{\mathbf{X}}
+%\newcommand{\qvec}{\mathbf{q}}
+\newcommand{\wvechat}{\hat{\mathbf{w}}}
+\newcommand{\Gbar}{\overline{G}}
+\newcommand{\Lmat}{\mbox{$\mathbf{L}$}}
+\newcommand{\Hmat}{\mbox{$\mathbf{H}$}}
+\newcommand{\Kmat}{\mbox{$\mathbf{K}$}}
+\newcommand{\Amat}{\mbox{$\mathbf{A}$}}
+\newcommand{\etal}{{\it et al.}}
+\newcommand{\ie}{{\it i.e.}}
+\newcommand{\eg}{{\it e.g.}}
+\newcommand\beqn{\begin{eqnarray*}}
+\newcommand\eeqn{\end{eqnarray*}}
 \newcommand{\defeq}{\stackrel{\mathrm{def}}{=}}
-\newcommand{\holds}[1]{\stackrel{?}{#1}}
-\newcommand{\conv}{\mathrm{conv}}
-\newcommand{\ie} {{\it i.e. }}
-\newcommand{\eg} {{\it e.g. }}
-\newcommand{\etal} {{\it et al. }}
-\newcommand{\E}{\mathbf{E}}
-\newcommand{\betavec}{\pmb{\beta}}
-\newcommand{\qvec}{\mathbf{q}}
-\newcommand{\lpf}{\mathbf{\psi}} %logpartition function
-\newcommand{\family}{\mathcal F} 
-\newcommand{\suff}{\pmb{\phi}} % suff stats
-\newcommand{\muvec}{\pmb{\mu}}
+\newcommand{\eqncom}[1]{\quad{\mbox{\texttt{(#1)}}}}
 
-% Document starts
+\newenvironment{proof*}{\noindent{\bf Proof:}}{}
+\newcommand{\note}[1]{\textcolor{gray}{#1}}
+\newcommand{\ccite}[1]{\textcolor{blue}{#1}}
+
+ \newtheorem{theorem}{Theorem}
+ \newtheorem{lemma}[theorem]{Lemma}
+ \newtheorem{prop}[theorem]{Proposition}
+ \newtheorem{definition}[theorem]{Definition}
+ \newtheorem{claim}[theorem]{Claim}
+ \newtheorem{corollary}[theorem]{Corollary}
+ \newcommand{\qed}{\hfill\rule{7pt}{7pt}}
+\newenvironment{proof}{\noindent{\bf Proof:}}{\qed\medskip}
+
+% If your paper is accepted, change the options for the package
+% aistats2e as follows:
+%
+%\usepackage[accepted]{aistats2e}
+%
+% This option will print headings for the title of your paper and
+% headings for the authors names, plus a copyright note at the end of
+% the first column of the first page.
 
 \begin{document}
-% Page heads
-\markboth{Abernethy et al.}{Maximum Entropy Prediction Markets}
-% Title portion
-\title{Maximum Entropy Prediction Markets}
-\author{JACOB ABERNETHY
-\affil{University of Michigan, Ann Arbor}
-SINDHU KUTTY
-\affil{University of Michigan, Ann Arbor}
-S\'{E}BASTIEN LAHAIE
-\affil{Microsoft Research, New York City}
-RAHUL SAMI
-\affil{University of Michigan, Ann Arbor}}
-%%\author{GANG ZHOU
-%%\affil{College of William and Mary}
-%%YAFENG WU
-%%\affil{University of Virginia}
-%%TING YAN
-%%\affil{Eaton Innovation Center}
-%%TIAN HE
-%%\affil{University of Minnesota}
-%%CHENGDU HUANG
-%%\affil{Google}
-%%JOHN A. STANKOVIC
-%%\affil{University of Virginia}
-%%TAREK F. ABDELZAHER
-%%\affil{University of Illinois at Urbana-Champaign}}
-% NOTE! Affiliations placed here should be for the institution where the
-%       BULK of the research was done. If the author has gone to a new
-%       institution, before publication, the (above) affiliation should NOT be changed.
-%       The authors 'current' address may be given in the "Author's addresses:" block (below).
-%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
-%       currently affiliated with NASA.
+% If your paper is accepted and the title of your paper is very long,
+% the style will print as headings an error message. Use the following
+% command to supply a shorter title of your paper so that it can be
+% used as headings.
+%
+%\runningtitle{I use this title instead because the last one was very long}
 
+% If your paper is accepted and the number of authors is large, the
+% style will print as headings an error message. Use the following
+% command to supply a shorter version of the authors names so that
+% they can be used as headings (for example, use only the surnames)
+%
+%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}
+\title{Prediction Markets for Exponential Families}
+\author{ Sindhu Kutty\\ Department of Computer Science and Engineering\\University of Michigan, Ann Arbor
+\and
+Rahul Sami \\School of Information\\University of Michigan, Ann Arbor} 
+\date\today
+\maketitle
 \begin{abstract}
-In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets.
-We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data.
-Similarly,  the market state can be interpreted as a distribution over the outcome space.
-In particular, we show that a proper scoring rule can be derived from maximum entropy distributions.
-This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces.
-In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility.
-We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function.
-These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders.
-In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market.
-Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
+  %The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
+%  both left and right-hand margins. Use 10~point type, with a vertical
+%  spacing of 11~points.  The {\bf Abstract} heading must be centered,
+%  bold, and in point size 12. Two line spaces precede the
+%  Abstract. The Abstract must be limited to one paragraph.
+ This paper identifies and develops new connections between two approaches to aggregating information from multiple sources: statistical learning techniques and prediction markets that rely on profit-seeking traders to update prices.
+ We present a new rich class of prediction markets for infinite outcome spaces that are based on parametric distribution families.
+For these markets, we prove that a sequence of traders maximizing their expected profit will result in price sequences analogous to sequential maximum likelihood estimation from statistics.
+Further, we show that even when some traders are adversarial attackers rather strategic profit-maximizers, these prediction markets continue to have attractive learning properties: The damage that the attackers can cause is bounded by their initial budgets, and unlike attackers causing damage, informative traders' budgets will grow in expectation as they trade in more markets.
 \end{abstract}
 
-\category{J.4}{Social and Behavioral Sciences}{Economics}
-\category{I.2.6} {Artificial Intelligence}{Learning}
+\section{INTRODUCTION}
+Prediction markets are aggregation mechanisms that allow market prices to be interpreted as predictive probabilities on an event.
+Each trader in the market is assumed to have some private information that he uses to make a prediction on the outcome of the event.
+The traders are assumed to be strategic in the sense that every trader wants to maximize his profit by participating in the market.
+Further, since the trades are done sequentially, the trader is allowed to observe all past trades in the market and update his private information based on this information.
 
-\terms{Prediction Market Design, Machine Learning Algorithms}
+Traders are allowed to report their beliefs on the outcome of the event by allowing them to buy and sell securities whose value depends on the outcome of this future event.
+This will effect the state of the market, thus updating the predictive probabilities for the event.
+For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins.
+Traders can see the past history of trades, so the price at which a current trader is willing to buy and sell these securities can be interpreted as an aggregate ``market probability'' of a particular candidate winning the election.
 
-\keywords{Exponential Families, Bayesian Learning, Maximum Entropy Distributions, generalized LMSR}
+Currently, most deployed prediction markets have binary or finite discrete outcome spaces.
+This means that the outcome of every event can be characterized by a finite set of mutually exclusive exhaustive outcomes.
+The securities in these market are usually tied to each outcome so that buying (respectively, selling) a security related to an outcome, indicates a belief in an increased (respectively, decreased) probability of occurrence of the outcome.
 
-\acmformat{Jacob Abernethy, Sindhu Kutty, S\'{e}bastien Lahaie, and Rahul Sami, 2014. Maximum Entropy Prediction Markets}
-%%Gang Zhou, Yafeng Wu, Ting Yan, Tian He, Chengdu Huang, John A. Stankovic, and Tarek F. Abdelzaher, 2010. A multifrequency MAC specially designed for  wireless sensor network applications.
+One form of prediction markets, which is rapidly gaining in popularity, is the market scoring rule~\cite{hanson03}.
+A market scoring rule considers all trades as a single chronological sequence.
+Traders earn rewards proportional to the incremental reduction in prediction loss caused by their trades in comparison to the previous trade.
+In other words, their rewards depend on the change in market probabilities caused by their trade, as well as on the eventual outcome.
+Thus, each trader has an
+incentive to minimize the prediction loss.
+In this format, the {\em market maker} who runs the market can suffer an overall loss, but Hanson~\cite{hanson03} showed that, for market scoring rules on finite outcome spaces, the loss of the market maker can be bounded.
 
-% At a minimum you need to supply the author names, year and a title.
-% IMPORTANT:
-% Full first names whenever they are known, surname last, followed by a period.
-% In the case of two authors, 'and' is placed between them.
-% In the case of three or more authors, the serial comma is used, that is, all author names
-% except the last one but including the penultimate author's name are followed by a comma,
-% and then 'and' is placed before the final author's name.
-% If only first and middle initials are known, then each initial
-% is followed by a period and they are separated by a space.
-% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.
+%Currently, prediction markets are predominantly designed for predicting finite outcomes.
+In \cite{Gao} the authors extend the finite outcome prediction market to an infinite outcome space.
+For the logarithmic market scoring rule (LMSR) they show that the market maker's worst case loss in unbounded for an infinite outcome space.
 
-\begin{bottomstuff}
-%%This work is supported by the National Science Foundation, under
-%%grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.
-%%
-%%Author's addresses: G. Zhou, Computer Science Department,
-%%College of William and Mary; Y. Wu  {and} J. A. Stankovic,
-%%Computer Science Department, University of Virginia; T. Yan,
-%%Eaton Innovation Center; T. He, Computer Science Department,
-%%University of Minnesota; C. Huang, Google; T. F. Abdelzaher,
-%%(Current address) NASA Ames Research Center, Moffett Field, California 94035.
-\end{bottomstuff}
+Directly applying techniques used for binary outcome spaces to large outcome spaces is computationally infeasible.
+In fact, it has been shown that computing and updating prices for combinatorial outcomes spaces is $\#$P-hard \cite{Chen08}.
+Infinite outcome spaces pose a particular challenge in prediction markets since it is not clear how to define securities in such spaces.
+The most direct way of discretizing the space may lead to loss of information from the traders.
+Extending prediction market techniques to this space has been a subject of recent interest.
+Various techniques to bound market maker loss in both a combinatorial and infinite outcome space have been studied (see, for instance, \cite{Abernethy11,OthmanS11}).
 
-\maketitle
+In this paper, we 
+use a powerful tool from statistics for handling infinite outcome spaces: the use of parametric distributions.
+Specifically, we consider markets for possibly infinite outcome spaces, with the restriction that the true distribution over 
+the outcomes belongs to an exponential family of distributions.
 
-\section{Introduction}
-We have dual goals in this paper.
-On the one hand, we highlight a structural similarity between prediction markets and exponential families.
-We see this as the syntax of our prediction market mechanism.
-On the other, this formulation has rich semantics as well: it allows for analysis of market behavior under various environments.
-For instance, we analyze market behavior with budget limited traders, traders that are exponential-utility-maximizers.
-\section{Notation and Definitions}
-convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesi
-\section{Maximum Entropy Market Making}
+For such parametric families, we show that we can design a prediction market such that, when traders are strategically maximizing their own expected profit but are not adversarial, the market state can provably be interpreted as the maximum likelihood estimate of the natural parameters of the distribution.
+Also, this market has the desirable property of being {\em arbitrage-free}: There is no trade or sequence of trades that traders can exploit to make a guaranteed profit.
 
-The purpose of a prediction market is to elicit and aggregate the beliefs (i.e., subjective probabilities) of agents over a space of \emph{outcomes}.
-In a single-agent setting, a scoring rule is used to elicit the agent's beliefs.
-In a multi-agent setting, an information market is used to aggregate the agent's beliefs.
-Hanson~\cite{} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both so that they can perform well in both thin and thick markets.
-In this work we derive a wide-ranging generalization of the logarithmic market scoring rule.
-Using a maximum entropy approach, we explain how a market scoring rule can be developed for outcome spaces both discrete and continuous, and for generic properties of the underlying distribution (e.g., mean and variance), rather than just the probabilities of individual outcomes.
+Next, we consider a more realistic situation in which their are both adversarial and strategic traders.
+We present a budget limited version of the prediction market that can be shown to suffer limited damage under concerted attacks by adversarial traders.
+We also exploit an information-theoretic interpretation of the particular form of the scoring rule that we use to show that an informative trader will eventually be unconstrained and he will be allowed to carry out an unrestricted trade.
+Taken together, we will have shown that while the market suffers limited damage from malicious traders, it is also able to make use of all the information from informative traders in the long run.
 
-%%%
-%%%
-%%%
+Apart from its value as a new form of prediction market, we believe that this connection between prediction markets and exponential families can be exploited further to design high performance online learning algorithms.
+In section~\ref{sec:conclusion}, we suggest future work in this direction.
+%Preliminary results were reported in \cite{KuttySami11}.
 
-\subsection{The Model}
-\label{model}
+\subsection{Related Work}
+Designing prediction markets to handle a large outcome space is an active area of research.
+In \cite{Chen08}, the authors use a restricted betting language to design efficient markets for a combinatorial outcome space.
+This technique is generalized by \cite{Pennock11}.
+\cite{Gao09} consider extending various automated market makers to an infinite outcome space.
+For the logarithmic market scoring rule they show that unbounded market maker loss can result in this setting.
+\cite{Abernethy11} and \cite{OthmanS11} specify frameworks under which they design cost function based markets that satisfy the desirable property of bounded market maker loss even in infinite outcome spaces.
 
-Let $\mX$ be the outcome space, which may be discrete or continuous.
-Let $\mP$ denote the set of probability distributions over the outcome space.
-We represent a probability distribution as a density $p$ absolutely continuous with respect to some base measure $\nu$ (e.g., counting for discrete outcomes, or Lebesgue for continuous outcomes).% We will discuss the interpretation of the base measure in a later section.
+The connection between machine learning and prediction markets has been studied previously.
+\cite{Chen10} and \cite{Abernethy11} have previously explored the connection to learning algorithms to inform the design and understanding of prediction markets.
+In particular, \cite{Chen10}  consider the correspondence between prediction markets with market scoring rules and the Follow the Regularized Leader algorithm proposed by \cite{Kalai05} and thus provide insight into the aggregation mechanism of a prediction market.
 
-There is an unknown distribution $p$ over outcomes.
-We are interested in estimating the expected value of different outcome \emph{statistics} under this distribution by aggregating the beliefs of agents.
-A statistic is a real-valued function $\phi_s : \mX \rightarrow \bR$ over outcomes, independent of $p$; here $s$ belongs to a finite index set $\mS$ of size $d$.
-The collection of functions $\phi = (\phi_s,\, s \in \mS)$ can be viewed as a vector-valued statistic mapping outcomes into $\bR^d$.
-Formally, the aim is to estimate
+Independently of this work, Beygelzimer et al.~\cite{BLP12} have shown that, for a particular form of binary prediction markets and traders with log-utility, the long-run dynamics of trading activity and budgets over many prediction markets lead the markets to satisfy a bounded regret property with respect to the best single trader.
+Our results here, and the future work we have suggested, form a program to prove bounded regret properties of budget-limited prediction markets under much more general conditions.
+
+\section{EXPONENTIAL FAMILIES}
+A number of popular distributions are members of a class of a distributions called the exponential family.
+For instance, exponential, beta, Bernoulli and Poisson distributions are all members of the exponential family of distributions.
+%In fact, these are specific examples of exponential family distributions that have finite support.
+We restrict ourselves to these kinds of distributions here.
+%exponential, log-normal, gamma, chi-squared, beta, Dirichlet, Bernoulli, categorical, Poisson, geometric, inverse Gaussian, von Mises
+In this work, we are concerned with outcomes that are drawn from an exponential family distribution.
+
+A member of an exponential family $\family$, $P_{\betavec} \in \family$ assigns a probability density (with respect to some base measure) over $x$ defined over a space $\mathcal{X}$ and has the following form:
+\[
+  P_{\betavec}(x) = e^{\betavec . \suff(x) -\lpf(\betavec)}
+\]
+Note that $\mathcal{X}$ could be a multi-dimensional space.
+
+Here, the set of natural parameters $\betavec$ is drawn from some space $\mathcal{B}$. $\lpf(\betavec)$ is known as the {\em log-partition function} and is defined as 
+ $$\lpf(\betavec)\defeq\log\int e^{\betavec . \suff(x)}\mbox{d}x$$
+
+The vector of sufficient statistics $\suff(x)$ is some function of the data and is defined differently for different members of the exponential family.
+
+For example, the Bernoulli distribution is a discrete probability distribution that is a member of the exponential family.
+It takes value 1 with  probability $p$ and value 0 with  probability $1-p$.
+In this case, the vector of sufficient statistics,  natural parameters and log-partition function is given by:
+\begin{eqnarray*}
+\suff(x)&=&x\\
+\betavec&=&\log\frac{p}{1-p}\\
+\lpf(\betavec)&=&-\log(1-p)
+\end{eqnarray*}
+
+%Let us consider the univariate Gaussian distribution as a particular example. The univariate Gaussian is defined as 
+%$$\mathcal{N}_{\mu,\sigma^2}(x)= \frac{1}{\sqrt{2\pi \sigma^2}}\exp\left\{-\frac{(x-\mu)^2}{2\sigma^2}\right\}$$
 %
-\begin{equation} \label{constraint}
-\Exp_p[\phi(x)] = \mu.
-\end{equation}
-%
-We stress that we are only seek to elicit $\mu \in \bR^d$, not the full distribution $p$.
-As an example, suppose that $\mX = \bR$.
-If we were interested in eliciting the first two uncentered moments of $p$, we would include $\phi_1(x) = x$ and $\phi_2(x) = x^2$ as statistics.
-Note that the variance cannot be directly obtained through any single statistic: it corresponds to the expectation of $(x - \Exp_p[x])$, which depends on $p$, violating our definition of a statistic.
-Of course, the variance can be indirectly obtained by eliciting the first two moments separately.\footnote{This is related to the notion of \emph{elicitation complexity} of statistical properties introduced by Lambert et al.~\cite{Lambert}, but we do not pursue this connection here.}
+%In this case, the vector of sufficient statistics,  natural parameters and log-partition function is 
+%\begin{eqnarray*}
+%\suff(x)&=&\begin{pmatrix} x\\ x^2\end{pmatrix}\\
+%\betavec&=&\begin{pmatrix} \beta_1\\ \beta_2\end{pmatrix}=\begin{pmatrix} \frac{\mu}{\sigma^2}\\ \frac{-1}{2\sigma^2}\end{pmatrix}\\
+%\lpf(\betavec)&=& \sqrt{\frac{-\beta_2}{\pi}} \exp\left(\frac{\beta_1^2}{4\beta_2}\right)
+%\end{eqnarray*}
 
-There are two approaches to eliciting an agent's estimated $\mu$ in~(\ref{constraint}).
-The first approach is to incentivize the agent to directly report $\mu$ via a \emph{scoring rule}.
-The second approach is to form an \emph{information market} by issuing securities for each statistic $s \in \mS$, with payoff depending on the realized value of the statistic.
-(Such securities are known as contingent claims.) The amount of shares of each security acquired by the agent indirectly reveals it estimate $\mu$.
-In a market scoring rule~\cite{Hanson}, these two approaches are dual to each other in a formal sense.
+\section{PREDICTION MARKETS}
+A prediction market is a market where traders bet on the outcome of an event and are paid according to the accuracy of their forecast.
+The current market state indicates the net prediction on the probability of an event.
+Let $i\in\{1,\ldots,N\}$ be the possible (mutually exclusive, exhaustive) outcomes of the event.
+Let $\pvec$ indicate the current market state.
+Traders in the market influence the value of $\pvec$ and hence the predictive probability of the event.
 
-%%%
-%%%
-%%%
+Market scoring rules were developed as a way to allow sequential participation in a prediction market.
+In the Log Market Scoring Rule (LMSR), a trader who moves the current market estimate of an actual outcome $i$ from $p_{i-1}$ to $p_i$ receives a (possibly negative) payoff of $\log\frac{p_i}{p_{i-1}}$.
 
-\subsection{Scoring Rule}
+In a cost-function based prediction market, a trader pays for the movement of the market probability upfront \cite{hanson03}.
+This movement is associated with the purchase of securities.
+The cost of purchase is determined by a cost function.
+The trader is then paid according to a payoff function.
+%Both of these functions are non-negative.
+Ideally, the cost function is such that it covers any possible loss that the traders may incur on the trade, thereby preventing any bankruptcy problems at the time of payoff settlement.
 
-We first consider how to directly elicit $\mu$ via a scoring rule. 
-%Note that the usual definition of a scoring rule presumes that a full probability distribution is elicited, rather than just some of its properties (e.g., mean and variance), so below we will adapt the definition to our setting.
-The space of feasible reports from an agent corresponds to
-%
-\begin{equation} \label{marginal}
-\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}.
-\end{equation}
-%
-A scoring rule is a function $S : \mM \times \mX \rightarrow \bR$ that rewards an agent with $S(\mu,x)$ based on how its report $\mu$ agreed with the eventual outcome $x$.
-A scoring rule is \emph{proper} if for all $\mu \in \mM$, and $p \in \mP$ such that $\Exp_p[\phi(x)] = \mu$, we have
-%
-\begin{equation} \label{scoring}
-\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)]
-\end{equation}
-%
-for all $\hmu \neq \mu$; the rule is \emph{strictly proper} if the inequality is strict.
-Note that in the literature, scoring rules are defined over entire probability distributions, not just properties of those distributions.
-Our definition imposes a minimum of informational requirements on agents, because an agent does not need to assess the full distribution $p$ that might have lead to its estimates $\mu$ in order to realize that truthful reporting is an optimal strategy.
+LMSR can also be specified as a cost-function based prediction market.
+Let $\{s_{i}\}$ be the securities issued in this market.
+Then $q_{i}$ represents the number of securities of type $s_{i}$ held by all the traders in the market and $\qvec$ is the vector of all securities held.
+The cost function $C(\qvec)$ is a differentiable function that depends on the market state $\qvec$.
+A trader pays $C(\qvec')-C(\qvec)$ to move the market probabilities from $\qvec$ to $\qvec'$.
+The partial derivative of $C(\qvec)$ with respect to $q_{i}$ represents the instantaneous price of buying an infinitesimally small quantity of security $s_{i}$ at the current market state $\qvec$.
+Typically, the payoff of each security is fixed at $\$1$ for an outcome  that occurs.
 
-Our scoring rule is implicitly defined via the solution of a mathematical program. The program computes the maximum entropy distribution consistent with the agent's reported beliefs $\mu$.
-%
-\begin{eqnarray}
-\max_{p \geq 0} & \ds - \int_{x \in \mX} p(x) \log p(x) \nu(dx) & \label{obj} \\
-\mbox{s.t.} & \ds \int_{x \in \mX} \phi(x) p(x) \nu(dx) = \mu & \label{mean} \\
- & \ds \int_{x \in \mX} p(x) \nu(dx) = 1 & \label{normalize}
-\end{eqnarray}
-%
-The non-negativity constraints together with~(\ref{normalize}) ensure that the solution is a probability distribution.
-Constraints~(\ref{mean}), which correspond to $d$ separate constraints, ensure consistency with the agent's report.
-The objective~(\ref{obj}) is the entropy of the solution $p$ with respect to the measure $\nu$.
-The program has a convex objective and linear constraints.
-%For continuous outcome spaces we will see in the next section that the optimal solution can still be cleanly characterized.
-%
-\begin{theorem} \label{thm:scoring}
-Let $p(x;\mu)$ be the optimal solution to the maximum entropy program, given report $\mu \in \bR^d$. The scoring rule defined by
-%
-\begin{equation} \label{scoring}
-S(\mu,x) = a_x + b\log p(x;\mu),
-\end{equation}
-%
-where $b,a_x \in \bR$ and $b > 0$, is strictly proper.
-\end{theorem}
-%
-Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$
+A necessary condition for the market to be arbitrage-free is that the sum of the instantaneous price of all securities be exactly equal to the payoff of each security (in this case $\$1$) for all states of the market.
+To see this we note that if this sum is less than 1 (respectively, greater than 1), then a trader could buy (respectively, sell) a bundle of securities for a guaranteed profit.
+Thus, arbitrage-freeness is a desirable property of a market.
+%For infinite outcome space, \cite{Gao} have proposed a modification to the finite outcome LMSR.
+We modify this definition to fit our particular purposes.
+We will show that this modified definition has some desirable properties.
 
-Note that the maximum entropy distribution is an exponential family distribution. 
- Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.
+We will now formalize the cost-function based market for a general payoff function $\Phi(x)$.
+Let $x$ be an event whose outcome is to be determined.
+Let us define securities $s_i$ whose payoff is given by $\phi_{i}(x)$; let $\Phi(x)$ be the payoff vector.
+Let $q_i$ be the  number of $s_i$ securities purchased; let  $\qvec$ be the securities vector.
 
- We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
-  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
-   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
-   Or 
-\begin{equation}\label{eq:proper}
-\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
-\end{equation}
+For a purchase of $\qvec$ securities, the cost function for LMSR is defined as 
+$$C(\qvec)\defeq \log \int e^{\qvec^{T}\Phi(x)}\ dx$$
+and the instantaneous price of the securities is given by $$p_i(\qvec)\defeq\frac{\partial C(\qvec)}{\partial q_i}$$
+Note that a perfectly rational, risk-neutral trader with infinite budget will purchase securities until the instantaneous price of the securities is equal to the payoff he believes he will obtain.
 
-   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
-   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
-   Rearranging the terms,
-   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
-   which is the first order condition on the convexity of $\psi(\cdot)$.
- \unproof
-
- \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters.
-That is,
-   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
-   \unrmk
-
-\noindent
-This rule represents a generalization of the logarithmic scoring rule for probability distributions to any properties of distributions that can be captured as expectations of statistics.
-As explained in the next section, the solution to the maximum entropy program takes the form of an \emph{exponential family} distribution.
-For many common properties of interest, these distributions are familiar; for example, the maximum entropy distribution with a given mean is an exponential distribution, and the maximum entropy distribution with a given mean and variance is a normal distribution~\cite{}.
-We provide here two examples in more depth.
-
-%%
-%%
-
-\paragraph{Multinomial Distribution}
-
-As a first example, consider the problem of estimating the individual probabilities of a finite set of outcomes.
-We have $\mX = \{1,\ldots,k\}$.
-The relevant statistics indicate which outcome actually occurs, so the index set is $\mS = \mX$.
-Statistic $\phi_x(x')$ is 1 if $x' = x$ and 0 otherwise.
-Observe that if $p$ is the distribution over outcomes, then $\Exp_p[\phi(x)] = p$.
-A feasible report from an agent is any non-negative $\pi = (\pi_1,\ldots,\pi_k)$ such that $\sum_{i=1}^k \pi_i = 1$.
-Given a report of $\mu = \pi$, the unique solution to the maximum entropy program is $p(x; \pi) = \pi$, and~(\ref{scoring}) corresponds to the classic logarithmic scoring rule of $S(\pi,x) = \log \pi_x$.
-\hfill{$\Box$}
-
-\paragraph{Normal Distribution}
-
-As a second example, suppose the outcome space is $\mX = \bR$ and that we are interested in estimating the mean $\mu$ and variance $\sigma^2$ of the underlying distribution.
-We choose the first and second (uncentered) moments, $\kappa_1 = \mu$ and  $\kappa_2 = \mu^2 + \sigma^2$ as our statistics.
-It is well-known that the maximum entropy distribution with a given mean and variance is the normal distribution, so the solution to the maximum entropy program (using Lebesgue as the base measure $\nu$) is
-%
-\[ p(x; \kappa) = \frac{1}{\sqrt{2\pi\sigma^2}} \exp\left[ -\frac{(x-\mu)^2}{2\sigma^2} \right]. \]
-%
-Applying Theorem~\ref{thm:scoring} and choosing appropriate constants, we find that the following scoring rule is strictly proper in this context. %
-\[ S(\kappa, x) = -\frac{(x-\mu)^2}{\sigma^2} - \log \sigma^2. \]
-%
-Note that the fact that this scoring rule is proper for the mean amounts to the well-known fact that reporting the mean is a Bayes act under quadratic loss~\cite{someone}. \hfill{$\Box$}
-
-\medskip\noindent
-A brief observation on parametrizations is in order.
-Note that the maximum entropy distribution here is parametrized by the uncentered moments $\kappa = (\kappa_1,\kappa_2)$ rather than $(\mu,\sigma^2)$ because the variance is not a valid statistic.
-However, it is clear that the agent does not have to directly report $\kappa$: the market maker itself can do the translation from reported mean and variance to uncentered moments.
-This is relevant because the $\kappa$ parametrization may be unintuitive for an agent, especially given the constraints $\kappa_2 - \kappa_1^2 \geq 0$ that must be enforced.
-In the $(\mu, \sigma^2)$ parametrization each parameter is unrestricted, and the parameters have intuitive interpretations.
-
-Our approach so far has been to identify relevant properties of the unknown distribution, and elicit them via a scoring rule computed through a maximum entropy program.
-However, it is the case that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}.
-Therefore, we see the following alternative approach.
-The market designer can begin by considering the outcome space $\mX$, and assessing what family of distributions over $\mX$ the unknown $p$ might come from.
-If this family is an exponential family, then Theorem~\ref{thm:scoring} immediately gives a proper scoring rule to elicit the parameters of the distribution.
-Many of the most familiar distributions are exponential families, including the normal, exponential, gamma, beta, binomial, Poisson, Weibull, and Dirichlet.
-To illustrate, we consider two examples in more depth.
-
-%%
-%%
-
-\paragraph{Exponential}
-Suppose the designer would like an estimate of the rate at which servers fail in a data center.
-An outcome is the time between failures, so $\mX = [0,+\infty)$.
-The usual distribution for failure times is the exponential distribution $p(x;\lambda) = \lambda e^{-\lambda x}$, which is parametrized by the rate $\lambda$ and has mean $\mu = 1/\lambda$.
-The exponential distribution is the maximum entropy distribution given a fixed mean $\mu$.
-From Theorem~\ref{thm:scoring} we obtain the following proper scoring rule for this setting:
-%
-\[ S(\mu, x) = \log \lambda - \lambda x. \]
-%
-It is simple to check using basic calculus that an agent maximizes the expected score by reporting $\lambda = 1/\hmu$, where here $\hmu$ is its estimate of the mean.
-In this case, it is unclear a priori which of the mean or rate parametrizations is most intuitive, but the maximum entropy approach easily allows for either.
-\hfill{$\Box$}
-
-%%
-%%
-
-\paragraph{Beta}
-Suppose the designer would like an estimate of the click-through rate of an online advertisement.
-The outcome space here is $\mX = (0,1)$.
-A typical prior over a probability like this is the beta distribution, which has two parameters $\alpha, \beta > 0$ and density function $p(x; \alpha, \beta) = x^{\alpha-1}(1-x)^{\beta-1}/B(\alpha,\beta)$, where $B$ is the beta function.
-By theorem~\ref{thm:scoring} this leads to the proper scoring rule
-%
-\[ S((\alpha, \beta), x) = (\alpha-1)\log x + (\beta-1)\log (1-x) - \log B(\alpha,\beta). \]
-%
-Although this parametrization is standard for the beta, it is perhaps more intuitive to use the parametrization $\mu = \alpha/(\alpha+\beta) \in (0,1)$ and $n = \alpha+\beta > 0$.
-Here $\mu$ is the mean of the distribution---corresponding to the agent's actual estimate of the click-through rate---and $n$ is a measure of the agent's confidence in the estimate.
-\hfill{$\Box$}
-
-%%%
-%%% 
-%%%
-
-\subsection{Information Market}
-
-We next consider how to indirectly elicit $\mu$ by setting up a market of contingent claim securities.
-Under this approach, the elements of the index set $\mS$ are interpreted not as statistics but as securities.
-The payoff from one share of security $s$ when outcome $x$ occurs is given by the mapping $\phi_s$.
-Thus if the vector of shares held by the agent is $\theta \in \bR^d$, where entry $\theta_s$ corresponds the number of shares of security $s$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \theta, \phi(x) \ra$.
-As a concrete example, recall our multinomial distribution example from the previous section, where $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise.
-This means that security $x \in \mS$ pays 1 dollar if outcome $x \in \mX$ occurs, and nothing otherwise.
-(Such securities are known as Arrow-Debreu securities.) In our normal distribution example, the statistic for the mean had the mapping $\phi_1(x) = x$.
-Therefore a share of the corresponding security has a payoff that is linear in the outcome.
-(Such securities amount to futures contracts.)
-
-To extract useful information from such a securities market, we set a pricing scheme and examine the number of shares the agent chooses to acquire.
-Let $\Omega$ be the set of all portfolios (i.e., vectors of shares) that the agent can feasibly hold; we will see how this domain is determined in an instant.
-Each security $s$ has an associated price function $c_s : \Omega \rightarrow \bR$, which gives the marginal price $c_s(\theta)$ of security $s$ when the agent holds portfolio $\theta$.
-(Note that the marginal price depends on the entire portfolio $\theta$, not just the number of shares $\theta_s$.) A risk-neutral agent will choose to acquire shares up to the point where, for each share, expected payoff equals marginal price.
-Formally, if the agent acquires portfolio $\theta$, then for each $s \in \mS$ we must have
-%
-\begin{equation} \label{market}
-\Exp_p[\phi_s(x)] = c_s(\theta).
-\end{equation}
-%
-In this way, by its choice of $\theta$, the agent reveals that its belief is $\mu = c(\theta)$.
-There are several important properties that the price function $c$ should have.
-To simplify the agent's portfolio acquisition process, it should not be the case that the total cost of acquiring $\theta$ depends on the order in which shares are bought.
-This means that there should exist a \emph{cost function} $C : \Omega \rightarrow \bR$ such that $c = \nabla C$.
-The gradient $\nabla C$ must be onto $\mM$ to ensure that~(\ref{market}) always has a solution---this is the most important but technical condition to realize.
-To ensure that the solution is unique, it would also be convenient if the cost function were strictly convex.
-%
-\begin{proposition}
-The following cost function is monotone, convex, and onto $\mM$: 
-%
-\begin{equation} \label{cost}
-C(\theta) = \log \int_{x \in \mX} \exp \left[\la \theta, \phi(x) \ra\right] \nu(dx).
-\end{equation}
-%
-\end{proposition}
-%
-\noindent
-Observe that in the context of our earlier multinomial example, cost function~(\ref{cost}) is exactly the cost function for Hanson's logarithmic market scoring rule.
-Our approach here provides a generalization of this market scoring rule to markets with contingent claim securities with arbitrary payoffs, designed to elicit specific properties of distributions, beyond just the probabilities of different outcomes.
-
-Because an agent would never select a portfolio with infinite cost, the effective domain of $C$ is $\Omega = \{\theta \:|\: C(\theta) < +\infty \}$.
-In the context of the multinomial distribution example, $\Omega = \bR^d$, so that shares of each security can always be bought or sold short.
-In the context of the normal distribution example the effective domain is $\Omega = \{(\theta_1,\theta_2) \in \bR^2 \:|\: \theta_2 > 0 \}$ if we re-define the second statistic to be $\phi_2(x) = -x^2$; this means that the security has a negative payoff for each share, and consequently the agent must be compensated to acquire such shares.
-Evaluating~(\ref{cost}), the cost function takes the form $C(\theta) = \frac{\theta_1^2}{4\theta_2} - \frac{1}{2}\log(2\theta_2)$.
-
-Now, in our previous elicitation approach that used a scoring rule, the process of computing the scoring rule also provided a complete distribution $p(x;\mu)$ over outcomes, which could be used to infer other properties beyond just the agent beliefs $\mu$.
-In the current market-based approach, the agent's chosen portfolio $\theta$ can also form the basis of a distribution over outcomes.
-Consider the following distribution, represented as a density with respect to a base measure $\nu$:
-%
-\begin{equation} \label{expfam}
-p(x;\theta) = \exp\left[ \la \theta, \phi(x) \ra - C(\theta) \right].
-\end{equation}
-%
-Meaning, the probability that a subset $X \subseteq \mX$ of the outcomes occurs is $\int_{x \in X} p(x;\theta)\nu(dx)$.
-Observe that by definition~(\ref{cost}), this probability density indeed integrates to 1 over $\mX$, and it is clear that the density is non-negative as required.
-
-In the statistical literature a distribution that takes the form~(\ref{expfam}) is known as an \emph{exponential family}.
-The mapping $\phi$ is known as the \emph{sufficient statistic}, $\theta$ is the \emph{natural parameter}, and $C$ is the log-partition or \emph{cumulant} function.
-%This connection between information markets and exponential families allows us to draw on a vast statistical literature to understand the market's properties, and a growing computer science literature to apply efficient algorithms to compute costs and perform inference.
-
-%%%
-%%%
-%%%
-
-\subsection{Duality}
-
-There is a well-known duality between the maximum entropy approach and exponential families, which translates into a duality between the scoring rule and information market just developed.
-The duality implies that the approach leads to a \emph{market scoring rule}, applicable to both thin and thick multi-agent settings.
-
-It is known that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}.
-To see this, let $\theta(\mu) \in \bR^d$ be the Lagrange multiplier corresponding to constraints~(\ref{mean}) when solving the maximum entropy program given the agent report $\mu$.
-(Our choice of notation is deliberately suggestive.) Let $A(\mu)$ be the Lagrange multiplier corresponding to~(\ref{normalize}).
-From the first order necessary conditions for optimality, we find that
-%
-\[ p(x;\mu) = \exp \left[ \la \theta(\mu), \phi(x) \ra - A(\mu) - 1 \right], \]
-%
-so that the solution is indeed in exponential family form.
-Conversely, given a cumulant $C$ from an exponential family and a parameter $\theta$, we see that $p(x;\theta)$ defined according to~(\ref{expfam}) is the maximum entropy distribution under constraints $\Exp_p[\phi(x)] = \mu$ where $\mu = \nabla C(\theta)$.
-We obtain the following.
-%
-\begin{proposition}
-Let $C$ be the cumulant (i.e, cost function) for the exponential family corresponding to $\phi$, and let $\theta(\mu)$ be the optimal Lagrange multiplier for the mean constraints given an agent report of $\mu \in \mM$.
-Then the following scoring rule is equivalent to~(\ref{scoring}):
-%
-\begin{equation} \label{expscore}
-S(\mu, x) = \la \theta(\mu), \phi(x) \ra - C(\theta(\mu)).
-\end{equation}
-%
-\end{proposition}
-%
-\noindent
-The scoring rule~(\ref{expscore}) decomposes neatly into payoff and cost functions.
-The first term $\la \theta(\mu), \phi(x) \ra$ defines the agent's outcome-contingent reward, while the second term~$C(\theta(\mu))$ is the cost of acquiring a portfolio $\theta(\mu)$.
-We see here the duality between the approach of directly reporting $\mu$, or acquiring shares $\theta$: an agent reporting beliefs $\mu$ under scoring rule~(\ref{scoring}) would choose to acquire shares $\theta(\mu)$ in the information market with cost function~(\ref{cost}).
-Since~(\ref{scoring}) is a proper scoring rule, the information market with cost function~(\ref{cost}) is based on a proper market scoring rule.
-
-\section{The Exponential Family Market Mechanism: A Maximum Likelihood Approach}
+\section{AGGREGATING BELIEFS OVER A CONTINUOUS OUTCOME SPACE}
 In this section, we will set up a prediction market that aggregates beliefs from traders where the outcome is a continuous random variable.
@@
 
-\subsection{The Exponential Family Market Mechanism in Adversarial Markets}
+\section{AGGREGATING BELIEFS OVER A CONTINUOUS OUTCOME SPACE UNDER ADVERSARIAL CONDITIONS}
 In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
@@
 
-\subsubsection{Budget-limited Aggregation}
+\subsection{Budget-limited Aggregation}
 Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
@@
 Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\qvec$.
-\subsubsection{Damage Bound}
+\subsection{Damage Bound}
 Given this restriction on the initial budget of traders, we can now show that the total loss that can be induced by malicious traders is bounded.
@@
 
-\subsubsection{An Information-Theoretic Interpretation}\label{sec:infth}
+\subsection{An Information-Theoretic Interpretation}\label{sec:infth}
  We observe a useful alternative view of the market 
@@
 
-\subsubsection{Budget of Informative Traders}\label{sec:inf}
+\subsection{Budget of Informative Traders}\label{sec:inf}
 Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
@@
 %To summarize our results for this in a market with budget-limited traders, the total loss induced by malicious traders is limited, while the budget of  informative traders grows.
-\section{The MaxEnt Market Mechanism: A Bayesian View}
-\subsubsection{Conjugate Priors}
 
-%In the previous section it's assumed that the agent has a belief $\q$, and moves the market state from initial state $\q_{init}$ to a convex combination $\tilde{\q} = \lambda \q + (1-\lambda)\q_{init}$.
-This note tries to justify this based on Bayesian updating, rather than budget limitations.
-(But it doesn't quite succeed.)
+\section{CONCLUSIONS AND FUTURE WORK}\label{sec:conclusion}
+%\subsection{Non-Myopic Attacks}
+We have defined a new class of cost-function based prediction markets for a continuous random variable distributed according to an exponential family.
+Our first result shows that if traders in this market believe sufficient statistics drawn from true distribution, then the market price provably represents the MLE of natural parameters.
+Under some restrictions on the distribution, the market maker will also have limited worst case loss.
+We then show that if we can budget limit the traders so that their influence on the market is constrained by their budgets, malicious traders can induce only bounded loss.
+We also show that eventually informative agents will be able to fully influence the market state in accordance with their beliefs.
 
-\medskip\noindent
-Let $p(x;\theta)$ denote a probability density drawn from an exponential family with sufficient statistic $\phi : \mX \rightarrow \bR^d$, where $\theta$ is the natural parameter:
-$$
-p(x;\theta) = \exp \left[ \la \theta, \phi(x) \ra \right],
-$$
-and
-$$
-g(\theta) = \log \int_{\mX} \exp\la\phi(x),\theta\ra d\!x.
-$$
-Recall that $\grad g(\theta) = \Exp[\phi(x)]$ and $\grad^2 g(\theta) = \Var[\phi(x)]$. The family of conjugate priors is also an exponential family and takes the form
-$$
-p(\theta; n,\nu) = \exp\left[ \la n\nu,\theta \ra - ng(\theta) - h(\nu,n) \right].
-$$
-Here the feature map is $\psi(\theta) = (\theta, -g(\theta))$, the natural parameter is $(n\nu, n)$ where $n \in \bR$ and $\nu \in \bR^d$.
-The normalizer $h(\nu, n)$ is convex in $(n\nu,n)$.
-It is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$.
-The justification for this is that
-$$
-\Exp_{\theta} \left[ \Exp_x \left[ \phi(x) | \theta \right]\right] 
-= \Exp_{\theta} \left[\grad g(\theta)\right]
-= \nu.
-$$
-%(The proof is straightforward but not obvious. I have a reference for this but it's impossible to read---it's a mathematical statistics paper.)
+This last result provides a technique that can be applied to constructing bounded regret learning algorithms.
+Online learning algorithms based on prediction markets are attractive for particular domains.
+Specifically, expert forecasts and recommendations in internet settings present the following challenges: Some forecasters may provide best-effort forecasts, but in some cases the experts may have vested interests in manipulating the system.
+Second, advice from different experts often arrive haphazardly over time allowing later experts to clone the predictions of earlier experts.
+Current expert advice algorithms cannot distinguish between genuine forecasters and clones, even though the clones are forced to make forecasts later than the genuine forecasters they copy.
+We believe that prediction markets can  be naturally applied to the sequential forecasting setting to design high-performance algorithms in this domain.
 
-Suppose we draw a sample $X = (x_1,\ldots,x_m)$ of size $m$, and denote the empirical mean by $\mu[X] = \sum_{i=1}^m \phi(x_i)$. The posterior distribution is then
-$$
-p(\theta|X) \propto p(X|\theta)p(\theta|n,\nu) \propto \exp\left[ \la\mu[X]+n\nu,\theta\ra - (m+n)g(\theta) \right],
-$$
-and so the posterior mean is 
-\begin{equation} \label{posterior-mean}
-\frac{m\mu[X] + n\nu}{m+n}.
-\end{equation}
-Thus the posterior mean is a convex combination of the prior and posterior means, and their relative weights depend on the phantom and empirical sample sizes.
+However, a scoring rule that only considers the impact of the expert on the predicted probability distribution on the target variable is bound to fail in the following way.
+An adversarial entity may control multiple experts that clone the reports of every honest expert.
+Consider such an expert who arrives after some honest expert in the sequence.
+Cloning the honest report will not lead to an increase in his score since he hasn't affected the prediction.
+However, although his imitation of this expert may not immediately lead to an expected increase in his score, he effectively reduces the impact on the prediction of any honest and informative expert arriving later in the sequence.
+Thus, if the adversarial entity now clones this later honest expert, he can siphon off some of the increase in budget due to her.
 
-In the context of prediction markets, one could imagine an agent who uses the current market estimate as its prior, and draws a empirical sample of size $m$.
-Its belief then takes the form~(\ref{posterior-mean}), where $n$ depends on the importance the agent places on the market estimate (perhaps based on how long the market has been running).
-However, note that the \emph{mean parameter} becomes a convex combination of market state and empirical belief, and this does not translate to the \emph{natural parameter}, which is what we would have liked.
-The new natural parameter is
-$$
-\grad g^{-1} \left(\frac{m\mu[X] + n\nu}{m+n}\right) = \grad g^* \left(\frac{m\mu[X] + n\nu}{m+n}\right).
-$$
-where $g^*$ is the convex conjugate of $g$.
+This attack is possible because the experts are scored based on their impact on the prediction.
+Further, they are not penalized if they adversely affect the impact of later (honest) experts in the sequence, so long as they do not move the prediction.
+Thus, this possible non-myopic attack needs to be considered in the assigned scores.
+We leave the details to future work.
 
-\subsubsection{Exponential Utility}
+%\subsubsection*{Acknowledgements}
+%We would like to thank Michael Wellman for insightful discussions.
+%Use unnumbered third level headings for the acknowledgements.  All
+%acknowledgements go at the end of the paper.  Be sure to omit any
+%identifying information in the initial double-blind submission!
+\bibliographystyle{apalike} %default bib style
+\bibliography{ExpFly}
+%\subsubsection*{References}
 
-Assume the agent's belief distribution $p$ belongs to an exponential family, so it takes the form
-%
-\[
-p(x;\theta) = \exp[\theta x - T(\theta)]
-\]
-%
-where $\theta$ is the natural parameter and $T$ is the log-partition function.
-(I'm assuming that the sufficient statistic is $\phi(x) = x$ just for simplicity.) Assume also that the agent has an exponential utility for money $w$:
-%
-\[
-U(w) = -\frac{1}{a} \exp(-aw).
-\]
-%
-Here $a$ is the coefficient of risk aversion (higher means more risk averse, and the utility function is more concave).
+%References follow the acknowledgements.  Use an unnumbered third level
+%heading for the references section.  Any choice of citation style is
+%acceptable as long as you are consistent.  Please use the same font
+%size for references as for the body of the paper---remember that
+%references do not count against your page length total.
 
-%
-\begin{proposition}
-An agent with exponential family belief with natural parameter $\htheta$, and exponential utility with coefficient $a$, makes a trade that moves the current market share vector $\theta$ to the convex combination $\frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$ assuming an LMSR cost function.
-\end{proposition}
-%
-%
-\begin{proof}
-Let $\delta$ be the vector of shares the agent trades.
-The payoff given eventual outcome $x$ is then $\delta x - C(\delta + \theta) + C(\theta)$.
-The utility for this payoff is as follows (recall that $\htheta$ is the agent's believed natural parameter).
-%
-\[ U( \delta x - C(\delta + \theta) + C(\theta) ) = -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)).
-\]
-%
-Taking the expected utility, we obtain
-%
-\begin{eqnarray*}
-&   & \Exp\left[ U( \delta x - C(\delta + \theta) + C(\theta) ) \right] \\
-& = & \int_{\mX} -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)) \exp[\htheta x - T(\htheta)]\, dx \\
-& = & -\frac{1}{a} \int_{\mX} \exp[(\htheta -a \delta) x + a C(\delta + \theta) - a C(\theta)) - T(\htheta)]\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)] \int_{\mX} \exp[(\htheta -a \delta) x - T(\htheta -a \delta)]\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)] \int_{\mX} p(x; \htheta -a \delta)\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)]\\
-& = & U\left(- C(\delta + \theta) + C(\theta)) - \frac{1}{a} T(\htheta - a\delta) + \frac{1}{a} T(\htheta)\right)
-\end{eqnarray*}
-%
-The second-last equality follows from the fact that $\int_{\mX} p(x; \htheta -a \delta)\, dx = 1$.
-Since utility $U$ is monotone increasing, it is maximized by maximizing its argument, which is a concave function of $\delta$ by convexity of $C$ and $T$.
-The optimality condition for the argument is
-%
-\begin{equation} \label{eq:optim}
-\grad C(\delta^* + \theta) = \grad T(\htheta - a\delta^*)
-\end{equation}
-%
-Now if the market maker is using LMSR, then $C$ is the log-partition function of the corresponding exponential family and $C = T$.
-Then~(\ref{eq:optim}) can be solved by equating the arguments.
-This leads to $\delta^* = (\htheta - \theta) / (1+a)$, which moves the share vector to $\theta + \delta^* = \frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$.
-\end{proof}
-%
-
-In the statement of the result I use the term ``LMSR cost function'' somewhat loosely, because we are not necessarily dealing with a market over exhaustive, mutually exclusive outcomes.
-What is meant is the cost function that arises by taking the dual to entropy of the maxent distribution with given mean parameter $\mu$.
-As we've discussed this seems like the right generalization of LMSR to arbitrary mean parameter spaces.
-
-Note that as $a \rightarrow 0$, we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$.
-As $a$ grows (agent grows more risk averse) the agent makes smaller and smaller trades that keep it closer to the current estimate $\theta$.
-
-So there is a kind of congruence between exponential family beliefs and exponential utility (as the names would suggest). 
-\subsection{Properties of the Maximum Entropy Market}
-Consider the dual of the cost function, $C^{*}(\mu)$ defined as
-$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
-This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
-$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
-Thus, the mean parameters are the dual variables to the natural parameters.
-
-\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
-\proof
-To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
-\begin{eqnarray*}
--H(p) &=& \int_{x}p(x)\log p(x)\\
-&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
-&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
-&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
-\end{eqnarray*}
-\unproof
-\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
-\rmk The LMSR is essentially a special case applied to a multinomial distribution.
-The LMSR is known to have bounded ($\log n$) market maker loss.
-Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
-\unrmk
-
-{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
-\proof
-Note that, for an exponential family distribution
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
-&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
-\end{eqnarray*}
-If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
-\unproof
-
-%{\cor The exponential family market maker has unbounded worst case loss.}
-%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
-%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
-%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
-\paragraph{Loss of the market maker}
-First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
-
-Recall the definition of the conjugate dual:
-$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
-The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
-$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
-This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
-So we have 
-$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
-Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
-
-Let $q_{0}$ be the initial and $q_{f}$ the final market state.
-Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
-\begin{eqnarray*}
-\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
-&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
-&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
-&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
-&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
-&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
-&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
-&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
-&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
-\end{eqnarray*}
-
-\subsubsection{Example: Gaussian Markets}
-{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
-In this case,  $\beta$ is a $2$-dimensional  vector.
-Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
-So
-\begin{eqnarray*}
-\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
-&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
-&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
-\end{eqnarray*}
-Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
-$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
-$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
-{\ex
-We will now derive the expression for the differential entropy for the normal distribution.
-Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
-&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
-\end{eqnarray*}
-Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
-Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
-
-{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
-Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
-Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
-Let's now work out the loss for the Gaussian market maker.
-First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
-\begin{eqnarray*}
-D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
-&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
-&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
-%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
-&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
-\end{eqnarray*}
-
-So the Gaussian market maker loss is 
-%parallels to quadratic market maker loss? 
-\begin{eqnarray*}
-D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
-&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
-&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
-\end{eqnarray*}¥
-Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
-\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
-
-\section{Connection to Mirror Descent?}
-
-\section{Conclusions}
+%J.~Alspector, B.~Gupta, and R.~B.~Allen (1989). Performance of a
+%stochastic learning microchip.  In D. S. Touretzky (ed.), {\it
+%  Advances in Neural Information Processing Systems 1}, 748-760.  San
+%Mateo, Calif.: Morgan Kaufmann.
 

### 2014-01-31 WriteUp/Original/exp-util.tex
tidied up main folder to just have submission doc

--- 2013-12-16 WriteUp/maxentpredmkts.tex
@@
-% Adjusted to add \usepackage[numbers]{natbib}
-% \bibliographystyle{acmsmall} and \documentclass[prodmode,acmec]{ec-acmsmall}
-% Jan 5, 2013 - David Parkes
-%
-% v2-acmsmall-sample.tex, dated March 6 2012
-% This is a sample file for ACM small trim journals
-%
-% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
-% (c) 2010 Association for Computing Machinery (ACM)
-%
-% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
-% Users can also go through the FAQs available on the journal's submission webpage.
-%
-% Steps to compile: latex, bibtex, latex latex
-%
-% For tracking purposes => this is v1.3 - March 2012
-
-\documentclass[prodmode,acmec]{ec-acmsmall} % Aptara syntax
-
-% Package to generate and customize Algorithm as per ACM style
-\usepackage[ruled]{algorithm2e}
-\renewcommand{\algorithmcfname}{ALGORITHM}
-\SetAlFnt{\small}
-\SetAlCapFnt{\small}
-\SetAlCapNameFnt{\small}
-\SetAlCapHSkip{0pt}
-\IncMargin{-\parindent}
-
-% Metadata Information
-\acmVolume{9}
-\acmNumber{4}
-\acmArticle{39}
-\acmYear{2013}
-\acmMonth{6}
-\usepackage[numbers]{natbib}
-\usepackage{amsmath}
-\usepackage{amssymb}
-%\usepackage{amsthm}
-\usepackage{parskip}
-\usepackage{booktabs}
-
-%new commands
-\newcommand{\la}{\langle}
-\newcommand{\ra}{\rangle}
-\newcommand{\mX}{\mathcal{X}}
-\newcommand{\bR}{\mathbf{R}}
-\newcommand{\grad}{\nabla}
-\newcommand{\Exp}{\mathbf{E}}
-\newcommand{\Var}{\mathrm{Var}}
-\newcommand{\q}{\mathbf{q}}
-\newcommand{\mM}{\ensuremath{\mathcal{M}}}
-\newcommand{\hmu}{\ensuremath{\hat{\mu}}}
-\newcommand{\mP}{\ensuremath{\mathcal{P}}}
-\newcommand{\mY}{\ensuremath{\mathcal{Y}}}
-\newcommand{\mS}{\ensuremath{\mathcal{S}}}
-\newcommand{\ds}{\displaystyle}
-\newcommand{\mF}{\mathcal{F}}
-\newcommand{\mD}{\mathcal{D}}
-\newcommand{\htheta}{\hat{\theta}}
-%---------------------------------------------------------------------------
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{prop}[thm]{Proposition}
-\newtheorem{lem}[thm]{Lemma}
-\newtheorem{con}[thm]{Conjecture}
-\newtheorem{cor}[thm]{Corollary}
-\newtheorem{ex}[thm]{Example}
-\newtheorem{qn}[thm]{Question}
-\newcommand{\defn}{\bigbreak\noindent{\bf Definition. }}
-\newcommand{\undefn}{\bigbreak}
-\newcommand{\rmk}{\bigbreak\noindent{\bf Remark. }}
-\newcommand{\unrmk}{\bigbreak}
-%\newcommand{\proof}{\noindent{\bf Proof. }}
-\newcommand{\unproof}{\hfill$\Box$\bigbreak}
-\newcommand{\defeq}{\stackrel{\mathrm{def}}{=}}
-\newcommand{\holds}[1]{\stackrel{?}{#1}}
-\newcommand{\conv}{\mathrm{conv}}
-\newcommand{\ie} {{\it i.e. }}
-\newcommand{\eg} {{\it e.g. }}
-\newcommand{\etal} {{\it et al. }}
-\newcommand{\E}{\mathbf{E}}
-\newcommand{\betavec}{\pmb{\beta}}
-\newcommand{\qvec}{\mathbf{q}}
-\newcommand{\lpf}{\mathbf{\psi}} %logpartition function
-\newcommand{\family}{\mathcal F} 
-\newcommand{\suff}{\pmb{\phi}} % suff stats
-\newcommand{\muvec}{\pmb{\mu}}
-
-% Document starts
-
-\begin{document}
-% Page heads
-\markboth{Abernethy et al.}{Maximum Entropy Prediction Markets}
-% Title portion
-\title{Maximum Entropy Prediction Markets}
-\author{JACOB ABERNETHY
-\affil{University of Michigan, Ann Arbor}
-SINDHU KUTTY
-\affil{University of Michigan, Ann Arbor}
-S\'{E}BASTIEN LAHAIE
-\affil{Microsoft Research, New York City}
-RAHUL SAMI
-\affil{University of Michigan, Ann Arbor}}
-%%\author{GANG ZHOU
-%%\affil{College of William and Mary}
-%%YAFENG WU
-%%\affil{University of Virginia}
-%%TING YAN
-%%\affil{Eaton Innovation Center}
-%%TIAN HE
-%%\affil{University of Minnesota}
-%%CHENGDU HUANG
-%%\affil{Google}
-%%JOHN A. STANKOVIC
-%%\affil{University of Virginia}
-%%TAREK F. ABDELZAHER
-%%\affil{University of Illinois at Urbana-Champaign}}
-% NOTE! Affiliations placed here should be for the institution where the
-%       BULK of the research was done. If the author has gone to a new
-%       institution, before publication, the (above) affiliation should NOT be changed.
-%       The authors 'current' address may be given in the "Author's addresses:" block (below).
-%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
-%       currently affiliated with NASA.
-
-\begin{abstract}
-In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets.
-We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data.
-Similarly,  the market state can be interpreted as a distribution over the outcome space.
-In particular, we show that a proper scoring rule can be derived from maximum entropy distributions.
-This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces.
-In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility.
-We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function.
-These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders.
-In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market.
-Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
-\end{abstract}
-
-\category{J.4}{Social and Behavioral Sciences}{Economics}
-\category{I.2.6} {Artificial Intelligence}{Learning}
-
-\terms{Prediction Market Design, Machine Learning Algorithms}
-
-\keywords{Exponential Families, Bayesian Learning, Maximum Entropy Distributions, generalized LMSR}
-
-\acmformat{Jacob Abernethy, Sindhu Kutty, S\'{e}bastien Lahaie, and Rahul Sami, 2014. Maximum Entropy Prediction Markets}
-%%Gang Zhou, Yafeng Wu, Ting Yan, Tian He, Chengdu Huang, John A. Stankovic, and Tarek F. Abdelzaher, 2010. A multifrequency MAC specially designed for  wireless sensor network applications.
-
-% At a minimum you need to supply the author names, year and a title.
-% IMPORTANT:
-% Full first names whenever they are known, surname last, followed by a period.
-% In the case of two authors, 'and' is placed between them.
-% In the case of three or more authors, the serial comma is used, that is, all author names
-% except the last one but including the penultimate author's name are followed by a comma,
-% and then 'and' is placed before the final author's name.
-% If only first and middle initials are known, then each initial
-% is followed by a period and they are separated by a space.
-% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.
-
-\begin{bottomstuff}
-%%This work is supported by the National Science Foundation, under
-%%grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.
-%%
-%%Author's addresses: G. Zhou, Computer Science Department,
-%%College of William and Mary; Y. Wu  {and} J. A. Stankovic,
-%%Computer Science Department, University of Virginia; T. Yan,
-%%Eaton Innovation Center; T. He, Computer Science Department,
-%%University of Minnesota; C. Huang, Google; T. F. Abdelzaher,
-%%(Current address) NASA Ames Research Center, Moffett Field, California 94035.
-\end{bottomstuff}
-
-\maketitle
-
-\section{Introduction}
-We have dual goals in this paper.
-On the one hand, we highlight a structural similarity between prediction markets and exponential families.
-We see this as the syntax of our prediction market mechanism.
-On the other, this formulation has rich semantics as well: it allows for analysis of market behavior under various environments.
-For instance, we analyze market behavior with budget limited traders, traders that are exponential-utility-maximizers.
-\section{Notation and Definitions}
-convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesi
-\section{Maximum Entropy Market Making}
-
-The purpose of a prediction market is to elicit and aggregate the beliefs (i.e., subjective probabilities) of agents over a space of \emph{outcomes}.
-In a single-agent setting, a scoring rule is used to elicit the agent's beliefs.
-In a multi-agent setting, an information market is used to aggregate the agent's beliefs.
-Hanson~\cite{} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both so that they can perform well in both thin and thick markets.
-In this work we derive a wide-ranging generalization of the logarithmic market scoring rule.
-Using a maximum entropy approach, we explain how a market scoring rule can be developed for outcome spaces both discrete and continuous, and for generic properties of the underlying distribution (e.g., mean and variance), rather than just the probabilities of individual outcomes.
-
 %%%
-%%%
+%%% Exponential Utility
 %%%
 
-\subsection{The Model}
-\label{model}
-
-Let $\mX$ be the outcome space, which may be discrete or continuous.
-Let $\mP$ denote the set of probability distributions over the outcome space.
-We represent a probability distribution as a density $p$ absolutely continuous with respect to some base measure $\nu$ (e.g., counting for discrete outcomes, or Lebesgue for continuous outcomes).% We will discuss the interpretation of the base measure in a later section.
-
-There is an unknown distribution $p$ over outcomes.
-We are interested in estimating the expected value of different outcome \emph{statistics} under this distribution by aggregating the beliefs of agents.
-A statistic is a real-valued function $\phi_s : \mX \rightarrow \bR$ over outcomes, independent of $p$; here $s$ belongs to a finite index set $\mS$ of size $d$.
-The collection of functions $\phi = (\phi_s,\, s \in \mS)$ can be viewed as a vector-valued statistic mapping outcomes into $\bR^d$.
-Formally, the aim is to estimate
-%
-\begin{equation} \label{constraint}
-\Exp_p[\phi(x)] = \mu.
-\end{equation}
-%
-We stress that we are only seek to elicit $\mu \in \bR^d$, not the full distribution $p$.
-As an example, suppose that $\mX = \bR$.
-If we were interested in eliciting the first two uncentered moments of $p$, we would include $\phi_1(x) = x$ and $\phi_2(x) = x^2$ as statistics.
-Note that the variance cannot be directly obtained through any single statistic: it corresponds to the expectation of $(x - \Exp_p[x])$, which depends on $p$, violating our definition of a statistic.
-Of course, the variance can be indirectly obtained by eliciting the first two moments separately.\footnote{This is related to the notion of \emph{elicitation complexity} of statistical properties introduced by Lambert et al.~\cite{Lambert}, but we do not pursue this connection here.}
-
-There are two approaches to eliciting an agent's estimated $\mu$ in~(\ref{constraint}).
-The first approach is to incentivize the agent to directly report $\mu$ via a \emph{scoring rule}.
-The second approach is to form an \emph{information market} by issuing securities for each statistic $s \in \mS$, with payoff depending on the realized value of the statistic.
-(Such securities are known as contingent claims.) The amount of shares of each security acquired by the agent indirectly reveals it estimate $\mu$.
-In a market scoring rule~\cite{Hanson}, these two approaches are dual to each other in a formal sense.
-
-%%%
-%%%
-%%%
-
-\subsection{Scoring Rule}
-
-We first consider how to directly elicit $\mu$ via a scoring rule. 
-%Note that the usual definition of a scoring rule presumes that a full probability distribution is elicited, rather than just some of its properties (e.g., mean and variance), so below we will adapt the definition to our setting.
-The space of feasible reports from an agent corresponds to
-%
-\begin{equation} \label{marginal}
-\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}.
-\end{equation}
-%
-A scoring rule is a function $S : \mM \times \mX \rightarrow \bR$ that rewards an agent with $S(\mu,x)$ based on how its report $\mu$ agreed with the eventual outcome $x$.
-A scoring rule is \emph{proper} if for all $\mu \in \mM$, and $p \in \mP$ such that $\Exp_p[\phi(x)] = \mu$, we have
-%
-\begin{equation} \label{scoring}
-\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)]
-\end{equation}
-%
-for all $\hmu \neq \mu$; the rule is \emph{strictly proper} if the inequality is strict.
-Note that in the literature, scoring rules are defined over entire probability distributions, not just properties of those distributions.
-Our definition imposes a minimum of informational requirements on agents, because an agent does not need to assess the full distribution $p$ that might have lead to its estimates $\mu$ in order to realize that truthful reporting is an optimal strategy.
-
-Our scoring rule is implicitly defined via the solution of a mathematical program. The program computes the maximum entropy distribution consistent with the agent's reported beliefs $\mu$.
-%
-\begin{eqnarray}
-\max_{p \geq 0} & \ds - \int_{x \in \mX} p(x) \log p(x) \nu(dx) & \label{obj} \\
-\mbox{s.t.} & \ds \int_{x \in \mX} \phi(x) p(x) \nu(dx) = \mu & \label{mean} \\
- & \ds \int_{x \in \mX} p(x) \nu(dx) = 1 & \label{normalize}
-\end{eqnarray}
-%
-The non-negativity constraints together with~(\ref{normalize}) ensure that the solution is a probability distribution.
-Constraints~(\ref{mean}), which correspond to $d$ separate constraints, ensure consistency with the agent's report.
-The objective~(\ref{obj}) is the entropy of the solution $p$ with respect to the measure $\nu$.
-The program has a convex objective and linear constraints.
-%For continuous outcome spaces we will see in the next section that the optimal solution can still be cleanly characterized.
-%
-\begin{theorem} \label{thm:scoring}
-Let $p(x;\mu)$ be the optimal solution to the maximum entropy program, given report $\mu \in \bR^d$. The scoring rule defined by
-%
-\begin{equation} \label{scoring}
-S(\mu,x) = a_x + b\log p(x;\mu),
-\end{equation}
-%
-where $b,a_x \in \bR$ and $b > 0$, is strictly proper.
-\end{theorem}
-%
-Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$
-
-Note that the maximum entropy distribution is an exponential family distribution. 
- Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.
-
- We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
-  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
-   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
-   Or 
-\begin{equation}\label{eq:proper}
-\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
-\end{equation}
-
-   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
-   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
-   Rearranging the terms,
-   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
-   which is the first order condition on the convexity of $\psi(\cdot)$.
- \unproof
-
- \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters.
-That is,
-   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
-   \unrmk
-
-\noindent
-This rule represents a generalization of the logarithmic scoring rule for probability distributions to any properties of distributions that can be captured as expectations of statistics.
-As explained in the next section, the solution to the maximum entropy program takes the form of an \emph{exponential family} distribution.
-For many common properties of interest, these distributions are familiar; for example, the maximum entropy distribution with a given mean is an exponential distribution, and the maximum entropy distribution with a given mean and variance is a normal distribution~\cite{}.
-We provide here two examples in more depth.
-
-%%
-%%
-
-\paragraph{Multinomial Distribution}
-
-As a first example, consider the problem of estimating the individual probabilities of a finite set of outcomes.
-We have $\mX = \{1,\ldots,k\}$.
-The relevant statistics indicate which outcome actually occurs, so the index set is $\mS = \mX$.
-Statistic $\phi_x(x')$ is 1 if $x' = x$ and 0 otherwise.
-Observe that if $p$ is the distribution over outcomes, then $\Exp_p[\phi(x)] = p$.
-A feasible report from an agent is any non-negative $\pi = (\pi_1,\ldots,\pi_k)$ such that $\sum_{i=1}^k \pi_i = 1$.
-Given a report of $\mu = \pi$, the unique solution to the maximum entropy program is $p(x; \pi) = \pi$, and~(\ref{scoring}) corresponds to the classic logarithmic scoring rule of $S(\pi,x) = \log \pi_x$.
-\hfill{$\Box$}
-
-\paragraph{Normal Distribution}
-
-As a second example, suppose the outcome space is $\mX = \bR$ and that we are interested in estimating the mean $\mu$ and variance $\sigma^2$ of the underlying distribution.
-We choose the first and second (uncentered) moments, $\kappa_1 = \mu$ and  $\kappa_2 = \mu^2 + \sigma^2$ as our statistics.
-It is well-known that the maximum entropy distribution with a given mean and variance is the normal distribution, so the solution to the maximum entropy program (using Lebesgue as the base measure $\nu$) is
-%
-\[ p(x; \kappa) = \frac{1}{\sqrt{2\pi\sigma^2}} \exp\left[ -\frac{(x-\mu)^2}{2\sigma^2} \right]. \]
-%
-Applying Theorem~\ref{thm:scoring} and choosing appropriate constants, we find that the following scoring rule is strictly proper in this context. %
-\[ S(\kappa, x) = -\frac{(x-\mu)^2}{\sigma^2} - \log \sigma^2. \]
-%
-Note that the fact that this scoring rule is proper for the mean amounts to the well-known fact that reporting the mean is a Bayes act under quadratic loss~\cite{someone}. \hfill{$\Box$}
-
-\medskip\noindent
-A brief observation on parametrizations is in order.
-Note that the maximum entropy distribution here is parametrized by the uncentered moments $\kappa = (\kappa_1,\kappa_2)$ rather than $(\mu,\sigma^2)$ because the variance is not a valid statistic.
-However, it is clear that the agent does not have to directly report $\kappa$: the market maker itself can do the translation from reported mean and variance to uncentered moments.
-This is relevant because the $\kappa$ parametrization may be unintuitive for an agent, especially given the constraints $\kappa_2 - \kappa_1^2 \geq 0$ that must be enforced.
-In the $(\mu, \sigma^2)$ parametrization each parameter is unrestricted, and the parameters have intuitive interpretations.
-
-Our approach so far has been to identify relevant properties of the unknown distribution, and elicit them via a scoring rule computed through a maximum entropy program.
-However, it is the case that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}.
-Therefore, we see the following alternative approach.
-The market designer can begin by considering the outcome space $\mX$, and assessing what family of distributions over $\mX$ the unknown $p$ might come from.
-If this family is an exponential family, then Theorem~\ref{thm:scoring} immediately gives a proper scoring rule to elicit the parameters of the distribution.
-Many of the most familiar distributions are exponential families, including the normal, exponential, gamma, beta, binomial, Poisson, Weibull, and Dirichlet.
-To illustrate, we consider two examples in more depth.
-
-%%
-%%
-
-\paragraph{Exponential}
-Suppose the designer would like an estimate of the rate at which servers fail in a data center.
-An outcome is the time between failures, so $\mX = [0,+\infty)$.
-The usual distribution for failure times is the exponential distribution $p(x;\lambda) = \lambda e^{-\lambda x}$, which is parametrized by the rate $\lambda$ and has mean $\mu = 1/\lambda$.
-The exponential distribution is the maximum entropy distribution given a fixed mean $\mu$.
-From Theorem~\ref{thm:scoring} we obtain the following proper scoring rule for this setting:
-%
-\[ S(\mu, x) = \log \lambda - \lambda x. \]
-%
-It is simple to check using basic calculus that an agent maximizes the expected score by reporting $\lambda = 1/\hmu$, where here $\hmu$ is its estimate of the mean.
-In this case, it is unclear a priori which of the mean or rate parametrizations is most intuitive, but the maximum entropy approach easily allows for either.
-\hfill{$\Box$}
-
-%%
-%%
-
-\paragraph{Beta}
-Suppose the designer would like an estimate of the click-through rate of an online advertisement.
-The outcome space here is $\mX = (0,1)$.
-A typical prior over a probability like this is the beta distribution, which has two parameters $\alpha, \beta > 0$ and density function $p(x; \alpha, \beta) = x^{\alpha-1}(1-x)^{\beta-1}/B(\alpha,\beta)$, where $B$ is the beta function.
-By theorem~\ref{thm:scoring} this leads to the proper scoring rule
-%
-\[ S((\alpha, \beta), x) = (\alpha-1)\log x + (\beta-1)\log (1-x) - \log B(\alpha,\beta). \]
-%
-Although this parametrization is standard for the beta, it is perhaps more intuitive to use the parametrization $\mu = \alpha/(\alpha+\beta) \in (0,1)$ and $n = \alpha+\beta > 0$.
-Here $\mu$ is the mean of the distribution---corresponding to the agent's actual estimate of the click-through rate---and $n$ is a measure of the agent's confidence in the estimate.
-\hfill{$\Box$}
-
-%%%
-%%% 
-%%%
-
-\subsection{Information Market}
-
-We next consider how to indirectly elicit $\mu$ by setting up a market of contingent claim securities.
-Under this approach, the elements of the index set $\mS$ are interpreted not as statistics but as securities.
-The payoff from one share of security $s$ when outcome $x$ occurs is given by the mapping $\phi_s$.
-Thus if the vector of shares held by the agent is $\theta \in \bR^d$, where entry $\theta_s$ corresponds the number of shares of security $s$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \theta, \phi(x) \ra$.
-As a concrete example, recall our multinomial distribution example from the previous section, where $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise.
-This means that security $x \in \mS$ pays 1 dollar if outcome $x \in \mX$ occurs, and nothing otherwise.
-(Such securities are known as Arrow-Debreu securities.) In our normal distribution example, the statistic for the mean had the mapping $\phi_1(x) = x$.
-Therefore a share of the corresponding security has a payoff that is linear in the outcome.
-(Such securities amount to futures contracts.)
-
-To extract useful information from such a securities market, we set a pricing scheme and examine the number of shares the agent chooses to acquire.
-Let $\Omega$ be the set of all portfolios (i.e., vectors of shares) that the agent can feasibly hold; we will see how this domain is determined in an instant.
-Each security $s$ has an associated price function $c_s : \Omega \rightarrow \bR$, which gives the marginal price $c_s(\theta)$ of security $s$ when the agent holds portfolio $\theta$.
-(Note that the marginal price depends on the entire portfolio $\theta$, not just the number of shares $\theta_s$.) A risk-neutral agent will choose to acquire shares up to the point where, for each share, expected payoff equals marginal price.
-Formally, if the agent acquires portfolio $\theta$, then for each $s \in \mS$ we must have
-%
-\begin{equation} \label{market}
-\Exp_p[\phi_s(x)] = c_s(\theta).
-\end{equation}
-%
-In this way, by its choice of $\theta$, the agent reveals that its belief is $\mu = c(\theta)$.
-There are several important properties that the price function $c$ should have.
-To simplify the agent's portfolio acquisition process, it should not be the case that the total cost of acquiring $\theta$ depends on the order in which shares are bought.
-This means that there should exist a \emph{cost function} $C : \Omega \rightarrow \bR$ such that $c = \nabla C$.
-The gradient $\nabla C$ must be onto $\mM$ to ensure that~(\ref{market}) always has a solution---this is the most important but technical condition to realize.
-To ensure that the solution is unique, it would also be convenient if the cost function were strictly convex.
-%
-\begin{proposition}
-The following cost function is monotone, convex, and onto $\mM$: 
-%
-\begin{equation} \label{cost}
-C(\theta) = \log \int_{x \in \mX} \exp \left[\la \theta, \phi(x) \ra\right] \nu(dx).
-\end{equation}
-%
-\end{proposition}
-%
-\noindent
-Observe that in the context of our earlier multinomial example, cost function~(\ref{cost}) is exactly the cost function for Hanson's logarithmic market scoring rule.
-Our approach here provides a generalization of this market scoring rule to markets with contingent claim securities with arbitrary payoffs, designed to elicit specific properties of distributions, beyond just the probabilities of different outcomes.
-
-Because an agent would never select a portfolio with infinite cost, the effective domain of $C$ is $\Omega = \{\theta \:|\: C(\theta) < +\infty \}$.
-In the context of the multinomial distribution example, $\Omega = \bR^d$, so that shares of each security can always be bought or sold short.
-In the context of the normal distribution example the effective domain is $\Omega = \{(\theta_1,\theta_2) \in \bR^2 \:|\: \theta_2 > 0 \}$ if we re-define the second statistic to be $\phi_2(x) = -x^2$; this means that the security has a negative payoff for each share, and consequently the agent must be compensated to acquire such shares.
-Evaluating~(\ref{cost}), the cost function takes the form $C(\theta) = \frac{\theta_1^2}{4\theta_2} - \frac{1}{2}\log(2\theta_2)$.
-
-Now, in our previous elicitation approach that used a scoring rule, the process of computing the scoring rule also provided a complete distribution $p(x;\mu)$ over outcomes, which could be used to infer other properties beyond just the agent beliefs $\mu$.
-In the current market-based approach, the agent's chosen portfolio $\theta$ can also form the basis of a distribution over outcomes.
-Consider the following distribution, represented as a density with respect to a base measure $\nu$:
-%
-\begin{equation} \label{expfam}
-p(x;\theta) = \exp\left[ \la \theta, \phi(x) \ra - C(\theta) \right].
-\end{equation}
-%
-Meaning, the probability that a subset $X \subseteq \mX$ of the outcomes occurs is $\int_{x \in X} p(x;\theta)\nu(dx)$.
-Observe that by definition~(\ref{cost}), this probability density indeed integrates to 1 over $\mX$, and it is clear that the density is non-negative as required.
-
-In the statistical literature a distribution that takes the form~(\ref{expfam}) is known as an \emph{exponential family}.
-The mapping $\phi$ is known as the \emph{sufficient statistic}, $\theta$ is the \emph{natural parameter}, and $C$ is the log-partition or \emph{cumulant} function.
-%This connection between information markets and exponential families allows us to draw on a vast statistical literature to understand the market's properties, and a growing computer science literature to apply efficient algorithms to compute costs and perform inference.
-
-%%%
-%%%
-%%%
-
-\subsection{Duality}
-
-There is a well-known duality between the maximum entropy approach and exponential families, which translates into a duality between the scoring rule and information market just developed.
-The duality implies that the approach leads to a \emph{market scoring rule}, applicable to both thin and thick multi-agent settings.
-
-It is known that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}.
-To see this, let $\theta(\mu) \in \bR^d$ be the Lagrange multiplier corresponding to constraints~(\ref{mean}) when solving the maximum entropy program given the agent report $\mu$.
-(Our choice of notation is deliberately suggestive.) Let $A(\mu)$ be the Lagrange multiplier corresponding to~(\ref{normalize}).
-From the first order necessary conditions for optimality, we find that
-%
-\[ p(x;\mu) = \exp \left[ \la \theta(\mu), \phi(x) \ra - A(\mu) - 1 \right], \]
-%
-so that the solution is indeed in exponential family form.
-Conversely, given a cumulant $C$ from an exponential family and a parameter $\theta$, we see that $p(x;\theta)$ defined according to~(\ref{expfam}) is the maximum entropy distribution under constraints $\Exp_p[\phi(x)] = \mu$ where $\mu = \nabla C(\theta)$.
-We obtain the following.
-%
-\begin{proposition}
-Let $C$ be the cumulant (i.e, cost function) for the exponential family corresponding to $\phi$, and let $\theta(\mu)$ be the optimal Lagrange multiplier for the mean constraints given an agent report of $\mu \in \mM$.
-Then the following scoring rule is equivalent to~(\ref{scoring}):
-%
-\begin{equation} \label{expscore}
-S(\mu, x) = \la \theta(\mu), \phi(x) \ra - C(\theta(\mu)).
-\end{equation}
-%
-\end{proposition}
-%
-\noindent
-The scoring rule~(\ref{expscore}) decomposes neatly into payoff and cost functions.
-The first term $\la \theta(\mu), \phi(x) \ra$ defines the agent's outcome-contingent reward, while the second term~$C(\theta(\mu))$ is the cost of acquiring a portfolio $\theta(\mu)$.
-We see here the duality between the approach of directly reporting $\mu$, or acquiring shares $\theta$: an agent reporting beliefs $\mu$ under scoring rule~(\ref{scoring}) would choose to acquire shares $\theta(\mu)$ in the information market with cost function~(\ref{cost}).
-Since~(\ref{scoring}) is a proper scoring rule, the information market with cost function~(\ref{cost}) is based on a proper market scoring rule.
-
-\section{The Exponential Family Market Mechanism: A Maximum Likelihood Approach}
-In this section, we will set up a prediction market that aggregates beliefs from traders where the outcome is a continuous random variable.
-In particular, we assume that the outcome is drawn from an exponential family distribution.
-Each of the traders has access to a series of points drawn from this distribution.
-In other words, the traders have access an empirical mean of the sufficient statistics of the exponential family.
-Every trader has infinite budget so that the current market price after a trader has traded in the prediction market, exactly reflects his beliefs.
-
-%We will now provide a more formal description of the problem statement.
-%\subsection{Problem Statement}
-%\begin{description}
-%\item[Setup: ] Learning proceeds in rounds $j=1,2,\ldots$ where at the end of each round the algorithm may be provided with $x_{j}$  drawn from a set $\mathcal{X}$ according to some distribution $\mathcal{D}_{j}$, where $\mathcal{D}_{j}$ belongs to an exponential family defined by $k$ sufficient statistics.
-In each round, some sequence of $n$ traders provide their estimates of the empirical means of each of the $k$ sufficient statistics drawn from $\mathcal{D}_{j}$.
-%\item[Define: ] A learning algorithm that learns this distribution by providing a maximum likelihood estimate of its unknown natural parameters based on the empirical means.  
-%\item[Proposed solution: ] Simulate a prediction market defined by a cost function $C$, securities $s_{0},s_{1},s_{2},\ldots, s_k$ and their payoffs $\phi_i(x)$ for $i=0,1,2,\ldots,k$ for the LMSR.
-Define a correspondence between the state of the prediction market and the output of the learning algorithm.
-%\end{description}
-
-For $x_{i}\sim P_{\betavec}$, recall that the likelihood function for independently drawn data $x_1,\ldots,x_n$ is given by $\prod_{i=1}^{n} P_{\betavec}(x_i)$.
-The maximum likelihood estimate of the natural parameters $\betavec$ is the value of the natural parameters that maximizes the likelihood function.
-
-We will now set up a prediction market with log market scoring rule (LMSR) and infinite budget traders such that the market state represents the maximum likelihood estimate (MLE) of the natural parameters of an exponential family distribution.
-%Given  under some interpretation of what information the participating agents have, and how they bet.
-
-For a given exponential family distribution, the prediction market is defined as follows:
-\begin{description}
-\item[Traders] The prediction market will simulate a trader $i$ corresponding to expert $i$.
-This trader processes all information samples available to her directly or inferred from previous trades, and  forms a belief distribution such that the believed means of the sufficient statistics matches the empirical means of the sufficient statistics from the information samples.
-The trader trades in the market to maximize her expected payoff under her believed distribution.
-\item[Securities and their payoff]
-For each $i=1,2,\ldots,k$, we define a security $s_{i}$ with payoff $\phi_{i}(x)$ where $x$ is the ultimate outcome and $\Phi()$ defines the vector of sufficient statistics of the exponential family distribution according to which $x$ is drawn.
-We define an additional security $s_{0}$ with payoff $\phi_{0}(x):=a-\sum_{i=1}^{k}\phi_{i}(x)$ where $a$ is an appropriately chosen constant dependent on the range of $\Phi$ so that the payoff of $s_{0}$ is non-negative.
-\end{description}
-
-We also note that since we want non-negative payoffs, we restrict the exponential families under consideration so that the sufficient statistics are lower bounded.
-If we have a constant lower bound on the sufficient statistics, then without loss of generality we can add a constant to each sufficient statistic without changing the exponential family in any way, as the constants will be absorbed in the log-partition function.
-
-\paragraph{Arbitrage-free Property}
-We can easily show that the prediction market we have defined is arbitrage-free: there is no sequence of trades that guarantees the trader a profit under all conditions.
-First, note that the incremental cost and payoff function are additive over multiple trades, so the net profit of a sequence of trades depends only on the initial and final market position, and is independent of the actual path along which the trade takes place.
-
-Now, consider any trade (or sequence of trades) that moves the market from an initial state of $\qvec$ to a final state of $\qvec + \qvec'$.
-The cost of this trade is $C(\qvec + \qvec') - C(\qvec)$.
-The cost function of our market is the log-partition function of an exponential family, and thus, $C(.)$ is a strictly convex function~\cite{WainJordan08}.
-Thus, we have the inequality:
-\[
-   C(\qvec + \qvec') - C(\qvec)  > \qvec' \nabla C|_\qvec = \qvec' \mu,
- \]
- where $\mu$ is the mean sufficient statistics vector for the distribution with parameters $\qvec$.
-The payoff due to this trade, with outcome $x$, is $\qvec'.\phi(x)$. 
-Now, we observe that under the distribution with parameters $\qvec$, the expected payoff is $\qvec' E[\Phi(x)] = \qvec' \mu$.
-Thus, under this distribution over outcomes $x$, the expected profit is strictly less than $0$.
-This is only possible if the
-realized profit is less than $0$ for at least one outcome $x$. As this statement is true for every $\qvec$ and $\qvec'$, and the profits are path-independent, the market is arbitrage-free.
-
-%Let $\qvec^*=(q^{*}_{0},q^{*}_{1},\ldots,q^{*}_{k})$ be the number of shares of each security held by the traders.
-First we note that under the assumption of perfectly rational, risk-neutral traders with infinite budget and  beliefs as indicated above, the gradient of the cost function at this point is
- %$$\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}=\mu_{i}$$
-%where $\mu_{i}$ is believed to be the expected payoff of the $i^{th}$ security by the last trader who traded in this market.
-We show below that the instantaneous price of security $i$, $\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$ at any $\qvec=\qvec^{*}$, can be written as $\left(\frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$.
-This means that the instantaneous price of the $i^{th}$ security only depends on the difference between the number of held securities of $i$ and 0.
-This ensures that if there exists an instantaneous price that reflects the belief of a trader at any vector $\qvec^{*}$, there must exist a $\qvec'$  that also reflects his beliefs and with the desirable property that $\forall i=0,1,2,\ldots,k\ q_{i}'\geq 0$.
-This additional property ensures that no trader will be required to short sell securities to move the market state to  reflect his beliefs.
-
-\paragraph{Market State and Natural Parameters}
-We define the interpretation function 
-\begin{eqnarray}\label{eqn:int}
-I(\qvec)&=&(q_{1}-q_{0},q_{2}-q_{0},\ldots,q_{k}-q_{0})\nonumber\\
-&=&(\beta_{1}, \beta_{2},\ldots, \beta_{k})=\betavec
-\end{eqnarray}
-This allows us to interpret the state of the market in terms of a prediction on the natural parameters of the distribution.
-
-\begin{theorem}
-The natural parameter vector corresponding to the interpretation of the market state given by Equation \ref{eqn:int} is the vector of their maximum likelihood estimates.
-Further, this interpretation is unique.
-\end{theorem}
-\begin{proof}
-We observe that for $i=1,\ldots,n$,
-\begin{eqnarray*}
-\mu_{i}&=&\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int \exp(\qvec^{T}\Phi(x))dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}\phi_{0}(x)+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}(a-\sum_{i=1}^{n}\phi_{i}(x))+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}a} e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log (e^{q_{0}a} \int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial (q_{0}a)}{\partial q_{i}} \frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\frac{\partial \beta_{i}}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\right)_{\rm \betavec=\betavec^{*}}
-\end{eqnarray*}
-Thus, this choice of parameters satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$. We will now argue that this vector is unique.
-
-If the exponential family is represented so that there is a unique parameter vector associated with each distribution, the representation is said to be minimal.
-The Bernoulli, Gaussian, and Poisson distributions all have minimal representations.
-Now, for an exponential family whose representation is minimal, the gradient is an injection \cite[Prop.
-3.2]{WainJordan08}.
-Thus, there is a unique parameter vector $\betavec$ that satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$, for each $i\in\{1,\ldots,n\}$.
-Thus, if the $\mu_{i}$'s correspond to the empirical means, since our choice of $\betavec$ satisfies the equality, it must also be the vector corresponding to the MLE.
-\end{proof}
-
-Thus, we have shown that this prediction market (along with its strategic traders) aggregates trader information in a way that the market state can be interpreted as a predictive distribution (the maximum likelihood distribution) over an infinite outcome space.
-
-It is worth noting here that if we can bound the parameter space and the sufficient statistics of the distribution, the market maker defined here has bounded worst case loss: The probability density, and hence log loss, would then be bounded for all market states that could be reached and all outcomes.
-In fact, for many practical applications, a distribution with unbounded parameter space is an approximation for a true space which itself may be bounded.
-In this case, worst case loss would thus be bounded.
-
-\subsection{The Exponential Family Market Mechanism in Adversarial Markets}
-In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
-
-In this section, we consider the the prediction market setup with traders that may be either informative or malicious.
-The malicious traders may want to inject faulty information into the market.
-The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.
-
-We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market.
-Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
-
-\subsubsection{Budget-limited Aggregation}
-Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
-We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market.
-This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader.
-Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.
-
-%There is distribution under which expected profit is $\leq 0$. Cost function is convex.
-
-%For any $\qvec$ there is a
-In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
-Let $\alpha$ be the budget of a trader in the market.
-Suppose with infinite budget, the trader would have moved the market state from $\qvec_{init}$ to $\qvec$, where $\qvec$ represents his true beliefs.
-Suppose further that $\alpha<C(\qvec)-C(\qvec_{init})$.
-In this case, we want to budget-limit the trader's influence on the market state.
-
-We define the budget-limited final market state as $\tilde{\qvec}$. 
-Here, we consider a specific functional form of $\tilde{\qvec}$: $$\tilde{\qvec}=\lambda\qvec+(1-\lambda)\qvec_{init}$$  where $$\lambda=\min\Big(1, \frac{\alpha}{C(\qvec) - C(\qvec_{init})}\Big)$$ We first show that this trade is feasible given the trader's budget:
-\begin{theorem}
-Let the current market state be given by the vector $\qvec_{init}$ and $\tilde{\qvec}=\lambda\qvec+(1-\lambda)\qvec_{init}$.
-For $\lambda=\min\Big(1, \frac{\alpha}{C(\qvec) - C(\qvec_{init})}\Big)$, the cost to the trader to move the market state from $\qvec_{init}$ to $\tilde{\qvec}$ is at most his budget $\alpha$.
-\end{theorem}
-\begin{proof}
-From the convexity of $C$, we have
- $$C(\tilde{\qvec})\leq(1-\lambda)C(\qvec_{init})+\lambda C(\qvec)$$
-
-Now 
-\begin{eqnarray*}
-C(\tilde{\qvec}) - C(\qvec_{init}) &\leq& (1-\lambda)C(\qvec_{init})\\&&+\lambda C(\qvec) - C(\qvec_{init})\\
-&=& \lambda\ (C(\qvec) - C(\qvec_{init}))
-\end{eqnarray*}
-
-%So when $\alpha<C(\qvec)-C(\qvec_{init})$, we pick $\lambda$ so that 
-%$$\lambda\leq \frac{\alpha}{(C(\qvec) - C(\qvec_{init}))}$$
-%
-%where we want $\lambda$ such that
- Thus, $C(\tilde{\qvec}) - C(\qvec_{init}) \leq \alpha$. 
- \end{proof}
-
- We note that moving to $\tilde{\qvec}$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
-In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\tilde{\qvec}$.
-Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\qvec$.
-\subsubsection{Damage Bound}
-Given this restriction on the initial budget of traders, we can now show that the total loss that can be induced by malicious traders is bounded.
-We define the loss function for $\qvec$ shares held as:
-$$L(\qvec,x)=-\log(p_{\betavec}(x))=\log\frac{\int e^{\qvec^{T}\Phi(x)}\ dx}{e^{\qvec^{T}\Phi(x)}}$$
-with the correspondence between $\betavec$ and $\qvec$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
-
-%{\bf Setup:} The market evolves in rounds $t=1,\ldots,T$.
-The end of a time segment occurs either when each expert has made his report or when the actual value of the random variable is received as input.
-Each expert $i$ reports the means of the sufficient statistics as available to him at time $t$.
-These are combined with the previously reported means to yield $\mu_{0}^{t}(x),\mu_{1}^{t}(x),\ldots,\mu_{k}^{t}(x)$ and an associated weight $w_{i}^{t}$ which can be interpreted as the number of points from which these values have been determined.
-These sufficient statistics correspond to a particular value of $B$ and hence of $\qvec$, the number of shares of each security that are held by traders.
-Since trader $i$ has finite budget $\alpha_{i}^{t}$ at time $t$, the current market state is moved to $\qvec'$ instead, so that cost of moving it to $\qvec'$ is within his budget.
-
-%For an initial positive budget of $\alpha_{i}^{0}$, we perform updates to the budgets of each agent as
-%$$\alpha_{i}^{t}:=\alpha_{i}^{t-1}+\lambda_{i}^{t}(L(\tilde{q}_{i-1}^{t},x^{t})-L({q}_{i}^{t},x^{t}))$$ where we define
-%$$\lambda_{i}^{t}:=\min\Big(1, \frac{\alpha_{i}^{t-1}}{C(\qvec_{i}) - C(\tilde{\qvec}_{i-1})}\Big)$$
-
-Suppose the prediction market runs over multiple rounds $t$.
-Let $\qvec_{0}^{t}$ be the initial number of shares of each security that are held.
-Let $\tilde{\qvec}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
-Let us assume that at this point we receive the value of the random variable $x^{t}$.
-
-%Let the loss function be $L(\cdot,\cdot)$, where the two arguments are the vector representing the number of shares held of each security and the actual value of the random variable.
-
-%We want to show:
-%$$L({\qvec}_{0}^{t},x^{t})-L(\tilde{\qvec}_{k}^{t},x^{t})+\sum_{i}\alpha_{i}^{t-1}\leq \sum_{i}\alpha_{i}^{t}$$
-%That is, the loss due to incorporating the budget-limited information of the traders, is not more than the increase in their budgets.
-Over multiple instances of the prediction market, we can track the change in budget of each trader. The change in budget for trader $i$ is
-\begin{eqnarray*}
-\alpha_{i}^{t}-\alpha_{i}^{t-1}&=&C(\tilde{\qvec}_{i-1}^{t})-C(\tilde{\qvec}_{i}^{t})-(\tilde{\qvec}_{i-1}^{t}-\tilde{\qvec}_{i}^{t})^{T}\Phi(x^{t})\\
-&=&L(\tilde{\qvec}_{i-1}^{t},x^{t})-L(\tilde{\qvec}_{i}^{t},x^{t})
-\end{eqnarray*}
-
-Define the myopic impact of a trader $i$ in segment $t$ as
-$$\Delta_{i}^{t}:=L(\tilde{\qvec}_{i-1}^{t},x^{t})-L(\tilde{\qvec}_{i}^{t},x^{t})$$
-Thus, the myopic impact captures incremental gain due to the trader in a round. Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
-
-The total myopic impact due to all $k$ active traders is given by
-$$\Delta^{t}=L(\qvec_{0}^{t},x^{t})-L(\tilde{\qvec}_{k}^{t},x^{t})$$
-Thus $-\Delta^{t}$ captures the incremental loss due to the predictive probability after aggregation of all $k$ traders.
-
-%The myopic impact of a trader $i$ in segment $t$, is given by
-%$$\Delta_{i}^{t}:=L(\tilde{B}_{i-1}^{t},x^{t})-L(\tilde{B}_{i}^{t},x^{t})$$
-%where $\tilde{B}_{i}$ is the vector of sufficient statistics corresponding to the the budget-limited state of the market as induced by trader $i$.
-%The total myopic impact due to all $k$ active traders is thus given by
-%$$\Delta^{t}=L(B_{0}^{t},x^{t})-L(\tilde{B}_{k}^{t},x^{t})$$
-
-%If the loss function is convex in the first argument, then we have:
-%\begin{eqnarray*}
-%\Delta_{i}^{t}&=&L(\tilde{B}_{i-1}^{t},x^{t})-L(\tilde{B}_{i}^{t},x^{t})\\
-%&=&L(\tilde{q}_{i-1}^{t},x^{t})-L(\tilde{q}_{i}^{t},x^{t})\\
-%&=&L(\tilde{q}_{i-1}^{t},x^{t})-L(\lambda_{i}{q}_{i}^{t}+(1-\lambda_{i})\tilde{q}_{i-1}^{t},x^{t})\\
-%&\geq&L(\tilde{q}_{i-1}^{t},x^{t})-\lambda_{i}L({q}_{i}^{t},x^{t})-(1-\lambda_{i})L(\tilde{q}_{i-1}^{t},x^{t})\\
-%&=&\lambda_{i}(L(\tilde{q}_{i-1}^{t},x^{t})-L({q}_{i}^{t},x^{t}))\\
-%&=& \alpha_{i}^{t} - \alpha_{i}^{t-1}
-%\end{eqnarray*}
-%Thus, the myopic damage caused by trader $i$ at round $t$ is upper-bounded by the change in his budget in that round.
-
-%If we pick a loss function that is also bounded in $[0,1]$, then we see that the budget never falls below zero. 
-\begin{theorem}
-A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
-\end{theorem}
-\begin{proof}
-Consider the myopic impact of a single trader $i$ after participating in the market $T$ times.
-Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
-$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
-
-Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
-\end{proof}
-
-This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders.
-
-In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation.
-We provide an information-theoretic justification for this claim in Section \ref{sec:infth}.
-The intuition behind this claim is that his prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
-
-\subsubsection{An Information-Theoretic Interpretation}\label{sec:infth}
- We observe a useful alternative view of the market 
-scoring rule prediction market for exponential family learning. We connect the cost, payoff and profit function to information-theoretic 
-quantities associated with the exponential family.
-
-The following result has been previously pointed out by Amari \cite{Amari-KL}. 
-\begin{theorem}\label{lem:profit_decomposition}{\bf (Profit Decomposition)}: Consider an exponential 
-family $\family$ of distributions over some set of statistics $\suff(x)$, with natural parameters $\betavec$. Let $\pi, \rho \in \family$ be any two probability distributions in the family. We use 
-$\betavec_\pi$ to denote the natural parameters of $\pi$ and $\muvec_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
-we can define $\betavec_\rho$ and  $\muvec_\rho$. We abuse notation slightly and let $\lpf(\rho)$ indicate the log partition function of $\rho$ which technically depends on its natural parameters.
-Let $H(\pi)$ denote the entropy of the distribution $\pi$, and 
- $K(\pi||\rho)$ denote the KL-divergence of $\rho$ relative to $\pi$.
-Then, the following equality holds:  
-\begin{equation}
- K(\pi||\rho) + H(\pi) = \lpf(\rho) - \betavec_\rho \cdot \muvec_\pi
-\label{eq:profit_decomposition}
-\end{equation}
-\end{theorem}
-\begin{proof}
-\begin{eqnarray*}
- K(\pi||\rho) + H(\pi) &=& \int_x \pi(x)\log{\frac{\pi(x)}{\rho(x)}}dx\\
- &&\quad - \int_x \pi(x)\log{\pi(x)}dx \\
-&=& - \int_x \pi(x)\log{\rho(x)}dx \\
-&=& - \int_x \pi(x)\left[ \betavec_\rho \cdot \suff(x) - \lpf(\rho)\right]dx\\
-&=& - \betavec_\rho \cdot \int_x \pi(x) \suff(x) dx\\
-&&\quad + \lpf(\rho) \int_x \pi(x) dx \\
-&=&  \lpf(\rho) - \betavec_\rho \cdot \muvec_\pi
-\end{eqnarray*}
-\end{proof}
-
-Equation~\ref{eq:profit_decomposition} gives us an alternative view of the
-market scoring rule construction. Assume that $\pi$ is the true distribution, and consider two arbitrary
-distributions $\rho_1, \rho_2 \in \family$.
-Note that $\lpf(\rho)$  is independent of $\pi$, and $\betavec_\rho \cdot \muvec_\pi$ is linear in the probabilities $\pi(x)$.
-If we want to measure loss by the KL-divergence, we can do so (in expectation) by setting a cost function that captures the first term, and defining security quantities ($\betavec$) and payoffs ($\suff$) to capture the second term.
-In particular, in a market with cost function $\lpf$, if the
-market price initially implies a distribution $\rho_1$, and a trader moves
-the market to price than implies a distribution $\rho_2$, then the cost she
-incurs is $\lpf(\rho_2) - \lpf(\rho_1)$.
-The number of securities bought to make this trade is given by the vector ($\betavec_{\rho_2} - \betavec_{\rho_1}$), and the expected payoff of the securities are given by $\muvec_\pi$.
-Thus, by equation \ref{eq:profit_decomposition}, the {\em net profit} of the trader is equal to $K(\pi||\rho_1) - K(\pi||\rho_2)$, {\it i.e.}, the reduction in KL-divergence with respect to the true distribution.
-We note one useful property of this construction: For a fixed vector $\betavec$ of purchased securities, the cost is independent of the outcome
-(and outcome distribution $\pi$), while the payoff is independent of the initial market state in which these securities were purchased.
-
-%Thus, we have shown that there exist deep connections between learning the natural parameters of an exponential family distribution and cost function based prediction markets.
-In Chapter \ref{nonmyopic} we exploit this connection and propose a bounded regret learning algorithm in a Bayesian model under less-than-ideal conditions: where the experts may be adversarial and strategic.
-
-\subsubsection{Budget of Informative Traders}\label{sec:inf}
-Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
-
-%For this, we need the following simple extension of Jensen's inequality that follows directly from the definition of convexity of a function.
-%\begin{lemma}\label{lem:convex}
-%For a strictly convex function $f$ with unique minimum $x_{min}$, $x\neq x_{min}$ in the domain of $f$, and $\lambda \in (0,1]$,
-%$$f(\lambda x_{min}+ (1-\lambda)x)< f(x)$$
-%\end{lemma}
-%\begin{proof}
-%\begin{eqnarray*}
-%f(\lambda x_{min}+ (1-\lambda)x) &<& \lambda f(x_{min})+ (1-\lambda)f(x)\\
-%&=& \lambda(f(x_{min})-f(x)) +f(x)\\
-%&\leq& f(x)
-%\end{eqnarray*}
-%\end{proof}
-
-We now characterize the expected change in budget for an informative trader. The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
-\begin{theorem}\label{thm:growth}
- Suppose that each informative trader gets a random sample of data, 
-resulting in a sequence of trader beliefs $\qvec_i$, and hence a sequence of budget-limited market positions
-$\tilde{\qvec}_i$, before a final outcome $x$.
-Then, the expectation, over trader $i$'s belief distribution $\qvec_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\tilde{\qvec}_{i-1}$.
-\end {theorem}
-\begin{proof}
-Trader $i$'s believed distribution is the distribution parametrized by $\qvec_i$.
-Therefore, the expected profit, over possible outcome values $x$, for a trader $i$ in a given round is given by
-\begin{eqnarray*}
-&& E[C(\tilde{\qvec}_{i-1})-C(\tilde{\qvec}_{i})-(\tilde{\qvec}_{i-1}-\tilde{\qvec}_{i})\Phi(x)]\\
-&=& K(\qvec_{i}||\tilde{\qvec}_{i-1})] + H(\qvec_{i}) - [K(\qvec_{i}||\tilde{\qvec}_{i}) + H(\pi)]]\\
-&=& K(\qvec_{i}||\tilde{\qvec}_{i-1}) - K(\qvec_{i}||\tilde{\qvec}_{i})
-\end{eqnarray*}
-We recall that $\tilde{\qvec}_{i}$ can be expanded as:
-$$\tilde{\qvec_{i}} = \lambda \qvec_{i} + (1- \lambda) \tilde{\qvec}_{i-1}$$ where $\lambda$ is strictly greater than $0$, but no more than $1$.
-It is a standard result from information theory that $K(.,.)$ is a convex function, and therefore, we have:
-\[
-K(\qvec_{i}||\tilde{\qvec}_{i-1}) \leq \lambda K(\qvec_{i}||\qvec_{i}) + (1-\lambda) K(\qvec_{i}||\tilde{\qvec}_{i-1})
-\]
-The first term on the right hand side is zero, and thus, we get:
-\[
-  K(\qvec_{i}||\tilde{\qvec}_{i-1}) - K(\qvec_{i}||\tilde{\qvec}_{i}) \geq \lambda K(\qvec_{i}||\tilde{\qvec}_{i-1})
-\]
-%Thus, it is sufficient to prove that E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t}] >= E_{\qvec}[K(\pi||\tilde{\qvec}_{i}^{t}].
- \end{proof}
-%\begin{theorem}
-%The expected budget of an informative trader increases when he participates in the market. Here, expectation is taken with respect to the true distribution from which the outcome is drawn.
-%We show that the budget of an informative trader grows in expectation:\end{theorem}
-%\begin{proof}
-%Let $\pi$ be the true distribution from which the outcome $x^{t}$ is drawn.
-Let $K(\pi||\qvec)$ be the relative entropy between the true distribution $\pi$ and the predictive distribution determined by the securities vector $\qvec$.
-Then the expected change in budget for a trader $i$ in round $t$ is given by
-%\begin{eqnarray*}
-%E_{x,\qvec}[\alpha_{i}^{t}&-&\alpha_{i}^{t-1}]\\
-%&=&E_{x,\qvec}[C(\tilde{\qvec}_{i-1}^{t})-C(\tilde{\qvec}_{i}^{t})-(\tilde{\qvec}_{i-1}^{t}-\tilde{\qvec}_{i}^{t})^{T}\Phi(x^{t})]\\
-%&=& E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t})] + H(\pi) - E[K(\pi||\tilde{\qvec}_{i}^{t}) + H(\pi)]]
-%\end{eqnarray*}
-%Thus, it is sufficient to prove that E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t}] >= E_{\qvec}[K(\pi||\tilde{\qvec}_{i}^{t}].
-
-%In the case that
-
-%It is well known that minimizing the relative entropy of a predictive distribution with respect to the true distribution is equivalent to maximizing the likelihood function.
-% (see, for instance, \cite{whom?}).
-%In other words, the MLE of the natural parameters of an exponential family distribution minimizes the relative entropy of a predictive distribution with respect to the true distribution.
-
-%Further, for a true distribution $\pi$ and predictive distribution determined by $\qvec$, $K(\pi||\qvec) + H(\pi)$ is a convex function of $\qvec$.
-In fact, it is strictly convex when the representation is minimal.
-Thus, it has a unique minimum.
-%Further, $K(\pi||\rho) + H(\pi)$ is non-negative.
-
-%Suppose the current market state is represented by $\tilde{\qvec}_{i-1}$ and $\qvec_{i}$ represents the belief of trader $i$.
-Since the trader is budget limited, he moves the market state to $\tilde{\qvec}_{i}$ instead where $\tilde{\qvec}_{i}=  \lambda \qvec_{i}+ (1-\lambda)\tilde{\qvec}_{i-1}$.
-For an informative trader $i$, $\qvec_{i}$ represents the MLE of the natural parameters of the distribution.
-Thus, $K(\pi||\qvec) + H(\pi)$ has a unique minimum at $\qvec_{i}$.
-
-%Thus,   from Lemma \ref{lem:convex}, $E[K(\pi||\tilde{\qvec}_{i-1}^{t})+ H(\pi)] > E[K(\pi||\tilde{\qvec}_{i}^{t})+ H(\pi)]$ and trader $i$'s expected budget increases.
-%\end{proof}
-
-For continuous distributions with a density, the probability that a trader with private information will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs.
-This result suggests that, eventually, every informative trader will have the ability to influence the market state in accordance with his beliefs, without being budget limited.
-
-We note one important aspect of Theorem~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
-This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure; depending on the distribution family and the prior distribution over parameter values, maximum likelihood estimation might not optimize the true expected score.
-If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
-%To summarize our results for this in a market with budget-limited traders, the total loss induced by malicious traders is limited, while the budget of  informative traders grows.
-\section{The MaxEnt Market Mechanism: A Bayesian View}
-\subsubsection{Conjugate Priors}
-
-%In the previous section it's assumed that the agent has a belief $\q$, and moves the market state from initial state $\q_{init}$ to a convex combination $\tilde{\q} = \lambda \q + (1-\lambda)\q_{init}$.
-This note tries to justify this based on Bayesian updating, rather than budget limitations.
-(But it doesn't quite succeed.)
-
-\medskip\noindent
-Let $p(x;\theta)$ denote a probability density drawn from an exponential family with sufficient statistic $\phi : \mX \rightarrow \bR^d$, where $\theta$ is the natural parameter:
-$$
-p(x;\theta) = \exp \left[ \la \theta, \phi(x) \ra \right],
-$$
-and
-$$
-g(\theta) = \log \int_{\mX} \exp\la\phi(x),\theta\ra d\!x.
-$$
-Recall that $\grad g(\theta) = \Exp[\phi(x)]$ and $\grad^2 g(\theta) = \Var[\phi(x)]$. The family of conjugate priors is also an exponential family and takes the form
-$$
-p(\theta; n,\nu) = \exp\left[ \la n\nu,\theta \ra - ng(\theta) - h(\nu,n) \right].
-$$
-Here the feature map is $\psi(\theta) = (\theta, -g(\theta))$, the natural parameter is $(n\nu, n)$ where $n \in \bR$ and $\nu \in \bR^d$.
-The normalizer $h(\nu, n)$ is convex in $(n\nu,n)$.
-It is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$.
-The justification for this is that
-$$
-\Exp_{\theta} \left[ \Exp_x \left[ \phi(x) | \theta \right]\right] 
-= \Exp_{\theta} \left[\grad g(\theta)\right]
-= \nu.
-$$
-%(The proof is straightforward but not obvious. I have a reference for this but it's impossible to read---it's a mathematical statistics paper.)
-
-Suppose we draw a sample $X = (x_1,\ldots,x_m)$ of size $m$, and denote the empirical mean by $\mu[X] = \sum_{i=1}^m \phi(x_i)$. The posterior distribution is then
-$$
-p(\theta|X) \propto p(X|\theta)p(\theta|n,\nu) \propto \exp\left[ \la\mu[X]+n\nu,\theta\ra - (m+n)g(\theta) \right],
-$$
-and so the posterior mean is 
-\begin{equation} \label{posterior-mean}
-\frac{m\mu[X] + n\nu}{m+n}.
-\end{equation}
-Thus the posterior mean is a convex combination of the prior and posterior means, and their relative weights depend on the phantom and empirical sample sizes.
-
-In the context of prediction markets, one could imagine an agent who uses the current market estimate as its prior, and draws a empirical sample of size $m$.
-Its belief then takes the form~(\ref{posterior-mean}), where $n$ depends on the importance the agent places on the market estimate (perhaps based on how long the market has been running).
-However, note that the \emph{mean parameter} becomes a convex combination of market state and empirical belief, and this does not translate to the \emph{natural parameter}, which is what we would have liked.
-The new natural parameter is
-$$
-\grad g^{-1} \left(\frac{m\mu[X] + n\nu}{m+n}\right) = \grad g^* \left(\frac{m\mu[X] + n\nu}{m+n}\right).
-$$
-where $g^*$ is the convex conjugate of $g$.
-
-\subsubsection{Exponential Utility}
+\section{Exponential Utility}
 
@@
 So there is a kind of congruence between exponential family beliefs and exponential utility (as the names would suggest).
-\subsection{Properties of the Maximum Entropy Market}
-Consider the dual of the cost function, $C^{*}(\mu)$ defined as
-$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
-This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
-$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
-Thus, the mean parameters are the dual variables to the natural parameters.
+It would be nice to have similar results for other utility-belief combinations.
+Table~\ref{tab:util-belief} lists many interesting/common utility functions, and the question is what beliefs to pair them with to get clear results on updating behavior.
 
-\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
-\proof
-To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
-\begin{eqnarray*}
--H(p) &=& \int_{x}p(x)\log p(x)\\
-&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
-&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
-&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
-\end{eqnarray*}
-\unproof
-\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
-\rmk The LMSR is essentially a special case applied to a multinomial distribution.
-The LMSR is known to have bounded ($\log n$) market maker loss.
-Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
-\unrmk
+%
+\begin{table}[ht]
+\begin{tabular}{llll}
+\toprule
+\multicolumn{1}{c}{Utility} &
+\multicolumn{1}{c}{Belief} & 
+\multicolumn{1}{c}{Cost function} & 
+\multicolumn{1}{c}{Result} \\ \midrule
+Linear & Arbitrary & Arbitrary & $\theta \rightarrow \htheta$ ($\mu \rightarrow \hmu$) \\
+Exponental (CARA) & Exponential Family & LMSR & $\theta \rightarrow \frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$ \\
+Mean-Variance &&& \\
+Quadratic &&& \\
+CRRA &&& \\
+Logarithmic &&& \\
+HARA &&& \\
+(Lindley) &&& \\
+\bottomrule
+\end{tabular}
+\caption{Behavior of agents with specific utility-belief combinations, in terms of how they update the market share vector $\theta$ when their own assessment is parameter $\htheta$.\label{tab:util-belief}}
+\end{table}
+%
 
-{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
-\proof
-Note that, for an exponential family distribution
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
-&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
-\end{eqnarray*}
-If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
-\unproof
-
-%{\cor The exponential family market maker has unbounded worst case loss.}
-%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
-%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
-%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
-\paragraph{Loss of the market maker}
-First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
-
-Recall the definition of the conjugate dual:
-$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
-The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
-$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
-This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
-So we have 
-$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
-Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
-
-Let $q_{0}$ be the initial and $q_{f}$ the final market state.
-Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
-\begin{eqnarray*}
-\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
-&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
-&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
-&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
-&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
-&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
-&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
-&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
-&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
-\end{eqnarray*}
-
-\subsubsection{Example: Gaussian Markets}
-{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
-In this case,  $\beta$ is a $2$-dimensional  vector.
-Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
-So
-\begin{eqnarray*}
-\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
-&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
-&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
-\end{eqnarray*}
-Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
-$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
-$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
-{\ex
-We will now derive the expression for the differential entropy for the normal distribution.
-Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
-&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
-\end{eqnarray*}
-Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
-Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
-
-{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
-Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
-Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
-Let's now work out the loss for the Gaussian market maker.
-First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
-\begin{eqnarray*}
-D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
-&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
-&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
-%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
-&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
-\end{eqnarray*}
-
-So the Gaussian market maker loss is 
-%parallels to quadratic market maker loss? 
-\begin{eqnarray*}
-D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
-&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
-&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
-\end{eqnarray*}¥
-Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
-\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
-
-\section{Connection to Mirror Descent?}
-
-\section{Conclusions}
+The behaviors in Table~\ref{tab:util-belief} have to do with how the share vector is updates.
+Recall that for a Bayesian agent with linear utility and an exponential family belief, it first updates its own mean estimate by $\hmu \rightarrow \frac{n_0}{n+n_0} \mu + \frac{n}{n+n_0} \hmu$, assuming the usual conjugate prior and that the agent takes the market estimate as prior mean.
+Here $n_0$ and $n$ are the ``phantom'' and actual samples sizes, in other words the relative weights on prior (market estimate) and private estimate.
+This belief update can occur before the further updates in the table.
+Note that the belief update occurs in mean parameter space, while the updates in the table occur in natural parameter (share vector) space, because they depend on properties of utility like risk aversion.
 

### 2014-01-31 WriteUp/Original/maxent-market.tex
tidied up main folder to just have submission doc

--- 2013-12-16 WriteUp/maxentpredmkts.tex
@@
-% Adjusted to add \usepackage[numbers]{natbib}
-% \bibliographystyle{acmsmall} and \documentclass[prodmode,acmec]{ec-acmsmall}
-% Jan 5, 2013 - David Parkes
-%
-% v2-acmsmall-sample.tex, dated March 6 2012
-% This is a sample file for ACM small trim journals
-%
-% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
-% (c) 2010 Association for Computing Machinery (ACM)
-%
-% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
-% Users can also go through the FAQs available on the journal's submission webpage.
-%
-% Steps to compile: latex, bibtex, latex latex
-%
-% For tracking purposes => this is v1.3 - March 2012
+%%%
+%%% Maximum Entropy Market Making
+%%%
 
-\documentclass[prodmode,acmec]{ec-acmsmall} % Aptara syntax
-
-% Package to generate and customize Algorithm as per ACM style
-\usepackage[ruled]{algorithm2e}
-\renewcommand{\algorithmcfname}{ALGORITHM}
-\SetAlFnt{\small}
-\SetAlCapFnt{\small}
-\SetAlCapNameFnt{\small}
-\SetAlCapHSkip{0pt}
-\IncMargin{-\parindent}
-
-% Metadata Information
-\acmVolume{9}
-\acmNumber{4}
-\acmArticle{39}
-\acmYear{2013}
-\acmMonth{6}
-\usepackage[numbers]{natbib}
-\usepackage{amsmath}
-\usepackage{amssymb}
-%\usepackage{amsthm}
-\usepackage{parskip}
-\usepackage{booktabs}
-
-%new commands
-\newcommand{\la}{\langle}
-\newcommand{\ra}{\rangle}
-\newcommand{\mX}{\mathcal{X}}
-\newcommand{\bR}{\mathbf{R}}
-\newcommand{\grad}{\nabla}
-\newcommand{\Exp}{\mathbf{E}}
-\newcommand{\Var}{\mathrm{Var}}
-\newcommand{\q}{\mathbf{q}}
-\newcommand{\mM}{\ensuremath{\mathcal{M}}}
-\newcommand{\hmu}{\ensuremath{\hat{\mu}}}
-\newcommand{\mP}{\ensuremath{\mathcal{P}}}
-\newcommand{\mY}{\ensuremath{\mathcal{Y}}}
-\newcommand{\mS}{\ensuremath{\mathcal{S}}}
-\newcommand{\ds}{\displaystyle}
-\newcommand{\mF}{\mathcal{F}}
-\newcommand{\mD}{\mathcal{D}}
-\newcommand{\htheta}{\hat{\theta}}
-%---------------------------------------------------------------------------
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{prop}[thm]{Proposition}
-\newtheorem{lem}[thm]{Lemma}
-\newtheorem{con}[thm]{Conjecture}
-\newtheorem{cor}[thm]{Corollary}
-\newtheorem{ex}[thm]{Example}
-\newtheorem{qn}[thm]{Question}
-\newcommand{\defn}{\bigbreak\noindent{\bf Definition. }}
-\newcommand{\undefn}{\bigbreak}
-\newcommand{\rmk}{\bigbreak\noindent{\bf Remark. }}
-\newcommand{\unrmk}{\bigbreak}
-%\newcommand{\proof}{\noindent{\bf Proof. }}
-\newcommand{\unproof}{\hfill$\Box$\bigbreak}
-\newcommand{\defeq}{\stackrel{\mathrm{def}}{=}}
-\newcommand{\holds}[1]{\stackrel{?}{#1}}
-\newcommand{\conv}{\mathrm{conv}}
-\newcommand{\ie} {{\it i.e. }}
-\newcommand{\eg} {{\it e.g. }}
-\newcommand{\etal} {{\it et al. }}
-\newcommand{\E}{\mathbf{E}}
-\newcommand{\betavec}{\pmb{\beta}}
-\newcommand{\qvec}{\mathbf{q}}
-\newcommand{\lpf}{\mathbf{\psi}} %logpartition function
-\newcommand{\family}{\mathcal F} 
-\newcommand{\suff}{\pmb{\phi}} % suff stats
-\newcommand{\muvec}{\pmb{\mu}}
-
-% Document starts
-
-\begin{document}
-% Page heads
-\markboth{Abernethy et al.}{Maximum Entropy Prediction Markets}
-% Title portion
-\title{Maximum Entropy Prediction Markets}
-\author{JACOB ABERNETHY
-\affil{University of Michigan, Ann Arbor}
-SINDHU KUTTY
-\affil{University of Michigan, Ann Arbor}
-S\'{E}BASTIEN LAHAIE
-\affil{Microsoft Research, New York City}
-RAHUL SAMI
-\affil{University of Michigan, Ann Arbor}}
-%%\author{GANG ZHOU
-%%\affil{College of William and Mary}
-%%YAFENG WU
-%%\affil{University of Virginia}
-%%TING YAN
-%%\affil{Eaton Innovation Center}
-%%TIAN HE
-%%\affil{University of Minnesota}
-%%CHENGDU HUANG
-%%\affil{Google}
-%%JOHN A. STANKOVIC
-%%\affil{University of Virginia}
-%%TAREK F. ABDELZAHER
-%%\affil{University of Illinois at Urbana-Champaign}}
-% NOTE! Affiliations placed here should be for the institution where the
-%       BULK of the research was done. If the author has gone to a new
-%       institution, before publication, the (above) affiliation should NOT be changed.
-%       The authors 'current' address may be given in the "Author's addresses:" block (below).
-%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
-%       currently affiliated with NASA.
-
-\begin{abstract}
-In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets.
-We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data.
-Similarly,  the market state can be interpreted as a distribution over the outcome space.
-In particular, we show that a proper scoring rule can be derived from maximum entropy distributions.
-This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces.
-In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility.
-We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function.
-These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders.
-In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market.
-Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
-\end{abstract}
-
-\category{J.4}{Social and Behavioral Sciences}{Economics}
-\category{I.2.6} {Artificial Intelligence}{Learning}
-
-\terms{Prediction Market Design, Machine Learning Algorithms}
-
-\keywords{Exponential Families, Bayesian Learning, Maximum Entropy Distributions, generalized LMSR}
-
-\acmformat{Jacob Abernethy, Sindhu Kutty, S\'{e}bastien Lahaie, and Rahul Sami, 2014. Maximum Entropy Prediction Markets}
-%%Gang Zhou, Yafeng Wu, Ting Yan, Tian He, Chengdu Huang, John A. Stankovic, and Tarek F. Abdelzaher, 2010. A multifrequency MAC specially designed for  wireless sensor network applications.
-
-% At a minimum you need to supply the author names, year and a title.
-% IMPORTANT:
-% Full first names whenever they are known, surname last, followed by a period.
-% In the case of two authors, 'and' is placed between them.
-% In the case of three or more authors, the serial comma is used, that is, all author names
-% except the last one but including the penultimate author's name are followed by a comma,
-% and then 'and' is placed before the final author's name.
-% If only first and middle initials are known, then each initial
-% is followed by a period and they are separated by a space.
-% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.
-
-\begin{bottomstuff}
-%%This work is supported by the National Science Foundation, under
-%%grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.
-%%
-%%Author's addresses: G. Zhou, Computer Science Department,
-%%College of William and Mary; Y. Wu  {and} J. A. Stankovic,
-%%Computer Science Department, University of Virginia; T. Yan,
-%%Eaton Innovation Center; T. He, Computer Science Department,
-%%University of Minnesota; C. Huang, Google; T. F. Abdelzaher,
-%%(Current address) NASA Ames Research Center, Moffett Field, California 94035.
-\end{bottomstuff}
-
-\maketitle
-
-\section{Introduction}
-We have dual goals in this paper.
-On the one hand, we highlight a structural similarity between prediction markets and exponential families.
-We see this as the syntax of our prediction market mechanism.
-On the other, this formulation has rich semantics as well: it allows for analysis of market behavior under various environments.
-For instance, we analyze market behavior with budget limited traders, traders that are exponential-utility-maximizers.
-\section{Notation and Definitions}
-convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesi
 \section{Maximum Entropy Market Making}
@@
 %
-Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$
-
-Note that the maximum entropy distribution is an exponential family distribution. 
- Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.
-
- We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
-  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
-   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
-   Or 
-\begin{equation}\label{eq:proper}
-\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
-\end{equation}
-
-   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
-   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
-   Rearranging the terms,
-   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
-   which is the first order condition on the convexity of $\psi(\cdot)$.
- \unproof
-
- \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters.
-That is,
-   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
-   \unrmk
-
 \noindent
@@
 
-\section{The Exponential Family Market Mechanism: A Maximum Likelihood Approach}
-In this section, we will set up a prediction market that aggregates beliefs from traders where the outcome is a continuous random variable.
-In particular, we assume that the outcome is drawn from an exponential family distribution.
-Each of the traders has access to a series of points drawn from this distribution.
-In other words, the traders have access an empirical mean of the sufficient statistics of the exponential family.
-Every trader has infinite budget so that the current market price after a trader has traded in the prediction market, exactly reflects his beliefs.
-
-%We will now provide a more formal description of the problem statement.
-%\subsection{Problem Statement}
-%\begin{description}
-%\item[Setup: ] Learning proceeds in rounds $j=1,2,\ldots$ where at the end of each round the algorithm may be provided with $x_{j}$  drawn from a set $\mathcal{X}$ according to some distribution $\mathcal{D}_{j}$, where $\mathcal{D}_{j}$ belongs to an exponential family defined by $k$ sufficient statistics.
-In each round, some sequence of $n$ traders provide their estimates of the empirical means of each of the $k$ sufficient statistics drawn from $\mathcal{D}_{j}$.
-%\item[Define: ] A learning algorithm that learns this distribution by providing a maximum likelihood estimate of its unknown natural parameters based on the empirical means.  
-%\item[Proposed solution: ] Simulate a prediction market defined by a cost function $C$, securities $s_{0},s_{1},s_{2},\ldots, s_k$ and their payoffs $\phi_i(x)$ for $i=0,1,2,\ldots,k$ for the LMSR.
-Define a correspondence between the state of the prediction market and the output of the learning algorithm.
-%\end{description}
-
-For $x_{i}\sim P_{\betavec}$, recall that the likelihood function for independently drawn data $x_1,\ldots,x_n$ is given by $\prod_{i=1}^{n} P_{\betavec}(x_i)$.
-The maximum likelihood estimate of the natural parameters $\betavec$ is the value of the natural parameters that maximizes the likelihood function.
-
-We will now set up a prediction market with log market scoring rule (LMSR) and infinite budget traders such that the market state represents the maximum likelihood estimate (MLE) of the natural parameters of an exponential family distribution.
-%Given  under some interpretation of what information the participating agents have, and how they bet.
-
-For a given exponential family distribution, the prediction market is defined as follows:
-\begin{description}
-\item[Traders] The prediction market will simulate a trader $i$ corresponding to expert $i$.
-This trader processes all information samples available to her directly or inferred from previous trades, and  forms a belief distribution such that the believed means of the sufficient statistics matches the empirical means of the sufficient statistics from the information samples.
-The trader trades in the market to maximize her expected payoff under her believed distribution.
-\item[Securities and their payoff]
-For each $i=1,2,\ldots,k$, we define a security $s_{i}$ with payoff $\phi_{i}(x)$ where $x$ is the ultimate outcome and $\Phi()$ defines the vector of sufficient statistics of the exponential family distribution according to which $x$ is drawn.
-We define an additional security $s_{0}$ with payoff $\phi_{0}(x):=a-\sum_{i=1}^{k}\phi_{i}(x)$ where $a$ is an appropriately chosen constant dependent on the range of $\Phi$ so that the payoff of $s_{0}$ is non-negative.
-\end{description}
-
-We also note that since we want non-negative payoffs, we restrict the exponential families under consideration so that the sufficient statistics are lower bounded.
-If we have a constant lower bound on the sufficient statistics, then without loss of generality we can add a constant to each sufficient statistic without changing the exponential family in any way, as the constants will be absorbed in the log-partition function.
-
-\paragraph{Arbitrage-free Property}
-We can easily show that the prediction market we have defined is arbitrage-free: there is no sequence of trades that guarantees the trader a profit under all conditions.
-First, note that the incremental cost and payoff function are additive over multiple trades, so the net profit of a sequence of trades depends only on the initial and final market position, and is independent of the actual path along which the trade takes place.
-
-Now, consider any trade (or sequence of trades) that moves the market from an initial state of $\qvec$ to a final state of $\qvec + \qvec'$.
-The cost of this trade is $C(\qvec + \qvec') - C(\qvec)$.
-The cost function of our market is the log-partition function of an exponential family, and thus, $C(.)$ is a strictly convex function~\cite{WainJordan08}.
-Thus, we have the inequality:
-\[
-   C(\qvec + \qvec') - C(\qvec)  > \qvec' \nabla C|_\qvec = \qvec' \mu,
- \]
- where $\mu$ is the mean sufficient statistics vector for the distribution with parameters $\qvec$.
-The payoff due to this trade, with outcome $x$, is $\qvec'.\phi(x)$. 
-Now, we observe that under the distribution with parameters $\qvec$, the expected payoff is $\qvec' E[\Phi(x)] = \qvec' \mu$.
-Thus, under this distribution over outcomes $x$, the expected profit is strictly less than $0$.
-This is only possible if the
-realized profit is less than $0$ for at least one outcome $x$. As this statement is true for every $\qvec$ and $\qvec'$, and the profits are path-independent, the market is arbitrage-free.
-
-%Let $\qvec^*=(q^{*}_{0},q^{*}_{1},\ldots,q^{*}_{k})$ be the number of shares of each security held by the traders.
-First we note that under the assumption of perfectly rational, risk-neutral traders with infinite budget and  beliefs as indicated above, the gradient of the cost function at this point is
- %$$\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}=\mu_{i}$$
-%where $\mu_{i}$ is believed to be the expected payoff of the $i^{th}$ security by the last trader who traded in this market.
-We show below that the instantaneous price of security $i$, $\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$ at any $\qvec=\qvec^{*}$, can be written as $\left(\frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$.
-This means that the instantaneous price of the $i^{th}$ security only depends on the difference between the number of held securities of $i$ and 0.
-This ensures that if there exists an instantaneous price that reflects the belief of a trader at any vector $\qvec^{*}$, there must exist a $\qvec'$  that also reflects his beliefs and with the desirable property that $\forall i=0,1,2,\ldots,k\ q_{i}'\geq 0$.
-This additional property ensures that no trader will be required to short sell securities to move the market state to  reflect his beliefs.
-
-\paragraph{Market State and Natural Parameters}
-We define the interpretation function 
-\begin{eqnarray}\label{eqn:int}
-I(\qvec)&=&(q_{1}-q_{0},q_{2}-q_{0},\ldots,q_{k}-q_{0})\nonumber\\
-&=&(\beta_{1}, \beta_{2},\ldots, \beta_{k})=\betavec
-\end{eqnarray}
-This allows us to interpret the state of the market in terms of a prediction on the natural parameters of the distribution.
-
-\begin{theorem}
-The natural parameter vector corresponding to the interpretation of the market state given by Equation \ref{eqn:int} is the vector of their maximum likelihood estimates.
-Further, this interpretation is unique.
-\end{theorem}
-\begin{proof}
-We observe that for $i=1,\ldots,n$,
-\begin{eqnarray*}
-\mu_{i}&=&\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int \exp(\qvec^{T}\Phi(x))dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}\phi_{0}(x)+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}(a-\sum_{i=1}^{n}\phi_{i}(x))+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}a} e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log (e^{q_{0}a} \int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial (q_{0}a)}{\partial q_{i}} \frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\frac{\partial \beta_{i}}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\right)_{\rm \betavec=\betavec^{*}}
-\end{eqnarray*}
-Thus, this choice of parameters satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$. We will now argue that this vector is unique.
-
-If the exponential family is represented so that there is a unique parameter vector associated with each distribution, the representation is said to be minimal.
-The Bernoulli, Gaussian, and Poisson distributions all have minimal representations.
-Now, for an exponential family whose representation is minimal, the gradient is an injection \cite[Prop.
-3.2]{WainJordan08}.
-Thus, there is a unique parameter vector $\betavec$ that satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$, for each $i\in\{1,\ldots,n\}$.
-Thus, if the $\mu_{i}$'s correspond to the empirical means, since our choice of $\betavec$ satisfies the equality, it must also be the vector corresponding to the MLE.
-\end{proof}
-
-Thus, we have shown that this prediction market (along with its strategic traders) aggregates trader information in a way that the market state can be interpreted as a predictive distribution (the maximum likelihood distribution) over an infinite outcome space.
-
-It is worth noting here that if we can bound the parameter space and the sufficient statistics of the distribution, the market maker defined here has bounded worst case loss: The probability density, and hence log loss, would then be bounded for all market states that could be reached and all outcomes.
-In fact, for many practical applications, a distribution with unbounded parameter space is an approximation for a true space which itself may be bounded.
-In this case, worst case loss would thus be bounded.
-
-\subsection{The Exponential Family Market Mechanism in Adversarial Markets}
-In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
-
-In this section, we consider the the prediction market setup with traders that may be either informative or malicious.
-The malicious traders may want to inject faulty information into the market.
-The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.
-
-We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market.
-Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
-
-\subsubsection{Budget-limited Aggregation}
-Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
-We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market.
-This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader.
-Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.
-
-%There is distribution under which expected profit is $\leq 0$. Cost function is convex.
-
-%For any $\qvec$ there is a
-In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
-Let $\alpha$ be the budget of a trader in the market.
-Suppose with infinite budget, the trader would have moved the market state from $\qvec_{init}$ to $\qvec$, where $\qvec$ represents his true beliefs.
-Suppose further that $\alpha<C(\qvec)-C(\qvec_{init})$.
-In this case, we want to budget-limit the trader's influence on the market state.
-
-We define the budget-limited final market state as $\tilde{\qvec}$. 
-Here, we consider a specific functional form of $\tilde{\qvec}$: $$\tilde{\qvec}=\lambda\qvec+(1-\lambda)\qvec_{init}$$  where $$\lambda=\min\Big(1, \frac{\alpha}{C(\qvec) - C(\qvec_{init})}\Big)$$ We first show that this trade is feasible given the trader's budget:
-\begin{theorem}
-Let the current market state be given by the vector $\qvec_{init}$ and $\tilde{\qvec}=\lambda\qvec+(1-\lambda)\qvec_{init}$.
-For $\lambda=\min\Big(1, \frac{\alpha}{C(\qvec) - C(\qvec_{init})}\Big)$, the cost to the trader to move the market state from $\qvec_{init}$ to $\tilde{\qvec}$ is at most his budget $\alpha$.
-\end{theorem}
-\begin{proof}
-From the convexity of $C$, we have
- $$C(\tilde{\qvec})\leq(1-\lambda)C(\qvec_{init})+\lambda C(\qvec)$$
-
-Now 
-\begin{eqnarray*}
-C(\tilde{\qvec}) - C(\qvec_{init}) &\leq& (1-\lambda)C(\qvec_{init})\\&&+\lambda C(\qvec) - C(\qvec_{init})\\
-&=& \lambda\ (C(\qvec) - C(\qvec_{init}))
-\end{eqnarray*}
-
-%So when $\alpha<C(\qvec)-C(\qvec_{init})$, we pick $\lambda$ so that 
-%$$\lambda\leq \frac{\alpha}{(C(\qvec) - C(\qvec_{init}))}$$
-%
-%where we want $\lambda$ such that
- Thus, $C(\tilde{\qvec}) - C(\qvec_{init}) \leq \alpha$. 
- \end{proof}
-
- We note that moving to $\tilde{\qvec}$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
-In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\tilde{\qvec}$.
-Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\qvec$.
-\subsubsection{Damage Bound}
-Given this restriction on the initial budget of traders, we can now show that the total loss that can be induced by malicious traders is bounded.
-We define the loss function for $\qvec$ shares held as:
-$$L(\qvec,x)=-\log(p_{\betavec}(x))=\log\frac{\int e^{\qvec^{T}\Phi(x)}\ dx}{e^{\qvec^{T}\Phi(x)}}$$
-with the correspondence between $\betavec$ and $\qvec$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
-
-%{\bf Setup:} The market evolves in rounds $t=1,\ldots,T$.
-The end of a time segment occurs either when each expert has made his report or when the actual value of the random variable is received as input.
-Each expert $i$ reports the means of the sufficient statistics as available to him at time $t$.
-These are combined with the previously reported means to yield $\mu_{0}^{t}(x),\mu_{1}^{t}(x),\ldots,\mu_{k}^{t}(x)$ and an associated weight $w_{i}^{t}$ which can be interpreted as the number of points from which these values have been determined.
-These sufficient statistics correspond to a particular value of $B$ and hence of $\qvec$, the number of shares of each security that are held by traders.
-Since trader $i$ has finite budget $\alpha_{i}^{t}$ at time $t$, the current market state is moved to $\qvec'$ instead, so that cost of moving it to $\qvec'$ is within his budget.
-
-%For an initial positive budget of $\alpha_{i}^{0}$, we perform updates to the budgets of each agent as
-%$$\alpha_{i}^{t}:=\alpha_{i}^{t-1}+\lambda_{i}^{t}(L(\tilde{q}_{i-1}^{t},x^{t})-L({q}_{i}^{t},x^{t}))$$ where we define
-%$$\lambda_{i}^{t}:=\min\Big(1, \frac{\alpha_{i}^{t-1}}{C(\qvec_{i}) - C(\tilde{\qvec}_{i-1})}\Big)$$
-
-Suppose the prediction market runs over multiple rounds $t$.
-Let $\qvec_{0}^{t}$ be the initial number of shares of each security that are held.
-Let $\tilde{\qvec}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
-Let us assume that at this point we receive the value of the random variable $x^{t}$.
-
-%Let the loss function be $L(\cdot,\cdot)$, where the two arguments are the vector representing the number of shares held of each security and the actual value of the random variable.
-
-%We want to show:
-%$$L({\qvec}_{0}^{t},x^{t})-L(\tilde{\qvec}_{k}^{t},x^{t})+\sum_{i}\alpha_{i}^{t-1}\leq \sum_{i}\alpha_{i}^{t}$$
-%That is, the loss due to incorporating the budget-limited information of the traders, is not more than the increase in their budgets.
-Over multiple instances of the prediction market, we can track the change in budget of each trader. The change in budget for trader $i$ is
-\begin{eqnarray*}
-\alpha_{i}^{t}-\alpha_{i}^{t-1}&=&C(\tilde{\qvec}_{i-1}^{t})-C(\tilde{\qvec}_{i}^{t})-(\tilde{\qvec}_{i-1}^{t}-\tilde{\qvec}_{i}^{t})^{T}\Phi(x^{t})\\
-&=&L(\tilde{\qvec}_{i-1}^{t},x^{t})-L(\tilde{\qvec}_{i}^{t},x^{t})
-\end{eqnarray*}
-
-Define the myopic impact of a trader $i$ in segment $t$ as
-$$\Delta_{i}^{t}:=L(\tilde{\qvec}_{i-1}^{t},x^{t})-L(\tilde{\qvec}_{i}^{t},x^{t})$$
-Thus, the myopic impact captures incremental gain due to the trader in a round. Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
-
-The total myopic impact due to all $k$ active traders is given by
-$$\Delta^{t}=L(\qvec_{0}^{t},x^{t})-L(\tilde{\qvec}_{k}^{t},x^{t})$$
-Thus $-\Delta^{t}$ captures the incremental loss due to the predictive probability after aggregation of all $k$ traders.
-
-%The myopic impact of a trader $i$ in segment $t$, is given by
-%$$\Delta_{i}^{t}:=L(\tilde{B}_{i-1}^{t},x^{t})-L(\tilde{B}_{i}^{t},x^{t})$$
-%where $\tilde{B}_{i}$ is the vector of sufficient statistics corresponding to the the budget-limited state of the market as induced by trader $i$.
-%The total myopic impact due to all $k$ active traders is thus given by
-%$$\Delta^{t}=L(B_{0}^{t},x^{t})-L(\tilde{B}_{k}^{t},x^{t})$$
-
-%If the loss function is convex in the first argument, then we have:
-%\begin{eqnarray*}
-%\Delta_{i}^{t}&=&L(\tilde{B}_{i-1}^{t},x^{t})-L(\tilde{B}_{i}^{t},x^{t})\\
-%&=&L(\tilde{q}_{i-1}^{t},x^{t})-L(\tilde{q}_{i}^{t},x^{t})\\
-%&=&L(\tilde{q}_{i-1}^{t},x^{t})-L(\lambda_{i}{q}_{i}^{t}+(1-\lambda_{i})\tilde{q}_{i-1}^{t},x^{t})\\
-%&\geq&L(\tilde{q}_{i-1}^{t},x^{t})-\lambda_{i}L({q}_{i}^{t},x^{t})-(1-\lambda_{i})L(\tilde{q}_{i-1}^{t},x^{t})\\
-%&=&\lambda_{i}(L(\tilde{q}_{i-1}^{t},x^{t})-L({q}_{i}^{t},x^{t}))\\
-%&=& \alpha_{i}^{t} - \alpha_{i}^{t-1}
-%\end{eqnarray*}
-%Thus, the myopic damage caused by trader $i$ at round $t$ is upper-bounded by the change in his budget in that round.
-
-%If we pick a loss function that is also bounded in $[0,1]$, then we see that the budget never falls below zero. 
-\begin{theorem}
-A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
-\end{theorem}
-\begin{proof}
-Consider the myopic impact of a single trader $i$ after participating in the market $T$ times.
-Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
-$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
-
-Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
-\end{proof}
-
-This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders.
-
-In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation.
-We provide an information-theoretic justification for this claim in Section \ref{sec:infth}.
-The intuition behind this claim is that his prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
-
-\subsubsection{An Information-Theoretic Interpretation}\label{sec:infth}
- We observe a useful alternative view of the market 
-scoring rule prediction market for exponential family learning. We connect the cost, payoff and profit function to information-theoretic 
-quantities associated with the exponential family.
-
-The following result has been previously pointed out by Amari \cite{Amari-KL}. 
-\begin{theorem}\label{lem:profit_decomposition}{\bf (Profit Decomposition)}: Consider an exponential 
-family $\family$ of distributions over some set of statistics $\suff(x)$, with natural parameters $\betavec$. Let $\pi, \rho \in \family$ be any two probability distributions in the family. We use 
-$\betavec_\pi$ to denote the natural parameters of $\pi$ and $\muvec_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
-we can define $\betavec_\rho$ and  $\muvec_\rho$. We abuse notation slightly and let $\lpf(\rho)$ indicate the log partition function of $\rho$ which technically depends on its natural parameters.
-Let $H(\pi)$ denote the entropy of the distribution $\pi$, and 
- $K(\pi||\rho)$ denote the KL-divergence of $\rho$ relative to $\pi$.
-Then, the following equality holds:  
-\begin{equation}
- K(\pi||\rho) + H(\pi) = \lpf(\rho) - \betavec_\rho \cdot \muvec_\pi
-\label{eq:profit_decomposition}
-\end{equation}
-\end{theorem}
-\begin{proof}
-\begin{eqnarray*}
- K(\pi||\rho) + H(\pi) &=& \int_x \pi(x)\log{\frac{\pi(x)}{\rho(x)}}dx\\
- &&\quad - \int_x \pi(x)\log{\pi(x)}dx \\
-&=& - \int_x \pi(x)\log{\rho(x)}dx \\
-&=& - \int_x \pi(x)\left[ \betavec_\rho \cdot \suff(x) - \lpf(\rho)\right]dx\\
-&=& - \betavec_\rho \cdot \int_x \pi(x) \suff(x) dx\\
-&&\quad + \lpf(\rho) \int_x \pi(x) dx \\
-&=&  \lpf(\rho) - \betavec_\rho \cdot \muvec_\pi
-\end{eqnarray*}
-\end{proof}
-
-Equation~\ref{eq:profit_decomposition} gives us an alternative view of the
-market scoring rule construction. Assume that $\pi$ is the true distribution, and consider two arbitrary
-distributions $\rho_1, \rho_2 \in \family$.
-Note that $\lpf(\rho)$  is independent of $\pi$, and $\betavec_\rho \cdot \muvec_\pi$ is linear in the probabilities $\pi(x)$.
-If we want to measure loss by the KL-divergence, we can do so (in expectation) by setting a cost function that captures the first term, and defining security quantities ($\betavec$) and payoffs ($\suff$) to capture the second term.
-In particular, in a market with cost function $\lpf$, if the
-market price initially implies a distribution $\rho_1$, and a trader moves
-the market to price than implies a distribution $\rho_2$, then the cost she
-incurs is $\lpf(\rho_2) - \lpf(\rho_1)$.
-The number of securities bought to make this trade is given by the vector ($\betavec_{\rho_2} - \betavec_{\rho_1}$), and the expected payoff of the securities are given by $\muvec_\pi$.
-Thus, by equation \ref{eq:profit_decomposition}, the {\em net profit} of the trader is equal to $K(\pi||\rho_1) - K(\pi||\rho_2)$, {\it i.e.}, the reduction in KL-divergence with respect to the true distribution.
-We note one useful property of this construction: For a fixed vector $\betavec$ of purchased securities, the cost is independent of the outcome
-(and outcome distribution $\pi$), while the payoff is independent of the initial market state in which these securities were purchased.
-
-%Thus, we have shown that there exist deep connections between learning the natural parameters of an exponential family distribution and cost function based prediction markets.
-In Chapter \ref{nonmyopic} we exploit this connection and propose a bounded regret learning algorithm in a Bayesian model under less-than-ideal conditions: where the experts may be adversarial and strategic.
-
-\subsubsection{Budget of Informative Traders}\label{sec:inf}
-Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
-
-%For this, we need the following simple extension of Jensen's inequality that follows directly from the definition of convexity of a function.
-%\begin{lemma}\label{lem:convex}
-%For a strictly convex function $f$ with unique minimum $x_{min}$, $x\neq x_{min}$ in the domain of $f$, and $\lambda \in (0,1]$,
-%$$f(\lambda x_{min}+ (1-\lambda)x)< f(x)$$
-%\end{lemma}
-%\begin{proof}
-%\begin{eqnarray*}
-%f(\lambda x_{min}+ (1-\lambda)x) &<& \lambda f(x_{min})+ (1-\lambda)f(x)\\
-%&=& \lambda(f(x_{min})-f(x)) +f(x)\\
-%&\leq& f(x)
-%\end{eqnarray*}
-%\end{proof}
-
-We now characterize the expected change in budget for an informative trader. The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
-\begin{theorem}\label{thm:growth}
- Suppose that each informative trader gets a random sample of data, 
-resulting in a sequence of trader beliefs $\qvec_i$, and hence a sequence of budget-limited market positions
-$\tilde{\qvec}_i$, before a final outcome $x$.
-Then, the expectation, over trader $i$'s belief distribution $\qvec_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\tilde{\qvec}_{i-1}$.
-\end {theorem}
-\begin{proof}
-Trader $i$'s believed distribution is the distribution parametrized by $\qvec_i$.
-Therefore, the expected profit, over possible outcome values $x$, for a trader $i$ in a given round is given by
-\begin{eqnarray*}
-&& E[C(\tilde{\qvec}_{i-1})-C(\tilde{\qvec}_{i})-(\tilde{\qvec}_{i-1}-\tilde{\qvec}_{i})\Phi(x)]\\
-&=& K(\qvec_{i}||\tilde{\qvec}_{i-1})] + H(\qvec_{i}) - [K(\qvec_{i}||\tilde{\qvec}_{i}) + H(\pi)]]\\
-&=& K(\qvec_{i}||\tilde{\qvec}_{i-1}) - K(\qvec_{i}||\tilde{\qvec}_{i})
-\end{eqnarray*}
-We recall that $\tilde{\qvec}_{i}$ can be expanded as:
-$$\tilde{\qvec_{i}} = \lambda \qvec_{i} + (1- \lambda) \tilde{\qvec}_{i-1}$$ where $\lambda$ is strictly greater than $0$, but no more than $1$.
-It is a standard result from information theory that $K(.,.)$ is a convex function, and therefore, we have:
-\[
-K(\qvec_{i}||\tilde{\qvec}_{i-1}) \leq \lambda K(\qvec_{i}||\qvec_{i}) + (1-\lambda) K(\qvec_{i}||\tilde{\qvec}_{i-1})
-\]
-The first term on the right hand side is zero, and thus, we get:
-\[
-  K(\qvec_{i}||\tilde{\qvec}_{i-1}) - K(\qvec_{i}||\tilde{\qvec}_{i}) \geq \lambda K(\qvec_{i}||\tilde{\qvec}_{i-1})
-\]
-%Thus, it is sufficient to prove that E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t}] >= E_{\qvec}[K(\pi||\tilde{\qvec}_{i}^{t}].
- \end{proof}
-%\begin{theorem}
-%The expected budget of an informative trader increases when he participates in the market. Here, expectation is taken with respect to the true distribution from which the outcome is drawn.
-%We show that the budget of an informative trader grows in expectation:\end{theorem}
-%\begin{proof}
-%Let $\pi$ be the true distribution from which the outcome $x^{t}$ is drawn.
-Let $K(\pi||\qvec)$ be the relative entropy between the true distribution $\pi$ and the predictive distribution determined by the securities vector $\qvec$.
-Then the expected change in budget for a trader $i$ in round $t$ is given by
-%\begin{eqnarray*}
-%E_{x,\qvec}[\alpha_{i}^{t}&-&\alpha_{i}^{t-1}]\\
-%&=&E_{x,\qvec}[C(\tilde{\qvec}_{i-1}^{t})-C(\tilde{\qvec}_{i}^{t})-(\tilde{\qvec}_{i-1}^{t}-\tilde{\qvec}_{i}^{t})^{T}\Phi(x^{t})]\\
-%&=& E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t})] + H(\pi) - E[K(\pi||\tilde{\qvec}_{i}^{t}) + H(\pi)]]
-%\end{eqnarray*}
-%Thus, it is sufficient to prove that E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t}] >= E_{\qvec}[K(\pi||\tilde{\qvec}_{i}^{t}].
-
-%In the case that
-
-%It is well known that minimizing the relative entropy of a predictive distribution with respect to the true distribution is equivalent to maximizing the likelihood function.
-% (see, for instance, \cite{whom?}).
-%In other words, the MLE of the natural parameters of an exponential family distribution minimizes the relative entropy of a predictive distribution with respect to the true distribution.
-
-%Further, for a true distribution $\pi$ and predictive distribution determined by $\qvec$, $K(\pi||\qvec) + H(\pi)$ is a convex function of $\qvec$.
-In fact, it is strictly convex when the representation is minimal.
-Thus, it has a unique minimum.
-%Further, $K(\pi||\rho) + H(\pi)$ is non-negative.
-
-%Suppose the current market state is represented by $\tilde{\qvec}_{i-1}$ and $\qvec_{i}$ represents the belief of trader $i$.
-Since the trader is budget limited, he moves the market state to $\tilde{\qvec}_{i}$ instead where $\tilde{\qvec}_{i}=  \lambda \qvec_{i}+ (1-\lambda)\tilde{\qvec}_{i-1}$.
-For an informative trader $i$, $\qvec_{i}$ represents the MLE of the natural parameters of the distribution.
-Thus, $K(\pi||\qvec) + H(\pi)$ has a unique minimum at $\qvec_{i}$.
-
-%Thus,   from Lemma \ref{lem:convex}, $E[K(\pi||\tilde{\qvec}_{i-1}^{t})+ H(\pi)] > E[K(\pi||\tilde{\qvec}_{i}^{t})+ H(\pi)]$ and trader $i$'s expected budget increases.
-%\end{proof}
-
-For continuous distributions with a density, the probability that a trader with private information will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs.
-This result suggests that, eventually, every informative trader will have the ability to influence the market state in accordance with his beliefs, without being budget limited.
-
-We note one important aspect of Theorem~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
-This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure; depending on the distribution family and the prior distribution over parameter values, maximum likelihood estimation might not optimize the true expected score.
-If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
-%To summarize our results for this in a market with budget-limited traders, the total loss induced by malicious traders is limited, while the budget of  informative traders grows.
-\section{The MaxEnt Market Mechanism: A Bayesian View}
-\subsubsection{Conjugate Priors}
-
-%In the previous section it's assumed that the agent has a belief $\q$, and moves the market state from initial state $\q_{init}$ to a convex combination $\tilde{\q} = \lambda \q + (1-\lambda)\q_{init}$.
-This note tries to justify this based on Bayesian updating, rather than budget limitations.
-(But it doesn't quite succeed.)
-
-\medskip\noindent
-Let $p(x;\theta)$ denote a probability density drawn from an exponential family with sufficient statistic $\phi : \mX \rightarrow \bR^d$, where $\theta$ is the natural parameter:
-$$
-p(x;\theta) = \exp \left[ \la \theta, \phi(x) \ra \right],
-$$
-and
-$$
-g(\theta) = \log \int_{\mX} \exp\la\phi(x),\theta\ra d\!x.
-$$
-Recall that $\grad g(\theta) = \Exp[\phi(x)]$ and $\grad^2 g(\theta) = \Var[\phi(x)]$. The family of conjugate priors is also an exponential family and takes the form
-$$
-p(\theta; n,\nu) = \exp\left[ \la n\nu,\theta \ra - ng(\theta) - h(\nu,n) \right].
-$$
-Here the feature map is $\psi(\theta) = (\theta, -g(\theta))$, the natural parameter is $(n\nu, n)$ where $n \in \bR$ and $\nu \in \bR^d$.
-The normalizer $h(\nu, n)$ is convex in $(n\nu,n)$.
-It is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$.
-The justification for this is that
-$$
-\Exp_{\theta} \left[ \Exp_x \left[ \phi(x) | \theta \right]\right] 
-= \Exp_{\theta} \left[\grad g(\theta)\right]
-= \nu.
-$$
-%(The proof is straightforward but not obvious. I have a reference for this but it's impossible to read---it's a mathematical statistics paper.)
-
-Suppose we draw a sample $X = (x_1,\ldots,x_m)$ of size $m$, and denote the empirical mean by $\mu[X] = \sum_{i=1}^m \phi(x_i)$. The posterior distribution is then
-$$
-p(\theta|X) \propto p(X|\theta)p(\theta|n,\nu) \propto \exp\left[ \la\mu[X]+n\nu,\theta\ra - (m+n)g(\theta) \right],
-$$
-and so the posterior mean is 
-\begin{equation} \label{posterior-mean}
-\frac{m\mu[X] + n\nu}{m+n}.
-\end{equation}
-Thus the posterior mean is a convex combination of the prior and posterior means, and their relative weights depend on the phantom and empirical sample sizes.
-
-In the context of prediction markets, one could imagine an agent who uses the current market estimate as its prior, and draws a empirical sample of size $m$.
-Its belief then takes the form~(\ref{posterior-mean}), where $n$ depends on the importance the agent places on the market estimate (perhaps based on how long the market has been running).
-However, note that the \emph{mean parameter} becomes a convex combination of market state and empirical belief, and this does not translate to the \emph{natural parameter}, which is what we would have liked.
-The new natural parameter is
-$$
-\grad g^{-1} \left(\frac{m\mu[X] + n\nu}{m+n}\right) = \grad g^* \left(\frac{m\mu[X] + n\nu}{m+n}\right).
-$$
-where $g^*$ is the convex conjugate of $g$.
-
-\subsubsection{Exponential Utility}
-
-Assume the agent's belief distribution $p$ belongs to an exponential family, so it takes the form
-%
-\[
-p(x;\theta) = \exp[\theta x - T(\theta)]
-\]
-%
-where $\theta$ is the natural parameter and $T$ is the log-partition function.
-(I'm assuming that the sufficient statistic is $\phi(x) = x$ just for simplicity.) Assume also that the agent has an exponential utility for money $w$:
-%
-\[
-U(w) = -\frac{1}{a} \exp(-aw).
-\]
-%
-Here $a$ is the coefficient of risk aversion (higher means more risk averse, and the utility function is more concave).
-
-%
-\begin{proposition}
-An agent with exponential family belief with natural parameter $\htheta$, and exponential utility with coefficient $a$, makes a trade that moves the current market share vector $\theta$ to the convex combination $\frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$ assuming an LMSR cost function.
-\end{proposition}
-%
-%
-\begin{proof}
-Let $\delta$ be the vector of shares the agent trades.
-The payoff given eventual outcome $x$ is then $\delta x - C(\delta + \theta) + C(\theta)$.
-The utility for this payoff is as follows (recall that $\htheta$ is the agent's believed natural parameter).
-%
-\[ U( \delta x - C(\delta + \theta) + C(\theta) ) = -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)).
-\]
-%
-Taking the expected utility, we obtain
-%
-\begin{eqnarray*}
-&   & \Exp\left[ U( \delta x - C(\delta + \theta) + C(\theta) ) \right] \\
-& = & \int_{\mX} -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)) \exp[\htheta x - T(\htheta)]\, dx \\
-& = & -\frac{1}{a} \int_{\mX} \exp[(\htheta -a \delta) x + a C(\delta + \theta) - a C(\theta)) - T(\htheta)]\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)] \int_{\mX} \exp[(\htheta -a \delta) x - T(\htheta -a \delta)]\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)] \int_{\mX} p(x; \htheta -a \delta)\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)]\\
-& = & U\left(- C(\delta + \theta) + C(\theta)) - \frac{1}{a} T(\htheta - a\delta) + \frac{1}{a} T(\htheta)\right)
-\end{eqnarray*}
-%
-The second-last equality follows from the fact that $\int_{\mX} p(x; \htheta -a \delta)\, dx = 1$.
-Since utility $U$ is monotone increasing, it is maximized by maximizing its argument, which is a concave function of $\delta$ by convexity of $C$ and $T$.
-The optimality condition for the argument is
-%
-\begin{equation} \label{eq:optim}
-\grad C(\delta^* + \theta) = \grad T(\htheta - a\delta^*)
-\end{equation}
-%
-Now if the market maker is using LMSR, then $C$ is the log-partition function of the corresponding exponential family and $C = T$.
-Then~(\ref{eq:optim}) can be solved by equating the arguments.
-This leads to $\delta^* = (\htheta - \theta) / (1+a)$, which moves the share vector to $\theta + \delta^* = \frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$.
-\end{proof}
-%
-
-In the statement of the result I use the term ``LMSR cost function'' somewhat loosely, because we are not necessarily dealing with a market over exhaustive, mutually exclusive outcomes.
-What is meant is the cost function that arises by taking the dual to entropy of the maxent distribution with given mean parameter $\mu$.
-As we've discussed this seems like the right generalization of LMSR to arbitrary mean parameter spaces.
-
-Note that as $a \rightarrow 0$, we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$.
-As $a$ grows (agent grows more risk averse) the agent makes smaller and smaller trades that keep it closer to the current estimate $\theta$.
-
-So there is a kind of congruence between exponential family beliefs and exponential utility (as the names would suggest). 
-\subsection{Properties of the Maximum Entropy Market}
-Consider the dual of the cost function, $C^{*}(\mu)$ defined as
-$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
-This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
-$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
-Thus, the mean parameters are the dual variables to the natural parameters.
-
-\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
-\proof
-To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
-\begin{eqnarray*}
--H(p) &=& \int_{x}p(x)\log p(x)\\
-&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
-&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
-&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
-\end{eqnarray*}
-\unproof
-\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
-\rmk The LMSR is essentially a special case applied to a multinomial distribution.
-The LMSR is known to have bounded ($\log n$) market maker loss.
-Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
-\unrmk
-
-{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
-\proof
-Note that, for an exponential family distribution
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
-&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
-\end{eqnarray*}
-If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
-\unproof
-
-%{\cor The exponential family market maker has unbounded worst case loss.}
-%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
-%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
-%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
-\paragraph{Loss of the market maker}
-First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
-
-Recall the definition of the conjugate dual:
-$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
-The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
-$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
-This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
-So we have 
-$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
-Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
-
-Let $q_{0}$ be the initial and $q_{f}$ the final market state.
-Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
-\begin{eqnarray*}
-\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
-&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
-&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
-&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
-&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
-&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
-&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
-&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
-&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
-\end{eqnarray*}
-
-\subsubsection{Example: Gaussian Markets}
-{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
-In this case,  $\beta$ is a $2$-dimensional  vector.
-Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
-So
-\begin{eqnarray*}
-\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
-&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
-&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
-\end{eqnarray*}
-Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
-$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
-$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
-{\ex
-We will now derive the expression for the differential entropy for the normal distribution.
-Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
-&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
-\end{eqnarray*}
-Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
-Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
-
-{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
-Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
-Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
-Let's now work out the loss for the Gaussian market maker.
-First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
-\begin{eqnarray*}
-D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
-&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
-&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
-%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
-&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
-\end{eqnarray*}
-
-So the Gaussian market maker loss is 
-%parallels to quadratic market maker loss? 
-\begin{eqnarray*}
-D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
-&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
-&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
-\end{eqnarray*}¥
-Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
-\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
-
-\section{Connection to Mirror Descent?}
-
-\section{Conclusions}
-
-\end{document}
+%\footnote{There is a lack of rigor here, because there might be more than one optimal Lagrange multiplier for a given $\mu$, so the function $\theta(\mu)$ may not be well-defined.
+it is well-defined if the representation $\phi$ is minimal.
+The mapping from $\theta$ to $\mu$ is well-defined and is given by $\nabla C$.}

### 2014-01-31 WriteUp/Original/maxlike-scoring.tex
tidied up main folder to just have submission doc

--- 2013-12-16 WriteUp/maxentpredmkts.tex
@@
-% Adjusted to add \usepackage[numbers]{natbib}
-% \bibliographystyle{acmsmall} and \documentclass[prodmode,acmec]{ec-acmsmall}
-% Jan 5, 2013 - David Parkes
-%
-% v2-acmsmall-sample.tex, dated March 6 2012
-% This is a sample file for ACM small trim journals
-%
-% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
-% (c) 2010 Association for Computing Machinery (ACM)
-%
-% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
-% Users can also go through the FAQs available on the journal's submission webpage.
-%
-% Steps to compile: latex, bibtex, latex latex
-%
-% For tracking purposes => this is v1.3 - March 2012
-
-\documentclass[prodmode,acmec]{ec-acmsmall} % Aptara syntax
-
-% Package to generate and customize Algorithm as per ACM style
-\usepackage[ruled]{algorithm2e}
-\renewcommand{\algorithmcfname}{ALGORITHM}
-\SetAlFnt{\small}
-\SetAlCapFnt{\small}
-\SetAlCapNameFnt{\small}
-\SetAlCapHSkip{0pt}
-\IncMargin{-\parindent}
-
-% Metadata Information
-\acmVolume{9}
-\acmNumber{4}
-\acmArticle{39}
-\acmYear{2013}
-\acmMonth{6}
-\usepackage[numbers]{natbib}
-\usepackage{amsmath}
-\usepackage{amssymb}
-%\usepackage{amsthm}
-\usepackage{parskip}
-\usepackage{booktabs}
-
-%new commands
-\newcommand{\la}{\langle}
-\newcommand{\ra}{\rangle}
-\newcommand{\mX}{\mathcal{X}}
-\newcommand{\bR}{\mathbf{R}}
-\newcommand{\grad}{\nabla}
-\newcommand{\Exp}{\mathbf{E}}
-\newcommand{\Var}{\mathrm{Var}}
-\newcommand{\q}{\mathbf{q}}
-\newcommand{\mM}{\ensuremath{\mathcal{M}}}
-\newcommand{\hmu}{\ensuremath{\hat{\mu}}}
-\newcommand{\mP}{\ensuremath{\mathcal{P}}}
-\newcommand{\mY}{\ensuremath{\mathcal{Y}}}
-\newcommand{\mS}{\ensuremath{\mathcal{S}}}
-\newcommand{\ds}{\displaystyle}
-\newcommand{\mF}{\mathcal{F}}
-\newcommand{\mD}{\mathcal{D}}
-\newcommand{\htheta}{\hat{\theta}}
-%---------------------------------------------------------------------------
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{prop}[thm]{Proposition}
-\newtheorem{lem}[thm]{Lemma}
-\newtheorem{con}[thm]{Conjecture}
-\newtheorem{cor}[thm]{Corollary}
-\newtheorem{ex}[thm]{Example}
-\newtheorem{qn}[thm]{Question}
-\newcommand{\defn}{\bigbreak\noindent{\bf Definition. }}
-\newcommand{\undefn}{\bigbreak}
-\newcommand{\rmk}{\bigbreak\noindent{\bf Remark. }}
-\newcommand{\unrmk}{\bigbreak}
-%\newcommand{\proof}{\noindent{\bf Proof. }}
-\newcommand{\unproof}{\hfill$\Box$\bigbreak}
-\newcommand{\defeq}{\stackrel{\mathrm{def}}{=}}
-\newcommand{\holds}[1]{\stackrel{?}{#1}}
-\newcommand{\conv}{\mathrm{conv}}
-\newcommand{\ie} {{\it i.e. }}
-\newcommand{\eg} {{\it e.g. }}
-\newcommand{\etal} {{\it et al. }}
-\newcommand{\E}{\mathbf{E}}
-\newcommand{\betavec}{\pmb{\beta}}
-\newcommand{\qvec}{\mathbf{q}}
-\newcommand{\lpf}{\mathbf{\psi}} %logpartition function
-\newcommand{\family}{\mathcal F} 
-\newcommand{\suff}{\pmb{\phi}} % suff stats
-\newcommand{\muvec}{\pmb{\mu}}
-
-% Document starts
-
-\begin{document}
-% Page heads
-\markboth{Abernethy et al.}{Maximum Entropy Prediction Markets}
-% Title portion
-\title{Maximum Entropy Prediction Markets}
-\author{JACOB ABERNETHY
-\affil{University of Michigan, Ann Arbor}
-SINDHU KUTTY
-\affil{University of Michigan, Ann Arbor}
-S\'{E}BASTIEN LAHAIE
-\affil{Microsoft Research, New York City}
-RAHUL SAMI
-\affil{University of Michigan, Ann Arbor}}
-%%\author{GANG ZHOU
-%%\affil{College of William and Mary}
-%%YAFENG WU
-%%\affil{University of Virginia}
-%%TING YAN
-%%\affil{Eaton Innovation Center}
-%%TIAN HE
-%%\affil{University of Minnesota}
-%%CHENGDU HUANG
-%%\affil{Google}
-%%JOHN A. STANKOVIC
-%%\affil{University of Virginia}
-%%TAREK F. ABDELZAHER
-%%\affil{University of Illinois at Urbana-Champaign}}
-% NOTE! Affiliations placed here should be for the institution where the
-%       BULK of the research was done. If the author has gone to a new
-%       institution, before publication, the (above) affiliation should NOT be changed.
-%       The authors 'current' address may be given in the "Author's addresses:" block (below).
-%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
-%       currently affiliated with NASA.
-
-\begin{abstract}
-In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets.
-We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data.
-Similarly,  the market state can be interpreted as a distribution over the outcome space.
-In particular, we show that a proper scoring rule can be derived from maximum entropy distributions.
-This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces.
-In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility.
-We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function.
-These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders.
-In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market.
-Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
-\end{abstract}
-
-\category{J.4}{Social and Behavioral Sciences}{Economics}
-\category{I.2.6} {Artificial Intelligence}{Learning}
-
-\terms{Prediction Market Design, Machine Learning Algorithms}
-
-\keywords{Exponential Families, Bayesian Learning, Maximum Entropy Distributions, generalized LMSR}
-
-\acmformat{Jacob Abernethy, Sindhu Kutty, S\'{e}bastien Lahaie, and Rahul Sami, 2014. Maximum Entropy Prediction Markets}
-%%Gang Zhou, Yafeng Wu, Ting Yan, Tian He, Chengdu Huang, John A. Stankovic, and Tarek F. Abdelzaher, 2010. A multifrequency MAC specially designed for  wireless sensor network applications.
-
-% At a minimum you need to supply the author names, year and a title.
-% IMPORTANT:
-% Full first names whenever they are known, surname last, followed by a period.
-% In the case of two authors, 'and' is placed between them.
-% In the case of three or more authors, the serial comma is used, that is, all author names
-% except the last one but including the penultimate author's name are followed by a comma,
-% and then 'and' is placed before the final author's name.
-% If only first and middle initials are known, then each initial
-% is followed by a period and they are separated by a space.
-% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.
-
-\begin{bottomstuff}
-%%This work is supported by the National Science Foundation, under
-%%grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.
-%%
-%%Author's addresses: G. Zhou, Computer Science Department,
-%%College of William and Mary; Y. Wu  {and} J. A. Stankovic,
-%%Computer Science Department, University of Virginia; T. Yan,
-%%Eaton Innovation Center; T. He, Computer Science Department,
-%%University of Minnesota; C. Huang, Google; T. F. Abdelzaher,
-%%(Current address) NASA Ames Research Center, Moffett Field, California 94035.
-\end{bottomstuff}
-
-\maketitle
-
-\section{Introduction}
-We have dual goals in this paper.
-On the one hand, we highlight a structural similarity between prediction markets and exponential families.
-We see this as the syntax of our prediction market mechanism.
-On the other, this formulation has rich semantics as well: it allows for analysis of market behavior under various environments.
-For instance, we analyze market behavior with budget limited traders, traders that are exponential-utility-maximizers.
-\section{Notation and Definitions}
-convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesi
-\section{Maximum Entropy Market Making}
-
-The purpose of a prediction market is to elicit and aggregate the beliefs (i.e., subjective probabilities) of agents over a space of \emph{outcomes}.
-In a single-agent setting, a scoring rule is used to elicit the agent's beliefs.
-In a multi-agent setting, an information market is used to aggregate the agent's beliefs.
-Hanson~\cite{} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both so that they can perform well in both thin and thick markets.
-In this work we derive a wide-ranging generalization of the logarithmic market scoring rule.
-Using a maximum entropy approach, we explain how a market scoring rule can be developed for outcome spaces both discrete and continuous, and for generic properties of the underlying distribution (e.g., mean and variance), rather than just the probabilities of individual outcomes.
+\section{Maximum Likelihood Scoring Rules}
 
@@
 
-\subsection{The Model}
-\label{model}
+\subsection{Preliminaries}
 
-Let $\mX$ be the outcome space, which may be discrete or continuous.
-Let $\mP$ denote the set of probability distributions over the outcome space.
-We represent a probability distribution as a density $p$ absolutely continuous with respect to some base measure $\nu$ (e.g., counting for discrete outcomes, or Lebesgue for continuous outcomes).% We will discuss the interpretation of the base measure in a later section.
+We consider a state space $\mX$ which may be discrete or continuous, and a measure $\nu$ over sets of states.
+We will typically take the state space to be  $\bR$ or $\bR^d$ together with the Lebesgue measure (perhaps restricted to some subset of states), or some finite set together with the counting measure.\footnote{We will not need to make reference to the underlying $\sigma$-algebra, but to complete the specification of the measure space we take it to be the collection of Borel sets in the case of $\bR$ and $\bR^d$, and the power set in the case of a discrete set.} A \emph{probability density} is a $\nu$-measurable function $p : \mX \rightarrow \bR_+$ that integrates to 1 over $\mX$.
 
-There is an unknown distribution $p$ over outcomes.
-We are interested in estimating the expected value of different outcome \emph{statistics} under this distribution by aggregating the beliefs of agents.
-A statistic is a real-valued function $\phi_s : \mX \rightarrow \bR$ over outcomes, independent of $p$; here $s$ belongs to a finite index set $\mS$ of size $d$.
-The collection of functions $\phi = (\phi_s,\, s \in \mS)$ can be viewed as a vector-valued statistic mapping outcomes into $\bR^d$.
-Formally, the aim is to estimate
+Let $\mP$ denote the set of all probability densities over the given measure space.
+We are interested in eliciting information about some unknown density $p \in \mP$ which corresponds to the beliefs of an expert.
+The expert may be familiar with $p$ in its entirety, or only with certain properties of $p$.
+Given a set of densities $\mD \subseteq \mP$, a \emph{property}  is a real (possibly vector-valued) function $\Gamma : \mD \rightarrow \bR^k$.
+With a slight abuse of terminology we call elements of its image $\Theta = \Gamma(\mD)$ \emph{properties} and we say that $p$ has property $\theta$ when $\Gamma(p) = \theta$.
+If $\Gamma$ is one-to-one, we instead call $\Gamma$ a \emph{parametrization} and refer to the elements of its image $\Theta$ as \emph{parameters}.
+
+To be concrete, our outcome space could for instance be $\bR$ with $\nu$ the Lebesgue measure restricted to $[0,+\infty)$.
+The set $\mP$ then consists of all probability densities with support contained in the non-negative reals, and $\mD \subset \mP$ might consist of the densities of exponential distributions.
+An example of a property over $\mP$ is the mapping from a density to its mean.\footnote{The property may not be defined for all $p \in \mP$.
+In this case we implicitly restrict it to the subset of $\mP$ for which it is defined whenever its domain $\mD$ is not made explicit.} When restricted to the exponential densities $\mD$, this mapping is a parametrization.
+In much of this work, as in this example, the purpose of the measure $\nu$ is to implicitly encode bounds on the support of possible densities.
+The reason we use this approach, rather than explicitly stating bounds, is that it allows for clearer connections with maximum entropy estimation later on.
+
+Given the set of relevant densities $\mD \subseteq \mP$, which contains the expert's belief $p$, and a property $\Gamma$ over this set, the expert is asked to report $\theta = \Gamma(p)$.
+The expert is rewarded according to a \emph{scoring rule} $S : \Theta \times \mX \rightarrow \bR \cup \{-\infty\}$ which pays it $S(\htheta,x)$ according to how well its report $\htheta \in \Theta$ agrees with the eventual outcome $x \in \mX$.
+The idea is to design the scoring rule so that the expert is incentivized to report its true belief, setting $\htheta = \theta$.
 %
-\begin{equation} \label{constraint}
-\Exp_p[\phi(x)] = \mu.
+\begin{definition}
+Given a property $\Gamma : \mD \rightarrow \bR^k$, a scoring rule $S$ is  \emph{proper} over domain $\mD$ if for each $\theta \in \Theta$ and $p \in \mD$ with property $\theta$, we have
+%
+\begin{equation} \label{eq-proper}
+\Exp_p[S(\theta,x)] > \Exp_p[S(\htheta,x)]
 \end{equation}
 %
-We stress that we are only seek to elicit $\mu \in \bR^d$, not the full distribution $p$.
-As an example, suppose that $\mX = \bR$.
-If we were interested in eliciting the first two uncentered moments of $p$, we would include $\phi_1(x) = x$ and $\phi_2(x) = x^2$ as statistics.
-Note that the variance cannot be directly obtained through any single statistic: it corresponds to the expectation of $(x - \Exp_p[x])$, which depends on $p$, violating our definition of a statistic.
-Of course, the variance can be indirectly obtained by eliciting the first two moments separately.\footnote{This is related to the notion of \emph{elicitation complexity} of statistical properties introduced by Lambert et al.~\cite{Lambert}, but we do not pursue this connection here.}
+for all $\htheta \neq \theta$. If the domain is $\mD = \mP$, the set of all possible densities, then we say the scoring rule is \emph{strongly proper}.
+\end{definition}
+%
 
-There are two approaches to eliciting an agent's estimated $\mu$ in~(\ref{constraint}).
-The first approach is to incentivize the agent to directly report $\mu$ via a \emph{scoring rule}.
-The second approach is to form an \emph{information market} by issuing securities for each statistic $s \in \mS$, with payoff depending on the realized value of the statistic.
-(Such securities are known as contingent claims.) The amount of shares of each security acquired by the agent indirectly reveals it estimate $\mu$.
-In a market scoring rule~\cite{Hanson}, these two approaches are dual to each other in a formal sense.
+\begin{itemize}
+\item Note that the inequality is strict, in the literature these rules are called \emph{strictly} proper, but we only consider strictly proper rules here so don't make this qualification.
+\item Lambert et al.'s rules are actually strongly proper, however we consider scoring rules over vector valued properties.
+Lambert et al only develop these by adding together different scoring rules for each property.
+We go further in our constructions and develop rules that are not necessarily separable.
+\item It is useful to consider rules that are not strongly proper because we might have constructions for these for properties where no strongly proper rule is known.
+\item It is important to understand the informational requirements placed on the expert.
+Under a proper scoring rule, the expert does not need to be aware of the entire density $p$ to be incentivized to report its property $\theta$, since $\theta$ is an optimal report no matter what the density might be.
+The agent only needs to agree that the density indeed lies in $\mD$.
+\item For instance, if we are considering densities with support on $[0,+\infty)$, and $\mD$ is the set of lognormal densities, then a proper scoring rule over $\mD$ for the mean does not require the agent to know anything beyond the mean (e.g., it does not need to know the variance).
+\item A strongly proper scoring rule goes further; with such a rule we do not require the expert to believe the density is drawn from some subset $\mD$, only that it agrees with the support implied by the base measure $\nu$.
+Therefore, a strongly proper scoring rule for the mean here would elicit the mean no matter what kind of density over the positive reals might apply (e.g., exponential, Pareto, lognormal).
+\item In practical scenarios this seems like an important extension, because while it might be clear which states are possible, it might be difficult to ensure the expert agrees that the probability density for states falls within some restricted family $\mD$.
+\end{itemize}
 
@@
 
-\subsection{Scoring Rule}
+\subsection{Maximum Likelihood}
 
-We first consider how to directly elicit $\mu$ via a scoring rule. 
-%Note that the usual definition of a scoring rule presumes that a full probability distribution is elicited, rather than just some of its properties (e.g., mean and variance), so below we will adapt the definition to our setting.
-The space of feasible reports from an agent corresponds to
 %
-\begin{equation} \label{marginal}
-\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}.
+\begin{theorem} \label{maxlike}
+The logarithmic scoring rule defined by
+%
+\[ S(\theta, x) = \log p(x;\theta) \]
+%
+is weakly proper for $\mF$ over $\Theta$ if and only if the maximum likelihood estimate for $\theta$ is consistent.
+Furthermore, if $\mM$ is an alternative parameter space such that there exists a bijection $t : \mM \rightarrow \Theta$, then $\bar{S}(\mu, x) = S(t(\mu), x)$ is weakly proper for $\mF$ over $\mM$.
+\end{theorem}
+%
+
+%
+\begin{example}
+Suppose $X$ is supported on $[0,+\infty)$ and follows an exponential distribution with density $f(x;\lambda) = \lambda e^{-\lambda x}$ parametrized by a rate $\lambda > 0$.
+The mean $\mu$ is related to the rate via the one-to-one mapping $\mu = 1/\lambda$, so the density can alternatively be parametrized by the mean.
+According to Theorem~\ref{maxlike} the following is a weakly proper scoring rule for the mean of an exponential distribution:
+%
+\begin{equation} \label{exp-scoring}
+S(\mu, x) = -\frac{x}{\mu} - \log \mu.
 \end{equation}
 %
-A scoring rule is a function $S : \mM \times \mX \rightarrow \bR$ that rewards an agent with $S(\mu,x)$ based on how its report $\mu$ agreed with the eventual outcome $x$.
-A scoring rule is \emph{proper} if for all $\mu \in \mM$, and $p \in \mP$ such that $\Exp_p[\phi(x)] = \mu$, we have
+To verify this, suppose the agent believes $X$ follows an exponential distribution with rate $\lambda_0$, equivalently mean $\mu_0 = 1/\lambda_0$.
+By the one-to-one relation between rate and mean, choosing $\mu$ to maximize the expected score~(\ref{exp-scoring}) is equivalent to choosing $\lambda$ to maximize $\Exp{-\lambda x + \log \lambda} = -\lambda \mu_0 + \log \lambda$.
+The latter is strictly concave in $\lambda$, and using basic calculus its unique global maximum is $\lambda^* = 1/\mu_0$, which therefore leads the agent to report $\mu = \mu_0 = 1/\lambda^*$ under scoring rule~(\ref{exp-scoring}).
+\end{example}
 %
-\begin{equation} \label{scoring}
-\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)]
+
+%
+\begin{example}
+Suppose that $X$ is supported on $[1,+\infty)$ and follows a Pareto distribution with density $f(x;\alpha) = \alpha/x^{\alpha+1}$ parametrized by an index $\alpha > 0$.
+The mean $\mu$ is related to the index via the one-to-one mapping $\mu = \frac{\alpha}{\alpha-1}$, so the density can alternatively be parametrized by the mean.
+Theorem~\ref{maxlike} gives the following weakly proper scoring rule for the mean of a Pareto distribution with support on $[1,+\infty)$:
+%
+\begin{equation} \label{pareto-scoring}
+S(\mu,x) = \log \frac{\mu}{\mu-1} - \left(\frac{\mu}{\mu-1} + 1\right) \log x.
 \end{equation}
 %
-for all $\hmu \neq \mu$; the rule is \emph{strictly proper} if the inequality is strict.
-Note that in the literature, scoring rules are defined over entire probability distributions, not just properties of those distributions.
-Our definition imposes a minimum of informational requirements on agents, because an agent does not need to assess the full distribution $p$ that might have lead to its estimates $\mu$ in order to realize that truthful reporting is an optimal strategy.
-
-Our scoring rule is implicitly defined via the solution of a mathematical program. The program computes the maximum entropy distribution consistent with the agent's reported beliefs $\mu$.
-%
-\begin{eqnarray}
-\max_{p \geq 0} & \ds - \int_{x \in \mX} p(x) \log p(x) \nu(dx) & \label{obj} \\
-\mbox{s.t.} & \ds \int_{x \in \mX} \phi(x) p(x) \nu(dx) = \mu & \label{mean} \\
- & \ds \int_{x \in \mX} p(x) \nu(dx) = 1 & \label{normalize}
-\end{eqnarray}
-%
-The non-negativity constraints together with~(\ref{normalize}) ensure that the solution is a probability distribution.
-Constraints~(\ref{mean}), which correspond to $d$ separate constraints, ensure consistency with the agent's report.
-The objective~(\ref{obj}) is the entropy of the solution $p$ with respect to the measure $\nu$.
-The program has a convex objective and linear constraints.
-%For continuous outcome spaces we will see in the next section that the optimal solution can still be cleanly characterized.
-%
-\begin{theorem} \label{thm:scoring}
-Let $p(x;\mu)$ be the optimal solution to the maximum entropy program, given report $\mu \in \bR^d$. The scoring rule defined by
-%
-\begin{equation} \label{scoring}
-S(\mu,x) = a_x + b\log p(x;\mu),
-\end{equation}
-%
-where $b,a_x \in \bR$ and $b > 0$, is strictly proper.
-\end{theorem}
-%
-Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$
-
-Note that the maximum entropy distribution is an exponential family distribution. 
- Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.
-
- We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
-  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
-   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
-   Or 
-\begin{equation}\label{eq:proper}
-\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
-\end{equation}
-
-   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
-   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
-   Rearranging the terms,
-   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
-   which is the first order condition on the convexity of $\psi(\cdot)$.
- \unproof
-
- \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters.
-That is,
-   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
-   \unrmk
-
-\noindent
-This rule represents a generalization of the logarithmic scoring rule for probability distributions to any properties of distributions that can be captured as expectations of statistics.
-As explained in the next section, the solution to the maximum entropy program takes the form of an \emph{exponential family} distribution.
-For many common properties of interest, these distributions are familiar; for example, the maximum entropy distribution with a given mean is an exponential distribution, and the maximum entropy distribution with a given mean and variance is a normal distribution~\cite{}.
-We provide here two examples in more depth.
-
-%%
-%%
-
-\paragraph{Multinomial Distribution}
-
-As a first example, consider the problem of estimating the individual probabilities of a finite set of outcomes.
-We have $\mX = \{1,\ldots,k\}$.
-The relevant statistics indicate which outcome actually occurs, so the index set is $\mS = \mX$.
-Statistic $\phi_x(x')$ is 1 if $x' = x$ and 0 otherwise.
-Observe that if $p$ is the distribution over outcomes, then $\Exp_p[\phi(x)] = p$.
-A feasible report from an agent is any non-negative $\pi = (\pi_1,\ldots,\pi_k)$ such that $\sum_{i=1}^k \pi_i = 1$.
-Given a report of $\mu = \pi$, the unique solution to the maximum entropy program is $p(x; \pi) = \pi$, and~(\ref{scoring}) corresponds to the classic logarithmic scoring rule of $S(\pi,x) = \log \pi_x$.
-\hfill{$\Box$}
-
-\paragraph{Normal Distribution}
-
-As a second example, suppose the outcome space is $\mX = \bR$ and that we are interested in estimating the mean $\mu$ and variance $\sigma^2$ of the underlying distribution.
-We choose the first and second (uncentered) moments, $\kappa_1 = \mu$ and  $\kappa_2 = \mu^2 + \sigma^2$ as our statistics.
-It is well-known that the maximum entropy distribution with a given mean and variance is the normal distribution, so the solution to the maximum entropy program (using Lebesgue as the base measure $\nu$) is
-%
-\[ p(x; \kappa) = \frac{1}{\sqrt{2\pi\sigma^2}} \exp\left[ -\frac{(x-\mu)^2}{2\sigma^2} \right]. \]
-%
-Applying Theorem~\ref{thm:scoring} and choosing appropriate constants, we find that the following scoring rule is strictly proper in this context. %
-\[ S(\kappa, x) = -\frac{(x-\mu)^2}{\sigma^2} - \log \sigma^2. \]
-%
-Note that the fact that this scoring rule is proper for the mean amounts to the well-known fact that reporting the mean is a Bayes act under quadratic loss~\cite{someone}. \hfill{$\Box$}
-
-\medskip\noindent
-A brief observation on parametrizations is in order.
-Note that the maximum entropy distribution here is parametrized by the uncentered moments $\kappa = (\kappa_1,\kappa_2)$ rather than $(\mu,\sigma^2)$ because the variance is not a valid statistic.
-However, it is clear that the agent does not have to directly report $\kappa$: the market maker itself can do the translation from reported mean and variance to uncentered moments.
-This is relevant because the $\kappa$ parametrization may be unintuitive for an agent, especially given the constraints $\kappa_2 - \kappa_1^2 \geq 0$ that must be enforced.
-In the $(\mu, \sigma^2)$ parametrization each parameter is unrestricted, and the parameters have intuitive interpretations.
-
-Our approach so far has been to identify relevant properties of the unknown distribution, and elicit them via a scoring rule computed through a maximum entropy program.
-However, it is the case that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}.
-Therefore, we see the following alternative approach.
-The market designer can begin by considering the outcome space $\mX$, and assessing what family of distributions over $\mX$ the unknown $p$ might come from.
-If this family is an exponential family, then Theorem~\ref{thm:scoring} immediately gives a proper scoring rule to elicit the parameters of the distribution.
-Many of the most familiar distributions are exponential families, including the normal, exponential, gamma, beta, binomial, Poisson, Weibull, and Dirichlet.
-To illustrate, we consider two examples in more depth.
-
-%%
-%%
-
-\paragraph{Exponential}
-Suppose the designer would like an estimate of the rate at which servers fail in a data center.
-An outcome is the time between failures, so $\mX = [0,+\infty)$.
-The usual distribution for failure times is the exponential distribution $p(x;\lambda) = \lambda e^{-\lambda x}$, which is parametrized by the rate $\lambda$ and has mean $\mu = 1/\lambda$.
-The exponential distribution is the maximum entropy distribution given a fixed mean $\mu$.
-From Theorem~\ref{thm:scoring} we obtain the following proper scoring rule for this setting:
-%
-\[ S(\mu, x) = \log \lambda - \lambda x. \]
-%
-It is simple to check using basic calculus that an agent maximizes the expected score by reporting $\lambda = 1/\hmu$, where here $\hmu$ is its estimate of the mean.
-In this case, it is unclear a priori which of the mean or rate parametrizations is most intuitive, but the maximum entropy approach easily allows for either.
-\hfill{$\Box$}
-
-%%
-%%
-
-\paragraph{Beta}
-Suppose the designer would like an estimate of the click-through rate of an online advertisement.
-The outcome space here is $\mX = (0,1)$.
-A typical prior over a probability like this is the beta distribution, which has two parameters $\alpha, \beta > 0$ and density function $p(x; \alpha, \beta) = x^{\alpha-1}(1-x)^{\beta-1}/B(\alpha,\beta)$, where $B$ is the beta function.
-By theorem~\ref{thm:scoring} this leads to the proper scoring rule
-%
-\[ S((\alpha, \beta), x) = (\alpha-1)\log x + (\beta-1)\log (1-x) - \log B(\alpha,\beta). \]
-%
-Although this parametrization is standard for the beta, it is perhaps more intuitive to use the parametrization $\mu = \alpha/(\alpha+\beta) \in (0,1)$ and $n = \alpha+\beta > 0$.
-Here $\mu$ is the mean of the distribution---corresponding to the agent's actual estimate of the click-through rate---and $n$ is a measure of the agent's confidence in the estimate.
-\hfill{$\Box$}
-
-%%%
-%%% 
-%%%
-
-\subsection{Information Market}
-
-We next consider how to indirectly elicit $\mu$ by setting up a market of contingent claim securities.
-Under this approach, the elements of the index set $\mS$ are interpreted not as statistics but as securities.
-The payoff from one share of security $s$ when outcome $x$ occurs is given by the mapping $\phi_s$.
-Thus if the vector of shares held by the agent is $\theta \in \bR^d$, where entry $\theta_s$ corresponds the number of shares of security $s$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \theta, \phi(x) \ra$.
-As a concrete example, recall our multinomial distribution example from the previous section, where $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise.
-This means that security $x \in \mS$ pays 1 dollar if outcome $x \in \mX$ occurs, and nothing otherwise.
-(Such securities are known as Arrow-Debreu securities.) In our normal distribution example, the statistic for the mean had the mapping $\phi_1(x) = x$.
-Therefore a share of the corresponding security has a payoff that is linear in the outcome.
-(Such securities amount to futures contracts.)
-
-To extract useful information from such a securities market, we set a pricing scheme and examine the number of shares the agent chooses to acquire.
-Let $\Omega$ be the set of all portfolios (i.e., vectors of shares) that the agent can feasibly hold; we will see how this domain is determined in an instant.
-Each security $s$ has an associated price function $c_s : \Omega \rightarrow \bR$, which gives the marginal price $c_s(\theta)$ of security $s$ when the agent holds portfolio $\theta$.
-(Note that the marginal price depends on the entire portfolio $\theta$, not just the number of shares $\theta_s$.) A risk-neutral agent will choose to acquire shares up to the point where, for each share, expected payoff equals marginal price.
-Formally, if the agent acquires portfolio $\theta$, then for each $s \in \mS$ we must have
-%
-\begin{equation} \label{market}
-\Exp_p[\phi_s(x)] = c_s(\theta).
-\end{equation}
-%
-In this way, by its choice of $\theta$, the agent reveals that its belief is $\mu = c(\theta)$.
-There are several important properties that the price function $c$ should have.
-To simplify the agent's portfolio acquisition process, it should not be the case that the total cost of acquiring $\theta$ depends on the order in which shares are bought.
-This means that there should exist a \emph{cost function} $C : \Omega \rightarrow \bR$ such that $c = \nabla C$.
-The gradient $\nabla C$ must be onto $\mM$ to ensure that~(\ref{market}) always has a solution---this is the most important but technical condition to realize.
-To ensure that the solution is unique, it would also be convenient if the cost function were strictly convex.
-%
-\begin{proposition}
-The following cost function is monotone, convex, and onto $\mM$: 
-%
-\begin{equation} \label{cost}
-C(\theta) = \log \int_{x \in \mX} \exp \left[\la \theta, \phi(x) \ra\right] \nu(dx).
-\end{equation}
-%
-\end{proposition}
-%
-\noindent
-Observe that in the context of our earlier multinomial example, cost function~(\ref{cost}) is exactly the cost function for Hanson's logarithmic market scoring rule.
-Our approach here provides a generalization of this market scoring rule to markets with contingent claim securities with arbitrary payoffs, designed to elicit specific properties of distributions, beyond just the probabilities of different outcomes.
-
-Because an agent would never select a portfolio with infinite cost, the effective domain of $C$ is $\Omega = \{\theta \:|\: C(\theta) < +\infty \}$.
-In the context of the multinomial distribution example, $\Omega = \bR^d$, so that shares of each security can always be bought or sold short.
-In the context of the normal distribution example the effective domain is $\Omega = \{(\theta_1,\theta_2) \in \bR^2 \:|\: \theta_2 > 0 \}$ if we re-define the second statistic to be $\phi_2(x) = -x^2$; this means that the security has a negative payoff for each share, and consequently the agent must be compensated to acquire such shares.
-Evaluating~(\ref{cost}), the cost function takes the form $C(\theta) = \frac{\theta_1^2}{4\theta_2} - \frac{1}{2}\log(2\theta_2)$.
-
-Now, in our previous elicitation approach that used a scoring rule, the process of computing the scoring rule also provided a complete distribution $p(x;\mu)$ over outcomes, which could be used to infer other properties beyond just the agent beliefs $\mu$.
-In the current market-based approach, the agent's chosen portfolio $\theta$ can also form the basis of a distribution over outcomes.
-Consider the following distribution, represented as a density with respect to a base measure $\nu$:
-%
-\begin{equation} \label{expfam}
-p(x;\theta) = \exp\left[ \la \theta, \phi(x) \ra - C(\theta) \right].
-\end{equation}
-%
-Meaning, the probability that a subset $X \subseteq \mX$ of the outcomes occurs is $\int_{x \in X} p(x;\theta)\nu(dx)$.
-Observe that by definition~(\ref{cost}), this probability density indeed integrates to 1 over $\mX$, and it is clear that the density is non-negative as required.
-
-In the statistical literature a distribution that takes the form~(\ref{expfam}) is known as an \emph{exponential family}.
-The mapping $\phi$ is known as the \emph{sufficient statistic}, $\theta$ is the \emph{natural parameter}, and $C$ is the log-partition or \emph{cumulant} function.
-%This connection between information markets and exponential families allows us to draw on a vast statistical literature to understand the market's properties, and a growing computer science literature to apply efficient algorithms to compute costs and perform inference.
-
-%%%
-%%%
-%%%
-
-\subsection{Duality}
-
-There is a well-known duality between the maximum entropy approach and exponential families, which translates into a duality between the scoring rule and information market just developed.
-The duality implies that the approach leads to a \emph{market scoring rule}, applicable to both thin and thick multi-agent settings.
-
-It is known that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}.
-To see this, let $\theta(\mu) \in \bR^d$ be the Lagrange multiplier corresponding to constraints~(\ref{mean}) when solving the maximum entropy program given the agent report $\mu$.
-(Our choice of notation is deliberately suggestive.) Let $A(\mu)$ be the Lagrange multiplier corresponding to~(\ref{normalize}).
-From the first order necessary conditions for optimality, we find that
-%
-\[ p(x;\mu) = \exp \left[ \la \theta(\mu), \phi(x) \ra - A(\mu) - 1 \right], \]
-%
-so that the solution is indeed in exponential family form.
-Conversely, given a cumulant $C$ from an exponential family and a parameter $\theta$, we see that $p(x;\theta)$ defined according to~(\ref{expfam}) is the maximum entropy distribution under constraints $\Exp_p[\phi(x)] = \mu$ where $\mu = \nabla C(\theta)$.
-We obtain the following.
-%
-\begin{proposition}
-Let $C$ be the cumulant (i.e, cost function) for the exponential family corresponding to $\phi$, and let $\theta(\mu)$ be the optimal Lagrange multiplier for the mean constraints given an agent report of $\mu \in \mM$.
-Then the following scoring rule is equivalent to~(\ref{scoring}):
-%
-\begin{equation} \label{expscore}
-S(\mu, x) = \la \theta(\mu), \phi(x) \ra - C(\theta(\mu)).
-\end{equation}
-%
-\end{proposition}
-%
-\noindent
-The scoring rule~(\ref{expscore}) decomposes neatly into payoff and cost functions.
-The first term $\la \theta(\mu), \phi(x) \ra$ defines the agent's outcome-contingent reward, while the second term~$C(\theta(\mu))$ is the cost of acquiring a portfolio $\theta(\mu)$.
-We see here the duality between the approach of directly reporting $\mu$, or acquiring shares $\theta$: an agent reporting beliefs $\mu$ under scoring rule~(\ref{scoring}) would choose to acquire shares $\theta(\mu)$ in the information market with cost function~(\ref{cost}).
-Since~(\ref{scoring}) is a proper scoring rule, the information market with cost function~(\ref{cost}) is based on a proper market scoring rule.
-
-\section{The Exponential Family Market Mechanism: A Maximum Likelihood Approach}
-In this section, we will set up a prediction market that aggregates beliefs from traders where the outcome is a continuous random variable.
-In particular, we assume that the outcome is drawn from an exponential family distribution.
-Each of the traders has access to a series of points drawn from this distribution.
-In other words, the traders have access an empirical mean of the sufficient statistics of the exponential family.
-Every trader has infinite budget so that the current market price after a trader has traded in the prediction market, exactly reflects his beliefs.
-
-%We will now provide a more formal description of the problem statement.
-%\subsection{Problem Statement}
-%\begin{description}
-%\item[Setup: ] Learning proceeds in rounds $j=1,2,\ldots$ where at the end of each round the algorithm may be provided with $x_{j}$  drawn from a set $\mathcal{X}$ according to some distribution $\mathcal{D}_{j}$, where $\mathcal{D}_{j}$ belongs to an exponential family defined by $k$ sufficient statistics.
-In each round, some sequence of $n$ traders provide their estimates of the empirical means of each of the $k$ sufficient statistics drawn from $\mathcal{D}_{j}$.
-%\item[Define: ] A learning algorithm that learns this distribution by providing a maximum likelihood estimate of its unknown natural parameters based on the empirical means.  
-%\item[Proposed solution: ] Simulate a prediction market defined by a cost function $C$, securities $s_{0},s_{1},s_{2},\ldots, s_k$ and their payoffs $\phi_i(x)$ for $i=0,1,2,\ldots,k$ for the LMSR.
-Define a correspondence between the state of the prediction market and the output of the learning algorithm.
-%\end{description}
-
-For $x_{i}\sim P_{\betavec}$, recall that the likelihood function for independently drawn data $x_1,\ldots,x_n$ is given by $\prod_{i=1}^{n} P_{\betavec}(x_i)$.
-The maximum likelihood estimate of the natural parameters $\betavec$ is the value of the natural parameters that maximizes the likelihood function.
-
-We will now set up a prediction market with log market scoring rule (LMSR) and infinite budget traders such that the market state represents the maximum likelihood estimate (MLE) of the natural parameters of an exponential family distribution.
-%Given  under some interpretation of what information the participating agents have, and how they bet.
-
-For a given exponential family distribution, the prediction market is defined as follows:
-\begin{description}
-\item[Traders] The prediction market will simulate a trader $i$ corresponding to expert $i$.
-This trader processes all information samples available to her directly or inferred from previous trades, and  forms a belief distribution such that the believed means of the sufficient statistics matches the empirical means of the sufficient statistics from the information samples.
-The trader trades in the market to maximize her expected payoff under her believed distribution.
-\item[Securities and their payoff]
-For each $i=1,2,\ldots,k$, we define a security $s_{i}$ with payoff $\phi_{i}(x)$ where $x$ is the ultimate outcome and $\Phi()$ defines the vector of sufficient statistics of the exponential family distribution according to which $x$ is drawn.
-We define an additional security $s_{0}$ with payoff $\phi_{0}(x):=a-\sum_{i=1}^{k}\phi_{i}(x)$ where $a$ is an appropriately chosen constant dependent on the range of $\Phi$ so that the payoff of $s_{0}$ is non-negative.
-\end{description}
-
-We also note that since we want non-negative payoffs, we restrict the exponential families under consideration so that the sufficient statistics are lower bounded.
-If we have a constant lower bound on the sufficient statistics, then without loss of generality we can add a constant to each sufficient statistic without changing the exponential family in any way, as the constants will be absorbed in the log-partition function.
-
-\paragraph{Arbitrage-free Property}
-We can easily show that the prediction market we have defined is arbitrage-free: there is no sequence of trades that guarantees the trader a profit under all conditions.
-First, note that the incremental cost and payoff function are additive over multiple trades, so the net profit of a sequence of trades depends only on the initial and final market position, and is independent of the actual path along which the trade takes place.
-
-Now, consider any trade (or sequence of trades) that moves the market from an initial state of $\qvec$ to a final state of $\qvec + \qvec'$.
-The cost of this trade is $C(\qvec + \qvec') - C(\qvec)$.
-The cost function of our market is the log-partition function of an exponential family, and thus, $C(.)$ is a strictly convex function~\cite{WainJordan08}.
-Thus, we have the inequality:
-\[
-   C(\qvec + \qvec') - C(\qvec)  > \qvec' \nabla C|_\qvec = \qvec' \mu,
- \]
- where $\mu$ is the mean sufficient statistics vector for the distribution with parameters $\qvec$.
-The payoff due to this trade, with outcome $x$, is $\qvec'.\phi(x)$. 
-Now, we observe that under the distribution with parameters $\qvec$, the expected payoff is $\qvec' E[\Phi(x)] = \qvec' \mu$.
-Thus, under this distribution over outcomes $x$, the expected profit is strictly less than $0$.
-This is only possible if the
-realized profit is less than $0$ for at least one outcome $x$. As this statement is true for every $\qvec$ and $\qvec'$, and the profits are path-independent, the market is arbitrage-free.
-
-%Let $\qvec^*=(q^{*}_{0},q^{*}_{1},\ldots,q^{*}_{k})$ be the number of shares of each security held by the traders.
-First we note that under the assumption of perfectly rational, risk-neutral traders with infinite budget and  beliefs as indicated above, the gradient of the cost function at this point is
- %$$\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}=\mu_{i}$$
-%where $\mu_{i}$ is believed to be the expected payoff of the $i^{th}$ security by the last trader who traded in this market.
-We show below that the instantaneous price of security $i$, $\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$ at any $\qvec=\qvec^{*}$, can be written as $\left(\frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$.
-This means that the instantaneous price of the $i^{th}$ security only depends on the difference between the number of held securities of $i$ and 0.
-This ensures that if there exists an instantaneous price that reflects the belief of a trader at any vector $\qvec^{*}$, there must exist a $\qvec'$  that also reflects his beliefs and with the desirable property that $\forall i=0,1,2,\ldots,k\ q_{i}'\geq 0$.
-This additional property ensures that no trader will be required to short sell securities to move the market state to  reflect his beliefs.
-
-\paragraph{Market State and Natural Parameters}
-We define the interpretation function 
-\begin{eqnarray}\label{eqn:int}
-I(\qvec)&=&(q_{1}-q_{0},q_{2}-q_{0},\ldots,q_{k}-q_{0})\nonumber\\
-&=&(\beta_{1}, \beta_{2},\ldots, \beta_{k})=\betavec
-\end{eqnarray}
-This allows us to interpret the state of the market in terms of a prediction on the natural parameters of the distribution.
-
-\begin{theorem}
-The natural parameter vector corresponding to the interpretation of the market state given by Equation \ref{eqn:int} is the vector of their maximum likelihood estimates.
-Further, this interpretation is unique.
-\end{theorem}
-\begin{proof}
-We observe that for $i=1,\ldots,n$,
-\begin{eqnarray*}
-\mu_{i}&=&\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int \exp(\qvec^{T}\Phi(x))dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}\phi_{0}(x)+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}(a-\sum_{i=1}^{n}\phi_{i}(x))+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}a} e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log (e^{q_{0}a} \int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial (q_{0}a)}{\partial q_{i}} \frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\frac{\partial \beta_{i}}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\right)_{\rm \betavec=\betavec^{*}}
-\end{eqnarray*}
-Thus, this choice of parameters satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$. We will now argue that this vector is unique.
-
-If the exponential family is represented so that there is a unique parameter vector associated with each distribution, the representation is said to be minimal.
-The Bernoulli, Gaussian, and Poisson distributions all have minimal representations.
-Now, for an exponential family whose representation is minimal, the gradient is an injection \cite[Prop.
-3.2]{WainJordan08}.
-Thus, there is a unique parameter vector $\betavec$ that satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$, for each $i\in\{1,\ldots,n\}$.
-Thus, if the $\mu_{i}$'s correspond to the empirical means, since our choice of $\betavec$ satisfies the equality, it must also be the vector corresponding to the MLE.
-\end{proof}
-
-Thus, we have shown that this prediction market (along with its strategic traders) aggregates trader information in a way that the market state can be interpreted as a predictive distribution (the maximum likelihood distribution) over an infinite outcome space.
-
-It is worth noting here that if we can bound the parameter space and the sufficient statistics of the distribution, the market maker defined here has bounded worst case loss: The probability density, and hence log loss, would then be bounded for all market states that could be reached and all outcomes.
-In fact, for many practical applications, a distribution with unbounded parameter space is an approximation for a true space which itself may be bounded.
-In this case, worst case loss would thus be bounded.
-
-\subsection{The Exponential Family Market Mechanism in Adversarial Markets}
-In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
-
-In this section, we consider the the prediction market setup with traders that may be either informative or malicious.
-The malicious traders may want to inject faulty information into the market.
-The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.
-
-We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market.
-Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
-
-\subsubsection{Budget-limited Aggregation}
-Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
-We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market.
-This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader.
-Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.
-
-%There is distribution under which expected profit is $\leq 0$. Cost function is convex.
-
-%For any $\qvec$ there is a
-In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
-Let $\alpha$ be the budget of a trader in the market.
-Suppose with infinite budget, the trader would have moved the market state from $\qvec_{init}$ to $\qvec$, where $\qvec$ represents his true beliefs.
-Suppose further that $\alpha<C(\qvec)-C(\qvec_{init})$.
-In this case, we want to budget-limit the trader's influence on the market state.
-
-We define the budget-limited final market state as $\tilde{\qvec}$. 
-Here, we consider a specific functional form of $\tilde{\qvec}$: $$\tilde{\qvec}=\lambda\qvec+(1-\lambda)\qvec_{init}$$  where $$\lambda=\min\Big(1, \frac{\alpha}{C(\qvec) - C(\qvec_{init})}\Big)$$ We first show that this trade is feasible given the trader's budget:
-\begin{theorem}
-Let the current market state be given by the vector $\qvec_{init}$ and $\tilde{\qvec}=\lambda\qvec+(1-\lambda)\qvec_{init}$.
-For $\lambda=\min\Big(1, \frac{\alpha}{C(\qvec) - C(\qvec_{init})}\Big)$, the cost to the trader to move the market state from $\qvec_{init}$ to $\tilde{\qvec}$ is at most his budget $\alpha$.
-\end{theorem}
-\begin{proof}
-From the convexity of $C$, we have
- $$C(\tilde{\qvec})\leq(1-\lambda)C(\qvec_{init})+\lambda C(\qvec)$$
-
-Now 
-\begin{eqnarray*}
-C(\tilde{\qvec}) - C(\qvec_{init}) &\leq& (1-\lambda)C(\qvec_{init})\\&&+\lambda C(\qvec) - C(\qvec_{init})\\
-&=& \lambda\ (C(\qvec) - C(\qvec_{init}))
-\end{eqnarray*}
-
-%So when $\alpha<C(\qvec)-C(\qvec_{init})$, we pick $\lambda$ so that 
-%$$\lambda\leq \frac{\alpha}{(C(\qvec) - C(\qvec_{init}))}$$
-%
-%where we want $\lambda$ such that
- Thus, $C(\tilde{\qvec}) - C(\qvec_{init}) \leq \alpha$. 
- \end{proof}
-
- We note that moving to $\tilde{\qvec}$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
-In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\tilde{\qvec}$.
-Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\qvec$.
-\subsubsection{Damage Bound}
-Given this restriction on the initial budget of traders, we can now show that the total loss that can be induced by malicious traders is bounded.
-We define the loss function for $\qvec$ shares held as:
-$$L(\qvec,x)=-\log(p_{\betavec}(x))=\log\frac{\int e^{\qvec^{T}\Phi(x)}\ dx}{e^{\qvec^{T}\Phi(x)}}$$
-with the correspondence between $\betavec$ and $\qvec$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
-
-%{\bf Setup:} The market evolves in rounds $t=1,\ldots,T$.
-The end of a time segment occurs either when each expert has made his report or when the actual value of the random variable is received as input.
-Each expert $i$ reports the means of the sufficient statistics as available to him at time $t$.
-These are combined with the previously reported means to yield $\mu_{0}^{t}(x),\mu_{1}^{t}(x),\ldots,\mu_{k}^{t}(x)$ and an associated weight $w_{i}^{t}$ which can be interpreted as the number of points from which these values have been determined.
-These sufficient statistics correspond to a particular value of $B$ and hence of $\qvec$, the number of shares of each security that are held by traders.
-Since trader $i$ has finite budget $\alpha_{i}^{t}$ at time $t$, the current market state is moved to $\qvec'$ instead, so that cost of moving it to $\qvec'$ is within his budget.
-
-%For an initial positive budget of $\alpha_{i}^{0}$, we perform updates to the budgets of each agent as
-%$$\alpha_{i}^{t}:=\alpha_{i}^{t-1}+\lambda_{i}^{t}(L(\tilde{q}_{i-1}^{t},x^{t})-L({q}_{i}^{t},x^{t}))$$ where we define
-%$$\lambda_{i}^{t}:=\min\Big(1, \frac{\alpha_{i}^{t-1}}{C(\qvec_{i}) - C(\tilde{\qvec}_{i-1})}\Big)$$
-
-Suppose the prediction market runs over multiple rounds $t$.
-Let $\qvec_{0}^{t}$ be the initial number of shares of each security that are held.
-Let $\tilde{\qvec}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
-Let us assume that at this point we receive the value of the random variable $x^{t}$.
-
-%Let the loss function be $L(\cdot,\cdot)$, where the two arguments are the vector representing the number of shares held of each security and the actual value of the random variable.
-
-%We want to show:
-%$$L({\qvec}_{0}^{t},x^{t})-L(\tilde{\qvec}_{k}^{t},x^{t})+\sum_{i}\alpha_{i}^{t-1}\leq \sum_{i}\alpha_{i}^{t}$$
-%That is, the loss due to incorporating the budget-limited information of the traders, is not more than the increase in their budgets.
-Over multiple instances of the prediction market, we can track the change in budget of each trader. The change in budget for trader $i$ is
-\begin{eqnarray*}
-\alpha_{i}^{t}-\alpha_{i}^{t-1}&=&C(\tilde{\qvec}_{i-1}^{t})-C(\tilde{\qvec}_{i}^{t})-(\tilde{\qvec}_{i-1}^{t}-\tilde{\qvec}_{i}^{t})^{T}\Phi(x^{t})\\
-&=&L(\tilde{\qvec}_{i-1}^{t},x^{t})-L(\tilde{\qvec}_{i}^{t},x^{t})
-\end{eqnarray*}
-
-Define the myopic impact of a trader $i$ in segment $t$ as
-$$\Delta_{i}^{t}:=L(\tilde{\qvec}_{i-1}^{t},x^{t})-L(\tilde{\qvec}_{i}^{t},x^{t})$$
-Thus, the myopic impact captures incremental gain due to the trader in a round. Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
-
-The total myopic impact due to all $k$ active traders is given by
-$$\Delta^{t}=L(\qvec_{0}^{t},x^{t})-L(\tilde{\qvec}_{k}^{t},x^{t})$$
-Thus $-\Delta^{t}$ captures the incremental loss due to the predictive probability after aggregation of all $k$ traders.
-
-%The myopic impact of a trader $i$ in segment $t$, is given by
-%$$\Delta_{i}^{t}:=L(\tilde{B}_{i-1}^{t},x^{t})-L(\tilde{B}_{i}^{t},x^{t})$$
-%where $\tilde{B}_{i}$ is the vector of sufficient statistics corresponding to the the budget-limited state of the market as induced by trader $i$.
-%The total myopic impact due to all $k$ active traders is thus given by
-%$$\Delta^{t}=L(B_{0}^{t},x^{t})-L(\tilde{B}_{k}^{t},x^{t})$$
-
-%If the loss function is convex in the first argument, then we have:
-%\begin{eqnarray*}
-%\Delta_{i}^{t}&=&L(\tilde{B}_{i-1}^{t},x^{t})-L(\tilde{B}_{i}^{t},x^{t})\\
-%&=&L(\tilde{q}_{i-1}^{t},x^{t})-L(\tilde{q}_{i}^{t},x^{t})\\
-%&=&L(\tilde{q}_{i-1}^{t},x^{t})-L(\lambda_{i}{q}_{i}^{t}+(1-\lambda_{i})\tilde{q}_{i-1}^{t},x^{t})\\
-%&\geq&L(\tilde{q}_{i-1}^{t},x^{t})-\lambda_{i}L({q}_{i}^{t},x^{t})-(1-\lambda_{i})L(\tilde{q}_{i-1}^{t},x^{t})\\
-%&=&\lambda_{i}(L(\tilde{q}_{i-1}^{t},x^{t})-L({q}_{i}^{t},x^{t}))\\
-%&=& \alpha_{i}^{t} - \alpha_{i}^{t-1}
-%\end{eqnarray*}
-%Thus, the myopic damage caused by trader $i$ at round $t$ is upper-bounded by the change in his budget in that round.
-
-%If we pick a loss function that is also bounded in $[0,1]$, then we see that the budget never falls below zero. 
-\begin{theorem}
-A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
-\end{theorem}
-\begin{proof}
-Consider the myopic impact of a single trader $i$ after participating in the market $T$ times.
-Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
-$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
-
-Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
-\end{proof}
-
-This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders.
-
-In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation.
-We provide an information-theoretic justification for this claim in Section \ref{sec:infth}.
-The intuition behind this claim is that his prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
-
-\subsubsection{An Information-Theoretic Interpretation}\label{sec:infth}
- We observe a useful alternative view of the market 
-scoring rule prediction market for exponential family learning. We connect the cost, payoff and profit function to information-theoretic 
-quantities associated with the exponential family.
-
-The following result has been previously pointed out by Amari \cite{Amari-KL}. 
-\begin{theorem}\label{lem:profit_decomposition}{\bf (Profit Decomposition)}: Consider an exponential 
-family $\family$ of distributions over some set of statistics $\suff(x)$, with natural parameters $\betavec$. Let $\pi, \rho \in \family$ be any two probability distributions in the family. We use 
-$\betavec_\pi$ to denote the natural parameters of $\pi$ and $\muvec_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
-we can define $\betavec_\rho$ and  $\muvec_\rho$. We abuse notation slightly and let $\lpf(\rho)$ indicate the log partition function of $\rho$ which technically depends on its natural parameters.
-Let $H(\pi)$ denote the entropy of the distribution $\pi$, and 
- $K(\pi||\rho)$ denote the KL-divergence of $\rho$ relative to $\pi$.
-Then, the following equality holds:  
-\begin{equation}
- K(\pi||\rho) + H(\pi) = \lpf(\rho) - \betavec_\rho \cdot \muvec_\pi
-\label{eq:profit_decomposition}
-\end{equation}
-\end{theorem}
-\begin{proof}
-\begin{eqnarray*}
- K(\pi||\rho) + H(\pi) &=& \int_x \pi(x)\log{\frac{\pi(x)}{\rho(x)}}dx\\
- &&\quad - \int_x \pi(x)\log{\pi(x)}dx \\
-&=& - \int_x \pi(x)\log{\rho(x)}dx \\
-&=& - \int_x \pi(x)\left[ \betavec_\rho \cdot \suff(x) - \lpf(\rho)\right]dx\\
-&=& - \betavec_\rho \cdot \int_x \pi(x) \suff(x) dx\\
-&&\quad + \lpf(\rho) \int_x \pi(x) dx \\
-&=&  \lpf(\rho) - \betavec_\rho \cdot \muvec_\pi
-\end{eqnarray*}
-\end{proof}
-
-Equation~\ref{eq:profit_decomposition} gives us an alternative view of the
-market scoring rule construction. Assume that $\pi$ is the true distribution, and consider two arbitrary
-distributions $\rho_1, \rho_2 \in \family$.
-Note that $\lpf(\rho)$  is independent of $\pi$, and $\betavec_\rho \cdot \muvec_\pi$ is linear in the probabilities $\pi(x)$.
-If we want to measure loss by the KL-divergence, we can do so (in expectation) by setting a cost function that captures the first term, and defining security quantities ($\betavec$) and payoffs ($\suff$) to capture the second term.
-In particular, in a market with cost function $\lpf$, if the
-market price initially implies a distribution $\rho_1$, and a trader moves
-the market to price than implies a distribution $\rho_2$, then the cost she
-incurs is $\lpf(\rho_2) - \lpf(\rho_1)$.
-The number of securities bought to make this trade is given by the vector ($\betavec_{\rho_2} - \betavec_{\rho_1}$), and the expected payoff of the securities are given by $\muvec_\pi$.
-Thus, by equation \ref{eq:profit_decomposition}, the {\em net profit} of the trader is equal to $K(\pi||\rho_1) - K(\pi||\rho_2)$, {\it i.e.}, the reduction in KL-divergence with respect to the true distribution.
-We note one useful property of this construction: For a fixed vector $\betavec$ of purchased securities, the cost is independent of the outcome
-(and outcome distribution $\pi$), while the payoff is independent of the initial market state in which these securities were purchased.
-
-%Thus, we have shown that there exist deep connections between learning the natural parameters of an exponential family distribution and cost function based prediction markets.
-In Chapter \ref{nonmyopic} we exploit this connection and propose a bounded regret learning algorithm in a Bayesian model under less-than-ideal conditions: where the experts may be adversarial and strategic.
-
-\subsubsection{Budget of Informative Traders}\label{sec:inf}
-Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
-
-%For this, we need the following simple extension of Jensen's inequality that follows directly from the definition of convexity of a function.
-%\begin{lemma}\label{lem:convex}
-%For a strictly convex function $f$ with unique minimum $x_{min}$, $x\neq x_{min}$ in the domain of $f$, and $\lambda \in (0,1]$,
-%$$f(\lambda x_{min}+ (1-\lambda)x)< f(x)$$
-%\end{lemma}
-%\begin{proof}
-%\begin{eqnarray*}
-%f(\lambda x_{min}+ (1-\lambda)x) &<& \lambda f(x_{min})+ (1-\lambda)f(x)\\
-%&=& \lambda(f(x_{min})-f(x)) +f(x)\\
-%&\leq& f(x)
-%\end{eqnarray*}
-%\end{proof}
-
-We now characterize the expected change in budget for an informative trader. The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
-\begin{theorem}\label{thm:growth}
- Suppose that each informative trader gets a random sample of data, 
-resulting in a sequence of trader beliefs $\qvec_i$, and hence a sequence of budget-limited market positions
-$\tilde{\qvec}_i$, before a final outcome $x$.
-Then, the expectation, over trader $i$'s belief distribution $\qvec_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\tilde{\qvec}_{i-1}$.
-\end {theorem}
-\begin{proof}
-Trader $i$'s believed distribution is the distribution parametrized by $\qvec_i$.
-Therefore, the expected profit, over possible outcome values $x$, for a trader $i$ in a given round is given by
-\begin{eqnarray*}
-&& E[C(\tilde{\qvec}_{i-1})-C(\tilde{\qvec}_{i})-(\tilde{\qvec}_{i-1}-\tilde{\qvec}_{i})\Phi(x)]\\
-&=& K(\qvec_{i}||\tilde{\qvec}_{i-1})] + H(\qvec_{i}) - [K(\qvec_{i}||\tilde{\qvec}_{i}) + H(\pi)]]\\
-&=& K(\qvec_{i}||\tilde{\qvec}_{i-1}) - K(\qvec_{i}||\tilde{\qvec}_{i})
-\end{eqnarray*}
-We recall that $\tilde{\qvec}_{i}$ can be expanded as:
-$$\tilde{\qvec_{i}} = \lambda \qvec_{i} + (1- \lambda) \tilde{\qvec}_{i-1}$$ where $\lambda$ is strictly greater than $0$, but no more than $1$.
-It is a standard result from information theory that $K(.,.)$ is a convex function, and therefore, we have:
-\[
-K(\qvec_{i}||\tilde{\qvec}_{i-1}) \leq \lambda K(\qvec_{i}||\qvec_{i}) + (1-\lambda) K(\qvec_{i}||\tilde{\qvec}_{i-1})
-\]
-The first term on the right hand side is zero, and thus, we get:
-\[
-  K(\qvec_{i}||\tilde{\qvec}_{i-1}) - K(\qvec_{i}||\tilde{\qvec}_{i}) \geq \lambda K(\qvec_{i}||\tilde{\qvec}_{i-1})
-\]
-%Thus, it is sufficient to prove that E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t}] >= E_{\qvec}[K(\pi||\tilde{\qvec}_{i}^{t}].
- \end{proof}
-%\begin{theorem}
-%The expected budget of an informative trader increases when he participates in the market. Here, expectation is taken with respect to the true distribution from which the outcome is drawn.
-%We show that the budget of an informative trader grows in expectation:\end{theorem}
-%\begin{proof}
-%Let $\pi$ be the true distribution from which the outcome $x^{t}$ is drawn.
-Let $K(\pi||\qvec)$ be the relative entropy between the true distribution $\pi$ and the predictive distribution determined by the securities vector $\qvec$.
-Then the expected change in budget for a trader $i$ in round $t$ is given by
-%\begin{eqnarray*}
-%E_{x,\qvec}[\alpha_{i}^{t}&-&\alpha_{i}^{t-1}]\\
-%&=&E_{x,\qvec}[C(\tilde{\qvec}_{i-1}^{t})-C(\tilde{\qvec}_{i}^{t})-(\tilde{\qvec}_{i-1}^{t}-\tilde{\qvec}_{i}^{t})^{T}\Phi(x^{t})]\\
-%&=& E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t})] + H(\pi) - E[K(\pi||\tilde{\qvec}_{i}^{t}) + H(\pi)]]
-%\end{eqnarray*}
-%Thus, it is sufficient to prove that E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t}] >= E_{\qvec}[K(\pi||\tilde{\qvec}_{i}^{t}].
-
-%In the case that
-
-%It is well known that minimizing the relative entropy of a predictive distribution with respect to the true distribution is equivalent to maximizing the likelihood function.
-% (see, for instance, \cite{whom?}).
-%In other words, the MLE of the natural parameters of an exponential family distribution minimizes the relative entropy of a predictive distribution with respect to the true distribution.
-
-%Further, for a true distribution $\pi$ and predictive distribution determined by $\qvec$, $K(\pi||\qvec) + H(\pi)$ is a convex function of $\qvec$.
-In fact, it is strictly convex when the representation is minimal.
-Thus, it has a unique minimum.
-%Further, $K(\pi||\rho) + H(\pi)$ is non-negative.
-
-%Suppose the current market state is represented by $\tilde{\qvec}_{i-1}$ and $\qvec_{i}$ represents the belief of trader $i$.
-Since the trader is budget limited, he moves the market state to $\tilde{\qvec}_{i}$ instead where $\tilde{\qvec}_{i}=  \lambda \qvec_{i}+ (1-\lambda)\tilde{\qvec}_{i-1}$.
-For an informative trader $i$, $\qvec_{i}$ represents the MLE of the natural parameters of the distribution.
-Thus, $K(\pi||\qvec) + H(\pi)$ has a unique minimum at $\qvec_{i}$.
-
-%Thus,   from Lemma \ref{lem:convex}, $E[K(\pi||\tilde{\qvec}_{i-1}^{t})+ H(\pi)] > E[K(\pi||\tilde{\qvec}_{i}^{t})+ H(\pi)]$ and trader $i$'s expected budget increases.
-%\end{proof}
-
-For continuous distributions with a density, the probability that a trader with private information will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs.
-This result suggests that, eventually, every informative trader will have the ability to influence the market state in accordance with his beliefs, without being budget limited.
-
-We note one important aspect of Theorem~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
-This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure; depending on the distribution family and the prior distribution over parameter values, maximum likelihood estimation might not optimize the true expected score.
-If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
-%To summarize our results for this in a market with budget-limited traders, the total loss induced by malicious traders is limited, while the budget of  informative traders grows.
-\section{The MaxEnt Market Mechanism: A Bayesian View}
-\subsubsection{Conjugate Priors}
-
-%In the previous section it's assumed that the agent has a belief $\q$, and moves the market state from initial state $\q_{init}$ to a convex combination $\tilde{\q} = \lambda \q + (1-\lambda)\q_{init}$.
-This note tries to justify this based on Bayesian updating, rather than budget limitations.
-(But it doesn't quite succeed.)
-
-\medskip\noindent
-Let $p(x;\theta)$ denote a probability density drawn from an exponential family with sufficient statistic $\phi : \mX \rightarrow \bR^d$, where $\theta$ is the natural parameter:
-$$
-p(x;\theta) = \exp \left[ \la \theta, \phi(x) \ra \right],
-$$
-and
-$$
-g(\theta) = \log \int_{\mX} \exp\la\phi(x),\theta\ra d\!x.
-$$
-Recall that $\grad g(\theta) = \Exp[\phi(x)]$ and $\grad^2 g(\theta) = \Var[\phi(x)]$. The family of conjugate priors is also an exponential family and takes the form
-$$
-p(\theta; n,\nu) = \exp\left[ \la n\nu,\theta \ra - ng(\theta) - h(\nu,n) \right].
-$$
-Here the feature map is $\psi(\theta) = (\theta, -g(\theta))$, the natural parameter is $(n\nu, n)$ where $n \in \bR$ and $\nu \in \bR^d$.
-The normalizer $h(\nu, n)$ is convex in $(n\nu,n)$.
-It is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$.
-The justification for this is that
-$$
-\Exp_{\theta} \left[ \Exp_x \left[ \phi(x) | \theta \right]\right] 
-= \Exp_{\theta} \left[\grad g(\theta)\right]
-= \nu.
-$$
-%(The proof is straightforward but not obvious. I have a reference for this but it's impossible to read---it's a mathematical statistics paper.)
-
-Suppose we draw a sample $X = (x_1,\ldots,x_m)$ of size $m$, and denote the empirical mean by $\mu[X] = \sum_{i=1}^m \phi(x_i)$. The posterior distribution is then
-$$
-p(\theta|X) \propto p(X|\theta)p(\theta|n,\nu) \propto \exp\left[ \la\mu[X]+n\nu,\theta\ra - (m+n)g(\theta) \right],
-$$
-and so the posterior mean is 
-\begin{equation} \label{posterior-mean}
-\frac{m\mu[X] + n\nu}{m+n}.
-\end{equation}
-Thus the posterior mean is a convex combination of the prior and posterior means, and their relative weights depend on the phantom and empirical sample sizes.
-
-In the context of prediction markets, one could imagine an agent who uses the current market estimate as its prior, and draws a empirical sample of size $m$.
-Its belief then takes the form~(\ref{posterior-mean}), where $n$ depends on the importance the agent places on the market estimate (perhaps based on how long the market has been running).
-However, note that the \emph{mean parameter} becomes a convex combination of market state and empirical belief, and this does not translate to the \emph{natural parameter}, which is what we would have liked.
-The new natural parameter is
-$$
-\grad g^{-1} \left(\frac{m\mu[X] + n\nu}{m+n}\right) = \grad g^* \left(\frac{m\mu[X] + n\nu}{m+n}\right).
-$$
-where $g^*$ is the convex conjugate of $g$.
-
-\subsubsection{Exponential Utility}
-
-Assume the agent's belief distribution $p$ belongs to an exponential family, so it takes the form
-%
-\[
-p(x;\theta) = \exp[\theta x - T(\theta)]
-\]
-%
-where $\theta$ is the natural parameter and $T$ is the log-partition function.
-(I'm assuming that the sufficient statistic is $\phi(x) = x$ just for simplicity.) Assume also that the agent has an exponential utility for money $w$:
-%
-\[
-U(w) = -\frac{1}{a} \exp(-aw).
-\]
-%
-Here $a$ is the coefficient of risk aversion (higher means more risk averse, and the utility function is more concave).
-
-%
-\begin{proposition}
-An agent with exponential family belief with natural parameter $\htheta$, and exponential utility with coefficient $a$, makes a trade that moves the current market share vector $\theta$ to the convex combination $\frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$ assuming an LMSR cost function.
-\end{proposition}
-%
-%
-\begin{proof}
-Let $\delta$ be the vector of shares the agent trades.
-The payoff given eventual outcome $x$ is then $\delta x - C(\delta + \theta) + C(\theta)$.
-The utility for this payoff is as follows (recall that $\htheta$ is the agent's believed natural parameter).
-%
-\[ U( \delta x - C(\delta + \theta) + C(\theta) ) = -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)).
-\]
-%
-Taking the expected utility, we obtain
-%
-\begin{eqnarray*}
-&   & \Exp\left[ U( \delta x - C(\delta + \theta) + C(\theta) ) \right] \\
-& = & \int_{\mX} -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)) \exp[\htheta x - T(\htheta)]\, dx \\
-& = & -\frac{1}{a} \int_{\mX} \exp[(\htheta -a \delta) x + a C(\delta + \theta) - a C(\theta)) - T(\htheta)]\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)] \int_{\mX} \exp[(\htheta -a \delta) x - T(\htheta -a \delta)]\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)] \int_{\mX} p(x; \htheta -a \delta)\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)]\\
-& = & U\left(- C(\delta + \theta) + C(\theta)) - \frac{1}{a} T(\htheta - a\delta) + \frac{1}{a} T(\htheta)\right)
-\end{eqnarray*}
-%
-The second-last equality follows from the fact that $\int_{\mX} p(x; \htheta -a \delta)\, dx = 1$.
-Since utility $U$ is monotone increasing, it is maximized by maximizing its argument, which is a concave function of $\delta$ by convexity of $C$ and $T$.
-The optimality condition for the argument is
-%
-\begin{equation} \label{eq:optim}
-\grad C(\delta^* + \theta) = \grad T(\htheta - a\delta^*)
-\end{equation}
-%
-Now if the market maker is using LMSR, then $C$ is the log-partition function of the corresponding exponential family and $C = T$.
-Then~(\ref{eq:optim}) can be solved by equating the arguments.
-This leads to $\delta^* = (\htheta - \theta) / (1+a)$, which moves the share vector to $\theta + \delta^* = \frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$.
-\end{proof}
+To verify this, suppose the agent believes $X$ follows a Pareto distribution with index $\alpha_0$, equivalently mean $\mu_0 = \frac{\alpha_0}{\alpha_0-1}$.
+By the one-to-one relation between the index and mean, choosing $\mu$ to maximize the expected score~(\ref{pareto-scoring}) is equivalent to choosing $\alpha$ to maximize $\Exp{\log \alpha - (\alpha+1)\log x} = \log \alpha - (\alpha+1)\Exp{\log x}$.
+The latter is strictly concave in $\alpha$, and using basic calculus its unique global maximum is $\alpha^* = 1/\Exp{\log X} = \alpha_0$, where the last equality follows from the fact that $\log X$ has an exponential distribution with rate $\alpha$ when $X$ is Pareto with rate $\alpha$.
+Therefore~(\ref{pareto-scoring}) leads the agent to report $\mu = \mu_0 = \frac{\alpha^*}{\alpha^* - 1}$.
+\end{example}
 %
 
-In the statement of the result I use the term ``LMSR cost function'' somewhat loosely, because we are not necessarily dealing with a market over exhaustive, mutually exclusive outcomes.
-What is meant is the cost function that arises by taking the dual to entropy of the maxent distribution with given mean parameter $\mu$.
-As we've discussed this seems like the right generalization of LMSR to arbitrary mean parameter spaces.
-
-Note that as $a \rightarrow 0$, we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$.
-As $a$ grows (agent grows more risk averse) the agent makes smaller and smaller trades that keep it closer to the current estimate $\theta$.
-
-So there is a kind of congruence between exponential family beliefs and exponential utility (as the names would suggest). 
-\subsection{Properties of the Maximum Entropy Market}
-Consider the dual of the cost function, $C^{*}(\mu)$ defined as
-$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
-This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
-$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
-Thus, the mean parameters are the dual variables to the natural parameters.
-
-\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
-\proof
-To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
-\begin{eqnarray*}
--H(p) &=& \int_{x}p(x)\log p(x)\\
-&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
-&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
-&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
-\end{eqnarray*}
-\unproof
-\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
-\rmk The LMSR is essentially a special case applied to a multinomial distribution.
-The LMSR is known to have bounded ($\log n$) market maker loss.
-Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
-\unrmk
-
-{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
-\proof
-Note that, for an exponential family distribution
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
-&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
-\end{eqnarray*}
-If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
-\unproof
-
-%{\cor The exponential family market maker has unbounded worst case loss.}
-%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
-%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
-%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
-\paragraph{Loss of the market maker}
-First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
-
-Recall the definition of the conjugate dual:
-$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
-The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
-$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
-This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
-So we have 
-$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
-Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
-
-Let $q_{0}$ be the initial and $q_{f}$ the final market state.
-Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
-\begin{eqnarray*}
-\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
-&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
-&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
-&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
-&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
-&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
-&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
-&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
-&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
-\end{eqnarray*}
-
-\subsubsection{Example: Gaussian Markets}
-{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
-In this case,  $\beta$ is a $2$-dimensional  vector.
-Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
-So
-\begin{eqnarray*}
-\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
-&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
-&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
-\end{eqnarray*}
-Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
-$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
-$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
-{\ex
-We will now derive the expression for the differential entropy for the normal distribution.
-Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
-&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
-\end{eqnarray*}
-Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
-Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
-
-{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
-Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
-Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
-Let's now work out the loss for the Gaussian market maker.
-First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
-\begin{eqnarray*}
-D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
-&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
-&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
-%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
-&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
-\end{eqnarray*}
-
-So the Gaussian market maker loss is 
-%parallels to quadratic market maker loss? 
-\begin{eqnarray*}
-D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
-&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
-&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
-\end{eqnarray*}¥
-Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
-\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
-
-\section{Connection to Mirror Descent?}
-
-\section{Conclusions}
-
-\end{document}
+\begin{itemize}
+\item Explain that the scoring rule from Example 2 does not elicit the mean when the agent's distribution is exponential (with change of variable so that it lies in $[1,+\infty)$. 
+\item Observe that scoring rule from Example does elicit the mean when the agent's distribution is exponential, Pareto, or in fact any other distribution with support on $[0,+\infty)$ such as the lognormal.
+\item This occurs because the scoring rule is linear in $x$, so its expectation is the same for any distribution with mean $\mu$. 
+\item Another feature to note is that under an appropriate parametrization (rate instead of mean), the scoring rule (1) is convex in the parameter being elicited. 
+\item These two observations taken together lay the groundwork for the characterization of strongly proper scoring rules for distribution statistics. 
+\item Before we proceed should note that the arguments behind Theorem~\ref{maxlike} apply not just to maximum likelihood but to any \emph{m-estimator}, which are notions from robust statistics.
+Explain what an m-estimator looks like and how there are general consistency results for those as well.
+\item The point of these alternative estimators is to be more robust to outliers in the sample, and so these kinds of estimators could translate into scoring rule more robust to inaccuracies in the agent's assessment of the underlying probability distribution or its parameters.
+(Just a hunch.)
+\end{itemize}

### 2014-01-31 WriteUp/Original/sindhu-outline.tex
tidied up main folder to just have submission doc

%%%
%%% Sindhu's Outline
%%%

\section{Sindhu's Outline}

In the notes on Maximum Entropy Market Maker, Lahaie constructs a scoring rule (and its sequential form: a market maker) based on maximum entropy distributions. Here is the problem: Suppose you want to elicit sufficient statistics from a trader (or group of traders: in this latter case you want to aggregate their information). Then if you find a probability distribution that is consistent with their belief of the expected sufficient statistics, you can construct a proper scoring rule (that is, the true distribution maximizes payoff). However, the solution to this consistent distribution is not unique. Hence, to pick amongst the distributions, you could choose the one that maximizes entropy (which is a notion of uncertainty) -- intuitively, this assumes the least about the distribution generating the data. Now, it turns out that the distribution that maximizes entropy is an exponential family distribution. So, if it is known that the true probability is an exponential family distribution, eliciting just the statistics allows you to characterize the actual probability distribution.

In the prediction markets and exponential families write-up, we have shown that that the various parts of the exponential family correspond to securities, payoffs and cost in a prediction market. We have shown that a sequence of traders seeking to maximize profit will essentially be performing a MLE of the natural parameters of an exponential family, when the beliefs of the traders take the form of an exponential family distribution (it turns out that this last assumption is not necesary in light of the above result). Further, in a budget-limited setting, adversarial traders can cause limited damage and informative traders' budgets grow in expectation over time when trading in multiple markets. The budget-limited results have the following issues. First, the movement of the share vector is not completely justified. Second, as it stands currently, the growth in budget is shown in expectation over a belief distribution of the informative trader, and this distribution is assumed to be an exponential family (this may be fixable).

So broadly, the outline is as follows: The main goal is to construct a prediction market that aggregates information from traders to determine the sufficient statistics of some unknown distribution. First, we can show that in a thin market with a single trader this can be done by constructing a (proper) scoring rule using a maximum entropy approach. This approach can then be extended to define a market scoring rule for a prediction market to aggregate information from multiple traders. This follows from the duality that exists between the maximum entropy approach and exponential families. We know that the updates to the mean parameters in a Bayesian setting are a convex combination of the empirical mean and the prior.  We can then show that the share vectors of this prediction market follow the Bayesian updates of the mean parameters when each trader treats the current market state as the prior and his beliefs as the empirical mean. Since we can interpret the share vector as the natural parameters of the exponential family, the updates essentially happen in a dual space.  This process of mapping to a dual space to perform an update is reminiscent of the online mirror descent algorithm; what exactly is the mapping? Also, we would like to analyze the prediction market so defined, when the traders are budget limited.  Is the step size of the algorithm somehow related to the budgets of the traders?

### 2014-02-04 WriteUp/maxentpredmkts.tex
Filled in first part of scoring rules section
Filled in the material for risk-neutral agents, need to do risk averse
still.

@@
 \newcommand{\muvec}{\pmb{\mu}}
+\newcommand{\st}{\hspace{10pt}\mathrm{s.t.}\hspace{10pt}}
+\newcommand{\norm}[1]{|| #1 ||}
+\newcommand{\tp}{\tilde{p}}
 
@@
 convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesian updates, conjugate prior.
-\section{Maximum Likelihood Scoring Rules}
-
-%%%
-%%%
-%%%
-
-\subsection{Preliminaries}
-
-We consider a state space $\mX$ which may be discrete or continuous, and a measure $\nu$ over sets of states.
-We will typically take the state space to be  $\bR$ or $\bR^d$ together with the Lebesgue measure (perhaps restricted to some subset of states), or some finite set together with the counting measure.\footnote{We will not need to make reference to the underlying $\sigma$-algebra, but to complete the specification of the measure space we take it to be the collection of Borel sets in the case of $\bR$ and $\bR^d$, and the power set in the case of a discrete set.} A \emph{probability density} is a $\nu$-measurable function $p : \mX \rightarrow \bR_+$ that integrates to 1 over $\mX$.
-
-Let $\mP$ denote the set of all probability densities over the given measure space.
-We are interested in eliciting information about some unknown density $p \in \mP$ which corresponds to the beliefs of an expert.
-The expert may be familiar with $p$ in its entirety, or only with certain properties of $p$.
-Given a set of densities $\mD \subseteq \mP$, a \emph{property}  is a real (possibly vector-valued) function $\Gamma : \mD \rightarrow \bR^k$.
-With a slight abuse of terminology we call elements of its image $\Theta = \Gamma(\mD)$ \emph{properties} and we say that $p$ has property $\theta$ when $\Gamma(p) = \theta$.
-If $\Gamma$ is one-to-one, we instead call $\Gamma$ a \emph{parametrization} and refer to the elements of its image $\Theta$ as \emph{parameters}.
-
-To be concrete, our outcome space could for instance be $\bR$ with $\nu$ the Lebesgue measure restricted to $[0,+\infty)$.
-The set $\mP$ then consists of all probability densities with support contained in the non-negative reals, and $\mD \subset \mP$ might consist of the densities of exponential distributions.
-An example of a property over $\mP$ is the mapping from a density to its mean.\footnote{The property may not be defined for all $p \in \mP$.
-In this case we implicitly restrict it to the subset of $\mP$ for which it is defined whenever its domain $\mD$ is not made explicit.} When restricted to the exponential densities $\mD$, this mapping is a parametrization.
-In much of this work, as in this example, the purpose of the measure $\nu$ is to implicitly encode bounds on the support of possible densities.
-The reason we use this approach, rather than explicitly stating bounds, is that it allows for clearer connections with maximum entropy estimation later on.
-
-Given the set of relevant densities $\mD \subseteq \mP$, which contains the expert's belief $p$, and a property $\Gamma$ over this set, the expert is asked to report $\theta = \Gamma(p)$.
-The expert is rewarded according to a \emph{scoring rule} $S : \Theta \times \mX \rightarrow \bR \cup \{-\infty\}$ which pays it $S(\htheta,x)$ according to how well its report $\htheta \in \Theta$ agrees with the eventual outcome $x \in \mX$.
-The idea is to design the scoring rule so that the expert is incentivized to report its true belief, setting $\htheta = \theta$.
+
+\pagebreak
+
+\section{Generalized Log Scoring Rules}
+
+We consider a measurable space that consists of a set of outcomes $\mX$ together with a $\sigma$-algebra $\mF$.
+An agent or expert has a \emph{belief} over potential outcomes taking the form of a probability measure absolutely continuous with respect to a base measure~$\nu$.\footnote{Recall that a measure $P$ is absolutely continuous with respect to~$\nu$ if $P(A) = 0$ for every $A \in \mF$ for which $\nu(A) = 0$.
+In essence the base measure~$\nu$ restricts the support of $P$.
+In our examples $\nu$ will typically be a restriction of the Lebesgue measure for continuous outcomes or the counting measure for discrete outcomes.} Throughout we represent the belief as the corresponding density $p$ with respect to~$\nu$.
+Let $\mP$ denote the set of all such probability densities, and let $\mD \subseteq \mP$ be a subset from which the beliefs are drawn.
+
+We are interested in eliciting information about the agent's belief, in particular expectation information.
+Let $\phi: \mX \rightarrow \bR^d$ be a vector-valued random variable or \emph{statistic}, where $d$ is finite.
+The aim is to elicit $\mu = \E_{p}[\phi(x)]$ where $x$ is the random outcome.
+A \emph{scoring rule} is a device for this purpose.
+Let $$\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}$$ be the set of realizable statistic expectations.
+A scoring rule $S : \mM \times \mX \rightarrow \bR \cup \{-\infty\}$ pays the agent $S(\hmu,x)$ according to how well its report $\hmu \in \mM$ agrees with the eventual outcome $x \in \mX$.
 %
 \begin{definition}
-Given a property $\Gamma : \mD \rightarrow \bR^k$, a scoring rule $S$ is  \emph{proper} over domain $\mD$ if for each $\theta \in \Theta$ and $p \in \mD$ with property $\theta$, we have
+A scoring rule $S$ is  \emph{proper over domain $\mD$} for statistic $\phi$ if for each $\mu \in \mM$ and $p \in \mD$ with expected statistic $\mu$, we have
 %
 \begin{equation} \label{eq-proper}
-\Exp_p[S(\theta,x)] > \Exp_p[S(\htheta,x)]
-\end{equation}
-%
-for all $\htheta \neq \theta$. If the domain is $\mD = \mP$, the set of all possible densities, then we say the scoring rule is \emph{strongly proper}.
+\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)]
+\end{equation}
+%
+for all alternative $\hmu$. If the domain is $\mD = \mP$, the set of all possible densities, then we simply say the scoring rule is \emph{proper}.
 \end{definition}
@@
 
-\begin{itemize}
-\item Note that the inequality is strict, in the literature these rules are called \emph{strictly} proper, but we only consider strictly proper rules here so don't make this qualification.
-\item Lambert et al.'s rules are actually strongly proper, however we consider scoring rules over vector valued properties.
-Lambert et al only develop these by adding together different scoring rules for each property.
-We go further in our constructions and develop rules that are not necessarily separable.
-\item It is useful to consider rules that are not strongly proper because we might have constructions for these for properties where no strongly proper rule is known.
-\item It is important to understand the informational requirements placed on the expert.
-Under a proper scoring rule, the expert does not need to be aware of the entire density $p$ to be incentivized to report its property $\theta$, since $\theta$ is an optimal report no matter what the density might be.
-The agent only needs to agree that the density indeed lies in $\mD$.
-\item For instance, if we are considering densities with support on $[0,+\infty)$, and $\mD$ is the set of lognormal densities, then a proper scoring rule over $\mD$ for the mean does not require the agent to know anything beyond the mean (e.g., it does not need to know the variance).
-\item A strongly proper scoring rule goes further; with such a rule we do not require the expert to believe the density is drawn from some subset $\mD$, only that it agrees with the support implied by the base measure $\nu$.
-Therefore, a strongly proper scoring rule for the mean here would elicit the mean no matter what kind of density over the positive reals might apply (e.g., exponential, Pareto, lognormal).
-\item In practical scenarios this seems like an important extension, because while it might be clear which states are possible, it might be difficult to ensure the expert agrees that the probability density for states falls within some restricted family $\mD$.
-\end{itemize}
-
-%%%
-%%%
-%%%
-
-\subsection{Maximum Likelihood}
-
-%
-\begin{theorem} \label{maxlike}
-The logarithmic scoring rule defined by
-%
-\[ S(\theta, x) = \log p(x;\theta) \]
-%
-is weakly proper for $\mF$ over $\Theta$ if and only if the maximum likelihood estimate for $\theta$ is consistent.
-Furthermore, if $\mM$ is an alternative parameter space such that there exists a bijection $t : \mM \rightarrow \Theta$, then $\bar{S}(\mu, x) = S(t(\mu), x)$ is weakly proper for $\mF$ over $\mM$.
+\noindent
+Note that given a proper scoring rule $S$ any affine transformation $\tilde{S}(\mu,x) = aS(\mu,x) + b(x)$ of the rule, with $a > 0$ and $b$ an arbitrary real-valued function of the outcomes, again yields a proper scoring rule termed \emph{equivalent}.
+Throughout we will often apply such affine transformations to obtain the clearest version of the scoring rule.
+We will also focus on scoring rules where inequality~(\ref{eq-proper}) is strict to avoid trivial cases such as constant scoring rules.
+
+Our definition is more general than classical scoring rules in two respects.
+Classically, scoring rules take in the entire density $p$ rather than just some statistic, and incentive compatibility must hold over all of $\mP$.
+When the outcome space is large or infinite, it is not feasible to directly communicate $p$, so our definition allows for summary information of the belief.
+If there is a bijection between the sets $\mD$ and $\mM$, then we say that $\mM$ parametrizes $\mD$ and write $p(\cdot\,;\mu)$ for the density mapping to $\mu$.
+
+Note that this definition of a scoring rule places only mild information requirements on the part of the agent to ensure truthful reporting.
+Because condition~(\ref{eq-proper}) holds for all $p$ consistent with expectation $\mu$, it is enough for the agent to simply know the latter and not the complete density to be properly incentivized.
+However, the agent must also agree that the density is actually drawn from domain $\mD$; if the scoring rule is proper, it is enough that the agent agree with the support of the density as implicitly defined by base measure $\nu$.
+
+When the outcome space is finite we can recover classical scoring rules from the definition by taking $\mD = \mP$ and using the statistic $\phi : \mX \rightarrow \{0,1\}^{\mX}$ that maps an outcome $x$ to a unit vector with a 1 in the component corresponding to $x$.
+The expectation of $\phi$ is then exactly the probability mass function.
+
+\subsection{Proper Scoring from Consistency}
+
+Our starting point for designing proper scoring rules is the classic logarithmic scoring rule for eliciting probabilities in the case of finite outcomes.
+This rule is simply $S(p,x) = \log p(x)$ and there are several ways to understand why it is proper, with different implications for generalization.
+
+First observe that the scoring rule compensates the agent with the log likelihood it assigned to the outcome, so the agent is maximizing the expected log likelihood via its report.
+Now it is a fundamental result in statistics that, under a variety of different sufficient conditions, the maximum likelihood estimator is statistically consistent; see~\cite{} for an overview of such conditions.
+This means that in the limit (as the size of the i.i.d.\ empirical sample increases) the log likelihood is maximized by the true parameters.
+By this reasoning, the following result follows immediately from the definition of consistency and the law of large numbers.
+%
+\begin{theorem} \label{max-like}
+Suppose that $\mM$ parametrizes $\mD$. The logarithmic scoring rule defined by
+%
+\begin{equation} \label{log-score}
+S(\mu, x) = \log p(x;\mu),
+\end{equation}
+%
+where $\mu \in \mM$, is proper over $\mD$ if and only if the maximum likelihood estimator for $\mu$ is consistent.
 \end{theorem}
 %
-
+We note that although our focus is on parametrizations by expectations, the preceding theorem in fact holds for more generic parameter spaces.
+The range of scoring rules it provides is broader than those characterized by~\cite{}, because it allows for a restriction over the set of beliefs.
+On the other hand, its applicability is limited to those domains $\mD$ that can be parametrized by the relevant statistics.
+The following provide concrete examples of these points.
 %
 \begin{example}
-Suppose $X$ is supported on $[0,+\infty)$ and follows an exponential distribution with density $f(x;\lambda) = \lambda e^{-\lambda x}$ parametrized by a rate $\lambda > 0$.
-The mean $\mu$ is related to the rate via the one-to-one mapping $\mu = 1/\lambda$, so the density can alternatively be parametrized by the mean.
-According to Theorem~\ref{maxlike} the following is a weakly proper scoring rule for the mean of an exponential distribution:
+Suppose the outcomes are supported on $[1,+\infty)$ and follow a Pareto distribution with density $f(x;\alpha) = \alpha/x^{\alpha+1}$ parametrized by an index $\alpha > 0$.
+The mean $\mu$ is related to the index via the one-to-one mapping $\mu = \frac{\alpha}{\alpha-1}$, so the density can alternatively be parametrized by the mean.
+Theorem~\ref{max-like} gives the following proper scoring rule for the mean of a Pareto distribution with support on $[1,+\infty)$:
+%
+\begin{equation} \label{pareto-scoring}
+S(\mu,x) = \log \frac{\mu}{\mu-1} - \left(\frac{\mu}{\mu-1} + 1\right) \log x.
+\end{equation}
+%
+We stress that the rule can only elicit the mean assuming the agent knows its belief is a Pareto distribution over $[1,+\infty]$.
+It does not elicit the mean of other families of densities parametrized by the mean (e.g., the exponential distribution).
+\end{example}
+%
+%
+\begin{example}
+Suppose the outcomes are supported on $(-\infty,+\infty)$ and follow a Cauchy distribution with density $f(x;m) = 1/\pi[1+(x-m)^2]$, parametrized by the median $m$.
+Theorem~\ref{max-like} leads to the following proper scoring rule for the median:
+%
+\begin{equation} \label{cauchy-scoring}
+S(m,x) = -\log[1+(x-m)^2].
+\end{equation}
+%
+Note that the median cannot be obtained as the expectation of any statistic. This highlights the range of parameters than may be elicited by the log score in different circumstances.
+\end{example}
+%
+
+\subsection{Proper Scoring from Maximum Entropy}
+
+We now turn to scoring rules that are proper over the entire domain of densities $\mP$.
+Our construction is the same as in the previous section: we take the log likelihood of a density parametrized by the relevant statistics.
+If the family of densities is chosen appropriately, the resulting rule will in fact hold over all $\mP$.
+To achieve this we draw on a well-known duality between maximum likelihood and maximum entropy.
+
+\subsubsection*{Exponential Families}
+
+We let $p(x;\mu)$ be the maximum entropy distribution with expected statistic $\mu$.
+Specifically, it is the solution to the following mathematical program:\footnote{We assume throughout that the minimum is finite and achieved for all $\mu \in \mM$.
+Some care is needed to ensure this holds for specific statistics and outcome spaces.
+For example, taking outcomes to be the real numbers, there is no maximum entropy distribution with a given mean $\mu$ (one can take densities tending towards the uniform distribution over the reals), but there is always a solution if we constrain both the mean and variance.}
+%
+\begin{equation} \label{maxent-prog}
+ \min_{p \in \mP} F(p) \st \Exp_{p}[\phi(x)] = \mu,
+\end{equation}
+%
+where the objective function is the negative entropy of the distribution, namely
+%
+\[ F(p) = \int_{x \in \mX} p(x) \log p(x)\, d\nu(x).
+\]
+%
+Note that the explicit set of constraints in~(\ref{maxent-prog}) are linear, and to stress this fact we find it helpful to re-write them as $A_{\phi} p = \mu$, where $A_{\phi}$ is the expectation operator of statistics $\phi$.
+We let $G : \mM \rightarrow \bR$ be the optimal value function of~(\ref{maxent-prog}), meaning $G(\mu)$ is the (negative) entropy of the maximum entropy distribution with expected statistics $\mu$.
+
+It is well-known that solutions to~(\ref{maxent-prog}) are \emph{exponential family} distributions, whose densities with respect to $\nu$ take the form
+%
+\begin{equation} \label{exp-fam}
+p(x;\theta) = \exp( \la \theta, \phi(x) \ra - T(\theta) ).
+\end{equation}
+%
+The density is stated here in terms of its \emph{natural} parametrization $\theta \in \bR^d$, where $\theta$ arises as the Lagrange multiplier associated with linear constraints $A_{\phi}p = \mu$.
+The term $T(\theta)$ essentially arises as the multiplier for the normalization constraint (the density must integrate to 1), and so ensures that~(\ref{exp-fam}) is normalized:
+%
+\begin{equation} \label{log-part}
+T(\theta) = \log \int_{\mX} \exp \la \theta, \phi(x) \ra \,d\nu(x).
+\end{equation}
+%
+The function $T$ is known as the \emph{log-partition} or \emph{cumulant} function corresponding to the exponential family.
+Its domain is $\Theta = \{ \theta \in \bR^d : T(\theta) < +\infty \}$, called the natural parameter space.
+The exponential family is \emph{regular} if $\Theta$ is open---almost all exponential families of interest, and all those we consider in this work, are regular.
+The family is \emph{minimal} if there is no $\alpha \in \Theta$ such that $\la \alpha, \phi(x) \ra$ is a constant over $\mX$ ($\nu$-almost everywhere); minimality is a property of the associated statistic $\phi$, usually called the \emph{sufficient statistic} in the literature.
+
+The following proposition collects the relevant results on regular exponential families and their associated parameter domains and functions.
+A convex function $T$ is of \emph{Legendre type} if it is proper, closed, strictly convex and differentiable on the interior of its domain, and $\lim_{\theta \rightarrow \bar{\theta}} \norm{\grad T(\theta)} = +\infty$ when $\bar{\theta}$ lies on the boundary of the domain.
+%
+\begin{prop}
+Consider a regular exponential family with minimal sufficient statistic. The following properties hold:
+\begin{enumerate}
+\item $T$ and $G$ are of Legendre type, and $T = G^*$ (equivalently $G = T^*$). 
+\item The gradient map $\grad T$ is one-to-one and onto the interior of $\mM$. Its inverse is $\grad G$ which is one-to-one and onto the interior of $\Theta$.
+\item The exponential family distribution with natural parameter $\theta \in \Theta$ has expected statistic $\mu = \Exp_p[\phi(x)] = \grad T(\theta)$.
+\item The maximum entropy distribution with expected statistic $\mu$ is the exponential family distribution with natural parameter $\theta = \grad G(\mu)$. 
+\end{enumerate}
+\end{prop}
+%
+In the above $T^*$ denotes as usual the convex conjugate of $T$, which here can be evaluated as $T^*(\mu) = \sup_{\theta \in \Theta} \la \theta, \mu \ra - T(\theta)$.
+Similarly, $G^*(\theta) = \sup_{\mu \in \mM} \la \theta, \mu \ra - G(\mu)$.
+
+\subsubsection*{Proper Log Scoring}
+
+We are now in a position to analyze the log scoring rule under exponential family distributions.
+From our discussion so far, we have that an exponential family density can be parametrized either by the natural parameter $\theta$, or by the mean parameter $\mu$, and that the two are related by the invertible gradient map $\mu = \grad T(\theta)$.
+We will write $p(x;\theta)$ or $p(x;\mu)$ given the parametrization used, which should be clear from context.
+
+The following observation is crucial.
+Let $\tp \in \mP$ be a density (not necessarily an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\mu' \in \mM$ be an alternative report.
+Then note that
+%
+\begin{equation} \label{equalizer-rule}
+\Exp_{\tp}[\log p(x;\mu')] = \Exp_{p(\cdot;\mu)}[\log p(x;\mu')] = \la \theta', \mu \ra - T(\theta'),
+\end{equation}
+%
+where $\theta' = \grad G(\mu')$ is the natural parameter for the exponential family with statistic $\mu'$.
+We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, so when reasoning about the agent's reporting incentives we can just as well assume its beliefs are from an exponential family for simplicity.
+%
+\begin{theorem} \label{maxent-score}
+Consider the logarithmic scoring rule $S(\mu,x) = \log p(x;\mu)$ defined over a set of densities $\mD$ parametrized by $\mM$.
+The scoring rule is proper over the entire domain of densities $\mP$ if and only if $\mD$ is the exponential family with statistic $\phi$.
+\end{theorem}
+%
+%
+\begin{proof}
+Let $\mu,\mu' \in \mM$ be the agent's true belief and an alternative report, and let $p,p' \in \mP$ be densities consistent with each belief.
+Let $\theta = \grad G(\mu)$ and $\theta' = \grad G(\mu')$, and note that $\mu = \grad T(\theta)$.
+We have
+%
+\begin{eqnarray*}
+& & \Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\mu')] \\
+& = & \la \theta, \mu \ra - T(\theta) - \la \theta', \mu \ra + T(\theta') \\
+& = & T(\theta') - T(\theta) - \la \theta' - \theta, \mu \ra \\
+& = & T(\theta') - T(\theta) - \la \theta' - \theta, \grad T(\theta) \ra.
+\end{eqnarray*}
+%
+The latter is positive by the strict convexity of $T$, which shows that the log score is proper.
+Now assume the defined log score is proper.
+By the Savage characterization of proper scoring rules for expectations (see~\cite{}), we must have
+%
+$$
+S(\mu,x) = G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra
+$$
+for some strictly convex function $G$. Let $T = G^*$, so that $\grad G = {\grad T}^{-1}$, and let $\theta = \grad G(\mu)$. Then the above takes the form
+%
+\begin{eqnarray*}
+\log p(x;\mu) & = & G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra \\
+& = & \la \theta, \mu \ra - T(\theta) - \la \theta, \mu - \phi(x) \ra \\
+& = & \la \theta, \phi(x) \ra - T(\theta),
+\end{eqnarray*}
+%
+which shows that $p(x;\mu)$ takes the form of an exponential family.
+\end{proof}
+%
+This theorem leads to a straightforward procedure for constructing score rules for expectations: define the relevant statistic, and consider the maximum entropy (equivalently, exponential family) distribution consistent with the agent's reported mean $\mu$.
+The agent is compensated according to the log likelihood of the eventual outcome according to the latter.
+The interpretation is that the agent is only providing partial information about the underlying density, so the principal first infers a full density according to the principle of maximum entropy, and then scores the agent using the usual log score.
+
+An advantage of this generalization of the log score is that, for many statistics of interest, it leads to simple \emph{closed-form} formulas for the scoring rule.
+The following examples illustrate the construction.
+%
+\begin{example}
+As base measure we take the Lebesgue restricted to $[0,+\infty)$, and we consider the statistic $\phi(x) = x$ so that we are simply eliciting the mean.
+The maximum entropy distribution with a given mean $\mu$ is the exponential distribution, and taking its log density gives the scoring rule
 %
@@
 %
-To verify this, suppose the agent believes $X$ follows an exponential distribution with rate $\lambda_0$, equivalently mean $\mu_0 = 1/\lambda_0$.
-By the one-to-one relation between rate and mean, choosing $\mu$ to maximize the expected score~(\ref{exp-scoring}) is equivalent to choosing $\lambda$ to maximize $\Exp{-\lambda x + \log \lambda} = -\lambda \mu_0 + \log \lambda$.
-The latter is strictly concave in $\lambda$, and using basic calculus its unique global maximum is $\lambda^* = 1/\mu_0$, which therefore leads the agent to report $\mu = \mu_0 = 1/\lambda^*$ under scoring rule~(\ref{exp-scoring}).
+We stress that although this rule is derived from the exponential distribution, Theorem~\ref{maxent-score} implies that it elicits the mean of any distribution supported on the non-negative reals (e.g., Pareto, lognormal), in contrast to~(\ref{pareto-scoring}).
+Indeed, it is easy to see that the expected score~(\ref{exp-scoring}) depends only on the mean of the agent's belief because it is linear in $x$.
+%
+Typically the exponential distribution is parametrized by its rate $\lambda = 1/\mu$.
+In that case the rule takes the form $S(\mu, x) = -\lambda x + \log \lambda$, and we see that $\lambda$ corresponds to the natural parameter.
 \end{example}
 %
-
 %
 \begin{example}
-Suppose that $X$ is supported on $[1,+\infty)$ and follows a Pareto distribution with density $f(x;\alpha) = \alpha/x^{\alpha+1}$ parametrized by an index $\alpha > 0$.
-The mean $\mu$ is related to the index via the one-to-one mapping $\mu = \frac{\alpha}{\alpha-1}$, so the density can alternatively be parametrized by the mean.
-Theorem~\ref{maxlike} gives the following weakly proper scoring rule for the mean of a Pareto distribution with support on $[1,+\infty)$:
-%
-\begin{equation} \label{pareto-scoring}
-S(\mu,x) = \log \frac{\mu}{\mu-1} - \left(\frac{\mu}{\mu-1} + 1\right) \log x.
-\end{equation}
-%
-To verify this, suppose the agent believes $X$ follows a Pareto distribution with index $\alpha_0$, equivalently mean $\mu_0 = \frac{\alpha_0}{\alpha_0-1}$.
-By the one-to-one relation between the index and mean, choosing $\mu$ to maximize the expected score~(\ref{pareto-scoring}) is equivalent to choosing $\alpha$ to maximize $\Exp{\log \alpha - (\alpha+1)\log x} = \log \alpha - (\alpha+1)\Exp{\log x}$.
-The latter is strictly concave in $\alpha$, and using basic calculus its unique global maximum is $\alpha^* = 1/\Exp{\log X} = \alpha_0$, where the last equality follows from the fact that $\log X$ has an exponential distribution with rate $\alpha$ when $X$ is Pareto with rate $\alpha$.
-Therefore~(\ref{pareto-scoring}) leads the agent to report $\mu = \mu_0 = \frac{\alpha^*}{\alpha^* - 1}$.
+As a base measure we take the Lebesgue over the real numbers $\bR$.
+We are interested in eliciting the mean $\mu$ and variance $\sigma^2$, so as a statistic we take $\phi(x) = (x, x^2)$ for which $\Exp_p[\phi(x)] = (\mu, \mu^2+\sigma^2)$.
+The maximum entropy distribution for a given mean and variance is the normal distribution, and taking its log density gives the scoring rule
+%
+\begin{equation}
+S((\mu, \sigma^2), x) = -\frac{(x-\mu)^2}{\sigma^2} - \log \sigma^2.
+\end{equation}
+%
+Again, we stress that this scoring rule elicits the mean and variance of any density over the real numbers, not just those of a normal distribution.
+The construction easily generalizes to a multi-dimensional outcome space by taking the log density of the multivariate normal:
+%
+\begin{equation}
+S(({\boldsymbol \mu}, {\boldsymbol \Sigma}), \boldsymbol{x}) = -( \boldsymbol{x} - {\boldsymbol \mu})'{\boldsymbol \Sigma}^{-1}( \boldsymbol{x} - {\boldsymbol \mu}) - \log |{\boldsymbol \Sigma}|.
+\end{equation}
+%
+Here the statistics being elicited are the mean vector ${\boldsymbol \mu}$ and the covariance matrix ${\boldsymbol \Sigma}$.
+These scoring rules have been studied by Dawid and Sebastiani~\cite{} as rules that only depend on the mean and variance of the reported density.
+They note that these rules are weakly proper (because they do not distinguish between densities with the same first and second moments), but do not make the point that knowledge of the full density is not necessary on the part of the agent.
 \end{example}
@@
 
-\begin{itemize}
-\item Explain that the scoring rule from Example 2 does not elicit the mean when the agent's distribution is exponential (with change of variable so that it lies in $[1,+\infty)$. 
-\item Observe that scoring rule from Example does elicit the mean when the agent's distribution is exponential, Pareto, or in fact any other distribution with support on $[0,+\infty)$ such as the lognormal.
-\item This occurs because the scoring rule is linear in $x$, so its expectation is the same for any distribution with mean $\mu$. 
-\item Another feature to note is that under an appropriate parametrization (rate instead of mean), the scoring rule (1) is convex in the parameter being elicited. 
-\item These two observations taken together lay the groundwork for the characterization of strongly proper scoring rules for distribution statistics. 
-\item Before we proceed should note that the arguments behind Theorem~\ref{maxlike} apply not just to maximum likelihood but to any \emph{m-estimator}, which are notions from robust statistics.
-Explain what an m-estimator looks like and how there are general consistency results for those as well.
-\item The point of these alternative estimators is to be more robust to outliers in the sample, and so these kinds of estimators could translate into scoring rule more robust to inaccuracies in the agent's assessment of the underlying probability distribution or its parameters.
-(Just a hunch.)
-\end{itemize}
+\pagebreak
 \section{Maximum Entropy Market Making}
@@
 Using a maximum entropy approach, we explain how a market scoring rule can be developed for outcome spaces both discrete and continuous, and for generic properties of the underlying distribution (e.g., mean and variance), rather than just the probabilities of individual outcomes.
-
-%%%
-%%%
-%%%
-
-\subsection{The Model}
-\label{model}
-
-Let $\mX$ be the outcome space, which may be discrete or continuous.
-Let $\mP$ denote the set of probability distributions over the outcome space.
-We represent a probability distribution as a density $p$ absolutely continuous with respect to some base measure $\nu$ (e.g., counting for discrete outcomes, or Lebesgue for continuous outcomes).% We will discuss the interpretation of the base measure in a later section.
-
-There is an unknown distribution $p$ over outcomes.
-We are interested in estimating the expected value of different outcome \emph{statistics} under this distribution by aggregating the beliefs of agents.
-A statistic is a real-valued function $\phi_s : \mX \rightarrow \bR$ over outcomes, independent of $p$; here $s$ belongs to a finite index set $\mS$ of size $d$.
-The collection of functions $\phi = (\phi_s,\, s \in \mS)$ can be viewed as a vector-valued statistic mapping outcomes into $\bR^d$.
-Formally, the aim is to estimate
-%
-\begin{equation} \label{constraint}
-\Exp_p[\phi(x)] = \mu.
-\end{equation}
-%
-We stress that we are only seek to elicit $\mu \in \bR^d$, not the full distribution $p$.
-As an example, suppose that $\mX = \bR$.
-If we were interested in eliciting the first two uncentered moments of $p$, we would include $\phi_1(x) = x$ and $\phi_2(x) = x^2$ as statistics.
-Note that the variance cannot be directly obtained through any single statistic: it corresponds to the expectation of $(x - \Exp_p[x])$, which depends on $p$, violating our definition of a statistic.
-Of course, the variance can be indirectly obtained by eliciting the first two moments separately.\footnote{This is related to the notion of \emph{elicitation complexity} of statistical properties introduced by Lambert et al.~\cite{Lambert}, but we do not pursue this connection here.}
-
-There are two approaches to eliciting an agent's estimated $\mu$ in~(\ref{constraint}).
-The first approach is to incentivize the agent to directly report $\mu$ via a \emph{scoring rule}.
-The second approach is to form an \emph{information market} by issuing securities for each statistic $s \in \mS$, with payoff depending on the realized value of the statistic.
-(Such securities are known as contingent claims.) The amount of shares of each security acquired by the agent indirectly reveals it estimate $\mu$.
-In a market scoring rule~\cite{Hanson}, these two approaches are dual to each other in a formal sense.
-
-%%%
-%%%
-%%%
-
-\subsection{Scoring Rule}
-
-We first consider how to directly elicit $\mu$ via a scoring rule. 
-%Note that the usual definition of a scoring rule presumes that a full probability distribution is elicited, rather than just some of its properties (e.g., mean and variance), so below we will adapt the definition to our setting.
-The space of feasible reports from an agent corresponds to
-%
-\begin{equation} \label{marginal}
-\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}.
-\end{equation}
-%
-A scoring rule is a function $S : \mM \times \mX \rightarrow \bR$ that rewards an agent with $S(\mu,x)$ based on how its report $\mu$ agreed with the eventual outcome $x$.
-A scoring rule is \emph{proper} if for all $\mu \in \mM$, and $p \in \mP$ such that $\Exp_p[\phi(x)] = \mu$, we have
-%
-\begin{equation} \label{scoring}
-\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)]
-\end{equation}
-%
-for all $\hmu \neq \mu$; the rule is \emph{strictly proper} if the inequality is strict.
-Note that in the literature, scoring rules are defined over entire probability distributions, not just properties of those distributions.
-Our definition imposes a minimum of informational requirements on agents, because an agent does not need to assess the full distribution $p$ that might have lead to its estimates $\mu$ in order to realize that truthful reporting is an optimal strategy.
-
-Our scoring rule is implicitly defined via the solution of a mathematical program. The program computes the maximum entropy distribution consistent with the agent's reported beliefs $\mu$.
-%
-\begin{eqnarray}
-\max_{p \geq 0} & \ds - \int_{x \in \mX} p(x) \log p(x) \nu(dx) & \label{obj} \\
-\mbox{s.t.} & \ds \int_{x \in \mX} \phi(x) p(x) \nu(dx) = \mu & \label{mean} \\
- & \ds \int_{x \in \mX} p(x) \nu(dx) = 1 & \label{normalize}
-\end{eqnarray}
-%
-The non-negativity constraints together with~(\ref{normalize}) ensure that the solution is a probability distribution.
-Constraints~(\ref{mean}), which correspond to $d$ separate constraints, ensure consistency with the agent's report.
-The objective~(\ref{obj}) is the entropy of the solution $p$ with respect to the measure $\nu$.
-The program has a convex objective and linear constraints.
-%For continuous outcome spaces we will see in the next section that the optimal solution can still be cleanly characterized.
-%
-\begin{theorem} \label{thm:scoring}
-Let $p(x;\mu)$ be the optimal solution to the maximum entropy program, given report $\mu \in \bR^d$. The scoring rule defined by
-%
-\begin{equation} \label{scoring}
-S(\mu,x) = a_x + b\log p(x;\mu),
-\end{equation}
-%
-where $b,a_x \in \bR$ and $b > 0$, is strictly proper.
-\end{theorem}
-%
-Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$
-
-Note that the maximum entropy distribution is an exponential family distribution. 
- Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.
-
- We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
-  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
-   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
-   Or 
-\begin{equation}\label{eq:proper}
-\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
-\end{equation}
-
-   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
-   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
-   Rearranging the terms,
-   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
-   which is the first order condition on the convexity of $\psi(\cdot)$.
- \unproof
-
- \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters.
-That is,
-   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
-   \unrmk
-
-\noindent
-This rule represents a generalization of the logarithmic scoring rule for probability distributions to any properties of distributions that can be captured as expectations of statistics.
-As explained in the next section, the solution to the maximum entropy program takes the form of an \emph{exponential family} distribution.
-For many common properties of interest, these distributions are familiar; for example, the maximum entropy distribution with a given mean is an exponential distribution, and the maximum entropy distribution with a given mean and variance is a normal distribution~\cite{}.
-We provide here two examples in more depth.
-
-%%
-%%
-
-\paragraph{Multinomial Distribution}
-
-As a first example, consider the problem of estimating the individual probabilities of a finite set of outcomes.
-We have $\mX = \{1,\ldots,k\}$.
-The relevant statistics indicate which outcome actually occurs, so the index set is $\mS = \mX$.
-Statistic $\phi_x(x')$ is 1 if $x' = x$ and 0 otherwise.
-Observe that if $p$ is the distribution over outcomes, then $\Exp_p[\phi(x)] = p$.
-A feasible report from an agent is any non-negative $\pi = (\pi_1,\ldots,\pi_k)$ such that $\sum_{i=1}^k \pi_i = 1$.
-Given a report of $\mu = \pi$, the unique solution to the maximum entropy program is $p(x; \pi) = \pi$, and~(\ref{scoring}) corresponds to the classic logarithmic scoring rule of $S(\pi,x) = \log \pi_x$.
-\hfill{$\Box$}
-
-\paragraph{Normal Distribution}
-
-As a second example, suppose the outcome space is $\mX = \bR$ and that we are interested in estimating the mean $\mu$ and variance $\sigma^2$ of the underlying distribution.
-We choose the first and second (uncentered) moments, $\kappa_1 = \mu$ and  $\kappa_2 = \mu^2 + \sigma^2$ as our statistics.
-It is well-known that the maximum entropy distribution with a given mean and variance is the normal distribution, so the solution to the maximum entropy program (using Lebesgue as the base measure $\nu$) is
-%
-\[ p(x; \kappa) = \frac{1}{\sqrt{2\pi\sigma^2}} \exp\left[ -\frac{(x-\mu)^2}{2\sigma^2} \right]. \]
-%
-Applying Theorem~\ref{thm:scoring} and choosing appropriate constants, we find that the following scoring rule is strictly proper in this context. %
-\[ S(\kappa, x) = -\frac{(x-\mu)^2}{\sigma^2} - \log \sigma^2. \]
-%
-Note that the fact that this scoring rule is proper for the mean amounts to the well-known fact that reporting the mean is a Bayes act under quadratic loss~\cite{someone}. \hfill{$\Box$}
-
-\medskip\noindent
-A brief observation on parametrizations is in order.
-Note that the maximum entropy distribution here is parametrized by the uncentered moments $\kappa = (\kappa_1,\kappa_2)$ rather than $(\mu,\sigma^2)$ because the variance is not a valid statistic.
-However, it is clear that the agent does not have to directly report $\kappa$: the market maker itself can do the translation from reported mean and variance to uncentered moments.
-This is relevant because the $\kappa$ parametrization may be unintuitive for an agent, especially given the constraints $\kappa_2 - \kappa_1^2 \geq 0$ that must be enforced.
-In the $(\mu, \sigma^2)$ parametrization each parameter is unrestricted, and the parameters have intuitive interpretations.
-
-Our approach so far has been to identify relevant properties of the unknown distribution, and elicit them via a scoring rule computed through a maximum entropy program.
-However, it is the case that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}.
-Therefore, we see the following alternative approach.
-The market designer can begin by considering the outcome space $\mX$, and assessing what family of distributions over $\mX$ the unknown $p$ might come from.
-If this family is an exponential family, then Theorem~\ref{thm:scoring} immediately gives a proper scoring rule to elicit the parameters of the distribution.
-Many of the most familiar distributions are exponential families, including the normal, exponential, gamma, beta, binomial, Poisson, Weibull, and Dirichlet.
-To illustrate, we consider two examples in more depth.
-
-%%
-%%
-
-\paragraph{Exponential}
-Suppose the designer would like an estimate of the rate at which servers fail in a data center.
-An outcome is the time between failures, so $\mX = [0,+\infty)$.
-The usual distribution for failure times is the exponential distribution $p(x;\lambda) = \lambda e^{-\lambda x}$, which is parametrized by the rate $\lambda$ and has mean $\mu = 1/\lambda$.
-The exponential distribution is the maximum entropy distribution given a fixed mean $\mu$.
-From Theorem~\ref{thm:scoring} we obtain the following proper scoring rule for this setting:
-%
-\[ S(\mu, x) = \log \lambda - \lambda x. \]
-%
-It is simple to check using basic calculus that an agent maximizes the expected score by reporting $\lambda = 1/\hmu$, where here $\hmu$ is its estimate of the mean.
-In this case, it is unclear a priori which of the mean or rate parametrizations is most intuitive, but the maximum entropy approach easily allows for either.
-\hfill{$\Box$}
-
-%%
-%%
-
-\paragraph{Beta}
-Suppose the designer would like an estimate of the click-through rate of an online advertisement.
-The outcome space here is $\mX = (0,1)$.
-A typical prior over a probability like this is the beta distribution, which has two parameters $\alpha, \beta > 0$ and density function $p(x; \alpha, \beta) = x^{\alpha-1}(1-x)^{\beta-1}/B(\alpha,\beta)$, where $B$ is the beta function.
-By theorem~\ref{thm:scoring} this leads to the proper scoring rule
-%
-\[ S((\alpha, \beta), x) = (\alpha-1)\log x + (\beta-1)\log (1-x) - \log B(\alpha,\beta). \]
-%
-Although this parametrization is standard for the beta, it is perhaps more intuitive to use the parametrization $\mu = \alpha/(\alpha+\beta) \in (0,1)$ and $n = \alpha+\beta > 0$.
-Here $\mu$ is the mean of the distribution---corresponding to the agent's actual estimate of the click-through rate---and $n$ is a measure of the agent's confidence in the estimate.
-\hfill{$\Box$}
-
-%%%
-%%% 
-%%%
 

### 2014-02-04 WriteUp/maxentpredmkts.bak.tex
moved my changes to backup file

--- 2014-02-04 WriteUp/maxentpredmkts.tex
@@
 %	- generalized lmsr
-%	- does modularity (a la Hanson) still hold?
+%	[- does modularity (a la Hanson) still hold?]
 %- prediction market setup
@@
 %- budgets and adversarial markets
+% any liquidity results?
 
@@
 
-\section{The Exponential Family Market Mechanism: A Maximum Likelihood Approach}
-In this section, we will set up a prediction market that aggregates beliefs from traders where the outcome is a continuous random variable.
-In particular, we assume that the outcome is drawn from an exponential family distribution.
-Each of the traders has access to a series of points drawn from this distribution.
-In other words, the traders have access an empirical mean of the sufficient statistics of the exponential family.
-Every trader has infinite budget so that the current market price after a trader has traded in the prediction market, exactly reflects his beliefs.
-
-%We will now provide a more formal description of the problem statement.
-%\subsection{Problem Statement}
-%\begin{description}
-%\item[Setup: ] Learning proceeds in rounds $j=1,2,\ldots$ where at the end of each round the algorithm may be provided with $x_{j}$  drawn from a set $\mathcal{X}$ according to some distribution $\mathcal{D}_{j}$, where $\mathcal{D}_{j}$ belongs to an exponential family defined by $k$ sufficient statistics.
-In each round, some sequence of $n$ traders provide their estimates of the empirical means of each of the $k$ sufficient statistics drawn from $\mathcal{D}_{j}$.
-%\item[Define: ] A learning algorithm that learns this distribution by providing a maximum likelihood estimate of its unknown natural parameters based on the empirical means.  
-%\item[Proposed solution: ] Simulate a prediction market defined by a cost function $C$, securities $s_{0},s_{1},s_{2},\ldots, s_k$ and their payoffs $\phi_i(x)$ for $i=0,1,2,\ldots,k$ for the LMSR.
-Define a correspondence between the state of the prediction market and the output of the learning algorithm.
-%\end{description}
-
-For $x_{i}\sim P_{\betavec}$, recall that the likelihood function for independently drawn data $x_1,\ldots,x_n$ is given by $\prod_{i=1}^{n} P_{\betavec}(x_i)$.
-The maximum likelihood estimate of the natural parameters $\betavec$ is the value of the natural parameters that maximizes the likelihood function.
-
-We will now set up a prediction market with log market scoring rule (LMSR) and infinite budget traders such that the market state represents the maximum likelihood estimate (MLE) of the natural parameters of an exponential family distribution.
-%Given  under some interpretation of what information the participating agents have, and how they bet.
-
-For a given exponential family distribution, the prediction market is defined as follows:
-\begin{description}
-\item[Traders] The prediction market will simulate a trader $i$ corresponding to expert $i$.
-This trader processes all information samples available to her directly or inferred from previous trades, and  forms a belief distribution such that the believed means of the sufficient statistics matches the empirical means of the sufficient statistics from the information samples.
-The trader trades in the market to maximize her expected payoff under her believed distribution.
-\item[Securities and their payoff]
-For each $i=1,2,\ldots,k$, we define a security $s_{i}$ with payoff $\phi_{i}(x)$ where $x$ is the ultimate outcome and $\Phi()$ defines the vector of sufficient statistics of the exponential family distribution according to which $x$ is drawn.
-We define an additional security $s_{0}$ with payoff $\phi_{0}(x):=a-\sum_{i=1}^{k}\phi_{i}(x)$ where $a$ is an appropriately chosen constant dependent on the range of $\Phi$ so that the payoff of $s_{0}$ is non-negative.
-\end{description}
-
-We also note that since we want non-negative payoffs, we restrict the exponential families under consideration so that the sufficient statistics are lower bounded.
-If we have a constant lower bound on the sufficient statistics, then without loss of generality we can add a constant to each sufficient statistic without changing the exponential family in any way, as the constants will be absorbed in the log-partition function.
-
-\paragraph{Arbitrage-free Property}
-We can easily show that the prediction market we have defined is arbitrage-free: there is no sequence of trades that guarantees the trader a profit under all conditions.
-First, note that the incremental cost and payoff function are additive over multiple trades, so the net profit of a sequence of trades depends only on the initial and final market position, and is independent of the actual path along which the trade takes place.
-
-Now, consider any trade (or sequence of trades) that moves the market from an initial state of $\qvec$ to a final state of $\qvec + \qvec'$.
-The cost of this trade is $C(\qvec + \qvec') - C(\qvec)$.
-The cost function of our market is the log-partition function of an exponential family, and thus, $C(.)$ is a strictly convex function~\cite{WainJordan08}.
-Thus, we have the inequality:
-\[
-   C(\qvec + \qvec') - C(\qvec)  > \qvec' \nabla C|_\qvec = \qvec' \mu,
- \]
- where $\mu$ is the mean sufficient statistics vector for the distribution with parameters $\qvec$.
-The payoff due to this trade, with outcome $x$, is $\qvec'.\phi(x)$. 
-Now, we observe that under the distribution with parameters $\qvec$, the expected payoff is $\qvec' E[\Phi(x)] = \qvec' \mu$.
-Thus, under this distribution over outcomes $x$, the expected profit is strictly less than $0$.
-This is only possible if the
-realized profit is less than $0$ for at least one outcome $x$. As this statement is true for every $\qvec$ and $\qvec'$, and the profits are path-independent, the market is arbitrage-free.
-
-%Let $\qvec^*=(q^{*}_{0},q^{*}_{1},\ldots,q^{*}_{k})$ be the number of shares of each security held by the traders.
-First we note that under the assumption of perfectly rational, risk-neutral traders with infinite budget and  beliefs as indicated above, the gradient of the cost function at this point is
- %$$\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}=\mu_{i}$$
-%where $\mu_{i}$ is believed to be the expected payoff of the $i^{th}$ security by the last trader who traded in this market.
-We show below that the instantaneous price of security $i$, $\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$ at any $\qvec=\qvec^{*}$, can be written as $\left(\frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$.
-This means that the instantaneous price of the $i^{th}$ security only depends on the difference between the number of held securities of $i$ and 0.
-This ensures that if there exists an instantaneous price that reflects the belief of a trader at any vector $\qvec^{*}$, there must exist a $\qvec'$  that also reflects his beliefs and with the desirable property that $\forall i=0,1,2,\ldots,k\ q_{i}'\geq 0$.
-This additional property ensures that no trader will be required to short sell securities to move the market state to  reflect his beliefs.
-
-\paragraph{Market State and Natural Parameters}
-We define the interpretation function 
-\begin{eqnarray}\label{eqn:int}
-I(\qvec)&=&(q_{1}-q_{0},q_{2}-q_{0},\ldots,q_{k}-q_{0})\nonumber\\
-&=&(\beta_{1}, \beta_{2},\ldots, \beta_{k})=\betavec
-\end{eqnarray}
-This allows us to interpret the state of the market in terms of a prediction on the natural parameters of the distribution.
-
-\begin{theorem}
-The natural parameter vector corresponding to the interpretation of the market state given by Equation \ref{eqn:int} is the vector of their maximum likelihood estimates.
-Further, this interpretation is unique.
-\end{theorem}
-\begin{proof}
-We observe that for $i=1,\ldots,n$,
-\begin{eqnarray*}
-\mu_{i}&=&\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int \exp(\qvec^{T}\Phi(x))dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}\phi_{0}(x)+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}(a-\sum_{i=1}^{n}\phi_{i}(x))+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}a} e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log (e^{q_{0}a} \int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial (q_{0}a)}{\partial q_{i}} \frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\frac{\partial \beta_{i}}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\right)_{\rm \betavec=\betavec^{*}}
-\end{eqnarray*}
-Thus, this choice of parameters satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$. We will now argue that this vector is unique.
-
-If the exponential family is represented so that there is a unique parameter vector associated with each distribution, the representation is said to be minimal.
-The Bernoulli, Gaussian, and Poisson distributions all have minimal representations.
-Now, for an exponential family whose representation is minimal, the gradient is an injection \cite[Prop.
-3.2]{WainJordan08}.
-Thus, there is a unique parameter vector $\betavec$ that satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$, for each $i\in\{1,\ldots,n\}$.
-Thus, if the $\mu_{i}$'s correspond to the empirical means, since our choice of $\betavec$ satisfies the equality, it must also be the vector corresponding to the MLE.
-\end{proof}
-
-Thus, we have shown that this prediction market (along with its strategic traders) aggregates trader information in a way that the market state can be interpreted as a predictive distribution (the maximum likelihood distribution) over an infinite outcome space.
-
-It is worth noting here that if we can bound the parameter space and the sufficient statistics of the distribution, the market maker defined here has bounded worst case loss: The probability density, and hence log loss, would then be bounded for all market states that could be reached and all outcomes.
-In fact, for many practical applications, a distribution with unbounded parameter space is an approximation for a true space which itself may be bounded.
-In this case, worst case loss would thus be bounded.
-
-\subsection{The Exponential Family Market Mechanism in Adversarial Markets}
+\section{The Exponential Family Market Mechanism in Adversarial Markets}
 In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
@@
 
-%There is distribution under which expected profit is $\leq 0$. Cost function is convex.
-
-%For any $\qvec$ there is a
 In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 

### 2014-02-04 WriteUp/maxentpredmkts.tex
added modified budget section

@@
 
-\section{The Exponential Family Market Mechanism: A Maximum Likelihood Approach}
-In this section, we will set up a prediction market that aggregates beliefs from traders where the outcome is a continuous random variable.
-In particular, we assume that the outcome is drawn from an exponential family distribution.
-Each of the traders has access to a series of points drawn from this distribution.
-In other words, the traders have access an empirical mean of the sufficient statistics of the exponential family.
-Every trader has infinite budget so that the current market price after a trader has traded in the prediction market, exactly reflects his beliefs.
-
-%We will now provide a more formal description of the problem statement.
-%\subsection{Problem Statement}
-%\begin{description}
-%\item[Setup: ] Learning proceeds in rounds $j=1,2,\ldots$ where at the end of each round the algorithm may be provided with $x_{j}$  drawn from a set $\mathcal{X}$ according to some distribution $\mathcal{D}_{j}$, where $\mathcal{D}_{j}$ belongs to an exponential family defined by $k$ sufficient statistics.
-In each round, some sequence of $n$ traders provide their estimates of the empirical means of each of the $k$ sufficient statistics drawn from $\mathcal{D}_{j}$.
-%\item[Define: ] A learning algorithm that learns this distribution by providing a maximum likelihood estimate of its unknown natural parameters based on the empirical means.  
-%\item[Proposed solution: ] Simulate a prediction market defined by a cost function $C$, securities $s_{0},s_{1},s_{2},\ldots, s_k$ and their payoffs $\phi_i(x)$ for $i=0,1,2,\ldots,k$ for the LMSR.
-Define a correspondence between the state of the prediction market and the output of the learning algorithm.
-%\end{description}
-
-For $x_{i}\sim P_{\betavec}$, recall that the likelihood function for independently drawn data $x_1,\ldots,x_n$ is given by $\prod_{i=1}^{n} P_{\betavec}(x_i)$.
-The maximum likelihood estimate of the natural parameters $\betavec$ is the value of the natural parameters that maximizes the likelihood function.
-
-We will now set up a prediction market with log market scoring rule (LMSR) and infinite budget traders such that the market state represents the maximum likelihood estimate (MLE) of the natural parameters of an exponential family distribution.
-%Given  under some interpretation of what information the participating agents have, and how they bet.
-
-For a given exponential family distribution, the prediction market is defined as follows:
-\begin{description}
-\item[Traders] The prediction market will simulate a trader $i$ corresponding to expert $i$.
-This trader processes all information samples available to her directly or inferred from previous trades, and  forms a belief distribution such that the believed means of the sufficient statistics matches the empirical means of the sufficient statistics from the information samples.
-The trader trades in the market to maximize her expected payoff under her believed distribution.
-\item[Securities and their payoff]
-For each $i=1,2,\ldots,k$, we define a security $s_{i}$ with payoff $\phi_{i}(x)$ where $x$ is the ultimate outcome and $\Phi()$ defines the vector of sufficient statistics of the exponential family distribution according to which $x$ is drawn.
-We define an additional security $s_{0}$ with payoff $\phi_{0}(x):=a-\sum_{i=1}^{k}\phi_{i}(x)$ where $a$ is an appropriately chosen constant dependent on the range of $\Phi$ so that the payoff of $s_{0}$ is non-negative.
-\end{description}
-
-We also note that since we want non-negative payoffs, we restrict the exponential families under consideration so that the sufficient statistics are lower bounded.
-If we have a constant lower bound on the sufficient statistics, then without loss of generality we can add a constant to each sufficient statistic without changing the exponential family in any way, as the constants will be absorbed in the log-partition function.
-
-\paragraph{Arbitrage-free Property}
-We can easily show that the prediction market we have defined is arbitrage-free: there is no sequence of trades that guarantees the trader a profit under all conditions.
-First, note that the incremental cost and payoff function are additive over multiple trades, so the net profit of a sequence of trades depends only on the initial and final market position, and is independent of the actual path along which the trade takes place.
-
-Now, consider any trade (or sequence of trades) that moves the market from an initial state of $\qvec$ to a final state of $\qvec + \qvec'$.
-The cost of this trade is $C(\qvec + \qvec') - C(\qvec)$.
-The cost function of our market is the log-partition function of an exponential family, and thus, $C(.)$ is a strictly convex function~\cite{WainJordan08}.
-Thus, we have the inequality:
-\[
-   C(\qvec + \qvec') - C(\qvec)  > \qvec' \nabla C|_\qvec = \qvec' \mu,
- \]
- where $\mu$ is the mean sufficient statistics vector for the distribution with parameters $\qvec$.
-The payoff due to this trade, with outcome $x$, is $\qvec'.\phi(x)$. 
-Now, we observe that under the distribution with parameters $\qvec$, the expected payoff is $\qvec' E[\Phi(x)] = \qvec' \mu$.
-Thus, under this distribution over outcomes $x$, the expected profit is strictly less than $0$.
-This is only possible if the
-realized profit is less than $0$ for at least one outcome $x$. As this statement is true for every $\qvec$ and $\qvec'$, and the profits are path-independent, the market is arbitrage-free.
-
-%Let $\qvec^*=(q^{*}_{0},q^{*}_{1},\ldots,q^{*}_{k})$ be the number of shares of each security held by the traders.
-First we note that under the assumption of perfectly rational, risk-neutral traders with infinite budget and  beliefs as indicated above, the gradient of the cost function at this point is
- %$$\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}=\mu_{i}$$
-%where $\mu_{i}$ is believed to be the expected payoff of the $i^{th}$ security by the last trader who traded in this market.
-We show below that the instantaneous price of security $i$, $\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$ at any $\qvec=\qvec^{*}$, can be written as $\left(\frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$.
-This means that the instantaneous price of the $i^{th}$ security only depends on the difference between the number of held securities of $i$ and 0.
-This ensures that if there exists an instantaneous price that reflects the belief of a trader at any vector $\qvec^{*}$, there must exist a $\qvec'$  that also reflects his beliefs and with the desirable property that $\forall i=0,1,2,\ldots,k\ q_{i}'\geq 0$.
-This additional property ensures that no trader will be required to short sell securities to move the market state to  reflect his beliefs.
-
-\paragraph{Market State and Natural Parameters}
-We define the interpretation function 
-\begin{eqnarray}\label{eqn:int}
-I(\qvec)&=&(q_{1}-q_{0},q_{2}-q_{0},\ldots,q_{k}-q_{0})\nonumber\\
-&=&(\beta_{1}, \beta_{2},\ldots, \beta_{k})=\betavec
-\end{eqnarray}
-This allows us to interpret the state of the market in terms of a prediction on the natural parameters of the distribution.
-
-\begin{theorem}
-The natural parameter vector corresponding to the interpretation of the market state given by Equation \ref{eqn:int} is the vector of their maximum likelihood estimates.
-Further, this interpretation is unique.
-\end{theorem}
-\begin{proof}
-We observe that for $i=1,\ldots,n$,
-\begin{eqnarray*}
-\mu_{i}&=&\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int \exp(\qvec^{T}\Phi(x))dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}\phi_{0}(x)+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}(a-\sum_{i=1}^{n}\phi_{i}(x))+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log\int e^{q_{0}a} e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial\log (e^{q_{0}a} \int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial (q_{0}a)}{\partial q_{i}} \frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\frac{\partial \beta_{i}}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
-&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\right)_{\rm \betavec=\betavec^{*}}
-\end{eqnarray*}
-Thus, this choice of parameters satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$. We will now argue that this vector is unique.
-
-If the exponential family is represented so that there is a unique parameter vector associated with each distribution, the representation is said to be minimal.
-The Bernoulli, Gaussian, and Poisson distributions all have minimal representations.
-Now, for an exponential family whose representation is minimal, the gradient is an injection \cite[Prop.
-3.2]{WainJordan08}.
-Thus, there is a unique parameter vector $\betavec$ that satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$, for each $i\in\{1,\ldots,n\}$.
-Thus, if the $\mu_{i}$'s correspond to the empirical means, since our choice of $\betavec$ satisfies the equality, it must also be the vector corresponding to the MLE.
-\end{proof}
-
-Thus, we have shown that this prediction market (along with its strategic traders) aggregates trader information in a way that the market state can be interpreted as a predictive distribution (the maximum likelihood distribution) over an infinite outcome space.
-
-It is worth noting here that if we can bound the parameter space and the sufficient statistics of the distribution, the market maker defined here has bounded worst case loss: The probability density, and hence log loss, would then be bounded for all market states that could be reached and all outcomes.
-In fact, for many practical applications, a distribution with unbounded parameter space is an approximation for a true space which itself may be bounded.
-In this case, worst case loss would thus be bounded.
-
-\subsection{The Exponential Family Market Mechanism in Adversarial Markets}
+\section{The Exponential Family Market Mechanism with Budgets}
 In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
@@
 
-%There is distribution under which expected profit is $\leq 0$. Cost function is convex.
-
-%For any $\qvec$ there is a
 In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
 Let $\alpha$ be the budget of a trader in the market.
-Suppose with infinite budget, the trader would have moved the market state from $\qvec_{init}$ to $\qvec$, where $\qvec$ represents his true beliefs.
-Suppose further that $\alpha<C(\qvec)-C(\qvec_{init})$.
+Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
+Suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
 In this case, we want to budget-limit the trader's influence on the market state.
 
-We define the budget-limited final market state as $\tilde{\qvec}$. 
-Here, we consider a specific functional form of $\tilde{\qvec}$: $$\tilde{\qvec}=\lambda\qvec+(1-\lambda)\qvec_{init}$$  where $$\lambda=\min\Big(1, \frac{\alpha}{C(\qvec) - C(\qvec_{init})}\Big)$$ We first show that this trade is feasible given the trader's budget:
+We define the budget-limited final market state as $\theta'$. 
+Here, we consider a specific functional form of $\theta'$: $$\theta'=\lambda\htheta+(1-\lambda)\theta$$  where $$\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$$
+
+First, we  show that this trade is feasible given the trader's budget:
 \begin{theorem}
-Let the current market state be given by the vector $\qvec_{init}$ and $\tilde{\qvec}=\lambda\qvec+(1-\lambda)\qvec_{init}$.
-For $\lambda=\min\Big(1, \frac{\alpha}{C(\qvec) - C(\qvec_{init})}\Big)$, the cost to the trader to move the market state from $\qvec_{init}$ to $\tilde{\qvec}$ is at most his budget $\alpha$.
+Let the current market state be given by $\theta$.
+Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
+The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$.
 \end{theorem}
@@
 From the convexity of $C$, we have
- $$C(\tilde{\qvec})\leq(1-\lambda)C(\qvec_{init})+\lambda C(\qvec)$$
+ $$C(\theta')\leq(1-\lambda)C(\theta)+\lambda C(\htheta)$$
 
@@
 \begin{eqnarray*}
-C(\tilde{\qvec}) - C(\qvec_{init}) &\leq& (1-\lambda)C(\qvec_{init})\\&&+\lambda C(\qvec) - C(\qvec_{init})\\
-&=& \lambda\ (C(\qvec) - C(\qvec_{init}))
-\end{eqnarray*}
-
-%So when $\alpha<C(\qvec)-C(\qvec_{init})$, we pick $\lambda$ so that 
-%$$\lambda\leq \frac{\alpha}{(C(\qvec) - C(\qvec_{init}))}$$
-%
-%where we want $\lambda$ such that
- Thus, $C(\tilde{\qvec}) - C(\qvec_{init}) \leq \alpha$. 
+C(\theta') - C(\theta) &\leq& (1-\lambda)C(\theta)\\
+&&+\lambda C(\htheta) - C(\theta)\\
+&=& \lambda\ (C(\htheta) - C(\theta))
+\end{eqnarray*}
+
+ Thus, $C(\theta') - C(\theta) \leq \alpha$. 
  \end{proof}
 
- We note that moving to $\tilde{\qvec}$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
-In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\tilde{\qvec}$.
-Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\qvec$.
-\subsubsection{Damage Bound}
-Given this restriction on the initial budget of traders, we can now show that the total loss that can be induced by malicious traders is bounded.
-We define the loss function for $\qvec$ shares held as:
-$$L(\qvec,x)=-\log(p_{\betavec}(x))=\log\frac{\int e^{\qvec^{T}\Phi(x)}\ dx}{e^{\qvec^{T}\Phi(x)}}$$
-with the correspondence between $\betavec$ and $\qvec$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
-
-%{\bf Setup:} The market evolves in rounds $t=1,\ldots,T$.
-The end of a time segment occurs either when each expert has made his report or when the actual value of the random variable is received as input.
-Each expert $i$ reports the means of the sufficient statistics as available to him at time $t$.
-These are combined with the previously reported means to yield $\mu_{0}^{t}(x),\mu_{1}^{t}(x),\ldots,\mu_{k}^{t}(x)$ and an associated weight $w_{i}^{t}$ which can be interpreted as the number of points from which these values have been determined.
-These sufficient statistics correspond to a particular value of $B$ and hence of $\qvec$, the number of shares of each security that are held by traders.
-Since trader $i$ has finite budget $\alpha_{i}^{t}$ at time $t$, the current market state is moved to $\qvec'$ instead, so that cost of moving it to $\qvec'$ is within his budget.
-
-%For an initial positive budget of $\alpha_{i}^{0}$, we perform updates to the budgets of each agent as
-%$$\alpha_{i}^{t}:=\alpha_{i}^{t-1}+\lambda_{i}^{t}(L(\tilde{q}_{i-1}^{t},x^{t})-L({q}_{i}^{t},x^{t}))$$ where we define
-%$$\lambda_{i}^{t}:=\min\Big(1, \frac{\alpha_{i}^{t-1}}{C(\qvec_{i}) - C(\tilde{\qvec}_{i-1})}\Big)$$
+ We note that moving to $\theta'$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
+In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\theta$.
+Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\htheta$.
+\subsection{Damage Bound}
+In this section, we will quantify the error in prediction that the market maker might have to endure as a result of malicious entities entering the market.
+We assume that these malicious entities trade in multiple instances of the market; thus the exposure of the market maker is over several \emph{rounds}.
+We use the standard log loss to measure this error in terms of the initial budget of traders.
+
+We define the loss function for $\theta$ shares held:
+$$L(\theta,x)=-\log(P_{\theta}(x))=\log\int \exp\{\theta^{T}\phi(x)\}\ dx-\theta^{T}\phi(x)=C(\theta)-\theta^{T}\phi(x)$$
+%with the correspondence between $\beta$ and $\theta$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
 
 Suppose the prediction market runs over multiple rounds $t$.
-Let $\qvec_{0}^{t}$ be the initial number of shares of each security that are held.
-Let $\tilde{\qvec}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
-Let us assume that at this point we receive the value of the random variable $x^{t}$.
-
-%Let the loss function be $L(\cdot,\cdot)$, where the two arguments are the vector representing the number of shares held of each security and the actual value of the random variable.
-
-%We want to show:
-%$$L({\qvec}_{0}^{t},x^{t})-L(\tilde{\qvec}_{k}^{t},x^{t})+\sum_{i}\alpha_{i}^{t-1}\leq \sum_{i}\alpha_{i}^{t}$$
-%That is, the loss due to incorporating the budget-limited information of the traders, is not more than the increase in their budgets.
-Over multiple instances of the prediction market, we can track the change in budget of each trader. The change in budget for trader $i$ is
-\begin{eqnarray*}
-\alpha_{i}^{t}-\alpha_{i}^{t-1}&=&C(\tilde{\qvec}_{i-1}^{t})-C(\tilde{\qvec}_{i}^{t})-(\tilde{\qvec}_{i-1}^{t}-\tilde{\qvec}_{i}^{t})^{T}\Phi(x^{t})\\
-&=&L(\tilde{\qvec}_{i-1}^{t},x^{t})-L(\tilde{\qvec}_{i}^{t},x^{t})
+Let $\theta_{0}^{t}$ be the initial number of shares of each security that are held.
+Let $\hat{\theta}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
+Let us assume that at this point the outcome is revealed; that is, we receive the value of the random variable $x^{t}$.
+
+Over multiple instances of the prediction market, we can track the change in budget of each trader.
+Let the budget at rounds $t$ and $t-1$ be $\alpha^{t}$ and $\alpha^{t-1}$ respectively.
+The change in budget for the trader is
+\begin{eqnarray*}
+\alpha^{t}-\alpha^{t-1}&=&C(\theta^{t})-C(\theta'^{t})-(\theta^{t}-\theta'^{t})^{T}\phi(x^{t})\\
+&=&L(\theta,x^{t})-L(\theta',x^{t})
 \end{eqnarray*}
@@
 Define the myopic impact of a trader $i$ in segment $t$ as
-$$\Delta_{i}^{t}:=L(\tilde{\qvec}_{i-1}^{t},x^{t})-L(\tilde{\qvec}_{i}^{t},x^{t})$$
-Thus, the myopic impact captures incremental gain due to the trader in a round. Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
+$$\Delta_{i}^{t}:=L(\hat{\theta}_{i-1}^{t},x^{t})-L(\hat{\theta}_{i}^{t},x^{t})$$
+Thus, the myopic impact captures incremental gain in prediction due to the trader in a round.
+Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
 
 The total myopic impact due to all $k$ active traders is given by
-$$\Delta^{t}=L(\qvec_{0}^{t},x^{t})-L(\tilde{\qvec}_{k}^{t},x^{t})$$
-Thus $-\Delta^{t}$ captures the incremental loss due to the predictive probability after aggregation of all $k$ traders.
-
-%The myopic impact of a trader $i$ in segment $t$, is given by
-%$$\Delta_{i}^{t}:=L(\tilde{B}_{i-1}^{t},x^{t})-L(\tilde{B}_{i}^{t},x^{t})$$
-%where $\tilde{B}_{i}$ is the vector of sufficient statistics corresponding to the the budget-limited state of the market as induced by trader $i$.
-%The total myopic impact due to all $k$ active traders is thus given by
-%$$\Delta^{t}=L(B_{0}^{t},x^{t})-L(\tilde{B}_{k}^{t},x^{t})$$
-
-%If the loss function is convex in the first argument, then we have:
-%\begin{eqnarray*}
-%\Delta_{i}^{t}&=&L(\tilde{B}_{i-1}^{t},x^{t})-L(\tilde{B}_{i}^{t},x^{t})\\
-%&=&L(\tilde{q}_{i-1}^{t},x^{t})-L(\tilde{q}_{i}^{t},x^{t})\\
-%&=&L(\tilde{q}_{i-1}^{t},x^{t})-L(\lambda_{i}{q}_{i}^{t}+(1-\lambda_{i})\tilde{q}_{i-1}^{t},x^{t})\\
-%&\geq&L(\tilde{q}_{i-1}^{t},x^{t})-\lambda_{i}L({q}_{i}^{t},x^{t})-(1-\lambda_{i})L(\tilde{q}_{i-1}^{t},x^{t})\\
-%&=&\lambda_{i}(L(\tilde{q}_{i-1}^{t},x^{t})-L({q}_{i}^{t},x^{t}))\\
-%&=& \alpha_{i}^{t} - \alpha_{i}^{t-1}
-%\end{eqnarray*}
-%Thus, the myopic damage caused by trader $i$ at round $t$ is upper-bounded by the change in his budget in that round.
-
-%If we pick a loss function that is also bounded in $[0,1]$, then we see that the budget never falls below zero. 
+$$\Delta^{t}=L(\theta_{0}^{t},x^{t})-L(\hat{\theta}_{k}^{t},x^{t})$$
+Thus $-\Delta^{t}$ captures the incremental loss of the market prediction after aggregation of all $k$ traders.
+
 \begin{theorem}
@@
 
-This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders.
+This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders. %imp
 
@@
 \begin{theorem}\label{lem:profit_decomposition}{\bf (Profit Decomposition)}: Consider an exponential 
-family $\family$ of distributions over some set of statistics $\suff(x)$, with natural parameters $\betavec$. Let $\pi, \rho \in \family$ be any two probability distributions in the family. We use 
-$\betavec_\pi$ to denote the natural parameters of $\pi$ and $\muvec_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
-we can define $\betavec_\rho$ and  $\muvec_\rho$. We abuse notation slightly and let $\lpf(\rho)$ indicate the log partition function of $\rho$ which technically depends on its natural parameters.
+family $\family$ of distributions over some set of statistics $\suff(x)$, with natural parameters $\beta$. Let $\pi, \rho \in \family$ be any two probability distributions in the family. We use 
+$\beta_\pi$ to denote the natural parameters of $\pi$ and $\mu_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
+we can define $\beta_\rho$ and  $\mu_\rho$. Let $\lpf(\beta_{\rho})$ indicate the log partition function of $\rho$ which technically depends on its natural parameters.
 Let $H(\pi)$ denote the entropy of the distribution $\pi$, and 
@@
 \begin{equation}
- K(\pi||\rho) + H(\pi) = \lpf(\rho) - \betavec_\rho \cdot \muvec_\pi
+ K(\pi||\rho) + H(\pi) = \lpf(\beta_{\rho}) - \beta_\rho \cdot \mu_\pi
 \label{eq:profit_decomposition}
@@
 &=& - \int_x \pi(x)\log{\rho(x)}dx \\
-&=& - \int_x \pi(x)\left[ \betavec_\rho \cdot \suff(x) - \lpf(\rho)\right]dx\\
-&=& - \betavec_\rho \cdot \int_x \pi(x) \suff(x) dx\\
-&&\quad + \lpf(\rho) \int_x \pi(x) dx \\
-&=&  \lpf(\rho) - \betavec_\rho \cdot \muvec_\pi
+&=& - \int_x \pi(x)\left[ \beta_\rho \cdot \suff(x) - \lpf(\beta_{\rho})\right]dx\\
+&=& - \beta_\rho \cdot \int_x \pi(x) \suff(x) dx\\
+&&\quad + \lpf(\beta_{\rho}) \int_x \pi(x) dx \\
+&=&  \lpf(\beta_{\rho}) - \beta_\rho \cdot \mu_\pi
 \end{eqnarray*}
@@
 
-%Thus, we have shown that there exist deep connections between learning the natural parameters of an exponential family distribution and cost function based prediction markets.
-In Chapter \ref{nonmyopic} we exploit this connection and propose a bounded regret learning algorithm in a Bayesian model under less-than-ideal conditions: where the experts may be adversarial and strategic.
-
 \subsubsection{Budget of Informative Traders}\label{sec:inf}
@@
 
-%For this, we need the following simple extension of Jensen's inequality that follows directly from the definition of convexity of a function.
-%\begin{lemma}\label{lem:convex}
-%For a strictly convex function $f$ with unique minimum $x_{min}$, $x\neq x_{min}$ in the domain of $f$, and $\lambda \in (0,1]$,
-%$$f(\lambda x_{min}+ (1-\lambda)x)< f(x)$$
-%\end{lemma}
-%\begin{proof}
-%\begin{eqnarray*}
-%f(\lambda x_{min}+ (1-\lambda)x) &<& \lambda f(x_{min})+ (1-\lambda)f(x)\\
-%&=& \lambda(f(x_{min})-f(x)) +f(x)\\
-%&\leq& f(x)
-%\end{eqnarray*}
-%\end{proof}
-
 We now characterize the expected change in budget for an informative trader. The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
 \begin{theorem}\label{thm:growth}
- Suppose that each informative trader gets a random sample of data, 
-resulting in a sequence of trader beliefs $\qvec_i$, and hence a sequence of budget-limited market positions
-$\tilde{\qvec}_i$, before a final outcome $x$.
-Then, the expectation, over trader $i$'s belief distribution $\qvec_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\tilde{\qvec}_{i-1}$.
+Let $\theta$ be the current market state.
+Suppose that an  informative trader with belief distribution  parametrized by $\theta'$  moves the market state to the budget-limited state $\htheta=\lambda \theta' + (1-\lambda) \theta$.
+Then, the expectation (over the trader's belief) of the trader's profit  is greater than zero whenever his budget is positive and his belief differs from the previous market position $\theta$.
+% Suppose that each informative trader gets a random sample of data, 
+%resulting in a sequence of trader beliefs $\theta_i$, and hence a sequence of budget-limited market positions
+%$\hat{\theta}_i$, before a final outcome $x$.
+Then, the expectation, over trader $i$'s belief distribution $\theta_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\hat{\theta}_{i-1}$.
 \end {theorem}
 \begin{proof}
-Trader $i$'s believed distribution is the distribution parametrized by $\qvec_i$.
-Therefore, the expected profit, over possible outcome values $x$, for a trader $i$ in a given round is given by
-\begin{eqnarray*}
-&& E[C(\tilde{\qvec}_{i-1})-C(\tilde{\qvec}_{i})-(\tilde{\qvec}_{i-1}-\tilde{\qvec}_{i})\Phi(x)]\\
-&=& K(\qvec_{i}||\tilde{\qvec}_{i-1})] + H(\qvec_{i}) - [K(\qvec_{i}||\tilde{\qvec}_{i}) + H(\pi)]]\\
-&=& K(\qvec_{i}||\tilde{\qvec}_{i-1}) - K(\qvec_{i}||\tilde{\qvec}_{i})
-\end{eqnarray*}
-We recall that $\tilde{\qvec}_{i}$ can be expanded as:
-$$\tilde{\qvec_{i}} = \lambda \qvec_{i} + (1- \lambda) \tilde{\qvec}_{i-1}$$ where $\lambda$ is strictly greater than $0$, but no more than $1$.
-It is a standard result from information theory that $K(.,.)$ is a convex function, and therefore, we have:
-\[
-K(\qvec_{i}||\tilde{\qvec}_{i-1}) \leq \lambda K(\qvec_{i}||\qvec_{i}) + (1-\lambda) K(\qvec_{i}||\tilde{\qvec}_{i-1})
-\]
-The first term on the right hand side is zero, and thus, we get:
-\[
-  K(\qvec_{i}||\tilde{\qvec}_{i-1}) - K(\qvec_{i}||\tilde{\qvec}_{i}) \geq \lambda K(\qvec_{i}||\tilde{\qvec}_{i-1})
-\]
-%Thus, it is sufficient to prove that E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t}] >= E_{\qvec}[K(\pi||\tilde{\qvec}_{i}^{t}].
+Let $\theta$ be the current market state.
+We have shown that an exponential utility trader with belief distribution  parametrized by $\beta$ will move the market state to $\theta'=\frac{1}{1+a} \beta + \frac{a}{1+a} \theta$.
+Therefore, the trader's expected net payoff is given by
+\begin{eqnarray*}
+&& \E_{x\sim P_{\beta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
+&=& \psi(\theta)-\theta\E_{x\sim P_{\beta}}[\phi(x)]-(\psi(\theta')-\theta' \E_{x\sim P_{\beta}}[\phi(x)])\\
+&=& \psi(\theta)-\theta\nabla\psi(\beta)-(\psi(\theta')-\theta' \nabla\psi(\beta))\\
+&=& \psi(\theta)-\psi(\beta)-\nabla\psi(\beta)(\theta-\beta)-(\psi(\theta')-\psi(\beta)- \nabla\psi(\beta)(\theta'-\beta))\\
+&=& D_{\psi}(\theta,\beta)-D_{\psi}(\theta',\beta)\\
+&\geq&\frac{1}{a}D_{\psi}(\theta',\beta)\geq0
+\end{eqnarray*}
+The second to last inequality holds since $D_{\psi}(\theta',\beta)$ is convex in $\theta'$ and we have:
+\begin{eqnarray*}
+D_{\psi}(\theta',\beta)&=&D_{\psi}\left(\frac{1}{1+a} \beta + \frac{a}{1+a} \theta,\beta\right)\\
+&\leq&\frac{1}{1+a}D_{\psi}( \beta ,\beta)+ \frac{a}{1+a} D_{\psi}( \theta ,\beta)\\
+&=&\frac{a}{1+a} D_{\psi}( \theta ,\beta)
+%\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
+\end{eqnarray*}¥
+Thus, a trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{\psi}(\theta',\beta)$.
  \end{proof}
-%\begin{theorem}
-%The expected budget of an informative trader increases when he participates in the market. Here, expectation is taken with respect to the true distribution from which the outcome is drawn.
-%We show that the budget of an informative trader grows in expectation:\end{theorem}
-%\begin{proof}
-%Let $\pi$ be the true distribution from which the outcome $x^{t}$ is drawn.
-Let $K(\pi||\qvec)$ be the relative entropy between the true distribution $\pi$ and the predictive distribution determined by the securities vector $\qvec$.
-Then the expected change in budget for a trader $i$ in round $t$ is given by
-%\begin{eqnarray*}
-%E_{x,\qvec}[\alpha_{i}^{t}&-&\alpha_{i}^{t-1}]\\
-%&=&E_{x,\qvec}[C(\tilde{\qvec}_{i-1}^{t})-C(\tilde{\qvec}_{i}^{t})-(\tilde{\qvec}_{i-1}^{t}-\tilde{\qvec}_{i}^{t})^{T}\Phi(x^{t})]\\
-%&=& E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t})] + H(\pi) - E[K(\pi||\tilde{\qvec}_{i}^{t}) + H(\pi)]]
-%\end{eqnarray*}
-%Thus, it is sufficient to prove that E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t}] >= E_{\qvec}[K(\pi||\tilde{\qvec}_{i}^{t}].
-
-%In the case that
-
-%It is well known that minimizing the relative entropy of a predictive distribution with respect to the true distribution is equivalent to maximizing the likelihood function.
-% (see, for instance, \cite{whom?}).
-%In other words, the MLE of the natural parameters of an exponential family distribution minimizes the relative entropy of a predictive distribution with respect to the true distribution.
-
-%Further, for a true distribution $\pi$ and predictive distribution determined by $\qvec$, $K(\pi||\qvec) + H(\pi)$ is a convex function of $\qvec$.
-In fact, it is strictly convex when the representation is minimal.
-Thus, it has a unique minimum.
-%Further, $K(\pi||\rho) + H(\pi)$ is non-negative.
-
-%Suppose the current market state is represented by $\tilde{\qvec}_{i-1}$ and $\qvec_{i}$ represents the belief of trader $i$.
-Since the trader is budget limited, he moves the market state to $\tilde{\qvec}_{i}$ instead where $\tilde{\qvec}_{i}=  \lambda \qvec_{i}+ (1-\lambda)\tilde{\qvec}_{i-1}$.
-For an informative trader $i$, $\qvec_{i}$ represents the MLE of the natural parameters of the distribution.
-Thus, $K(\pi||\qvec) + H(\pi)$ has a unique minimum at $\qvec_{i}$.
-
-%Thus,   from Lemma \ref{lem:convex}, $E[K(\pi||\tilde{\qvec}_{i-1}^{t})+ H(\pi)] > E[K(\pi||\tilde{\qvec}_{i}^{t})+ H(\pi)]$ and trader $i$'s expected budget increases.
-%\end{proof}
 
@@
 
+Notice that Theorem~\ref{thm:growth} only required that the market state to which the trader moves, be representable as a convex combination of the current market state and his belief.
+This means that the result holds for exponential utility traders aiming to maximize their utility.
+In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{\psi}(\theta',\beta)$ where is the exponential utility parameter.
+
 We note one important aspect of Theorem~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
-This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure; depending on the distribution family and the prior distribution over parameter values, maximum likelihood estimation might not optimize the true expected score.
+This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure.
 If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
-%To summarize our results for this in a market with budget-limited traders, the total loss induced by malicious traders is limited, while the budget of  informative traders grows.
+
 \section{The MaxEnt Market Mechanism: A Bayesian View}
@@
 \frac{\theta+\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
-Let us modify the cost function so that it includes a liquidity parameter:
-$$C_{\lambda}(q)\defeq\lambda \psi(q/\lambda)$$
-
-\subsubsection{Lower Bound on Expected Payoff for Exponential Utility Traders}
-Let $\theta$ be the current market state.
-We have shown that an exponential utility trader with belief distribution  parametrized by $\beta$ will move the market state to $\theta'=\frac{1}{1+a} \beta + \frac{a}{1+a} \theta$.
-Therefore, the trader's expected net payoff is given by
-\begin{eqnarray*}
-&& \E_{x\sim P_{\beta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
-&=& \psi(\theta)-\theta\E_{x\sim P_{\beta}}[\phi(x)]-(\psi(\theta')-\theta' \E_{x\sim P_{\beta}}[\phi(x)])\\
-&=& \psi(\theta)-\theta\nabla\psi(\beta)-(\psi(\theta')-\theta' \nabla\psi(\beta))\\
-&=& \psi(\theta)-\psi(\beta)-\nabla\psi(\beta)(\theta-\beta)-(\psi(\theta')-\psi(\beta)- \nabla\psi(\beta)(\theta'-\beta))\\
-&=& D_{\psi}(\theta,\beta)-D_{\psi}(\theta',\beta)\\
-&\geq&\frac{1}{a}D_{\psi}(\theta',\beta)\geq0
-\end{eqnarray*}
-The second to last inequality holds since $D_{\psi}(\theta',\beta)$ is convex in $\theta'$ and we have:
-\begin{eqnarray*}
-D_{\psi}(\theta',\beta)&=&D_{\psi}\left(\frac{1}{1+a} \beta + \frac{a}{1+a} \theta,\beta\right)\\
-&\leq&\frac{1}{1+a}D_{\psi}( \beta ,\beta)+ \frac{a}{1+a} D_{\psi}( \theta ,\beta)\\
-&=&\frac{a}{1+a} D_{\psi}( \theta ,\beta)
-%\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
-\end{eqnarray*}¥
-Thus, a trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{\psi}(\theta',\beta)$.
+
 \section{Conclusions}

### 2014-02-04 WriteUp/maxentpredmkts.tex
some clean-up on budgets

@@
 %------------------------------------------------------------
-%- exp family review
-%- maxent scoring rule
+%-introduction [Sindhu; Thursday]
+%-literature review
+%- exp family review [Sindhu; Tuesday]
+%- maxent scoring rule [Sebastien]
 %	- generalized lmsr
-%	- does modularity (a la Hanson) still hold?
-%- prediction market setup
+%	[- does modularity (a la Hanson) still hold?]
+%- prediction market setup [Sebastien]
 %	- MLE with risk neutrality
-%- conjugate prior market
+%- conjugate prior market [Sindhu; Wednesday]
 %	- also on hyperparam?
-%- exp-util
+%- exp-util [Sebastien]
 %- equilibrium for repeated trades
-%- budgets and adversarial markets
+% 	[- rate of convergence?]
+%- budgets and adversarial markets [Sindhu; Monday]
+% [any liquidity results?] [Sindhu: weekend]
+% conclusion
+% bibliography
 
@@
 
+In this section, we assume that the traders have exponential family beliefs.
+The cost function has the same form as the log partition function $\psi$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
 \subsection{Budget-limited Aggregation}
@@
 Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
-Suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
+Let $C$ be the cost function.
+Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
 In this case, we want to budget-limit the trader's influence on the market state.
@@
 $\beta_\pi$ to denote the natural parameters of $\pi$ and $\mu_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
-we can define $\beta_\rho$ and  $\mu_\rho$. Let $\lpf(\beta_{\rho})$ indicate the log partition function of $\rho$ which technically depends on its natural parameters.
+we can define $\beta_\rho$ and  $\mu_\rho$. Let $\lpf(\beta_{\rho})$ indicate the log partition function of $\rho$.
 Let $H(\pi)$ denote the entropy of the distribution $\pi$, and 
@@
 \begin{theorem}\label{thm:growth}
-Let $\theta$ be the current market state.
+Let $\theta$ be the current market state in the exponential family prediction market.
 Suppose that an  informative trader with belief distribution  parametrized by $\theta'$  moves the market state to the budget-limited state $\htheta=\lambda \theta' + (1-\lambda) \theta$.
@@
 \begin{proof}
-Let $\theta$ be the current market state.
-We have shown that an exponential utility trader with belief distribution  parametrized by $\beta$ will move the market state to $\theta'=\frac{1}{1+a} \beta + \frac{a}{1+a} \theta$.
-Therefore, the trader's expected net payoff is given by
-\begin{eqnarray*}
-&& \E_{x\sim P_{\beta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
-&=& \psi(\theta)-\theta\E_{x\sim P_{\beta}}[\phi(x)]-(\psi(\theta')-\theta' \E_{x\sim P_{\beta}}[\phi(x)])\\
-&=& \psi(\theta)-\theta\nabla\psi(\beta)-(\psi(\theta')-\theta' \nabla\psi(\beta))\\
-&=& \psi(\theta)-\psi(\beta)-\nabla\psi(\beta)(\theta-\beta)-(\psi(\theta')-\psi(\beta)- \nabla\psi(\beta)(\theta'-\beta))\\
-&=& D_{\psi}(\theta,\beta)-D_{\psi}(\theta',\beta)\\
-&\geq&\frac{1}{a}D_{\psi}(\theta',\beta)\geq0
-\end{eqnarray*}
-The second to last inequality holds since $D_{\psi}(\theta',\beta)$ is convex in $\theta'$ and we have:
-\begin{eqnarray*}
-D_{\psi}(\theta',\beta)&=&D_{\psi}\left(\frac{1}{1+a} \beta + \frac{a}{1+a} \theta,\beta\right)\\
-&\leq&\frac{1}{1+a}D_{\psi}( \beta ,\beta)+ \frac{a}{1+a} D_{\psi}( \theta ,\beta)\\
-&=&\frac{a}{1+a} D_{\psi}( \theta ,\beta)
+Let the cost function $C$ be equal to the log partition function $\psi$ of the belief distribution.
+The payoff is given by the sufficient statistics $\phi(x)$.
+Then, the trader's expected net payoff is given by
+\begin{eqnarray*}
+&& \E_{x\sim P_{\htheta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
+&=& \psi(\theta)-\theta\E_{x\sim P_{\htheta}}[\phi(x)]-(\psi(\theta')-\theta' \E_{x\sim P_{\htheta}}[\phi(x)])\\
+&=& \psi(\theta)-\theta\nabla\psi(\htheta)-(\psi(\theta')-\theta' \nabla\psi(\htheta))\\
+&=& \psi(\theta)-\psi(\htheta)-\nabla\psi(\htheta)(\theta-\htheta)-(\psi(\theta')-\psi(\htheta)- \nabla\psi(\htheta)(\theta'-\htheta))\\
+&=& D_{\psi}(\theta,\htheta)-D_{\psi}(\theta',\htheta)\\
+&\geq&\lambda D_{\psi}(\theta,\htheta)\geq0
+\end{eqnarray*}
+The second to last inequality holds since $D_{\psi}(\theta',\htheta)$ is convex in $\theta'$ and we have:
+\begin{eqnarray*}
+D_{\psi}(\theta',\htheta)&=&D_{\psi}\left(\lambda\htheta+(1-\lambda)\theta,\htheta\right)\\
+&\leq&\lambda D_{\psi}( \htheta ,\htheta)+ (1-\lambda) D_{\psi}( \theta ,\htheta)\\
+&=&(1-\lambda) D_{\psi}( \theta ,\htheta)
 %\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
 \end{eqnarray*}¥
-Thus, a trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{\psi}(\theta',\beta)$.
+Thus, a trader who moves the market state can expect his profit to be positive and at least $\lambda D_{\psi}(\theta,\htheta)$.
  \end{proof}
@@
 This means that the result holds for exponential utility traders aiming to maximize their utility.
-In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{\psi}(\theta',\beta)$ where is the exponential utility parameter.
+In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{\psi}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
 

### 2014-02-04 WriteUp/maxentpredmkts.tex
added pagebreaks
cleaned up \section{Reinterpreting repeated trades}

@@
 %	- MLE with risk neutrality
+% 	need a definition of term 'exponential family prediction market'
 %- conjugate prior market [Sindhu; Wednesday]
@@
 %- exp-util [Sebastien]
-%- equilibrium for repeated trades
+%- equilibrium for repeated trades [Sindhu; Tuesday]
 % 	[- rate of convergence?]
@@
 % bibliography
+
+%------------------------------------------------------------
+% Naming conventions:
+%------------------------------------------------------------
+%htheta belief
+%theta current market state
+%thetaÕ final state
+%T is the log partition function
 
@@
 Since~(\ref{scoring}) is a proper scoring rule, the information market with cost function~(\ref{cost}) is based on a proper market scoring rule.
+\pagebreak
 
@@
 In this section, we assume that the traders have exponential family beliefs.
-The cost function has the same form as the log partition function $\psi$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
+The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
 \subsection{Budget-limited Aggregation}
@@
 \begin{proof}
-Let the cost function $C$ be equal to the log partition function $\psi$ of the belief distribution.
+Let the cost function $C$ be equal to the log partition function $T$ of the belief distribution.
 The payoff is given by the sufficient statistics $\phi(x)$.
@@
 && \E_{x\sim P_{\htheta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
-&=& \psi(\theta)-\theta\E_{x\sim P_{\htheta}}[\phi(x)]-(\psi(\theta')-\theta' \E_{x\sim P_{\htheta}}[\phi(x)])\\
-&=& \psi(\theta)-\theta\nabla\psi(\htheta)-(\psi(\theta')-\theta' \nabla\psi(\htheta))\\
-&=& \psi(\theta)-\psi(\htheta)-\nabla\psi(\htheta)(\theta-\htheta)-(\psi(\theta')-\psi(\htheta)- \nabla\psi(\htheta)(\theta'-\htheta))\\
-&=& D_{\psi}(\theta,\htheta)-D_{\psi}(\theta',\htheta)\\
-&\geq&\lambda D_{\psi}(\theta,\htheta)\geq0
-\end{eqnarray*}
-The second to last inequality holds since $D_{\psi}(\theta',\htheta)$ is convex in $\theta'$ and we have:
-\begin{eqnarray*}
-D_{\psi}(\theta',\htheta)&=&D_{\psi}\left(\lambda\htheta+(1-\lambda)\theta,\htheta\right)\\
-&\leq&\lambda D_{\psi}( \htheta ,\htheta)+ (1-\lambda) D_{\psi}( \theta ,\htheta)\\
-&=&(1-\lambda) D_{\psi}( \theta ,\htheta)
+&=& T(\theta)-\theta\E_{x\sim P_{\htheta}}[\phi(x)]-(T(\theta')-\theta' \E_{x\sim P_{\htheta}}[\phi(x)])\\
+&=& T(\theta)-\theta\nabla T(\htheta)-(T(\theta')-\theta' \nabla T(\htheta))\\
+&=& T(\theta)-T(\htheta)-\nabla T(\htheta)(\theta-\htheta)-(T(\theta')-T(\htheta)- \nabla T(\htheta)(\theta'-\htheta))\\
+&=& D_{T}(\theta,\htheta)-D_{T}(\theta',\htheta)\\
+&\geq&\lambda D_{T}(\theta,\htheta)\geq0
+\end{eqnarray*}
+The second to last inequality holds since $D_{T}(\theta',\htheta)$ is convex in $\theta'$ and we have:
+\begin{eqnarray*}
+D_{T}(\theta',\htheta)&=&D_{T}\left(\lambda\htheta+(1-\lambda)\theta,\htheta\right)\\
+&\leq&\lambda D_{T}( \htheta ,\htheta)+ (1-\lambda) D_{T}( \theta ,\htheta)\\
+&=&(1-\lambda) D_{T}( \theta ,\htheta)
 %\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
 \end{eqnarray*}¥
-Thus, a trader who moves the market state can expect his profit to be positive and at least $\lambda D_{\psi}(\theta,\htheta)$.
+Thus, a trader who moves the market state can expect his profit to be positive and at least $\lambda D_{T}(\theta,\htheta)$.
  \end{proof}
@@
 This means that the result holds for exponential utility traders aiming to maximize their utility.
-In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{\psi}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
+In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{T}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
 
@@
 If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
+\pagebreak
 
@@
 %
-\begin{proposition}
+\begin{proposition}\label{prop:exp-util}
 An agent with exponential family belief with natural parameter $\htheta$, and exponential utility with coefficient $a$, makes a trade that moves the current market share vector $\theta$ to the convex combination $\frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$ assuming an LMSR cost function.
@@
 \end{eqnarray*}¥
-
-\section{Reinterpreting repeated trades in a market}
-Let us suppose that traders in this market have exponential utility \[
+\pagebreak
+
+\section{Reinterpreting repeated trades}
+In previous sections we have analyzed trader behavior as if it is his first entry into the market.
+In this section we will quantify how prior exposure in the market affects a trader's choices.
+In particular, we will show that a trader who has previously purchased shares in a market will, on subsequent entry, behave as if this purchase has updated his private belief.
+This result implies that any financial trade made in a market is equivalent to changing the trader's effective belief.
+
+As before, we consider the exponential family prediction market where traders have exponential belief. Let us also suppose that traders in this market have exponential utility \[
 U(w) = -\frac{1}{a} \exp(-aw).
@@
 
-Let $\delta_{1}^{*}$ be the optimal vector of shares the agent decides to trade on first entering the market with exponential family belief parametrized by natural parameter $\hat{\theta}$.
-Thus, his belief distribution is given by the pdf $$\exp\{\hat{\theta} \phi(x) - T(\hat{\theta})\}]$$ where $T(\hat{\theta}) =\int_{\mathcal{X}}\exp\{\hat{\theta} \phi(x)\}\ dx$ is the log partition function and $\phi(x)$ are the sufficient statistics.
-On a subsequent entry into this market when the market state is $\theta'$, his optimal purchase $\delta_{2}^{*}$ is given by the solution of
-$$\arg\max_{\delta_{2}} \E_{x\sim \hat{\theta}} U[\mathrm{payoff}]$$
-Here, the payoff given eventual outcome $x$ is  $[\delta_{1}^{*}+\delta_{2}^{*} ]\phi(x) - C(\delta_{1}^{*} + \theta) + C(\theta)- C(\delta_{2}^{*} + \theta') + C(\theta')$.
-Here $\theta' = \theta+\delta_{1}^{*}+\delta'$ where $\delta'$ captures the movement of the share vector by other traders in the market.
-The expected utility for this payoff is as follows.
-%%
-%\[ U( \delta x - C(\delta + \theta) + C(\theta) ) = -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)).
-%\]
-%%
-\begin{eqnarray*}
-&   & \arg\max_{\delta_{2}}\E\left[ (\delta_{1}^{*}+\delta_{2} )\phi(x) - C(\delta_{1}^{*} + \theta) + C(\theta)- C(\delta_{2} + \theta') + C(\theta') \right] \\
+Suppose an agent has exponential family belief parametrized by natural parameter $\hat{\theta}$.
+Based on this belief, let $\delta_{1}^{*}$ be the optimal vector of shares the agent decides to trade on first entering the market.
+Thus, his belief distribution is given by the density $$p(x;\htheta)=\exp(\la\hat{\theta}, \phi(x)\ra - T(\hat{\theta}))$$ where $T(\hat{\theta}) =\int_{\mathcal{X}}\exp\{\la\hat{\theta}, \phi(x)\ra\}\ dx$ is the log partition function and $\phi(x)$ are the sufficient statistics of the outcome $x\in \mathcal{X}$.
+On a subsequent entry into this market with market state $\theta'$, his optimal purchase $\delta_{2}^{*}$ is given by the solution of
+%Here $\theta' = \theta+\delta_{1}^{*}+\delta'$ where $\delta'$ captures the movement of the share vector by other traders in the market.
+\begin{eqnarray*}
+&   & \arg\max_{\delta_{2}} \E_{x\sim p(x;\htheta)} U\left[ (\delta_{1}^{*}+\delta_{2} )\phi(x) - C(\delta_{1}^{*} + \theta) + C(\theta)- C(\delta_{2} + \theta') + C(\theta') \right] \\
 & = & \arg\max_{\delta_{2}}\int_{\mathcal{X}} -\frac{1}{a} [\exp(-a (\delta_{1}^{*}+\delta_{2} )\phi(x) \\
@@
 
-Here the first equality follows from the fact that we are taking expectation over the traders belief parameter $\hat{\theta}$ and the second equality follows simply from rearranging the factors of $\phi(x)$.
+Here, the first equality follows from the fact that we are taking expectation over the agent's belief parameter $\hat{\theta}$ and the second equality follows simply from rearranging the factors of $\phi(x)$.
 And lastly we have written $-\frac{1}{a} [\exp\{-a \delta_{2} \phi(x)+a C(\delta_{2} + \theta') -a C(\theta')]$ as $U[N(\delta_{2},\theta')]$ where $N(\delta_{2},\theta')$ is the net payoff when $\delta_{2}$ shares are purchased when the current market state is $\theta'$.
 
-This is equivalent to maximizing expected utility of $N(\delta_{2},\theta')$, where expectation is taken with respect to an exponential family distribution over $\mathcal{X}$ parametrized by $\hat{\theta}-a \delta_{1}^{*}$, so long as
-$$\int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}]\, dx = c\times \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)-T(\hat{\theta}-a \delta_{1}^{*})\} \, dx$$
-
-%$$\int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)-\psi(\hat{\theta}-a \delta_{1}^{*})\} \, dx =1$$
-%That is
-%$$\psi(\hat{\theta}-a \delta_{1}^{*})= \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)\}\ dx$$
-In other words we want the following statement to hold.
-\begin{eqnarray*}
-\exists c: c\times\exp\{- T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}&=&
-T(\hat{\theta}-a \delta_{1}^{*})\\
-&=& \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)\}\ dx
-\end{eqnarray*}
-
-where $c$ is a constant with respect to $\delta_{2}$ and $x$ (but could potentially depend on $\hat{\theta}$, $a$ and $\delta_{1}^{*}$).
-This is true because the right hand side of the equation is a function of only $\hat{\theta}$, $a$ and $\delta_{1}^{*}$.
-Let $\Theta\defeq\hat{\theta}-a \delta_{1}^{*}  $ be the effective belief.
-Thus we have that the trader chooses his share vector as follows.
+Note that since $C=T$, $- T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)$ is proportional to $T(\hat{\theta}-a \delta_{1}^{*})$, it follows that $\int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}]\, dx$ is proportional to the density with parameter $\hat{\theta}-a \delta_{1}^{*}$.
+Thus, picking the optimal share vector is equivalent to maximizing expected utility of $N(\delta_{2},\theta')$, where expectation is taken with respect to an exponential family distribution over $\mathcal{X}$ parametrized by $\hat{\theta}-a \delta_{1}^{*}$.
+
+Let $\Theta\defeq\hat{\theta}-a \delta_{1}^{*}  $ be the effective belief. Thus we have that the trader chooses his share vector as follows.
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-This follows from the argument detailed in in exp-util and simple substitution.
-Further,  the maximizer is $\delta_{2}^* = (\Theta - \theta') / (1+a)$, which moves the share vector to $\theta' + \delta_{2}^* = \frac{1}{1+a} \Theta + \frac{a}{1+a} \theta'$ which is a convex combination of the effective belief and the current market state.
-
-In other words, an exponential utility maximizing trader who has belief $\hat{\theta}$ with prior exposure $\delta$ in a market  will behave identically as an exponential utility maximizing trader with belief $\hat{\theta}-a \delta$ and no prior exposure in the market.
+This follows from Proposition \ref{prop:exp-util}.
+The maximizer is $\delta_{2}^* = (\Theta - \theta') / (1+a)$, which moves the share vector to $\theta' + \delta_{2}^* = \frac{1}{1+a} \Theta + \frac{a}{1+a} \theta'$ which is a convex combination of the effective belief and the current market state.
+
+In other words, an exponential utility maximizing trader who has belief $\hat{\theta}$ with prior exposure $\delta$ in a market  will behave identically to an exponential utility maximizing trader with belief $\hat{\theta}-a \delta$ and no prior exposure in the market.
 Here $a$ is the utility parameter.
 This means that financial exposure can be equivalently understood as changing the privately held beliefs.
-
-Note that as the trader becomes more risk averse, the  belief is updated more aggressively.
-As $a \rightarrow 0$, the trader becomes more risk neutral and his effective belief is closer to his true belief.
-%
-%----------------------------------------------------------------------------------------
-%	BIBLIOGRAPHY
-%----------------------------------------------------------------------------------------
-
-\subsection{Equilibrium in a market with multiple traders}
+\pagebreak
+
+\section{Equilibrium Market State for Exponential Utility Agents}
 We have shown that every exponential-utility maximizing trader picks the share vector $\delta_{i}$ so as to maximize the utility of $- C(\delta_{i} + \theta) + C(\theta) - \frac{1}{a_{i}} T(\beta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\beta_{i})$ where $\theta$ is the initial market state, $a_{i}$ the utility parameter and $\beta_{i}$ the belief parameter of  trader $i$.
@@
 \section{Conclusions}
+%
+%----------------------------------------------------------------------------------------
+%	BIBLIOGRAPHY
+%----------------------------------------------------------------------------------------
 

### 2014-02-04 WriteUp/maxentpredmkts.tex
cleaned up section on equilibrium state

@@
 For instance, we analyze market behavior with budget limited traders, traders that are exponential-utility-maximizers.
-\section{Notation and Definitions}
-convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesian updates, conjugate prior.
+%\section{Notation and Definitions}
+%convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesian updates, conjugate prior.
 
@@
 
-\subsubsection{Exponential Utility}
+\section{Exponential Utility}
 
@@
 So there is a kind of congruence between exponential family beliefs and exponential utility (as the names would suggest). 
-\subsection{Properties of the Maximum Entropy Market}
+\section{Properties of the Maximum Entropy Market}
 Consider the dual of the cost function, $C^{*}(\mu)$ defined as
@@
 
-\subsubsection{Example: Gaussian Markets}
+\subsection{Example: Gaussian Markets}
 {\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
@@
 \section{Equilibrium Market State for Exponential Utility Agents}
-We have shown that every exponential-utility maximizing trader picks the share vector $\delta_{i}$ so as to maximize the utility of $- C(\delta_{i} + \theta) + C(\theta) - \frac{1}{a_{i}} T(\beta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\beta_{i})$ where $\theta$ is the initial market state, $a_{i}$ the utility parameter and $\beta_{i}$ the belief parameter of  trader $i$.
-To maximize utility, it suffices to maximize the argument, assuming that $C()=T()=\psi()$.
-
-Now consider a market with multiple exponential-utility maximizing traders.
-We wish to compute the equilibrium state.
-Let us assume that equilibrium is reached at a market state $\theta+\sum_{i}\delta_{i}$ where $\theta$ is the initial market state before trading begins and $\delta_{i}$ is each trader's purchase.
-
-%Thus, each trader picks $\delta_{i}$ as follows.
-%\begin{eqnarray*}
-%&&\arg\max_{\delta_{i}}\left(- \psi(\theta+\sum_{i}\delta_{i}) + \psi(\theta) - \frac{1}{a_{i}} \psi(\beta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} \psi(\beta_{i})\right)\\
-%&\iff&\arg\min_{\delta_{i}}\left( \psi(\theta+\sum_{i}\delta_{i})+\sum_{i}\frac{1}{a_{i}}\psi(\beta_{i}-a_{i}\delta_{i})\right)
-%\end{eqnarray*}¥
-Recall the following result from game theory.
+We have shown that every exponential-utility maximizing trader picks the share vector $\delta$ so that the eventual market state can be represented as a convex combination of the initial market state and the natural parameter of his (exponential family) belief distribution.
+In this section we will compute the equilibrium state in an exponential family market with multiple exponential-utility maximizing traders.
+
+Recall the following result from game theory. %need a reference/name for this result?
 \begin{thm}\label{thm:potl}
 Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}\defeq \delta_{1},\ldots,\delta_{i},\ldots,\delta_{n}$.
-If there exists a potential function $\phi(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=\phi(\vec{\delta})-\phi(\vec{\delta}_{-i},\delta_{i}')$$
-then when $\phi(\vec{\delta})$ is maximized, $\vec{\delta}$ is an equilibrium.
+If there exists a potential function $g(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=g(\vec{\delta})-g(\vec{\delta}_{-i},\delta_{i}')$$
+then when $g(\vec{\delta})$ is maximized, $\vec{\delta}$ is an equilibrium.
 \end{thm}¥
-
-Let $$\phi(\vec{\delta})\defeq  \psi(\theta+\sum_{i}\delta_{i})+\sum_{i}\frac{1}{a_{i}}\psi(\beta_{i}-a_{i}\delta_{i})$$
-Now $U_{i}(\vec{\delta})=- \psi(\theta+\sum_{j}\delta_{j}) + \psi(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} \psi(\beta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} \psi(\beta_{i})$.
-Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $\phi(\vec{\delta})$ for each $\delta_{i}$.
-\begin{eqnarray*}
-\nabla_{\delta_{i}}\phi(\vec{\delta})&=&\nabla\psi(\theta+\sum_{j=1}^{n}\delta_{i})-\nabla\psi(\beta_{i}-a_{i}\delta_{i})\\
+In the exponential family market, the cost function $C$ is identical to the log partition function $T$.
+Let $\vec{\delta}$ be the vector of vectors of shares purchased by every trader in the market when the market has reached equilibrium, $\theta$ the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his utility parameter.
+
+Define a potential function as  $$g(\vec{\delta})\defeq  T(\theta+\sum_{i}\delta_{i})+\sum_{i}\frac{1}{a_{i}}T(\htheta_{i}-a_{i}\delta_{i})$$
+
+%is the following expression for utility correct?
+Now the utility of trader $i$ is $U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i})$.
+Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $g(\vec{\delta})$ for each $\delta_{i}$.
+\begin{eqnarray*}
+\nabla_{\delta_{i}}g(\vec{\delta})&=&\nabla T(\theta+\sum_{j=1}^{n}\delta_{i})-\nabla T(\htheta_{i}-a_{i}\delta_{i})\\
 &=&0
@@
 \begin{equation}\label{eq:eqbm}
-\beta_{i}-a_{i}\delta_{i}=\theta+\sum_{j=1}^{n}\delta_{j}
+\htheta_{i}-a_{i}\delta_{i}=\theta+\sum_{j=1}^{n}\delta_{j}
 \end{equation}¥
 Rewriting, we have for each trader $i$, 
-$$\frac{\beta_{i}}{a_{i}}-\delta_{i}=\frac{1}{a_{i}}(\theta+\sum_{j=1}^{n}\delta_{j})$$
+$$\frac{\htheta_{i}}{a_{i}}-\delta_{i}=\frac{1}{a_{i}}(\theta+\sum_{j=1}^{n}\delta_{j})$$
 Thus,
-$$\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)-\sum_{i=1}^{n}\delta_{i}=\left(\theta+\sum_{j=1}^{n}\delta_{j}\right)\sum_{i=1}^{n}\frac{1}{a_{i}}$$
+$$\sum_{i=1}^{n}\left(\frac{\htheta_{i}}{a_{i}}\right)-\sum_{i=1}^{n}\delta_{i}=\left(\theta+\sum_{j=1}^{n}\delta_{j}\right)\sum_{i=1}^{n}\frac{1}{a_{i}}$$
 And
 $$\sum_{j=1}^{n}\delta_{j}=
-\frac{\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)-\theta\sum_{i=1}^{n}\left(\frac{1}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
+\frac{\sum_{i=1}^{n}\left(\frac{\htheta_{i}}{a_{i}}\right)-\theta\sum_{i=1}^{n}\left(\frac{1}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
 Substituting in Equation \ref{eq:eqbm} we have the following expression for the final market state.
 $$\theta+\sum_{j=1}^{n}\delta_{j}=
-\frac{\theta+\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
-
+\frac{\theta+\sum_{i=1}^{n}\left(\frac{\htheta_{i}}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
+Thus the equilibrium state is a convex combination of trader beliefs and initial market state.
 \section{Conclusions}

### 2014-02-04 WriteUp/maxentpredmkts.tex
minor edits
rearranged some sections

@@
 %-literature review
-%- exp family review [Sindhu; Tuesday]
 %- maxent scoring rule [Sebastien]
 %	- generalized lmsr
+%	- exp family review
 %	[- does modularity (a la Hanson) still hold?]
@@
 %- exp-util [Sebastien]
-%- equilibrium for repeated trades [Sindhu; Tuesday]
+%- equilibrium for repeated trades [Sindhu; Tuesday] done
 % 	[- rate of convergence?]
-%- budgets and adversarial markets [Sindhu; Monday]
+%- budgets and adversarial markets [Sindhu; Monday] done
 % [any liquidity results?] [Sindhu: weekend]
@@
 
+%%%%%%%%%%%%
+%This bit needs editing
+%%%%%%%%%%%%
+\subsection{Generalized Log Scoring Rule}
+Suppose that an agent holds a belief in the expected value of the moments of a random variable $x$.
+The maximum entropy distribution that is consistent with these beliefs is an exponential family distribution.
+ \defn For an outcome $x$ with predicted probability $p$, the generalized log scoring rule is defined as $$S(p,x)\defeq a+b\log p(x)$$ where $a$ and $b$ are constants that we will set to $0$ and $1$ respectively in these notes.\undefn
+
+\begin{thm} The generalized log scoring rule for the maximum entropy distribution is proper.
+\end{thm}
+\proof
+Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$
+
+Note that the maximum entropy distribution is an exponential family distribution. 
+ Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.
+
+ We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
+  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
+   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
+   Or 
+\begin{equation}\label{eq:proper}
+\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
+\end{equation}
+
+   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
+   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
+   Rearranging the terms,
+   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
+   which is the first order condition on the convexity of $\psi(\cdot)$.
+ \unproof
+
+ \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters.
+That is,
+   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
+   \unrmk
+
+\subsection{Exponential Smoothing}
+One way to model the exponential family prediction market might be through exponential smoothing.
+\begin{itemize}
+\item A \emph{time series} is a set of data points sampled at periodic instances.
+\item \emph{Smoothing} is the process of creating an approximation function based on time series that while capturing  patterns in the data is relatively insensitive to noise.
+\item Natural first step is to use moving average. But average on how many samples (say $k$)? What to do until the first $k$ samples have been received?
+\item
+Next attempt might be to use weighted moving average. Advantage is that you can give higher weight to more recent terms. But still same disadvantage as simple moving average technique.
+\item
+\emph{Exponential Smoothing} attempts to remove this disadvantage.
+\end{itemize}
+
+Exponential smoothing is defined iteratively for $\alpha\in(0,1)$ as
+\begin{eqnarray*}
+s_{1} &=& x_{0}\\
+s_{t}&=&\alpha x_{t-1}+(1-\alpha)s_{t-1}
+\end{eqnarray*}
+
+Smaller values of $\alpha$ means greater smoothing e.g., $\alpha=0$ gives a constant function ($=x_{0}$).
+Note that the selection of $x_{0}$ is important -- and increases in importance as $\alpha$.
+Some advantages over moving average: all data points matter (decreasing importance with time) and computationally only last data point need be stored.
 \pagebreak
@@
 Since~(\ref{scoring}) is a proper scoring rule, the information market with cost function~(\ref{cost}) is based on a proper market scoring rule.
+\subsection{Properties of the Maximum Entropy Market}
+Consider the dual of the cost function, $C^{*}(\mu)$ defined as
+$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
+This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
+$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
+Thus, the mean parameters are the dual variables to the natural parameters.
+
+\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
+\proof
+To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
+\begin{eqnarray*}
+-H(p) &=& \int_{x}p(x)\log p(x)\\
+&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
+&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
+&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
+\end{eqnarray*}
+\unproof
+\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
+\rmk The LMSR is essentially a special case applied to a multinomial distribution.
+The LMSR is known to have bounded ($\log n$) market maker loss.
+Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
+\unrmk
+
+{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
+\proof
+Note that, for an exponential family distribution
+\begin{eqnarray*}
+H(p) &=& -\int_{x}p(x)\log p(x)dx\\
+&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
+&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
+\end{eqnarray*}
+If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
+\unproof
+
+%{\cor The exponential family market maker has unbounded worst case loss.}
+%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
+%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
+%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
+\paragraph{Loss of the market maker}
+First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
+
+Recall the definition of the conjugate dual:
+$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
+The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
+$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
+This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
+So we have 
+$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
+Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
+
+Let $q_{0}$ be the initial and $q_{f}$ the final market state.
+Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
+\begin{eqnarray*}
+\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
+&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
+&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
+&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
+&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
+&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
+&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
+&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
+&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
+\end{eqnarray*}
+
+\subsection{Example: Gaussian Markets}
+{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
+In this case,  $\beta$ is a $2$-dimensional  vector.
+Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
+So
+\begin{eqnarray*}
+\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
+&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
+&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
+\end{eqnarray*}
+Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
+$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
+$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
+{\ex
+We will now derive the expression for the differential entropy for the normal distribution.
+Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
+\begin{eqnarray*}
+H(p) &=& -\int_{x}p(x)\log p(x)dx\\
+&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
+&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
+&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
+&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
+&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
+\end{eqnarray*}
+Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
+Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
+
+{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
+Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
+Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
+Let's now work out the loss for the Gaussian market maker.
+First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
+\begin{eqnarray*}
+D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
+&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
+&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
+&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
+&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
+%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
+&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
+\end{eqnarray*}
+
+So the Gaussian market maker loss is 
+%parallels to quadratic market maker loss? 
+\begin{eqnarray*}
+D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
+&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
+&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
+\end{eqnarray*}¥
+Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
+\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
 \pagebreak
@@
 \section{The MaxEnt Market Mechanism: A Bayesian View}
-\subsubsection{Conjugate Priors}
+\subsection{Conjugate Priors}
 
@@
 where $g^*$ is the convex conjugate of $g$.
+\subsection{Bayesian Traders}
+Suppose data is drawn from an exponential family distribution.
+Further suppose, following a Bayesian model, that each trader has some prior belief distribution over the parameters of the data distribution.
+Assuming a conjugate prior, this prior distribution is also an exponential family distribution.
+
+The data distribution is given by $\exp\{\beta\cdot\phi(x)-\psi(\beta)\}$ where $\psi()$ is the log partition function and $\phi()$ are the sufficient statistics.
+Then the conjugate prior parametrized by $b_{0}=(n\nu, n)$ is given by $\exp\{b_{0}\cdot(\beta,\psi(\beta))-\Psi(b_{0})\}$where $\Psi()$ is the corresponding log partition function.
+The posterior distribution is given by $b=(n\nu+m\hmu, n+m)$ where $\hmu$ is the empirical mean of the sufficient statistics of the $m$ data points drawn from the data distribution, $\hmu=\sum_{i=1}^{m}\phi(x_{i})$.
+We also have $\E_{\beta\sim b_{0}}\E_{x\sim \beta}[\phi(x)]=\nu$.
+
+\subsection{Linear Utility}
+
+Suppose the trader wishes to maximize his expected payoff.
+Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by $$\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]$$
+But, $\E_{\beta\sim b}\E_{x\sim \beta}[\phi(x)]=\frac{n\nu+m\hmu}{ n+m}$.
+To obtain the maximum, we set the gradient to $0$.
+Thus, we have $$\nabla C(\delta + \theta)=\frac{n\nu+m\hmu}{ n+m}$$ For an exponential family prediction market, the final market state is given by
+$$\nabla C^{*}\left(\frac{n\nu+m\hmu}{ n+m}\right)$$
+
+\subsection{Exponential Utility}
+Now suppose the trader has exponential utility, given by $$U(w) = -\frac{1}{a} \exp(-aw)$$ where $a$ is the coefficient of risk aversion.
+Suppose the trader wishes to maximize his expected utility.
+Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by
+\begin{eqnarray*}
+&&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}U\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]\\
+&=&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}-\frac{1}{a}\exp\left[ -a\delta\phi(x) +a C(\delta + \theta) -a C(\theta)\right]
+\end{eqnarray*}¥
+
+\begin{eqnarray*}
+\E_{\beta\sim b}\E_{x\sim \beta}\exp\left[-a\delta\phi(x)\right]&=&\int_{\mathcal{B}}\int_{\mathcal{X}}\exp[-a\delta\phi(x)]\, \exp\{\beta\cdot\phi(x)-\psi(\beta)\}\, dx\\
+&&\qquad\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\, d\beta\\
+&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
+&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\int_{\mathcal{X}}\exp[(\beta-a\delta)\cdot\phi(x)-\psi(\beta-a\delta)]\, dx\\
+&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
+&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\\
+\end{eqnarray*}¥
 
@@
 So there is a kind of congruence between exponential family beliefs and exponential utility (as the names would suggest).
-\section{Properties of the Maximum Entropy Market}
-Consider the dual of the cost function, $C^{*}(\mu)$ defined as
-$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
-This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
-$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
-Thus, the mean parameters are the dual variables to the natural parameters.
-
-\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
-\proof
-To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
-\begin{eqnarray*}
--H(p) &=& \int_{x}p(x)\log p(x)\\
-&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
-&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
-&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
-\end{eqnarray*}
-\unproof
-\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
-\rmk The LMSR is essentially a special case applied to a multinomial distribution.
-The LMSR is known to have bounded ($\log n$) market maker loss.
-Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
-\unrmk
-
-{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
-\proof
-Note that, for an exponential family distribution
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
-&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
-\end{eqnarray*}
-If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
-\unproof
-
-%{\cor The exponential family market maker has unbounded worst case loss.}
-%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
-%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
-%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
-\paragraph{Loss of the market maker}
-First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
-
-Recall the definition of the conjugate dual:
-$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
-The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
-$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
-This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
-So we have 
-$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
-Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
-
-Let $q_{0}$ be the initial and $q_{f}$ the final market state.
-Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
-\begin{eqnarray*}
-\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
-&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
-&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
-&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
-&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
-&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
-&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
-&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
-&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
-\end{eqnarray*}
-
-\subsection{Example: Gaussian Markets}
-{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
-In this case,  $\beta$ is a $2$-dimensional  vector.
-Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
-So
-\begin{eqnarray*}
-\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
-&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
-&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
-\end{eqnarray*}
-Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
-$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
-$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
-{\ex
-We will now derive the expression for the differential entropy for the normal distribution.
-Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
-&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
-\end{eqnarray*}
-Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
-Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
-
-{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
-Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
-Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
-Let's now work out the loss for the Gaussian market maker.
-First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
-\begin{eqnarray*}
-D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
-&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
-&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
-%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
-&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
-\end{eqnarray*}
-
-So the Gaussian market maker loss is 
-%parallels to quadratic market maker loss? 
-\begin{eqnarray*}
-D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
-&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
-&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
-\end{eqnarray*}¥
-Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
-\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
-\section{Generalized Log Scoring Rule}
-Suppose that an agent holds a belief in the expected value of the moments of a random variable $x$.
-The maximum entropy distribution that is consistent with these beliefs is an exponential family distribution.
- \defn For an outcome $x$ with predicted probability $p$, the generalized log scoring rule is defined as $$S(p,x)\defeq a+b\log p(x)$$ where $a$ and $b$ are constants that we will set to $0$ and $1$ respectively in these notes.\undefn
-
-\begin{thm} The generalized log scoring rule for the maximum entropy distribution is proper.
-\end{thm}
-\proof
-Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$
-
-Note that the maximum entropy distribution is an exponential family distribution. 
- Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.
-
- We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
-  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
-   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
-   Or 
-\begin{equation}\label{eq:proper}
-\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
-\end{equation}
-
-   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
-   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
-   Rearranging the terms,
-   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
-   which is the first order condition on the convexity of $\psi(\cdot)$.
- \unproof
-
- \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters.
-That is,
-   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
-   \unrmk
-
- \subsection{Exponential Smoothing}
-One way to model the exponential family prediction market might be through exponential smoothing.
-\begin{itemize}
-\item A \emph{time series} is a set of data points sampled at periodic instances.
-\item \emph{Smoothing} is the process of creating an approximation function based on time series that while capturing  patterns in the data is relatively insensitive to noise.
-\item Natural first step is to use moving average. But average on how many samples (say $k$)? What to do until the first $k$ samples have been received?
-\item
-Next attempt might be to use weighted moving average. Advantage is that you can give higher weight to more recent terms. But still same disadvantage as simple moving average technique.
-\item
-\emph{Exponential Smoothing} attempts to remove this disadvantage.
-\end{itemize}
-
-Exponential smoothing is defined iteratively for $\alpha\in(0,1)$ as
-\begin{eqnarray*}
-s_{1} &=& x_{0}\\
-s_{t}&=&\alpha x_{t-1}+(1-\alpha)s_{t-1}
-\end{eqnarray*}
-
-Smaller values of $\alpha$ means greater smoothing e.g., $\alpha=0$ gives a constant function ($=x_{0}$).
-Note that the selection of $x_{0}$ is important -- and increases in importance as $\alpha$.
-Some advantages over moving average: all data points matter (decreasing importance with time) and computationally only last data point need be stored.
-
-\section{Bayesian Traders}
-Suppose data is drawn from an exponential family distribution.
-Further suppose, following a Bayesian model, that each trader has some prior belief distribution over the parameters of the data distribution.
-Assuming a conjugate prior, this prior distribution is also an exponential family distribution.
-
-The data distribution is given by $\exp\{\beta\cdot\phi(x)-\psi(\beta)\}$ where $\psi()$ is the log partition function and $\phi()$ are the sufficient statistics.
-Then the conjugate prior parametrized by $b_{0}=(n\nu, n)$ is given by $\exp\{b_{0}\cdot(\beta,\psi(\beta))-\Psi(b_{0})\}$where $\Psi()$ is the corresponding log partition function.
-The posterior distribution is given by $b=(n\nu+m\hmu, n+m)$ where $\hmu$ is the empirical mean of the sufficient statistics of the $m$ data points drawn from the data distribution, $\hmu=\sum_{i=1}^{m}\phi(x_{i})$.
-We also have $\E_{\beta\sim b_{0}}\E_{x\sim \beta}[\phi(x)]=\nu$.
-
-\subsection{Linear Utility}
-
-Suppose the trader wishes to maximize his expected payoff.
-Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by $$\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]$$
-But, $\E_{\beta\sim b}\E_{x\sim \beta}[\phi(x)]=\frac{n\nu+m\hmu}{ n+m}$.
-To obtain the maximum, we set the gradient to $0$.
-Thus, we have $$\nabla C(\delta + \theta)=\frac{n\nu+m\hmu}{ n+m}$$ For an exponential family prediction market, the final market state is given by
-$$\nabla C^{*}\left(\frac{n\nu+m\hmu}{ n+m}\right)$$
-
-\subsection{Exponential Utility}
-Now suppose the trader has exponential utility, given by $$U(w) = -\frac{1}{a} \exp(-aw)$$ where $a$ is the coefficient of risk aversion.
-Suppose the trader wishes to maximize his expected utility.
-Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by
-\begin{eqnarray*}
-&&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}U\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]\\
-&=&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}-\frac{1}{a}\exp\left[ -a\delta\phi(x) +a C(\delta + \theta) -a C(\theta)\right]
-\end{eqnarray*}¥
-
-\begin{eqnarray*}
-\E_{\beta\sim b}\E_{x\sim \beta}\exp\left[-a\delta\phi(x)\right]&=&\int_{\mathcal{B}}\int_{\mathcal{X}}\exp[-a\delta\phi(x)]\, \exp\{\beta\cdot\phi(x)-\psi(\beta)\}\, dx\\
-&&\qquad\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\, d\beta\\
-&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
-&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\int_{\mathcal{X}}\exp[(\beta-a\delta)\cdot\phi(x)-\psi(\beta-a\delta)]\, dx\\
-&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
-&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\\
-\end{eqnarray*}¥
+
 \pagebreak

### 2014-02-05 WriteUp/maxentpredmkts.tex
minor edits

@@
 
+ We note that the expected difference between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters.
+That is,
+   $$\E_{x\sim p}[S(p,x)]- \E_{x\sim p}[S(p',x)] = D_{T}(\theta,\theta')$$
+
 An advantage of this generalization of the log score is that, for many statistics of interest, it leads to simple \emph{closed-form} formulas for the scoring rule.
@@
 
-%%%%%%%%%%%%
-%This bit needs editing
-%%%%%%%%%%%%
-\subsection{Generalized Log Scoring Rule}
-Suppose that an agent holds a belief in the expected value of the moments of a random variable $x$.
-The maximum entropy distribution that is consistent with these beliefs is an exponential family distribution.
- \defn For an outcome $x$ with predicted probability $p$, the generalized log scoring rule is defined as $$S(p,x)\defeq a+b\log p(x)$$ where $a$ and $b$ are constants that we will set to $0$ and $1$ respectively in these notes.\undefn
-
-\begin{thm} The generalized log scoring rule for the maximum entropy distribution is proper.
-\end{thm}
-\proof
-Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$
-
-Note that the maximum entropy distribution is an exponential family distribution. 
- Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.
-
- We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
-  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
-   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
-   Or 
-\begin{equation}\label{eq:proper}
-\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
-\end{equation}
-
-   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
-   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
-   Rearranging the terms,
-   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
-   which is the first order condition on the convexity of $\psi(\cdot)$.
- \unproof
-
- \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters.
-That is,
-   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
-   \unrmk
-
-\subsection{Exponential Smoothing}
-One way to model the exponential family prediction market might be through exponential smoothing.
-\begin{itemize}
-\item A \emph{time series} is a set of data points sampled at periodic instances.
-\item \emph{Smoothing} is the process of creating an approximation function based on time series that while capturing  patterns in the data is relatively insensitive to noise.
-\item Natural first step is to use moving average. But average on how many samples (say $k$)? What to do until the first $k$ samples have been received?
-\item
-Next attempt might be to use weighted moving average. Advantage is that you can give higher weight to more recent terms. But still same disadvantage as simple moving average technique.
-\item
-\emph{Exponential Smoothing} attempts to remove this disadvantage.
-\end{itemize}
-
-Exponential smoothing is defined iteratively for $\alpha\in(0,1)$ as
-\begin{eqnarray*}
-s_{1} &=& x_{0}\\
-s_{t}&=&\alpha x_{t-1}+(1-\alpha)s_{t-1}
-\end{eqnarray*}
-
-Smaller values of $\alpha$ means greater smoothing e.g., $\alpha=0$ gives a constant function ($=x_{0}$).
-Note that the selection of $x_{0}$ is important -- and increases in importance as $\alpha$.
+%\subsection{Exponential Smoothing}
+%One way to model the exponential family prediction market might be through exponential smoothing.
+%\begin{itemize}
+%\item A \emph{time series} is a set of data points sampled at periodic instances.
+%\item \emph{Smoothing} is the process of creating an approximation function based on time series that while capturing  patterns in the data is relatively insensitive to noise.
+%\item Natural first step is to use moving average. But average on how many samples (say $k$)? What to do until the first $k$ samples have been received?
+%\item
+%Next attempt might be to use weighted moving average. Advantage is that you can give higher weight to more recent terms. But still same disadvantage as simple moving average technique.
+%\item
+%\emph{Exponential Smoothing} attempts to remove this disadvantage.
+%\end{itemize}
+%
+%Exponential smoothing is defined iteratively for $\alpha\in(0,1)$ as
+%\begin{eqnarray*}
+%s_{1} &=& x_{0}\\
+%s_{t}&=&\alpha x_{t-1}+(1-\alpha)s_{t-1}
+%\end{eqnarray*}
+%
+%Smaller values of $\alpha$ means greater smoothing e.g., $\alpha=0$ gives a constant function ($=x_{0}$).
+%Note that the selection of $x_{0}$ is important -- and increases in importance as $\alpha$.
 Some advantages over moving average: all data points matter (decreasing importance with time) and computationally only last data point need be stored.
@@
 Since~(\ref{scoring}) is a proper scoring rule, the information market with cost function~(\ref{cost}) is based on a proper market scoring rule.
-\subsection{Properties of the Maximum Entropy Market}
-Consider the dual of the cost function, $C^{*}(\mu)$ defined as
-$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
-This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
-$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
-Thus, the mean parameters are the dual variables to the natural parameters.
-
-\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
-\proof
-To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
-\begin{eqnarray*}
--H(p) &=& \int_{x}p(x)\log p(x)\\
-&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
-&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
-&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
-\end{eqnarray*}
-\unproof
-\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
-\rmk The LMSR is essentially a special case applied to a multinomial distribution.
-The LMSR is known to have bounded ($\log n$) market maker loss.
-Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
-\unrmk
-
-{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
-\proof
-Note that, for an exponential family distribution
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
-&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
-\end{eqnarray*}
-If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
-\unproof
-
-%{\cor The exponential family market maker has unbounded worst case loss.}
-%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
-%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
-%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
-\paragraph{Loss of the market maker}
-First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
-
-Recall the definition of the conjugate dual:
-$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
-The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
-$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
-This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
-So we have 
-$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
-Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
-
-Let $q_{0}$ be the initial and $q_{f}$ the final market state.
-Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
-\begin{eqnarray*}
-\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
-&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
-&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
-&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
-&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
-&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
-&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
-&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
-&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
-\end{eqnarray*}
-
-\subsection{Example: Gaussian Markets}
-{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
-In this case,  $\beta$ is a $2$-dimensional  vector.
-Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
-So
-\begin{eqnarray*}
-\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
-&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
-&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
-\end{eqnarray*}
-Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
-$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
-$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
-{\ex
-We will now derive the expression for the differential entropy for the normal distribution.
-Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
-&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
-\end{eqnarray*}
-Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
-Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
-
-{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
-Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
-Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
-Let's now work out the loss for the Gaussian market maker.
-First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
-\begin{eqnarray*}
-D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
-&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
-&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
-%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
-&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
-\end{eqnarray*}
-
-So the Gaussian market maker loss is 
-%parallels to quadratic market maker loss? 
-\begin{eqnarray*}
-D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
-&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
-&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
-\end{eqnarray*}¥
-Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
-\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
-\pagebreak
-
-\section{The Exponential Family Market Mechanism with Budgets}
-In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
-
-In this section, we consider the the prediction market setup with traders that may be either informative or malicious.
-The malicious traders may want to inject faulty information into the market.
-The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.
-
-We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market.
-Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
-
-In this section, we assume that the traders have exponential family beliefs.
-The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
-\subsection{Budget-limited Aggregation}
-Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
-We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market.
-This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader.
-Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.
-
-In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
-Let $\alpha$ be the budget of a trader in the market.
-Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
-Let $C$ be the cost function.
-Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
-In this case, we want to budget-limit the trader's influence on the market state.
-
-We define the budget-limited final market state as $\theta'$. 
-Here, we consider a specific functional form of $\theta'$: $$\theta'=\lambda\htheta+(1-\lambda)\theta$$  where $$\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$$
-
-First, we  show that this trade is feasible given the trader's budget:
-\begin{theorem}
-Let the current market state be given by $\theta$.
-Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
-The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$.
-\end{theorem}
-\begin{proof}
-From the convexity of $C$, we have
- $$C(\theta')\leq(1-\lambda)C(\theta)+\lambda C(\htheta)$$
-
-Now 
-\begin{eqnarray*}
-C(\theta') - C(\theta) &\leq& (1-\lambda)C(\theta)\\
-&&+\lambda C(\htheta) - C(\theta)\\
-&=& \lambda\ (C(\htheta) - C(\theta))
-\end{eqnarray*}
-
- Thus, $C(\theta') - C(\theta) \leq \alpha$. 
- \end{proof}
-
- We note that moving to $\theta'$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
-In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\theta$.
-Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\htheta$.
-\subsection{Damage Bound}
-In this section, we will quantify the error in prediction that the market maker might have to endure as a result of malicious entities entering the market.
-We assume that these malicious entities trade in multiple instances of the market; thus the exposure of the market maker is over several \emph{rounds}.
-We use the standard log loss to measure this error in terms of the initial budget of traders.
-
-We define the loss function for $\theta$ shares held:
-$$L(\theta,x)=-\log(P_{\theta}(x))=\log\int \exp\{\theta^{T}\phi(x)\}\ dx-\theta^{T}\phi(x)=C(\theta)-\theta^{T}\phi(x)$$
-%with the correspondence between $\beta$ and $\theta$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
-
-Suppose the prediction market runs over multiple rounds $t$.
-Let $\theta_{0}^{t}$ be the initial number of shares of each security that are held.
-Let $\hat{\theta}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
-Let us assume that at this point the outcome is revealed; that is, we receive the value of the random variable $x^{t}$.
-
-Over multiple instances of the prediction market, we can track the change in budget of each trader.
-Let the budget at rounds $t$ and $t-1$ be $\alpha^{t}$ and $\alpha^{t-1}$ respectively.
-The change in budget for the trader is
-\begin{eqnarray*}
-\alpha^{t}-\alpha^{t-1}&=&C(\theta^{t})-C(\theta'^{t})-(\theta^{t}-\theta'^{t})^{T}\phi(x^{t})\\
-&=&L(\theta,x^{t})-L(\theta',x^{t})
-\end{eqnarray*}
-
-Define the myopic impact of a trader $i$ in segment $t$ as
-$$\Delta_{i}^{t}:=L(\hat{\theta}_{i-1}^{t},x^{t})-L(\hat{\theta}_{i}^{t},x^{t})$$
-Thus, the myopic impact captures incremental gain in prediction due to the trader in a round.
-Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
-
-The total myopic impact due to all $k$ active traders is given by
-$$\Delta^{t}=L(\theta_{0}^{t},x^{t})-L(\hat{\theta}_{k}^{t},x^{t})$$
-Thus $-\Delta^{t}$ captures the incremental loss of the market prediction after aggregation of all $k$ traders.
-
-\begin{theorem}
-A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
-\end{theorem}
-\begin{proof}
-Consider the myopic impact of a single trader $i$ after participating in the market $T$ times.
-Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
-$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
-
-Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
-\end{proof}
-
-This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders. %imp
-
-In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation.
-We provide an information-theoretic justification for this claim in Section \ref{sec:infth}.
-The intuition behind this claim is that his prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
-
-\subsubsection{An Information-Theoretic Interpretation}\label{sec:infth}
+\subsection{An Information-Theoretic Interpretation}\label{sec:infth}
  We observe a useful alternative view of the market 
@@
 (and outcome distribution $\pi$), while the payoff is independent of the initial market state in which these securities were purchased.
-
-\subsubsection{Budget of Informative Traders}\label{sec:inf}
+\subsection{Properties of the Maximum Entropy Market}
+Consider the dual of the cost function, $C^{*}(\mu)$ defined as
+$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
+This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
+$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
+Thus, the mean parameters are the dual variables to the natural parameters.
+
+\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
+\proof
+To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
+\begin{eqnarray*}
+-H(p) &=& \int_{x}p(x)\log p(x)\\
+&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
+&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
+&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
+\end{eqnarray*}
+\unproof
+\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
+\rmk The LMSR is essentially a special case applied to a multinomial distribution.
+The LMSR is known to have bounded ($\log n$) market maker loss.
+Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
+\unrmk
+
+{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
+\proof
+Note that, for an exponential family distribution
+\begin{eqnarray*}
+H(p) &=& -\int_{x}p(x)\log p(x)dx\\
+&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
+&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
+\end{eqnarray*}
+If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
+\unproof
+
+%{\cor The exponential family market maker has unbounded worst case loss.}
+%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
+%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
+%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
+\paragraph{Loss of the market maker}
+First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
+
+Recall the definition of the conjugate dual:
+$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
+The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
+$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
+This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
+So we have 
+$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
+Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
+
+Let $q_{0}$ be the initial and $q_{f}$ the final market state.
+Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
+\begin{eqnarray*}
+\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
+&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
+&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
+&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
+&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
+&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
+&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
+&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
+&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
+\end{eqnarray*}
+
+\subsection{Example: Gaussian Markets}
+{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
+In this case,  $\beta$ is a $2$-dimensional  vector.
+Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
+So
+\begin{eqnarray*}
+\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
+&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
+&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
+\end{eqnarray*}
+Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
+$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
+$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
+{\ex
+We will now derive the expression for the differential entropy for the normal distribution.
+Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
+\begin{eqnarray*}
+H(p) &=& -\int_{x}p(x)\log p(x)dx\\
+&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
+&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
+&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
+&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
+&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
+\end{eqnarray*}
+Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
+Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
+
+{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
+Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
+Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
+Let's now work out the loss for the Gaussian market maker.
+First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
+\begin{eqnarray*}
+D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
+&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
+&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
+&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
+&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
+%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
+&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
+\end{eqnarray*}
+
+So the Gaussian market maker loss is 
+%parallels to quadratic market maker loss? 
+\begin{eqnarray*}
+D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
+&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
+&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
+\end{eqnarray*}¥
+Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
+\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
+\pagebreak
+
+\section{The Exponential Family Market Mechanism with Budgets}
+In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
+
+In this section, we consider the the prediction market setup with traders that may be either informative or malicious.
+The malicious traders may want to inject faulty information into the market.
+The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.
+
+We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market.
+Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
+
+In this section, we assume that the traders have exponential family beliefs.
+The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
+\subsection{Budget-limited Aggregation}
+Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
+We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market.
+This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader.
+Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.
+
+In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
+Let $\alpha$ be the budget of a trader in the market.
+Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
+Let $C$ be the cost function.
+Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
+In this case, we want to budget-limit the trader's influence on the market state.
+
+We define the budget-limited final market state as $\theta'$. 
+Here, we consider a specific functional form of $\theta'$: $$\theta'=\lambda\htheta+(1-\lambda)\theta$$  where $$\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$$
+
+First, we  show that this trade is feasible given the trader's budget:
+\begin{theorem}
+Let the current market state be given by $\theta$.
+Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
+The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$.
+\end{theorem}
+\begin{proof}
+From the convexity of $C$, we have
+ $$C(\theta')\leq(1-\lambda)C(\theta)+\lambda C(\htheta)$$
+
+Now 
+\begin{eqnarray*}
+C(\theta') - C(\theta) &\leq& (1-\lambda)C(\theta)\\
+&&+\lambda C(\htheta) - C(\theta)\\
+&=& \lambda\ (C(\htheta) - C(\theta))
+\end{eqnarray*}
+
+ Thus, $C(\theta') - C(\theta) \leq \alpha$. 
+ \end{proof}
+
+ We note that moving to $\theta'$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
+In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\theta$.
+Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\htheta$.
+\subsection{Damage Bound}
+In this section, we will quantify the error in prediction that the market maker might have to endure as a result of malicious entities entering the market.
+We assume that these malicious entities trade in multiple instances of the market; thus the exposure of the market maker is over several \emph{rounds}.
+We use the standard log loss to measure this error in terms of the initial budget of traders.
+
+We define the loss function for $\theta$ shares held:
+$$L(\theta,x)=-\log(P_{\theta}(x))=\log\int \exp\{\theta^{T}\phi(x)\}\ dx-\theta^{T}\phi(x)=C(\theta)-\theta^{T}\phi(x)$$
+%with the correspondence between $\beta$ and $\theta$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
+
+Suppose the prediction market runs over multiple rounds $t$.
+Let $\theta_{0}^{t}$ be the initial number of shares of each security that are held.
+Let $\hat{\theta}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
+Let us assume that at this point the outcome is revealed; that is, we receive the value of the random variable $x^{t}$.
+
+Over multiple instances of the prediction market, we can track the change in budget of each trader.
+Let the budget at rounds $t$ and $t-1$ be $\alpha^{t}$ and $\alpha^{t-1}$ respectively.
+The change in budget for the trader is
+\begin{eqnarray*}
+\alpha^{t}-\alpha^{t-1}&=&C(\theta^{t})-C(\theta'^{t})-(\theta^{t}-\theta'^{t})^{T}\phi(x^{t})\\
+&=&L(\theta,x^{t})-L(\theta',x^{t})
+\end{eqnarray*}
+
+Define the myopic impact of a trader $i$ in segment $t$ as
+$$\Delta_{i}^{t}:=L(\hat{\theta}_{i-1}^{t},x^{t})-L(\hat{\theta}_{i}^{t},x^{t})$$
+Thus, the myopic impact captures incremental gain in prediction due to the trader in a round.
+Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
+
+The total myopic impact due to all $k$ active traders is given by
+$$\Delta^{t}=L(\theta_{0}^{t},x^{t})-L(\hat{\theta}_{k}^{t},x^{t})$$
+Thus $-\Delta^{t}$ captures the incremental loss of the market prediction after aggregation of all $k$ traders.
+
+\begin{theorem}
+A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
+\end{theorem}
+\begin{proof}
+Consider the myopic impact of a single trader $i$ after participating in the market $T$ times.
+Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
+$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
+
+Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
+\end{proof}
+
+This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders. %imp
+
+In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation. 
+%We provide an information-theoretic justification for this claim in Section \ref{sec:infth}. 
+The intuition behind this  is that a trader's prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
+
+\subsection{Budget of Informative Traders}\label{sec:inf}
 Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
@@
 
-\section{The MaxEnt Market Mechanism: A Bayesian View}
-\subsection{Conjugate Priors}
+\pagebreak
+\section{Bayesian Traders with Linear Utility}
+%\subsection{Conjugate Priors}
 
@@
 
-\subsection{Exponential Utility}
-Now suppose the trader has exponential utility, given by $$U(w) = -\frac{1}{a} \exp(-aw)$$ where $a$ is the coefficient of risk aversion.
+%\subsection{Exponential Utility}
+%Now suppose the trader has exponential utility, given by $$U(w) = -\frac{1}{a} \exp(-aw)$$ where $a$ is the coefficient of risk aversion.
 Suppose the trader wishes to maximize his expected utility.
 Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by
-\begin{eqnarray*}
-&&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}U\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]\\
-&=&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}-\frac{1}{a}\exp\left[ -a\delta\phi(x) +a C(\delta + \theta) -a C(\theta)\right]
-\end{eqnarray*}¥
-
-\begin{eqnarray*}
-\E_{\beta\sim b}\E_{x\sim \beta}\exp\left[-a\delta\phi(x)\right]&=&\int_{\mathcal{B}}\int_{\mathcal{X}}\exp[-a\delta\phi(x)]\, \exp\{\beta\cdot\phi(x)-\psi(\beta)\}\, dx\\
-&&\qquad\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\, d\beta\\
-&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
-&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\int_{\mathcal{X}}\exp[(\beta-a\delta)\cdot\phi(x)-\psi(\beta-a\delta)]\, dx\\
-&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
-&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\\
-\end{eqnarray*}¥
+%\begin{eqnarray*}
+%&&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}U\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]\\
+%&=&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}-\frac{1}{a}\exp\left[ -a\delta\phi(x) +a C(\delta + \theta) -a C(\theta)\right]
+%\end{eqnarray*}¥
+%
+%\begin{eqnarray*}
+%\E_{\beta\sim b}\E_{x\sim \beta}\exp\left[-a\delta\phi(x)\right]&=&\int_{\mathcal{B}}\int_{\mathcal{X}}\exp[-a\delta\phi(x)]\, \exp\{\beta\cdot\phi(x)-\psi(\beta)\}\, dx\\
+%&&\qquad\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\, d\beta\\
+%&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
+%&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\int_{\mathcal{X}}\exp[(\beta-a\delta)\cdot\phi(x)-\psi(\beta-a\delta)]\, dx\\
+%&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
+%&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\\
+%\end{eqnarray*}¥
+\pagebreak
 

### 2014-02-05 WriteUp/maxentpredmkts.tex
edited section on conjugate priors

@@
 (But it doesn't quite succeed.)
-
-\medskip\noindent
-Let $p(x;\theta)$ denote a probability density drawn from an exponential family with sufficient statistic $\phi : \mX \rightarrow \bR^d$, where $\theta$ is the natural parameter:
-$$
-p(x;\theta) = \exp \left[ \la \theta, \phi(x) \ra \right],
-$$
-and
-$$
-g(\theta) = \log \int_{\mX} \exp\la\phi(x),\theta\ra d\!x.
-$$
-Recall that $\grad g(\theta) = \Exp[\phi(x)]$ and $\grad^2 g(\theta) = \Var[\phi(x)]$. The family of conjugate priors is also an exponential family and takes the form
-$$
-p(\theta; n,\nu) = \exp\left[ \la n\nu,\theta \ra - ng(\theta) - h(\nu,n) \right].
-$$
-Here the feature map is $\psi(\theta) = (\theta, -g(\theta))$, the natural parameter is $(n\nu, n)$ where $n \in \bR$ and $\nu \in \bR^d$.
+In this section, we lay the foundation for interpreting trader behavior in terms of Bayesian updates.
+
+As before, we are interested in eliciting the sufficient statistics of the data.
+We assume that the outcome is drawn from an exponential family distribution; the prediction market is setup as before with the cost function corresponding to the log partition function and the payoff function corresponding to the sufficient statistics.
+Thus, the market state provides an estimate on the natural parameter of the distribution from which the outcome is drawn.
+Additionally, we assume that the market also makes public the total number of traders that have traded in the market.
+
+The goal is to aggregate information from risk neutral agents who have a belief distribution over the natural parameters.
+This prior distribution is updated by the agents based on the current market state, They also each have to access to the empirical mean of sufficient statistics based on a fixed number  $m$ of data points.
+Assuming a conjugate prior, both the prior and posterior belief distributions on the natural parameters are also an exponential family distributions.
+
+Now we will define the setup more formally.
+
+Let the data distribution be given by $p(x;\beta)=\exp\{\beta\cdot\phi(x)-T(\beta)\}$ where $T()$ is the log partition function and $\phi()$ are the sufficient statistics.
+Then the conjugate prior parametrized by $b_{0}=(n\nu, n)$ is given by $p(\beta;b_{0})=\exp\{b_{0}\cdot(\beta,\psi(\beta))-\psi(b_{0})\}$where $\psi()$ is the corresponding log partition function.
+The posterior distribution on the natural parameters is given by $p(\beta;b)$ where $b=(n\nu+m\hmu, n+m)$ where $\hmu$ is the empirical mean of the sufficient statistics of the $m$ data points drawn from the data distribution; that is, $\hmu=\sum_{i=1}^{m}\phi(x_{i})$.
+It turns out that $\E_{\beta\sim b_{0}}\E_{x\sim \beta}[\phi(x)]=\nu$; thus, it is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$.
+Thus the posterior mean is a convex combination of the prior and posterior means, and their relative weights depend on the phantom and empirical sample sizes.% need a reference here
+
+Suppose the current market state is $\theta$ and $i$ traders have traded in the market when trader $i+1$ with prior belief distribution $p(\beta;b^{i}_{0})$ enters the market.
+Here $b^{i}_{0} = (n\nu_{i}, n)$.
+This trader also has access to private information in the form of empirical sufficient statistics $\hmu_{i}$ from $m$ data points.
+Thus, he updates his belief as $p(\beta;b_{i})$ where $b=(m\hmu+mi \nabla T(\theta) + n\nu_{i}, n+mi+m)$.
+
+Suppose the trader wishes to maximize his expected payoff.
+Then the number of shares $\delta_{i}$ that he purchases when the current market state is $\theta$ is given by $$\arg\max_{\delta_{i}}\E_{\beta\sim b_{i}}\E_{x\sim \beta}\left[ \delta_{i}\phi(x) - T(\delta_{i} + \theta) + T(\theta)\right]$$
+But, $\E_{\beta\sim b}\E_{x\sim \beta}[\phi(x)]=\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}$.
+To obtain the maximum, we set the gradient of the above expression with respect to $\delta_{i}$ to $0$.
+Thus, we have for the optimal number of shares $\delta_{i}^{*}$ $$\nabla T(\delta_{i}^{*} + \theta)=\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}$$ For an exponential family prediction market, the final market state is given by
+$$\nabla T^{*}\left(\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}\right)$$ where $T^*$ is the convex conjugate of $T$.
+Thus, the final market state is a convex combination of prior, posterior and market means.
+
+%Let $p(x;\theta)$ denote a probability density drawn from an exponential family with sufficient statistic $\phi : \mX \rightarrow \bR^d$, where $\theta$ is the natural parameter:
+%$$
+%p(x;\theta) = \exp \left[ \la \theta, \phi(x) \ra \right],
+%$$
+%and
+%$$
+%g(\theta) = \log \int_{\mX} \exp\la\phi(x),\theta\ra d\!x.
+%$$
+%Recall that $\grad g(\theta) = \Exp[\phi(x)]$ and $\grad^2 g(\theta) = \Var[\phi(x)]$. The family of conjugate priors is also an exponential family and takes the form
+%$$
+%p(\theta; n,\nu) = \exp\left[ \la n\nu,\theta \ra - ng(\theta) - h(\nu,n) \right].
+%$$
+%Here the feature map is $\psi(\theta) = (\theta, -g(\theta))$, the natural parameter is $(n\nu, n)$ where $n \in \bR$ and $\nu \in \bR^d$.
 The normalizer $h(\nu, n)$ is convex in $(n\nu,n)$.
@@
 The justification for this is that
-$$
-\Exp_{\theta} \left[ \Exp_x \left[ \phi(x) | \theta \right]\right] 
-= \Exp_{\theta} \left[\grad g(\theta)\right]
-= \nu.
-$$
+%$$
+%\Exp_{\theta} \left[ \Exp_x \left[ \phi(x) | \theta \right]\right] 
+%= \Exp_{\theta} \left[\grad g(\theta)\right]
+%= \nu.
+%$$
 %(The proof is straightforward but not obvious. I have a reference for this but it's impossible to read---it's a mathematical statistics paper.)
 
-Suppose we draw a sample $X = (x_1,\ldots,x_m)$ of size $m$, and denote the empirical mean by $\mu[X] = \sum_{i=1}^m \phi(x_i)$. The posterior distribution is then
-$$
-p(\theta|X) \propto p(X|\theta)p(\theta|n,\nu) \propto \exp\left[ \la\mu[X]+n\nu,\theta\ra - (m+n)g(\theta) \right],
-$$
-and so the posterior mean is 
-\begin{equation} \label{posterior-mean}
-\frac{m\mu[X] + n\nu}{m+n}.
-\end{equation}
-Thus the posterior mean is a convex combination of the prior and posterior means, and their relative weights depend on the phantom and empirical sample sizes.
-
-In the context of prediction markets, one could imagine an agent who uses the current market estimate as its prior, and draws a empirical sample of size $m$.
+%Suppose we draw a sample $X = (x_1,\ldots,x_m)$ of size $m$, and denote the empirical mean by $\mu[X] = \sum_{i=1}^m \phi(x_i)$. The posterior distribution is then
+%$$
+%p(\theta|X) \propto p(X|\theta)p(\theta|n,\nu) \propto \exp\left[ \la\mu[X]+n\nu,\theta\ra - (m+n)g(\theta) \right],
+%$$
+%and so the posterior mean is 
+%\begin{equation} \label{posterior-mean}
+%\frac{m\mu[X] + n\nu}{m+n}.
+%\end{equation}
+
+%In the context of prediction markets, one could imagine an agent who uses the current market estimate as its prior, and draws a empirical sample of size $m$.
 Its belief then takes the form~(\ref{posterior-mean}), where $n$ depends on the importance the agent places on the market estimate (perhaps based on how long the market has been running).
@@
 The new natural parameter is
-$$
-\grad g^{-1} \left(\frac{m\mu[X] + n\nu}{m+n}\right) = \grad g^* \left(\frac{m\mu[X] + n\nu}{m+n}\right).
-$$
-where $g^*$ is the convex conjugate of $g$.
-\subsection{Bayesian Traders}
-Suppose data is drawn from an exponential family distribution.
-Further suppose, following a Bayesian model, that each trader has some prior belief distribution over the parameters of the data distribution.
-Assuming a conjugate prior, this prior distribution is also an exponential family distribution.
-
-The data distribution is given by $\exp\{\beta\cdot\phi(x)-\psi(\beta)\}$ where $\psi()$ is the log partition function and $\phi()$ are the sufficient statistics.
-Then the conjugate prior parametrized by $b_{0}=(n\nu, n)$ is given by $\exp\{b_{0}\cdot(\beta,\psi(\beta))-\Psi(b_{0})\}$where $\Psi()$ is the corresponding log partition function.
-The posterior distribution is given by $b=(n\nu+m\hmu, n+m)$ where $\hmu$ is the empirical mean of the sufficient statistics of the $m$ data points drawn from the data distribution, $\hmu=\sum_{i=1}^{m}\phi(x_{i})$.
-We also have $\E_{\beta\sim b_{0}}\E_{x\sim \beta}[\phi(x)]=\nu$.
-
-\subsection{Linear Utility}
-
-Suppose the trader wishes to maximize his expected payoff.
-Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by $$\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]$$
-But, $\E_{\beta\sim b}\E_{x\sim \beta}[\phi(x)]=\frac{n\nu+m\hmu}{ n+m}$.
-To obtain the maximum, we set the gradient to $0$.
-Thus, we have $$\nabla C(\delta + \theta)=\frac{n\nu+m\hmu}{ n+m}$$ For an exponential family prediction market, the final market state is given by
-$$\nabla C^{*}\left(\frac{n\nu+m\hmu}{ n+m}\right)$$
+%$$
+%\grad g^{-1} \left(\frac{m\mu[X] + n\nu}{m+n}\right) = \grad g^* \left(\frac{m\mu[X] + n\nu}{m+n}\right).
+%$$
+%where $g^*$ is the convex conjugate of $g$.
 

### 2014-02-05 WriteUp/maxentpredmkts.tex
some more edits to section on conj prior

@@
 %
-\begin{prop}
+\begin{prop}\label{prop:exp}
 Consider a regular exponential family with minimal sufficient statistic. The following properties hold:
@@
 The posterior distribution on the natural parameters is given by $p(\beta;b)$ where $b=(n\nu+m\hmu, n+m)$ where $\hmu$ is the empirical mean of the sufficient statistics of the $m$ data points drawn from the data distribution; that is, $\hmu=\sum_{i=1}^{m}\phi(x_{i})$.
-It turns out that $\E_{\beta\sim b_{0}}\E_{x\sim \beta}[\phi(x)]=\nu$; thus, it is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$.
+It turns out that
+\begin{equation}\label{eq:prior}
+\E_{\beta\sim b_{0}}\E_{x\sim \beta}[\phi(x)]=\nu
+\end{equation}¥ Thus, it is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$.
 Thus the posterior mean is a convex combination of the prior and posterior means, and their relative weights depend on the phantom and empirical sample sizes.% need a reference here
@@
 This trader also has access to private information in the form of empirical sufficient statistics $\hmu_{i}$ from $m$ data points.
+Recall from Proposition \ref{prop:exp} that natural parameter $\theta$ corresponds to expected statistics $\nabla T(\theta)$.
 Thus, he updates his belief as $p(\beta;b_{i})$ where $b=(m\hmu+mi \nabla T(\theta) + n\nu_{i}, n+mi+m)$.
@@
 Then the number of shares $\delta_{i}$ that he purchases when the current market state is $\theta$ is given by $$\arg\max_{\delta_{i}}\E_{\beta\sim b_{i}}\E_{x\sim \beta}\left[ \delta_{i}\phi(x) - T(\delta_{i} + \theta) + T(\theta)\right]$$
-But, $\E_{\beta\sim b}\E_{x\sim \beta}[\phi(x)]=\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}$.
+But, from Equation \ref{eq:prior} we have $\E_{\beta\sim b}\E_{x\sim \beta}[\phi(x)]=\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}$.
 To obtain the maximum, we set the gradient of the above expression with respect to $\delta_{i}$ to $0$.
-Thus, we have for the optimal number of shares $\delta_{i}^{*}$ $$\nabla T(\delta_{i}^{*} + \theta)=\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}$$ For an exponential family prediction market, the final market state is given by
-$$\nabla T^{*}\left(\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}\right)$$ where $T^*$ is the convex conjugate of $T$.
+Thus, we have for the optimal number of shares $\delta_{i}^{*}$ $$\nabla T(\delta_{i}^{*} + \theta)=\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}$$ Thus, from Proposition \ref{prop:exp} we have that for an exponential family prediction market, the final market state is given by
+$$\nabla G\left(\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}\right)$$ where $G$ is the convex conjugate of $T$.
 Thus, the final market state is a convex combination of prior, posterior and market means.

### 2014-02-06 WriteUp/maxentpredmkts.tex
Progress on information markets section

@@
 %
-\begin{eqnarray*}
-& & \Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\mu')] \\
-& = & \la \theta, \mu \ra - T(\theta) - \la \theta', \mu \ra + T(\theta') \\
-& = & T(\theta') - T(\theta) - \la \theta' - \theta, \mu \ra \\
-& = & T(\theta') - T(\theta) - \la \theta' - \theta, \grad T(\theta) \ra.
-\end{eqnarray*}
+\begin{eqnarray}
+& & \Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\mu')] \nonumber\\
+& = & \la \theta, \mu \ra - T(\theta) - \la \theta', \mu \ra + T(\theta') \nonumber\\
+& = & T(\theta') - T(\theta) - \la \theta' - \theta, \mu \ra \nonumber\\
+& = & T(\theta') - T(\theta) - \la \theta' - \theta, \grad T(\theta) \ra. \label{breg-div}
+\end{eqnarray}
 %
 The latter is positive by the strict convexity of $T$, which shows that the log score is proper.
-Now assume the defined log score is proper.
+For the converse, assume the defined log score is proper.
 By the Savage characterization of proper scoring rules for expectations (see~\cite{}), we must have
@@
 %
-This theorem leads to a straightforward procedure for constructing score rules for expectations: define the relevant statistic, and consider the maximum entropy (equivalently, exponential family) distribution consistent with the agent's reported mean $\mu$.
+As further intuition for the result, note that~(\ref{breg-div}) is the definition of the Bregman divergence with respect to strictly convex function $T$. Therefore we have 
+%
+\begin{equation*}
+\Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\mu')] = D_T(\theta',\theta) = D_G(\mu, \mu'),
+\end{equation*}
+%
+where the last equality is a well-known identity relating the Bregman divergences of $T$ and $T^* = G$~\citep{}.
+The equation states that the agent's regret from misreporting its mean parameter does not depend on the full density $p$, only the mean $\mu$.
+
+Theorem~\ref{maxent-score} leads to a straightforward procedure for constructing score rules for expectations: define the relevant statistic, and consider the maximum entropy (equivalently, exponential family) distribution consistent with the agent's reported mean $\mu$.
 The agent is compensated according to the log likelihood of the eventual outcome according to the latter.
@@
 
- We note that the expected difference between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters.
-That is,
-   $$\E_{x\sim p}[S(p,x)]- \E_{x\sim p}[S(p',x)] = D_{T}(\theta,\theta')$$
-
 An advantage of this generalization of the log score is that, for many statistics of interest, it leads to simple \emph{closed-form} formulas for the scoring rule.
@@
 %
-\begin{example}
+\begin{example} \label{ex:exponential}
 As base measure we take the Lebesgue restricted to $[0,+\infty)$, and we consider the statistic $\phi(x) = x$ so that we are simply eliciting the mean.
@@
 Here the statistics being elicited are the mean vector ${\boldsymbol \mu}$ and the covariance matrix ${\boldsymbol \Sigma}$.
-These scoring rules have been studied by Dawid and Sebastiani~\cite{} as rules that only depend on the mean and variance of the reported density.
+These scoring rules have been studied by~\citet{} as rules that only depend on the mean and variance of the reported density.
 They note that these rules are weakly proper (because they do not distinguish between densities with the same first and second moments), but do not make the point that knowledge of the full density is not necessary on the part of the agent.
@@
 Some advantages over moving average: all data points matter (decreasing importance with time) and computationally only last data point need be stored.
+
+\vfill
 \pagebreak
@@
 
-The purpose of a prediction market is to elicit and aggregate the beliefs (i.e., subjective probabilities) of agents over a space of \emph{outcomes}.
-In a single-agent setting, a scoring rule is used to elicit the agent's beliefs.
-In a multi-agent setting, an information market is used to aggregate the agent's beliefs.
-Hanson~\cite{} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both so that they can perform well in both thin and thick markets.
-In this work we derive a wide-ranging generalization of the logarithmic market scoring rule.
-Using a maximum entropy approach, we explain how a market scoring rule can be developed for outcome spaces both discrete and continuous, and for generic properties of the underlying distribution (e.g., mean and variance), rather than just the probabilities of individual outcomes.
+In a single-agent setting, a scoring rule is used to \emph{elicit} the agent's belief.
+In a multi-agent setting, an information market can be used to \emph{aggregate} the beliefs of the agents.
+In his seminal paper~\citet{} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both in order to perform well in thin or thick markets.
+In this section, we adapt the generalized log scoring rule to a market scoring rule which leads to markets with simple closed-form cost functions for many statistics of interest.
+The exponential families framework also allows us to move beyond risk-neutral agents to understand the aggregation properties of information markets under alternative behaviors, such as Bayesian updating and risk aversion.
 
@@
 
-We next consider how to indirectly elicit $\mu$ by setting up a market of contingent claim securities.
-Under this approach, the elements of the index set $\mS$ are interpreted not as statistics but as securities.
-The payoff from one share of security $s$ when outcome $x$ occurs is given by the mapping $\phi_s$.
-Thus if the vector of shares held by the agent is $\theta \in \bR^d$, where entry $\theta_s$ corresponds the number of shares of security $s$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \theta, \phi(x) \ra$.
-As a concrete example, recall our multinomial distribution example from the previous section, where $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise.
-This means that security $x \in \mS$ pays 1 dollar if outcome $x \in \mX$ occurs, and nothing otherwise.
-(Such securities are known as Arrow-Debreu securities.) In our normal distribution example, the statistic for the mean had the mapping $\phi_1(x) = x$.
-Therefore a share of the corresponding security has a payoff that is linear in the outcome.
-(Such securities amount to futures contracts.)
-
-To extract useful information from such a securities market, we set a pricing scheme and examine the number of shares the agent chooses to acquire.
-Let $\Omega$ be the set of all portfolios (i.e., vectors of shares) that the agent can feasibly hold; we will see how this domain is determined in an instant.
-Each security $s$ has an associated price function $c_s : \Omega \rightarrow \bR$, which gives the marginal price $c_s(\theta)$ of security $s$ when the agent holds portfolio $\theta$.
-(Note that the marginal price depends on the entire portfolio $\theta$, not just the number of shares $\theta_s$.) A risk-neutral agent will choose to acquire shares up to the point where, for each share, expected payoff equals marginal price.
-Formally, if the agent acquires portfolio $\theta$, then for each $s \in \mS$ we must have
-%
-\begin{equation} \label{market}
-\Exp_p[\phi_s(x)] = c_s(\theta).
-\end{equation}
-%
-In this way, by its choice of $\theta$, the agent reveals that its belief is $\mu = c(\theta)$.
-There are several important properties that the price function $c$ should have.
-To simplify the agent's portfolio acquisition process, it should not be the case that the total cost of acquiring $\theta$ depends on the order in which shares are bought.
-This means that there should exist a \emph{cost function} $C : \Omega \rightarrow \bR$ such that $c = \nabla C$.
-The gradient $\nabla C$ must be onto $\mM$ to ensure that~(\ref{market}) always has a solution---this is the most important but technical condition to realize.
-To ensure that the solution is unique, it would also be convenient if the cost function were strictly convex.
-%
-\begin{proposition}
-The following cost function is monotone, convex, and onto $\mM$: 
-%
-\begin{equation} \label{cost}
-C(\theta) = \log \int_{x \in \mX} \exp \left[\la \theta, \phi(x) \ra\right] \nu(dx).
-\end{equation}
-%
-\end{proposition}
-%
-\noindent
+In an information market an agent's expected belief $\mu$ is elicited indirectly through the purchase and sale of contingent claim securities.
+Under this approach, each component~$i$ of the statistic $\phi$ is interpreted as the payoff function of a security; that is, a single share of security $i$ pays off $\phi_i(x)$ when $x \in \mX$ occurs.
+Thus if the portfolio of shares held by the agent is $\delta \in \bR^d$, where entry $\delta_i$ corresponds to the number of shares of security $i$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \delta, \phi(x) \ra$.
+
+As a concrete example, recall that in the classic finite-outcome case the statistic has a component for each outcome $x$ such that $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise.
+Therefore the corresponding security pays 1 dollar if outcome $x$ occurs.
+(These are known as Arrow-Debreu securities.) In Example~\ref{ex:exponential} the one-dimensional statistic is $\phi(x) = x$, corresponding to a security whose payoff is linear in the outcome $x \in \bR_+$.
+(This amounts to a futures contract.)
+
+The standard way to implement an information market in the literature, due to~\citet{}, is via a centralized market maker.
+The market maker maintains a convex, differentiable cost function $C : \bR^d \rightarrow (-\infty,+\infty]$, where $C(\theta)$ records the revenue collected when the vector of outstanding shares is $\theta$.
+The cost to an agent of purchasing portfolio $\delta$ under a market state of $\theta$ is $C(\theta + \delta) - C(\theta)$, and therefore the instantaneous prices of the securities are given by the gradient $\grad C(\theta)$.
+
+A risk-neutral agent will choose to acquire shares up to the point where, for each share, expected payoff equals marginal price. Formally, if the agent acquires portfolio $\delta$, then we must have
+%
+\begin{equation} \label{risk-neutral-agent}
+\Exp_p[\phi(x)] = \grad C(\theta + \delta).
+\end{equation}
+%
+In this way, by its choice of $\delta$, the agent reveals that its expected belief is $\mu = \grad C(\theta + \delta)$.
+We stress that this observation relies on the assumptions that 1) the agent is risk-neutral, 2) the agent does not incorporate the market's information into its own beliefs, and 3) the agent is not budget constrained.
+
+\subsection{Information-Theoretic Interpretation}
+
+In the remainder of this paper we focus on the following cost function, which arises from the ``generalized'' logarithmic market scoring rule (LMSR):
+%
+\begin{equation} \label{lmsr-cost}
+C(\theta) = \log \int_{\mX} \exp \la \theta, \phi(x) \ra \,d\nu(x).
+\end{equation}
+%
+This is of course exactly the log-partition function~(\ref{log-part}) for the exponential family with sufficient statistic $\phi$, and we recover the classic LMSR using outcome indicator vectors as statistics.
+Because an agent would never select a portfolio with infinite cost, the effective domain (i.e., the possible vectors of outstanding shares) of $C$ is $\Theta = \{\theta \in \bR^d : C(\theta) < +\infty \}$.
+
 Observe that in the context of our earlier multinomial example, cost function~(\ref{cost}) is exactly the cost function for Hanson's logarithmic market scoring rule.
-Our approach here provides a generalization of this market scoring rule to markets with contingent claim securities with arbitrary payoffs, designed to elicit specific properties of distributions, beyond just the probabilities of different outcomes.
-
-Because an agent would never select a portfolio with infinite cost, the effective domain of $C$ is $\Omega = \{\theta \:|\: C(\theta) < +\infty \}$.
+
+.
 In the context of the multinomial distribution example, $\Omega = \bR^d$, so that shares of each security can always be bought or sold short.
@@
 
-Now, in our previous elicitation approach that used a scoring rule, the process of computing the scoring rule also provided a complete distribution $p(x;\mu)$ over outcomes, which could be used to infer other properties beyond just the agent beliefs $\mu$.
  In the current market-based approach, the agent's chosen portfolio $\theta$ can also form the basis of a distribution over outcomes.
-Consider the following distribution, represented as a density with respect to a base measure $\nu$:
-%
-\begin{equation} \label{expfam}
-p(x;\theta) = \exp\left[ \la \theta, \phi(x) \ra - C(\theta) \right].
-\end{equation}
-%
-Meaning, the probability that a subset $X \subseteq \mX$ of the outcomes occurs is $\int_{x \in X} p(x;\theta)\nu(dx)$.
-Observe that by definition~(\ref{cost}), this probability density indeed integrates to 1 over $\mX$, and it is clear that the density is non-negative as required.
-
-In the statistical literature a distribution that takes the form~(\ref{expfam}) is known as an \emph{exponential family}.
-The mapping $\phi$ is known as the \emph{sufficient statistic}, $\theta$ is the \emph{natural parameter}, and $C$ is the log-partition or \emph{cumulant} function.
-%This connection between information markets and exponential families allows us to draw on a vast statistical literature to understand the market's properties, and a growing computer science literature to apply efficient algorithms to compute costs and perform inference.
-
-%%%
-%%%
-%%%
-
-\subsection{Duality}
-
-There is a well-known duality between the maximum entropy approach and exponential families, which translates into a duality between the scoring rule and information market just developed.
-The duality implies that the approach leads to a \emph{market scoring rule}, applicable to both thin and thick multi-agent settings.
-
-It is known that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}.
-To see this, let $\theta(\mu) \in \bR^d$ be the Lagrange multiplier corresponding to constraints~(\ref{mean}) when solving the maximum entropy program given the agent report $\mu$.
-(Our choice of notation is deliberately suggestive.) Let $A(\mu)$ be the Lagrange multiplier corresponding to~(\ref{normalize}).
-From the first order necessary conditions for optimality, we find that
-%
-\[ p(x;\mu) = \exp \left[ \la \theta(\mu), \phi(x) \ra - A(\mu) - 1 \right], \]
-%
-so that the solution is indeed in exponential family form.
-Conversely, given a cumulant $C$ from an exponential family and a parameter $\theta$, we see that $p(x;\theta)$ defined according to~(\ref{expfam}) is the maximum entropy distribution under constraints $\Exp_p[\phi(x)] = \mu$ where $\mu = \nabla C(\theta)$.
-We obtain the following.
-%
-\begin{proposition}
-Let $C$ be the cumulant (i.e, cost function) for the exponential family corresponding to $\phi$, and let $\theta(\mu)$ be the optimal Lagrange multiplier for the mean constraints given an agent report of $\mu \in \mM$.
-Then the following scoring rule is equivalent to~(\ref{scoring}):
-%
-\begin{equation} \label{expscore}
-S(\mu, x) = \la \theta(\mu), \phi(x) \ra - C(\theta(\mu)).
-\end{equation}
-%
-\end{proposition}
-%
-\noindent
+
 The scoring rule~(\ref{expscore}) decomposes neatly into payoff and cost functions.
@@
 Since~(\ref{scoring}) is a proper scoring rule, the information market with cost function~(\ref{cost}) is based on a proper market scoring rule.
+
+%
+\begin{proposition}
+Let $C$ be the cumulant (i.e, cost function) for the exponential family corresponding to $\phi$, and let $\theta(\mu)$ be the optimal Lagrange multiplier for the mean constraints given an agent report of $\mu \in \mM$.
+Then the following scoring rule is equivalent to~(\ref{scoring}):
+%
+\begin{equation} \label{expscore}
+S(\mu, x) = \la \theta(\mu), \phi(x) \ra - C(\theta(\mu)).
+\end{equation}
+%
+\end{proposition}
+%
+
+\vfill
+\pagebreak
+
 \subsection{An Information-Theoretic Interpretation}\label{sec:infth}

### 2014-02-06 WriteUp/maxentpredmkts.tex
minor edits
Revert "Progress on information markets section"
This reverts commit 4b4cb0559fce0de653a6299aa65828367c5d65e4.
first draft of intro

--- 2014-02-05 WriteUp/maxentpredmkts.tex
@@
 % 	need a definition of term 'exponential family prediction market'
-%- conjugate prior market [Sindhu; Wednesday]
+%- conjugate prior market [Sindhu; Wednesday] done
 %	- also on hyperparam?
@@
 
+%- prediction market construction
+%	- exponential failies
+%	- scoring rules
+%	- information market
+%- Traders
+%	- aggregation process
+%		- bayesian traders
+%		- frequentist
+%	- utility
+%		- linear
+%		- exponential
+%	- budgets
+%- multiple trades
+%- multiple traders
+Prediction markets are aggregation mechanisms that allow market prices to be interpreted as predictive probabilities on an event.
+Each trader in the market is assumed to have some private information that he uses to make a prediction on the outcome of the event.
+The traders are assumed to be strategic in the sense that every trader wants to maximize his profit by participating in the market.
+Further, since the trades are done sequentially, the trader is allowed to observe all past trades in the market and update his private information based on this information.
+
+Traders are allowed to report their beliefs on the outcome of the event by allowing them to buy and sell securities whose value depends on the outcome of this future event.
+This will effect the state of the market, thus updating the predictive probabilities for the event.
+For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins.
+Traders can see the past history of trades, so the price at which a current trader is willing to buy and sell these securities can be interpreted as an aggregate ``market probability'' of a particular candidate winning the election.
+
+Currently, most deployed prediction markets have binary or finite discrete outcome spaces.
+This means that the outcome of every event can be characterized by a finite set of mutually exclusive exhaustive outcomes.
+The securities in these market are usually tied to each outcome so that buying (respectively, selling) a security related to an outcome, indicates a belief in an increased (respectively, decreased) probability of occurrence of the outcome.
+
+Directly applying techniques used for binary outcome spaces to large outcome spaces is computationally infeasible.
+In fact, it has been shown that computing and updating prices for combinatorial outcomes spaces is $\#$P-hard \cite{Chen08}.
+Infinite outcome spaces pose a particular challenge in prediction markets since it is not clear how to define securities in such spaces.
+The most direct way of discretizing the space may lead to loss of information from the traders.
+Extending prediction market techniques to this space has been a subject of recent interest.
+Various techniques to bound market maker loss in both a combinatorial and infinite outcome space have been studied (see, for instance, \cite{Abernethy13,OthmanS11}).
+
+One popular form of prediction markets is the market scoring rule~\cite{hanson03}.
+A market scoring rule considers all trades as a single chronological sequence.
+Traders earn rewards proportional to the incremental reduction in prediction loss caused by their trades in comparison to the previous trade.
+In other words, their rewards depend on the change in market probabilities caused by their trade, as well as on the eventual outcome.
+Thus, each trader has an
+incentive to minimize the prediction loss.
+In this format, the {\em market maker} who runs the market can suffer an overall loss, but Hanson~\cite{hanson03} showed that, for market scoring rules on finite outcome spaces, the loss of the market maker can be bounded.
+
+In this paper, we propose a general version of the popular logarithmic scoring rule extended to continuous outcome spaces.
+Our version of the scoring rule uses parametric distributions to elicit traders' expectation on some function (called statistics) of the data.
+It turns out that the maximum entropy distribution given the statistics are  exponential family distributions.
+
 We have dual goals in this paper.
@@
 We see this as the syntax of our prediction market mechanism.
-On the other, this formulation has rich semantics as well: it allows for analysis of market behavior under various environments.
-For instance, we analyze market behavior with budget limited traders, traders that are exponential-utility-maximizers.
+On the other, this formulation has rich semantics as well: it allows for analysis of the evolution of the market under various  models of trader behavior.
+
+Traders with limited budgets are constrained in how much they influence the market.
+Their purchases may not reflect their true beliefs.
+We analyze the market  under these circumstances; we are able to show that traders with good information can expect to profit and their influence over the market state increases over time whereas malicious traders have limited impact on the market.
+
+We also analyze the behavior of traders with prior exposure in the market.
+Our results formalize the intuitive notion that private belief and financial exposure are just two sides of the same coin; trades in a prediction market can be seen as updating privately held beliefs.
+
+For risk averse traders, we show how a single trader interacts with the market as well as when multiple traders achieve equilibrium.
+The eventual market state turns out to be a weighted combination of trader beliefs and initial market state; the weights depend on the risk aversion parameter of the individual traders.
+Trader behavior varies depending on whether they assimilate information as Bayesians or as frequentists.
+Interestingly, we show that although these two kinds of traders pick the eventual market state as a convex combination of their private belief and current market state, they do so in dual spaces.
+
 %\section{Notation and Definitions}

### 2014-02-06 WriteUp/maxentpredmkts.tex
More progress on information markets section

--- 2014-02-06 WriteUp/maxentpredmkts.tex
@@
 \section{Generalized Log Scoring Rules}
+\label{sec:scoring}
 
@@
 %
-\begin{definition}
+\begin{definition} \label{def:proper-score}
 A scoring rule $S$ is  \emph{proper over domain $\mD$} for statistic $\phi$ if for each $\mu \in \mM$ and $p \in \mD$ with expected statistic $\mu$, we have
@@
 The following observation is crucial.
-Let $\tp \in \mP$ be a density (not necessarily an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\mu' \in \mM$ be an alternative report.
-Then note that
+Let $\tp \in \mP$ be a density (not necessarily from an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\mu' \in \mM$ be an alternative report.
+Then note from~(\ref{exp-fam}) that
 %
@@
 where $\theta' = \grad G(\mu')$ is the natural parameter for the exponential family with statistic $\mu'$.
-We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, so when reasoning about the agent's reporting incentives we can just as well assume its beliefs are from an exponential family for simplicity.
+We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, which is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
 %
@@
 $$
-for some strictly convex function $G$. Let $T = G^*$, so that $\grad G = {\grad T}^{-1}$, and let $\theta = \grad G(\mu)$. Then the above takes the form
+for some strictly convex function $G$. Let $T = G^*$, so that $\grad G = {\grad T}^{-1}$, and let $\theta = \grad G(\mu)$. Then the above can be written as
 %
@@
 %
-As further intuition for the result, note that~(\ref{breg-div}) is the definition of the Bregman divergence with respect to strictly convex function $T$. Therefore we have 
+As further intuition for the result, note that~(\ref{breg-div}) is the definition of the so-called Bregman divergence with respect to strictly convex function $T$. Therefore we have 
 %
@@
 Theorem~\ref{maxent-score} leads to a straightforward procedure for constructing score rules for expectations: define the relevant statistic, and consider the maximum entropy (equivalently, exponential family) distribution consistent with the agent's reported mean $\mu$.
-The agent is compensated according to the log likelihood of the eventual outcome according to the latter.
+The scoring rule compensates the agent according to the log likelihood of the eventual outcome according to the latter.
 The interpretation is that the agent is only providing partial information about the underlying density, so the principal first infers a full density according to the principle of maximum entropy, and then scores the agent using the usual log score.
 
-An advantage of this generalization of the log score is that, for many statistics of interest, it leads to simple \emph{closed-form} formulas for the scoring rule.
+An advantage of this generalization of the log score is that, for many expectations of interest, it leads to simple \emph{closed-form} formulas for the scoring rule (including for multi-dimensional outcomes in some cases).
+This covers for instance several different combinations of the arithmetic, geometric, and harmonic means, as well as higher order moments.
 The following examples illustrate the construction.
@@
 %
-Typically the exponential distribution is parametrized by its rate $\lambda = 1/\mu$.
+%Typically the exponential distribution is parametrized by its rate $\lambda = 1/\mu$.
 In that case the rule takes the form $S(\mu, x) = -\lambda x + \log \lambda$, and we see that $\lambda$ corresponds to the natural parameter.
+%
+As a generalization of this example, the maximum entropy distribution for the $k$-th moment $\phi(x) = x^k$ with respect to the same base measure is the Weibull distribution.
+Taking its log density leads to the following equivalent scoring rule (after scaling and eliminating unnecessary additive constants):
+%
+\begin{equation*}
+S(\mu, x) = (k-1) \log x - k \log \mu - \Gamma\left(1+\frac{1}{k}\right)^k \left(\frac{x}{\mu}\right)^k,
+\end{equation*}
+%
+where $\Gamma$ denotes the usual gamma function (the extension of the factorial to the reals).
 \end{example}
@@
 %
-\begin{example}
+\begin{example} \label{ex:gaussian}
 As a base measure we take the Lebesgue over the real numbers $\bR$.
@@
 %
+
+In the above, Example~\ref{ex:gaussian} illustrates an important point about parametrizations of the elicited expectations.
+The variance $\sigma^2$ cannot be written as $\Exp[\phi(x)]$ for any $\phi$, because the mean $\mu$ enters the definition of $\sigma^2$ but is not available when $\phi$ is defined (indeed it is elicited in tandem with the variance).\footnote{This is an intuitive but far from formal explanation for the fact that the dimension of the message space, or \emph{elicitation complexity}, for eliciting the variance is at least 2~\citep{}.} Instead one must use the first two \emph{uncentered} moments $\Exp[x]$ and $\Exp[x^2]$.
+These are in bijection with $\mu$ and $\sigma^2$, so the resulting scoring rule can be re-written in terms of the latter.
+Therefore, it is possible to elicit not just expectations but also bijective transformations of expectations.
+
+ VON-MISES FISHER
 
@@
 
-A risk-neutral agent will choose to acquire shares up to the point where, for each share, expected payoff equals marginal price. Formally, if the agent acquires portfolio $\delta$, then we must have
+A risk-neutral agent will choose to acquire shares up to the point where, for each share, expected payoff equals marginal price.
+Formally, if the agent acquires portfolio $\delta$, moving the market state vector to $\theta' = \theta + \delta$, then we must have
 %
 \begin{equation} \label{risk-neutral-agent}
-\Exp_p[\phi(x)] = \grad C(\theta + \delta).
-\end{equation}
-%
-In this way, by its choice of $\delta$, the agent reveals that its expected belief is $\mu = \grad C(\theta + \delta)$.
+\Exp_p[\phi(x)] = \grad C(\theta').
+\end{equation}
+%
+In this way, by its choice of $\delta$, the agent reveals that its expected belief is $\mu = \grad C(\theta')$.
 We stress that this observation relies on the assumptions that 1) the agent is risk-neutral, 2) the agent does not incorporate the market's information into its own beliefs, and 3) the agent is not budget constrained.
+We will examine relaxations of each assumption in later sections.
 
@@
 This is of course exactly the log-partition function~(\ref{log-part}) for the exponential family with sufficient statistic $\phi$, and we recover the classic LMSR using outcome indicator vectors as statistics.
-Because an agent would never select a portfolio with infinite cost, the effective domain (i.e., the possible vectors of outstanding shares) of $C$ is $\Theta = \{\theta \in \bR^d : C(\theta) < +\infty \}$.
+Because an agent would never select a portfolio with infinite cost, the effective domain (i.e., the possible vectors of outstanding shares) of $C$ is $\Theta = \{\theta \in \bR^d : C(\theta) < +\infty \}$, which gives an economic interpretation to the natural parameter space of an exponential family.
+
+The correspondence between the cost function~({\ref{lmsr-cost}) and the log-partition function~(\ref{log-part}) suggests the following interpretation.
+The market maker maintains an exponential family distribution over the state space $\mX$ parametrized by share vectors that lie in $\Theta$.
+When an agent buys shares, it moves the distribution's natural parameter so that the market prices matches its beliefs, or in other words the market's mean parametrization matches the agent's expectation.
+
+There is a well-known duality between scoring rules and cost-function based markets. To see this in our context, recall from~(\ref{equalizer-rule}) that
+%
+\begin{equation*}
+\Exp_{\tp}[\log p(x;\mu')] = \la \theta', \mu \ra - T(\theta')
+\end{equation*}
+%
+where $\tp$ is the agent's belief and $\mu'$ the agent's report.
+The expected log score from reporting $\mu'$ is exactly the same as the expected payoff from buying portfolio of shares $\theta' = \grad C(\mu')$ (assuming an initial market state of 0), as $\la \theta', \mu \ra$ is the expected revenue and $T(\theta')$ is the cost.
+As in Section~\ref{sec:scoring} this reasoning relies on the assumption of risk-neutrality (and no budget constraints), not on any specific form for the agent's belief.
+
+The agent's expected profit from moving the share vector from $\theta$ to $\theta'$ is
+%
+\begin{eqnarray*}
+& & \la \theta' - \theta, \mu \ra - C(\theta') + C(\theta) \\
+& = & C(\theta) - C(\theta') - \la \theta - \theta', \grad C(\theta) \rangle \\
+& = & D_C(\theta, \theta') = D_{C^*}(\mu', \mu),
+\end{eqnarray*}
+%
+recalling~(\ref{breg-div}).
+Now~\citet{} have observed (among others) that the Kullback-Leibler divergence between two exponential family distributions is the Bregman divergence, with repect to the log-partition function, between their natural parameters.
+The agent's expected profit is therefore the KL divergence between the market's implied expectation and the exponential family corresponding to the agent's expectation, a well-known property from the classical LMSR~\citep{}.
+
+%* DIFFERENTIAL ENTROPY
+%* UNBOUNDED LOSS?
+%* EXAMPLE OF LOSS FOR EXPONENTIAL/GAUSSIAN
+
+EXPONENTIAL EXAMPLE: EFFECTIVE DOMAIN AND COST FUNCTION
+GAUSSIAN EXAMPLE
+VON MISES EXAMPLE
 
@@
 Evaluating~(\ref{cost}), the cost function takes the form $C(\theta) = \frac{\theta_1^2}{4\theta_2} - \frac{1}{2}\log(2\theta_2)$.
-
- In the current market-based approach, the agent's chosen portfolio $\theta$ can also form the basis of a distribution over outcomes.
-
-The scoring rule~(\ref{expscore}) decomposes neatly into payoff and cost functions.
-The first term $\la \theta(\mu), \phi(x) \ra$ defines the agent's outcome-contingent reward, while the second term~$C(\theta(\mu))$ is the cost of acquiring a portfolio $\theta(\mu)$.
-We see here the duality between the approach of directly reporting $\mu$, or acquiring shares $\theta$: an agent reporting beliefs $\mu$ under scoring rule~(\ref{scoring}) would choose to acquire shares $\theta(\mu)$ in the information market with cost function~(\ref{cost}).
-Since~(\ref{scoring}) is a proper scoring rule, the information market with cost function~(\ref{cost}) is based on a proper market scoring rule.
-
-%
-\begin{proposition}
-Let $C$ be the cumulant (i.e, cost function) for the exponential family corresponding to $\phi$, and let $\theta(\mu)$ be the optimal Lagrange multiplier for the mean constraints given an agent report of $\mu \in \mM$.
-Then the following scoring rule is equivalent to~(\ref{scoring}):
-%
-\begin{equation} \label{expscore}
-S(\mu, x) = \la \theta(\mu), \phi(x) \ra - C(\theta(\mu)).
-\end{equation}
-%
-\end{proposition}
-%
-
-\vfill
-\pagebreak
-
-\subsection{An Information-Theoretic Interpretation}\label{sec:infth}
- We observe a useful alternative view of the market 
-scoring rule prediction market for exponential family learning. We connect the cost, payoff and profit function to information-theoretic 
-quantities associated with the exponential family.
-
-The following result has been previously pointed out by Amari \cite{Amari-KL}. 
-\begin{theorem}\label{lem:profit_decomposition}{\bf (Profit Decomposition)}: Consider an exponential 
-family $\family$ of distributions over some set of statistics $\suff(x)$, with natural parameters $\beta$. Let $\pi, \rho \in \family$ be any two probability distributions in the family. We use 
-$\beta_\pi$ to denote the natural parameters of $\pi$ and $\mu_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
-we can define $\beta_\rho$ and  $\mu_\rho$. Let $\lpf(\beta_{\rho})$ indicate the log partition function of $\rho$.
-Let $H(\pi)$ denote the entropy of the distribution $\pi$, and 
- $K(\pi||\rho)$ denote the KL-divergence of $\rho$ relative to $\pi$.
-Then, the following equality holds:  
-\begin{equation}
- K(\pi||\rho) + H(\pi) = \lpf(\beta_{\rho}) - \beta_\rho \cdot \mu_\pi
-\label{eq:profit_decomposition}
-\end{equation}
-\end{theorem}
-\begin{proof}
-\begin{eqnarray*}
- K(\pi||\rho) + H(\pi) &=& \int_x \pi(x)\log{\frac{\pi(x)}{\rho(x)}}dx\\
- &&\quad - \int_x \pi(x)\log{\pi(x)}dx \\
-&=& - \int_x \pi(x)\log{\rho(x)}dx \\
-&=& - \int_x \pi(x)\left[ \beta_\rho \cdot \suff(x) - \lpf(\beta_{\rho})\right]dx\\
-&=& - \beta_\rho \cdot \int_x \pi(x) \suff(x) dx\\
-&&\quad + \lpf(\beta_{\rho}) \int_x \pi(x) dx \\
-&=&  \lpf(\beta_{\rho}) - \beta_\rho \cdot \mu_\pi
-\end{eqnarray*}
-\end{proof}
-
-Equation~\ref{eq:profit_decomposition} gives us an alternative view of the
-market scoring rule construction. Assume that $\pi$ is the true distribution, and consider two arbitrary
-distributions $\rho_1, \rho_2 \in \family$.
-Note that $\lpf(\rho)$  is independent of $\pi$, and $\betavec_\rho \cdot \muvec_\pi$ is linear in the probabilities $\pi(x)$.
-If we want to measure loss by the KL-divergence, we can do so (in expectation) by setting a cost function that captures the first term, and defining security quantities ($\betavec$) and payoffs ($\suff$) to capture the second term.
-In particular, in a market with cost function $\lpf$, if the
-market price initially implies a distribution $\rho_1$, and a trader moves
-the market to price than implies a distribution $\rho_2$, then the cost she
-incurs is $\lpf(\rho_2) - \lpf(\rho_1)$.
-The number of securities bought to make this trade is given by the vector ($\betavec_{\rho_2} - \betavec_{\rho_1}$), and the expected payoff of the securities are given by $\muvec_\pi$.
-Thus, by equation \ref{eq:profit_decomposition}, the {\em net profit} of the trader is equal to $K(\pi||\rho_1) - K(\pi||\rho_2)$, {\it i.e.}, the reduction in KL-divergence with respect to the true distribution.
-We note one useful property of this construction: For a fixed vector $\betavec$ of purchased securities, the cost is independent of the outcome
-(and outcome distribution $\pi$), while the payoff is independent of the initial market state in which these securities were purchased.
-\subsection{Properties of the Maximum Entropy Market}
-Consider the dual of the cost function, $C^{*}(\mu)$ defined as
-$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
-This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
-$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
-Thus, the mean parameters are the dual variables to the natural parameters.
-
-\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
-\proof
-To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
-\begin{eqnarray*}
--H(p) &=& \int_{x}p(x)\log p(x)\\
-&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
-&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
-&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
-\end{eqnarray*}
-\unproof
-\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
-\rmk The LMSR is essentially a special case applied to a multinomial distribution.
-The LMSR is known to have bounded ($\log n$) market maker loss.
-Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
-\unrmk
-
-{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
-\proof
-Note that, for an exponential family distribution
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
-&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
-\end{eqnarray*}
-If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
-\unproof
-
-%{\cor The exponential family market maker has unbounded worst case loss.}
-%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
-%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
-%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
-\paragraph{Loss of the market maker}
-First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
-
-Recall the definition of the conjugate dual:
-$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
-The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
-$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
-This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
-So we have 
-$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
-Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
-
-Let $q_{0}$ be the initial and $q_{f}$ the final market state.
-Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
-\begin{eqnarray*}
-\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
-&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
-&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
-&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
-&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
-&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
-&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
-&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
-&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
-\end{eqnarray*}
 
@@
 \end{document}
+
+%----------------------------------------------------------------------------------------
+
+\subsection{Properties of the Maximum Entropy Market}
+Consider the dual of the cost function, $C^{*}(\mu)$ defined as
+$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
+This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
+$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
+Thus, the mean parameters are the dual variables to the natural parameters.
+
+\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
+\proof
+To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
+\begin{eqnarray*}
+-H(p) &=& \int_{x}p(x)\log p(x)\\
+&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
+&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
+&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
+\end{eqnarray*}
+\unproof
+\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
+\rmk The LMSR is essentially a special case applied to a multinomial distribution.
+The LMSR is known to have bounded ($\log n$) market maker loss.
+Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
+\unrmk
+
+Consider the dual of the cost function, $C^{*}(\mu)$ defined as
+$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
+This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
+$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
+Thus, the mean parameters are the dual variables to the natural parameters.
+
+\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
+\proof
+To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
+\begin{eqnarray*}
+-H(p) &=& \int_{x}p(x)\log p(x)\\
+&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
+&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
+&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
+\end{eqnarray*}
+\unproof
+\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
+\rmk The LMSR is essentially a special case applied to a multinomial distribution.
+The LMSR is known to have bounded ($\log n$) market maker loss.
+Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
+\unrmk
+
+\subsection{An Information-Theoretic Interpretation}
+\label{sec:infth}
+
+ We observe a useful alternative view of the market 
+scoring rule prediction market for exponential family learning. We connect the cost, payoff and profit function to information-theoretic 
+quantities associated with the exponential family.
+
+The following result has been previously pointed out by Amari \cite{Amari-KL}. 
+\begin{theorem}\label{lem:profit_decomposition}{\bf (Profit Decomposition)}: Consider an exponential 
+family $\family$ of distributions over some set of statistics $\suff(x)$, with natural parameters $\beta$. Let $\pi, \rho \in \family$ be any two probability distributions in the family. We use 
+$\beta_\pi$ to denote the natural parameters of $\pi$ and $\mu_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
+we can define $\beta_\rho$ and  $\mu_\rho$. Let $\lpf(\beta_{\rho})$ indicate the log partition function of $\rho$.
+Let $H(\pi)$ denote the entropy of the distribution $\pi$, and 
+ $K(\pi||\rho)$ denote the KL-divergence of $\rho$ relative to $\pi$.
+Then, the following equality holds:  
+\begin{equation}
+ K(\pi||\rho) + H(\pi) = \lpf(\beta_{\rho}) - \beta_\rho \cdot \mu_\pi
+\label{eq:profit_decomposition}
+\end{equation}
+\end{theorem}
+\begin{proof}
+\begin{eqnarray*}
+ K(\pi||\rho) + H(\pi) &=& \int_x \pi(x)\log{\frac{\pi(x)}{\rho(x)}}dx\\
+ &&\quad - \int_x \pi(x)\log{\pi(x)}dx \\
+&=& - \int_x \pi(x)\log{\rho(x)}dx \\
+&=& - \int_x \pi(x)\left[ \beta_\rho \cdot \suff(x) - \lpf(\beta_{\rho})\right]dx\\
+&=& - \beta_\rho \cdot \int_x \pi(x) \suff(x) dx\\
+&&\quad + \lpf(\beta_{\rho}) \int_x \pi(x) dx \\
+&=&  \lpf(\beta_{\rho}) - \beta_\rho \cdot \mu_\pi
+\end{eqnarray*}
+\end{proof}
+
+Equation~\ref{eq:profit_decomposition} gives us an alternative view of the
+market scoring rule construction. Assume that $\pi$ is the true distribution, and consider two arbitrary
+distributions $\rho_1, \rho_2 \in \family$.
+Note that $\lpf(\rho)$  is independent of $\pi$, and $\betavec_\rho \cdot \muvec_\pi$ is linear in the probabilities $\pi(x)$.
+If we want to measure loss by the KL-divergence, we can do so (in expectation) by setting a cost function that captures the first term, and defining security quantities ($\betavec$) and payoffs ($\suff$) to capture the second term.
+In particular, in a market with cost function $\lpf$, if the
+market price initially implies a distribution $\rho_1$, and a trader moves
+the market to price than implies a distribution $\rho_2$, then the cost she
+incurs is $\lpf(\rho_2) - \lpf(\rho_1)$.
+The number of securities bought to make this trade is given by the vector ($\betavec_{\rho_2} - \betavec_{\rho_1}$), and the expected payoff of the securities are given by $\muvec_\pi$.
+Thus, by equation \ref{eq:profit_decomposition}, the {\em net profit} of the trader is equal to $K(\pi||\rho_1) - K(\pi||\rho_2)$, {\it i.e.}, the reduction in KL-divergence with respect to the true distribution.
+We note one useful property of this construction: For a fixed vector $\betavec$ of purchased securities, the cost is independent of the outcome
+(and outcome distribution $\pi$), while the payoff is independent of the initial market state in which these securities were purchased.
+
+\subsection{Properties of the Maximum Entropy Market}
+
+{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
+\proof
+Note that, for an exponential family distribution
+\begin{eqnarray*}
+H(p) &=& -\int_{x}p(x)\log p(x)dx\\
+&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
+&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
+\end{eqnarray*}
+If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
+\unproof
+
+%{\cor The exponential family market maker has unbounded worst case loss.}
+%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
+%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
+%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
+\paragraph{Loss of the market maker}
+First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
+
+Recall the definition of the conjugate dual:
+$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
+The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
+$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
+This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
+So we have 
+$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
+Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
+
+Let $q_{0}$ be the initial and $q_{f}$ the final market state.
+Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
+\begin{eqnarray*}
+\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
+&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
+&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
+&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
+&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
+&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
+&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
+&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
+&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
+\end{eqnarray*}

### 2014-02-07 WriteUp/maxentpredmkts.tex
More progress on information markets section

@@
 %
-\begin{eqnarray}
-& & \Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\mu')] \nonumber\\
-& = & \la \theta, \mu \ra - T(\theta) - \la \theta', \mu \ra + T(\theta') \nonumber\\
-& = & T(\theta') - T(\theta) - \la \theta' - \theta, \mu \ra \nonumber\\
-& = & T(\theta') - T(\theta) - \la \theta' - \theta, \grad T(\theta) \ra. \label{breg-div}
-\end{eqnarray}
+\begin{eqnarray*}
+& & \Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\mu')] \\
+& = & \la \theta, \mu \ra - T(\theta) - \la \theta', \mu \ra + T(\theta') \\
+& = & T(\theta') - T(\theta) - \la \theta' - \theta, \mu \ra \\
+& = & T(\theta') - T(\theta) - \la \theta' - \theta, \grad T(\theta) \ra.
+\end{eqnarray*}
 %
 The latter is positive by the strict convexity of $T$, which shows that the log score is proper.
-For the converse, assume the defined log score is proper.
+Now assume the defined log score is proper.
 By the Savage characterization of proper scoring rules for expectations (see~\cite{}), we must have
@@
 %
+
 As further intuition for the result, note that~(\ref{breg-div}) is the definition of the so-called Bregman divergence with respect to strictly convex function $T$. Therefore we have 
@@
 %
-\begin{example} \label{ex:exponential}
+\begin{example}
 As base measure we take the Lebesgue restricted to $[0,+\infty)$, and we consider the statistic $\phi(x) = x$ so that we are simply eliciting the mean.
@@
 Here the statistics being elicited are the mean vector ${\boldsymbol \mu}$ and the covariance matrix ${\boldsymbol \Sigma}$.
-These scoring rules have been studied by~\citet{} as rules that only depend on the mean and variance of the reported density.
+These scoring rules have been studied by Dawid and Sebastiani~\cite{} as rules that only depend on the mean and variance of the reported density.
 They note that these rules are weakly proper (because they do not distinguish between densities with the same first and second moments), but do not make the point that knowledge of the full density is not necessary on the part of the agent.
@@
 Some advantages over moving average: all data points matter (decreasing importance with time) and computationally only last data point need be stored.
-
-\vfill
 \pagebreak
@@
 
-In a single-agent setting, a scoring rule is used to \emph{elicit} the agent's belief.
-In a multi-agent setting, an information market can be used to \emph{aggregate} the beliefs of the agents.
-In his seminal paper~\citet{} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both in order to perform well in thin or thick markets.
-In this section, we adapt the generalized log scoring rule to a market scoring rule which leads to markets with simple closed-form cost functions for many statistics of interest.
-The exponential families framework also allows us to move beyond risk-neutral agents to understand the aggregation properties of information markets under alternative behaviors, such as Bayesian updating and risk aversion.
+The purpose of a prediction market is to elicit and aggregate the beliefs (i.e., subjective probabilities) of agents over a space of \emph{outcomes}.
+In a single-agent setting, a scoring rule is used to elicit the agent's beliefs.
+In a multi-agent setting, an information market is used to aggregate the agent's beliefs.
+Hanson~\cite{} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both so that they can perform well in both thin and thick markets.
+In this work we derive a wide-ranging generalization of the logarithmic market scoring rule.
+Using a maximum entropy approach, we explain how a market scoring rule can be developed for outcome spaces both discrete and continuous, and for generic properties of the underlying distribution (e.g., mean and variance), rather than just the probabilities of individual outcomes.
 
@@
 
-In an information market an agent's expected belief $\mu$ is elicited indirectly through the purchase and sale of contingent claim securities.
-Under this approach, each component~$i$ of the statistic $\phi$ is interpreted as the payoff function of a security; that is, a single share of security $i$ pays off $\phi_i(x)$ when $x \in \mX$ occurs.
-Thus if the portfolio of shares held by the agent is $\delta \in \bR^d$, where entry $\delta_i$ corresponds to the number of shares of security $i$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \delta, \phi(x) \ra$.
-
-As a concrete example, recall that in the classic finite-outcome case the statistic has a component for each outcome $x$ such that $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise.
-Therefore the corresponding security pays 1 dollar if outcome $x$ occurs.
-(These are known as Arrow-Debreu securities.) In Example~\ref{ex:exponential} the one-dimensional statistic is $\phi(x) = x$, corresponding to a security whose payoff is linear in the outcome $x \in \bR_+$.
-(This amounts to a futures contract.)
-
-The standard way to implement an information market in the literature, due to~\citet{}, is via a centralized market maker.
-The market maker maintains a convex, differentiable cost function $C : \bR^d \rightarrow (-\infty,+\infty]$, where $C(\theta)$ records the revenue collected when the vector of outstanding shares is $\theta$.
-The cost to an agent of purchasing portfolio $\delta$ under a market state of $\theta$ is $C(\theta + \delta) - C(\theta)$, and therefore the instantaneous prices of the securities are given by the gradient $\grad C(\theta)$.
+We next consider how to indirectly elicit $\mu$ by setting up a market of contingent claim securities.
+Under this approach, the elements of the index set $\mS$ are interpreted not as statistics but as securities.
+The payoff from one share of security $s$ when outcome $x$ occurs is given by the mapping $\phi_s$.
+Thus if the vector of shares held by the agent is $\theta \in \bR^d$, where entry $\theta_s$ corresponds the number of shares of security $s$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \theta, \phi(x) \ra$.
+As a concrete example, recall our multinomial distribution example from the previous section, where $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise.
+This means that security $x \in \mS$ pays 1 dollar if outcome $x \in \mX$ occurs, and nothing otherwise.
+(Such securities are known as Arrow-Debreu securities.) In our normal distribution example, the statistic for the mean had the mapping $\phi_1(x) = x$.
+Therefore a share of the corresponding security has a payoff that is linear in the outcome.
+(Such securities amount to futures contracts.)
 
@@
 %
-\begin{equation} \label{lmsr-cost}
-C(\theta) = \log \int_{\mX} \exp \la \theta, \phi(x) \ra \,d\nu(x).
+\begin{equation} \label{cost}
+C(\theta) = \log \int_{x \in \mX} \exp \left[\la \theta, \phi(x) \ra\right] \nu(dx).
 \end{equation}
@@
 VON MISES EXAMPLE
-
-Observe that in the context of our earlier multinomial example, cost function~(\ref{cost}) is exactly the cost function for Hanson's logarithmic market scoring rule.
 

### 2014-02-07 WriteUp/maxentpredmkts.tex
jake has made a number of edits to intro

@@
 Each trader in the market is assumed to have some private information that he uses to make a prediction on the outcome of the event.
-The traders are assumed to be strategic in the sense that every trader wants to maximize his profit by participating in the market.
-Further, since the trades are done sequentially, the trader is allowed to observe all past trades in the market and update his private information based on this information.
-
 Traders are allowed to report their beliefs on the outcome of the event by allowing them to buy and sell securities whose value depends on the outcome of this future event.
 This will effect the state of the market, thus updating the predictive probabilities for the event.
-For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins.
-Traders can see the past history of trades, so the price at which a current trader is willing to buy and sell these securities can be interpreted as an aggregate ``market probability'' of a particular candidate winning the election.
-
-Currently, most deployed prediction markets have binary or finite discrete outcome spaces.
+Further, since the trades are done sequentially, the trader is allowed to observe all past trades in the market and update his private information based on this information.
+Traders can see the past history of trades, so the price at which a current trader is willing to buy and sell these securities can be interpreted as an aggregate ``consensus probability forecast'' of a particular candidate winning the election.
+
+% For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins.
+
+% JAKE: I think most people reading this will have enough background on markets.
+% Currently, most deployed prediction markets have binary or finite discrete outcome spaces.
 This means that the outcome of every event can be characterized by a finite set of mutually exclusive exhaustive outcomes.
@@
 
-Directly applying techniques used for binary outcome spaces to large outcome spaces is computationally infeasible.
+% JAKE: removing this paragraph, we don't want to harp too much on the efficiency question
+%       as this has already been adressed in a number of papers 
+% Directly applying techniques used for binary outcome spaces to large outcome spaces is computationally infeasible.
 In fact, it has been shown that computing and updating prices for combinatorial outcomes spaces is $\#$P-hard \cite{Chen08}.
@@
 
-One popular form of prediction markets is the market scoring rule~\cite{hanson03}.
+(JAKE: I want to use this paragraph for a general overview of mm frameworks) One popular form of prediction markets is the market scoring rule~\cite{hanson03}.
 A market scoring rule considers all trades as a single chronological sequence.
@@
 In other words, their rewards depend on the change in market probabilities caused by their trade, as well as on the eventual outcome.
-Thus, each trader has an
-incentive to minimize the prediction loss.
+Thus, each trader has an incentive to minimize the prediction loss.
 In this format, the {\em market maker} who runs the market can suffer an overall loss, but Hanson~\cite{hanson03} showed that, for market scoring rules on finite outcome spaces, the loss of the market maker can be bounded.
 
-In this paper, we propose a general version of the popular logarithmic scoring rule extended to continuous outcome spaces.
+Much the work on prediction market frameworks has focused primarily on structural properties of the mechanism: incentive compatibility, the market maker loss, the available liquidity, the fluctuations of the prices as a function of the trading volume, to name a few.
+Absent from much of the literature is a corresponding \emph{semantics} of the market behavior or the observed prices.
+That is, how can we interpret the equilibrium market state when we have a number of traders with diverse beliefs on the underlying state of the world?
+In what sense is the market an aggregation mechanism?
+Do price changes relate to our usual Bayesian notion of information incorporation via posterior updating?
+
+In the present work we show that a number of classical statistical tools can be leveraged to design a prediction market framework in the mold of an \emph{exponential family distributions} that possesses a number of attractive properties.
+Common concepts in statistics \emph{entropy maximization}, \emph{log loss}, and \emph{bayesian inference}, relate to natural aspects of our class of mechanisms.
+In particular, the central objects in our market framework can be interpreted as elements of exponential families:
+\begin{itemize}
+	\item the \emph{payoff function} of the market corresponds to the \emph{sufficient statistics} of the probabiilty distribution;
+	\item the vector of \emph{oustanding shares} in the market corresponds to the \emph{natural parameter vector} of the distribution;
+	\item the market prices correspond to \emph{mean parameters};
+	\item the profit potential for a trader corresponds to a \emph{Kullback-Leibler divergence} between the trader's belief and that of the market.
+\end{itemize}
+
+In addition to showing this syntactic relationship between exponential families and prediction markets, we also explore the semantic implications as well.
+In particular, we show that our formulation allows us to analyze the evolution of the market under various  models of trader behavior:
+\begin{itemize}
+	\item We consider \emph{budget-limited traders} who are constrained in how much they influence the market.
+We analyze the market  under these circumstances; we are able to show that traders with good information can expect to profit and their influence over the market state increases over time whereas malicious traders have limited impact on the market.
+	\item We consider when our agents are risk-averse in an interestins special case, that is under the assumption they utilize \emph{exponential utility} to optimize their bets.
+In this case we can characterize precisely how a single trader interacts with the market as well as the equilibrium reached given multiple traders.
+The eventual market state turns out to be a weighted combination of trader beliefs and initial market state; the weights depend on the risk aversion parameter of the individual traders.
+	\item In addition, we observe that trader behavior varies depending on whether they assimilate information as Bayesians or as frequentists.
+Interestingly, we show that although these two kinds of traders pick the eventual market state as a convex combination of their private belief and current market state, they do so in dual spaces.
+\end{itemize}
+
+% In this paper, we propose a general version of the popular logarithmic scoring rule extended to continuous outcome spaces.
 Our version of the scoring rule uses parametric distributions to elicit traders' expectation on some function (called statistics) of the data.
@@
 
-We have dual goals in this paper.
+% The connection we establish between markets and exponential  We have dual goals in this paper.
 On the one hand, we highlight a structural similarity between prediction markets and exponential families.
 We see this as the syntax of our prediction market mechanism.
-On the other, this formulation has rich semantics as well: it allows for analysis of the evolution of the market under various  models of trader behavior.
-
-Traders with limited budgets are constrained in how much they influence the market.
-Their purchases may not reflect their true beliefs.
-We analyze the market  under these circumstances; we are able to show that traders with good information can expect to profit and their influence over the market state increases over time whereas malicious traders have limited impact on the market.
-
-We also analyze the behavior of traders with prior exposure in the market.
+On the other,
+
+% We also analyze the behavior of traders with prior exposure in the market.
 Our results formalize the intuitive notion that private belief and financial exposure are just two sides of the same coin; trades in a prediction market can be seen as updating privately held beliefs.
-
-For risk averse traders, we show how a single trader interacts with the market as well as when multiple traders achieve equilibrium.
-The eventual market state turns out to be a weighted combination of trader beliefs and initial market state; the weights depend on the risk aversion parameter of the individual traders.
-Trader behavior varies depending on whether they assimilate information as Bayesians or as frequentists.
-Interestingly, we show that although these two kinds of traders pick the eventual market state as a convex combination of their private belief and current market state, they do so in dual spaces.
 
@@
 
-\pagebreak
+%\pagebreak
 
@@
 Some advantages over moving average: all data points matter (decreasing importance with time) and computationally only last data point need be stored.
-\pagebreak
+%\pagebreak
 \section{Maximum Entropy Market Making}
@@
 \rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
-\pagebreak
+%\pagebreak
 
@@
 If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
-\pagebreak
-
-\pagebreak
+%\pagebreak
+
+%\pagebreak
 \section{Bayesian Traders with Linear Utility}
@@
 %\end{eqnarray*}¥
-\pagebreak
+%\pagebreak
 
@@
 This means that financial exposure can be equivalently understood as changing the privately held beliefs.
-\pagebreak
+%\pagebreak
 

### 2014-02-07 WriteUp/liquidity.tex
related works section; bib files; liquidity write-up

--- 2014-02-04 WriteUp/maxentpredmkts.bak.tex
@@
-% Adjusted to add \usepackage[numbers]{natbib}
-% \bibliographystyle{acmsmall} and \documentclass[prodmode,acmec]{ec-acmsmall}
-% Jan 5, 2013 - David Parkes
-%
-% v2-acmsmall-sample.tex, dated March 6 2012
-% This is a sample file for ACM small trim journals
-%
-% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
-% (c) 2010 Association for Computing Machinery (ACM)
-%
-% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
-% Users can also go through the FAQs available on the journal's submission webpage.
-%
-% Steps to compile: latex, bibtex, latex latex
-%
-% For tracking purposes => this is v1.3 - March 2012
-
 \documentclass[prodmode,acmec]{ec-acmsmall} % Aptara syntax
@@
 \usepackage{booktabs}
-
 %new commands
@@
 \newcommand{\muvec}{\pmb{\mu}}
-\newcommand{\st}{\hspace{10pt}\mathrm{s.t.}\hspace{10pt}}
-\newcommand{\norm}[1]{|| #1 ||}
-\newcommand{\tp}{\tilde{p}}
 
@@
 \begin{document}
-% Page heads
 \markboth{Abernethy et al.}{Maximum Entropy Prediction Markets}
@@
 \affil{University of Michigan, Ann Arbor}}
-%%\author{GANG ZHOU
-%%\affil{College of William and Mary}
-%%YAFENG WU
-%%\affil{University of Virginia}
-%%TING YAN
-%%\affil{Eaton Innovation Center}
-%%TIAN HE
-%%\affil{University of Minnesota}
-%%CHENGDU HUANG
-%%\affil{Google}
-%%JOHN A. STANKOVIC
-%%\affil{University of Virginia}
-%%TAREK F. ABDELZAHER
-%%\affil{University of Illinois at Urbana-Champaign}}
-% NOTE! Affiliations placed here should be for the institution where the
-%       BULK of the research was done. If the author has gone to a new
-%       institution, before publication, the (above) affiliation should NOT be changed.
-%       The authors 'current' address may be given in the "Author's addresses:" block (below).
-%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
-%       currently affiliated with NASA.
 
-\begin{abstract}
-In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets.
-We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data.
-Similarly,  the market state can be interpreted as a distribution over the outcome space.
-In particular, we show that a proper scoring rule can be derived from maximum entropy distributions.
-This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces.
-In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility.
-We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function.
-These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders.
-In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market.
-Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
-\end{abstract}
+\section{Cost Function with a Liquidity Parameter}
+We may introduce a liquidity parameter to the cost function as $C_{\lambda}(\theta)=\lambda C\left(\frac{\theta}{\lambda}\right)$.
+We would like to find an expression for a lower bound on expected payoff.
+We would also like to find an expression for final market state for exponential utility traders with exponential family beliefs.
+\subsection{lower bound on expected payoff}
+The prediction market is set up with cost function $C_{\lambda}(\theta)$  and payoff for a unit share vector as $\phi(x)$. To find a lower bound on expected payoff, we first note that
+\begin{eqnarray*}
+\E_{x\sim p(x;\beta)}[C_{\lambda}(\theta)-\theta\phi(x)]&=&C_{\lambda}(\theta)-\theta\nabla T(\beta)\\
+&=&\lambda T(\theta/\lambda)-\theta\nabla T(\beta)\\
+&=&\lambda [T(\theta/\lambda)-\theta/\lambda\nabla T(\beta)-T(\beta)+\beta\nabla T(\beta)]+\lambda T(\beta)-\theta\lambda\nabla T(\beta)\\
+&=&\lambda D_{T}(\theta/\lambda,\beta)+\lambda [T(\beta)-\theta\nabla T(\beta)]
+\end{eqnarray*}¥ 
+where $D_{T}(x_{1},x_{2}) = T(x_{1})-T(x_{2})-\nabla T(x_{1}-x_{2})$ is the Bregman Divergence between $x_{1}$ and $x_{2}$ based on $T$.
 
-\category{J.4}{Social and Behavioral Sciences}{Economics}
-\category{I.2.6} {Artificial Intelligence}{Learning}
-
-\terms{Prediction Market Design, Machine Learning Algorithms}
-
-\keywords{Exponential Families, Bayesian Learning, Maximum Entropy Distributions, generalized LMSR}
-
-\acmformat{Jacob Abernethy, Sindhu Kutty, S\'{e}bastien Lahaie, and Rahul Sami, 2014. Maximum Entropy Prediction Markets}
-%%Gang Zhou, Yafeng Wu, Ting Yan, Tian He, Chengdu Huang, John A. Stankovic, and Tarek F. Abdelzaher, 2010. A multifrequency MAC specially designed for  wireless sensor network applications.
-
-% At a minimum you need to supply the author names, year and a title.
-% IMPORTANT:
-% Full first names whenever they are known, surname last, followed by a period.
-% In the case of two authors, 'and' is placed between them.
-% In the case of three or more authors, the serial comma is used, that is, all author names
-% except the last one but including the penultimate author's name are followed by a comma,
-% and then 'and' is placed before the final author's name.
-% If only first and middle initials are known, then each initial
-% is followed by a period and they are separated by a space.
-% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.
-
-\begin{bottomstuff}
-%%This work is supported by the National Science Foundation, under
-%%grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.
-%%
-%%Author's addresses: G. Zhou, Computer Science Department,
-%%College of William and Mary; Y. Wu  {and} J. A. Stankovic,
-%%Computer Science Department, University of Virginia; T. Yan,
-%%Eaton Innovation Center; T. He, Computer Science Department,
-%%University of Minnesota; C. Huang, Google; T. F. Abdelzaher,
-%%(Current address) NASA Ames Research Center, Moffett Field, California 94035.
-\end{bottomstuff}
-
-\maketitle
-
-\section{Introduction}
-%------------------------------------------------------------
-%Outline
-%------------------------------------------------------------
-%- exp family review
-%- maxent scoring rule
-%	- generalized lmsr
-%	[- does modularity (a la Hanson) still hold?]
-%- prediction market setup
-%	- MLE with risk neutrality
-%- conjugate prior market
-%	- also on hyperparam?
-%- exp-util
-%- equilibrium for repeated trades
-%- budgets and adversarial markets
-% any liquidity results?
-
-We have dual goals in this paper.
-On the one hand, we highlight a structural similarity between prediction markets and exponential families.
-We see this as the syntax of our prediction market mechanism.
-On the other, this formulation has rich semantics as well: it allows for analysis of market behavior under various environments.
-For instance, we analyze market behavior with budget limited traders, traders that are exponential-utility-maximizers.
-\section{Notation and Definitions}
-convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesian updates, conjugate prior.
-
-\pagebreak
-
-\section{Generalized Log Scoring Rules}
-
-We consider a measurable space that consists of a set of outcomes $\mX$ together with a $\sigma$-algebra $\mF$.
-An agent or expert has a \emph{belief} over potential outcomes taking the form of a probability measure absolutely continuous with respect to a base measure~$\nu$.\footnote{Recall that a measure $P$ is absolutely continuous with respect to~$\nu$ if $P(A) = 0$ for every $A \in \mF$ for which $\nu(A) = 0$.
-In essence the base measure~$\nu$ restricts the support of $P$.
-In our examples $\nu$ will typically be a restriction of the Lebesgue measure for continuous outcomes or the counting measure for discrete outcomes.} Throughout we represent the belief as the corresponding density $p$ with respect to~$\nu$.
-Let $\mP$ denote the set of all such probability densities, and let $\mD \subseteq \mP$ be a subset from which the beliefs are drawn.
-
-We are interested in eliciting information about the agent's belief, in particular expectation information.
-Let $\phi: \mX \rightarrow \bR^d$ be a vector-valued random variable or \emph{statistic}, where $d$ is finite.
-The aim is to elicit $\mu = \E_{p}[\phi(x)]$ where $x$ is the random outcome.
-A \emph{scoring rule} is a device for this purpose.
-Let $$\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}$$ be the set of realizable statistic expectations.
-A scoring rule $S : \mM \times \mX \rightarrow \bR \cup \{-\infty\}$ pays the agent $S(\hmu,x)$ according to how well its report $\hmu \in \mM$ agrees with the eventual outcome $x \in \mX$.
-%
-\begin{definition}
-A scoring rule $S$ is  \emph{proper over domain $\mD$} for statistic $\phi$ if for each $\mu \in \mM$ and $p \in \mD$ with expected statistic $\mu$, we have
-%
-\begin{equation} \label{eq-proper}
-\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)]
-\end{equation}
-%
-for all alternative $\hmu$. If the domain is $\mD = \mP$, the set of all possible densities, then we simply say the scoring rule is \emph{proper}.
-\end{definition}
-%
-
-\noindent
-Note that given a proper scoring rule $S$ any affine transformation $\tilde{S}(\mu,x) = aS(\mu,x) + b(x)$ of the rule, with $a > 0$ and $b$ an arbitrary real-valued function of the outcomes, again yields a proper scoring rule termed \emph{equivalent}.
-Throughout we will often apply such affine transformations to obtain the clearest version of the scoring rule.
-We will also focus on scoring rules where inequality~(\ref{eq-proper}) is strict to avoid trivial cases such as constant scoring rules.
-
-Our definition is more general than classical scoring rules in two respects.
-Classically, scoring rules take in the entire density $p$ rather than just some statistic, and incentive compatibility must hold over all of $\mP$.
-When the outcome space is large or infinite, it is not feasible to directly communicate $p$, so our definition allows for summary information of the belief.
-If there is a bijection between the sets $\mD$ and $\mM$, then we say that $\mM$ parametrizes $\mD$ and write $p(\cdot\,;\mu)$ for the density mapping to $\mu$.
-
-Note that this definition of a scoring rule places only mild information requirements on the part of the agent to ensure truthful reporting.
-Because condition~(\ref{eq-proper}) holds for all $p$ consistent with expectation $\mu$, it is enough for the agent to simply know the latter and not the complete density to be properly incentivized.
-However, the agent must also agree that the density is actually drawn from domain $\mD$; if the scoring rule is proper, it is enough that the agent agree with the support of the density as implicitly defined by base measure $\nu$.
-
-When the outcome space is finite we can recover classical scoring rules from the definition by taking $\mD = \mP$ and using the statistic $\phi : \mX \rightarrow \{0,1\}^{\mX}$ that maps an outcome $x$ to a unit vector with a 1 in the component corresponding to $x$.
-The expectation of $\phi$ is then exactly the probability mass function.
-
-\subsection{Proper Scoring from Consistency}
-
-Our starting point for designing proper scoring rules is the classic logarithmic scoring rule for eliciting probabilities in the case of finite outcomes.
-This rule is simply $S(p,x) = \log p(x)$ and there are several ways to understand why it is proper, with different implications for generalization.
-
-First observe that the scoring rule compensates the agent with the log likelihood it assigned to the outcome, so the agent is maximizing the expected log likelihood via its report.
-Now it is a fundamental result in statistics that, under a variety of different sufficient conditions, the maximum likelihood estimator is statistically consistent; see~\cite{} for an overview of such conditions.
-This means that in the limit (as the size of the i.i.d.\ empirical sample increases) the log likelihood is maximized by the true parameters.
-By this reasoning, the following result follows immediately from the definition of consistency and the law of large numbers.
-%
-\begin{theorem} \label{max-like}
-Suppose that $\mM$ parametrizes $\mD$. The logarithmic scoring rule defined by
-%
-\begin{equation} \label{log-score}
-S(\mu, x) = \log p(x;\mu),
-\end{equation}
-%
-where $\mu \in \mM$, is proper over $\mD$ if and only if the maximum likelihood estimator for $\mu$ is consistent.
-\end{theorem}
-%
-We note that although our focus is on parametrizations by expectations, the preceding theorem in fact holds for more generic parameter spaces.
-The range of scoring rules it provides is broader than those characterized by~\cite{}, because it allows for a restriction over the set of beliefs.
-On the other hand, its applicability is limited to those domains $\mD$ that can be parametrized by the relevant statistics.
-The following provide concrete examples of these points.
-%
-\begin{example}
-Suppose the outcomes are supported on $[1,+\infty)$ and follow a Pareto distribution with density $f(x;\alpha) = \alpha/x^{\alpha+1}$ parametrized by an index $\alpha > 0$.
-The mean $\mu$ is related to the index via the one-to-one mapping $\mu = \frac{\alpha}{\alpha-1}$, so the density can alternatively be parametrized by the mean.
-Theorem~\ref{max-like} gives the following proper scoring rule for the mean of a Pareto distribution with support on $[1,+\infty)$:
-%
-\begin{equation} \label{pareto-scoring}
-S(\mu,x) = \log \frac{\mu}{\mu-1} - \left(\frac{\mu}{\mu-1} + 1\right) \log x.
-\end{equation}
-%
-We stress that the rule can only elicit the mean assuming the agent knows its belief is a Pareto distribution over $[1,+\infty]$.
-It does not elicit the mean of other families of densities parametrized by the mean (e.g., the exponential distribution).
-\end{example}
-%
-%
-\begin{example}
-Suppose the outcomes are supported on $(-\infty,+\infty)$ and follow a Cauchy distribution with density $f(x;m) = 1/\pi[1+(x-m)^2]$, parametrized by the median $m$.
-Theorem~\ref{max-like} leads to the following proper scoring rule for the median:
-%
-\begin{equation} \label{cauchy-scoring}
-S(m,x) = -\log[1+(x-m)^2].
-\end{equation}
-%
-Note that the median cannot be obtained as the expectation of any statistic. This highlights the range of parameters than may be elicited by the log score in different circumstances.
-\end{example}
-%
-
-\subsection{Proper Scoring from Maximum Entropy}
-
-We now turn to scoring rules that are proper over the entire domain of densities $\mP$.
-Our construction is the same as in the previous section: we take the log likelihood of a density parametrized by the relevant statistics.
-If the family of densities is chosen appropriately, the resulting rule will in fact hold over all $\mP$.
-To achieve this we draw on a well-known duality between maximum likelihood and maximum entropy.
-
-\subsubsection*{Exponential Families}
-
-We let $p(x;\mu)$ be the maximum entropy distribution with expected statistic $\mu$.
-Specifically, it is the solution to the following mathematical program:\footnote{We assume throughout that the minimum is finite and achieved for all $\mu \in \mM$.
-Some care is needed to ensure this holds for specific statistics and outcome spaces.
-For example, taking outcomes to be the real numbers, there is no maximum entropy distribution with a given mean $\mu$ (one can take densities tending towards the uniform distribution over the reals), but there is always a solution if we constrain both the mean and variance.}
-%
-\begin{equation} \label{maxent-prog}
- \min_{p \in \mP} F(p) \st \Exp_{p}[\phi(x)] = \mu,
-\end{equation}
-%
-where the objective function is the negative entropy of the distribution, namely
-%
-\[ F(p) = \int_{x \in \mX} p(x) \log p(x)\, d\nu(x).
-\]
-%
-Note that the explicit set of constraints in~(\ref{maxent-prog}) are linear, and to stress this fact we find it helpful to re-write them as $A_{\phi} p = \mu$, where $A_{\phi}$ is the expectation operator of statistics $\phi$.
-We let $G : \mM \rightarrow \bR$ be the optimal value function of~(\ref{maxent-prog}), meaning $G(\mu)$ is the (negative) entropy of the maximum entropy distribution with expected statistics $\mu$.
-
-It is well-known that solutions to~(\ref{maxent-prog}) are \emph{exponential family} distributions, whose densities with respect to $\nu$ take the form
-%
-\begin{equation} \label{exp-fam}
-p(x;\theta) = \exp( \la \theta, \phi(x) \ra - T(\theta) ).
-\end{equation}
-%
-The density is stated here in terms of its \emph{natural} parametrization $\theta \in \bR^d$, where $\theta$ arises as the Lagrange multiplier associated with linear constraints $A_{\phi}p = \mu$.
-The term $T(\theta)$ essentially arises as the multiplier for the normalization constraint (the density must integrate to 1), and so ensures that~(\ref{exp-fam}) is normalized:
-%
-\begin{equation} \label{log-part}
-T(\theta) = \log \int_{\mX} \exp \la \theta, \phi(x) \ra \,d\nu(x).
-\end{equation}
-%
-The function $T$ is known as the \emph{log-partition} or \emph{cumulant} function corresponding to the exponential family.
-Its domain is $\Theta = \{ \theta \in \bR^d : T(\theta) < +\infty \}$, called the natural parameter space.
-The exponential family is \emph{regular} if $\Theta$ is open---almost all exponential families of interest, and all those we consider in this work, are regular.
-The family is \emph{minimal} if there is no $\alpha \in \Theta$ such that $\la \alpha, \phi(x) \ra$ is a constant over $\mX$ ($\nu$-almost everywhere); minimality is a property of the associated statistic $\phi$, usually called the \emph{sufficient statistic} in the literature.
-
-The following proposition collects the relevant results on regular exponential families and their associated parameter domains and functions.
-A convex function $T$ is of \emph{Legendre type} if it is proper, closed, strictly convex and differentiable on the interior of its domain, and $\lim_{\theta \rightarrow \bar{\theta}} \norm{\grad T(\theta)} = +\infty$ when $\bar{\theta}$ lies on the boundary of the domain.
-%
-\begin{prop}
-Consider a regular exponential family with minimal sufficient statistic. The following properties hold:
-\begin{enumerate}
-\item $T$ and $G$ are of Legendre type, and $T = G^*$ (equivalently $G = T^*$). 
-\item The gradient map $\grad T$ is one-to-one and onto the interior of $\mM$. Its inverse is $\grad G$ which is one-to-one and onto the interior of $\Theta$.
-\item The exponential family distribution with natural parameter $\theta \in \Theta$ has expected statistic $\mu = \Exp_p[\phi(x)] = \grad T(\theta)$.
-\item The maximum entropy distribution with expected statistic $\mu$ is the exponential family distribution with natural parameter $\theta = \grad G(\mu)$. 
-\end{enumerate}
-\end{prop}
-%
-In the above $T^*$ denotes as usual the convex conjugate of $T$, which here can be evaluated as $T^*(\mu) = \sup_{\theta \in \Theta} \la \theta, \mu \ra - T(\theta)$.
-Similarly, $G^*(\theta) = \sup_{\mu \in \mM} \la \theta, \mu \ra - G(\mu)$.
-
-\subsubsection*{Proper Log Scoring}
-
-We are now in a position to analyze the log scoring rule under exponential family distributions.
-From our discussion so far, we have that an exponential family density can be parametrized either by the natural parameter $\theta$, or by the mean parameter $\mu$, and that the two are related by the invertible gradient map $\mu = \grad T(\theta)$.
-We will write $p(x;\theta)$ or $p(x;\mu)$ given the parametrization used, which should be clear from context.
-
-The following observation is crucial.
-Let $\tp \in \mP$ be a density (not necessarily an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\mu' \in \mM$ be an alternative report.
-Then note that
-%
-\begin{equation} \label{equalizer-rule}
-\Exp_{\tp}[\log p(x;\mu')] = \Exp_{p(\cdot;\mu)}[\log p(x;\mu')] = \la \theta', \mu \ra - T(\theta'),
-\end{equation}
-%
-where $\theta' = \grad G(\mu')$ is the natural parameter for the exponential family with statistic $\mu'$.
-We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, so when reasoning about the agent's reporting incentives we can just as well assume its beliefs are from an exponential family for simplicity.
-%
-\begin{theorem} \label{maxent-score}
-Consider the logarithmic scoring rule $S(\mu,x) = \log p(x;\mu)$ defined over a set of densities $\mD$ parametrized by $\mM$.
-The scoring rule is proper over the entire domain of densities $\mP$ if and only if $\mD$ is the exponential family with statistic $\phi$.
-\end{theorem}
-%
-%
-\begin{proof}
-Let $\mu,\mu' \in \mM$ be the agent's true belief and an alternative report, and let $p,p' \in \mP$ be densities consistent with each belief.
-Let $\theta = \grad G(\mu)$ and $\theta' = \grad G(\mu')$, and note that $\mu = \grad T(\theta)$.
-We have
-%
+Thus, the expected payoff  for a trader with exponential family belief parameter $\beta$ who moves the market state from $\theta$ to $\theta'$ can be written as
 \begin{eqnarray*}
-& & \Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\mu')] \\
-& = & \la \theta, \mu \ra - T(\theta) - \la \theta', \mu \ra + T(\theta') \\
-& = & T(\theta') - T(\theta) - \la \theta' - \theta, \mu \ra \\
-& = & T(\theta') - T(\theta) - \la \theta' - \theta, \grad T(\theta) \ra.
-\end{eqnarray*}
-%
-The latter is positive by the strict convexity of $T$, which shows that the log score is proper.
-Now assume the defined log score is proper.
-By the Savage characterization of proper scoring rules for expectations (see~\cite{}), we must have
-%
-$$
-S(\mu,x) = G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra
-$$
-for some strictly convex function $G$. Let $T = G^*$, so that $\grad G = {\grad T}^{-1}$, and let $\theta = \grad G(\mu)$. Then the above takes the form
-%
+\E_{x\sim p(x;\beta)}\left[C_{\lambda}(\theta)-\theta\phi(x)-(C_{\lambda}(\theta')-\theta'\phi(x))\right]=\lambda[D_{T}(\theta/\lambda,\beta)-D_{T}(\theta'/\lambda,\beta)]
+\end{eqnarray*}¥
+If $\theta'=a\beta+(1-a)\theta$ for some $a\in[0,1]$ we can write
 \begin{eqnarray*}
-\log p(x;\mu) & = & G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra \\
-& = & \la \theta, \mu \ra - T(\theta) - \la \theta, \mu - \phi(x) \ra \\
-& = & \la \theta, \phi(x) \ra - T(\theta),
-\end{eqnarray*}
-%
-which shows that $p(x;\mu)$ takes the form of an exponential family.
-\end{proof}
-%
-This theorem leads to a straightforward procedure for constructing score rules for expectations: define the relevant statistic, and consider the maximum entropy (equivalently, exponential family) distribution consistent with the agent's reported mean $\mu$.
-The agent is compensated according to the log likelihood of the eventual outcome according to the latter.
-The interpretation is that the agent is only providing partial information about the underlying density, so the principal first infers a full density according to the principle of maximum entropy, and then scores the agent using the usual log score.
-
-An advantage of this generalization of the log score is that, for many statistics of interest, it leads to simple \emph{closed-form} formulas for the scoring rule.
-The following examples illustrate the construction.
-%
-\begin{example}
-As base measure we take the Lebesgue restricted to $[0,+\infty)$, and we consider the statistic $\phi(x) = x$ so that we are simply eliciting the mean.
-The maximum entropy distribution with a given mean $\mu$ is the exponential distribution, and taking its log density gives the scoring rule
-%
-\begin{equation} \label{exp-scoring}
-S(\mu, x) = -\frac{x}{\mu} - \log \mu.
-\end{equation}
-%
-We stress that although this rule is derived from the exponential distribution, Theorem~\ref{maxent-score} implies that it elicits the mean of any distribution supported on the non-negative reals (e.g., Pareto, lognormal), in contrast to~(\ref{pareto-scoring}).
-Indeed, it is easy to see that the expected score~(\ref{exp-scoring}) depends only on the mean of the agent's belief because it is linear in $x$.
-%
-Typically the exponential distribution is parametrized by its rate $\lambda = 1/\mu$.
-In that case the rule takes the form $S(\mu, x) = -\lambda x + \log \lambda$, and we see that $\lambda$ corresponds to the natural parameter.
-\end{example}
-%
-%
-\begin{example}
-As a base measure we take the Lebesgue over the real numbers $\bR$.
-We are interested in eliciting the mean $\mu$ and variance $\sigma^2$, so as a statistic we take $\phi(x) = (x, x^2)$ for which $\Exp_p[\phi(x)] = (\mu, \mu^2+\sigma^2)$.
-The maximum entropy distribution for a given mean and variance is the normal distribution, and taking its log density gives the scoring rule
-%
-\begin{equation}
-S((\mu, \sigma^2), x) = -\frac{(x-\mu)^2}{\sigma^2} - \log \sigma^2.
-\end{equation}
-%
-Again, we stress that this scoring rule elicits the mean and variance of any density over the real numbers, not just those of a normal distribution.
-The construction easily generalizes to a multi-dimensional outcome space by taking the log density of the multivariate normal:
-%
-\begin{equation}
-S(({\boldsymbol \mu}, {\boldsymbol \Sigma}), \boldsymbol{x}) = -( \boldsymbol{x} - {\boldsymbol \mu})'{\boldsymbol \Sigma}^{-1}( \boldsymbol{x} - {\boldsymbol \mu}) - \log |{\boldsymbol \Sigma}|.
-\end{equation}
-%
-Here the statistics being elicited are the mean vector ${\boldsymbol \mu}$ and the covariance matrix ${\boldsymbol \Sigma}$.
-These scoring rules have been studied by Dawid and Sebastiani~\cite{} as rules that only depend on the mean and variance of the reported density.
-They note that these rules are weakly proper (because they do not distinguish between densities with the same first and second moments), but do not make the point that knowledge of the full density is not necessary on the part of the agent.
-\end{example}
-%
-
-\pagebreak
-\section{Maximum Entropy Market Making}
-
-The purpose of a prediction market is to elicit and aggregate the beliefs (i.e., subjective probabilities) of agents over a space of \emph{outcomes}.
-In a single-agent setting, a scoring rule is used to elicit the agent's beliefs.
-In a multi-agent setting, an information market is used to aggregate the agent's beliefs.
-Hanson~\cite{} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both so that they can perform well in both thin and thick markets.
-In this work we derive a wide-ranging generalization of the logarithmic market scoring rule.
-Using a maximum entropy approach, we explain how a market scoring rule can be developed for outcome spaces both discrete and continuous, and for generic properties of the underlying distribution (e.g., mean and variance), rather than just the probabilities of individual outcomes.
-
-\subsection{Information Market}
-
-We next consider how to indirectly elicit $\mu$ by setting up a market of contingent claim securities.
-Under this approach, the elements of the index set $\mS$ are interpreted not as statistics but as securities.
-The payoff from one share of security $s$ when outcome $x$ occurs is given by the mapping $\phi_s$.
-Thus if the vector of shares held by the agent is $\theta \in \bR^d$, where entry $\theta_s$ corresponds the number of shares of security $s$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \theta, \phi(x) \ra$.
-As a concrete example, recall our multinomial distribution example from the previous section, where $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise.
-This means that security $x \in \mS$ pays 1 dollar if outcome $x \in \mX$ occurs, and nothing otherwise.
-(Such securities are known as Arrow-Debreu securities.) In our normal distribution example, the statistic for the mean had the mapping $\phi_1(x) = x$.
-Therefore a share of the corresponding security has a payoff that is linear in the outcome.
-(Such securities amount to futures contracts.)
-
-To extract useful information from such a securities market, we set a pricing scheme and examine the number of shares the agent chooses to acquire.
-Let $\Omega$ be the set of all portfolios (i.e., vectors of shares) that the agent can feasibly hold; we will see how this domain is determined in an instant.
-Each security $s$ has an associated price function $c_s : \Omega \rightarrow \bR$, which gives the marginal price $c_s(\theta)$ of security $s$ when the agent holds portfolio $\theta$.
-(Note that the marginal price depends on the entire portfolio $\theta$, not just the number of shares $\theta_s$.) A risk-neutral agent will choose to acquire shares up to the point where, for each share, expected payoff equals marginal price.
-Formally, if the agent acquires portfolio $\theta$, then for each $s \in \mS$ we must have
-%
-\begin{equation} \label{market}
-\Exp_p[\phi_s(x)] = c_s(\theta).
-\end{equation}
-%
-In this way, by its choice of $\theta$, the agent reveals that its belief is $\mu = c(\theta)$.
-There are several important properties that the price function $c$ should have.
-To simplify the agent's portfolio acquisition process, it should not be the case that the total cost of acquiring $\theta$ depends on the order in which shares are bought.
-This means that there should exist a \emph{cost function} $C : \Omega \rightarrow \bR$ such that $c = \nabla C$.
-The gradient $\nabla C$ must be onto $\mM$ to ensure that~(\ref{market}) always has a solution---this is the most important but technical condition to realize.
-To ensure that the solution is unique, it would also be convenient if the cost function were strictly convex.
-%
-\begin{proposition}
-The following cost function is monotone, convex, and onto $\mM$: 
-%
-\begin{equation} \label{cost}
-C(\theta) = \log \int_{x \in \mX} \exp \left[\la \theta, \phi(x) \ra\right] \nu(dx).
-\end{equation}
-%
-\end{proposition}
-%
-\noindent
-Observe that in the context of our earlier multinomial example, cost function~(\ref{cost}) is exactly the cost function for Hanson's logarithmic market scoring rule.
-Our approach here provides a generalization of this market scoring rule to markets with contingent claim securities with arbitrary payoffs, designed to elicit specific properties of distributions, beyond just the probabilities of different outcomes.
-
-Because an agent would never select a portfolio with infinite cost, the effective domain of $C$ is $\Omega = \{\theta \:|\: C(\theta) < +\infty \}$.
-In the context of the multinomial distribution example, $\Omega = \bR^d$, so that shares of each security can always be bought or sold short.
-In the context of the normal distribution example the effective domain is $\Omega = \{(\theta_1,\theta_2) \in \bR^2 \:|\: \theta_2 > 0 \}$ if we re-define the second statistic to be $\phi_2(x) = -x^2$; this means that the security has a negative payoff for each share, and consequently the agent must be compensated to acquire such shares.
-Evaluating~(\ref{cost}), the cost function takes the form $C(\theta) = \frac{\theta_1^2}{4\theta_2} - \frac{1}{2}\log(2\theta_2)$.
-
-Now, in our previous elicitation approach that used a scoring rule, the process of computing the scoring rule also provided a complete distribution $p(x;\mu)$ over outcomes, which could be used to infer other properties beyond just the agent beliefs $\mu$.
-In the current market-based approach, the agent's chosen portfolio $\theta$ can also form the basis of a distribution over outcomes.
-Consider the following distribution, represented as a density with respect to a base measure $\nu$:
-%
-\begin{equation} \label{expfam}
-p(x;\theta) = \exp\left[ \la \theta, \phi(x) \ra - C(\theta) \right].
-\end{equation}
-%
-Meaning, the probability that a subset $X \subseteq \mX$ of the outcomes occurs is $\int_{x \in X} p(x;\theta)\nu(dx)$.
-Observe that by definition~(\ref{cost}), this probability density indeed integrates to 1 over $\mX$, and it is clear that the density is non-negative as required.
-
-In the statistical literature a distribution that takes the form~(\ref{expfam}) is known as an \emph{exponential family}.
-The mapping $\phi$ is known as the \emph{sufficient statistic}, $\theta$ is the \emph{natural parameter}, and $C$ is the log-partition or \emph{cumulant} function.
-%This connection between information markets and exponential families allows us to draw on a vast statistical literature to understand the market's properties, and a growing computer science literature to apply efficient algorithms to compute costs and perform inference.
-
-%%%
-%%%
-%%%
-
-\subsection{Duality}
-
-There is a well-known duality between the maximum entropy approach and exponential families, which translates into a duality between the scoring rule and information market just developed.
-The duality implies that the approach leads to a \emph{market scoring rule}, applicable to both thin and thick multi-agent settings.
-
-It is known that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}.
-To see this, let $\theta(\mu) \in \bR^d$ be the Lagrange multiplier corresponding to constraints~(\ref{mean}) when solving the maximum entropy program given the agent report $\mu$.
-(Our choice of notation is deliberately suggestive.) Let $A(\mu)$ be the Lagrange multiplier corresponding to~(\ref{normalize}).
-From the first order necessary conditions for optimality, we find that
-%
-\[ p(x;\mu) = \exp \left[ \la \theta(\mu), \phi(x) \ra - A(\mu) - 1 \right], \]
-%
-so that the solution is indeed in exponential family form.
-Conversely, given a cumulant $C$ from an exponential family and a parameter $\theta$, we see that $p(x;\theta)$ defined according to~(\ref{expfam}) is the maximum entropy distribution under constraints $\Exp_p[\phi(x)] = \mu$ where $\mu = \nabla C(\theta)$.
-We obtain the following.
-%
-\begin{proposition}
-Let $C$ be the cumulant (i.e, cost function) for the exponential family corresponding to $\phi$, and let $\theta(\mu)$ be the optimal Lagrange multiplier for the mean constraints given an agent report of $\mu \in \mM$.
-Then the following scoring rule is equivalent to~(\ref{scoring}):
-%
-\begin{equation} \label{expscore}
-S(\mu, x) = \la \theta(\mu), \phi(x) \ra - C(\theta(\mu)).
-\end{equation}
-%
-\end{proposition}
-%
-\noindent
-The scoring rule~(\ref{expscore}) decomposes neatly into payoff and cost functions.
-The first term $\la \theta(\mu), \phi(x) \ra$ defines the agent's outcome-contingent reward, while the second term~$C(\theta(\mu))$ is the cost of acquiring a portfolio $\theta(\mu)$.
-We see here the duality between the approach of directly reporting $\mu$, or acquiring shares $\theta$: an agent reporting beliefs $\mu$ under scoring rule~(\ref{scoring}) would choose to acquire shares $\theta(\mu)$ in the information market with cost function~(\ref{cost}).
-Since~(\ref{scoring}) is a proper scoring rule, the information market with cost function~(\ref{cost}) is based on a proper market scoring rule.
-
-\section{The Exponential Family Market Mechanism in Adversarial Markets}
-In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
-
-In this section, we consider the the prediction market setup with traders that may be either informative or malicious.
-The malicious traders may want to inject faulty information into the market.
-The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.
-
-We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market.
-Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
-
-\subsubsection{Budget-limited Aggregation}
-Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
-We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market.
-This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader.
-Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.
-
-In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
-Let $\alpha$ be the budget of a trader in the market.
-Suppose with infinite budget, the trader would have moved the market state from $\qvec_{init}$ to $\qvec$, where $\qvec$ represents his true beliefs.
-Suppose further that $\alpha<C(\qvec)-C(\qvec_{init})$.
-In this case, we want to budget-limit the trader's influence on the market state.
-
-We define the budget-limited final market state as $\tilde{\qvec}$. 
-Here, we consider a specific functional form of $\tilde{\qvec}$: $$\tilde{\qvec}=\lambda\qvec+(1-\lambda)\qvec_{init}$$  where $$\lambda=\min\Big(1, \frac{\alpha}{C(\qvec) - C(\qvec_{init})}\Big)$$ We first show that this trade is feasible given the trader's budget:
-\begin{theorem}
-Let the current market state be given by the vector $\qvec_{init}$ and $\tilde{\qvec}=\lambda\qvec+(1-\lambda)\qvec_{init}$.
-For $\lambda=\min\Big(1, \frac{\alpha}{C(\qvec) - C(\qvec_{init})}\Big)$, the cost to the trader to move the market state from $\qvec_{init}$ to $\tilde{\qvec}$ is at most his budget $\alpha$.
-\end{theorem}
-\begin{proof}
-From the convexity of $C$, we have
- $$C(\tilde{\qvec})\leq(1-\lambda)C(\qvec_{init})+\lambda C(\qvec)$$
-
-Now 
+D_{T}(\theta'/\lambda,\beta)&\leq& a D_{T}(\beta/\lambda,\beta)+(1-a) D_{T}(\theta/\lambda,\beta)\\
+D_{T}(\theta/\lambda,\beta)-D(\theta'/\lambda,\beta)&\geq& a D_{T}(\theta/\lambda,\beta)-a D_{T}(\beta/\lambda,\beta)
+\end{eqnarray*}¥
+%Note that 
+%\begin{eqnarray*}
+%\nabla_{\theta} C_{\lambda} (\theta)&=&\nabla_{\theta}\lambda C\left(\frac{\theta}{\lambda}\right)\\
+%&=&\nabla_{\theta/\lambda}C\left(\frac{\theta}{\lambda}\right)\\
+%&=&\E_{\theta/\lambda}[\phi(x)]
+%\end{eqnarray*}¥
+\subsection{exponential utility traders with exponential family beliefs}
+As before the prediction market is set up with cost function $C_{\lambda}(\theta)$  and payoff for a unit share vector as $\phi(x)$.
+Thus, the  payoff of a trader who moves the market state from $\theta$ to $\theta'$ is given by $C_{\lambda}(\theta)-C_{\lambda}(\theta')+\delta\phi(x)$ where $\delta=\theta'-\theta$ and the expected utility of the payoff for a trader with exponential belief parameter $\beta$ is given by
 \begin{eqnarray*}
-C(\tilde{\qvec}) - C(\qvec_{init}) &\leq& (1-\lambda)C(\qvec_{init})\\&&+\lambda C(\qvec) - C(\qvec_{init})\\
-&=& \lambda\ (C(\qvec) - C(\qvec_{init}))
+&&\E_{x\sim p(x;\beta)}U[C_{\lambda}(\theta)-C_{\lambda}(\theta')+\delta\phi(x)]\\
+&=&U[C_{\lambda}(\theta)-C_{\lambda}(\theta+\delta)+\frac{1}{a}(T(\beta)-T(\beta-a\delta))]\\
+&=&U[\lambda C(\theta/\lambda)-\lambda C\left(\frac{\theta+\delta}{\lambda}\right)+\frac{1}{a}(T(\beta)-T(\beta-a\delta))]\\
+&=&U[\lambda T(\theta/\lambda)-\lambda T\left(\frac{\theta+\delta}{\lambda}\right)+\frac{1}{a}(T(\beta)-T(\beta-a\delta))]
+\end{eqnarray*}¥
+Here $T$ is the log partition function; for the exponential family prediction market $C=T$. To maximize the utility, we set the gradient of the argument with respect to $\delta$ to 0. 
+$$\nabla T\left(\frac{\theta+\delta}{\lambda}\right) = \nabla T(\beta-a\delta)$$
+So we have for an optimal choice of $\delta$, 
+\begin{eqnarray*}
+\frac{\theta+\delta}{\lambda} &=& \beta-a\delta\\
+\implies \delta&=&\frac{\beta\lambda-\theta}{1+a\lambda}\\
+\implies \theta+\delta&=&\lambda\frac{\beta+a\delta}{1+a\lambda}
 \end{eqnarray*}
 
-%So when $\alpha<C(\qvec)-C(\qvec_{init})$, we pick $\lambda$ so that 
-%$$\lambda\leq \frac{\alpha}{(C(\qvec) - C(\qvec_{init}))}$$
-%
-%where we want $\lambda$ such that
- Thus, $C(\tilde{\qvec}) - C(\qvec_{init}) \leq \alpha$. 
- \end{proof}
-
- We note that moving to $\tilde{\qvec}$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
-In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\tilde{\qvec}$.
-Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\qvec$.
-\subsubsection{Damage Bound}
-Given this restriction on the initial budget of traders, we can now show that the total loss that can be induced by malicious traders is bounded.
-We define the loss function for $\qvec$ shares held as:
-$$L(\qvec,x)=-\log(p_{\betavec}(x))=\log\frac{\int e^{\qvec^{T}\Phi(x)}\ dx}{e^{\qvec^{T}\Phi(x)}}$$
-with the correspondence between $\betavec$ and $\qvec$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
-
-%{\bf Setup:} The market evolves in rounds $t=1,\ldots,T$.
-The end of a time segment occurs either when each expert has made his report or when the actual value of the random variable is received as input.
-Each expert $i$ reports the means of the sufficient statistics as available to him at time $t$.
-These are combined with the previously reported means to yield $\mu_{0}^{t}(x),\mu_{1}^{t}(x),\ldots,\mu_{k}^{t}(x)$ and an associated weight $w_{i}^{t}$ which can be interpreted as the number of points from which these values have been determined.
-These sufficient statistics correspond to a particular value of $B$ and hence of $\qvec$, the number of shares of each security that are held by traders.
-Since trader $i$ has finite budget $\alpha_{i}^{t}$ at time $t$, the current market state is moved to $\qvec'$ instead, so that cost of moving it to $\qvec'$ is within his budget.
-
-%For an initial positive budget of $\alpha_{i}^{0}$, we perform updates to the budgets of each agent as
-%$$\alpha_{i}^{t}:=\alpha_{i}^{t-1}+\lambda_{i}^{t}(L(\tilde{q}_{i-1}^{t},x^{t})-L({q}_{i}^{t},x^{t}))$$ where we define
-%$$\lambda_{i}^{t}:=\min\Big(1, \frac{\alpha_{i}^{t-1}}{C(\qvec_{i}) - C(\tilde{\qvec}_{i-1})}\Big)$$
-
-Suppose the prediction market runs over multiple rounds $t$.
-Let $\qvec_{0}^{t}$ be the initial number of shares of each security that are held.
-Let $\tilde{\qvec}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
-Let us assume that at this point we receive the value of the random variable $x^{t}$.
-
-%Let the loss function be $L(\cdot,\cdot)$, where the two arguments are the vector representing the number of shares held of each security and the actual value of the random variable.
-
-%We want to show:
-%$$L({\qvec}_{0}^{t},x^{t})-L(\tilde{\qvec}_{k}^{t},x^{t})+\sum_{i}\alpha_{i}^{t-1}\leq \sum_{i}\alpha_{i}^{t}$$
-%That is, the loss due to incorporating the budget-limited information of the traders, is not more than the increase in their budgets.
-Over multiple instances of the prediction market, we can track the change in budget of each trader. The change in budget for trader $i$ is
-\begin{eqnarray*}
-\alpha_{i}^{t}-\alpha_{i}^{t-1}&=&C(\tilde{\qvec}_{i-1}^{t})-C(\tilde{\qvec}_{i}^{t})-(\tilde{\qvec}_{i-1}^{t}-\tilde{\qvec}_{i}^{t})^{T}\Phi(x^{t})\\
-&=&L(\tilde{\qvec}_{i-1}^{t},x^{t})-L(\tilde{\qvec}_{i}^{t},x^{t})
-\end{eqnarray*}
-
-Define the myopic impact of a trader $i$ in segment $t$ as
-$$\Delta_{i}^{t}:=L(\tilde{\qvec}_{i-1}^{t},x^{t})-L(\tilde{\qvec}_{i}^{t},x^{t})$$
-Thus, the myopic impact captures incremental gain due to the trader in a round. Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
-
-The total myopic impact due to all $k$ active traders is given by
-$$\Delta^{t}=L(\qvec_{0}^{t},x^{t})-L(\tilde{\qvec}_{k}^{t},x^{t})$$
-Thus $-\Delta^{t}$ captures the incremental loss due to the predictive probability after aggregation of all $k$ traders.
-
-%The myopic impact of a trader $i$ in segment $t$, is given by
-%$$\Delta_{i}^{t}:=L(\tilde{B}_{i-1}^{t},x^{t})-L(\tilde{B}_{i}^{t},x^{t})$$
-%where $\tilde{B}_{i}$ is the vector of sufficient statistics corresponding to the the budget-limited state of the market as induced by trader $i$.
-%The total myopic impact due to all $k$ active traders is thus given by
-%$$\Delta^{t}=L(B_{0}^{t},x^{t})-L(\tilde{B}_{k}^{t},x^{t})$$
-
-%If the loss function is convex in the first argument, then we have:
-%\begin{eqnarray*}
-%\Delta_{i}^{t}&=&L(\tilde{B}_{i-1}^{t},x^{t})-L(\tilde{B}_{i}^{t},x^{t})\\
-%&=&L(\tilde{q}_{i-1}^{t},x^{t})-L(\tilde{q}_{i}^{t},x^{t})\\
-%&=&L(\tilde{q}_{i-1}^{t},x^{t})-L(\lambda_{i}{q}_{i}^{t}+(1-\lambda_{i})\tilde{q}_{i-1}^{t},x^{t})\\
-%&\geq&L(\tilde{q}_{i-1}^{t},x^{t})-\lambda_{i}L({q}_{i}^{t},x^{t})-(1-\lambda_{i})L(\tilde{q}_{i-1}^{t},x^{t})\\
-%&=&\lambda_{i}(L(\tilde{q}_{i-1}^{t},x^{t})-L({q}_{i}^{t},x^{t}))\\
-%&=& \alpha_{i}^{t} - \alpha_{i}^{t-1}
-%\end{eqnarray*}
-%Thus, the myopic damage caused by trader $i$ at round $t$ is upper-bounded by the change in his budget in that round.
-
-%If we pick a loss function that is also bounded in $[0,1]$, then we see that the budget never falls below zero. 
-\begin{theorem}
-A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
-\end{theorem}
-\begin{proof}
-Consider the myopic impact of a single trader $i$ after participating in the market $T$ times.
-Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
-$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
-
-Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
-\end{proof}
-
-This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders.
-
-In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation.
-We provide an information-theoretic justification for this claim in Section \ref{sec:infth}.
-The intuition behind this claim is that his prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
-
-\subsubsection{An Information-Theoretic Interpretation}\label{sec:infth}
- We observe a useful alternative view of the market 
-scoring rule prediction market for exponential family learning. We connect the cost, payoff and profit function to information-theoretic 
-quantities associated with the exponential family.
-
-The following result has been previously pointed out by Amari \cite{Amari-KL}. 
-\begin{theorem}\label{lem:profit_decomposition}{\bf (Profit Decomposition)}: Consider an exponential 
-family $\family$ of distributions over some set of statistics $\suff(x)$, with natural parameters $\betavec$. Let $\pi, \rho \in \family$ be any two probability distributions in the family. We use 
-$\betavec_\pi$ to denote the natural parameters of $\pi$ and $\muvec_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
-we can define $\betavec_\rho$ and  $\muvec_\rho$. We abuse notation slightly and let $\lpf(\rho)$ indicate the log partition function of $\rho$ which technically depends on its natural parameters.
-Let $H(\pi)$ denote the entropy of the distribution $\pi$, and 
- $K(\pi||\rho)$ denote the KL-divergence of $\rho$ relative to $\pi$.
-Then, the following equality holds:  
-\begin{equation}
- K(\pi||\rho) + H(\pi) = \lpf(\rho) - \betavec_\rho \cdot \muvec_\pi
-\label{eq:profit_decomposition}
-\end{equation}
-\end{theorem}
-\begin{proof}
-\begin{eqnarray*}
- K(\pi||\rho) + H(\pi) &=& \int_x \pi(x)\log{\frac{\pi(x)}{\rho(x)}}dx\\
- &&\quad - \int_x \pi(x)\log{\pi(x)}dx \\
-&=& - \int_x \pi(x)\log{\rho(x)}dx \\
-&=& - \int_x \pi(x)\left[ \betavec_\rho \cdot \suff(x) - \lpf(\rho)\right]dx\\
-&=& - \betavec_\rho \cdot \int_x \pi(x) \suff(x) dx\\
-&&\quad + \lpf(\rho) \int_x \pi(x) dx \\
-&=&  \lpf(\rho) - \betavec_\rho \cdot \muvec_\pi
-\end{eqnarray*}
-\end{proof}
-
-Equation~\ref{eq:profit_decomposition} gives us an alternative view of the
-market scoring rule construction. Assume that $\pi$ is the true distribution, and consider two arbitrary
-distributions $\rho_1, \rho_2 \in \family$.
-Note that $\lpf(\rho)$  is independent of $\pi$, and $\betavec_\rho \cdot \muvec_\pi$ is linear in the probabilities $\pi(x)$.
-If we want to measure loss by the KL-divergence, we can do so (in expectation) by setting a cost function that captures the first term, and defining security quantities ($\betavec$) and payoffs ($\suff$) to capture the second term.
-In particular, in a market with cost function $\lpf$, if the
-market price initially implies a distribution $\rho_1$, and a trader moves
-the market to price than implies a distribution $\rho_2$, then the cost she
-incurs is $\lpf(\rho_2) - \lpf(\rho_1)$.
-The number of securities bought to make this trade is given by the vector ($\betavec_{\rho_2} - \betavec_{\rho_1}$), and the expected payoff of the securities are given by $\muvec_\pi$.
-Thus, by equation \ref{eq:profit_decomposition}, the {\em net profit} of the trader is equal to $K(\pi||\rho_1) - K(\pi||\rho_2)$, {\it i.e.}, the reduction in KL-divergence with respect to the true distribution.
-We note one useful property of this construction: For a fixed vector $\betavec$ of purchased securities, the cost is independent of the outcome
-(and outcome distribution $\pi$), while the payoff is independent of the initial market state in which these securities were purchased.
-
-%Thus, we have shown that there exist deep connections between learning the natural parameters of an exponential family distribution and cost function based prediction markets.
-In Chapter \ref{nonmyopic} we exploit this connection and propose a bounded regret learning algorithm in a Bayesian model under less-than-ideal conditions: where the experts may be adversarial and strategic.
-
-\subsubsection{Budget of Informative Traders}\label{sec:inf}
-Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
-
-%For this, we need the following simple extension of Jensen's inequality that follows directly from the definition of convexity of a function.
-%\begin{lemma}\label{lem:convex}
-%For a strictly convex function $f$ with unique minimum $x_{min}$, $x\neq x_{min}$ in the domain of $f$, and $\lambda \in (0,1]$,
-%$$f(\lambda x_{min}+ (1-\lambda)x)< f(x)$$
-%\end{lemma}
-%\begin{proof}
-%\begin{eqnarray*}
-%f(\lambda x_{min}+ (1-\lambda)x) &<& \lambda f(x_{min})+ (1-\lambda)f(x)\\
-%&=& \lambda(f(x_{min})-f(x)) +f(x)\\
-%&\leq& f(x)
-%\end{eqnarray*}
-%\end{proof}
-
-We now characterize the expected change in budget for an informative trader. The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
-\begin{theorem}\label{thm:growth}
- Suppose that each informative trader gets a random sample of data, 
-resulting in a sequence of trader beliefs $\qvec_i$, and hence a sequence of budget-limited market positions
-$\tilde{\qvec}_i$, before a final outcome $x$.
-Then, the expectation, over trader $i$'s belief distribution $\qvec_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\tilde{\qvec}_{i-1}$.
-\end {theorem}
-\begin{proof}
-Trader $i$'s believed distribution is the distribution parametrized by $\qvec_i$.
-Therefore, the expected profit, over possible outcome values $x$, for a trader $i$ in a given round is given by
-\begin{eqnarray*}
-&& E[C(\tilde{\qvec}_{i-1})-C(\tilde{\qvec}_{i})-(\tilde{\qvec}_{i-1}-\tilde{\qvec}_{i})\Phi(x)]\\
-&=& K(\qvec_{i}||\tilde{\qvec}_{i-1})] + H(\qvec_{i}) - [K(\qvec_{i}||\tilde{\qvec}_{i}) + H(\pi)]]\\
-&=& K(\qvec_{i}||\tilde{\qvec}_{i-1}) - K(\qvec_{i}||\tilde{\qvec}_{i})
-\end{eqnarray*}
-We recall that $\tilde{\qvec}_{i}$ can be expanded as:
-$$\tilde{\qvec_{i}} = \lambda \qvec_{i} + (1- \lambda) \tilde{\qvec}_{i-1}$$ where $\lambda$ is strictly greater than $0$, but no more than $1$.
-It is a standard result from information theory that $K(.,.)$ is a convex function, and therefore, we have:
-\[
-K(\qvec_{i}||\tilde{\qvec}_{i-1}) \leq \lambda K(\qvec_{i}||\qvec_{i}) + (1-\lambda) K(\qvec_{i}||\tilde{\qvec}_{i-1})
-\]
-The first term on the right hand side is zero, and thus, we get:
-\[
-  K(\qvec_{i}||\tilde{\qvec}_{i-1}) - K(\qvec_{i}||\tilde{\qvec}_{i}) \geq \lambda K(\qvec_{i}||\tilde{\qvec}_{i-1})
-\]
-%Thus, it is sufficient to prove that E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t}] >= E_{\qvec}[K(\pi||\tilde{\qvec}_{i}^{t}].
- \end{proof}
-%\begin{theorem}
-%The expected budget of an informative trader increases when he participates in the market. Here, expectation is taken with respect to the true distribution from which the outcome is drawn.
-%We show that the budget of an informative trader grows in expectation:\end{theorem}
-%\begin{proof}
-%Let $\pi$ be the true distribution from which the outcome $x^{t}$ is drawn.
-Let $K(\pi||\qvec)$ be the relative entropy between the true distribution $\pi$ and the predictive distribution determined by the securities vector $\qvec$.
-Then the expected change in budget for a trader $i$ in round $t$ is given by
-%\begin{eqnarray*}
-%E_{x,\qvec}[\alpha_{i}^{t}&-&\alpha_{i}^{t-1}]\\
-%&=&E_{x,\qvec}[C(\tilde{\qvec}_{i-1}^{t})-C(\tilde{\qvec}_{i}^{t})-(\tilde{\qvec}_{i-1}^{t}-\tilde{\qvec}_{i}^{t})^{T}\Phi(x^{t})]\\
-%&=& E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t})] + H(\pi) - E[K(\pi||\tilde{\qvec}_{i}^{t}) + H(\pi)]]
-%\end{eqnarray*}
-%Thus, it is sufficient to prove that E_{\qvec}[K(\pi||\tilde{\qvec}_{i-1}^{t}] >= E_{\qvec}[K(\pi||\tilde{\qvec}_{i}^{t}].
-
-%In the case that
-
-%It is well known that minimizing the relative entropy of a predictive distribution with respect to the true distribution is equivalent to maximizing the likelihood function.
-% (see, for instance, \cite{whom?}).
-%In other words, the MLE of the natural parameters of an exponential family distribution minimizes the relative entropy of a predictive distribution with respect to the true distribution.
-
-%Further, for a true distribution $\pi$ and predictive distribution determined by $\qvec$, $K(\pi||\qvec) + H(\pi)$ is a convex function of $\qvec$.
-In fact, it is strictly convex when the representation is minimal.
-Thus, it has a unique minimum.
-%Further, $K(\pi||\rho) + H(\pi)$ is non-negative.
-
-%Suppose the current market state is represented by $\tilde{\qvec}_{i-1}$ and $\qvec_{i}$ represents the belief of trader $i$.
-Since the trader is budget limited, he moves the market state to $\tilde{\qvec}_{i}$ instead where $\tilde{\qvec}_{i}=  \lambda \qvec_{i}+ (1-\lambda)\tilde{\qvec}_{i-1}$.
-For an informative trader $i$, $\qvec_{i}$ represents the MLE of the natural parameters of the distribution.
-Thus, $K(\pi||\qvec) + H(\pi)$ has a unique minimum at $\qvec_{i}$.
-
-%Thus,   from Lemma \ref{lem:convex}, $E[K(\pi||\tilde{\qvec}_{i-1}^{t})+ H(\pi)] > E[K(\pi||\tilde{\qvec}_{i}^{t})+ H(\pi)]$ and trader $i$'s expected budget increases.
-%\end{proof}
-
-For continuous distributions with a density, the probability that a trader with private information will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs.
-This result suggests that, eventually, every informative trader will have the ability to influence the market state in accordance with his beliefs, without being budget limited.
-
-We note one important aspect of Theorem~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
-This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure; depending on the distribution family and the prior distribution over parameter values, maximum likelihood estimation might not optimize the true expected score.
-If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
-%To summarize our results for this in a market with budget-limited traders, the total loss induced by malicious traders is limited, while the budget of  informative traders grows.
-\section{The MaxEnt Market Mechanism: A Bayesian View}
-\subsubsection{Conjugate Priors}
-
-%In the previous section it's assumed that the agent has a belief $\q$, and moves the market state from initial state $\q_{init}$ to a convex combination $\tilde{\q} = \lambda \q + (1-\lambda)\q_{init}$.
-This note tries to justify this based on Bayesian updating, rather than budget limitations.
-(But it doesn't quite succeed.)
-
-\medskip\noindent
-Let $p(x;\theta)$ denote a probability density drawn from an exponential family with sufficient statistic $\phi : \mX \rightarrow \bR^d$, where $\theta$ is the natural parameter:
-$$
-p(x;\theta) = \exp \left[ \la \theta, \phi(x) \ra \right],
-$$
-and
-$$
-g(\theta) = \log \int_{\mX} \exp\la\phi(x),\theta\ra d\!x.
-$$
-Recall that $\grad g(\theta) = \Exp[\phi(x)]$ and $\grad^2 g(\theta) = \Var[\phi(x)]$. The family of conjugate priors is also an exponential family and takes the form
-$$
-p(\theta; n,\nu) = \exp\left[ \la n\nu,\theta \ra - ng(\theta) - h(\nu,n) \right].
-$$
-Here the feature map is $\psi(\theta) = (\theta, -g(\theta))$, the natural parameter is $(n\nu, n)$ where $n \in \bR$ and $\nu \in \bR^d$.
-The normalizer $h(\nu, n)$ is convex in $(n\nu,n)$.
-It is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$.
-The justification for this is that
-$$
-\Exp_{\theta} \left[ \Exp_x \left[ \phi(x) | \theta \right]\right] 
-= \Exp_{\theta} \left[\grad g(\theta)\right]
-= \nu.
-$$
-%(The proof is straightforward but not obvious. I have a reference for this but it's impossible to read---it's a mathematical statistics paper.)
-
-Suppose we draw a sample $X = (x_1,\ldots,x_m)$ of size $m$, and denote the empirical mean by $\mu[X] = \sum_{i=1}^m \phi(x_i)$. The posterior distribution is then
-$$
-p(\theta|X) \propto p(X|\theta)p(\theta|n,\nu) \propto \exp\left[ \la\mu[X]+n\nu,\theta\ra - (m+n)g(\theta) \right],
-$$
-and so the posterior mean is 
-\begin{equation} \label{posterior-mean}
-\frac{m\mu[X] + n\nu}{m+n}.
-\end{equation}
-Thus the posterior mean is a convex combination of the prior and posterior means, and their relative weights depend on the phantom and empirical sample sizes.
-
-In the context of prediction markets, one could imagine an agent who uses the current market estimate as its prior, and draws a empirical sample of size $m$.
-Its belief then takes the form~(\ref{posterior-mean}), where $n$ depends on the importance the agent places on the market estimate (perhaps based on how long the market has been running).
-However, note that the \emph{mean parameter} becomes a convex combination of market state and empirical belief, and this does not translate to the \emph{natural parameter}, which is what we would have liked.
-The new natural parameter is
-$$
-\grad g^{-1} \left(\frac{m\mu[X] + n\nu}{m+n}\right) = \grad g^* \left(\frac{m\mu[X] + n\nu}{m+n}\right).
-$$
-where $g^*$ is the convex conjugate of $g$.
-
-\subsubsection{Exponential Utility}
-
-Assume the agent's belief distribution $p$ belongs to an exponential family, so it takes the form
-%
-\[
-p(x;\theta) = \exp[\theta x - T(\theta)]
-\]
-%
-where $\theta$ is the natural parameter and $T$ is the log-partition function.
-(I'm assuming that the sufficient statistic is $\phi(x) = x$ just for simplicity.) Assume also that the agent has an exponential utility for money $w$:
-%
-\[
-U(w) = -\frac{1}{a} \exp(-aw).
-\]
-%
-Here $a$ is the coefficient of risk aversion (higher means more risk averse, and the utility function is more concave).
-
-%
-\begin{proposition}
-An agent with exponential family belief with natural parameter $\htheta$, and exponential utility with coefficient $a$, makes a trade that moves the current market share vector $\theta$ to the convex combination $\frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$ assuming an LMSR cost function.
-\end{proposition}
-%
-%
-\begin{proof}
-Let $\delta$ be the vector of shares the agent trades.
-The payoff given eventual outcome $x$ is then $\delta x - C(\delta + \theta) + C(\theta)$.
-The utility for this payoff is as follows (recall that $\htheta$ is the agent's believed natural parameter).
-%
-\[ U( \delta x - C(\delta + \theta) + C(\theta) ) = -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)).
-\]
-%
-Taking the expected utility, we obtain
-%
-\begin{eqnarray*}
-&   & \Exp\left[ U( \delta x - C(\delta + \theta) + C(\theta) ) \right] \\
-& = & \int_{\mX} -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)) \exp[\htheta x - T(\htheta)]\, dx \\
-& = & -\frac{1}{a} \int_{\mX} \exp[(\htheta -a \delta) x + a C(\delta + \theta) - a C(\theta)) - T(\htheta)]\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)] \int_{\mX} \exp[(\htheta -a \delta) x - T(\htheta -a \delta)]\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)] \int_{\mX} p(x; \htheta -a \delta)\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)]\\
-& = & U\left(- C(\delta + \theta) + C(\theta)) - \frac{1}{a} T(\htheta - a\delta) + \frac{1}{a} T(\htheta)\right)
-\end{eqnarray*}
-%
-The second-last equality follows from the fact that $\int_{\mX} p(x; \htheta -a \delta)\, dx = 1$.
-Since utility $U$ is monotone increasing, it is maximized by maximizing its argument, which is a concave function of $\delta$ by convexity of $C$ and $T$.
-The optimality condition for the argument is
-%
-\begin{equation} \label{eq:optim}
-\grad C(\delta^* + \theta) = \grad T(\htheta - a\delta^*)
-\end{equation}
-%
-Now if the market maker is using LMSR, then $C$ is the log-partition function of the corresponding exponential family and $C = T$.
-Then~(\ref{eq:optim}) can be solved by equating the arguments.
-This leads to $\delta^* = (\htheta - \theta) / (1+a)$, which moves the share vector to $\theta + \delta^* = \frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$.
-\end{proof}
-%
-
-In the statement of the result I use the term ``LMSR cost function'' somewhat loosely, because we are not necessarily dealing with a market over exhaustive, mutually exclusive outcomes.
-What is meant is the cost function that arises by taking the dual to entropy of the maxent distribution with given mean parameter $\mu$.
-As we've discussed this seems like the right generalization of LMSR to arbitrary mean parameter spaces.
-
-Note that as $a \rightarrow 0$, we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$.
-As $a$ grows (agent grows more risk averse) the agent makes smaller and smaller trades that keep it closer to the current estimate $\theta$.
-
-So there is a kind of congruence between exponential family beliefs and exponential utility (as the names would suggest). 
-\subsection{Properties of the Maximum Entropy Market}
-Consider the dual of the cost function, $C^{*}(\mu)$ defined as
-$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
-This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
-$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
-Thus, the mean parameters are the dual variables to the natural parameters.
-
-\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
-\proof
-To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
-\begin{eqnarray*}
--H(p) &=& \int_{x}p(x)\log p(x)\\
-&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
-&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
-&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
-\end{eqnarray*}
-\unproof
-\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
-\rmk The LMSR is essentially a special case applied to a multinomial distribution.
-The LMSR is known to have bounded ($\log n$) market maker loss.
-Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
-\unrmk
-
-{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
-\proof
-Note that, for an exponential family distribution
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
-&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
-\end{eqnarray*}
-If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
-\unproof
-
-%{\cor The exponential family market maker has unbounded worst case loss.}
-%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
-%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
-%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
-\paragraph{Loss of the market maker}
-First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
-
-Recall the definition of the conjugate dual:
-$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
-The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
-$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
-This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
-So we have 
-$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
-Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
-
-Let $q_{0}$ be the initial and $q_{f}$ the final market state.
-Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
-\begin{eqnarray*}
-\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
-&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
-&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
-&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
-&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
-&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
-&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
-&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
-&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
-\end{eqnarray*}
-
-\subsubsection{Example: Gaussian Markets}
-{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
-In this case,  $\beta$ is a $2$-dimensional  vector.
-Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
-So
-\begin{eqnarray*}
-\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
-&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
-&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
-\end{eqnarray*}
-Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
-$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
-$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
-{\ex
-We will now derive the expression for the differential entropy for the normal distribution.
-Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
-&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
-\end{eqnarray*}
-Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
-Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
-
-{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
-Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
-Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
-Let's now work out the loss for the Gaussian market maker.
-First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
-\begin{eqnarray*}
-D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
-&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
-&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
-%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
-&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
-\end{eqnarray*}
-
-So the Gaussian market maker loss is 
-%parallels to quadratic market maker loss? 
-\begin{eqnarray*}
-D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
-&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
-&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
-\end{eqnarray*}¥
-Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
-\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
-\section{Generalized Log Scoring Rule}
-Suppose that an agent holds a belief in the expected value of the moments of a random variable $x$.
-The maximum entropy distribution that is consistent with these beliefs is an exponential family distribution.
- \defn For an outcome $x$ with predicted probability $p$, the generalized log scoring rule is defined as $$S(p,x)\defeq a+b\log p(x)$$ where $a$ and $b$ are constants that we will set to $0$ and $1$ respectively in these notes.\undefn
-
-\begin{thm} The generalized log scoring rule for the maximum entropy distribution is proper.
-\end{thm}
-\proof
-Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$
-
-Note that the maximum entropy distribution is an exponential family distribution. 
- Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.
-
- We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
-  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
-   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
-   Or 
-\begin{equation}\label{eq:proper}
-\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
-\end{equation}
-
-   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
-   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
-   Rearranging the terms,
-   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
-   which is the first order condition on the convexity of $\psi(\cdot)$.
- \unproof
-
- \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters.
-That is,
-   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
-   \unrmk
-
- \subsection{Exponential Smoothing}
-One way to model the exponential family prediction market might be through exponential smoothing.
-\begin{itemize}
-\item A \emph{time series} is a set of data points sampled at periodic instances.
-\item \emph{Smoothing} is the process of creating an approximation function based on time series that while capturing  patterns in the data is relatively insensitive to noise.
-\item Natural first step is to use moving average. But average on how many samples (say $k$)? What to do until the first $k$ samples have been received?
-\item
-Next attempt might be to use weighted moving average. Advantage is that you can give higher weight to more recent terms. But still same disadvantage as simple moving average technique.
-\item
-\emph{Exponential Smoothing} attempts to remove this disadvantage.
-\end{itemize}
-
-Exponential smoothing is defined iteratively for $\alpha\in(0,1)$ as
-\begin{eqnarray*}
-s_{1} &=& x_{0}\\
-s_{t}&=&\alpha x_{t-1}+(1-\alpha)s_{t-1}
-\end{eqnarray*}
-
-Smaller values of $\alpha$ means greater smoothing e.g., $\alpha=0$ gives a constant function ($=x_{0}$).
-Note that the selection of $x_{0}$ is important -- and increases in importance as $\alpha$.
-Some advantages over moving average: all data points matter (decreasing importance with time) and computationally only last data point need be stored.
-
-\section{Bayesian Traders}
-Suppose data is drawn from an exponential family distribution.
-Further suppose, following a Bayesian model, that each trader has some prior belief distribution over the parameters of the data distribution.
-Assuming a conjugate prior, this prior distribution is also an exponential family distribution.
-
-The data distribution is given by $\exp\{\beta\cdot\phi(x)-\psi(\beta)\}$ where $\psi()$ is the log partition function and $\phi()$ are the sufficient statistics.
-Then the conjugate prior parametrized by $b_{0}=(n\nu, n)$ is given by $\exp\{b_{0}\cdot(\beta,\psi(\beta))-\Psi(b_{0})\}$where $\Psi()$ is the corresponding log partition function.
-The posterior distribution is given by $b=(n\nu+m\hmu, n+m)$ where $\hmu$ is the empirical mean of the sufficient statistics of the $m$ data points drawn from the data distribution, $\hmu=\sum_{i=1}^{m}\phi(x_{i})$.
-We also have $\E_{\beta\sim b_{0}}\E_{x\sim \beta}[\phi(x)]=\nu$.
-
-\subsection{Linear Utility}
-
-Suppose the trader wishes to maximize his expected payoff.
-Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by $$\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]$$
-But, $\E_{\beta\sim b}\E_{x\sim \beta}[\phi(x)]=\frac{n\nu+m\hmu}{ n+m}$.
-To obtain the maximum, we set the gradient to $0$.
-Thus, we have $$\nabla C(\delta + \theta)=\frac{n\nu+m\hmu}{ n+m}$$ For an exponential family prediction market, the final market state is given by
-$$\nabla C^{*}\left(\frac{n\nu+m\hmu}{ n+m}\right)$$
-
-\subsection{Exponential Utility}
-Now suppose the trader has exponential utility, given by $$U(w) = -\frac{1}{a} \exp(-aw)$$ where $a$ is the coefficient of risk aversion.
-Suppose the trader wishes to maximize his expected utility.
-Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by
-\begin{eqnarray*}
-&&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}U\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]\\
-&=&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}-\frac{1}{a}\exp\left[ -a\delta\phi(x) +a C(\delta + \theta) -a C(\theta)\right]
-\end{eqnarray*}¥
-
-\begin{eqnarray*}
-\E_{\beta\sim b}\E_{x\sim \beta}\exp\left[-a\delta\phi(x)\right]&=&\int_{\mathcal{B}}\int_{\mathcal{X}}\exp[-a\delta\phi(x)]\, \exp\{\beta\cdot\phi(x)-\psi(\beta)\}\, dx\\
-&&\qquad\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\, d\beta\\
-&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
-&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\int_{\mathcal{X}}\exp[(\beta-a\delta)\cdot\phi(x)-\psi(\beta-a\delta)]\, dx\\
-&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
-&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\\
-\end{eqnarray*}¥
-
-\section{Reinterpreting repeated trades in a market}
-Let us suppose that traders in this market have exponential utility \[
-U(w) = -\frac{1}{a} \exp(-aw).
-\]
-%
-Here $a$ is the coefficient of risk aversion (higher means more risk averse, and the utility function is more concave).
-
-Let $\delta_{1}^{*}$ be the optimal vector of shares the agent decides to trade on first entering the market with exponential family belief parametrized by natural parameter $\hat{\theta}$.
-Thus, his belief distribution is given by the pdf $$\exp\{\hat{\theta} \phi(x) - T(\hat{\theta})\}]$$ where $T(\hat{\theta}) =\int_{\mathcal{X}}\exp\{\hat{\theta} \phi(x)\}\ dx$ is the log partition function and $\phi(x)$ are the sufficient statistics.
-On a subsequent entry into this market when the market state is $\theta'$, his optimal purchase $\delta_{2}^{*}$ is given by the solution of
-$$\arg\max_{\delta_{2}} \E_{x\sim \hat{\theta}} U[\mathrm{payoff}]$$
-Here, the payoff given eventual outcome $x$ is  $[\delta_{1}^{*}+\delta_{2}^{*} ]\phi(x) - C(\delta_{1}^{*} + \theta) + C(\theta)- C(\delta_{2}^{*} + \theta') + C(\theta')$.
-Here $\theta' = \theta+\delta_{1}^{*}+\delta'$ where $\delta'$ captures the movement of the share vector by other traders in the market.
-The expected utility for this payoff is as follows.
-%%
-%\[ U( \delta x - C(\delta + \theta) + C(\theta) ) = -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)).
-%\]
-%%
-\begin{eqnarray*}
-&   & \arg\max_{\delta_{2}}\E\left[ (\delta_{1}^{*}+\delta_{2} )\phi(x) - C(\delta_{1}^{*} + \theta) + C(\theta)- C(\delta_{2} + \theta') + C(\theta') \right] \\
-& = & \arg\max_{\delta_{2}}\int_{\mathcal{X}} -\frac{1}{a} [\exp(-a (\delta_{1}^{*}+\delta_{2} )\phi(x) \\
-&  & + a C(\delta_{1}^{*} + \theta) -a C(\theta)+a C(\delta_{2} + \theta') -a C(\theta')) \exp\{\hat{\theta} \phi(x) - T(\hat{\theta})\}]\, dx \\
-& = & \arg\max_{\delta_{2}}\int_{\mathcal{X}} -\frac{1}{a} [\exp\{-a \delta_{2} \phi(x)+a C(\delta_{2} + \theta') -a C(\theta') \\
-&  & + a C(\delta_{1}^{*} + \theta) -a C(\theta)\} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})\}]\, dx \\
-& = & \arg\max_{\delta_{2}}\int_{\mathcal{X}} U[N(\delta_{2},\theta')] \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}]\, dx 
-\end{eqnarray*}
-
-Here the first equality follows from the fact that we are taking expectation over the traders belief parameter $\hat{\theta}$ and the second equality follows simply from rearranging the factors of $\phi(x)$.
-And lastly we have written $-\frac{1}{a} [\exp\{-a \delta_{2} \phi(x)+a C(\delta_{2} + \theta') -a C(\theta')]$ as $U[N(\delta_{2},\theta')]$ where $N(\delta_{2},\theta')$ is the net payoff when $\delta_{2}$ shares are purchased when the current market state is $\theta'$.
-
-This is equivalent to maximizing expected utility of $N(\delta_{2},\theta')$, where expectation is taken with respect to an exponential family distribution over $\mathcal{X}$ parametrized by $\hat{\theta}-a \delta_{1}^{*}$, so long as
-$$\int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}]\, dx = c\times \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)-T(\hat{\theta}-a \delta_{1}^{*})\} \, dx$$
-
-%$$\int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)-\psi(\hat{\theta}-a \delta_{1}^{*})\} \, dx =1$$
-%That is
-%$$\psi(\hat{\theta}-a \delta_{1}^{*})= \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)\}\ dx$$
-In other words we want the following statement to hold.
-\begin{eqnarray*}
-\exists c: c\times\exp\{- T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}&=&
-T(\hat{\theta}-a \delta_{1}^{*})\\
-&=& \int_{\mathcal{X}} \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x)\}\ dx
-\end{eqnarray*}
-
-where $c$ is a constant with respect to $\delta_{2}$ and $x$ (but could potentially depend on $\hat{\theta}$, $a$ and $\delta_{1}^{*}$).
-This is true because the right hand side of the equation is a function of only $\hat{\theta}$, $a$ and $\delta_{1}^{*}$.
-Let $\Theta\defeq\hat{\theta}-a \delta_{1}^{*}  $ be the effective belief.
-Thus we have that the trader chooses his share vector as follows.
-\begin{eqnarray*}
-&&  \arg\max_{\delta_{2}}\int_{\mathcal{X}} U[N(\delta_{2},\theta')] \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta})+ a C(\delta_{1}^{*} + \theta) -a C(\theta)\}\, dx \\
-& = &\arg\max_{\delta_{2}}\int_{\mathcal{X}} U[N(\delta_{2},\theta')] \exp\{(\hat{\theta}-a \delta_{1}^{*}) \phi(x) - T(\hat{\theta}-a \delta_{1}^{*})\}]\, dx\\
-& = & \arg\max_{\delta_{2}} \int_{\mathcal{X}} -\frac{1}{a} \exp\{-a \delta_{2} \phi(x)+a C(\delta_{2} + \theta') -a C(\theta')\} \exp\{\Theta\cdot \phi(x) - T(\Theta)\}\, dx\\
-& = & \arg\max_{\delta_{2}} U\left[- C(\delta_{2} + \theta') + C(\theta') - \frac{1}{a} T(\Theta - a\delta_{2}) + \frac{1}{a} T(\Theta)\right]\\
-\end{eqnarray*}
-This follows from the argument detailed in in exp-util and simple substitution.
-Further,  the maximizer is $\delta_{2}^* = (\Theta - \theta') / (1+a)$, which moves the share vector to $\theta' + \delta_{2}^* = \frac{1}{1+a} \Theta + \frac{a}{1+a} \theta'$ which is a convex combination of the effective belief and the current market state.
-
-In other words, an exponential utility maximizing trader who has belief $\hat{\theta}$ with prior exposure $\delta$ in a market  will behave identically as an exponential utility maximizing trader with belief $\hat{\theta}-a \delta$ and no prior exposure in the market.
-Here $a$ is the utility parameter.
-This means that financial exposure can be equivalently understood as changing the privately held beliefs.
-
-Note that as the trader becomes more risk averse, the  belief is updated more aggressively.
-As $a \rightarrow 0$, the trader becomes more risk neutral and his effective belief is closer to his true belief.
-%
-%----------------------------------------------------------------------------------------
-%	BIBLIOGRAPHY
-%----------------------------------------------------------------------------------------
-
-\subsection{Equilibrium in a market with multiple traders}
-We have shown that every exponential-utility maximizing trader picks the share vector $\delta_{i}$ so as to maximize the utility of $- C(\delta_{i} + \theta) + C(\theta) - \frac{1}{a_{i}} T(\beta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\beta_{i})$ where $\theta$ is the initial market state, $a_{i}$ the utility parameter and $\beta_{i}$ the belief parameter of  trader $i$.
-To maximize utility, it suffices to maximize the argument, assuming that $C()=T()=\psi()$.
-
-Now consider a market with multiple exponential-utility maximizing traders.
-We wish to compute the equilibrium state.
-Let us assume that equilibrium is reached at a market state $\theta+\sum_{i}\delta_{i}$ where $\theta$ is the initial market state before trading begins and $\delta_{i}$ is each trader's purchase.
-
-%Thus, each trader picks $\delta_{i}$ as follows.
-%\begin{eqnarray*}
-%&&\arg\max_{\delta_{i}}\left(- \psi(\theta+\sum_{i}\delta_{i}) + \psi(\theta) - \frac{1}{a_{i}} \psi(\beta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} \psi(\beta_{i})\right)\\
-%&\iff&\arg\min_{\delta_{i}}\left( \psi(\theta+\sum_{i}\delta_{i})+\sum_{i}\frac{1}{a_{i}}\psi(\beta_{i}-a_{i}\delta_{i})\right)
-%\end{eqnarray*}¥
-Recall the following result from game theory.
-\begin{thm}\label{thm:potl}
-Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}\defeq \delta_{1},\ldots,\delta_{i},\ldots,\delta_{n}$.
-If there exists a potential function $\phi(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=\phi(\vec{\delta})-\phi(\vec{\delta}_{-i},\delta_{i}')$$
-then when $\phi(\vec{\delta})$ is maximized, $\vec{\delta}$ is an equilibrium.
-\end{thm}¥
-
-Let $$\phi(\vec{\delta})\defeq  \psi(\theta+\sum_{i}\delta_{i})+\sum_{i}\frac{1}{a_{i}}\psi(\beta_{i}-a_{i}\delta_{i})$$
-Now $U_{i}(\vec{\delta})=- \psi(\theta+\sum_{j}\delta_{j}) + \psi(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} \psi(\beta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} \psi(\beta_{i})$.
-Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $\phi(\vec{\delta})$ for each $\delta_{i}$.
-\begin{eqnarray*}
-\nabla_{\delta_{i}}\phi(\vec{\delta})&=&\nabla\psi(\theta+\sum_{j=1}^{n}\delta_{i})-\nabla\psi(\beta_{i}-a_{i}\delta_{i})\\
-&=&0
-\end{eqnarray*}¥
-This can be achieved by equating the arguments. That is, for each trader $i$, 
-\begin{equation}\label{eq:eqbm}
-\beta_{i}-a_{i}\delta_{i}=\theta+\sum_{j=1}^{n}\delta_{j}
-\end{equation}¥
-Rewriting, we have for each trader $i$, 
-$$\frac{\beta_{i}}{a_{i}}-\delta_{i}=\frac{1}{a_{i}}(\theta+\sum_{j=1}^{n}\delta_{j})$$
-Thus,
-$$\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)-\sum_{i=1}^{n}\delta_{i}=\left(\theta+\sum_{j=1}^{n}\delta_{j}\right)\sum_{i=1}^{n}\frac{1}{a_{i}}$$
-And
-$$\sum_{j=1}^{n}\delta_{j}=
-\frac{\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)-\theta\sum_{i=1}^{n}\left(\frac{1}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
-Substituting in Equation \ref{eq:eqbm} we have the following expression for the final market state.
-$$\theta+\sum_{j=1}^{n}\delta_{j}=
-\frac{\theta+\sum_{i=1}^{n}\left(\frac{\beta_{i}}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
-Let us modify the cost function so that it includes a liquidity parameter:
-$$C_{\lambda}(q)\defeq\lambda \psi(q/\lambda)$$
-
-\subsubsection{Lower Bound on Expected Payoff for Exponential Utility Traders}
-Let $\theta$ be the current market state.
-We have shown that an exponential utility trader with belief distribution  parametrized by $\beta$ will move the market state to $\theta'=\frac{1}{1+a} \beta + \frac{a}{1+a} \theta$.
-Therefore, the trader's expected net payoff is given by
-\begin{eqnarray*}
-&& \E_{x\sim P_{\beta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
-&=& \psi(\theta)-\theta\E_{x\sim P_{\beta}}[\phi(x)]-(\psi(\theta')-\theta' \E_{x\sim P_{\beta}}[\phi(x)])\\
-&=& \psi(\theta)-\theta\nabla\psi(\beta)-(\psi(\theta')-\theta' \nabla\psi(\beta))\\
-&=& \psi(\theta)-\psi(\beta)-\nabla\psi(\beta)(\theta-\beta)-(\psi(\theta')-\psi(\beta)- \nabla\psi(\beta)(\theta'-\beta))\\
-&=& D_{\psi}(\theta,\beta)-D_{\psi}(\theta',\beta)\\
-&\geq&\frac{1}{a}D_{\psi}(\theta',\beta)\geq0
-\end{eqnarray*}
-The second to last inequality holds since $D_{\psi}(\theta',\beta)$ is convex in $\theta'$ and we have:
-\begin{eqnarray*}
-D_{\psi}(\theta',\beta)&=&D_{\psi}\left(\frac{1}{1+a} \beta + \frac{a}{1+a} \theta,\beta\right)\\
-&\leq&\frac{1}{1+a}D_{\psi}( \beta ,\beta)+ \frac{a}{1+a} D_{\psi}( \theta ,\beta)\\
-&=&\frac{a}{1+a} D_{\psi}( \theta ,\beta)
-%\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
-\end{eqnarray*}¥
-Thus, a trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{\psi}(\theta',\beta)$.
-\section{Conclusions}
-
 \end{document}

### 2014-02-07 WriteUp/maxentpredmkts.tex
related works section; bib files; liquidity write-up

@@
 
+\subsection{Related Work}
+Designing prediction markets to handle a large outcome space is an active area of research.
+In \cite{Chen08}, the authors use a restricted betting language to design efficient markets for a combinatorial outcome space.
+This technique is generalized by \cite{Pennock11}.
+\cite{Gao09} consider extending various automated market makers to an infinite outcome space.
+For the logarithmic market scoring rule they show that unbounded market maker loss can result in this setting.
+\cite{Abernethy11} and \cite{OthmanS11} specify frameworks under which they design cost function based markets that satisfy the desirable property of bounded market maker loss even in infinite outcome spaces.
+
+The connection between machine learning and prediction markets has been studied previously.
+\cite{Chen10} and \cite{Abernethy11} have previously explored the connection to learning algorithms to inform the design and understanding of prediction markets.
+In particular, \cite{Chen10}  consider the correspondence between prediction markets with market scoring rules and the Follow the Regularized Leader algorithm proposed by \cite{Kalai05} and thus provide insight into the aggregation mechanism of a prediction market.
+
+Independently of this work, Beygelzimer et al.~\cite{BLP12} have shown that, for a particular form of binary prediction markets and traders with log-utility, the long-run dynamics of trading activity and budgets over many prediction markets lead the markets to satisfy a bounded regret property with respect to the best single trader.
+Our results here, and the future work we have suggested, form a program to prove bounded regret properties of budget-limited prediction markets under much more general conditions.
+
 % In this paper, we propose a general version of the popular logarithmic scoring rule extended to continuous outcome spaces.
@@
 %----------------------------------------------------------------------------------------
-
+\bibliographystyle{acmsmall}
+\bibliography{maxentpredmkts}
 \end{document}

### 2014-02-08 WriteUp/maxentpredmkts.tex
small edits
completed first draft of info markets section

@@
 %
-\begin{eqnarray*}
-& & \Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\mu')] \\
-& = & \la \theta, \mu \ra - T(\theta) - \la \theta', \mu \ra + T(\theta') \\
-& = & T(\theta') - T(\theta) - \la \theta' - \theta, \mu \ra \\
-& = & T(\theta') - T(\theta) - \la \theta' - \theta, \grad T(\theta) \ra.
-\end{eqnarray*}
+\begin{eqnarray}
+& & \Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\mu')] \nonumber\\
+& = & \la \theta, \mu \ra - T(\theta) - \la \theta', \mu \ra + T(\theta') \nonumber\\
+& = & T(\theta') - T(\theta) - \la \theta' - \theta, \mu \ra \nonumber\\
+& = & T(\theta') - T(\theta) - \la \theta' - \theta, \grad T(\theta) \ra. \label{breg-div}
+\end{eqnarray}
 %
@@
 As a generalization of this example, the maximum entropy distribution for the $k$-th moment $\phi(x) = x^k$ with respect to the same base measure is the Weibull distribution.
-Taking its log density leads to the following equivalent scoring rule (after scaling and eliminating unnecessary additive constants):
+Taking its log density leads to the following equivalent scoring rule:
 %
@@
 Therefore, it is possible to elicit not just expectations but also bijective transformations of expectations.
-
- VON-MISES FISHER
 
@@
 
-We next consider how to indirectly elicit $\mu$ by setting up a market of contingent claim securities.
-Under this approach, the elements of the index set $\mS$ are interpreted not as statistics but as securities.
-The payoff from one share of security $s$ when outcome $x$ occurs is given by the mapping $\phi_s$.
-Thus if the vector of shares held by the agent is $\theta \in \bR^d$, where entry $\theta_s$ corresponds the number of shares of security $s$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \theta, \phi(x) \ra$.
-As a concrete example, recall our multinomial distribution example from the previous section, where $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise.
-This means that security $x \in \mS$ pays 1 dollar if outcome $x \in \mX$ occurs, and nothing otherwise.
-(Such securities are known as Arrow-Debreu securities.) In our normal distribution example, the statistic for the mean had the mapping $\phi_1(x) = x$.
-Therefore a share of the corresponding security has a payoff that is linear in the outcome.
-(Such securities amount to futures contracts.)
+In an information market an agent's expected belief $\mu$ is elicited indirectly through the purchase and sale of contingent claim securities.
+Under this approach, each component~$i$ of the statistic $\phi$ is interpreted as the payoff function of a security; that is, a single share of security $i$ pays off $\phi_i(x)$ when $x \in \mX$ occurs.
+Thus if the portfolio of shares held by the agent is $\delta \in \bR^d$, where entry $\delta_i$ corresponds to the number of shares of security $i$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \delta, \phi(x) \ra$.
+
+As a concrete example, recall that in the classic finite-outcome case the statistic has a component for each outcome $x$ such that $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise.
+Therefore the corresponding security pays 1 dollar if outcome $x$ occurs.
+(These are known as Arrow-Debreu securities.) In Example~\ref{ex:exponential} the one-dimensional statistic is $\phi(x) = x$, corresponding to a security whose payoff is linear in the outcome $x \in \bR_+$.
+(This amounts to a futures contract.)
+
+The standard way to implement an information market in the literature, due to~\citet{}, is via a centralized market maker.
+The market maker maintains a convex, differentiable cost function $C : \bR^d \rightarrow (-\infty,+\infty]$, where $C(\theta)$ records the revenue collected when the vector of outstanding shares is $\theta$.
+The cost to an agent of purchasing portfolio $\delta$ under a market state of $\theta$ is $C(\theta + \delta) - C(\theta)$, and therefore the instantaneous prices of the securities are given by the gradient $\grad C(\theta)$.
 
@@
 %
-\begin{equation} \label{cost}
+\begin{equation} \label{lmsr-cost}
 C(\theta) = \log \int_{x \in \mX} \exp \left[\la \theta, \phi(x) \ra\right] \nu(dx).
@@
 
-EXPONENTIAL EXAMPLE: EFFECTIVE DOMAIN AND COST FUNCTION
-GAUSSIAN EXAMPLE
-VON MISES EXAMPLE
-
-.
-In the context of the multinomial distribution example, $\Omega = \bR^d$, so that shares of each security can always be bought or sold short.
-In the context of the normal distribution example the effective domain is $\Omega = \{(\theta_1,\theta_2) \in \bR^2 \:|\: \theta_2 > 0 \}$ if we re-define the second statistic to be $\phi_2(x) = -x^2$; this means that the security has a negative payoff for each share, and consequently the agent must be compensated to acquire such shares.
-Evaluating~(\ref{cost}), the cost function takes the form $C(\theta) = \frac{\theta_1^2}{4\theta_2} - \frac{1}{2}\log(2\theta_2)$.
-
-\subsection{Example: Gaussian Markets}
-{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
-In this case,  $\beta$ is a $2$-dimensional  vector.
-Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
-So
-\begin{eqnarray*}
-\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
-&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
-&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
-\end{eqnarray*}
-Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
-$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
-$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
-{\ex
-We will now derive the expression for the differential entropy for the normal distribution.
-Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
-&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
-\end{eqnarray*}
-Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
-Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
-
-{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
-Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
-Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
-Let's now work out the loss for the Gaussian market maker.
-First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
-\begin{eqnarray*}
-D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
-&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
-&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
-%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
-&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
-\end{eqnarray*}
-
-So the Gaussian market maker loss is 
-%parallels to quadratic market maker loss? 
-\begin{eqnarray*}
-D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
-&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
-&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
-\end{eqnarray*}¥
-Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
-\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
-%\pagebreak
+\subsection{Examples: Real Line and the Sphere}
+
+Let us now revisit our scoring rules examples from Section~\ref{sec:scoring} in the context of prediction markets.
+The relevant entities now are the payoff function, the effective domain of shares, and the cost function.
+%
+\begin{example} \label{ex:exp-market}
+We consider outcomes over the positive reals $bR_+$ and set up a market for the expected outcome, consisting of a single security that pays off $\phi(x) = x$.
+The log partition funciton of the exponential distribution leads to the following cost function:
+%
+$$
+C(\theta) = -\log(-\theta).
+$$
+%
+The effective domain is $\Theta = \{\theta \in \bR : \theta < 0\}$.
+This means the market must start with a negative number of outstanding shares for the security, and the number shares must stay negative.
+The market maker need not explicitly enforce this, because by the Legendre property of $C$ the cost tends to $+\infty$ as the outstanding shares approach the boundary, which is straightforward to see in this example.
+\end{example}
+%
+
+%
+\begin{example} \label{ex:gauss-market}
+We consider outcomes over the real line $\bR$ and set up a market with securities corresponding to the first two uncentered moments (i.e, agents are betting on the return and volatility).
+The securities are defined by the payoffs $\phi(x) = (x, x^2)$.
+The log partition function of the normal distribution, under its natural parametrization, leads to the following cost function:
+%
+$$ 
+C(\theta) = -\frac{\theta_1^2}{4\theta_2} - \frac{1}{2}\log(-2\theta_2).
+$$
+The effective domain is $\Theta = \{(\theta_1,\theta_2) \in \bR^2 : \theta_2 < 0 \}$.
+Again, we have here an instance where it is not possible for the number of outstanding shares of the second security to exceed 0.
+However, an arbitrary amount of the securities can be sold short, which corresponds to increasing the variance of the market's estimate.
+\end{example}
+%
+
+%
+\begin{example} \label{ex:vonmises-market}
+As another example let the outcome space be the $d$-dimensional unit sphere.
+This setting was considered by~\citet{} who provide a cost function implicitly defined through a variational characterization.
+The maximum entropy approach leads to another alternative.
+We have a security for each of the $d$ dimensions, and security $i$ simply pays off $\phi_i(x) = x_i$, where $x \in \bR^d$ is the unit-norm outcome.
+The maximum entropy distribution over the sphere with such sufficient statistics is the von Mises-Fisher distribution.
+The log partition function corresponds to the cost function
+%
+$$
+C(\theta) = I_{\frac{d}{2}-1}(\norm{\theta}) - \left(\frac{d}{2} - 1 \right)\log \norm{\theta},
+$$
+%
+where $I_r$ refers to the modified Bessel function of first kind and order $r$.
+The effective domain of $\theta$ is the positive orthant in $\bR^d$.
+The mean parametrization of the von Mises-Fisher distribution gives a generalized log scoring rule for the expected outcome components, but it is unwieldy and involves several special functions.
+\end{example}
+%
 
@@
 \end{eqnarray*}
+
+\subsection{Example: Gaussian Markets}
+{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
+In this case,  $\beta$ is a $2$-dimensional  vector.
+Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
+So
+\begin{eqnarray*}
+\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
+&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
+&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
+\end{eqnarray*}
+Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
+$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
+$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
+{\ex
+We will now derive the expression for the differential entropy for the normal distribution.
+Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
+\begin{eqnarray*}
+H(p) &=& -\int_{x}p(x)\log p(x)dx\\
+&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
+&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
+&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
+&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
+&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
+\end{eqnarray*}
+Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
+Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
+
+{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
+Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
+Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
+Let's now work out the loss for the Gaussian market maker.
+First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
+\begin{eqnarray*}
+D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
+&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
+&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
+&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
+&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
+%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
+&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
+\end{eqnarray*}
+
+So the Gaussian market maker loss is 
+%parallels to quadratic market maker loss? 
+\begin{eqnarray*}
+D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
+&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
+&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
+\end{eqnarray*}¥
+Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
+\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
+%\pagebreak

### 2014-02-08 WriteUp/maxentpredmkts.tex
completed first draft of exp utility section

@@
 \newcommand{\tp}{\tilde{p}}
+\newcommand{\CE}{C\!E}
 
@@
 
-\section{Exponential Utility}
-
-Assume the agent's belief distribution $p$ belongs to an exponential family, so it takes the form
-%
-\[
-p(x;\theta) = \exp[\theta x - T(\theta)]
-\]
-%
-where $\theta$ is the natural parameter and $T$ is the log-partition function.
-(I'm assuming that the sufficient statistic is $\phi(x) = x$ just for simplicity.) Assume also that the agent has an exponential utility for money $w$:
-%
-\[
-U(w) = -\frac{1}{a} \exp(-aw).
-\]
-%
-Here $a$ is the coefficient of risk aversion (higher means more risk averse, and the utility function is more concave).
-
-%
-\begin{proposition}\label{prop:exp-util}
-An agent with exponential family belief with natural parameter $\htheta$, and exponential utility with coefficient $a$, makes a trade that moves the current market share vector $\theta$ to the convex combination $\frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$ assuming an LMSR cost function.
-\end{proposition}
+\section{Risk Averse Traders with Exponential Utility}
+
+In this section we relax that standard assumption that agents in the market are risk-neutral.
+We show that, with sufficient extra structure to the agents' beliefs and utilities, the market performs a clean aggregation of the agents' expectations in the form of a simple weighted averages.
+Assume that the agent has an exponential utility function for money $w$:
+%
+\begin{equation} \label{exp-util}
+U_a(w) = -\frac{1}{a} \exp(-aw).
+\end{equation}
+%
+Here $a$ controls the level of risk aversion: the agent is more risk averse as $a$ increases, and as $a$ tends to 0 we approach linear utility (risk-neutrality).
+Specifically, $a$ is the Arrow-Pratt coefficient of absolute risk aversion, and exponential utilities of the form~(\ref{exp-util}) are the unique utilities that exhibit constant absolute risk aversion (CARA).
+
+If wealth is distributed according to a probability measure $P$, then the \emph{certainty equivalent} of a random amount of wealth is defined as
+$$
+\CE(w) = U_a^{-1} \Exp_P\left[ U_a(w) \right].
+$$
+Suppose as before that the agent's belief over outcomes takes the form of a density $p$ with respect to base measure $\nu$.
+There is a close relationship between the log-partition function and the certainty equivalent under exponential utility.
+%
+\begin{lemma} \label{lem:exp-util}
+The certainty equivalent of the agent's expected profit, under exponential utility, when acquiring shares $\delta$ under a market state of $\theta$ is
+\begin{equation}
+\log a - T_p(-a\delta) - aC(\theta+\delta) + aC(\theta),
+\end{equation}
+where $T_p$ is the log partition function~(\ref{log-part}) with a base measure of $p\,d\nu$. Furthermore, if the agent's belief is an exponential family with parameter $\htheta$, we have
+%
+\begin{equation*}
+T_p(\delta) = T(\htheta + \delta) - T(\htheta),
+\end{equation*}
+%
+where $T$ is the usual log partition function with base measure $\nu$.
+\end{lemma}
 %
@@
 \begin{proof}
-Let $\delta$ be the vector of shares the agent trades.
-The payoff given eventual outcome $x$ is then $\delta x - C(\delta + \theta) + C(\theta)$.
-The utility for this payoff is as follows (recall that $\htheta$ is the agent's believed natural parameter).
-%
-\[ U( \delta x - C(\delta + \theta) + C(\theta) ) = -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)).
-\]
-%
-Taking the expected utility, we obtain
-%
-\begin{eqnarray*}
-&   & \Exp\left[ U( \delta x - C(\delta + \theta) + C(\theta) ) \right] \\
-& = & \int_{\mX} -\frac{1}{a} \exp(-a \delta x + a C(\delta + \theta) - a C(\theta)) \exp[\htheta x - T(\htheta)]\, dx \\
-& = & -\frac{1}{a} \int_{\mX} \exp[(\htheta -a \delta) x + a C(\delta + \theta) - a C(\theta)) - T(\htheta)]\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)] \int_{\mX} \exp[(\htheta -a \delta) x - T(\htheta -a \delta)]\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)] \int_{\mX} p(x; \htheta -a \delta)\, dx \\
-& = & -\frac{1}{a} \exp[a C(\delta + \theta) - a C(\theta)) + T(\htheta - a\delta) - T(\htheta)]\\
-& = & U\left(- C(\delta + \theta) + C(\theta)) - \frac{1}{a} T(\htheta - a\delta) + \frac{1}{a} T(\htheta)\right)
-\end{eqnarray*}
-%
-The second-last equality follows from the fact that $\int_{\mX} p(x; \htheta -a \delta)\, dx = 1$.
-Since utility $U$ is monotone increasing, it is maximized by maximizing its argument, which is a concave function of $\delta$ by convexity of $C$ and $T$.
-The optimality condition for the argument is
-%
-\begin{equation} \label{eq:optim}
-\grad C(\delta^* + \theta) = \grad T(\htheta - a\delta^*)
-\end{equation}
-%
-Now if the market maker is using LMSR, then $C$ is the log-partition function of the corresponding exponential family and $C = T$.
-Then~(\ref{eq:optim}) can be solved by equating the arguments.
-This leads to $\delta^* = (\htheta - \theta) / (1+a)$, which moves the share vector to $\theta + \delta^* = \frac{1}{1+a} \htheta + \frac{a}{1+a} \theta$.
+Explicitly, the certainty equivalent of the profit is
+%
+\begin{eqnarray*}
+& & \CE(\,\la\delta,\phi(x)\ra - [C(\theta+\delta) - C(\theta)]\,)\\
+& = & -\log \int_{\mX} \frac{1}{a} \exp\left( \la-a\delta,\phi(x)\ra + a [C(\theta+\delta) - C(\theta)]  \right) p(x)d\nu(x)\\
+& = & \log a - a [C(\theta+\delta) - C(\theta)] - \log \int_{\mX} \exp\left( \la-a\delta,\phi(x)\ra \right)  p(x)d\nu(x)\\
+& = & \log a - a [C(\theta+\delta) - C(\theta)] - T_p(-a\delta).
+\end{eqnarray*}
+%
+For the second part of the result, we have
+%
+\begin{eqnarray*}
+T_p(\delta) & = & \log \int_{\mX} \exp(\la\delta,\phi(x)\ra)\,p(x;\htheta)\, d\nu(x)\\
+& = &  \log \int_{\mX} \exp( \la\delta+\htheta,\phi(x)\ra - T(\htheta) )\, d\nu(x)\\
+& = &  T(\htheta+\delta) - T(\theta) + \log \int_{\mX} \exp( \la\delta+\htheta,\phi(x)\ra - T(\htheta+\delta) )\, d\nu(x)\\
+& = &  T(\htheta+\delta) - T(\theta) + \log \int_{\mX} p(x;\htheta+\delta) \, d\nu(x)\\
+& = &  T(\htheta+\delta) - T(\theta),
+\end{eqnarray*}
+%
+where the last line follows from the fact that density $p(x;\htheta+\delta)$ integrates to 1.
 \end{proof}
-%
-
-In the statement of the result I use the term ``LMSR cost function'' somewhat loosely, because we are not necessarily dealing with a market over exhaustive, mutually exclusive outcomes.
-What is meant is the cost function that arises by taking the dual to entropy of the maxent distribution with given mean parameter $\mu$.
-As we've discussed this seems like the right generalization of LMSR to arbitrary mean parameter spaces.
-
-Note that as $a \rightarrow 0$, we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$.
-As $a$ grows (agent grows more risk averse) the agent makes smaller and smaller trades that keep it closer to the current estimate $\theta$.
-
-So there is a kind of congruence between exponential family beliefs and exponential utility (as the names would suggest).
+
+\noindent
+Recall that for the generalized LMSR, the cost function $C$ is exactly the log partition function $T$.
+We are therefore lead to the following clear understanding of a risk-averse agent's behavior in such a market.
+%
+\begin{theorem}
+Suppose an agent has exponential utility with coefficient $a$ and exponential family beliefs with natural parameter $\htheta$.
+In the generalized LMSR market with current market state $\theta$, the agent's optimal trade $\delta$ moves the state vector to
+%
+\begin{equation} \label{exp-update}
+\theta + \delta = \frac{1}{1+a} \htheta + \frac{a}{1+a} \theta.
+\end{equation}
+%
+\end{theorem}
+%
+%
+\begin{proof}
+The agent's optimal trade maximizes its expected utility, or equivalently the certainty equivalent. From Lemma~\ref{lem:exp-util} and the fact that $T = C$, the agent therefore maximizes
+%
+$$
+\log a - T(\htheta - a\delta) + T(\htheta) - aT(\theta+\delta) + aT(\theta). 
+$$
+This objective is strictly concave, from the strict convexity of $T$. The optimum is therefore characterized by the first-order conditions:
+$$
+\grad T(\htheta - a\delta) = \grad T(\theta+\delta).
+$$
+As the gradient map $\grad T$ is one-to-one, this is solved by equating the arguments, which leads to $\delta = (\htheta - \theta) / (1+a)$ and~(\ref{exp-update}).
+% 
+\end{proof}
+%
+
+Note that as, $a$ tends to 0, we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$.
+As $a$ grows larger (agent grows more risk averse) the agent makes smaller trades to reduce it exposure, and the final state stays closer to the current state $\theta$.
+Update~(\ref{exp-update}) implies that under the conditions of the theorem, a market that receives a sequence of myopic traders aggregates their natural parameters in the form of an exponentially weighted moving average.
+The final market estimates (i.e., prices) are obtained by applying $\grad T$ to this average.
 

### 2014-02-08 WriteUp/maxentpredmkts.tex
revision of scoring rules section

@@
 
-We consider a measurable space that consists of a set of outcomes $\mX$ together with a $\sigma$-algebra $\mF$.
+We consider a measurable space consisting of a set of outcomes $\mX$ together with a $\sigma$-algebra $\mF$.
 An agent or expert has a \emph{belief} over potential outcomes taking the form of a probability measure absolutely continuous with respect to a base measure~$\nu$.\footnote{Recall that a measure $P$ is absolutely continuous with respect to~$\nu$ if $P(A) = 0$ for every $A \in \mF$ for which $\nu(A) = 0$.
@@
 In our examples $\nu$ will typically be a restriction of the Lebesgue measure for continuous outcomes or the counting measure for discrete outcomes.} Throughout we represent the belief as the corresponding density $p$ with respect to~$\nu$.
-Let $\mP$ denote the set of all such probability densities, and let $\mD \subseteq \mP$ be a subset from which the beliefs are drawn.
+Let $\mP$ denote the set of all such probability densities.
 
@@
 A scoring rule $S : \mM \times \mX \rightarrow \bR \cup \{-\infty\}$ pays the agent $S(\hmu,x)$ according to how well its report $\hmu \in \mM$ agrees with the eventual outcome $x \in \mX$.
+The following definition is due to~\citet{}.
 %
 \begin{definition} \label{def:proper-score}
-A scoring rule $S$ is  \emph{proper over domain $\mD$} for statistic $\phi$ if for each $\mu \in \mM$ and $p \in \mD$ with expected statistic $\mu$, we have
+A scoring rule $S$ is  \emph{proper} for statistic $\phi$ if for each $\mu \in \mM$ and $p \in \mP$ with expected statistic $\mu$, we have
 %
@@
 %
-for all alternative $\hmu$. If the domain is $\mD = \mP$, the set of all possible densities, then we simply say the scoring rule is \emph{proper}.
+for all alternative $\hmu \neq \mu$.% If the domain is $\mD = \mP$, the set of all possible densities, then we simply say the scoring rule is \emph{proper}.
 \end{definition}
@@
 \noindent
-Note that given a proper scoring rule $S$ any affine transformation $\tilde{S}(\mu,x) = aS(\mu,x) + b(x)$ of the rule, with $a > 0$ and $b$ an arbitrary real-valued function of the outcomes, again yields a proper scoring rule termed \emph{equivalent}.
-Throughout we will often apply such affine transformations to obtain the clearest version of the scoring rule.
+Note that given a proper scoring rule $S$ any affine transformation $\tilde{S}(\mu,x) = aS(\mu,x) + b(x)$ of the rule, with $a > 0$ and $b$ an arbitrary real-valued function of the outcomes, again yields a proper scoring rule termed \emph{equivalent}~\citep{}.
+Throughout we will implicitly apply such affine transformations to obtain the clearest version of the scoring rule.
 We will also focus on scoring rules where inequality~(\ref{eq-proper}) is strict to avoid trivial cases such as constant scoring rules.
 
-Our definition is more general than classical scoring rules in two respects.
  Classically, scoring rules take in the entire density $p$ rather than just some statistic, and incentive compatibility must hold over all of $\mP$.
-When the outcome space is large or infinite, it is not feasible to directly communicate $p$, so our definition allows for summary information of the belief.
+When the outcome space is large or infinite, it is not feasible to directly communicate $p$, so the definition allows for summary information of the belief.
+
+Note that Definition~\ref{def:proper-score} places only mild information requirements on the part of the agent to ensure truthful reporting.
+Because condition~(\ref{eq-proper}) holds for all $p$ consistent with expectation $\mu$, it is enough for the agent to simply know the latter and not the complete density to be properly incentivized.
+However, the agent must also agree with the support of the density as implicitly defined by base measure $\nu$.
+
+When the outcome space is finite we can recover classical scoring rules from the definition by using the statistic $\phi : \mX \rightarrow \{0,1\}^{\mX}$ that maps an outcome $x$ to a unit vector with a 1 in the component corresponding to $x$.
+The expectation of $\phi$ is then exactly the probability mass function.
+
+\subsection{Proper Scoring from Maximum Entropy}
+
+Our starting point for designing proper scoring rules is the classic logarithmic scoring rule for eliciting probabilities in the case of finite outcomes.
+This rule is simply $S(p,x) = \log p(x)$, namely we take the log likelihood of the reported density at the eventual outcome.
+To generalize the rule to expected statistics rather than full densities, we consider a subset of densities $\mD \subseteq \mP$.
 If there is a bijection between the sets $\mD$ and $\mM$, then we say that $\mM$ parametrizes $\mD$ and write $p(\cdot\,;\mu)$ for the density mapping to $\mu$.
-
-Note that this definition of a scoring rule places only mild information requirements on the part of the agent to ensure truthful reporting.
-Because condition~(\ref{eq-proper}) holds for all $p$ consistent with expectation $\mu$, it is enough for the agent to simply know the latter and not the complete density to be properly incentivized.
-However, the agent must also agree that the density is actually drawn from domain $\mD$; if the scoring rule is proper, it is enough that the agent agree with the support of the density as implicitly defined by base measure $\nu$.
-
-When the outcome space is finite we can recover classical scoring rules from the definition by taking $\mD = \mP$ and using the statistic $\phi : \mX \rightarrow \{0,1\}^{\mX}$ that maps an outcome $x$ to a unit vector with a 1 in the component corresponding to $x$.
-The expectation of $\phi$ is then exactly the probability mass function.
-
-\subsection{Proper Scoring from Consistency}
-
-Our starting point for designing proper scoring rules is the classic logarithmic scoring rule for eliciting probabilities in the case of finite outcomes.
-This rule is simply $S(p,x) = \log p(x)$ and there are several ways to understand why it is proper, with different implications for generalization.
-
-First observe that the scoring rule compensates the agent with the log likelihood it assigned to the outcome, so the agent is maximizing the expected log likelihood via its report.
-Now it is a fundamental result in statistics that, under a variety of different sufficient conditions, the maximum likelihood estimator is statistically consistent; see~\cite{} for an overview of such conditions.
-This means that in the limit (as the size of the i.i.d.\ empirical sample increases) the log likelihood is maximized by the true parameters.
-By this reasoning, the following result follows immediately from the definition of consistency and the law of large numbers.
-%
-\begin{theorem} \label{max-like}
-Suppose that $\mM$ parametrizes $\mD$. The logarithmic scoring rule defined by
+Given such a family parametrized be the relevant statistics, the generalized log scoring rule is then
 %
 \begin{equation} \label{log-score}
-S(\mu, x) = \log p(x;\mu),
-\end{equation}
-%
-where $\mu \in \mM$, is proper over $\mD$ if and only if the maximum likelihood estimator for $\mu$ is consistent.
-\end{theorem}
-%
-We note that although our focus is on parametrizations by expectations, the preceding theorem in fact holds for more generic parameter spaces.
-The range of scoring rules it provides is broader than those characterized by~\cite{}, because it allows for a restriction over the set of beliefs.
-On the other hand, its applicability is limited to those domains $\mD$ that can be parametrized by the relevant statistics.
-The following provide concrete examples of these points.
-%
-\begin{example}
-Suppose the outcomes are supported on $[1,+\infty)$ and follow a Pareto distribution with density $f(x;\alpha) = \alpha/x^{\alpha+1}$ parametrized by an index $\alpha > 0$.
-The mean $\mu$ is related to the index via the one-to-one mapping $\mu = \frac{\alpha}{\alpha-1}$, so the density can alternatively be parametrized by the mean.
-Theorem~\ref{max-like} gives the following proper scoring rule for the mean of a Pareto distribution with support on $[1,+\infty)$:
-%
-\begin{equation} \label{pareto-scoring}
-S(\mu,x) = \log \frac{\mu}{\mu-1} - \left(\frac{\mu}{\mu-1} + 1\right) \log x.
-\end{equation}
-%
-We stress that the rule can only elicit the mean assuming the agent knows its belief is a Pareto distribution over $[1,+\infty]$.
-It does not elicit the mean of other families of densities parametrized by the mean (e.g., the exponential distribution).
-\end{example}
-%
-%
-\begin{example}
-Suppose the outcomes are supported on $(-\infty,+\infty)$ and follow a Cauchy distribution with density $f(x;m) = 1/\pi[1+(x-m)^2]$, parametrized by the median $m$.
-Theorem~\ref{max-like} leads to the following proper scoring rule for the median:
-%
-\begin{equation} \label{cauchy-scoring}
-S(m,x) = -\log[1+(x-m)^2].
-\end{equation}
-%
-Note that the median cannot be obtained as the expectation of any statistic. This highlights the range of parameters than may be elicited by the log score in different circumstances.
-\end{example}
-%
-
-\subsection{Proper Scoring from Maximum Entropy}
-
-We now turn to scoring rules that are proper over the entire domain of densities $\mP$.
-Our construction is the same as in the previous section: we take the log likelihood of a density parametrized by the relevant statistics.
-If the family of densities is chosen appropriately, the resulting rule will in fact hold over all $\mP$.
-To achieve this we draw on a well-known duality between maximum likelihood and maximum entropy.
+S(\mu, x) = \log p(x;\mu).
+\end{equation}
+%
+Even though the log score is only applied to densities from $\mD$, according to Definition~\ref{def:proper-score} it must work over all densities in $\mP$.
+It turns out this is possible if $\mD$ is chosen appropriately, drawing on a well-known duality between maximum likelihood and maximum entropy~\citep{}.
 
@@
 %
-Note that the explicit set of constraints in~(\ref{maxent-prog}) are linear, and to stress this fact we find it helpful to re-write them as $A_{\phi} p = \mu$, where $A_{\phi}$ is the expectation operator of statistics $\phi$.
-We let $G : \mM \rightarrow \bR$ be the optimal value function of~(\ref{maxent-prog}), meaning $G(\mu)$ is the (negative) entropy of the maximum entropy distribution with expected statistics $\mu$.
+Note that the explicit set of constraints in~(\ref{maxent-prog}) are linear, whereas the objective is convex.
+We let $G : \mM \rightarrow \bR$ be the optimal value function of~(\ref{maxent-prog}), meaning $G(\mu)$ is the negative entropy of the maximum entropy distribution with expected statistics $\mu$.
 
@@
 %
-The density is stated here in terms of its \emph{natural} parametrization $\theta \in \bR^d$, where $\theta$ arises as the Lagrange multiplier associated with linear constraints $A_{\phi}p = \mu$.
+The density is stated here in terms of its \emph{natural} parametrization $\theta \in \bR^d$, where $\theta$ arises as the Lagrange multiplier associated with the linear constraints in~(\ref{maxent-prog}).
 The term $T(\theta)$ essentially arises as the multiplier for the normalization constraint (the density must integrate to 1), and so ensures that~(\ref{exp-fam}) is normalized:
@@
 
-The following proposition collects the relevant results on regular exponential families and their associated parameter domains and functions.
+The following proposition collects the relevant results on regular exponential families; proofs may be found in~\citep{}.
 A convex function $T$ is of \emph{Legendre type} if it is proper, closed, strictly convex and differentiable on the interior of its domain, and $\lim_{\theta \rightarrow \bar{\theta}} \norm{\grad T(\theta)} = +\infty$ when $\bar{\theta}$ lies on the boundary of the domain.
@@
 %
-In the above $T^*$ denotes as usual the convex conjugate of $T$, which here can be evaluated as $T^*(\mu) = \sup_{\theta \in \Theta} \la \theta, \mu \ra - T(\theta)$.
+In the above $T^*$ denotes the convex conjugate of $T$, which here can be evaluated as $T^*(\mu) = \sup_{\theta \in \Theta} \la \theta, \mu \ra - T(\theta)$.
 Similarly, $G^*(\theta) = \sup_{\mu \in \mM} \la \theta, \mu \ra - G(\mu)$.
@@
 The following observation is crucial.
-Let $\tp \in \mP$ be a density (not necessarily from an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\mu' \in \mM$ be an alternative report.
+Let $\tp \in \mP$ be a density (not necessarily from an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\hmu \in \mM$ be an alternative report.
 Then note from~(\ref{exp-fam}) that
@@
 \begin{equation} \label{equalizer-rule}
-\Exp_{\tp}[\log p(x;\mu')] = \Exp_{p(\cdot;\mu)}[\log p(x;\mu')] = \la \theta', \mu \ra - T(\theta'),
-\end{equation}
-%
-where $\theta' = \grad G(\mu')$ is the natural parameter for the exponential family with statistic $\mu'$.
+\Exp_{\tp}[\log p(x;\hmu)] = \Exp_{p(\cdot;\mu)}[\log p(x;\hmu)] = \la \htheta, \mu \ra - T(\htheta),
+\end{equation}
+%
+where $\htheta = \grad G(\hmu)$ is the natural parameter for the exponential family with statistic $\hmu$.
 We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, which is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
@@
 Consider the logarithmic scoring rule $S(\mu,x) = \log p(x;\mu)$ defined over a set of densities $\mD$ parametrized by $\mM$.
-The scoring rule is proper over the entire domain of densities $\mP$ if and only if $\mD$ is the exponential family with statistic $\phi$.
+The scoring rule is proper if and only if $\mD$ is the exponential family with statistic $\phi$.
 \end{theorem}
@@
 \begin{proof}
-Let $\mu,\mu' \in \mM$ be the agent's true belief and an alternative report, and let $p,p' \in \mP$ be densities consistent with each belief.
-Let $\theta = \grad G(\mu)$ and $\theta' = \grad G(\mu')$, and note that $\mu = \grad T(\theta)$.
+Let $\mu,\hmu \in \mM$ be the agent's true belief and an alternative report, and let $p \in \mP$ be a density consistent with $\mu$.
+Let $\theta = \grad G(\mu)$ and $\htheta = \grad G(\hmu)$, and note that $\mu = \grad T(\theta)$.
 We have
@@
 \begin{eqnarray}
-& & \Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\mu')] \nonumber\\
-& = & \la \theta, \mu \ra - T(\theta) - \la \theta', \mu \ra + T(\theta') \nonumber\\
-& = & T(\theta') - T(\theta) - \la \theta' - \theta, \mu \ra \nonumber\\
-& = & T(\theta') - T(\theta) - \la \theta' - \theta, \grad T(\theta) \ra. \label{breg-div}
+& & \Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\hmu)] \nonumber\\
+& = & \la \theta, \mu \ra - T(\theta) - \la \htheta, \mu \ra + T(\htheta) \nonumber\\
+& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \mu \ra \nonumber\\
+& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \grad T(\theta) \ra. \label{breg-div}
 \end{eqnarray}
@@
 The latter is positive by the strict convexity of $T$, which shows that the log score is proper.
-Now assume the defined log score is proper.
+For the converse, assume the defined log score is proper.
 By the Savage characterization of proper scoring rules for expectations (see~\cite{}), we must have
@@
 
-As further intuition for the result, note that~(\ref{breg-div}) is the definition of the so-called Bregman divergence with respect to strictly convex function $T$. Therefore we have 
+As further intuition for the result, note that~(\ref{breg-div}) is the definition of the `Bregman divergence' with respect to strictly convex function $T$~\citep{}. Therefore we have 
 %
 \begin{equation*}
-\Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\mu')] = D_T(\theta',\theta) = D_G(\mu, \mu'),
+\Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\hmu)] = D_T(\htheta,\theta) = D_G(\mu, \hmu),
 \end{equation*}
@@
 
+\subsection{Examples: Moments over the Real Line}
+
 Theorem~\ref{maxent-score} leads to a straightforward procedure for constructing score rules for expectations: define the relevant statistic, and consider the maximum entropy (equivalently, exponential family) distribution consistent with the agent's reported mean $\mu$.
-The scoring rule compensates the agent according to the log likelihood of the eventual outcome according to the latter.
+The scoring rule compensates the agent according to the log likelihood of the eventual outcome according to this distribution.
 The interpretation is that the agent is only providing partial information about the underlying density, so the principal first infers a full density according to the principle of maximum entropy, and then scores the agent using the usual log score.
 
-An advantage of this generalization of the log score is that, for many expectations of interest, it leads to simple \emph{closed-form} formulas for the scoring rule (including for multi-dimensional outcomes in some cases).
-This covers for instance several different combinations of the arithmetic, geometric, and harmonic means, as well as higher order moments.
+An advantage of this generalization of the log score is that, for many domains (multi-dimensional included) and expectations of interest, it leads to novel closed-form scoring rules.
+By examining the log densities of various exponential families, we can for instance obtain scoring rules for several different combinations of the arithmetic, geometric, and harmonic means, as well as higher order moments.
 The following examples illustrate the construction.
@@
 %
-We stress that although this rule is derived from the exponential distribution, Theorem~\ref{maxent-score} implies that it elicits the mean of any distribution supported on the non-negative reals (e.g., Pareto, lognormal), in contrast to~(\ref{pareto-scoring}).
+We stress that although this rule is derived from the exponential distribution, Theorem~\ref{maxent-score} implies that it elicits the mean of any distribution supported on the non-negative reals (e.g., Pareto, lognormal).
 Indeed, it is easy to see that the expected score~(\ref{exp-scoring}) depends only on the mean of the agent's belief because it is linear in $x$.
@@
 %
-\begin{equation*}
+\begin{equation} \label{weibull-scoring}
 S(\mu, x) = (k-1) \log x - k \log \mu - \Gamma\left(1+\frac{1}{k}\right)^k \left(\frac{x}{\mu}\right)^k,
-\end{equation*}
-%
-where $\Gamma$ denotes the usual gamma function (the extension of the factorial to the reals).
+\end{equation}
+%
+where $\Gamma$ denotes the usual gamma function (the extension of the factorial to the reals). We have not found either scoring rule~(\ref{exp-scoring}) or~(\ref{weibull-scoring}) in the literature.
 \end{example}
@@
 \begin{equation}
-S(({\boldsymbol \mu}, {\boldsymbol \Sigma}), \boldsymbol{x}) = -( \boldsymbol{x} - {\boldsymbol \mu})'{\boldsymbol \Sigma}^{-1}( \boldsymbol{x} - {\boldsymbol \mu}) - \log |{\boldsymbol \Sigma}|.
-\end{equation}
-%
-Here the statistics being elicited are the mean vector ${\boldsymbol \mu}$ and the covariance matrix ${\boldsymbol \Sigma}$.
-These scoring rules have been studied by Dawid and Sebastiani~\cite{} as rules that only depend on the mean and variance of the reported density.
+S((\mu, \Sigma), x) = -(x - \mu)'\Sigma^{-1}(x - \mu) - \log |\Sigma|.
+\end{equation}
+%
+Here the statistics being elicited are the mean vector $\mu$ and the covariance matrix $\Sigma$.
+These scoring rules have been studied by~\citet{} as rules that only depend on the mean and variance of the reported density.
 They note that these rules are weakly proper (because they do not distinguish between densities with the same first and second moments), but do not make the point that knowledge of the full density is not necessary on the part of the agent.
@@
 A risk-neutral agent will choose to acquire shares up to the point where, for each share, expected payoff equals marginal price.
-Formally, if the agent acquires portfolio $\delta$, moving the market state vector to $\theta' = \theta + \delta$, then we must have
+Formally, if the agent acquires portfolio $\delta$, moving the market state vector to $\htheta = \theta + \delta$, then we must have
 %
 \begin{equation} \label{risk-neutral-agent}
-\Exp_p[\phi(x)] = \grad C(\theta').
-\end{equation}
-%
-In this way, by its choice of $\delta$, the agent reveals that its expected belief is $\mu = \grad C(\theta')$.
+\Exp_p[\phi(x)] = \grad C(\htheta).
+\end{equation}
+%
+In this way, by its choice of $\delta$, the agent reveals that its expected belief is $\mu = \grad C(\htheta)$.
 We stress that this observation relies on the assumptions that 1) the agent is risk-neutral, 2) the agent does not incorporate the market's information into its own beliefs, and 3) the agent is not budget constrained.
@@
 \begin{equation*}
-\Exp_{\tp}[\log p(x;\mu')] = \la \theta', \mu \ra - T(\theta')
+\Exp_{\tp}[\log p(x;\hmu)] = \la \htheta, \mu \ra - T(\htheta)
 \end{equation*}
 %
-where $\tp$ is the agent's belief and $\mu'$ the agent's report.
-The expected log score from reporting $\mu'$ is exactly the same as the expected payoff from buying portfolio of shares $\theta' = \grad C(\mu')$ (assuming an initial market state of 0), as $\la \theta', \mu \ra$ is the expected revenue and $T(\theta')$ is the cost.
+where $\tp$ is the agent's belief and $\hmu$ the agent's report.
+The expected log score from reporting $\hmu$ is exactly the same as the expected payoff from buying portfolio of shares $\htheta = \grad C(\hmu)$ (assuming an initial market state of 0), as $\la \htheta, \mu \ra$ is the expected revenue and $T(\htheta)$ is the cost.
 As in Section~\ref{sec:scoring} this reasoning relies on the assumption of risk-neutrality (and no budget constraints), not on any specific form for the agent's belief.
@@
 %\pagebreak
+
+\subsection{Proper Scoring from Consistency}
+
+Our starting point for designing proper scoring rules is the classic logarithmic scoring rule for eliciting probabilities in the case of finite outcomes.
+This rule is simply $S(p,x) = \log p(x)$ and there are several ways to understand why it is proper, with different implications for generalization.
+
+First observe that the scoring rule compensates the agent with the log likelihood it assigned to the outcome, so the agent is maximizing the expected log likelihood via its report.
+Now it is a fundamental result in statistics that, under a variety of different sufficient conditions, the maximum likelihood estimator is statistically consistent; see~\cite{} for an overview of such conditions.
+This means that in the limit (as the size of the i.i.d.\ empirical sample increases) the log likelihood is maximized by the true parameters.
+By this reasoning, the following result follows immediately from the definition of consistency and the law of large numbers.
+%
+\begin{theorem} \label{max-like}
+Suppose that $\mM$ parametrizes $\mD$. The logarithmic scoring rule defined by
+%
+\begin{equation} \label{log-score}
+S(\mu, x) = \log p(x;\mu),
+\end{equation}
+%
+where $\mu \in \mM$, is proper over $\mD$ if and only if the maximum likelihood estimator for $\mu$ is consistent.
+\end{theorem}
+%
+We note that although our focus is on parametrizations by expectations, the preceding theorem in fact holds for more generic parameter spaces.
+The range of scoring rules it provides is broader than those characterized by~\cite{}, because it allows for a restriction over the set of beliefs.
+On the other hand, its applicability is limited to those domains $\mD$ that can be parametrized by the relevant statistics.
+The following provide concrete examples of these points.
+%
+\begin{example}
+Suppose the outcomes are supported on $[1,+\infty)$ and follow a Pareto distribution with density $f(x;\alpha) = \alpha/x^{\alpha+1}$ parametrized by an index $\alpha > 0$.
+The mean $\mu$ is related to the index via the one-to-one mapping $\mu = \frac{\alpha}{\alpha-1}$, so the density can alternatively be parametrized by the mean.
+Theorem~\ref{max-like} gives the following proper scoring rule for the mean of a Pareto distribution with support on $[1,+\infty)$:
+%
+\begin{equation} \label{pareto-scoring}
+S(\mu,x) = \log \frac{\mu}{\mu-1} - \left(\frac{\mu}{\mu-1} + 1\right) \log x.
+\end{equation}
+%
+We stress that the rule can only elicit the mean assuming the agent knows its belief is a Pareto distribution over $[1,+\infty]$.
+It does not elicit the mean of other families of densities parametrized by the mean (e.g., the exponential distribution).
+\end{example}
+%
+%
+\begin{example}
+Suppose the outcomes are supported on $(-\infty,+\infty)$ and follow a Cauchy distribution with density $f(x;m) = 1/\pi[1+(x-m)^2]$, parametrized by the median $m$.
+Theorem~\ref{max-like} leads to the following proper scoring rule for the median:
+%
+\begin{equation} \label{cauchy-scoring}
+S(m,x) = -\log[1+(x-m)^2].
+\end{equation}
+%
+Note that the median cannot be obtained as the expectation of any statistic. This highlights the range of parameters than may be elicited by the log score in different circumstances.
+\end{example}
+%

### 2014-02-08 WriteUp/maxentpredmkts.tex
revision of maxent market section

@@
 In a single-agent setting, a scoring rule is used to \emph{elicit} the agent's belief.
-In a multi-agent setting, an information market can be used to \emph{aggregate} the beliefs of the agents.
+In a multi-agent setting, a prediction market can be used to \emph{aggregate} the beliefs of the agents.
 In his seminal paper~\citet{} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both in order to perform well in thin or thick markets.
-In this section, we adapt the generalized log scoring rule to a market scoring rule which leads to markets with simple closed-form cost functions for many statistics of interest.
-The exponential families framework also allows us to move beyond risk-neutral agents to understand the aggregation properties of information markets under alternative behaviors, such as Bayesian updating and risk aversion.
-
-\subsection{Information Market}
-
-In an information market an agent's expected belief $\mu$ is elicited indirectly through the purchase and sale of contingent claim securities.
+In this section, we adapt the generalized log scoring rule to a market scoring rule which leads to markets with simple closed-form cost functions for many statistics of interest.%The exponential families framework also allows us to move beyond risk-neutral agents to understand the aggregation properties of prediction markets under alternative behaviors, such as Bayesian updating and risk aversion.
+
+\subsection{Prediction Market}
+
+In a prediction market an agent's expected belief $\mu$ is elicited indirectly through the purchase and sale of contingent claim securities.
 Under this approach, each component~$i$ of the statistic $\phi$ is interpreted as the payoff function of a security; that is, a single share of security $i$ pays off $\phi_i(x)$ when $x \in \mX$ occurs.
@@
 
-The standard way to implement an information market in the literature, due to~\citet{}, is via a centralized market maker.
+The standard way to implement a prediction market in the literature, due to~\citet{}, is via a centralized market maker.
 The market maker maintains a convex, differentiable cost function $C : \bR^d \rightarrow (-\infty,+\infty]$, where $C(\theta)$ records the revenue collected when the vector of outstanding shares is $\theta$.
@@
 A risk-neutral agent will choose to acquire shares up to the point where, for each share, expected payoff equals marginal price.
-Formally, if the agent acquires portfolio $\delta$, moving the market state vector to $\htheta = \theta + \delta$, then we must have
+Formally, if the agent acquires portfolio $\delta$, moving the market state vector to $\theta' = \theta + \delta$, then we must have
 %
 \begin{equation} \label{risk-neutral-agent}
-\Exp_p[\phi(x)] = \grad C(\htheta).
-\end{equation}
-%
-In this way, by its choice of $\delta$, the agent reveals that its expected belief is $\mu = \grad C(\htheta)$.
+\Exp_p[\phi(x)] = \grad C(\theta').
+\end{equation}
+%
+In this way, by its choice of $\delta$, the agent reveals that its expected belief is $\mu = \grad C(\theta')$.
 We stress that this observation relies on the assumptions that 1) the agent is risk-neutral, 2) the agent does not incorporate the market's information into its own beliefs, and 3) the agent is not budget constrained.
@@
 The expected log score from reporting $\hmu$ is exactly the same as the expected payoff from buying portfolio of shares $\htheta = \grad C(\hmu)$ (assuming an initial market state of 0), as $\la \htheta, \mu \ra$ is the expected revenue and $T(\htheta)$ is the cost.
-As in Section~\ref{sec:scoring} this reasoning relies on the assumption of risk-neutrality (and no budget constraints), not on any specific form for the agent's belief.
+As in Section~\ref{sec:scoring} this reasoning relies on the assumption of risk-neutrality, not on any specific form for the agent's belief.
 
@@
 \begin{example} \label{ex:exp-market}
-We consider outcomes over the positive reals $bR_+$ and set up a market for the expected outcome, consisting of a single security that pays off $\phi(x) = x$.
-The log partition funciton of the exponential distribution leads to the following cost function:
+We consider outcomes over the positive reals $\bR_+$ and set up a market for the expected outcome, consisting of a single security that pays off $\phi(x) = x$.
+The log partition function of the exponential distribution leads to the following cost function:
 %
@@
 The effective domain is $\Theta = \{\theta \in \bR : \theta < 0\}$.
-This means the market must start with a negative number of outstanding shares for the security, and the number shares must stay negative.
+This means the market must start with a negative number of outstanding shares for the security, and the number of shares must stay negative.
 The market maker need not explicitly enforce this, because by the Legendre property of $C$ the cost tends to $+\infty$ as the outstanding shares approach the boundary, which is straightforward to see in this example.
@@
 %
-where $I_r$ refers to the modified Bessel function of first kind and order $r$.
+where $I_r$ refers to the modified Bessel function of first kind and order $r$; see~\citep{} for an explanation of these quantities.
 The effective domain of $\theta$ is the positive orthant in $\bR^d$.

### 2014-02-09 WriteUp/liquidity.tex
added comments to liquidity write-up

@@
 
+\subsection{Introducing Liquidity (S\'{e}bastien Remarks)}
+
+We have been assuming that the cost function is the same as the log partition function, but this assumes a liquidity setting of 1.
+Consider the parametrized  cost $C_{\lambda}(\theta)=\frac{1}{\lambda} C\left(\lambda\theta\right)$.
+Here $\lambda$ is construed as the \emph{inverse liquidity}, or price responsiveness.
+Higher $\lambda$ means fewer shares need to be bought to reach the same prices.
+
+(Discussion, you can skip this paragraph.) This is because for higher $\lambda$ we are making the cost function ``more convex''.
+Specifically, for $\lambda_1 > \lambda_2$ it's the case that $C_{\lambda_1}$ is an increasing convex transformation of $C(\lambda_2)$.
+This holds because the ``coefficient of risk aversion'' $A(\theta) = C''(\theta)/C'(\theta)$ increases with $\lambda$ at every point.
+I've written it assuming single-dimensional $\theta$ but there are more complicated looking generalizations for vector $\theta$.
+In fact, this coefficient might be the ``right'' notion of liquidity at a point, not the second derivative or largest eigenvalue of the Hessian.
+We should discuss.
+
+\newcommand{\ttheta}{\tilde{\theta}}
+As we're dealing with LMSR I write $T$ rather than $C$ now where $T$ is the log partition function.
+The cost function is the liquidity-adjusted version $T_{\lambda}$.
+Let $\hmu$ be the agent's mean belief and let $\htheta = \grad T^{-1}(\hmu)$ be the corresponding natural parameter.
+A risk-neutral agent moves the share vector so that the prices match its mean parameter.
+Therefore, define the \emph{target shares} as $\ttheta = \grad T^{-1}_{\lambda}(\hmu)$.
+So $\htheta$ and $\ttheta$ are related by
+%
+$$
+\grad T(\htheta) = \grad T_{\lambda}(\ttheta).
+$$
+%
+It's easy to check that $\grad T_{\lambda}(\ttheta) = \grad T(\lambda \ttheta)$ so we have
+%
+\begin{equation} \label{target}
+\ttheta = \htheta / \lambda.
+\end{equation}
+%
+Now using the liquidity-adjusted cost function, Equation~(17) from Theorem~6.2 in the write-up becomes the following, where $\theta$ is the current market state:
+%
+\begin{equation} \label{update}
+\frac{\lambda}{\lambda+a} \ttheta + \frac{a}{\lambda+a} \theta.
+\end{equation}
+%
+As $\lambda$ grows large, the agent moves the market state closer to the \emph{target shares}, rather than its true natural parameter.
+Note that the target shares themselves depend on $\lambda$, because of~(\ref{target}).
+This make sense because higher price responsiveness means the agent needs to buy fewer shares to make the market prices match its own expectations.
+In that sense, it may be conceptually cleaner to write~(\ref{update}) as
+%
+\begin{equation}
+\frac{\lambda}{\lambda+a} \grad T^{-1}_{\lambda}(\hmu) + \frac{a}{\lambda+a} \grad T^{-1}_{\lambda}(\mu).
+\end{equation}
+%
+where $\mu$ are the current market prices.
+
 \end{document}

### 2014-02-09 WriteUp/maxentpredmkts.tex
polished up Bayesian traders section

@@
 \subsection{Prediction Market}
+\label{sec:pred-market}
 
@@
 
-%\pagebreak
 \section{Bayesian Traders with Linear Utility}
-%\subsection{Conjugate Priors}
-
-%In the previous section it's assumed that the agent has a belief $\q$, and moves the market state from initial state $\q_{init}$ to a convex combination $\tilde{\q} = \lambda \q + (1-\lambda)\q_{init}$.
-This note tries to justify this based on Bayesian updating, rather than budget limitations.
-(But it doesn't quite succeed.)
-In this section, we lay the foundation for interpreting trader behavior in terms of Bayesian updates.
-
-As before, we are interested in eliciting the sufficient statistics of the data.
+
+In the standard model of cost-function based prediction markets, a sequence of myopic, risk-neutral agents arrive and trade in the market~\citep{}.
+As we saw in Section~\ref{sec:pred-market}, such a trader moves the prices to its own expectation $\mu$.
+However, this means that the market does not perform any meaningful aggregation of the agent's belief, as the final prices are simply the final agent's expectation.
+
+In this section we examine the aggregation behavior of the market when agents are Bayesian and take into account the current market state when forming their beliefs.
+This requires more structure to the agents' beliefs.
+For this section and the remainder of the paper, we will assume that agents have \emph{exponential family beliefs}.
+
+%As before, we are interested in eliciting the sufficient statistics of the data.
 We assume that the outcome is drawn from an exponential family distribution; the prediction market is setup as before with the cost function corresponding to the log partition function and the payoff function corresponding to the sufficient statistics.
@@
 
-The goal is to aggregate information from risk neutral agents who have a belief distribution over the natural parameters.
+%The goal is to aggregate information from risk neutral agents who have a belief distribution over the natural parameters.
 This prior distribution is updated by the agents based on the current market state, They also each have to access to the empirical mean of sufficient statistics based on a fixed number  $m$ of data points.
@@
 
-Now we will define the setup more formally.
-
-Let the data distribution be given by $p(x;\beta)=\exp\{\beta\cdot\phi(x)-T(\beta)\}$ where $T()$ is the log partition function and $\phi()$ are the sufficient statistics.
-Then the conjugate prior parametrized by $b_{0}=(n\nu, n)$ is given by $p(\beta;b_{0})=\exp\{b_{0}\cdot(\beta,\psi(\beta))-\psi(b_{0})\}$where $\psi()$ is the corresponding log partition function.
-The posterior distribution on the natural parameters is given by $p(\beta;b)$ where $b=(n\nu+m\hmu, n+m)$ where $\hmu$ is the empirical mean of the sufficient statistics of the $m$ data points drawn from the data distribution; that is, $\hmu=\sum_{i=1}^{m}\phi(x_{i})$.
-It turns out that
+The exponential families framework is well-suited to reasoning about Bayesian updates.
+As before let the data distribution be given by $p(x;\theta)=\exp(\la\theta,\phi(x)\ra-T(\theta))$ where $T$ is the log partition function and $\phi$ are the sufficient statistics.
+Instead of direct beliefs about the data distribution the agent maintains a conjugate prior over the parameters $\theta$.
+Every exponential family admits a conjugate prior of the form
+%
+$$p(\theta;b_{0})=\exp(\la n\nu,\theta\ra + nT(\theta) - \psi(\nu,n)).$$
+%
+Note that this is also an exponential family with natural parameter $b_0 = (n\nu,n)$ where $\nu \in \bR^d$ and $n$ is a positive integer.
+The sufficient statistic maps $\theta$ to $(\theta, T(\theta))$, and the log partition function $\psi$ is defined as the normalizer as usual.
+For a complete treatment of exponential families conjugate priors, see for instance~\citep{}.
+Now~\citet{} and~\citet{} have shown that
+%
 \begin{equation}\label{eq:prior}
-\E_{\beta\sim b_{0}}\E_{x\sim \beta}[\phi(x)]=\nu
-\end{equation}¥ Thus, it is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$.
+\E_{\theta\sim b_{0}}\E_{x\sim \theta}[\phi(x)]=\nu,
+\end{equation}
+%
+meaning that $\nu = n \nu / n$ is the posterior mean.
+Thus, it is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$.
+Suppose now that the agent observes an empirical sample with mean $\hmu$ and size $m$.
+By a standard derivation~\citep{}, the posterior conjugate prior parameters become $n\nu \leftarrow n\nu + m\hmu$ and $n \leftarrow n+m$, and the posterior expectation~(\ref{eq:prior}) evaluates to
+%
+\begin{equation} \label{posterior}
+\frac{n\nu + m\hmu}{n+m}.
+\end{equation}
+%
 Thus the posterior mean is a convex combination of the prior and posterior means, and their relative weights depend on the phantom and empirical sample sizes.% need a reference here
 
-Suppose the current market state is $\theta$ and $i$ traders have traded in the market when trader $i+1$ with prior belief distribution $p(\beta;b^{i}_{0})$ enters the market.
+Consider Bayesian agents maintaning and exponential family conjugate prior over the data model's natural parameters (equivalently, the expected security payoffs).
+Each agent has access to a private sample of the data of size $m$ with mean statistic $\hmu$.
+If $n$ agents have arrived before to trade, then the current market prices $\mu$ correspond to the phantom sample, and the phantom sample size is $nm$.
+After forming the posterior~(\ref{posterior}) with these substitutions, the (risk-neutral) agent purchases shares $\delta$ to move the current market share vector to
+%
+$$
+\grad C(\theta + \delta) = \frac{n\nu + \hmu}{n+1}.
+$$
+%
+As a result, the final market prices under this behavior are a simple average of the agent's mean parameters and the initial market prices.
+We note that to facilitate such belief updating, the market should post the number of trades since initialization.
+
+%Suppose the current market state is $\theta$ and $i$ traders have traded in the market when trader $i+1$ with prior belief distribution $p(\theta;b^{i}_{0})$ enters the market.
 Here $b^{i}_{0} = (n\nu_{i}, n)$.
@@
 Recall from Proposition \ref{prop:exp} that natural parameter $\theta$ corresponds to expected statistics $\nabla T(\theta)$.
-Thus, he updates his belief as $p(\beta;b_{i})$ where $b=(m\hmu+mi \nabla T(\theta) + n\nu_{i}, n+mi+m)$.
-
-Suppose the trader wishes to maximize his expected payoff.
-Then the number of shares $\delta_{i}$ that he purchases when the current market state is $\theta$ is given by $$\arg\max_{\delta_{i}}\E_{\beta\sim b_{i}}\E_{x\sim \beta}\left[ \delta_{i}\phi(x) - T(\delta_{i} + \theta) + T(\theta)\right]$$
-But, from Equation \ref{eq:prior} we have $\E_{\beta\sim b}\E_{x\sim \beta}[\phi(x)]=\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}$.
+Thus, he updates his belief as $p(\theta;b_{i})$ where $b=(m\hmu+mi \nabla T(\theta) + n\nu_{i}, n+mi+m)$.
+
+%Suppose the trader wishes to maximize his expected payoff.
+Then the number of shares $\delta_{i}$ that he purchases when the current market state is $\theta$ is given by $$\arg\max_{\delta_{i}}\E_{\theta\sim b_{i}}\E_{x\sim \theta}\left[ \delta_{i}\phi(x) - T(\delta_{i} + \theta) + T(\theta)\right]$$
+%But, from Equation \ref{eq:prior} we have $\E_{\theta\sim b}\E_{x\sim \theta}[\phi(x)]=\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}$.
 To obtain the maximum, we set the gradient of the above expression with respect to $\delta_{i}$ to $0$.
-Thus, we have for the optimal number of shares $\delta_{i}^{*}$ $$\nabla T(\delta_{i}^{*} + \theta)=\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}$$ Thus, from Proposition \ref{prop:exp} we have that for an exponential family prediction market, the final market state is given by
-$$\nabla G\left(\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}\right)$$ where $G$ is the convex conjugate of $T$.
+Thus, we have for the optimal number of shares $\delta_{i}^{*}$ $$\nabla T(\delta_{i}^{*} + \theta)=\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}$$ Thus, from Proposition \ref{prop:exp} we have that for an exponential family prediction market, the final market state is given by $$\nabla G\left(\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}\right)$$ where $G$ is the convex conjugate of $T$.
 Thus, the final market state is a convex combination of prior, posterior and market means.

### 2014-02-10 WriteUp/maxentpredmkts.tex
moved Bayesian section to earlier in paper
fixed some preamble stuff
moved sections around, added ability to make comments

@@
 % Metadata Information
-\acmVolume{9}
-\acmNumber{4}
-\acmArticle{39}
-\acmYear{2013}
-\acmMonth{6}
+\acmVolume{X}
+\acmNumber{X}
+\acmArticle{X}
+\acmYear{2014}
+\acmMonth{2}
+
 \usepackage[numbers]{natbib}
@@
 
+\newcommand{\Comments}{1}
+\newcommand{\mynote}[2]{\ifnum\Comments=1\textcolor{#1}{#2}\fi}
+\newcommand{\jake}[1]{\mynote{blue}{[JA: #1]}}
+\newcommand{\sindhu}[1]{\mynote{red}{[SK: #1]}}
+\newcommand{\sebastien}[1]{\mynote{magenta}{[SL: #1]}}
+
 % Document starts
@@
 
-\terms{Prediction Market Design, Machine Learning Algorithms}
-
-\keywords{Exponential Families, Bayesian Learning, Maximum Entropy Distributions, generalized LMSR}
+\terms{Algorithms, Economics}
+
+\keywords{logarithmic score, exponential family, maximum entropy, risk aversion, budget constraints}
 
@@
 %- multiple traders
+
+\jake{This is a comment from jake}
+\sindhu{This is a comment from sindhu}
+\sebastien{This is a comment from sebastien}
 Prediction markets are aggregation mechanisms that allow market prices to be interpreted as predictive probabilities on an event.
@@
 
-(JAKE: I want to use this paragraph for a general overview of mm frameworks) One popular form of prediction markets is the market scoring rule~\cite{hanson03}.
+\jake{I want to use this paragraph for a general overview of mm frameworks} One popular form of prediction markets is the market scoring rule~\cite{hanson03}.
 A market scoring rule considers all trades as a single chronological sequence.
@@
 \subsection{Related Work}
+\jake{I'll add more here. Sebastien: do you have anything to put in here regarding the exponential utility stuff?}
 Designing prediction markets to handle a large outcome space is an active area of research.
@@
 %
-
-\section{The Exponential Family Market Mechanism with Budgets}
-In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
-
-In this section, we consider the the prediction market setup with traders that may be either informative or malicious.
-The malicious traders may want to inject faulty information into the market.
-The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.
-
-We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market.
-Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
-
-In this section, we assume that the traders have exponential family beliefs.
-The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
-\subsection{Budget-limited Aggregation}
-Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
-We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market.
-This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader.
-Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.
-
-In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
-Let $\alpha$ be the budget of a trader in the market.
-Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
-Let $C$ be the cost function.
-Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
-In this case, we want to budget-limit the trader's influence on the market state.
-
-We define the budget-limited final market state as $\theta'$. 
-Here, we consider a specific functional form of $\theta'$: $$\theta'=\lambda\htheta+(1-\lambda)\theta$$  where $$\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$$
-
-First, we  show that this trade is feasible given the trader's budget:
-\begin{theorem}
-Let the current market state be given by $\theta$.
-Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
-The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$.
-\end{theorem}
-\begin{proof}
-From the convexity of $C$, we have
- $$C(\theta')\leq(1-\lambda)C(\theta)+\lambda C(\htheta)$$
-
-Now 
-\begin{eqnarray*}
-C(\theta') - C(\theta) &\leq& (1-\lambda)C(\theta)\\
-&&+\lambda C(\htheta) - C(\theta)\\
-&=& \lambda\ (C(\htheta) - C(\theta))
-\end{eqnarray*}
-
- Thus, $C(\theta') - C(\theta) \leq \alpha$. 
- \end{proof}
-
- We note that moving to $\theta'$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
-In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\theta$.
-Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\htheta$.
-\subsection{Damage Bound}
-In this section, we will quantify the error in prediction that the market maker might have to endure as a result of malicious entities entering the market.
-We assume that these malicious entities trade in multiple instances of the market; thus the exposure of the market maker is over several \emph{rounds}.
-We use the standard log loss to measure this error in terms of the initial budget of traders.
-
-We define the loss function for $\theta$ shares held:
-$$L(\theta,x)=-\log(P_{\theta}(x))=\log\int \exp\{\theta^{T}\phi(x)\}\ dx-\theta^{T}\phi(x)=C(\theta)-\theta^{T}\phi(x)$$
-%with the correspondence between $\beta$ and $\theta$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
-
-Suppose the prediction market runs over multiple rounds $t$.
-Let $\theta_{0}^{t}$ be the initial number of shares of each security that are held.
-Let $\hat{\theta}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
-Let us assume that at this point the outcome is revealed; that is, we receive the value of the random variable $x^{t}$.
-
-Over multiple instances of the prediction market, we can track the change in budget of each trader.
-Let the budget at rounds $t$ and $t-1$ be $\alpha^{t}$ and $\alpha^{t-1}$ respectively.
-The change in budget for the trader is
-\begin{eqnarray*}
-\alpha^{t}-\alpha^{t-1}&=&C(\theta^{t})-C(\theta'^{t})-(\theta^{t}-\theta'^{t})^{T}\phi(x^{t})\\
-&=&L(\theta,x^{t})-L(\theta',x^{t})
-\end{eqnarray*}
-
-Define the myopic impact of a trader $i$ in segment $t$ as
-$$\Delta_{i}^{t}:=L(\hat{\theta}_{i-1}^{t},x^{t})-L(\hat{\theta}_{i}^{t},x^{t})$$
-Thus, the myopic impact captures incremental gain in prediction due to the trader in a round.
-Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
-
-The total myopic impact due to all $k$ active traders is given by
-$$\Delta^{t}=L(\theta_{0}^{t},x^{t})-L(\hat{\theta}_{k}^{t},x^{t})$$
-Thus $-\Delta^{t}$ captures the incremental loss of the market prediction after aggregation of all $k$ traders.
-
-\begin{theorem}
-A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
-\end{theorem}
-\begin{proof}
-Consider the myopic impact of a single trader $i$ after participating in the market $T$ times.
-Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
-$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
-
-Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
-\end{proof}
-
-This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders. %imp
-
-In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation. 
-%We provide an information-theoretic justification for this claim in Section \ref{sec:infth}. 
-The intuition behind this  is that a trader's prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
-
-\subsection{Budget of Informative Traders}\label{sec:inf}
-Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
-
-We now characterize the expected change in budget for an informative trader. The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
-\begin{theorem}\label{thm:growth}
-Let $\theta$ be the current market state in the exponential family prediction market.
-Suppose that an  informative trader with belief distribution  parametrized by $\theta'$  moves the market state to the budget-limited state $\htheta=\lambda \theta' + (1-\lambda) \theta$.
-Then, the expectation (over the trader's belief) of the trader's profit  is greater than zero whenever his budget is positive and his belief differs from the previous market position $\theta$.
-% Suppose that each informative trader gets a random sample of data, 
-%resulting in a sequence of trader beliefs $\theta_i$, and hence a sequence of budget-limited market positions
-%$\hat{\theta}_i$, before a final outcome $x$.
-Then, the expectation, over trader $i$'s belief distribution $\theta_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\hat{\theta}_{i-1}$.
-\end {theorem}
-\begin{proof}
-Let the cost function $C$ be equal to the log partition function $T$ of the belief distribution.
-The payoff is given by the sufficient statistics $\phi(x)$.
-Then, the trader's expected net payoff is given by
-\begin{eqnarray*}
-&& \E_{x\sim P_{\htheta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
-&=& T(\theta)-\theta\E_{x\sim P_{\htheta}}[\phi(x)]-(T(\theta')-\theta' \E_{x\sim P_{\htheta}}[\phi(x)])\\
-&=& T(\theta)-\theta\nabla T(\htheta)-(T(\theta')-\theta' \nabla T(\htheta))\\
-&=& T(\theta)-T(\htheta)-\nabla T(\htheta)(\theta-\htheta)-(T(\theta')-T(\htheta)- \nabla T(\htheta)(\theta'-\htheta))\\
-&=& D_{T}(\theta,\htheta)-D_{T}(\theta',\htheta)\\
-&\geq&\lambda D_{T}(\theta,\htheta)\geq0
-\end{eqnarray*}
-The second to last inequality holds since $D_{T}(\theta',\htheta)$ is convex in $\theta'$ and we have:
-\begin{eqnarray*}
-D_{T}(\theta',\htheta)&=&D_{T}\left(\lambda\htheta+(1-\lambda)\theta,\htheta\right)\\
-&\leq&\lambda D_{T}( \htheta ,\htheta)+ (1-\lambda) D_{T}( \theta ,\htheta)\\
-&=&(1-\lambda) D_{T}( \theta ,\htheta)
-%\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
-\end{eqnarray*}¥
-Thus, a trader who moves the market state can expect his profit to be positive and at least $\lambda D_{T}(\theta,\htheta)$.
- \end{proof}
-
-For continuous distributions with a density, the probability that a trader with private information will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs.
-This result suggests that, eventually, every informative trader will have the ability to influence the market state in accordance with his beliefs, without being budget limited.
-
-Notice that Theorem~\ref{thm:growth} only required that the market state to which the trader moves, be representable as a convex combination of the current market state and his belief.
-This means that the result holds for exponential utility traders aiming to maximize their utility.
-In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{T}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
-
-We note one important aspect of Theorem~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
-This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure.
-If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
-%\pagebreak
 
@@
 
-%\subsection{Exponential Utility}
-%Now suppose the trader has exponential utility, given by $$U(w) = -\frac{1}{a} \exp(-aw)$$ where $a$ is the coefficient of risk aversion.
-Suppose the trader wishes to maximize his expected utility.
-Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by
-%\begin{eqnarray*}
-%&&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}U\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]\\
-%&=&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}-\frac{1}{a}\exp\left[ -a\delta\phi(x) +a C(\delta + \theta) -a C(\theta)\right]
-%\end{eqnarray*}¥
-%
-%\begin{eqnarray*}
-%\E_{\beta\sim b}\E_{x\sim \beta}\exp\left[-a\delta\phi(x)\right]&=&\int_{\mathcal{B}}\int_{\mathcal{X}}\exp[-a\delta\phi(x)]\, \exp\{\beta\cdot\phi(x)-\psi(\beta)\}\, dx\\
-%&&\qquad\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\, d\beta\\
-%&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
-%&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\int_{\mathcal{X}}\exp[(\beta-a\delta)\cdot\phi(x)-\psi(\beta-a\delta)]\, dx\\
-%&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
-%&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\\
-%\end{eqnarray*}¥
-%\pagebreak
-
-\section{Risk Averse Traders with Exponential Utility}
+\section{Risk Aversion, Exponential Utility, and Market Equilibrium}
+
+\subsection{Behavior of Traders endowed with Exponential Utility}
 
@@
 
-\section{Reinterpreting repeated trades}
+\subsection{Reinterpreting repeated trades}
 In previous sections we have analyzed trader behavior as if it is his first entry into the market.
@@
 Based on this belief, let $\delta_{1}^{*}$ be the optimal vector of shares the agent decides to trade on first entering the market.
-Thus, his belief distribution is given by the density $$p(x;\htheta)=\exp(\la\hat{\theta}, \phi(x)\ra - T(\hat{\theta}))$$ where $T(\hat{\theta}) =\int_{\mathcal{X}}\exp\{\la\hat{\theta}, \phi(x)\ra\}\ dx$ is the log partition function and $\phi(x)$ are the sufficient statistics of the outcome $x\in \mathcal{X}$.
+Thus, his belief distribution is given by the density $p(x;\htheta)=\exp(\la\hat{\theta}, \phi(x)\ra - T(\hat{\theta})),$ where $T(\hat{\theta}) =\int_{\mathcal{X}}\exp\{\la\hat{\theta}, \phi(x)\ra\}\ dx$ is the log partition function and $\phi(x)$ are the sufficient statistics of the outcome $x\in \mathcal{X}$.
 On a subsequent entry into this market with market state $\theta'$, his optimal purchase $\delta_{2}^{*}$ is given by the solution of
@@
 
-\section{Equilibrium Market State for Exponential Utility Agents}
+\subsection{Equilibrium Market State for Exponential Utility Agents}
 We have shown that every exponential-utility maximizing trader picks the share vector $\delta$ so that the eventual market state can be represented as a convex combination of the initial market state and the natural parameter of his (exponential family) belief distribution.
@@
 Thus the equilibrium state is a convex combination of trader beliefs and initial market state.
+
+\section{The Exponential Family Market Mechanism with Budgets}
+In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
+
+In this section, we consider the the prediction market setup with traders that may be either informative or malicious.
+The malicious traders may want to inject faulty information into the market.
+The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.
+
+We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market.
+Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
+
+In this section, we assume that the traders have exponential family beliefs.
+The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
+\subsection{Budget-limited Aggregation}
+Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
+We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market.
+This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader.
+Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.
+
+In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
+Let $\alpha$ be the budget of a trader in the market.
+Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
+Let $C$ be the cost function.
+Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
+In this case, we want to budget-limit the trader's influence on the market state.
+
+We define the budget-limited final market state as $\theta'$. 
+Here, we consider a specific functional form of $\theta'$: $$\theta'=\lambda\htheta+(1-\lambda)\theta$$  where $$\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$$
+
+First, we  show that this trade is feasible given the trader's budget:
+\begin{theorem}
+Let the current market state be given by $\theta$.
+Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
+The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$.
+\end{theorem}
+\begin{proof}
+From the convexity of $C$, we have
+ $$C(\theta')\leq(1-\lambda)C(\theta)+\lambda C(\htheta)$$
+
+Now 
+\begin{eqnarray*}
+C(\theta') - C(\theta) &\leq& (1-\lambda)C(\theta)\\
+&&+\lambda C(\htheta) - C(\theta)\\
+&=& \lambda\ (C(\htheta) - C(\theta))
+\end{eqnarray*}
+
+ Thus, $C(\theta') - C(\theta) \leq \alpha$. 
+ \end{proof}
+
+ We note that moving to $\theta'$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
+In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\theta$.
+Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\htheta$.
+\subsection{Damage Bound}
+In this section, we will quantify the error in prediction that the market maker might have to endure as a result of malicious entities entering the market.
+We assume that these malicious entities trade in multiple instances of the market; thus the exposure of the market maker is over several \emph{rounds}.
+We use the standard log loss to measure this error in terms of the initial budget of traders.
+
+We define the loss function for $\theta$ shares held:
+$$L(\theta,x)=-\log(P_{\theta}(x))=\log\int \exp\{\theta^{T}\phi(x)\}\ dx-\theta^{T}\phi(x)=C(\theta)-\theta^{T}\phi(x)$$
+%with the correspondence between $\beta$ and $\theta$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
+
+Suppose the prediction market runs over multiple rounds $t$.
+Let $\theta_{0}^{t}$ be the initial number of shares of each security that are held.
+Let $\hat{\theta}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
+Let us assume that at this point the outcome is revealed; that is, we receive the value of the random variable $x^{t}$.
+
+Over multiple instances of the prediction market, we can track the change in budget of each trader.
+Let the budget at rounds $t$ and $t-1$ be $\alpha^{t}$ and $\alpha^{t-1}$ respectively.
+The change in budget for the trader is
+\begin{eqnarray*}
+\alpha^{t}-\alpha^{t-1}&=&C(\theta^{t})-C(\theta'^{t})-(\theta^{t}-\theta'^{t})^{T}\phi(x^{t})\\
+&=&L(\theta,x^{t})-L(\theta',x^{t})
+\end{eqnarray*}
+
+Define the myopic impact of a trader $i$ in segment $t$ as
+$$\Delta_{i}^{t}:=L(\hat{\theta}_{i-1}^{t},x^{t})-L(\hat{\theta}_{i}^{t},x^{t})$$
+Thus, the myopic impact captures incremental gain in prediction due to the trader in a round.
+Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
+
+The total myopic impact due to all $k$ active traders is given by
+$$\Delta^{t}=L(\theta_{0}^{t},x^{t})-L(\hat{\theta}_{k}^{t},x^{t})$$
+Thus $-\Delta^{t}$ captures the incremental loss of the market prediction after aggregation of all $k$ traders.
+
+\begin{theorem}
+A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
+\end{theorem}
+\begin{proof}
+Consider the myopic impact of a single trader $i$ after participating in the market $T$ times.
+Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
+$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
+
+Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
+\end{proof}
+
+This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders. %imp
+
+In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation. 
+%We provide an information-theoretic justification for this claim in Section \ref{sec:infth}. 
+The intuition behind this  is that a trader's prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
+
+\subsection{Budget of Informative Traders}\label{sec:inf}
+Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
+
+We now characterize the expected change in budget for an informative trader. The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
+\begin{theorem}\label{thm:growth}
+Let $\theta$ be the current market state in the exponential family prediction market.
+Suppose that an  informative trader with belief distribution  parametrized by $\theta'$  moves the market state to the budget-limited state $\htheta=\lambda \theta' + (1-\lambda) \theta$.
+Then, the expectation (over the trader's belief) of the trader's profit  is greater than zero whenever his budget is positive and his belief differs from the previous market position $\theta$.
+% Suppose that each informative trader gets a random sample of data, 
+%resulting in a sequence of trader beliefs $\theta_i$, and hence a sequence of budget-limited market positions
+%$\hat{\theta}_i$, before a final outcome $x$.
+Then, the expectation, over trader $i$'s belief distribution $\theta_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\hat{\theta}_{i-1}$.
+\end {theorem}
+\begin{proof}
+Let the cost function $C$ be equal to the log partition function $T$ of the belief distribution.
+The payoff is given by the sufficient statistics $\phi(x)$.
+Then, the trader's expected net payoff is given by
+\begin{eqnarray*}
+&& \E_{x\sim P_{\htheta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
+&=& T(\theta)-\theta\E_{x\sim P_{\htheta}}[\phi(x)]-(T(\theta')-\theta' \E_{x\sim P_{\htheta}}[\phi(x)])\\
+&=& T(\theta)-\theta\nabla T(\htheta)-(T(\theta')-\theta' \nabla T(\htheta))\\
+&=& T(\theta)-T(\htheta)-\nabla T(\htheta)(\theta-\htheta)-(T(\theta')-T(\htheta)- \nabla T(\htheta)(\theta'-\htheta))\\
+&=& D_{T}(\theta,\htheta)-D_{T}(\theta',\htheta)\\
+&\geq&\lambda D_{T}(\theta,\htheta)\geq0
+\end{eqnarray*}
+The second to last inequality holds since $D_{T}(\theta',\htheta)$ is convex in $\theta'$ and we have:
+\begin{eqnarray*}
+D_{T}(\theta',\htheta)&=&D_{T}\left(\lambda\htheta+(1-\lambda)\theta,\htheta\right)\\
+&\leq&\lambda D_{T}( \htheta ,\htheta)+ (1-\lambda) D_{T}( \theta ,\htheta)\\
+&=&(1-\lambda) D_{T}( \theta ,\htheta)
+%\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
+\end{eqnarray*}¥
+Thus, a trader who moves the market state can expect his profit to be positive and at least $\lambda D_{T}(\theta,\htheta)$.
+ \end{proof}
+
+For continuous distributions with a density, the probability that a trader with private information will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs.
+This result suggests that, eventually, every informative trader will have the ability to influence the market state in accordance with his beliefs, without being budget limited.
+
+Notice that Theorem~\ref{thm:growth} only required that the market state to which the trader moves, be representable as a convex combination of the current market state and his belief.
+This means that the result holds for exponential utility traders aiming to maximize their utility.
+In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{T}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
+
+We note one important aspect of Theorem~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
+This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure.
+If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
+%\pagebreak
+
+%\subsection{Exponential Utility}
+%Now suppose the trader has exponential utility, given by $$U(w) = -\frac{1}{a} \exp(-aw)$$ where $a$ is the coefficient of risk aversion.
+Suppose the trader wishes to maximize his expected utility.
+Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by
+%\begin{eqnarray*}
+%&&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}U\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]\\
+%&=&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}-\frac{1}{a}\exp\left[ -a\delta\phi(x) +a C(\delta + \theta) -a C(\theta)\right]
+%\end{eqnarray*}¥
+%
+%\begin{eqnarray*}
+%\E_{\beta\sim b}\E_{x\sim \beta}\exp\left[-a\delta\phi(x)\right]&=&\int_{\mathcal{B}}\int_{\mathcal{X}}\exp[-a\delta\phi(x)]\, \exp\{\beta\cdot\phi(x)-\psi(\beta)\}\, dx\\
+%&&\qquad\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\, d\beta\\
+%&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
+%&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\int_{\mathcal{X}}\exp[(\beta-a\delta)\cdot\phi(x)-\psi(\beta-a\delta)]\, dx\\
+%&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
+%&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\\
+%\end{eqnarray*}¥
+%\pagebreak
+
 \section{Conclusions}

### 2014-02-10 WriteUp/maxentpredmkts.tex
adding more notes and some fixes

@@
 
+\jake{I want to add more to this list. It's ok but doesn't totally reflect what's in the last few sections.}
 \subsection{Related Work}
@@
 
-If wealth is distributed according to a probability measure $P$, then the \emph{certainty equivalent} of a random amount of wealth is defined as
-$$
-\CE(w) = U_a^{-1} \Exp_P\left[ U_a(w) \right].
+\jake{This set of ideas is nice and needs to be sold more strongly in the intro} If wealth is distributed according to a probability measure $P$, then the \emph{certainty equivalent} of a random amount of wealth is defined as
+$$
+\CE(w) = U_a^{-1} (\Exp_P\left[ U_a(w) \right]).
 $$
@@
 %
-For the second part of the result, we have
+For the second part of the result, we have\jake{Is the $\hat \cdot$ missing from the computations below?}
 %
@@
 
-\subsection{Reinterpreting repeated trades}
+\subsection{Repeated Trading and the Effective Belief}
+\jake{I need to clean up a few things in this section.}
 In previous sections we have analyzed trader behavior as if it is his first entry into the market.
@@
 
-As before, we consider the exponential family prediction market where traders have exponential belief. Let us also suppose that traders in this market have exponential utility \[
-U(w) = -\frac{1}{a} \exp(-aw).
-\]
-%
-Here $a$ is the coefficient of risk aversion (higher means more risk averse, and the utility function is more concave).
+% As before, we consider the exponential family prediction market where traders have exponential belief. Let us also suppose that traders in this market have exponential utility \[
+% U(w) = -\frac{1}{a} \exp(-aw).
+% \]
+% %
+% Here $a$ is the coefficient of risk aversion (higher means more risk averse, and the utility function is more concave).
 
@@
 \subsection{Equilibrium Market State for Exponential Utility Agents}
+\jake{need to mention the connection to potential games in the intro, it's a nice little extra bit.}
 We have shown that every exponential-utility maximizing trader picks the share vector $\delta$ so that the eventual market state can be represented as a convex combination of the initial market state and the natural parameter of his (exponential family) belief distribution.
@@
 
-Recall the following result from game theory. %need a reference/name for this result?
+We now review a well-known result from game theory regarding the class of \emph{potential games} \cite{monderer1996potential}.
+Note that we say a function $f(\vec{x})$ is at a \emph{local optimum} if changing any coordinate of $\vec{x}$ does not increase the value of $f$.
 \begin{thm}\label{thm:potl}
-Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}\defeq \delta_{1},\ldots,\delta_{i},\ldots,\delta_{n}$.
-If there exists a potential function $g(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=g(\vec{\delta})-g(\vec{\delta}_{-i},\delta_{i}')$$
-then when $g(\vec{\delta})$ is maximized, $\vec{\delta}$ is an equilibrium.
+Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}\defeq (\delta_{1},\ldots,\delta_{i},\ldots,\delta_{n})$.
+If there exists a potential function $\Phi(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=\Phi(\vec{\delta})-\Phi(\vec{\delta}_{-i},\delta_{i}')$$
+then $\vec{\delta}$ is a Nash equilibrium if and only if $\Phi(\vec{\delta})$ is at a local optimum.
 \end{thm}¥
@@
 
-Define a potential function as  $$g(\vec{\delta})\defeq  T(\theta+\sum_{i}\delta_{i})+\sum_{i}\frac{1}{a_{i}}T(\htheta_{i}-a_{i}\delta_{i})$$
+Define a potential function as  $$\Phi(\vec{\delta})\defeq  T(\theta+\sum_{i}\delta_{i})+\sum_{i}\frac{1}{a_{i}}T(\htheta_{i}-a_{i}\delta_{i})$$
 
@@
 Now the utility of trader $i$ is $U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i})$.
-Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $g(\vec{\delta})$ for each $\delta_{i}$.
-\begin{eqnarray*}
-\nabla_{\delta_{i}}g(\vec{\delta})&=&\nabla T(\theta+\sum_{j=1}^{n}\delta_{i})-\nabla T(\htheta_{i}-a_{i}\delta_{i})\\
-&=&0
+Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $\Phi(\vec{\delta})$ for each $\delta_{i}$.
+\begin{eqnarray*}
+\nabla_{\delta_{i}}\Phi(\vec{\delta})&=&\nabla T(\theta+\sum_{j=1}^{n}\delta_{i})-\nabla T(\htheta_{i}-a_{i}\delta_{i}) = 0
 \end{eqnarray*}¥

### 2014-02-10 WriteUp/maxentpredmkts.tex
added small note for sindhu
removed unnecessary character

@@
 \end{eqnarray*}
-This follows from Proposition \ref{prop:exp-util}.
+This follows from Proposition \ref{prop:exp-util} \jake{Sindhu this ref isn't in the doc}.
 The maximizer is $\delta_{2}^* = (\Theta - \theta') / (1+a)$, which moves the share vector to $\theta' + \delta_{2}^* = \frac{1}{1+a} \Theta + \frac{a}{1+a} \theta'$ which is a convex combination of the effective belief and the current market state.
@@
 then $\vec{\delta}$ is a Nash equilibrium if and only if $\Phi(\vec{\delta})$ is at a local optimum.
-\end{thm}¥
+\end{thm}
 In the exponential family market, the cost function $C$ is identical to the log partition function $T$.
 Let $\vec{\delta}$ be the vector of vectors of shares purchased by every trader in the market when the market has reached equilibrium, $\theta$ the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his utility parameter.
+\jake{I need to fix a small technical detail with the fact that we are using the log utilities to use the potential argument.}
 

### 2014-02-10 WriteUp/maxentpredmkts.tex
more edits to intro, labelled lots of sections so I could refer to them

@@
 In the present work we show that a number of classical statistical tools can be leveraged to design a prediction market framework in the mold of an \emph{exponential family distributions} that possesses a number of attractive properties.
-Common concepts in statistics \emph{entropy maximization}, \emph{log loss}, and \emph{bayesian inference}, relate to natural aspects of our class of mechanisms.
-In particular, the central objects in our market framework can be interpreted as elements of exponential families:
+Common concepts in statistics, e.g.
+\emph{entropy maximization}, \emph{log loss}, and \emph{bayesian inference}, relate to natural aspects of our class of mechanisms.
+In particular, the central objects in our market framework can be interpreted via conceptual objects used to define exponential families:
 \begin{itemize}
@@
 
-In addition to showing this syntactic relationship between exponential families and prediction markets, we also explore the semantic implications as well.
+We begin with a discussion of scoring rules in Section~\ref{sec:scoring} for exponential family distributions, and in particular we show that the logarithmic scoring rule is the only score that is \emph{proper} for the class of exponential families.
+We turn our attention to market design in Section~\ref{sec:maxentmarkets} and give a full description of our proposed mechanisms.
+In addition to showing the syntactic relationship between exponential families and prediction markets, we also explore the semantic implications as well.
 In particular, we show that our formulation allows us to analyze the evolution of the market under various models of trader behavior:
 \begin{itemize}
-	\item We consider \emph{budget-limited traders} who are constrained in how much they influence the market.
+	\item Trader behavior varies depending on how they assimilate information; for example, should we consider our agents as Bayesians or frequentists.
+In Section~\ref{sec:bayesians} We consider traders that use a conjugate prior to update their beliefs, and we study how their trades would affect the market state.
+\jake{Do we show the following?} We show that although these two kinds of traders pick the eventual market state as a convex combination of their private belief and current market state, they do so in dual spaces.
+	\item In Section~\ref{sec:exputil} we consider when our agents are risk-averse in an interestins special case, that is under the assumption they utilize \emph{exponential utility} to optimize their bets.
+We show how this dovetails nicely with the notion of a \emph{certainty equivalent} which is often used in risk-aversion settings.
+In this case we can characterize precisely how a single trader interacts with the market as well as the equilibrium reached given multiple traders; this result is achieved via using a \emph{potential game} argument.
+The eventual market state turns out to be a weighted combination of trader beliefs and initial market state; the weights depend on the risk aversion parameter of the individual traders.
+	\item In Section~\ref{sec:budgets} we consider \emph{budget-limited traders} who are constrained in how much they influence the market.
 We analyze the market  under these circumstances; we are able to show that traders with good information can expect to profit and their influence over the market state increases over time whereas malicious traders have limited impact on the market.
-	\item We consider when our agents are risk-averse in an interestins special case, that is under the assumption they utilize \emph{exponential utility} to optimize their bets.
-In this case we can characterize precisely how a single trader interacts with the market as well as the equilibrium reached given multiple traders.
-The eventual market state turns out to be a weighted combination of trader beliefs and initial market state; the weights depend on the risk aversion parameter of the individual traders.
-	\item In addition, we observe that trader behavior varies depending on whether they assimilate information as Bayesians or as frequentists.
-Interestingly, we show that although these two kinds of traders pick the eventual market state as a convex combination of their private belief and current market state, they do so in dual spaces.
 \end{itemize}
@@
 %\pagebreak
-\section{Maximum Entropy Market Making}
+\section{Maximum Entropy Prediction Markets via Exponential Families}
+\label{sec:maxentmarkets}
 
@@
 
-\section{Bayesian Traders with Linear Utility}
+\section{Bayesian Traders with Linear Utility} \label{sec:bayesians}
 
@@
 
-\section{Risk Aversion, Exponential Utility, and Market Equilibrium}
+\section{Risk Aversion, Exponential Utility, and Market Equilibrium} \label{sec:exputil}
 
@@
 
-Define a potential function as  $$\Phi(\vec{\delta})\defeq  T(\theta+\sum_{i}\delta_{i})+\sum_{i}\frac{1}{a_{i}}T(\htheta_{i}-a_{i}\delta_{i})$$
+Define a potential function as  $$\Phi(\vec{\delta})\defeq  T\left(\theta+\sum_{i}\delta_{i}\right)+\sum_{i}\frac{1}{a_{i}}T(\htheta_{i}-a_{i}\delta_{i})$$
 
@@
 \begin{eqnarray*}
-\nabla_{\delta_{i}}\Phi(\vec{\delta})&=&\nabla T(\theta+\sum_{j=1}^{n}\delta_{i})-\nabla T(\htheta_{i}-a_{i}\delta_{i}) = 0
+\nabla_{\delta_{i}}\Phi(\vec{\delta})&=&\nabla T\left(\theta+\sum_{j=1}^{n}\delta_{i}\right)-\nabla T(\htheta_{i}-a_{i}\delta_{i}) = 0
 \end{eqnarray*}¥
@@
 Rewriting, we have for each trader $i$, 
-$$\frac{\htheta_{i}}{a_{i}}-\delta_{i}=\frac{1}{a_{i}}(\theta+\sum_{j=1}^{n}\delta_{j})$$
+$$\frac{\htheta_{i}}{a_{i}}-\delta_{i}=\frac{1}{a_{i}}\left(\theta+\sum_{j=1}^{n}\delta_{j}\right)$$
 Thus,
@@
 
-\section{The Exponential Family Market Mechanism with Budgets}
+\section{The Exponential Family Market Mechanism with Budgets} \label{sec:budgets}
 In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.

### 2014-02-10 WriteUp/maxentpredmkts.tex
liquidity notes; minor edits and comments to main doc

@@
 RAHUL SAMI
-\affil{University of Michigan, Ann Arbor}}
+\affil{Google India}}
 %%\author{GANG ZHOU
@@
 \begin{abstract}
+\sindhu{this needs to change once we finalize the sections}
 In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets.
@@
 Further, since the trades are done sequentially, the trader is allowed to observe all past trades in the market and update his private information based on this information.
-Traders can see the past history of trades, so the price at which a current trader is willing to buy and sell these securities can be interpreted as an aggregate ``consensus probability forecast'' of a particular candidate winning the election.
-
+Traders can see the past history of trades, so the price at which a current trader is willing to buy and sell these securities can be interpreted as an aggregate ``consensus probability forecast'' of a particular event occurring.
 % For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins.
@@
 
-Much the work on prediction market frameworks has focused primarily on structural properties of the mechanism: incentive compatibility, the market maker loss, the available liquidity, the fluctuations of the prices as a function of the trading volume, to name a few.
+Much of the work on prediction market frameworks has focused primarily on structural properties of the mechanism: incentive compatibility, the market maker loss, the available liquidity, the fluctuations of the prices as a function of the trading volume, to name a few.
 Absent from much of the literature is a corresponding \emph{semantics} of the market behavior or the observed prices.
@@
 	\item Trader behavior varies depending on how they assimilate information; for example, should we consider our agents as Bayesians or frequentists.
-In Section~\ref{sec:bayesians} We consider traders that use a conjugate prior to update their beliefs, and we study how their trades would affect the market state.
+In Section~\ref{sec:bayesians} we consider traders that use a conjugate prior to update their beliefs, and we study how their trades would affect the market state.
 \jake{Do we show the following?} We show that although these two kinds of traders pick the eventual market state as a convex combination of their private belief and current market state, they do so in dual spaces.
-	\item In Section~\ref{sec:exputil} we consider when our agents are risk-averse in an interestins special case, that is under the assumption they utilize \emph{exponential utility} to optimize their bets.
+	\item In Section~\ref{sec:exputil} we consider when our agents are risk-averse in an interesting special case, that is under the assumption they utilize \emph{exponential utility} to optimize their bets.
 We show how this dovetails nicely with the notion of a \emph{certainty equivalent} which is often used in risk-aversion settings.
@@
 Note that given a proper scoring rule $S$ any affine transformation $\tilde{S}(\mu,x) = aS(\mu,x) + b(x)$ of the rule, with $a > 0$ and $b$ an arbitrary real-valued function of the outcomes, again yields a proper scoring rule termed \emph{equivalent}~\citep{}.
-Throughout we will implicitly apply such affine transformations to obtain the clearest version of the scoring rule.
+\sindhu{this one?
+\cite{dawid2006geometry}}Throughout we will implicitly apply such affine transformations to obtain the clearest version of the scoring rule.
 We will also focus on scoring rules where inequality~(\ref{eq-proper}) is strict to avoid trivial cases such as constant scoring rules.
@@
 \subsection{Proper Scoring from Maximum Entropy}
-
+\sindhu{\cite{GrunwaldDawid}}
 Our starting point for designing proper scoring rules is the classic logarithmic scoring rule for eliciting probabilities in the case of finite outcomes.
@@
 \subsubsection*{Exponential Families}
-
+\sindhu{\cite{WainJordan08}}
 We let $p(x;\mu)$ be the maximum entropy distribution with expected statistic $\mu$.
@@
 
-Consider Bayesian agents maintaning and exponential family conjugate prior over the data model's natural parameters (equivalently, the expected security payoffs).
+Consider Bayesian agents maintaining and exponential family conjugate prior over the data model's natural parameters (equivalently, the expected security payoffs).
 Each agent has access to a private sample of the data of size $m$ with mean statistic $\hmu$.
@@
 \end{eqnarray*}
-This follows from Proposition \ref{prop:exp-util} \jake{Sindhu this ref isn't in the doc}.
+This follows from Proposition \ref{prop:exp-util} \jake{Sindhu this ref isn't in the doc}\sindhu{it looks like the exp-util section has been rewritten since I edited this: I'll need to check this reference}.
 The maximizer is $\delta_{2}^* = (\Theta - \theta') / (1+a)$, which moves the share vector to $\theta' + \delta_{2}^* = \frac{1}{1+a} \Theta + \frac{a}{1+a} \theta'$ which is a convex combination of the effective belief and the current market state.
@@
 \section{The Exponential Family Market Mechanism with Budgets} \label{sec:budgets}
+\sindhu{\cite{Azoury01}}
 In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
@@
 From the convexity of $C$, we have
- $$C(\theta')\leq(1-\lambda)C(\theta)+\lambda C(\htheta)$$
-
-Now 
-\begin{eqnarray*}
-C(\theta') - C(\theta) &\leq& (1-\lambda)C(\theta)\\
-&&+\lambda C(\htheta) - C(\theta)\\
-&=& \lambda\ (C(\htheta) - C(\theta))
-\end{eqnarray*}
-
- Thus, $C(\theta') - C(\theta) \leq \alpha$. 
- \end{proof}
+\begin{eqnarray*}
+C(\theta')&\leq&(1-\lambda)C(\theta)+\lambda C(\htheta)\\
+\mbox{Thus, } C(\theta') - C(\theta) &\leq& \lambda\ (C(\htheta) - C(\theta))\\
+\mbox{And, } C(\theta') - C(\theta) &\leq& \alpha 
+ \end{eqnarray*}\end{proof}
 

### 2014-02-10 WriteUp/maxentpredmkts.tex
some more minor edits
added liquidity adjustment to exp util section

@@
 %
-\begin{theorem}
+\begin{theorem} \label{thm:exp-util}
 Suppose an agent has exponential utility with coefficient $a$ and exponential family beliefs with natural parameter $\htheta$.
@@
 
-%\pagebreak
+\subsubsection*{Liquidity Adjustment}
+
+In practice the centralized market maker allows itself some control over the \emph{liquidity} in the market, which captures how responsive prices are to trades.
+To adjust liquidity we consider the parametrized cost $C_{\lambda}(\theta)=\frac{1}{\lambda} C\left(\lambda\theta\right)$.
+Here $\lambda$ is construed as the \emph{inverse liquidity}, or price responsiveness.
+A larger setting of $\lambda$ means fewer shares need to be bought to reach the same prices.\footnote{Observe that the liquidity adjustment to the cost function takes the same form as the risk-aversion adjustment to the exponential utility in~(\ref{exp-util}).
+In convex analysis, this transformation is known as the perspective function~\citep{}.}
+
+%(Discussion, you can skip this paragraph.) This is because for higher $\lambda$ we are making the cost function ``more convex''.
+Specifically, for $\lambda_1 > \lambda_2$ it's the case that $C_{\lambda_1}$ is an increasing convex transformation of $C(\lambda_2)$.
+This holds because the ``coefficient of risk aversion'' $A(\theta) = C''(\theta)/C'(\theta)$ increases with $\lambda$ at every point.
+I've written it assuming single-dimensional $\theta$ but there are more complicated looking generalizations for vector $\theta$.
+In fact, this coefficient might be the ``right'' notion of liquidity at a point, not the second derivative or largest eigenvalue of the Hessian.
+We should discuss.
+
+\newcommand{\ttheta}{\tilde{\theta}}
+In the context of the generalized LMSR we write $T$ rather than $C$ where $T$ is the log partition function, with liquidity-adjusted version $T_{\lambda}$.
+Let $\hmu$ be the agent's mean belief with corresponding natural parameter $\htheta = \grad T^{-1}(\hmu)$.
+Recall that a risk-neutral agent moves the share vector so that the prices match its mean parameter.
+Therefore, define the \emph{target shares} as $\ttheta = \grad T^{-1}_{\lambda}(\hmu)$.
+The target shares $\htheta$ and natural parameter $\ttheta$ are related by
+%
+$$
+\grad T(\htheta) = \grad T_{\lambda}(\ttheta) = \hmu.
+$$
+%
+It is straightforward to check that $\grad T_{\lambda}(\ttheta) = \grad T(\lambda \ttheta)$ so we have
+%
+\begin{equation} \label{target}
+\ttheta = \htheta / \lambda.
+\end{equation}
+%
+Higher price responsiveness means fewer shares must be bought to make the market prices match the agent's expectation, so the natural parameter is scaled down accordingly.
+With a liquidity adjustment, the analysis of Theorem~\ref{thm:exp-util} can be extended and yields the following result.
+%
+\begin{corollary}
+Under the conditions of Theorem~\ref{thm:exp-util} and an inverse liquidity of $\lambda$, the agent's optimal trade $\delta$ moves the state vector to
+%
+\begin{equation} \label{liqu-update}
+\theta + \delta = \frac{\lambda}{\lambda+a} \ttheta + \frac{a}{\lambda+a} \theta
+= \frac{\lambda}{\lambda+a} \grad T^{-1}_{\lambda}(\hmu) + \frac{a}{\lambda+a} \grad T^{-1}_{\lambda}(\mu).
+\end{equation}
+%
+\end{corollary}
+%
+According to~(\ref{liqu-update}), as $\lambda$ grows large the agent moves the market state closer to the \emph{target shares}, rather than its true natural parameter.
+Note that the target shares themselves depend on $\lambda$ by~(\ref{target}), but the update can be directly written in terms of the agent's beliefs as in the right-hand side of~(\ref{liqu-update}).
 
@@
 Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
+This idea of a dual analysis has been used in a recommender system context to compute reputations in \cite{ressami07}.
 

### 2014-02-10 WriteUp/maxentpredmkts.tex
new command in preamble

@@
 \newcommand{\htheta}{\hat{\theta}}
+\newcommand{\ttheta}{\tilde{\theta}}
 %---------------------------------------------------------------------------
@@
 
-\newcommand{\ttheta}{\tilde{\theta}}
 In the context of the generalized LMSR we write $T$ rather than $C$ where $T$ is the log partition function, with liquidity-adjusted version $T_{\lambda}$.

### 2014-02-10 WriteUp/maxentpredmkts.tex
some more comments on budgets

@@
 \sindhu{\cite{Azoury01}}
-In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
+%In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
 
@@
 In this section, we assume that the traders have exponential family beliefs.
-The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
+The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$; the vector of sufficient statistics is assumed to be non-negative.
+\sindhu{is this ok?}.
 \subsection{Budget-limited Aggregation}
@@
 In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget.
+\sindhu{this assumption is also very strong but is needed}
 Let $\alpha$ be the budget of a trader in the market.
@@
 First, we  show that this trade is feasible given the trader's budget:
-\begin{theorem}
+\begin{lemma}
 Let the current market state be given by $\theta$.
@@
 The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$.
-\end{theorem}
+\end{lemma}
 \begin{proof}

### 2014-02-10 WriteUp/maxentpredmkts.tex

--- 2014-02-10 WriteUp/maxentpredmkts.tex
@@
 A scoring rule $S : \mM \times \mX \rightarrow \bR \cup \{-\infty\}$ pays the agent $S(\hmu,x)$ according to how well its report $\hmu \in \mM$ agrees with the eventual outcome $x \in \mX$.
-The following definition is due to~\citet{}.
+The following definition is due to~\citet{lambert2008eliciting}.
 %
@@
 \noindent
-Note that given a proper scoring rule $S$ any affine transformation $\tilde{S}(\mu,x) = aS(\mu,x) + b(x)$ of the rule, with $a > 0$ and $b$ an arbitrary real-valued function of the outcomes, again yields a proper scoring rule termed \emph{equivalent}~\citep{}.
-\sindhu{this one?
-\cite{dawid2006geometry}}Throughout we will implicitly apply such affine transformations to obtain the clearest version of the scoring rule.
+Note that given a proper scoring rule $S$ any affine transformation $\tilde{S}(\mu,x) = aS(\mu,x) + b(x)$ of the rule, with $a > 0$ and $b$ an arbitrary real-valued function of the outcomes, again yields a proper scoring rule termed \emph{equivalent}~\citep{Dawid98,GneitingRa07}.
+Throughout we will implicitly apply such affine transformations to obtain the clearest version of the scoring rule.
 We will also focus on scoring rules where inequality~(\ref{eq-proper}) is strict to avoid trivial cases such as constant scoring rules.
@@
 Even though the log score is only applied to densities from $\mD$, according to Definition~\ref{def:proper-score} it must work over all densities in $\mP$.
-It turns out this is possible if $\mD$ is chosen appropriately, drawing on a well-known duality between maximum likelihood and maximum entropy~\citep{}.
+It turns out this is possible if $\mD$ is chosen appropriately, drawing on a well-known duality between maximum likelihood and maximum entropy~\citep{GrunwaldDawid}.
 
@@
 
-The following proposition collects the relevant results on regular exponential families; proofs may be found in~\citep{}.
+The following proposition collects the relevant results on regular exponential families; proofs may be found in~\citep{BanerjeeDhGh05b,Barndorff78,WainJordan08}.
 A convex function $T$ is of \emph{Legendre type} if it is proper, closed, strictly convex and differentiable on the interior of its domain, and $\lim_{\theta \rightarrow \bar{\theta}} \norm{\grad T(\theta)} = +\infty$ when $\bar{\theta}$ lies on the boundary of the domain.
@@
 
-As further intuition for the result, note that~(\ref{breg-div}) is the definition of the `Bregman divergence' with respect to strictly convex function $T$~\citep{}. Therefore we have 
+As further intuition for the result, note that~(\ref{breg-div}) is the definition of the `Bregman divergence' with respect to strictly convex function $T$. Therefore we have 
 %
@@
 %
-where the last equality is a well-known identity relating the Bregman divergences of $T$ and $T^* = G$~\citep{}.
+where the last equality is a well-known identity relating the Bregman divergences of $T$ and $T^* = G$~\citep{BanerjeeDhGh05b}.
 The equation states that the agent's regret from misreporting its mean parameter does not depend on the full density $p$, only the mean $\mu$.
@@
 Here the statistics being elicited are the mean vector $\mu$ and the covariance matrix $\Sigma$.
-These scoring rules have been studied by~\citet{} as rules that only depend on the mean and variance of the reported density.
+These scoring rules have been studied by~\citet{dawid99coherent} as rules that only depend on the mean and variance of the reported density.
 They note that these rules are weakly proper (because they do not distinguish between densities with the same first and second moments), but do not make the point that knowledge of the full density is not necessary on the part of the agent.
@@
 In the above, Example~\ref{ex:gaussian} illustrates an important point about parametrizations of the elicited expectations.
-The variance $\sigma^2$ cannot be written as $\Exp[\phi(x)]$ for any $\phi$, because the mean $\mu$ enters the definition of $\sigma^2$ but is not available when $\phi$ is defined (indeed it is elicited in tandem with the variance).\footnote{This is an intuitive but far from formal explanation for the fact that the dimension of the message space, or \emph{elicitation complexity}, for eliciting the variance is at least 2~\citep{}.} Instead one must use the first two \emph{uncentered} moments $\Exp[x]$ and $\Exp[x^2]$.
+The variance $\sigma^2$ cannot be written as $\Exp[\phi(x)]$ for any $\phi$, because the mean $\mu$ enters the definition of $\sigma^2$ but is not available when $\phi$ is defined (indeed it is elicited in tandem with the variance).\footnote{This is an intuitive but far from formal explanation for the fact that the dimension of the message space, or \emph{elicitation complexity}, for eliciting the variance is at least 2~\citep{lambert2008eliciting}.} Instead one must use the first two \emph{uncentered} moments $\Exp[x]$ and $\Exp[x^2]$.
 These are in bijection with $\mu$ and $\sigma^2$, so the resulting scoring rule can be re-written in terms of the latter.
@@
 In a multi-agent setting, a prediction market can be used to \emph{aggregate} the beliefs of the agents.
-In his seminal paper~\citet{} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both in order to perform well in thin or thick markets.
+In his seminal paper~\citet{H03} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both in order to perform well in thin or thick markets.
 In this section, we adapt the generalized log scoring rule to a market scoring rule which leads to markets with simple closed-form cost functions for many statistics of interest.%The exponential families framework also allows us to move beyond risk-neutral agents to understand the aggregation properties of prediction markets under alternative behaviors, such as Bayesian updating and risk aversion.
@@
 
-The standard way to implement a prediction market in the literature, due to~\citet{}, is via a centralized market maker.
+The standard way to implement a prediction market in the literature, due to~\citet{CP07}, is via a centralized market maker.
 The market maker maintains a convex, differentiable cost function $C : \bR^d \rightarrow (-\infty,+\infty]$, where $C(\theta)$ records the revenue collected when the vector of outstanding shares is $\theta$.
@@
 recalling~(\ref{breg-div}).
-Now~\citet{} have observed (among others) that the Kullback-Leibler divergence between two exponential family distributions is the Bregman divergence, with repect to the log-partition function, between their natural parameters.
-The agent's expected profit is therefore the KL divergence between the market's implied expectation and the exponential family corresponding to the agent's expectation, a well-known property from the classical LMSR~\citep{}.
+Now~\citet{BanerjeeDhGh05} have observed (among others) that the Kullback-Leibler divergence between two exponential family distributions is the Bregman divergence, with repect to the log-partition function, between their natural parameters.
+The agent's expected profit is therefore the KL divergence between the market's implied expectation and the exponential family corresponding to the agent's expectation, a well-known property from the classical LMSR~\citep{H07}.
 
@@
 As another example let the outcome space be the $d$-dimensional unit sphere.
-This setting was considered by~\citet{} who provide a cost function implicitly defined through a variational characterization.
+This setting was considered by~\citet{Abernethy11} who provide a cost function implicitly defined through a variational characterization.
 The maximum entropy approach leads to another alternative.
@@
 %
-where $I_r$ refers to the modified Bessel function of first kind and order $r$; see~\citep{} for an explanation of these quantities.
+where $I_r$ refers to the modified Bessel function of first kind and order $r$; see~\citep{BanerjeeDhGh05} for an explanation of these quantities.
 The effective domain of $\theta$ is the positive orthant in $\bR^d$.
@@
 
-In the standard model of cost-function based prediction markets, a sequence of myopic, risk-neutral agents arrive and trade in the market~\citep{}.
+In the standard model of cost-function based prediction markets, a sequence of myopic, risk-neutral agents arrive and trade in the market~\citep{CP07,Chen10}.
 As we saw in Section~\ref{sec:pred-market}, such a trader moves the prices to its own expectation $\mu$.
@@
 The sufficient statistic maps $\theta$ to $(\theta, T(\theta))$, and the log partition function $\psi$ is defined as the normalizer as usual.
-For a complete treatment of exponential families conjugate priors, see for instance~\citep{}.
-Now~\citet{} and~\citet{} have shown that
+For a complete treatment of exponential families conjugate priors, see for instance~\citep{Barndorff78}.
+Now~\citet{diaconis79} and~\citet{jewell1974credible} have shown that
 %
@@
 Suppose now that the agent observes an empirical sample with mean $\hmu$ and size $m$.
-By a standard derivation~\citep{}, the posterior conjugate prior parameters become $n\nu \leftarrow n\nu + m\hmu$ and $n \leftarrow n+m$, and the posterior expectation~(\ref{eq:prior}) evaluates to
+By a standard derivation~\citep{diaconis79}, the posterior conjugate prior parameters become $n\nu \leftarrow n\nu + m\hmu$ and $n \leftarrow n+m$, and the posterior expectation~(\ref{eq:prior}) evaluates to
 %
@@
 A larger setting of $\lambda$ means fewer shares need to be bought to reach the same prices.\footnote{Observe that the liquidity adjustment to the cost function takes the same form as the risk-aversion adjustment to the exponential utility in~(\ref{exp-util}).
-In convex analysis, this transformation is known as the perspective function~\citep{}.}
+In convex analysis, this transformation is known as the perspective function~\citep{HiriartLe00}.}
 

### 2014-02-10 WriteUp/maxentpredmkts.tex
merged files

@@
 \jake{I'll add more here. Sebastien: do you have anything to put in here regarding the exponential utility stuff?}
+Prediction markets have been known to perform well in practice (see, for instance, \cite{AGT07,Wol04,pennock2001real}).
+However, a sound theory for interpreting trader behavior and market prices is an ongoing area of study \cite{wolfers2006interpreting}.
+
+Scoring rules are a measure of merit of a predictor \cite{GneitingRa07}.
+\cite{hanson03} adapted scoring rules to markets with sequential participation of multiple traders.
+A cost function based formulation of markets was first introduced by \cite{CP07}.
+
 Designing prediction markets to handle a large outcome space is an active area of research.
@@
 For the logarithmic market scoring rule they show that unbounded market maker loss can result in this setting.
+\cite{Abernethy13} and \cite{OthmanS11} specify frameworks under which they design cost function based markets that satisfy the desirable property of bounded market maker loss even in infinite outcome spaces.
+
+In this paper, we use exponential family distributions to construct our cost function.
+For a comprehensive introduction to the class of exponential family distributions see \cite{WainJordan08}.
+\cite{Azoury01} consider learning exponential family distributions in a traditional online learning model and provide worst case loss bounds relative to using an offline algorithm.
+\cite{storkey11} constructs a prediction market where each participating agent is a learning algorithm with varying utility functions.
+
+%Designing prediction markets to handle a large outcome space is an active area of research.
+In \cite{Chen08}, the authors use a restricted betting language to design efficient markets for a combinatorial outcome space.
+This technique is generalized by \cite{Pennock11}.
+\cite{Gao09} consider extending various automated market makers to an infinite outcome space.
+For the logarithmic market scoring rule they show that unbounded market maker loss can result in this setting.
 \cite{Abernethy11} and \cite{OthmanS11} specify frameworks under which they design cost function based markets that satisfy the desirable property of bounded market maker loss even in infinite outcome spaces.
-
-The connection between machine learning and prediction markets has been studied previously.
+%
+%The connection between machine learning and prediction markets has been studied previously.
 \cite{Chen10} and \cite{Abernethy11} have previously explored the connection to learning algorithms to inform the design and understanding of prediction markets.
 In particular, \cite{Chen10}  consider the correspondence between prediction markets with market scoring rules and the Follow the Regularized Leader algorithm proposed by \cite{Kalai05} and thus provide insight into the aggregation mechanism of a prediction market.
-
-Independently of this work, Beygelzimer et al.~\cite{BLP12} have shown that, for a particular form of binary prediction markets and traders with log-utility, the long-run dynamics of trading activity and budgets over many prediction markets lead the markets to satisfy a bounded regret property with respect to the best single trader.
+%
+%Independently of this work, Beygelzimer et al.~\cite{BLP12} have shown that, for a particular form of binary prediction markets and traders with log-utility, the long-run dynamics of trading activity and budgets over many prediction markets lead the markets to satisfy a bounded regret property with respect to the best single trader.
 Our results here, and the future work we have suggested, form a program to prove bounded regret properties of budget-limited prediction markets under much more general conditions.
@@
 \subsection{Proper Scoring from Maximum Entropy}
-\sindhu{\cite{GrunwaldDawid}}
+
 Our starting point for designing proper scoring rules is the classic logarithmic scoring rule for eliciting probabilities in the case of finite outcomes.
@@
 \subsubsection*{Exponential Families}
-\sindhu{\cite{WainJordan08}}
+
 We let $p(x;\mu)$ be the maximum entropy distribution with expected statistic $\mu$.
@@
 As another example let the outcome space be the $d$-dimensional unit sphere.
-This setting was considered by~\citet{Abernethy11} who provide a cost function implicitly defined through a variational characterization.
+This setting was considered by~\citet{Abernethy13} who provide a cost function implicitly defined through a variational characterization.
 The maximum entropy approach leads to another alternative.

### 2014-02-10 WriteUp/maxentpredmkts.tex
cleaned up scoring rule and info market sections

--- 2014-02-10 WriteUp/maxentpredmkts.tex
@@
 \noindent
-Note that given a proper scoring rule $S$ any affine transformation $\tilde{S}(\mu,x) = aS(\mu,x) + b(x)$ of the rule, with $a > 0$ and $b$ an arbitrary real-valued function of the outcomes, again yields a proper scoring rule termed \emph{equivalent}~\citep{Dawid98,GneitingRa07}.
+Given a proper scoring rule $S$ any affine transformation $\tilde{S}(\mu,x) = aS(\mu,x) + b(x)$ of the rule, with $a > 0$ and $b$ an arbitrary real-valued function of the outcomes, again yields a proper scoring rule termed \emph{equivalent}~\citep{Dawid98,GneitingRa07}.
 Throughout we will implicitly apply such affine transformations to obtain the clearest version of the scoring rule.
@@
 If there is a bijection between the sets $\mD$ and $\mM$, then we say that $\mM$ parametrizes $\mD$ and write $p(\cdot\,;\mu)$ for the density mapping to $\mu$.
-Given such a family parametrized be the relevant statistics, the generalized log scoring rule is then
+Given such a family parametrized by the relevant statistics, the generalized log scoring rule is then
 %
@@
 \begin{equation} \label{maxent-prog}
- \min_{p \in \mP} F(p) \st \Exp_{p}[\phi(x)] = \mu,
+ \min_{p \in \mP} \:F(p) \st \Exp_{p}[\phi(x)] = \mu,
 \end{equation}
@@
 
-The following proposition collects the relevant results on regular exponential families; proofs may be found in~\citep{BanerjeeDhGh05b,Barndorff78,WainJordan08}.
+The following proposition collects the relevant results on regular exponential families; proofs may be found in~\citet[Prop.\ 3.1--3.2, Thm.\ 3.3--3.4]{WainJordan08} and see also~\citet[Lem.\ 1, Thm.\ 2]{BanerjeeDhGh05b}.
 A convex function $T$ is of \emph{Legendre type} if it is proper, closed, strictly convex and differentiable on the interior of its domain, and $\lim_{\theta \rightarrow \bar{\theta}} \norm{\grad T(\theta)} = +\infty$ when $\bar{\theta}$ lies on the boundary of the domain.
@@
 From our discussion so far, we have that an exponential family density can be parametrized either by the natural parameter $\theta$, or by the mean parameter $\mu$, and that the two are related by the invertible gradient map $\mu = \grad T(\theta)$.
-We will write $p(x;\theta)$ or $p(x;\mu)$ given the parametrization used, which should be clear from context.
+We will write $p(x;\theta)$ or $p(x;\mu)$ given the parametrization used.
 
@@
 For the converse, assume the defined log score is proper.
-By the Savage characterization of proper scoring rules for expectations (see~\cite{}), we must have
+By the Savage characterization of proper scoring rules for expectations (see~\citet[Thm.\ 1]{GneitingRa07} and~\citet{Savage71}), we must have
 %
@@
 
-As further intuition for the result, note that~(\ref{breg-div}) is the definition of the `Bregman divergence' with respect to strictly convex function $T$. Therefore we have 
+As further intuition for the result, note that~(\ref{breg-div}) is the definition of the `Bregman divergence' with respect to strictly convex function $T$, written $D_T$. Therefore we have 
 %
@@
 %
-where the last equality is a well-known identity relating the Bregman divergences of $T$ and $T^* = G$~\citep{BanerjeeDhGh05b}.
+where the last equality is a well-known identity relating the Bregman divergences of $T$ and $T^* = G$.
 The equation states that the agent's regret from misreporting its mean parameter does not depend on the full density $p$, only the mean $\mu$.
@@
 
-Theorem~\ref{maxent-score} leads to a straightforward procedure for constructing score rules for expectations: define the relevant statistic, and consider the maximum entropy (equivalently, exponential family) distribution consistent with the agent's reported mean $\mu$.
+Theorem~\ref{maxent-score} leads to a straightforward procedure for constructing score rules for expectations.
+Define the relevant statistic, and consider the maximum entropy (equivalently, exponential family) distribution consistent with the agent's reported mean $\mu$.
 The scoring rule compensates the agent according to the log likelihood of the eventual outcome according to this distribution.
@@
 %
-where $\Gamma$ denotes the usual gamma function (the extension of the factorial to the reals). We have not found either scoring rule~(\ref{exp-scoring}) or~(\ref{weibull-scoring}) in the literature.
+where $\Gamma$ denotes the gamma function (the extension of the factorial to the reals). We have not found either scoring rule~(\ref{exp-scoring}) or~(\ref{weibull-scoring}) in the literature.
 \end{example}
@@
 Here the statistics being elicited are the mean vector $\mu$ and the covariance matrix $\Sigma$.
-These scoring rules have been studied by~\citet{dawid99coherent} as rules that only depend on the mean and variance of the reported density.
+These scoring rules have been studied by~\citet{dawid1999coherent} as rules that only depend on the mean and variance of the reported density.
 They note that these rules are weakly proper (because they do not distinguish between densities with the same first and second moments), but do not make the point that knowledge of the full density is not necessary on the part of the agent.
@@
 \begin{equation} \label{lmsr-cost}
-C(\theta) = \log \int_{x \in \mX} \exp \left[\la \theta, \phi(x) \ra\right] \nu(dx).
+C(\theta) = \log \int_{x \in \mX} \exp \la \theta, \phi(x) \ra \,d\nu(x).
 \end{equation}
@@
 
-There is a well-known duality between scoring rules and cost-function based markets. To see this in our context, recall from~(\ref{equalizer-rule}) that
+There is a well-known duality between scoring rules and cost-function based markets~\citep{acv13,hanson03}. To see this in our context, recall from~(\ref{equalizer-rule}) that
 %
@@
 
-Let us now revisit our scoring rules examples from Section~\ref{sec:scoring} in the context of prediction markets.
+Let us now revisit our scoring rule examples from Section~\ref{sec:scoring} in the context of prediction markets.
 The relevant entities now are the payoff function, the effective domain of shares, and the cost function.
@@
 As another example let the outcome space be the $d$-dimensional unit sphere.
-This setting was considered by~\citet{Abernethy11} who provide a cost function implicitly defined through a variational characterization.
+This setting was considered by~\citet{acv13} who provide a cost function implicitly defined through a variational characterization.
 The maximum entropy approach leads to another alternative.
@@
 %
-where $I_r$ refers to the modified Bessel function of first kind and order $r$; see~\citep{BanerjeeDhGh05} for an explanation of these quantities.
+where $I_r$ refers to the modified Bessel function of first kind and order $r$; see~\citet{BanerjeeDhGh05} for an explanation of these quantities.
 The effective domain of $\theta$ is the positive orthant in $\bR^d$.

### 2014-02-10 WriteUp/maxentpredmkts.tex

--- 2014-02-10 WriteUp/maxentpredmkts.tex
@@
 \begin{example} \label{ex:vonmises-market}
+
 As another example let the outcome space be the $d$-dimensional unit sphere.
-This setting was considered by~\citet{Abernethy13} who provide a cost function implicitly defined through a variational characterization.
+This setting was considered by~\citet{acv13} who provide a cost function implicitly defined through a variational characterization.
 The maximum entropy approach leads to another alternative.

### 2014-02-10 WriteUp/maxentpredmkts.tex
Cleaned up bayesian section

@@
 
-In the standard model of cost-function based prediction markets, a sequence of myopic, risk-neutral agents arrive and trade in the market~\citep{CP07,Chen10}.
+In the standard model of cost-function based prediction markets, a sequence of myopic, risk-neutral agents arrive and trade in the market~\citep{CP07,chen2010new}.
 As we saw in Section~\ref{sec:pred-market}, such a trader moves the prices to its own expectation $\mu$.
@@
 In this section we examine the aggregation behavior of the market when agents are Bayesian and take into account the current market state when forming their beliefs.
-This requires more structure to the agents' beliefs.
+This requires more structure to their beliefs.
 For this section and the remainder of the paper, we will assume that agents have \emph{exponential family beliefs}.
@@
 The sufficient statistic maps $\theta$ to $(\theta, T(\theta))$, and the log partition function $\psi$ is defined as the normalizer as usual.
-For a complete treatment of exponential families conjugate priors, see for instance~\citep{Barndorff78}.
-Now~\citet{diaconis79} and~\citet{jewell1974credible} have shown that
+For a complete treatment of exponential families conjugate priors, see for instance~\citet{Barndorff78}.
+Now~\citet[Thm.\ 2]{diaconis79} and~\citet{jewell1974credible} have shown that
 %
@@
 Suppose now that the agent observes an empirical sample with mean $\hmu$ and size $m$.
-By a standard derivation~\citep{diaconis79}, the posterior conjugate prior parameters become $n\nu \leftarrow n\nu + m\hmu$ and $n \leftarrow n+m$, and the posterior expectation~(\ref{eq:prior}) evaluates to
+By a standard derivation~\citep[see][]{diaconis79}, the posterior conjugate prior parameters become $n\nu \leftarrow n\nu + m\hmu$ and $n \leftarrow n+m$, and the posterior expectation~(\ref{eq:prior}) evaluates to
 %
@@
 %
-Thus the posterior mean is a convex combination of the prior and posterior means, and their relative weights depend on the phantom and empirical sample sizes.% need a reference here
-
-Consider Bayesian agents maintaining and exponential family conjugate prior over the data model's natural parameters (equivalently, the expected security payoffs).
+Thus the posterior mean is a convex combination of the prior and empirical means, and their relative weights depend on the phantom and empirical sample sizes.% need a reference here
+
+Consider Bayesian agents maintaining an exponential family conjugate prior over the data model's natural parameters (equivalently, the expected security payoffs).
 Each agent has access to a private sample of the data of size $m$ with mean statistic $\hmu$.

### 2014-02-10 WriteUp/maxentpredmkts.tex
fixed some references

--- 2014-02-10 WriteUp/maxentpredmkts.tex
@@
 For the logarithmic market scoring rule they show that unbounded market maker loss can result in this setting.
-\cite{Abernethy13} and \cite{OthmanS11} specify frameworks under which they design cost function based markets that satisfy the desirable property of bounded market maker loss even in infinite outcome spaces.
+\cite{acv13} and \cite{OthmanS11} specify frameworks under which they design cost function based markets that satisfy the desirable property of bounded market maker loss even in infinite outcome spaces.
 

### 2014-02-10 WriteUp/maxentpredmkts.tex
Cleaned up first part of exp util section

@@
 
+\sebastien{Need to talk about certainty equivalents somewhere}
+
 %Designing prediction markets to handle a large outcome space is an active area of research.
@@
 
-\section{Risk Aversion, Exponential Utility, and Market Equilibrium} \label{sec:exputil}
-
-\subsection{Behavior of Traders endowed with Exponential Utility}
-
-In this section we relax that standard assumption that agents in the market are risk-neutral.
-We show that, with sufficient extra structure to the agents' beliefs and utilities, the market performs a clean aggregation of the agents' expectations in the form of a simple weighted averages.
-Assume that the agent has an exponential utility function for money $w$:
+%\section{Risk Aversion, Exponential Utility, and Market Equilibrium} \label{sec:exputil}
+%\subsection{Behavior of Traders endowed with Exponential Utility}
+
+\section{Risk-Averse Traders with Exponential Utility}
+
+In this section we relax the standard assumption that agents in the market are risk-neutral.
+We show that, with sufficient extra structure to the agents' beliefs and utilities, the market performs a clean aggregation of the agents' expectations in the form of a simple weighted average.
+Assume that the agent has an exponential utility function for wealth $w$:
 %
@@
 
-\jake{This set of ideas is nice and needs to be sold more strongly in the intro} If wealth is distributed according to a probability measure $P$, then the \emph{certainty equivalent} of a random amount of wealth is defined as
+If wealth is distributed according to a probability measure $P$, then the \emph{certainty equivalent} of a random amount of wealth is defined as
 $$
@@
 Suppose as before that the agent's belief over outcomes takes the form of a density $p$ with respect to base measure $\nu$.
-There is a close relationship between the log-partition function and the certainty equivalent under exponential utility.
+There is a close relationship between the log-partition function and the certainty equivalent under exponential utility~\citep[see][]{ben2007old}.
 %
 \begin{lemma} \label{lem:exp-util}
-The certainty equivalent of the agent's expected profit, under exponential utility, when acquiring shares $\delta$ under a market state of $\theta$ is
+The certainty equivalent of the agent's expected profit, with exponential utility, when acquiring shares $\delta$ under a market state of $\theta$ is
 \begin{equation}
-\log a - T_p(-a\delta) - aC(\theta+\delta) + aC(\theta),
-\end{equation}
-where $T_p$ is the log partition function~(\ref{log-part}) with a base measure of $p\,d\nu$. Furthermore, if the agent's belief is an exponential family with parameter $\htheta$, we have
+\log a - T^p(-a\delta) - aC(\theta+\delta) + aC(\theta),
+\end{equation}
+where $T^p$ is the log partition function~(\ref{log-part}) with a base measure of $p\,d\nu$. Furthermore, if the agent's belief is an exponential family with natural parameter $\htheta$, we have
 %
 \begin{equation*}
-T_p(\delta) = T(\htheta + \delta) - T(\htheta),
+T^p(\delta) = T(\htheta + \delta) - T(\htheta),
 \end{equation*}
@@
 & & \CE(\,\la\delta,\phi(x)\ra - [C(\theta+\delta) - C(\theta)]\,)\\
-& = & -\log \int_{\mX} \frac{1}{a} \exp\left( \la-a\delta,\phi(x)\ra + a [C(\theta+\delta) - C(\theta)]  \right) p(x)d\nu(x)\\
-& = & \log a - a [C(\theta+\delta) - C(\theta)] - \log \int_{\mX} \exp\left( \la-a\delta,\phi(x)\ra \right)  p(x)d\nu(x)\\
-& = & \log a - a [C(\theta+\delta) - C(\theta)] - T_p(-a\delta).
-\end{eqnarray*}
-%
-For the second part of the result, we have\jake{Is the $\hat \cdot$ missing from the computations below?}
-%
-\begin{eqnarray*}
-T_p(\delta) & = & \log \int_{\mX} \exp(\la\delta,\phi(x)\ra)\,p(x;\htheta)\, d\nu(x)\\
+& = & -\log \int_{\mX} \frac{1}{a} \exp \left(\la-a\delta,\phi(x)\ra + a [C(\theta+\delta) - C(\theta)]\right) p(x)d\nu(x)\\
+& = & \log a - a [C(\theta+\delta) - C(\theta)] - \log \int_{\mX} \exp\la-a\delta,\phi(x)\ra\, p(x)\,d\nu(x)\\
+& = & \log a - a [C(\theta+\delta) - C(\theta)] - T^p(-a\delta).
+\end{eqnarray*}
+%
+For the second part of the result, we have
+%
+\begin{eqnarray*}
+T^p(\delta) & = & \log \int_{\mX} \exp\la\delta,\phi(x)\ra\,p(x;\htheta)\, d\nu(x)\\
 & = &  \log \int_{\mX} \exp( \la\delta+\htheta,\phi(x)\ra - T(\htheta) )\, d\nu(x)\\
-& = &  T(\htheta+\delta) - T(\theta) + \log \int_{\mX} \exp( \la\delta+\htheta,\phi(x)\ra - T(\htheta+\delta) )\, d\nu(x)\\
-& = &  T(\htheta+\delta) - T(\theta) + \log \int_{\mX} p(x;\htheta+\delta) \, d\nu(x)\\
-& = &  T(\htheta+\delta) - T(\theta),
+& = &  T(\htheta+\delta) - T(\htheta) + \log \int_{\mX} \exp( \la\delta+\htheta,\phi(x)\ra - T(\htheta+\delta) )\, d\nu(x)\\
+& = &  T(\htheta+\delta) - T(\htheta) + \log \int_{\mX} p(x;\htheta+\delta) \, d\nu(x)\\
+& = &  T(\htheta+\delta) - T(\htheta),
 \end{eqnarray*}
@@
 Recall that for the generalized LMSR, the cost function $C$ is exactly the log partition function $T$.
-We are therefore lead to the following clear understanding of a risk-averse agent's behavior in such a market.
+We are therefore lead to the following understanding of a risk-averse agent's behavior in such a market.
 %
@@
 Note that as, $a$ tends to 0, we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$.
-As $a$ grows larger (agent grows more risk averse) the agent makes smaller trades to reduce it exposure, and the final state stays closer to the current state $\theta$.
-Update~(\ref{exp-update}) implies that under the conditions of the theorem, a market that receives a sequence of myopic traders aggregates their natural parameters in the form of an exponentially weighted moving average.
+As $a$ grows larger (the agent grows more risk averse) the agent makes smaller trades to reduce it exposure, and the final state stays closer to the current state $\theta$.
+Update~(\ref{exp-update}) implies that, under the conditions of the theorem, a market that receives a sequence of myopic traders aggregates their natural parameters in the form of an exponentially weighted moving average.
 The final market estimates (i.e., prices) are obtained by applying $\grad T$ to this average.
@@
 \begin{equation} \label{liqu-update}
-\theta + \delta = \frac{\lambda}{\lambda+a} \ttheta + \frac{a}{\lambda+a} \theta
-= \frac{\lambda}{\lambda+a} \grad T^{-1}_{\lambda}(\hmu) + \frac{a}{\lambda+a} \grad T^{-1}_{\lambda}(\mu).
+\theta + \delta \:\:=\:\: \frac{\lambda}{\lambda+a} \ttheta + \frac{a}{\lambda+a} \theta
+\:\:=\:\: \frac{\lambda}{\lambda+a} \grad T^{-1}_{\lambda}(\hmu) + \frac{a}{\lambda+a} \grad T^{-1}_{\lambda}(\mu).
 \end{equation}

### 2014-02-10 WriteUp/maxentpredmkts.tex
added comment

--- 2014-02-10 WriteUp/maxentpredmkts.tex
@@
 \subsection{Behavior of Traders endowed with Exponential Utility}
-
+\sindhu{\cite{norstad99util} for instance?}
 In this section we relax that standard assumption that agents in the market are risk-neutral.
@@
 \end{eqnarray*}
-This follows from Proposition \ref{prop:exp-util} \jake{Sindhu this ref isn't in the doc}\sindhu{it looks like the exp-util section has been rewritten since I edited this: I'll need to check this reference}.
 The maximizer is $\delta_{2}^* = (\Theta - \theta') / (1+a)$, which moves the share vector to $\theta' + \delta_{2}^* = \frac{1}{1+a} \Theta + \frac{a}{1+a} \theta'$ which is a convex combination of the effective belief and the current market state.

### 2014-02-10 WriteUp/maxentpredmkts.tex
cleaned up merge

--- 2014-02-10 WriteUp/maxentpredmkts.tex
@@
 \section{Risk-Averse Traders with Exponential Utility}
+\sindhu{\cite{norstad99util} for instance?}
 

### 2014-02-11 WriteUp/maxentpredmkts.tex
More polishing up

@@
 \noindent
-Given a proper scoring rule $S$ any affine transformation $\tilde{S}(\mu,x) = aS(\mu,x) + b(x)$ of the rule, with $a > 0$ and $b$ an arbitrary real-valued function of the outcomes, again yields a proper scoring rule termed \emph{equivalent}~\citep{Dawid98,GneitingRa07}.
+Given a proper scoring rule $S$ any affine transformation $\tilde{S}(\mu,x) = aS(\mu,x) + b(x)$ of the rule, with $a > 0$ and $b$ an arbitrary real-valued function of the outcomes, again yields a proper scoring rule termed \emph{equivalent}~\citep{Dawid98,gneiting2007strictly}.
 Throughout we will implicitly apply such affine transformations to obtain the clearest version of the scoring rule.
@@
 We let $p(x;\mu)$ be the maximum entropy distribution with expected statistic $\mu$.
-Specifically, it is the solution to the following mathematical program:\footnote{We assume throughout that the minimum is finite and achieved for all $\mu \in \mM$.
+Specifically, it is the solution to the following mathematical program:\footnote{We assume that the minimum is finite and achieved for all $\mu \in \mM$.
 Some care is needed to ensure this holds for specific statistics and outcome spaces.
@@
 For the converse, assume the defined log score is proper.
-By the Savage characterization of proper scoring rules for expectations (see~\citet[Thm.\ 1]{GneitingRa07} and~\citet{Savage71}), we must have
+By the Savage characterization of proper scoring rules for expectations (see~\citet[Thm.\ 1]{gneiting2007strictly} and~\citet{Savage71}), we must have
 %
@@
 
-There is a well-known duality between scoring rules and cost-function based markets~\citep{acv13,hanson03}. To see this in our context, recall from~(\ref{equalizer-rule}) that
+There is a well-known duality between scoring rules and cost-function based markets~\citep{acv13,H03}. To see this in our context, recall from~(\ref{equalizer-rule}) that
 %
@@
 \begin{example} \label{ex:vonmises-market}
-
 As another example let the outcome space be the $d$-dimensional unit sphere.
@@
 Here $\lambda$ is construed as the \emph{inverse liquidity}, or price responsiveness.
-A larger setting of $\lambda$ means fewer shares need to be bought to reach the same prices.\footnote{Observe that the liquidity adjustment to the cost function takes the same form as the risk-aversion adjustment to the exponential utility in~(\ref{exp-util}).
-In convex analysis, this transformation is known as the perspective function~\citep{HiriartLe00}.}
+A larger setting of $\lambda$ means fewer shares need to be bought to reach the same prices.\footnote{The liquidity adjustment to the cost function takes the same form as the risk-aversion adjustment to the exponential utility in~(\ref{exp-util}).
+In convex analysis, this transformation is known as the perspective function~\citep[p.\ 90]{HiriartLe00}.}
 

### 2014-02-11 WriteUp/maxentpredmkts.tex
added missing section label
fixed some broken references

@@
 
-\jake{I want to use this paragraph for a general overview of mm frameworks} One popular form of prediction markets is the market scoring rule~\cite{hanson03}.
+\jake{I want to use this paragraph for a general overview of mm frameworks} One popular form of prediction markets is the market scoring rule~\cite{H03}.
 A market scoring rule considers all trades as a single chronological sequence.
@@
 Thus, each trader has an incentive to minimize the prediction loss.
-In this format, the {\em market maker} who runs the market can suffer an overall loss, but Hanson~\cite{hanson03} showed that, for market scoring rules on finite outcome spaces, the loss of the market maker can be bounded.
+In this format, the {\em market maker} who runs the market can suffer an overall loss, but Hanson~\cite{H03} showed that, for market scoring rules on finite outcome spaces, the loss of the market maker can be bounded.
 
@@
 
-Scoring rules are a measure of merit of a predictor \cite{GneitingRa07}.
-\cite{hanson03} adapted scoring rules to markets with sequential participation of multiple traders.
-A cost function based formulation of markets was first introduced by \cite{CP07}.
+Scoring rules are a measure of merit of a predictor \cite{gneiting2007strictly}.
+\cite{H03} adapted scoring rules to markets with sequential participation of multiple traders.
+A cost function based formulation of markets was first introduced by \cite{chen2007utility}.
 
@@
 
-The standard way to implement a prediction market in the literature, due to~\citet{CP07}, is via a centralized market maker.
+The standard way to implement a prediction market in the literature, due to~\citet{chen2007utility}, is via a centralized market maker.
 The market maker maintains a convex, differentiable cost function $C : \bR^d \rightarrow (-\infty,+\infty]$, where $C(\theta)$ records the revenue collected when the vector of outstanding shares is $\theta$.
@@
 
-In the standard model of cost-function based prediction markets, a sequence of myopic, risk-neutral agents arrive and trade in the market~\citep{CP07,chen2010new}.
+In the standard model of cost-function based prediction markets, a sequence of myopic, risk-neutral agents arrive and trade in the market~\citep{chen2007utility,chen2010new}.
 As we saw in Section~\ref{sec:pred-market}, such a trader moves the prices to its own expectation $\mu$.
@@
 
-\section{Risk-Averse Traders with Exponential Utility}
+\section{Risk-Averse Traders with Exponential Utility} \label{sec:exputil}
 \sindhu{\cite{norstad99util} for instance?}

### 2014-02-11 WriteUp/maxentpredmkts.tex
edited section 5

@@
 \section{Risk-Averse Traders with Exponential Utility} \label{sec:exputil}
-\sindhu{\cite{norstad99util} for instance?}
+\sindhu{Sebastien, do you think we could cite \cite{norstad99util} for instance?}
 
@@
 
-Suppose an agent has exponential family belief parametrized by natural parameter $\hat{\theta}$.
+Suppose an exponential utility agent has exponential family belief parametrized by natural parameter $\hat{\theta}$.
 Based on this belief, let $\delta_{1}^{*}$ be the optimal vector of shares the agent decides to trade on first entering the market.
@@
 %Here $\theta' = \theta+\delta_{1}^{*}+\delta'$ where $\delta'$ captures the movement of the share vector by other traders in the market.
+$$\arg\max_{\delta_{2}} \E_{x\sim p(x;\htheta)} U\left[ (\delta_{1}^{*}+\delta_{2} )\phi(x) - C(\delta_{1}^{*} + \theta) + C(\theta)- C(\delta_{2} + \theta') + C(\theta') \right]$$
+Then if $\Theta\defeq\hat{\theta}-a \delta_{1}^{*}  $ is the effective belief, the trader optimal purchase is given by $\delta_{2}^* = (\Theta - \theta') / (1+a)$, which moves the share vector to $\theta' + \delta_{2}^* = \frac{1}{1+a} \Theta + \frac{a}{1+a} \theta'$ which is a convex combination of the effective belief and the current market state.
+\begin{theorem}\label{thm:reptrades}
+Suppose an exponential utility maximizing trader with utility parameter $a$ who has belief $\hat{\theta}$ makes a purchase $\delta$ in a market.
+On subsequently re-entering this market, he will behave identically to an exponential utility maximizing trader with belief $\hat{\theta}-a \delta$ and no prior exposure in the market.
+\end{theorem}
+Theorem \ref{thm:reptrades} implies that financial exposure can be equivalently understood as changing the privately held beliefs.
+
+\subsection{Equilibrium Market State for Exponential Utility Agents}
+\jake{need to mention the connection to potential games in the intro, it's a nice little extra bit.}
+We have shown that every exponential-utility maximizing trader picks the share vector $\delta$ so that the eventual market state can be represented as a convex combination of the initial market state and the natural parameter of his (exponential family) belief distribution.
+In this section we will compute the equilibrium state in an exponential family market with multiple exponential-utility maximizing traders.
+
+We now review a well-known result from game theory regarding the class of \emph{potential games} \cite{monderer1996potential}.
+Note that we say a function $f(\vec{x})$ is at a \emph{local optimum} if changing any coordinate of $\vec{x}$ does not increase the value of $f$.
+\begin{thm}\label{thm:potl}
+Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}\defeq (\delta_{1},\ldots,\delta_{i},\ldots,\delta_{n})$.
+If there exists a potential function $\Phi(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=\Phi(\vec{\delta})-\Phi(\vec{\delta}_{-i},\delta_{i}')$$
+then $\vec{\delta}$ is a Nash equilibrium if and only if $\Phi(\vec{\delta})$ is at a local optimum.
+\end{thm}
+In the exponential family market, the cost function $C$ is identical to the log partition function $T$ defined in (\ref{log-part}).
+Let $\vec{\delta}$ be the vector of vectors of shares purchased by every trader in the market when the market has reached equilibrium, $\theta$ the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his utility parameter.
+\jake{I need to fix a small technical detail with the fact that we are using the log utilities to use the potential argument.}
+
+Define a potential function as  $$\Phi(\vec{\delta})\defeq  T\left(\theta+\sum_{i}\delta_{i}\right)+\sum_{i}\frac{1}{a_{i}}T(\htheta_{i}-a_{i}\delta_{i})$$
+
+Now the utility of trader $i$ is $U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i})$.
+Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $\Phi(\vec{\delta})$ for each $\delta_{i}$.
+\begin{eqnarray*}
+\nabla_{\delta_{i}}\Phi(\vec{\delta})&=&\nabla T\left(\theta+\sum_{j=1}^{n}\delta_{i}\right)-\nabla T(\htheta_{i}-a_{i}\delta_{i}) = 0
+\end{eqnarray*}¥
+And we have the following expression for the final market state.
+$$\theta+\sum_{j=1}^{n}\delta_{j}=
+\frac{\theta+\sum_{i=1}^{n}\left(\frac{\htheta_{i}}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
+Thus, the equilibrium state is a convex combination of trader beliefs and initial market state.
+
+\section{The Exponential Family Market Mechanism with Budgets} \label{sec:budgets}
+\sindhu{\cite{Azoury01}}
+In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
+
+In this section, we consider the the prediction market setup with traders that may be either informative or malicious.
+The malicious traders may want to inject faulty information into the market.
+The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.
+
+We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market.
+Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
+This idea of a dual analysis has been used in a recommender system context to compute reputations in \cite{ressami07}.
+
+In this section, we assume that the traders have exponential family beliefs.
+The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
+\subsection{Budget-limited Aggregation}
+Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
+We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market.
+This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader.
+Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.
+
+In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
+Let $\alpha$ be the budget of a trader in the market.
+Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
+Let $C$ be the cost function.
+Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
+In this case, we want to budget-limit the trader's influence on the market state.
+
+We define the budget-limited final market state as $\theta'$. 
+Here, we consider a specific functional form of $\theta'$: $$\theta'=\lambda\htheta+(1-\lambda)\theta$$  where $$\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$$
+
+First, we  show that this trade is feasible given the trader's budget:
+\begin{theorem}
+Let the current market state be given by $\theta$.
+Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
+The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$.
+\end{theorem}
+\begin{proof}
+From the convexity of $C$, we have
+\begin{eqnarray*}
+C(\theta')&\leq&(1-\lambda)C(\theta)+\lambda C(\htheta)\\
+\mbox{Thus, } C(\theta') - C(\theta) &\leq& \lambda\ (C(\htheta) - C(\theta))\\
+\mbox{And, } C(\theta') - C(\theta) &\leq& \alpha 
+ \end{eqnarray*}\end{proof}
+
+ We note that moving to $\theta'$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
+In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\theta$.
+Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\htheta$.
+\subsection{Damage Bound}
+In this section, we will quantify the error in prediction that the market maker might have to endure as a result of malicious entities entering the market.
+We assume that these malicious entities trade in multiple instances of the market; thus the exposure of the market maker is over several \emph{rounds}.
+We use the standard log loss to measure this error in terms of the initial budget of traders.
+
+We define the loss function for $\theta$ shares held:
+$$L(\theta,x)=-\log(P_{\theta}(x))=\log\int \exp\{\theta^{T}\phi(x)\}\ dx-\theta^{T}\phi(x)=C(\theta)-\theta^{T}\phi(x)$$
+%with the correspondence between $\beta$ and $\theta$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
+
+Suppose the prediction market runs over multiple rounds $t$.
+Let $\theta_{0}^{t}$ be the initial number of shares of each security that are held.
+Let $\hat{\theta}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
+Let us assume that at this point the outcome is revealed; that is, we receive the value of the random variable $x^{t}$.
+
+Over multiple instances of the prediction market, we can track the change in budget of each trader.
+Let the budget at rounds $t$ and $t-1$ be $\alpha^{t}$ and $\alpha^{t-1}$ respectively.
+The change in budget for the trader is
+\begin{eqnarray*}
+\alpha^{t}-\alpha^{t-1}&=&C(\theta^{t})-C(\theta'^{t})-(\theta^{t}-\theta'^{t})^{T}\phi(x^{t})\\
+&=&L(\theta,x^{t})-L(\theta',x^{t})
+\end{eqnarray*}
+
+Define the myopic impact of a trader $i$ in segment $t$ as
+$$\Delta_{i}^{t}:=L(\hat{\theta}_{i-1}^{t},x^{t})-L(\hat{\theta}_{i}^{t},x^{t})$$
+Thus, the myopic impact captures incremental gain in prediction due to the trader in a round.
+Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
+
+The total myopic impact due to all $k$ active traders is given by
+$$\Delta^{t}=L(\theta_{0}^{t},x^{t})-L(\hat{\theta}_{k}^{t},x^{t})$$
+Thus $-\Delta^{t}$ captures the incremental loss of the market prediction after aggregation of all $k$ traders.
+
+\begin{theorem}
+A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
+\end{theorem}
+\begin{proof}
+Consider the myopic impact of a single trader $i$ after participating in the market $T$ times.
+Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
+$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
+
+Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
+\end{proof}
+
+This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders. %imp
+
+In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation. 
+%We provide an information-theoretic justification for this claim in Section \ref{sec:infth}. 
+The intuition behind this  is that a trader's prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
+
+\subsection{Budget of Informative Traders}\label{sec:inf}
+Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
+
+We now characterize the expected change in budget for an informative trader. The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
+\begin{theorem}\label{thm:growth}
+Let $\theta$ be the current market state in the exponential family prediction market.
+Suppose that an  informative trader with belief distribution  parametrized by $\theta'$  moves the market state to the budget-limited state $\htheta=\lambda \theta' + (1-\lambda) \theta$.
+Then, the expectation (over the trader's belief) of the trader's profit  is greater than zero whenever his budget is positive and his belief differs from the previous market position $\theta$.
+% Suppose that each informative trader gets a random sample of data, 
+%resulting in a sequence of trader beliefs $\theta_i$, and hence a sequence of budget-limited market positions
+%$\hat{\theta}_i$, before a final outcome $x$.
+Then, the expectation, over trader $i$'s belief distribution $\theta_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\hat{\theta}_{i-1}$.
+\end {theorem}
+\begin{proof}
+Let the cost function $C$ be equal to the log partition function $T$ of the belief distribution.
+The payoff is given by the sufficient statistics $\phi(x)$.
+Then, the trader's expected net payoff is given by
+\begin{eqnarray*}
+&& \E_{x\sim P_{\htheta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
+&=& T(\theta)-\theta\E_{x\sim P_{\htheta}}[\phi(x)]-(T(\theta')-\theta' \E_{x\sim P_{\htheta}}[\phi(x)])\\
+&=& T(\theta)-\theta\nabla T(\htheta)-(T(\theta')-\theta' \nabla T(\htheta))\\
+&=& T(\theta)-T(\htheta)-\nabla T(\htheta)(\theta-\htheta)-(T(\theta')-T(\htheta)- \nabla T(\htheta)(\theta'-\htheta))\\
+&=& D_{T}(\theta,\htheta)-D_{T}(\theta',\htheta)\\
+&\geq&\lambda D_{T}(\theta,\htheta)\geq0
+\end{eqnarray*}
+The second to last inequality holds since $D_{T}(\theta',\htheta)$ is convex in $\theta'$ and we have:
+\begin{eqnarray*}
+D_{T}(\theta',\htheta)&=&D_{T}\left(\lambda\htheta+(1-\lambda)\theta,\htheta\right)\\
+&\leq&\lambda D_{T}( \htheta ,\htheta)+ (1-\lambda) D_{T}( \theta ,\htheta)\\
+&=&(1-\lambda) D_{T}( \theta ,\htheta)
+%\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
+\end{eqnarray*}¥
+Thus, a trader who moves the market state can expect his profit to be positive and at least $\lambda D_{T}(\theta,\htheta)$.
+ \end{proof}
+
+For continuous distributions with a density, the probability that a trader with private information will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs.
+This result suggests that, eventually, every informative trader will have the ability to influence the market state in accordance with his beliefs, without being budget limited.
+
+Notice that Theorem~\ref{thm:growth} only required that the market state to which the trader moves, be representable as a convex combination of the current market state and his belief.
+This means that the result holds for exponential utility traders aiming to maximize their utility.
+In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{T}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
+
+We note one important aspect of Theorem~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
+This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure.
+If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
+%\pagebreak
+
+%\subsection{Exponential Utility}
+%Now suppose the trader has exponential utility, given by $$U(w) = -\frac{1}{a} \exp(-aw)$$ where $a$ is the coefficient of risk aversion.
+Suppose the trader wishes to maximize his expected utility.
+Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by
+%\begin{eqnarray*}
+%&&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}U\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]\\
+%&=&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}-\frac{1}{a}\exp\left[ -a\delta\phi(x) +a C(\delta + \theta) -a C(\theta)\right]
+%\end{eqnarray*}¥
+%
+%\begin{eqnarray*}
+%\E_{\beta\sim b}\E_{x\sim \beta}\exp\left[-a\delta\phi(x)\right]&=&\int_{\mathcal{B}}\int_{\mathcal{X}}\exp[-a\delta\phi(x)]\, \exp\{\beta\cdot\phi(x)-\psi(\beta)\}\, dx\\
+%&&\qquad\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\, d\beta\\
+%&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
+%&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\int_{\mathcal{X}}\exp[(\beta-a\delta)\cdot\phi(x)-\psi(\beta-a\delta)]\, dx\\
+%&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
+%&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\\
+%\end{eqnarray*}¥
+%\pagebreak
+
+\section{Conclusions}
+\sindhu{Do we need this section?}
+%
+%----------------------------------------------------------------------------------------
+%	BIBLIOGRAPHY
+%----------------------------------------------------------------------------------------
+\bibliographystyle{acmsmall}
+\bibliography{maxentpredmkts}
+\end{document}
+
+%----------------------------------------------------------------------------------------
+
+\subsection{Properties of the Maximum Entropy Market}
+Consider the dual of the cost function, $C^{*}(\mu)$ defined as
+$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
+This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
+$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
+Thus, the mean parameters are the dual variables to the natural parameters.
+
+\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
+\proof
+To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
+\begin{eqnarray*}
+-H(p) &=& \int_{x}p(x)\log p(x)\\
+&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
+&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
+&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
+\end{eqnarray*}
+\unproof
+\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
+\rmk The LMSR is essentially a special case applied to a multinomial distribution.
+The LMSR is known to have bounded ($\log n$) market maker loss.
+Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
+\unrmk
+
+Consider the dual of the cost function, $C^{*}(\mu)$ defined as
+$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
+This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
+$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
+Thus, the mean parameters are the dual variables to the natural parameters.
+
+\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
+\proof
+To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
+\begin{eqnarray*}
+-H(p) &=& \int_{x}p(x)\log p(x)\\
+&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
+&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
+&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
+\end{eqnarray*}
+\unproof
+\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
+\rmk The LMSR is essentially a special case applied to a multinomial distribution.
+The LMSR is known to have bounded ($\log n$) market maker loss.
+Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
+\unrmk
+
+\subsection{An Information-Theoretic Interpretation}
+\label{sec:infth}
+
+ We observe a useful alternative view of the market 
+scoring rule prediction market for exponential family learning. We connect the cost, payoff and profit function to information-theoretic 
+quantities associated with the exponential family.
+
+The following result has been previously pointed out by Amari \cite{Amari-KL}. 
+\begin{theorem}\label{lem:profit_decomposition}{\bf (Profit Decomposition)}: Consider an exponential 
+family $\family$ of distributions over some set of statistics $\suff(x)$, with natural parameters $\beta$. Let $\pi, \rho \in \family$ be any two probability distributions in the family. We use 
+$\beta_\pi$ to denote the natural parameters of $\pi$ and $\mu_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
+we can define $\beta_\rho$ and  $\mu_\rho$. Let $\lpf(\beta_{\rho})$ indicate the log partition function of $\rho$.
+Let $H(\pi)$ denote the entropy of the distribution $\pi$, and 
+ $K(\pi||\rho)$ denote the KL-divergence of $\rho$ relative to $\pi$.
+Then, the following equality holds:  
+\begin{equation}
+ K(\pi||\rho) + H(\pi) = \lpf(\beta_{\rho}) - \beta_\rho \cdot \mu_\pi
+\label{eq:profit_decomposition}
+\end{equation}
+\end{theorem}
+\begin{proof}
+\begin{eqnarray*}
+ K(\pi||\rho) + H(\pi) &=& \int_x \pi(x)\log{\frac{\pi(x)}{\rho(x)}}dx\\
+ &&\quad - \int_x \pi(x)\log{\pi(x)}dx \\
+&=& - \int_x \pi(x)\log{\rho(x)}dx \\
+&=& - \int_x \pi(x)\left[ \beta_\rho \cdot \suff(x) - \lpf(\beta_{\rho})\right]dx\\
+&=& - \beta_\rho \cdot \int_x \pi(x) \suff(x) dx\\
+&&\quad + \lpf(\beta_{\rho}) \int_x \pi(x) dx \\
+&=&  \lpf(\beta_{\rho}) - \beta_\rho \cdot \mu_\pi
+\end{eqnarray*}
+\end{proof}
+
+Equation~\ref{eq:profit_decomposition} gives us an alternative view of the
+market scoring rule construction. Assume that $\pi$ is the true distribution, and consider two arbitrary
+distributions $\rho_1, \rho_2 \in \family$.
+Note that $\lpf(\rho)$  is independent of $\pi$, and $\betavec_\rho \cdot \muvec_\pi$ is linear in the probabilities $\pi(x)$.
+If we want to measure loss by the KL-divergence, we can do so (in expectation) by setting a cost function that captures the first term, and defining security quantities ($\betavec$) and payoffs ($\suff$) to capture the second term.
+In particular, in a market with cost function $\lpf$, if the
+market price initially implies a distribution $\rho_1$, and a trader moves
+the market to price than implies a distribution $\rho_2$, then the cost she
+incurs is $\lpf(\rho_2) - \lpf(\rho_1)$.
+The number of securities bought to make this trade is given by the vector ($\betavec_{\rho_2} - \betavec_{\rho_1}$), and the expected payoff of the securities are given by $\muvec_\pi$.
+Thus, by equation \ref{eq:profit_decomposition}, the {\em net profit} of the trader is equal to $K(\pi||\rho_1) - K(\pi||\rho_2)$, {\it i.e.}, the reduction in KL-divergence with respect to the true distribution.
+We note one useful property of this construction: For a fixed vector $\betavec$ of purchased securities, the cost is independent of the outcome
+(and outcome distribution $\pi$), while the payoff is independent of the initial market state in which these securities were purchased.
+
+\subsection{Properties of the Maximum Entropy Market}
+
+{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
+\proof
+Note that, for an exponential family distribution
+\begin{eqnarray*}
+H(p) &=& -\int_{x}p(x)\log p(x)dx\\
+&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
+&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
+\end{eqnarray*}
+If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
+\unproof
+
+%{\cor The exponential family market maker has unbounded worst case loss.}
+%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
+%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
+%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
+\paragraph{Loss of the market maker}
+First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
+
+Recall the definition of the conjugate dual:
+$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
+The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
+$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
+This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
+So we have 
+$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
+Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
+
+Let $q_{0}$ be the initial and $q_{f}$ the final market state.
+Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
+\begin{eqnarray*}
+\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
+&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
+&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
+&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
+&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
+&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
+&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
+&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
+&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
+\end{eqnarray*}
+
+\subsection{Example: Gaussian Markets}
+{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
+In this case,  $\beta$ is a $2$-dimensional  vector.
+Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
+So
+\begin{eqnarray*}
+\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
+&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
+&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
+\end{eqnarray*}
+Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
+$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
+$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
+{\ex
+We will now derive the expression for the differential entropy for the normal distribution.
+Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
+\begin{eqnarray*}
+H(p) &=& -\int_{x}p(x)\log p(x)dx\\
+&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
+&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
+&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
+&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
+&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
+\end{eqnarray*}
+Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
+Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
+
+{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
+Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
+Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
+Let's now work out the loss for the Gaussian market maker.
+First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
+\begin{eqnarray*}
+D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
+&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
+&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
+&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
+&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
+%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
+&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
+\end{eqnarray*}
+
+So the Gaussian market maker loss is 
+%parallels to quadratic market maker loss? 
+\begin{eqnarray*}
+D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
+&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
+&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
+\end{eqnarray*}¥
+Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
+\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
+%\pagebreak
+
+\subsection{Proper Scoring from Consistency}
+
+Our starting point for designing proper scoring rules is the classic logarithmic scoring rule for eliciting probabilities in the case of finite outcomes.
+This rule is simply $S(p,x) = \log p(x)$ and there are several ways to understand why it is proper, with different implications for generalization.
+
+First observe that the scoring rule compensates the agent with the log likelihood it assigned to the outcome, so the agent is maximizing the expected log likelihood via its report.
+Now it is a fundamental result in statistics that, under a variety of different sufficient conditions, the maximum likelihood estimator is statistically consistent; see~\cite{} for an overview of such conditions.
+This means that in the limit (as the size of the i.i.d.\ empirical sample increases) the log likelihood is maximized by the true parameters.
+By this reasoning, the following result follows immediately from the definition of consistency and the law of large numbers.
+%
+\begin{theorem} \label{max-like}
+Suppose that $\mM$ parametrizes $\mD$. The logarithmic scoring rule defined by
+%
+\begin{equation} \label{log-score}
+S(\mu, x) = \log p(x;\mu),
+\end{equation}
+%
+where $\mu \in \mM$, is proper over $\mD$ if and only if the maximum likelihood estimator for $\mu$ is consistent.
+\end{theorem}
+%
+We note that although our focus is on parametrizations by expectations, the preceding theorem in fact holds for more generic parameter spaces.
+The range of scoring rules it provides is broader than those characterized by~\cite{}, because it allows for a restriction over the set of beliefs.
+On the other hand, its applicability is limited to those domains $\mD$ that can be parametrized by the relevant statistics.
+The following provide concrete examples of these points.
+%
+\begin{example}
+Suppose the outcomes are supported on $[1,+\infty)$ and follow a Pareto distribution with density $f(x;\alpha) = \alpha/x^{\alpha+1}$ parametrized by an index $\alpha > 0$.
+The mean $\mu$ is related to the index via the one-to-one mapping $\mu = \frac{\alpha}{\alpha-1}$, so the density can alternatively be parametrized by the mean.
+Theorem~\ref{max-like} gives the following proper scoring rule for the mean of a Pareto distribution with support on $[1,+\infty)$:
+%
+\begin{equation} \label{pareto-scoring}
+S(\mu,x) = \log \frac{\mu}{\mu-1} - \left(\frac{\mu}{\mu-1} + 1\right) \log x.
+\end{equation}
+%
+We stress that the rule can only elicit the mean assuming the agent knows its belief is a Pareto distribution over $[1,+\infty]$.
+It does not elicit the mean of other families of densities parametrized by the mean (e.g., the exponential distribution).
+\end{example}
+%
+%
+\begin{example}
+Suppose the outcomes are supported on $(-\infty,+\infty)$ and follow a Cauchy distribution with density $f(x;m) = 1/\pi[1+(x-m)^2]$, parametrized by the median $m$.
+Theorem~\ref{max-like} leads to the following proper scoring rule for the median:
+%
+\begin{equation} \label{cauchy-scoring}
+S(m,x) = -\log[1+(x-m)^2].
+\end{equation}
+%
+Note that the median cannot be obtained as the expectation of any statistic. This highlights the range of parameters than may be elicited by the log score in different circumstances.
+\end{example}
+%
+
+\subsection{Repeated Trading and the Effective Belief}
+On a subsequent entry into this market with market state $\theta'$, his optimal purchase $\delta_{2}^{*}$ is given by the solution of
+Here $\theta' = \theta+\delta_{1}^{*}+\delta'$ where $\delta'$ captures the movement of the share vector by other traders in the market.
+
+$$\arg\max_{\delta_{2}} \E_{x\sim p(x;\htheta)} U\left[ (\delta_{1}^{*}+\delta_{2} )\phi(x) - C(\delta_{1}^{*} + \theta) + C(\theta)- C(\delta_{2} + \theta') + C(\theta') \right]$$
+
 \begin{eqnarray*}
@@
 This means that financial exposure can be equivalently understood as changing the privately held beliefs.
-%\pagebreak
-
 \subsection{Equilibrium Market State for Exponential Utility Agents}
-\jake{need to mention the connection to potential games in the intro, it's a nice little extra bit.}
-We have shown that every exponential-utility maximizing trader picks the share vector $\delta$ so that the eventual market state can be represented as a convex combination of the initial market state and the natural parameter of his (exponential family) belief distribution.
-In this section we will compute the equilibrium state in an exponential family market with multiple exponential-utility maximizing traders.
-
-We now review a well-known result from game theory regarding the class of \emph{potential games} \cite{monderer1996potential}.
-Note that we say a function $f(\vec{x})$ is at a \emph{local optimum} if changing any coordinate of $\vec{x}$ does not increase the value of $f$.
-\begin{thm}\label{thm:potl}
-Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}\defeq (\delta_{1},\ldots,\delta_{i},\ldots,\delta_{n})$.
-If there exists a potential function $\Phi(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=\Phi(\vec{\delta})-\Phi(\vec{\delta}_{-i},\delta_{i}')$$
-then $\vec{\delta}$ is a Nash equilibrium if and only if $\Phi(\vec{\delta})$ is at a local optimum.
-\end{thm}
-In the exponential family market, the cost function $C$ is identical to the log partition function $T$.
-Let $\vec{\delta}$ be the vector of vectors of shares purchased by every trader in the market when the market has reached equilibrium, $\theta$ the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his utility parameter.
-\jake{I need to fix a small technical detail with the fact that we are using the log utilities to use the potential argument.}
-
-Define a potential function as  $$\Phi(\vec{\delta})\defeq  T\left(\theta+\sum_{i}\delta_{i}\right)+\sum_{i}\frac{1}{a_{i}}T(\htheta_{i}-a_{i}\delta_{i})$$
-
-%is the following expression for utility correct?
-Now the utility of trader $i$ is $U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i})$.
-Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $\Phi(\vec{\delta})$ for each $\delta_{i}$.
-\begin{eqnarray*}
-\nabla_{\delta_{i}}\Phi(\vec{\delta})&=&\nabla T\left(\theta+\sum_{j=1}^{n}\delta_{i}\right)-\nabla T(\htheta_{i}-a_{i}\delta_{i}) = 0
-\end{eqnarray*}¥
 This can be achieved by equating the arguments. That is, for each trader $i$, 
@@
 Thus the equilibrium state is a convex combination of trader beliefs and initial market state.
-
-\section{The Exponential Family Market Mechanism with Budgets} \label{sec:budgets}
-\sindhu{\cite{Azoury01}}
-In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
-
-In this section, we consider the the prediction market setup with traders that may be either informative or malicious.
-The malicious traders may want to inject faulty information into the market.
-The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.
-
-We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market.
-Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
-This idea of a dual analysis has been used in a recommender system context to compute reputations in \cite{ressami07}.
-
-In this section, we assume that the traders have exponential family beliefs.
-The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
-\subsection{Budget-limited Aggregation}
-Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
-We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market.
-This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader.
-Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.
-
-In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
-Let $\alpha$ be the budget of a trader in the market.
-Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
-Let $C$ be the cost function.
-Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
-In this case, we want to budget-limit the trader's influence on the market state.
-
-We define the budget-limited final market state as $\theta'$. 
-Here, we consider a specific functional form of $\theta'$: $$\theta'=\lambda\htheta+(1-\lambda)\theta$$  where $$\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$$
-
-First, we  show that this trade is feasible given the trader's budget:
-\begin{theorem}
-Let the current market state be given by $\theta$.
-Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
-The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$.
-\end{theorem}
-\begin{proof}
-From the convexity of $C$, we have
-\begin{eqnarray*}
-C(\theta')&\leq&(1-\lambda)C(\theta)+\lambda C(\htheta)\\
-\mbox{Thus, } C(\theta') - C(\theta) &\leq& \lambda\ (C(\htheta) - C(\theta))\\
-\mbox{And, } C(\theta') - C(\theta) &\leq& \alpha 
- \end{eqnarray*}\end{proof}
-
- We note that moving to $\theta'$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
-In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\theta$.
-Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\htheta$.
-\subsection{Damage Bound}
-In this section, we will quantify the error in prediction that the market maker might have to endure as a result of malicious entities entering the market.
-We assume that these malicious entities trade in multiple instances of the market; thus the exposure of the market maker is over several \emph{rounds}.
-We use the standard log loss to measure this error in terms of the initial budget of traders.
-
-We define the loss function for $\theta$ shares held:
-$$L(\theta,x)=-\log(P_{\theta}(x))=\log\int \exp\{\theta^{T}\phi(x)\}\ dx-\theta^{T}\phi(x)=C(\theta)-\theta^{T}\phi(x)$$
-%with the correspondence between $\beta$ and $\theta$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
-
-Suppose the prediction market runs over multiple rounds $t$.
-Let $\theta_{0}^{t}$ be the initial number of shares of each security that are held.
-Let $\hat{\theta}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
-Let us assume that at this point the outcome is revealed; that is, we receive the value of the random variable $x^{t}$.
-
-Over multiple instances of the prediction market, we can track the change in budget of each trader.
-Let the budget at rounds $t$ and $t-1$ be $\alpha^{t}$ and $\alpha^{t-1}$ respectively.
-The change in budget for the trader is
-\begin{eqnarray*}
-\alpha^{t}-\alpha^{t-1}&=&C(\theta^{t})-C(\theta'^{t})-(\theta^{t}-\theta'^{t})^{T}\phi(x^{t})\\
-&=&L(\theta,x^{t})-L(\theta',x^{t})
-\end{eqnarray*}
-
-Define the myopic impact of a trader $i$ in segment $t$ as
-$$\Delta_{i}^{t}:=L(\hat{\theta}_{i-1}^{t},x^{t})-L(\hat{\theta}_{i}^{t},x^{t})$$
-Thus, the myopic impact captures incremental gain in prediction due to the trader in a round.
-Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
-
-The total myopic impact due to all $k$ active traders is given by
-$$\Delta^{t}=L(\theta_{0}^{t},x^{t})-L(\hat{\theta}_{k}^{t},x^{t})$$
-Thus $-\Delta^{t}$ captures the incremental loss of the market prediction after aggregation of all $k$ traders.
-
-\begin{theorem}
-A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
-\end{theorem}
-\begin{proof}
-Consider the myopic impact of a single trader $i$ after participating in the market $T$ times.
-Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
-$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
-
-Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
-\end{proof}
-
-This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders. %imp
-
-In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation. 
-%We provide an information-theoretic justification for this claim in Section \ref{sec:infth}. 
-The intuition behind this  is that a trader's prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
-
-\subsection{Budget of Informative Traders}\label{sec:inf}
-Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
-
-We now characterize the expected change in budget for an informative trader. The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
-\begin{theorem}\label{thm:growth}
-Let $\theta$ be the current market state in the exponential family prediction market.
-Suppose that an  informative trader with belief distribution  parametrized by $\theta'$  moves the market state to the budget-limited state $\htheta=\lambda \theta' + (1-\lambda) \theta$.
-Then, the expectation (over the trader's belief) of the trader's profit  is greater than zero whenever his budget is positive and his belief differs from the previous market position $\theta$.
-% Suppose that each informative trader gets a random sample of data, 
-%resulting in a sequence of trader beliefs $\theta_i$, and hence a sequence of budget-limited market positions
-%$\hat{\theta}_i$, before a final outcome $x$.
-Then, the expectation, over trader $i$'s belief distribution $\theta_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\hat{\theta}_{i-1}$.
-\end {theorem}
-\begin{proof}
-Let the cost function $C$ be equal to the log partition function $T$ of the belief distribution.
-The payoff is given by the sufficient statistics $\phi(x)$.
-Then, the trader's expected net payoff is given by
-\begin{eqnarray*}
-&& \E_{x\sim P_{\htheta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
-&=& T(\theta)-\theta\E_{x\sim P_{\htheta}}[\phi(x)]-(T(\theta')-\theta' \E_{x\sim P_{\htheta}}[\phi(x)])\\
-&=& T(\theta)-\theta\nabla T(\htheta)-(T(\theta')-\theta' \nabla T(\htheta))\\
-&=& T(\theta)-T(\htheta)-\nabla T(\htheta)(\theta-\htheta)-(T(\theta')-T(\htheta)- \nabla T(\htheta)(\theta'-\htheta))\\
-&=& D_{T}(\theta,\htheta)-D_{T}(\theta',\htheta)\\
-&\geq&\lambda D_{T}(\theta,\htheta)\geq0
-\end{eqnarray*}
-The second to last inequality holds since $D_{T}(\theta',\htheta)$ is convex in $\theta'$ and we have:
-\begin{eqnarray*}
-D_{T}(\theta',\htheta)&=&D_{T}\left(\lambda\htheta+(1-\lambda)\theta,\htheta\right)\\
-&\leq&\lambda D_{T}( \htheta ,\htheta)+ (1-\lambda) D_{T}( \theta ,\htheta)\\
-&=&(1-\lambda) D_{T}( \theta ,\htheta)
-%\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
-\end{eqnarray*}¥
-Thus, a trader who moves the market state can expect his profit to be positive and at least $\lambda D_{T}(\theta,\htheta)$.
- \end{proof}
-
-For continuous distributions with a density, the probability that a trader with private information will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs.
-This result suggests that, eventually, every informative trader will have the ability to influence the market state in accordance with his beliefs, without being budget limited.
-
-Notice that Theorem~\ref{thm:growth} only required that the market state to which the trader moves, be representable as a convex combination of the current market state and his belief.
-This means that the result holds for exponential utility traders aiming to maximize their utility.
-In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{T}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
-
-We note one important aspect of Theorem~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
-This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure.
-If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
-%\pagebreak
-
-%\subsection{Exponential Utility}
-%Now suppose the trader has exponential utility, given by $$U(w) = -\frac{1}{a} \exp(-aw)$$ where $a$ is the coefficient of risk aversion.
-Suppose the trader wishes to maximize his expected utility.
-Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by
-%\begin{eqnarray*}
-%&&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}U\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]\\
-%&=&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}-\frac{1}{a}\exp\left[ -a\delta\phi(x) +a C(\delta + \theta) -a C(\theta)\right]
-%\end{eqnarray*}¥
-%
-%\begin{eqnarray*}
-%\E_{\beta\sim b}\E_{x\sim \beta}\exp\left[-a\delta\phi(x)\right]&=&\int_{\mathcal{B}}\int_{\mathcal{X}}\exp[-a\delta\phi(x)]\, \exp\{\beta\cdot\phi(x)-\psi(\beta)\}\, dx\\
-%&&\qquad\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\, d\beta\\
-%&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
-%&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\int_{\mathcal{X}}\exp[(\beta-a\delta)\cdot\phi(x)-\psi(\beta-a\delta)]\, dx\\
-%&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
-%&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\\
-%\end{eqnarray*}¥
-%\pagebreak
-
-\section{Conclusions}
-%
-%----------------------------------------------------------------------------------------
-%	BIBLIOGRAPHY
-%----------------------------------------------------------------------------------------
-\bibliographystyle{acmsmall}
-\bibliography{maxentpredmkts}
-\end{document}
-
-%----------------------------------------------------------------------------------------
-
-\subsection{Properties of the Maximum Entropy Market}
-Consider the dual of the cost function, $C^{*}(\mu)$ defined as
-$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
-This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
-$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
-Thus, the mean parameters are the dual variables to the natural parameters.
-
-\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
-\proof
-To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
-\begin{eqnarray*}
--H(p) &=& \int_{x}p(x)\log p(x)\\
-&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
-&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
-&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
-\end{eqnarray*}
-\unproof
-\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
-\rmk The LMSR is essentially a special case applied to a multinomial distribution.
-The LMSR is known to have bounded ($\log n$) market maker loss.
-Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
-\unrmk
-
-Consider the dual of the cost function, $C^{*}(\mu)$ defined as
-$$C^{*}(\mu) = \sup_{\beta} \beta\cdot\mu - C(\beta)$$
-This supremum is obtained at the value of $\beta$ for which $\mu = \nabla C(\beta)$; that is the natural parameter $\beta$ for which $\mu=\E_{p_{\beta}}[\phi(x)]$ is the mean parameter. Rewriting, 
-$$C^{*}(\E_{p_{\beta}}[\phi(x)]) = \beta\cdot\E_{p_{\beta}}[\phi(x)] - C(\beta)$$
-Thus, the mean parameters are the dual variables to the natural parameters.
-
-\begin{thm} The value of the dual of the cost function is the negative entropy of the exponential family distribution obtained from backward mapping the mean parameters. \end{thm}
-\proof
-To see this, we note that for $p(x)=\exp\{\beta\cdot\phi(x)-C(\beta)\}$
-\begin{eqnarray*}
--H(p) &=& \int_{x}p(x)\log p(x)\\
-&=& \int_{x}p(x)[\beta\cdot\phi(x)-C(\beta)]\\
-&=& \beta\cdot\int_{x}p(x)[\phi(x)]-\int_{x}p(x)C(\beta)\\
-&=& \beta\cdot\E_{p}[\phi(x)]-C(\beta)
-\end{eqnarray*}
-\unproof
-\rmk This result shows a nice parallel to LMSR, since the dual of the cost function of the LMSR is the negative entropy.\unrmk
-\rmk The LMSR is essentially a special case applied to a multinomial distribution.
-The LMSR is known to have bounded ($\log n$) market maker loss.
-Thus, while in general the exponential family LMSR does not guarantee bounded market maker loss, for some special cases it can.
-\unrmk
-
-\subsection{An Information-Theoretic Interpretation}
-\label{sec:infth}
-
- We observe a useful alternative view of the market 
-scoring rule prediction market for exponential family learning. We connect the cost, payoff and profit function to information-theoretic 
-quantities associated with the exponential family.
-
-The following result has been previously pointed out by Amari \cite{Amari-KL}. 
-\begin{theorem}\label{lem:profit_decomposition}{\bf (Profit Decomposition)}: Consider an exponential 
-family $\family$ of distributions over some set of statistics $\suff(x)$, with natural parameters $\beta$. Let $\pi, \rho \in \family$ be any two probability distributions in the family. We use 
-$\beta_\pi$ to denote the natural parameters of $\pi$ and $\mu_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
-we can define $\beta_\rho$ and  $\mu_\rho$. Let $\lpf(\beta_{\rho})$ indicate the log partition function of $\rho$.
-Let $H(\pi)$ denote the entropy of the distribution $\pi$, and 
- $K(\pi||\rho)$ denote the KL-divergence of $\rho$ relative to $\pi$.
-Then, the following equality holds:  
-\begin{equation}
- K(\pi||\rho) + H(\pi) = \lpf(\beta_{\rho}) - \beta_\rho \cdot \mu_\pi
-\label{eq:profit_decomposition}
-\end{equation}
-\end{theorem}
-\begin{proof}
-\begin{eqnarray*}
- K(\pi||\rho) + H(\pi) &=& \int_x \pi(x)\log{\frac{\pi(x)}{\rho(x)}}dx\\
- &&\quad - \int_x \pi(x)\log{\pi(x)}dx \\
-&=& - \int_x \pi(x)\log{\rho(x)}dx \\
-&=& - \int_x \pi(x)\left[ \beta_\rho \cdot \suff(x) - \lpf(\beta_{\rho})\right]dx\\
-&=& - \beta_\rho \cdot \int_x \pi(x) \suff(x) dx\\
-&&\quad + \lpf(\beta_{\rho}) \int_x \pi(x) dx \\
-&=&  \lpf(\beta_{\rho}) - \beta_\rho \cdot \mu_\pi
-\end{eqnarray*}
-\end{proof}
-
-Equation~\ref{eq:profit_decomposition} gives us an alternative view of the
-market scoring rule construction. Assume that $\pi$ is the true distribution, and consider two arbitrary
-distributions $\rho_1, \rho_2 \in \family$.
-Note that $\lpf(\rho)$  is independent of $\pi$, and $\betavec_\rho \cdot \muvec_\pi$ is linear in the probabilities $\pi(x)$.
-If we want to measure loss by the KL-divergence, we can do so (in expectation) by setting a cost function that captures the first term, and defining security quantities ($\betavec$) and payoffs ($\suff$) to capture the second term.
-In particular, in a market with cost function $\lpf$, if the
-market price initially implies a distribution $\rho_1$, and a trader moves
-the market to price than implies a distribution $\rho_2$, then the cost she
-incurs is $\lpf(\rho_2) - \lpf(\rho_1)$.
-The number of securities bought to make this trade is given by the vector ($\betavec_{\rho_2} - \betavec_{\rho_1}$), and the expected payoff of the securities are given by $\muvec_\pi$.
-Thus, by equation \ref{eq:profit_decomposition}, the {\em net profit} of the trader is equal to $K(\pi||\rho_1) - K(\pi||\rho_2)$, {\it i.e.}, the reduction in KL-divergence with respect to the true distribution.
-We note one useful property of this construction: For a fixed vector $\betavec$ of purchased securities, the cost is independent of the outcome
-(and outcome distribution $\pi$), while the payoff is independent of the initial market state in which these securities were purchased.
-
-\subsection{Properties of the Maximum Entropy Market}
-
-{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
-\proof
-Note that, for an exponential family distribution
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
-&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
-\end{eqnarray*}
-If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
-\unproof
-
-%{\cor The exponential family market maker has unbounded worst case loss.}
-%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
-%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
-%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
-\paragraph{Loss of the market maker}
-First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
-
-Recall the definition of the conjugate dual:
-$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
-The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
-$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
-This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
-So we have 
-$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
-Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
-
-Let $q_{0}$ be the initial and $q_{f}$ the final market state.
-Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
-\begin{eqnarray*}
-\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
-&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
-&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
-&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
-&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
-&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
-&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
-&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
-&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
-\end{eqnarray*}
-
-\subsection{Example: Gaussian Markets}
-{\ex We will now derive an expression for the dual of the log partition function for the Gaussian distribution.
-In this case,  $\beta$ is a $2$-dimensional  vector.
-Let $\beta=\begin{pmatrix}\beta_{1}\\ \beta_{2}\end{pmatrix}$ and $\mu=\begin{pmatrix}\mu_{1}\\ \mu_{2}\end{pmatrix}$.
-So
-\begin{eqnarray*}
-\pmb{\mu}&=&\nabla\psi(\pmb{\beta})\\
-&=&\nabla\left(-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})\right)\\
-&=&\begin{pmatrix}-\frac{\beta_{1}}{2\beta_{2}}\\ \frac{\beta_{1}^{2}}{4\beta_{2}^{2}}-\frac{1}{2\beta_{2}}\end{pmatrix}
-\end{eqnarray*}
-Thus, $\pmb{\beta}$ can be written in terms of $\pmb{\mu}$ as follows:
-$\beta_{1}=\frac{\mu_{1}}{\mu_{2}-\mu_{1}^{2}}$ and $\beta_{2}=\frac{1}{\mu_{1}^{2}-\mu_{2}}$. This leads to the following closed form expression for the dual of the log partition function:
-$$\psi^{*}(\mu)=-\frac{1}{2}-\frac{1}{2}\log(\mu_{2}-\mu_{1}^{2})$$}
-{\ex
-We will now derive the expression for the differential entropy for the normal distribution.
-Recall that for the normal distribution $$p(x) = \frac{1}{\sqrt{2\pi\sigma^{2}}}\exp\left\{\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right\}$$
-\begin{eqnarray*}
-H(p) &=& -\int_{x}p(x)\log p(x)dx\\
-&=& -\int_{x}p(x)\left[-\log(2\pi)-\log\sigma+\frac{-(x-\mu)^{2}}{2\sigma^{2}}\right]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\int_{x}p(x)(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\E_{p(x)}[(x-\mu)^{2}]dx\\
-&=& \log(2\pi)+\log\sigma+\frac{1}{2\sigma^{2}}\sigma^{2}dx\\
-&=& \log(2\pi)+\frac{1}{2}+\log\sigma 
-\end{eqnarray*}
-Recall that the value of the dual of the log partition function is the value of the negative entropy of the distribution for an exponential family distribution.
-Thus when the variance of the normal distribution $\sigma^{2} = 0$, the value of the dual of the cost function goes to $\infty$ and is thus unbounded.}
-
-{\ex For the normal distribution, the natural parameters $\pmb{\beta}=(\beta_{1},\beta_{2})$ can be written in terms of $\mu$ and $\sigma$ as $$\beta_{1}=\frac{\mu}{\sigma^{2}},\quad \beta_{2}=\frac{-1}{2\sigma^{2}}$$ and the log partition function is  $$\psi(\pmb{\beta})=-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$$
-Alternately, the log partition function can be written in terms of its variance and mean as $\frac{\mu^{2}}{2\sigma^{2}}+\log\sigma$.
-Also note that the mean parameters are $\mu$ and $\mu^{2}+\sigma^{2}$.}
-Let's now work out the loss for the Gaussian market maker.
-First note that for a Gaussian market for mean parameters $\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}$ and $\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}$, we have
-\begin{eqnarray*}
-D_{C^{*}}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix},\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)
-&=& C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)-C^{*}\left(\begin{pmatrix}\mu_{1}\\ \mu_{1}^{2}+\sigma_{1}^{2}\end{pmatrix}\right)\\
-&&-\nabla C^{*} \left(\begin{pmatrix}\mu_{2}\\ \mu_{2}^{2}+\sigma_{2}^{2}\end{pmatrix}\right)\begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\begin{pmatrix}\frac{\mu_{2}}{\sigma_{2}^{2}}\\ -\frac{1}{2\sigma_{2}^{2}}\end{pmatrix}\cdot \begin{pmatrix}\mu_{1}-\mu_{2}\\ \mu_{1}^{2}+\sigma_{1}^{2}-\mu_{2}^{2}-\sigma_{2}^{2}\end{pmatrix}\\
-&=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
-%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
-&=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
-\end{eqnarray*}
-
-So the Gaussian market maker loss is 
-%parallels to quadratic market maker loss? 
-\begin{eqnarray*}
-D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]&=& \log \frac{\sigma_{0}}{\sigma_{x}}+\frac{\sigma_{0}^{2}-\sigma_{x}^{2}}{2\sigma_{0}^{2}}+\frac{(\mu_{x}-\mu_{0})^{2}}{2\sigma_{0}^{2}}\\
-&&-\left[\log \frac{\sigma_{f}}{\sigma_{x}}+\frac{\sigma_{f}^{2}-\sigma_{x}^{2}}{2\sigma_{f}^{2}}+\frac{(\mu_{x}-\mu_{f})^{2}}{2\sigma_{f}^{2}}\right]\\
-&=& \log \frac{\sigma_{0}}{ \sigma_{f}}+\frac{(x-\mu_{0})^{2}}{2\sigma_{0}^{2}}-\frac{(x-\mu_{f})^{2}}{2\sigma_{f}^{2}}\\
-\end{eqnarray*}¥
-Here we have used the fact that $\mu_{x}=x$ and $\sigma_{x}=0$.
-\rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
-%\pagebreak
-
-\subsection{Proper Scoring from Consistency}
-
-Our starting point for designing proper scoring rules is the classic logarithmic scoring rule for eliciting probabilities in the case of finite outcomes.
-This rule is simply $S(p,x) = \log p(x)$ and there are several ways to understand why it is proper, with different implications for generalization.
-
-First observe that the scoring rule compensates the agent with the log likelihood it assigned to the outcome, so the agent is maximizing the expected log likelihood via its report.
-Now it is a fundamental result in statistics that, under a variety of different sufficient conditions, the maximum likelihood estimator is statistically consistent; see~\cite{} for an overview of such conditions.
-This means that in the limit (as the size of the i.i.d.\ empirical sample increases) the log likelihood is maximized by the true parameters.
-By this reasoning, the following result follows immediately from the definition of consistency and the law of large numbers.
-%
-\begin{theorem} \label{max-like}
-Suppose that $\mM$ parametrizes $\mD$. The logarithmic scoring rule defined by
-%
-\begin{equation} \label{log-score}
-S(\mu, x) = \log p(x;\mu),
-\end{equation}
-%
-where $\mu \in \mM$, is proper over $\mD$ if and only if the maximum likelihood estimator for $\mu$ is consistent.
-\end{theorem}
-%
-We note that although our focus is on parametrizations by expectations, the preceding theorem in fact holds for more generic parameter spaces.
-The range of scoring rules it provides is broader than those characterized by~\cite{}, because it allows for a restriction over the set of beliefs.
-On the other hand, its applicability is limited to those domains $\mD$ that can be parametrized by the relevant statistics.
-The following provide concrete examples of these points.
-%
-\begin{example}
-Suppose the outcomes are supported on $[1,+\infty)$ and follow a Pareto distribution with density $f(x;\alpha) = \alpha/x^{\alpha+1}$ parametrized by an index $\alpha > 0$.
-The mean $\mu$ is related to the index via the one-to-one mapping $\mu = \frac{\alpha}{\alpha-1}$, so the density can alternatively be parametrized by the mean.
-Theorem~\ref{max-like} gives the following proper scoring rule for the mean of a Pareto distribution with support on $[1,+\infty)$:
-%
-\begin{equation} \label{pareto-scoring}
-S(\mu,x) = \log \frac{\mu}{\mu-1} - \left(\frac{\mu}{\mu-1} + 1\right) \log x.
-\end{equation}
-%
-We stress that the rule can only elicit the mean assuming the agent knows its belief is a Pareto distribution over $[1,+\infty]$.
-It does not elicit the mean of other families of densities parametrized by the mean (e.g., the exponential distribution).
-\end{example}
-%
-%
-\begin{example}
-Suppose the outcomes are supported on $(-\infty,+\infty)$ and follow a Cauchy distribution with density $f(x;m) = 1/\pi[1+(x-m)^2]$, parametrized by the median $m$.
-Theorem~\ref{max-like} leads to the following proper scoring rule for the median:
-%
-\begin{equation} \label{cauchy-scoring}
-S(m,x) = -\log[1+(x-m)^2].
-\end{equation}
-%
-Note that the median cannot be obtained as the expectation of any statistic. This highlights the range of parameters than may be elicited by the log score in different circumstances.
-\end{example}
-%

### 2014-02-11 WriteUp/maxentpredmkts.tex
edited section 6

@@
 
-\section{The Exponential Family Market Mechanism with Budgets} \label{sec:budgets}
-\sindhu{\cite{Azoury01}}
-In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
-
-In this section, we consider the the prediction market setup with traders that may be either informative or malicious.
-The malicious traders may want to inject faulty information into the market.
-The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.
-
-We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market.
-Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
-This idea of a dual analysis has been used in a recommender system context to compute reputations in \cite{ressami07}.
-
-In this section, we assume that the traders have exponential family beliefs.
-The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
-\subsection{Budget-limited Aggregation}
-Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
-We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market.
-This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader.
-Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.
-
-In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
+\section{Budget-limited Aggregation} \label{sec:budgets}
+
+In this section, we consider the evolution of the market state when traders are budget-limited and their budgets are known to the market maker.
+We consider a cost-function based market with cost function $C$ and payoff $\phi(x)$ with non-negative payoff.
+\sindhu{this means no short selling}We assume that the traders trade in multiple instances of the market.
+As before, the market price is interpreted as a probability density over the outcome space by interpreting the share vector as the natural parameter of an exponential family distribution \sindhu{I think we need this for Lemma \ref{lem:damage}}.
+We measure the error in prediction using the standard log loss.
+
+We show that traders with faulty information can only induce a limited amount of additional loss to the market's prediction.
+Further, since informative traders experience an expected increase in budget,  they will eventually be unconstrained and allowed to carry out  unrestricted trades.
+Taken together, this means that while the market suffers limited damage from ill-informed traders, it is also able to make use of all the information from informative traders in the long run.
+\paragraph{Budget-limited trades}
 Let $\alpha$ be the budget of a trader in the market.
 Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
-Let $C$ be the cost function.
-Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
+Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is, the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
 In this case, we want to budget-limit the trader's influence on the market state.
 
-We define the budget-limited final market state as $\theta'$. 
-Here, we consider a specific functional form of $\theta'$: $$\theta'=\lambda\htheta+(1-\lambda)\theta$$  where $$\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$$
-
-First, we  show that this trade is feasible given the trader's budget:
-\begin{theorem}
+\begin{definition}
 Let the current market state be given by $\theta$.
 Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
-The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$.
-\end{theorem}
+The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$ and is called his \emph{budget-limited trade}.
+\end{definition}
+
+\paragraph{Limited Damage}
+We will now quantify the error in prediction that the market maker might have to endure as a result of ill-informed entities entering the market.
+We assume that these entities trade in multiple instances of the market; thus the exposure of the market maker is over several \emph{rounds}.
+We use the log loss to measure this error in terms of the initial budget of traders; the loss function for $\theta$ shares held is defined as: $L(\theta,x)=-\log(p(x;\theta))=C(\theta)-\theta^{T}\phi(x)$.
+
+\begin{lemma}\label{lem:damage}
+The loss induced on the market by an uninformative trader is bounded by his initial budget.
+\end{lemma}
 \begin{proof}
-From the convexity of $C$, we have
-\begin{eqnarray*}
-C(\theta')&\leq&(1-\lambda)C(\theta)+\lambda C(\htheta)\\
-\mbox{Thus, } C(\theta') - C(\theta) &\leq& \lambda\ (C(\htheta) - C(\theta))\\
-\mbox{And, } C(\theta') - C(\theta) &\leq& \alpha 
- \end{eqnarray*}\end{proof}
-
- We note that moving to $\theta'$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
-In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\theta$.
-Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\htheta$.
-\subsection{Damage Bound}
-In this section, we will quantify the error in prediction that the market maker might have to endure as a result of malicious entities entering the market.
-We assume that these malicious entities trade in multiple instances of the market; thus the exposure of the market maker is over several \emph{rounds}.
-We use the standard log loss to measure this error in terms of the initial budget of traders.
-
-We define the loss function for $\theta$ shares held:
-$$L(\theta,x)=-\log(P_{\theta}(x))=\log\int \exp\{\theta^{T}\phi(x)\}\ dx-\theta^{T}\phi(x)=C(\theta)-\theta^{T}\phi(x)$$
-%with the correspondence between $\beta$ and $\theta$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
-
-Suppose the prediction market runs over multiple rounds $t$.
-Let $\theta_{0}^{t}$ be the initial number of shares of each security that are held.
-Let $\hat{\theta}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
-Let us assume that at this point the outcome is revealed; that is, we receive the value of the random variable $x^{t}$.
-
-Over multiple instances of the prediction market, we can track the change in budget of each trader.
-Let the budget at rounds $t$ and $t-1$ be $\alpha^{t}$ and $\alpha^{t-1}$ respectively.
-The change in budget for the trader is
-\begin{eqnarray*}
-\alpha^{t}-\alpha^{t-1}&=&C(\theta^{t})-C(\theta'^{t})-(\theta^{t}-\theta'^{t})^{T}\phi(x^{t})\\
-&=&L(\theta,x^{t})-L(\theta',x^{t})
-\end{eqnarray*}
-
-Define the myopic impact of a trader $i$ in segment $t$ as
-$$\Delta_{i}^{t}:=L(\hat{\theta}_{i-1}^{t},x^{t})-L(\hat{\theta}_{i}^{t},x^{t})$$
-Thus, the myopic impact captures incremental gain in prediction due to the trader in a round.
-Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
-
-The total myopic impact due to all $k$ active traders is given by
-$$\Delta^{t}=L(\theta_{0}^{t},x^{t})-L(\hat{\theta}_{k}^{t},x^{t})$$
-Thus $-\Delta^{t}$ captures the incremental loss of the market prediction after aggregation of all $k$ traders.
-
-\begin{theorem}
-A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
-\end{theorem}
-\begin{proof}
-Consider the myopic impact of a single trader $i$ after participating in the market $T$ times.
+First consider the change in budget of a trader $i$ over multiple rounds of the prediction market.
+Let his budget at rounds $t$ and $t-1$ be $\alpha_{i}^{t}$ and $\alpha_{i}^{t-1}$ respectively.
+The change in budget for trader $i$ moving the market state from $\theta$ to $\theta'$ with outcome $x^{t}$ is
+\begin{eqnarray*}
+\alpha_{i}^{t}-\alpha_{i}^{t-1}&=&C(\theta)-C(\theta')-(\theta-\theta')^{T}\phi(x^{t})\\
+&=&L(\theta,x^{t})-L(\theta',x^{t})=\Delta_{i}^{t}
+\end{eqnarray*}
+
+Where $\Delta_{i}^{t}$ is called the myopic impact of a trader $i$ in round $t$.
+Thus, the myopic impact captures incremental gain in prediction due to the trader in a round and is equal to the change in his budget in that round.
+
 Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
 $$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
-
-Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
 \end{proof}
-
-This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders. %imp
-
-In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation. 
-%We provide an information-theoretic justification for this claim in Section \ref{sec:infth}. 
-The intuition behind this  is that a trader's prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
-
-\subsection{Budget of Informative Traders}\label{sec:inf}
-Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
-
-We now characterize the expected change in budget for an informative trader. The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
-\begin{theorem}\label{thm:growth}
-Let $\theta$ be the current market state in the exponential family prediction market.
+\paragraph{Budget of Informative Traders}
+%Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
+We now characterize the expected change in budget for an informative trader. %The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
+\begin{lemma}\label{thm:growth}
+Let $\theta$ be the current market state.
 Suppose that an  informative trader with belief distribution  parametrized by $\theta'$  moves the market state to the budget-limited state $\htheta=\lambda \theta' + (1-\lambda) \theta$.
@@
 Then, the expectation, over trader $i$'s belief distribution $\theta_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\hat{\theta}_{i-1}$.
-\end {theorem}
+\end {lemma}
 \begin{proof}
@@
 && \E_{x\sim P_{\htheta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
-&=& T(\theta)-\theta\E_{x\sim P_{\htheta}}[\phi(x)]-(T(\theta')-\theta' \E_{x\sim P_{\htheta}}[\phi(x)])\\
+%&=& T(\theta)-\theta\E_{x\sim P_{\htheta}}[\phi(x)]-(T(\theta')-\theta' \E_{x\sim P_{\htheta}}[\phi(x)])\\
 &=& T(\theta)-\theta\nabla T(\htheta)-(T(\theta')-\theta' \nabla T(\htheta))\\
-&=& T(\theta)-T(\htheta)-\nabla T(\htheta)(\theta-\htheta)-(T(\theta')-T(\htheta)- \nabla T(\htheta)(\theta'-\htheta))\\
+%&=& T(\theta)-T(\htheta)-\nabla T(\htheta)(\theta-\htheta)-(T(\theta')-T(\htheta)- \nabla T(\htheta)(\theta'-\htheta))\\
 &=& D_{T}(\theta,\htheta)-D_{T}(\theta',\htheta)\\
@@
 \end{eqnarray*}
-The second to last inequality holds since $D_{T}(\theta',\htheta)$ is convex in $\theta'$ and we have:
+where $D_{T}(\cdot,\cdot)$ is the Bregman divergence based on $T$. The second to last inequality holds since $D_{T}(\theta',\htheta)$ is convex in $\theta'$ and we have:
 \begin{eqnarray*}
@@
 
+We note one important aspect of Lemma~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
+This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure.
+If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
+
 For continuous distributions with a density, the probability that a trader with private information will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs.
@@
 
-Notice that Theorem~\ref{thm:growth} only required that the market state to which the trader moves, be representable as a convex combination of the current market state and his belief.
+Notice that Lemma~\ref{thm:growth} only required that the market state to which the trader moves, be representable as a convex combination of the current market state and his belief.
 This means that the result holds for exponential utility traders aiming to maximize their utility.
@@
 
-We note one important aspect of Theorem~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
-This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure.
-If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
-%\pagebreak
-
-%\subsection{Exponential Utility}
-%Now suppose the trader has exponential utility, given by $$U(w) = -\frac{1}{a} \exp(-aw)$$ where $a$ is the coefficient of risk aversion.
-Suppose the trader wishes to maximize his expected utility.
-Then the number of shares $\delta$ that he purchases when the current market state is $\theta$ is given by
-%\begin{eqnarray*}
-%&&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}U\left[ \delta\phi(x) - C(\delta + \theta) + C(\theta)\right]\\
-%&=&\arg\max_{\delta}\E_{\beta\sim b}\E_{x\sim \beta}-\frac{1}{a}\exp\left[ -a\delta\phi(x) +a C(\delta + \theta) -a C(\theta)\right]
-%\end{eqnarray*}¥
-%
-%\begin{eqnarray*}
-%\E_{\beta\sim b}\E_{x\sim \beta}\exp\left[-a\delta\phi(x)\right]&=&\int_{\mathcal{B}}\int_{\mathcal{X}}\exp[-a\delta\phi(x)]\, \exp\{\beta\cdot\phi(x)-\psi(\beta)\}\, dx\\
-%&&\qquad\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\, d\beta\\
-%&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
-%&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\int_{\mathcal{X}}\exp[(\beta-a\delta)\cdot\phi(x)-\psi(\beta-a\delta)]\, dx\\
-%&=&\int_{\mathcal{B}}\exp\{(n\nu+m\hmu)\cdot\beta-(n+m)\psi(\beta))-\Psi((n\nu+m\hmu, n+m))\}\\
-%&&\qquad\exp\{\psi(\beta-a\delta)-\psi(\beta)\}\,\, d\beta\\
-%\end{eqnarray*}¥
-%\pagebreak
+\sindhu{does a discussion of liquidity fit in here?}
+%Also note that if we consider a cost function with a liquidity parameter as in Section \ref{},
 
@@
 Thus the equilibrium state is a convex combination of trader beliefs and initial market state.
+
+\section{The Exponential Family Market Mechanism with Budgets} \label{sec:budgets}
+In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.
+
+In this section, we consider the the prediction market setup with traders that may be either informative or malicious.
+The malicious traders may want to inject faulty information into the market.
+The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.
+
+We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market.
+Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.
+This idea of a dual analysis has been used in a recommender system context to compute reputations in \cite{ressami07}.
+
+In this section, we assume that the traders have exponential family beliefs.
+The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
+\subsection{Budget-limited Aggregation}
+Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices.
+We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market.
+This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader.
+Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.
+
+In this section, we assume that the budget of each trader is known to the market maker, and that the market maker can directly limit the allowed trades based on a trader's budget. 
+Let $\alpha$ be the budget of a trader in the market.
+Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
+Let $C$ be the cost function.
+Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
+In this case, we want to budget-limit the trader's influence on the market state.
+
+We define the budget-limited final market state as $\theta'$. 
+Here, we consider a specific functional form of $\theta'$: $$\theta'=\lambda\htheta+(1-\lambda)\theta$$  where $$\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$$
+
+First, we  show that this trade is feasible given the trader's budget:
+\begin{theorem}
+Let the current market state be given by $\theta$.
+Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
+The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$.
+\end{theorem}
+\begin{proof}
+From the convexity of $C$, we have
+\begin{eqnarray*}
+C(\theta')&\leq&(1-\lambda)C(\theta)+\lambda C(\htheta)\\
+\mbox{Thus, } C(\theta') - C(\theta) &\leq& \lambda\ (C(\htheta) - C(\theta))\\
+\mbox{And, } C(\theta') - C(\theta) &\leq& \alpha 
+ \end{eqnarray*}\end{proof}
+
+ We note that moving to $\theta'$ as defined may not be the optimal trade for a rational trader maximizing her expected profit.
+In general, the inequality above is strict, and so a trader does not fully exhaust her budget by moving to $\theta$.
+Our results below will continue to hold in the case that strategic informative traders move to a position closer to their beliefs $\htheta$.
+\subsection{Damage Bound}
+In this section, we will quantify the error in prediction that the market maker might have to endure as a result of malicious entities entering the market.
+We assume that these malicious entities trade in multiple instances of the market; thus the exposure of the market maker is over several \emph{rounds}.
+We use the standard log loss to measure this error in terms of the initial budget of traders.
+
+We define the loss function for $\theta$ shares held:
+$$L(\theta,x)=-\log(P_{\theta}(x))=\log\int \exp\{\theta^{T}\phi(x)\}\ dx-\theta^{T}\phi(x)=C(\theta)-\theta^{T}\phi(x)$$
+%with the correspondence between $\beta$ and $\theta$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
+
+Suppose the prediction market runs over multiple rounds $t$.
+Let $\theta_{0}^{t}$ be the initial number of shares of each security that are held.
+Let $\hat{\theta}_{k}^{t}$ be the final values corresponding to the market state after the traders have made their reports.
+Let us assume that at this point the outcome is revealed; that is, we receive the value of the random variable $x^{t}$.
+
+Over multiple instances of the prediction market, we can track the change in budget of each trader.
+Let the budget at rounds $t$ and $t-1$ be $\alpha^{t}$ and $\alpha^{t-1}$ respectively.
+The change in budget for the trader is
+\begin{eqnarray*}
+\alpha^{t}-\alpha^{t-1}&=&C(\theta^{t})-C(\theta'^{t})-(\theta^{t}-\theta'^{t})^{T}\phi(x^{t})\\
+&=&L(\theta,x^{t})-L(\theta',x^{t})
+\end{eqnarray*}
+
+Define the myopic impact of a trader $i$ in segment $t$ as
+$$\Delta_{i}^{t}:=L(\hat{\theta}_{i-1}^{t},x^{t})-L(\hat{\theta}_{i}^{t},x^{t})$$
+Thus, the myopic impact captures incremental gain in prediction due to the trader in a round.
+Note that the myopic impact caused by trader $i$ at round $t$ is equal to the change in his budget in that round.
+
+The total myopic impact due to all $k$ active traders is given by
+$$\Delta^{t}=L(\theta_{0}^{t},x^{t})-L(\hat{\theta}_{k}^{t},x^{t})$$
+Thus $-\Delta^{t}$ captures the incremental loss of the market prediction after aggregation of all $k$ traders.
+
+\begin{theorem}
+A coalition of $b$ malicious traders can at most cause loss bounded by their initial budgets.
+\end{theorem}
+\begin{proof}
+Consider the myopic impact of a single trader $i$ after participating in the market $T$ times.
+Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
+$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
+
+Thus, any coalition of $b$ adversaries $\{1,\ldots,b\}$ can cause at most $\sum_{i=1}^{b}\alpha_{i}^{0}$ damage.
+\end{proof}
+
+This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders. %imp
+
+In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation. 
+%We provide an information-theoretic justification for this claim in Section \ref{sec:infth}. 
+The intuition behind this  is that a trader's prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
+
+\subsection{Budget of Informative Traders}\label{sec:inf}
+Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
+
+We now characterize the expected change in budget for an informative trader. The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
+\begin{theorem}\label{thm:growth}
+Let $\theta$ be the current market state in the exponential family prediction market.
+Suppose that an  informative trader with belief distribution  parametrized by $\theta'$  moves the market state to the budget-limited state $\htheta=\lambda \theta' + (1-\lambda) \theta$.
+Then, the expectation (over the trader's belief) of the trader's profit  is greater than zero whenever his budget is positive and his belief differs from the previous market position $\theta$.
+% Suppose that each informative trader gets a random sample of data, 
+%resulting in a sequence of trader beliefs $\theta_i$, and hence a sequence of budget-limited market positions
+%$\hat{\theta}_i$, before a final outcome $x$.
+Then, the expectation, over trader $i$'s belief distribution $\theta_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\hat{\theta}_{i-1}$.
+\end {theorem}
+\begin{proof}
+Let the cost function $C$ be equal to the log partition function $T$ of the belief distribution.
+The payoff is given by the sufficient statistics $\phi(x)$.
+Then, the trader's expected net payoff is given by
+\begin{eqnarray*}
+&& \E_{x\sim P_{\htheta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
+&=& T(\theta)-\theta\E_{x\sim P_{\htheta}}[\phi(x)]-(T(\theta')-\theta' \E_{x\sim P_{\htheta}}[\phi(x)])\\
+&=& T(\theta)-\theta\nabla T(\htheta)-(T(\theta')-\theta' \nabla T(\htheta))\\
+&=& T(\theta)-T(\htheta)-\nabla T(\htheta)(\theta-\htheta)-(T(\theta')-T(\htheta)- \nabla T(\htheta)(\theta'-\htheta))\\
+&=& D_{T}(\theta,\htheta)-D_{T}(\theta',\htheta)\\
+&\geq&\lambda D_{T}(\theta,\htheta)\geq0
+\end{eqnarray*}
+The second to last inequality holds since $D_{T}(\theta',\htheta)$ is convex in $\theta'$ and we have:
+\begin{eqnarray*}
+D_{T}(\theta',\htheta)&=&D_{T}\left(\lambda\htheta+(1-\lambda)\theta,\htheta\right)\\
+&\leq&\lambda D_{T}( \htheta ,\htheta)+ (1-\lambda) D_{T}( \theta ,\htheta)\\
+&=&(1-\lambda) D_{T}( \theta ,\htheta)
+%\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
+\end{eqnarray*}¥
+Thus, a trader who moves the market state can expect his profit to be positive and at least $\lambda D_{T}(\theta,\htheta)$.
+ \end{proof}
+
+For continuous distributions with a density, the probability that a trader with private information will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs.
+This result suggests that, eventually, every informative trader will have the ability to influence the market state in accordance with his beliefs, without being budget limited.
+
+Notice that Theorem~\ref{thm:growth} only required that the market state to which the trader moves, be representable as a convex combination of the current market state and his belief.
+This means that the result holds for exponential utility traders aiming to maximize their utility.
+In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{T}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
+
+We note one important aspect of Theorem~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
+This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure.
+If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.

### 2014-02-11 WriteUp/liquidity.tex
some edits to liquidity
fixed lb proof in liquidity write-up

@@
 \section{Cost Function with a Liquidity Parameter}
-We may introduce a liquidity parameter to the cost function as $C_{\lambda}(\theta)=\lambda C\left(\frac{\theta}{\lambda}\right)$.
+We may introduce a liquidity parameter to the cost function as $C_{\lambda}(\theta)=\frac{1}{\lambda} C(\lambda\theta)$.
 We would like to find an expression for a lower bound on expected payoff.
@@
 \E_{x\sim p(x;\beta)}[C_{\lambda}(\theta)-\theta\phi(x)]&=&C_{\lambda}(\theta)-\theta\nabla T(\beta)\\
-&=&\lambda T(\theta/\lambda)-\theta\nabla T(\beta)\\
-&=&\lambda [T(\theta/\lambda)-\theta/\lambda\nabla T(\beta)-T(\beta)+\beta\nabla T(\beta)]+\lambda T(\beta)-\theta\lambda\nabla T(\beta)\\
-&=&\lambda D_{T}(\theta/\lambda,\beta)+\lambda [T(\beta)-\theta\nabla T(\beta)]
+&=&\frac{1}{\lambda} T(\theta\lambda)-\theta\nabla T(\beta)\\
+&=&\frac{1}{\lambda} [T(\theta\lambda)-\theta\lambda\nabla T(\beta)-T(\beta)+\beta\nabla T(\beta)]+\frac{1}{\lambda} T(\beta)-\frac{\beta}{\lambda}\nabla T(\beta)\\
+&=&\frac{1}{\lambda} D_{T}(\theta\lambda,\beta)+\frac{1}{\lambda} [T(\beta)-\beta\nabla T(\beta)]
 \end{eqnarray*}¥ 
-where $D_{T}(x_{1},x_{2}) = T(x_{1})-T(x_{2})-\nabla T(x_{1}-x_{2})$ is the Bregman Divergence between $x_{1}$ and $x_{2}$ based on $T$.
+where $D_{T}(x_{1},x_{2}) = T(x_{1})-T(x_{2})-\nabla T(x_{2})(x_{1}-x_{2})$ is the Bregman Divergence between $x_{1}$ and $x_{2}$ based on $T$.
 
@@
 \begin{eqnarray*}
-\E_{x\sim p(x;\beta)}\left[C_{\lambda}(\theta)-\theta\phi(x)-(C_{\lambda}(\theta')-\theta'\phi(x))\right]=\lambda[D_{T}(\theta/\lambda,\beta)-D_{T}(\theta'/\lambda,\beta)]
+\E_{x\sim p(x;\beta)}\left[C_{\lambda}(\theta)-\theta\phi(x)-(C_{\lambda}(\theta')-\theta'\phi(x))\right]=\frac{1}{\lambda}[D_{T}(\theta\lambda,\beta)-D_{T}(\theta'\lambda,\beta)]
 \end{eqnarray*}¥
-If $\theta'=a\beta+(1-a)\theta$ for some $a\in[0,1]$ we can write
+We have $\theta'=\frac{\lambda}{\lambda+a}(\beta/\lambda)+\frac{a}{\lambda+a}\theta$ where $a$ is the utility parameter
 \begin{eqnarray*}
-D_{T}(\theta'/\lambda,\beta)&\leq& a D_{T}(\beta/\lambda,\beta)+(1-a) D_{T}(\theta/\lambda,\beta)\\
-D_{T}(\theta/\lambda,\beta)-D(\theta'/\lambda,\beta)&\geq& a D_{T}(\theta/\lambda,\beta)-a D_{T}(\beta/\lambda,\beta)
+D_{T}(\theta'\lambda,\beta)&\leq& \frac{\lambda}{\lambda+a} D_{T}(\beta,\beta)+\frac{a}{\lambda+a} D_{T}(\theta\lambda,\beta)\\
+D_{T}(\theta/\lambda,\beta)-D_{T}(\theta'/\lambda,\beta)&\geq& \frac{\lambda}{a} D_{T}(\theta'/\lambda,\beta)
 \end{eqnarray*}¥
+
+Thus, the expected payoff is at least $\frac{1}{a} D_{T}(\theta'/\lambda,\beta)$.
+
 %Note that 
@@
 \implies \delta&=&\frac{\beta\lambda-\theta}{1+a\lambda}\\
-\implies \theta+\delta&=&\lambda\frac{\beta+a\delta}{1+a\lambda}
+\implies \theta+\delta&=&\frac{1}{1+a\lambda}\beta\lambda+\frac{a\lambda}{1+a\lambda}\delta
 \end{eqnarray*}
+
+From the remarks below, we see that $\beta\lambda$ actually represents the agent's target share vector.
 
@@
 
-\newcommand{\ttheta}{\tilde{\theta}}
 As we're dealing with LMSR I write $T$ rather than $C$ now where $T$ is the log partition function.
@@
 where $\mu$ are the current market prices.
+\end{document}
 
-\end{document}
+\section{Cost Function with a Liquidity Parameter}
+We may introduce a liquidity parameter to the cost function as $C_{\lambda}(\theta)=\lambda C\left(\frac{\theta}{\lambda}\right)$.
+We would like to find an expression for a lower bound on expected payoff.
+We would also like to find an expression for final market state for exponential utility traders with exponential family beliefs.
+\subsection{lower bound on expected payoff}
+The prediction market is set up with cost function $C_{\lambda}(\theta)$  and payoff for a unit share vector as $\phi(x)$. To find a lower bound on expected payoff, we first note that
+\begin{eqnarray*}
+\E_{x\sim p(x;\beta)}[C_{\lambda}(\theta)-\theta\phi(x)]&=&C_{\lambda}(\theta)-\theta\nabla T(\beta)\\
+&=&\lambda T(\theta/\lambda)-\theta\nabla T(\beta)\\
+&=&\lambda [T(\theta/\lambda)-\theta/\lambda\nabla T(\beta)-T(\beta)+\beta\nabla T(\beta)]+\lambda T(\beta)-\theta\lambda\nabla T(\beta)\\
+&=&\lambda D_{T}(\theta/\lambda,\beta)+\lambda [T(\beta)-\theta\nabla T(\beta)]
+\end{eqnarray*}¥ 
+where $D_{T}(x_{1},x_{2}) = T(x_{1})-T(x_{2})-\nabla T(x_{1}-x_{2})$ is the Bregman Divergence between $x_{1}$ and $x_{2}$ based on $T$.
+
+Thus, the expected payoff  for a trader with exponential family belief parameter $\beta$ who moves the market state from $\theta$ to $\theta'$ can be written as
+\begin{eqnarray*}
+\E_{x\sim p(x;\beta)}\left[C_{\lambda}(\theta)-\theta\phi(x)-(C_{\lambda}(\theta')-\theta'\phi(x))\right]=\lambda[D_{T}(\theta/\lambda,\beta)-D_{T}(\theta'/\lambda,\beta)]
+\end{eqnarray*}¥
+If $\theta'=a\beta\lambda+(1-a)\theta$ for some $a\in[0,1]$ we can write
+\begin{eqnarray*}
+D_{T}(\theta'/\lambda,\beta)&\leq& a D_{T}(\beta\lambda/\lambda,\beta)+(1-a) D_{T}(\theta/\lambda,\beta)\\
+D_{T}(\theta/\lambda,\beta)-D_{T}(\theta'/\lambda,\beta)&\geq& a D_{T}(\theta/\lambda,\beta)-a D_{T}(\beta,\beta)
+\end{eqnarray*}¥

### 2014-02-11 WriteUp/maxentpredmkts.tex
minor edit
minor updates so sindhu can edit sec 6

@@
 %
-\begin{corollary}
+\begin{corollary}\jake{$\mu$ without a hat is never defined... (this is for sebastien)}
 Under the conditions of Theorem~\ref{thm:exp-util} and an inverse liquidity of $\lambda$, the agent's optimal trade $\delta$ moves the state vector to
@@
 \subsection{Repeated Trading and the Effective Belief}
-\jake{I need to clean up a few things in this section.}
+
 In previous sections we have analyzed trader behavior as if it is his first entry into the market.
-In this section we will quantify how prior exposure in the market affects a trader's choices.
-In particular, we will show that a trader who has previously purchased shares in a market will, on subsequent entry, behave as if this purchase has updated his private belief.
-This result implies that any financial trade made in a market is equivalent to changing the trader's effective belief.
+We now pose the question: how will a trader reason about a possible future investment when the trader holds an \emph{existing portfolio?} When the agent is exposed to the market given a prior investment this should clearly affect the agent's preference for additional risk.
+In the context of a trader possessing an exponential family belief and with exponential utility we show that we can analyze explicitly how an agent incorporates an existing portfolio.
+Indeed, the key conclusion we make in this section is that a trader holding a natural parameter vector $\hat \theta$ and holding a portfolio $\delta$ in the market will reason about future a future investment simply \emph{as though} his belief were $\hat \theta - a \delta$ and he had no prior investment.
 
@@
 We consider a cost-function based market with cost function $C$ and payoff $\phi(x)$ with non-negative payoff.
-\sindhu{this means no short selling}We assume that the traders trade in multiple instances of the market.
+We assume that the traders trade in multiple instances of the market.
 As before, the market price is interpreted as a probability density over the outcome space by interpreting the share vector as the natural parameter of an exponential family distribution \sindhu{I think we need this for Lemma \ref{lem:damage}}.
@@
 
-\sindhu{does a discussion of liquidity fit in here?}
-%Also note that if we consider a cost function with a liquidity parameter as in Section \ref{},
-
 \section{Conclusions}

### 2014-02-11 WriteUp/maxentpredmkts.tex
added title and a note

@@
 % Title portion
-\title{Maximum Entropy Prediction Markets}
+\title{Exponential Family Markets for Information Aggregation}
 \author{JACOB ABERNETHY
@@
 \begin{abstract}
-\sindhu{this needs to change once we finalize the sections}
+\jake{I will take care of this.}
 In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets.

### 2014-02-11 WriteUp/maxentpredmkts.tex
added liquidity to section 6; fixed theorem numbering

@@
 Note that we say a function $f(\vec{x})$ is at a \emph{local optimum} if changing any coordinate of $\vec{x}$ does not increase the value of $f$.
-\begin{thm}\label{thm:potl}
+\begin{theorem}\label{thm:potl}
 Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}\defeq (\delta_{1},\ldots,\delta_{i},\ldots,\delta_{n})$.
@@
 then $\vec{\delta}$ is a Nash equilibrium if and only if $\Phi(\vec{\delta})$ is at a local optimum.
-\end{thm}
+\end{theorem}
 In the exponential family market, the cost function $C$ is identical to the log partition function $T$ defined in (\ref{log-part}).
@@
 In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{T}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
-
-\section{Conclusions}
-\sindhu{Do we need this section?}
+When the cost function is adjusted with a liquidity parameter as $C_{l}(\theta)=(1/l) C(l\theta)$ as described in Section \ref{sec:exputil},  the trader receives an expected payoff of at least $\frac{1}{a} D_{T}(\theta'/l,\htheta)$.
+
+%\section{Conclusions}
+%\sindhu{Do we need this section?}
 %

### 2014-02-12 WriteUp/maxentpredmkts.tex
fixed a typo

@@
 In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{T}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
-When the cost function is adjusted with a liquidity parameter as $C_{l}(\theta)=(1/l) C(l\theta)$ as described in Section \ref{sec:exputil},  the trader receives an expected payoff of at least $\frac{1}{a} D_{T}(\theta'/l,\htheta)$.
+When the cost function is adjusted with a liquidity parameter as $C_{l}(\theta)=(1/l) C(l\theta)$ as described in Section \ref{sec:exputil},  the trader receives an expected payoff of at least $\frac{1}{a} D_{T}(l\theta',\htheta)$.
 

### 2014-02-12 WriteUp/maxentpredmkts.tex
added liquidity proof

@@
 \section{Budget-limited Aggregation} \label{sec:budgets}
-
 In this section, we consider the evolution of the market state when traders are budget-limited and their budgets are known to the market maker.
@@
 We assume that the traders trade in multiple instances of the market.
-As before, the market price is interpreted as a probability density over the outcome space by interpreting the share vector as the natural parameter of an exponential family distribution \sindhu{I think we need this for Lemma \ref{lem:damage}}.
+As before, the market price is interpreted as a probability density over the outcome space by interpreting the share vector as the natural parameter of an exponential family distribution.
 We measure the error in prediction using the standard log loss.
@@
 If we assume that the traders' belief formation is optimal, then this growth result will extend to the true distribution as well.
+
+\subsection{Cost Function with a Liquidity Parameter}
+We may introduce a liquidity parameter to the cost function as $C_{\lambda}(\theta)=\frac{1}{\lambda} C(\lambda\theta)$.
+We would like to find an expression for a lower bound on expected payoff.
+We would also like to find an expression for final market state for exponential utility traders with exponential family beliefs.
+\subsection{lower bound on expected payoff}
+The prediction market is set up with cost function $C_{\lambda}(\theta)$  and payoff for a unit share vector as $\phi(x)$. To find a lower bound on expected payoff, we first note that
+\begin{eqnarray*}
+\E_{x\sim p(x;\beta)}[C_{\lambda}(\theta)-\theta\phi(x)]&=&C_{\lambda}(\theta)-\theta\nabla T(\beta)\\
+&=&\frac{1}{\lambda} T(\theta\lambda)-\theta\nabla T(\beta)\\
+&=&\frac{1}{\lambda} [T(\theta\lambda)-\theta\lambda\nabla T(\beta)-T(\beta)+\beta\nabla T(\beta)]+\frac{1}{\lambda} T(\beta)-\frac{\beta}{\lambda}\nabla T(\beta)\\
+&=&\frac{1}{\lambda} D_{T}(\theta\lambda,\beta)+\frac{1}{\lambda} [T(\beta)-\beta\nabla T(\beta)]
+\end{eqnarray*}¥ 
+where $D_{T}(x_{1},x_{2}) = T(x_{1})-T(x_{2})-\nabla T(x_{2})(x_{1}-x_{2})$ is the Bregman Divergence between $x_{1}$ and $x_{2}$ based on $T$.
+
+Thus, the expected payoff  for a trader with exponential family belief parameter $\beta$ who moves the market state from $\theta$ to $\theta'$ can be written as
+\begin{eqnarray*}
+\E_{x\sim p(x;\beta)}\left[C_{\lambda}(\theta)-\theta\phi(x)-(C_{\lambda}(\theta')-\theta'\phi(x))\right]=\frac{1}{\lambda}[D_{T}(\theta\lambda,\beta)-D_{T}(\theta'\lambda,\beta)]
+\end{eqnarray*}¥
+We have $\theta'=\frac{\lambda}{\lambda+a}(\beta/\lambda)+\frac{a}{\lambda+a}\theta$ where $a$ is the utility parameter
+\begin{eqnarray*}
+D_{T}(\theta'\lambda,\beta)&\leq& \frac{\lambda}{\lambda+a} D_{T}(\beta,\beta)+\frac{a}{\lambda+a} D_{T}(\theta\lambda,\beta)\\
+D_{T}(\theta\lambda,\beta)-D_{T}(\theta'\lambda,\beta)&\geq& \frac{\lambda}{a} D_{T}(\theta'\lambda,\beta)
+\end{eqnarray*}¥
+
+Thus, the expected payoff is at least $\frac{1}{a} D_{T}(\theta'\lambda,\beta)$.

### 2014-02-12 WriteUp/maxentpredmkts.tex
minor edit
some comments

@@
 %
+\sindhu{sebastien/jake, should this be $\Exp_p(x;\theta')[\phi(x)]$?}
 In this way, by its choice of $\delta$, the agent reveals that its expected belief is $\mu = \grad C(\theta')$.
@@
 %
+\sindhu{Maybe this is nitpicking but $\E_{\theta\sim p(\theta;b_{0})}\E_{x\sim p(x;\theta)}[\phi(x)]$?}
 meaning that $\nu = n \nu / n$ is the posterior mean.
@@
 \subsection{Cost Function with a Liquidity Parameter}
-We may introduce a liquidity parameter to the cost function as $C_{\lambda}(\theta)=\frac{1}{\lambda} C(\lambda\theta)$.
-We would like to find an expression for a lower bound on expected payoff.
-We would also like to find an expression for final market state for exponential utility traders with exponential family beliefs.
+We may introduce a liquidity parameter to the cost function as $C_{\lambda}(\theta)=\frac{1}{\lambda} C(\lambda\theta)$. We would like to find an expression for a lower bound on expected payoff. 
 \subsection{lower bound on expected payoff}

### 2014-02-12 WriteUp/maxentpredmkts.tex
edits to sec 5

@@
 In the context of a trader possessing an exponential family belief and with exponential utility we show that we can analyze explicitly how an agent incorporates an existing portfolio.
-Indeed, the key conclusion we make in this section is that a trader holding a natural parameter vector $\hat \theta$ and holding a portfolio $\delta$ in the market will reason about future a future investment simply \emph{as though} his belief were $\hat \theta - a \delta$ and he had no prior investment.
+Indeed, the key conclusion we make in this section is that a trader will reason about a future investment simply \emph{as though} he had simply updated his belief and he had no prior investment.
 
@@
 Suppose an exponential utility agent has exponential family belief parametrized by natural parameter $\hat{\theta}$.
-Based on this belief, let $\delta_{1}^{*}$ be the optimal vector of shares the agent decides to trade on first entering the market.
+Based on this belief, let $\delta_1$ be the vector of shares the agent has purchased on first entry in the market.
 Thus, his belief distribution is given by the density $p(x;\htheta)=\exp(\la\hat{\theta}, \phi(x)\ra - T(\hat{\theta})),$ where $T(\hat{\theta}) =\int_{\mathcal{X}}\exp\{\la\hat{\theta}, \phi(x)\ra\}\ dx$ is the log partition function and $\phi(x)$ are the sufficient statistics of the outcome $x\in \mathcal{X}$.
 On a subsequent entry into this market with market state $\theta'$, his optimal purchase $\delta_{2}^{*}$ is given by the solution of
-%Here $\theta' = \theta+\delta_{1}^{*}+\delta'$ where $\delta'$ captures the movement of the share vector by other traders in the market.
-$$\arg\max_{\delta_{2}} \E_{x\sim p(x;\htheta)} U\left[ (\delta_{1}^{*}+\delta_{2} )\phi(x) - C(\delta_{1}^{*} + \theta) + C(\theta)- C(\delta_{2} + \theta') + C(\theta') \right]$$
-Then if $\Theta\defeq\hat{\theta}-a \delta_{1}^{*}  $ is the effective belief, the trader optimal purchase is given by $\delta_{2}^* = (\Theta - \theta') / (1+a)$, which moves the share vector to $\theta' + \delta_{2}^* = \frac{1}{1+a} \Theta + \frac{a}{1+a} \theta'$ which is a convex combination of the effective belief and the current market state.
+%Here $\theta' = \theta+\delta_1+\delta'$ where $\delta'$ captures the movement of the share vector by other traders in the market.
+$$\arg\max_{\delta_{2}} \E_{x\sim p(x;\htheta)} U\left[ (\delta_1+\delta_{2} )\phi(x) - C(\delta_1 + \theta) + C(\theta)- C(\delta_{2} + \theta') + C(\theta') \right]$$
+Then if $\Theta\defeq\hat{\theta}-a \delta_1  $ is the effective belief, the trader optimal purchase is given by $\delta_{2}^* = (\Theta - \theta') / (1+a)$, which moves the share vector to $\theta' + \delta_{2}^* = \frac{1}{1+a} \Theta + \frac{a}{1+a} \theta'$ which is a convex combination of the effective belief and the current market state.
 \begin{theorem}\label{thm:reptrades}
@@
 \subsection{Equilibrium Market State for Exponential Utility Agents}
-\jake{need to mention the connection to potential games in the intro, it's a nice little extra bit.}
+
 We have shown that every exponential-utility maximizing trader picks the share vector $\delta$ so that the eventual market state can be represented as a convex combination of the initial market state and the natural parameter of his (exponential family) belief distribution.
@@
 In the exponential family market, the cost function $C$ is identical to the log partition function $T$ defined in (\ref{log-part}).
-Let $\vec{\delta}$ be the vector of vectors of shares purchased by every trader in the market when the market has reached equilibrium, $\theta$ the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his utility parameter.
-\jake{I need to fix a small technical detail with the fact that we are using the log utilities to use the potential argument.}
-
-Define a potential function as  $$\Phi(\vec{\delta})\defeq  T\left(\theta+\sum_{i}\delta_{i}\right)+\sum_{i}\frac{1}{a_{i}}T(\htheta_{i}-a_{i}\delta_{i})$$
-
-Now the utility of trader $i$ is $U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i})$.
+Let $\vec{\delta}$ be the matrix of share vectors purchased by every trader in the market when the market has reached equilibrium, $\theta$ the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his utility parameter.
+% \jake{I need to fix a small technical detail with the fact that we are using the log utilities to use the potential argument.}
+
+Define a potential function as  $\Phi(\vec{\delta})\defeq  T\left(\theta+\sum_{i}\delta_{i}\right)+\sum_{i}\frac{1}{a_{i}}T(\htheta_{i}-a_{i}\delta_{i})$.
+
+Rather than work directly with the utilities of every trader, we will work with the log of their utility values\footnote{It is important to note that the potential function analysis still applies for any monotonically increasing transformation of the traders' utility functions}.
+Now the log-utility of trader $i$ is $U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\xtheta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i})$.
 Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $\Phi(\vec{\delta})$ for each $\delta_{i}$.
@@
 \frac{\theta+\sum_{i=1}^{n}\left(\frac{\htheta_{i}}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
-Thus, the equilibrium state is a convex combination of trader beliefs and initial market state.
+The equilibrium state is a convex combination of all beliefs and initial market state.
 

### 2014-02-12 WriteUp/maxentpredmkts.tex
fixed compile error

@@
 Rather than work directly with the utilities of every trader, we will work with the log of their utility values\footnote{It is important to note that the potential function analysis still applies for any monotonically increasing transformation of the traders' utility functions}.
-Now the log-utility of trader $i$ is $U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\xtheta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i})$.
+Now the log-utility of trader $i$ is $U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i})$.
 Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $\Phi(\vec{\delta})$ for each $\delta_{i}$.

### 2014-02-12 WriteUp/maxentpredmkts.tex
minor edits

@@
 \begin{itemize}
-	\item the \emph{payoff function} of the market corresponds to the \emph{sufficient statistics} of the probabiilty distribution;
-	\item the vector of \emph{oustanding shares} in the market corresponds to the \emph{natural parameter vector} of the distribution;
+	\item the \emph{payoff function} of the market corresponds to the \emph{sufficient statistics} of the probability distribution;
+	\item the vector of \emph{outstanding shares} in the market corresponds to the \emph{natural parameter vector} of the distribution;
 	\item the market prices correspond to \emph{mean parameters};
@@
 \end{itemize}
+\sindhu{I don't think we have the KL divergence result here -- is this the bregman divergence in the budget section?}
 

### 2014-02-12 WriteUp/maxentpredmkts.tex
still working on abstract

@@
 \begin{abstract}
-\jake{I will take care of this.}
-In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets.
+We consider the design of prediction market mechanisms known as automated market makers.
+We show that we can design these mechanisms via the mold of \emph{exponential family distributions}, a common distribution template used in statistics.
+We give a full development of this relationship and explore a range of benefits.
+We draw connections between the information aggregation of market prices and the belief aggregation of learning agents that relying on exponential family distributions.
+We develop a very natural analysis of the market behavior as well as the price equilibrium under the assumption that the traders elicit risk aversion according to exponential utility.
+We also consider similar phenomena in other scenarios, e.g.
+when the traders are budget limited.
+%
+% In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets.
 We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data.
@@
 
-\jake{This is a comment from jake}
-\sindhu{This is a comment from sindhu}
-\sebastien{This is a comment from sebastien}
 Prediction markets are aggregation mechanisms that allow market prices to be interpreted as predictive probabilities on an event.

### 2014-02-12 WriteUp/maxentpredmkts.tex
revised related work

@@
 \jake{I want to add more to this list. It's ok but doesn't totally reflect what's in the last few sections.}
-\subsection{Related Work}
-\jake{I'll add more here. Sebastien: do you have anything to put in here regarding the exponential utility stuff?}
-Prediction markets have been known to perform well in practice (see, for instance, \cite{AGT07,Wol04,pennock2001real}).
-However, a sound theory for interpreting trader behavior and market prices is an ongoing area of study \cite{wolfers2006interpreting}.
-
-Scoring rules are a measure of merit of a predictor \cite{gneiting2007strictly}.
-\cite{H03} adapted scoring rules to markets with sequential participation of multiple traders.
-A cost function based formulation of markets was first introduced by \cite{chen2007utility}.
-
-Designing prediction markets to handle a large outcome space is an active area of research.
-In \cite{Chen08}, the authors use a restricted betting language to design efficient markets for a combinatorial outcome space.
-This technique is generalized by \cite{Pennock11}.
-\cite{Gao09} consider extending various automated market makers to an infinite outcome space.
-For the logarithmic market scoring rule they show that unbounded market maker loss can result in this setting.
-\cite{acv13} and \cite{OthmanS11} specify frameworks under which they design cost function based markets that satisfy the desirable property of bounded market maker loss even in infinite outcome spaces.
-
-In this paper, we use exponential family distributions to construct our cost function.
-For a comprehensive introduction to the class of exponential family distributions see \cite{WainJordan08}.
-\cite{Azoury01} consider learning exponential family distributions in a traditional online learning model and provide worst case loss bounds relative to using an offline algorithm.
-\cite{storkey11} constructs a prediction market where each participating agent is a learning algorithm with varying utility functions.
-
-\sebastien{Need to talk about certainty equivalents somewhere}
-
-%Designing prediction markets to handle a large outcome space is an active area of research.
-In \cite{Chen08}, the authors use a restricted betting language to design efficient markets for a combinatorial outcome space.
-This technique is generalized by \cite{Pennock11}.
-\cite{Gao09} consider extending various automated market makers to an infinite outcome space.
-For the logarithmic market scoring rule they show that unbounded market maker loss can result in this setting.
-\cite{Abernethy11} and \cite{OthmanS11} specify frameworks under which they design cost function based markets that satisfy the desirable property of bounded market maker loss even in infinite outcome spaces.
-%
-%The connection between machine learning and prediction markets has been studied previously.
-\cite{Chen10} and \cite{Abernethy11} have previously explored the connection to learning algorithms to inform the design and understanding of prediction markets.
-In particular, \cite{Chen10}  consider the correspondence between prediction markets with market scoring rules and the Follow the Regularized Leader algorithm proposed by \cite{Kalai05} and thus provide insight into the aggregation mechanism of a prediction market.
-%
-%Independently of this work, Beygelzimer et al.~\cite{BLP12} have shown that, for a particular form of binary prediction markets and traders with log-utility, the long-run dynamics of trading activity and budgets over many prediction markets lead the markets to satisfy a bounded regret property with respect to the best single trader.
-Our results here, and the future work we have suggested, form a program to prove bounded regret properties of budget-limited prediction markets under much more general conditions.
-
-% In this paper, we propose a general version of the popular logarithmic scoring rule extended to continuous outcome spaces.
-Our version of the scoring rule uses parametric distributions to elicit traders' expectation on some function (called statistics) of the data.
-It turns out that the maximum entropy distribution given the statistics are  exponential family distributions.
-
-% The connection we establish between markets and exponential  We have dual goals in this paper.
-On the one hand, we highlight a structural similarity between prediction markets and exponential families.
-We see this as the syntax of our prediction market mechanism.
-On the other,
-
-% We also analyze the behavior of traders with prior exposure in the market.
-Our results formalize the intuitive notion that private belief and financial exposure are just two sides of the same coin; trades in a prediction market can be seen as updating privately held beliefs.
-
-%\section{Notation and Definitions}
-%convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesian updates, conjugate prior.
-
-%\pagebreak
+
+\subsection*{Related Work}
+
+The notion of an exponential family distribution is fundamental to this paper.
+For a comprehensive introductions to these distributions, see~\citep{Barndorff78,WainJordan08}.
+Exponential families are intimately tied to the notions of log loss and entropy, but can be generalized to other types of convex losses and information, as shown by~\citet{GrunwaldDawid}, who also make a connection to scoring rules.
+
+Scoring rules are a measure of prediction accuracy, and we are concerned here with scoring rules for statistic expectations, typically over infinite outcome spaces.
+Such rules have been characterized by~\citet{Savage71}; see also~\citep{frongillo2013elicitingmeans,lambert2008eliciting}.
+Our rules are of course special cases of this characterization, but it appears the range of elegant scoring rules that arise from exponential families has not been appreciated.
+Indeed,~\citet{gneiting2007strictly} observe that specific instances of scoring rules for continuous outcomes are lacking, and survey various possibilities.
+In an early work,~\citet{matheson1976scoring} designed scoring rules for real-valued outcomes from those for binary outcomes.
+
+In a seminal paper,~\citet{H03} showed how to form a prediction market based on a sequentially-shared scoring rule, and specifically proposed the logarithmic market scoring rule (LMSR) based on log loss for finite outcome spaces~\citep{H07}.
+The markets we introduce are direct generalizations of the LMSR to continuous outcomes, but take the form of cost-function based markets as introduced by~\citet{chen2007utility}.~\citet{Gao09} and~\citet{measurable13} also consider extending various market makers to infinite outcome spaces.
+
+Prediction markets have been known to perform well in practice~\citep{AGT07,Wol04,pennock2001real}.
+However, a sound theory for interpreting trader behavior and market prices is an ongoing area of study~\citep{wolfers2006interpreting}.
+At one extreme, agents are assumed myopic and risk-neutral, which implies that they move the market state to their belief~\citep{chen2010new}.
+At the other extreme, agents are strategic and the market fully incorporates all information~\citep{ostrovsky2012}.
+Strategic behavior in prediction markets is also addressed in~\citep{chen2010gaming,dimitrov08}.
+
+We are not aware of any works that consider risk-averse agents within cost-function based markets.
+However, risk aversion is a fundamental component of mathematical finance and portfolio optimization, and there are close connections between the notion of a cost function and that of a convex risk measure~\citep{follmerRisk,follmer2011entropic}.
+Indeed, they arise from the same axioms as noted by~\citet{othman2011liqu}.
+We see the potential to draw more on the mathematical finance literature to take into account risk aversion, as prediction markets are simply single-period financial markets~\citep[Part I]{follmerFinance}.
 

### 2014-02-12 WriteUp/maxentpredmkts.tex
addressed some comments

@@
 %\pagebreak
-\section{Maximum Entropy Prediction Markets via Exponential Families}
+\section{Exponential Family Markets}
 \label{sec:maxentmarkets}
@@
 %
-\sindhu{sebastien/jake, should this be $\Exp_p(x;\theta')[\phi(x)]$?}
 In this way, by its choice of $\delta$, the agent reveals that its expected belief is $\mu = \grad C(\theta')$.
@@
 %
-\sindhu{Maybe this is nitpicking but $\E_{\theta\sim p(\theta;b_{0})}\E_{x\sim p(x;\theta)}[\phi(x)]$?}
 meaning that $\nu = n \nu / n$ is the posterior mean.
@@
 \section{Risk-Averse Traders with Exponential Utility} \label{sec:exputil}
-\sindhu{Sebastien, do you think we could cite \cite{norstad99util} for instance?}
 
@@
 Here $a$ controls the level of risk aversion: the agent is more risk averse as $a$ increases, and as $a$ tends to 0 we approach linear utility (risk-neutrality).
-Specifically, $a$ is the Arrow-Pratt coefficient of absolute risk aversion, and exponential utilities of the form~(\ref{exp-util}) are the unique utilities that exhibit constant absolute risk aversion (CARA).
+Specifically, $a$ is the Arrow-Pratt coefficient of absolute risk aversion, and exponential utilities of the form~(\ref{exp-util}) are the unique utilities that exhibit constant absolute risk aversion~\citep[Chap.\ 11]{VarianMicro}.
 
@@
 Higher price responsiveness means fewer shares must be bought to make the market prices match the agent's expectation, so the natural parameter is scaled down accordingly.
-With a liquidity adjustment, the analysis of Theorem~\ref{thm:exp-util} can be extended and yields the following result.
-%
-\begin{corollary}\jake{$\mu$ without a hat is never defined... (this is for sebastien)}
+With a liquidity adjustment the analysis of Theorem~\ref{thm:exp-util} can be extended and yields the following result, where as before $\theta$ and $\mu$ are the market's outstanding shares and prices respectively.
+%
+\begin{corollary}
 Under the conditions of Theorem~\ref{thm:exp-util} and an inverse liquidity of $\lambda$, the agent's optimal trade $\delta$ moves the state vector to

### 2014-02-12 WriteUp/maxentpredmkts.tex
cleaned up 5.1 and 5.2

@@
 
-In previous sections we have analyzed trader behavior as if it is his first entry into the market.
-We now pose the question: how will a trader reason about a possible future investment when the trader holds an \emph{existing portfolio?} When the agent is exposed to the market given a prior investment this should clearly affect the agent's preference for additional risk.
-In the context of a trader possessing an exponential family belief and with exponential utility we show that we can analyze explicitly how an agent incorporates an existing portfolio.
-Indeed, the key conclusion we make in this section is that a trader will reason about a future investment simply \emph{as though} he had simply updated his belief and he had no prior investment.
+In previous sections we have analyzed trader behavior assuming it is his first entry into the market.
+We now pose the question: how will a trader reason about a possible future investment when the trader holds an existing portfolio?
+%When the agent is exposed to the market given a prior investment this should clearly affect the agent's preference for additional risk.
+In the context of a trader possessing an exponential family belief together with exponential utility, we show that we can explicitly analyze how an agent incorporates an existing portfolio.
+The key conclusion is that a trader will reason about a future investment simply as though he had updated his belief and had no prior investment.
 
@@
 Based on this belief, let $\delta_1$ be the vector of shares the agent has purchased on first entry in the market.
-Thus, his belief distribution is given by the density $p(x;\htheta)=\exp(\la\hat{\theta}, \phi(x)\ra - T(\hat{\theta})),$ where $T(\hat{\theta}) =\int_{\mathcal{X}}\exp\{\la\hat{\theta}, \phi(x)\ra\}\ dx$ is the log partition function and $\phi(x)$ are the sufficient statistics of the outcome $x\in \mathcal{X}$.
+%Thus, his belief distribution is given by the density $p(x;\htheta)=\exp(\la\hat{\theta}, \phi(x)\ra - T(\hat{\theta})),$ where $T(\hat{\theta}) =\int_{\mathcal{X}}\exp\{\la\hat{\theta}, \phi(x)\ra\}\ dx$ is the log partition function and $\phi(x)$ are the sufficient statistics of the outcome $x\in \mathcal{X}$.
 On a subsequent entry into this market with market state $\theta'$, his optimal purchase $\delta_{2}^{*}$ is given by the solution of
 %Here $\theta' = \theta+\delta_1+\delta'$ where $\delta'$ captures the movement of the share vector by other traders in the market.
-$$\arg\max_{\delta_{2}} \E_{x\sim p(x;\htheta)} U\left[ (\delta_1+\delta_{2} )\phi(x) - C(\delta_1 + \theta) + C(\theta)- C(\delta_{2} + \theta') + C(\theta') \right]$$
-Then if $\Theta\defeq\hat{\theta}-a \delta_1  $ is the effective belief, the trader optimal purchase is given by $\delta_{2}^* = (\Theta - \theta') / (1+a)$, which moves the share vector to $\theta' + \delta_{2}^* = \frac{1}{1+a} \Theta + \frac{a}{1+a} \theta'$ which is a convex combination of the effective belief and the current market state.
+%
+$$
+\arg\max_{\delta_{2}} \:\: \E_{x\sim p(x;\htheta)} U\left[ \la\delta_1+\delta_{2} ,\phi(x)\ra - C(\delta_1 + \theta) + C(\theta)- C(\delta_{2} + \theta') + C(\theta') \right].
+$$
+%
+Then if $\theta'' = \hat{\theta}-a \delta_1  $ is the effective belief, the trader's optimal purchase is given by $\delta_{2} = (\theta'' - \theta') / (1+a)$, moving the share vector to $\theta' + \delta_{2} = \frac{1}{1+a} \theta'' + \frac{a}{1+a} \theta'$, which is a convex combination of the effective belief and the current market state.
+%
 \begin{theorem}\label{thm:reptrades}
 Suppose an exponential utility maximizing trader with utility parameter $a$ who has belief $\hat{\theta}$ makes a purchase $\delta$ in a market.
-On subsequently re-entering this market, he will behave identically to an exponential utility maximizing trader with belief $\hat{\theta}-a \delta$ and no prior exposure in the market.
+On subsequently re-entering the market, he will behave identically to an exponential utility maximizing trader with belief $\hat{\theta}-a \delta$ and no prior exposure in the market.
 \end{theorem}
+%
 Theorem \ref{thm:reptrades} implies that financial exposure can be equivalently understood as changing the privately held beliefs.
@@
 
-We have shown that every exponential-utility maximizing trader picks the share vector $\delta$ so that the eventual market state can be represented as a convex combination of the initial market state and the natural parameter of his (exponential family) belief distribution.
-In this section we will compute the equilibrium state in an exponential family market with multiple exponential-utility maximizing traders.
-
-We now review a well-known result from game theory regarding the class of \emph{potential games} \cite{monderer1996potential}.
-Note that we say a function $f(\vec{x})$ is at a \emph{local optimum} if changing any coordinate of $\vec{x}$ does not increase the value of $f$.
-\begin{theorem}\label{thm:potl}
-Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}\defeq (\delta_{1},\ldots,\delta_{i},\ldots,\delta_{n})$.
+We have shown that every exponential-utility maximizing trader picks the share vector $\delta$ so that the eventual market state can be represented as a convex combination of the current market state and the natural parameter of his (exponential family) belief distribution.
+In this section we will compute the equilibrium state in an exponential family market with multiple such traders.
+
+We draw a well-known result from game theory regarding the class of \emph{potential games}.
+We say a function $f(\vec{x})$ is at a \emph{local optimum} if changing any coordinate of $\vec{x}$ does not increase the value of $f$.
+%
+\begin{theorem}[\citet{monderer1996potential}]\label{thm:potl}
+Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}= (\delta_{1},\ldots,\delta_{i},\ldots,\delta_{n})$.
 If there exists a potential function $\Phi(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=\Phi(\vec{\delta})-\Phi(\vec{\delta}_{-i},\delta_{i}')$$
@@
 \end{theorem}
+%
 In the exponential family market, the cost function $C$ is identical to the log partition function $T$ defined in (\ref{log-part}).
-Let $\vec{\delta}$ be the matrix of share vectors purchased by every trader in the market when the market has reached equilibrium, $\theta$ the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his utility parameter.
+Let $\vec{\delta}$ be the matrix of share vectors purchased by every trader in the market when the market has reached equilibrium.
+Let $\theta$ be the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his risk aversion parameter.
 % \jake{I need to fix a small technical detail with the fact that we are using the log utilities to use the potential argument.}
 
-Define a potential function as  $\Phi(\vec{\delta})\defeq  T\left(\theta+\sum_{i}\delta_{i}\right)+\sum_{i}\frac{1}{a_{i}}T(\htheta_{i}-a_{i}\delta_{i})$.
-
-Rather than work directly with the utilities of every trader, we will work with the log of their utility values\footnote{It is important to note that the potential function analysis still applies for any monotonically increasing transformation of the traders' utility functions}.
-Now the log-utility of trader $i$ is $U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i})$.
-Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $\Phi(\vec{\delta})$ for each $\delta_{i}$.
-\begin{eqnarray*}
-\nabla_{\delta_{i}}\Phi(\vec{\delta})&=&\nabla T\left(\theta+\sum_{j=1}^{n}\delta_{i}\right)-\nabla T(\htheta_{i}-a_{i}\delta_{i}) = 0
+Define a potential function as  $\Phi(\vec{\delta}) =  T\left(\theta+\sum_{i}\delta_{i}\right)+\sum_{i}\frac{1}{a_{i}}T(\htheta_{i}-a_{i}\delta_{i})$.
+
+Rather than working directly with the utilities of every trader, we will work with the log of their utility values.\footnote{It is important to note that the potential function analysis still applies for any monotonically increasing transformation of the traders' utility functions.} Now the log-utility of trader $i$ is $$U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i}).$$
+Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $\Phi(\vec{\delta})$ for each $\delta_{i}$:
+\begin{eqnarray*}
+\nabla_{\delta_{i}}\Phi(\vec{\delta})&=&\nabla T\left(\theta+\sum_{j=1}^{n}\delta_{i}\right)-\nabla T(\htheta_{i}-a_{i}\delta_{i}) = 0.
 \end{eqnarray*}¥
-And we have the following expression for the final market state.
+This leads to the following expression for the final market state.
 $$\theta+\sum_{j=1}^{n}\delta_{j}=
 \frac{\theta+\sum_{i=1}^{n}\left(\frac{\htheta_{i}}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
-The equilibrium state is a convex combination of all beliefs and initial market state.
+We see that the equilibrium state is a convex combination of all beliefs and the initial market state, where beliefs are weighted by the agents' risk tolerances (inverse risk aversion).
 

### 2014-02-12 WriteUp/maxentpredmkts.tex
changed the title!

@@
 % Page heads
-\markboth{Abernethy et al.}{Maximum Entropy Prediction Markets}
+\markboth{Abernethy et al.}{Information Aggregation in Exponential Family Markets}
 % Title portion
-\title{Exponential Family Markets for Information Aggregation}
+\title{Information Aggregation in Exponential Family Markets}
 \author{JACOB ABERNETHY
@@
 
-\acmformat{Jacob Abernethy, Sindhu Kutty, S\'{e}bastien Lahaie, and Rahul Sami, 2014. Maximum Entropy Prediction Markets}
+%\acmformat{Jacob Abernethy, Sindhu Kutty, S\'{e}bastien Lahaie, and Rahul Sami, 2014. Information Aggregation in Exponential Family Markets}
 %%Gang Zhou, Yafeng Wu, Ting Yan, Tian He, Chengdu Huang, John A. Stankovic, and Tarek F. Abdelzaher, 2010. A multifrequency MAC specially designed for  wireless sensor network applications.

### 2014-02-12 WriteUp/maxentpredmkts.tex
edits to intro and space optimizations

@@
 Each trader in the market is assumed to have some private information that he uses to make a prediction on the outcome of the event.
-Traders are allowed to report their beliefs on the outcome of the event by allowing them to buy and sell securities whose value depends on the outcome of this future event.
-This will effect the state of the market, thus updating the predictive probabilities for the event.
+Traders are allowed to report their beliefs by buying and selling securities whose ultimate payoff depends on the future outcome.
+This will affect the state of the market, thus updating the predictive probabilities for the event.
 Further, since the trades are done sequentially, the trader is allowed to observe all past trades in the market and update his private information based on this information.
-Traders can see the past history of trades, so the price at which a current trader is willing to buy and sell these securities can be interpreted as an aggregate ``consensus probability forecast'' of a particular event occurring.
+In some sense the market prices, which are in effect the prices at which the marginal trader is willing to buy or sell the available securities, can be interpreted as an aggregate ``consensus probability forecast'' of the event in question.
 % For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins.
@@
 
-\jake{I want to use this paragraph for a general overview of mm frameworks} One popular form of prediction markets is the market scoring rule~\cite{H03}.
+% \jake{I want to use this paragraph for a general overview of mm frameworks} One popular form of prediction markets is the market scoring rule~\cite{H03}.
 A market scoring rule considers all trades as a single chronological sequence.
@@
 
-Much of the work on prediction market frameworks has focused primarily on structural properties of the mechanism: incentive compatibility, the market maker loss, the available liquidity, the fluctuations of the prices as a function of the trading volume, to name a few.
+We give an brief overview of the work on prediction market design below.
+Much of this work has focused heavily on structural properties of the mechanism: incentive compatibility, the market maker loss, the available liquidity, the fluctuations of the prices as a function of the trading volume, to name a few.
 Absent from much of the literature is a corresponding \emph{semantics} of the market behavior or the observed prices.
@@
 
-In the present work we show that a number of classical statistical tools can be leveraged to design a prediction market framework in the mold of an \emph{exponential family distributions} that possesses a number of attractive properties.
-Common concepts in statistics, e.g.
-\emph{entropy maximization}, \emph{log loss}, and \emph{bayesian inference}, relate to natural aspects of our class of mechanisms.
+In the present work we show that a number of classical statistical tools can be leveraged to design a prediction market framework in the mold of \emph{exponential family distributions}; we show that this statistical framework leads to a number of attractive properties and interpretations.
+Common concepts in statistics -- e.g.
+\emph{entropy maximization}, \emph{log loss}, and \emph{bayesian inference} -- relate to natural aspects of our class of mechanisms.
 In particular, the central objects in our market framework can be interpreted via conceptual objects used to define exponential families:
@@
 \end{itemize}
-\sindhu{I don't think we have the KL divergence result here -- is this the bregman divergence in the budget section?}
+% \sindhu{I don't think we have the KL divergence result here -- is this the bregman divergence in the budget section?}
 
@@
 We turn our attention to market design in Section~\ref{sec:maxentmarkets} and give a full description of our proposed mechanisms.
-In addition to showing the syntactic relationship between exponential families and prediction markets, we also explore the semantic implications as well.
+In addition to showing the syntactic relationship between exponential families and prediction markets, we explore a number of rich semantic implications as well.
 In particular, we show that our formulation allows us to analyze the evolution of the market under various models of trader behavior:
@@
 The maximum entropy distribution over the sphere with such sufficient statistics is the von Mises-Fisher distribution.
-The log partition function corresponds to the cost function
+The log partition function corresponds to
 %
@@
 & = &  T(\htheta+\delta) - T(\htheta) + \log \int_{\mX} \exp( \la\delta+\htheta,\phi(x)\ra - T(\htheta+\delta) )\, d\nu(x)\\
-& = &  T(\htheta+\delta) - T(\htheta) + \log \int_{\mX} p(x;\htheta+\delta) \, d\nu(x)\\
-& = &  T(\htheta+\delta) - T(\htheta),
+& = &  T(\htheta+\delta) - T(\htheta) + \log \int_{\mX} p(x;\htheta+\delta) \, d\nu(x) =   T(\htheta+\delta) - T(\htheta),
 \end{eqnarray*}
@@
 $$
-This objective is strictly concave, from the strict convexity of $T$. The optimum is therefore characterized by the first-order conditions:
-$$
-\grad T(\htheta - a\delta) = \grad T(\theta+\delta).
-$$
+This objective is strictly concave, from the strict convexity of $T$.
+The optimum is therefore characterized by the first-order condition $\grad T(\htheta - a\delta) = \grad T(\theta+\delta)$.
 As the gradient map $\grad T$ is one-to-one, this is solved by equating the arguments, which leads to $\delta = (\htheta - \theta) / (1+a)$ and~(\ref{exp-update}).
@@
 We assume that the traders trade in multiple instances of the market.
-As before, the market price is interpreted as a probability density over the outcome space by interpreting the share vector as the natural parameter of an exponential family distribution.
+As before, the market price is interpreted as a probability density over the outcome space by viewing the share vector as the natural parameters of an exponential family distribution.
 We measure the error in prediction using the standard log loss.
@@
 Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
-Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is, the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
+Now suppose that $\alpha<C(\htheta)-C(\theta)$; that is, the trader's budget precludes purchasing enough shares to move the market state to his belief.
 In this case, we want to budget-limit the trader's influence on the market state.
@@
 \begin{eqnarray*}
-&& \E_{x\sim P_{\htheta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]\\
-%&=& T(\theta)-\theta\E_{x\sim P_{\htheta}}[\phi(x)]-(T(\theta')-\theta' \E_{x\sim P_{\htheta}}[\phi(x)])\\
+\E_{x\sim P_{\htheta}}[C(\theta)-C(\theta')-(\theta-\theta')\phi(x)]
 &=& T(\theta)-\theta\nabla T(\htheta)-(T(\theta')-\theta' \nabla T(\htheta))\\
 %&=& T(\theta)-T(\htheta)-\nabla T(\htheta)(\theta-\htheta)-(T(\theta')-T(\htheta)- \nabla T(\htheta)(\theta'-\htheta))\\
-&=& D_{T}(\theta,\htheta)-D_{T}(\theta',\htheta)\\
-&\geq&\lambda D_{T}(\theta,\htheta)\geq0
+&=& D_{T}(\theta,\htheta)-D_{T}(\theta',\htheta) \geq \lambda D_{T}(\theta,\htheta)\geq0
 \end{eqnarray*}
@@
 D_{T}(\theta',\htheta)&=&D_{T}\left(\lambda\htheta+(1-\lambda)\theta,\htheta\right)\\
-&\leq&\lambda D_{T}( \htheta ,\htheta)+ (1-\lambda) D_{T}( \theta ,\htheta)\\
-&=&(1-\lambda) D_{T}( \theta ,\htheta)
+&\leq&\lambda D_{T}( \htheta ,\htheta)+ (1-\lambda) D_{T}( \theta ,\htheta) 
+= (1-\lambda) D_{T}( \theta ,\htheta)
 %\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
 \end{eqnarray*}¥
-Thus, a trader who moves the market state can expect his profit to be positive and at least $\lambda D_{T}(\theta,\htheta)$.
+A trader who adjust the market state may expect positive profit $\geq \lambda D_{T}(\theta,\htheta)$.
  \end{proof}
@@
 
-For continuous distributions with a density, the probability that a trader with private information will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs.
+Given a continuous density the probability a trader will form exactly the same beliefs as the current market position is $0$, and thus, each trader will have positive expected profit on almost all sequences of observed samples and beliefs.
 This result suggests that, eventually, every informative trader will have the ability to influence the market state in accordance with his beliefs, without being budget limited.

### 2014-02-12 WriteUp/maxentpredmkts.tex
polished section 6

--- 2014-02-12 WriteUp/maxentpredmkts.tex
@@
 \section{Budget-limited Aggregation} \label{sec:budgets}
-In this section, we consider the evolution of the market state when traders are budget-limited and their budgets are known to the market maker.
-We consider a cost-function based market with cost function $C$ and payoff $\phi(x)$ with non-negative payoff.
+In this section, we consider the evolution of the market state when traders are budget-limited.
+% and their budgets are known to the market maker. We consider a cost-function based market with cost function $C$ and payoff $\phi(x)$ with non-negative payoff. 
 We assume that the traders trade in multiple instances of the market.
-As before, the market price is interpreted as a probability density over the outcome space by interpreting the share vector as the natural parameter of an exponential family distribution.
-We measure the error in prediction using the standard log loss.
-
-We show that traders with faulty information can only induce a limited amount of additional loss to the market's prediction.
+As before, the market price is interpreted as a probability density over the outcome space and the share vector as the natural parameter of an exponential family distribution.
+Consistent with the connections drawn in Section~\ref{sec:scoring} and throughout, we measure the error in prediction using the standard log loss.
+
+We show that traders with faulty information can only impose a limited amount of additional loss to the market's prediction.
 Further, since informative traders experience an expected increase in budget,  they will eventually be unconstrained and allowed to carry out  unrestricted trades.
@@
 Let $\alpha$ be the budget of a trader in the market.
-Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
+Suppose that with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
 Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is, the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
@@
 
-\begin{definition}
-Let the current market state be given by $\theta$.
-Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
+Let the current market state be given by $\theta$ and let the final market state be $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
 The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$ and is called his \emph{budget-limited trade}.
-\end{definition}
 
@@
 We will now quantify the error in prediction that the market maker might have to endure as a result of ill-informed entities entering the market.
-We assume that these entities trade in multiple instances of the market; thus the exposure of the market maker is over several \emph{rounds}.
-We use the log loss to measure this error in terms of the initial budget of traders; the loss function for $\theta$ shares held is defined as: $L(\theta,x)=-\log(p(x;\theta))=C(\theta)-\theta^{T}\phi(x)$.
+We assume that these entities trade in multiple instances of the market; thus the exposure of the market maker is over several rounds.
+The log loss function for $\theta$ shares held is defined as $L(\theta,x)=-\log p(x;\theta)=C(\theta)-\la \theta\phi(x)\ra$.
 
@@
 \end{proof}
+
+An interesting aspect of Lemma~\ref{lem:damage} is that the log loss can be quantified in the same units as the traders' budgets.
+
 \paragraph{Budget of Informative Traders}
@@
 Let $\theta$ be the current market state.
-Suppose that an  informative trader with belief distribution  parametrized by $\theta'$  moves the market state to the budget-limited state $\htheta=\lambda \theta' + (1-\lambda) \theta$.
-Then, the expectation (over the trader's belief) of the trader's profit  is greater than zero whenever his budget is positive and his belief differs from the previous market position $\theta$.
+Suppose that an  informative trader with belief distribution  parametrized by $\htheta$  moves the market state to the budget-limited state $\theta'=\lambda \htheta + (1-\lambda) \theta$.
+Then, the expectation (over the trader's belief) of the trader's profit  is strictly positive whenever his budget is positive and his belief differs from the previous market position $\theta$.
 % Suppose that each informative trader gets a random sample of data, 
@@
 
-We note one important aspect of Lemma~\ref{thm:growth}: The expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
+We note one important aspect of Lemma~\ref{thm:growth}: the expectation is taken with respect to each trader's belief at the time of trade, rather than with respect to the true distribution.
 This is needed because we have made no assumptions about the optimality of the traders' belief updating procedure.
@@
 
-Notice that Lemma~\ref{thm:growth} only required that the market state to which the trader moves, be representable as a convex combination of the current market state and his belief.
-This means that the result holds for exponential utility traders aiming to maximize their utility.
+Notice that Lemma~\ref{thm:growth} only required that the market state to which the trader moves be representable as a convex combination of the current market state and his belief.
+This means that the result holds for exponential utility traders aiming to maximize their utility by Theorem~\ref{thm:exp-util}.
 In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{T}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
-When the cost function is adjusted with a liquidity parameter as $C_{l}(\theta)=(1/l) C(l\theta)$ as described in Section \ref{sec:exputil},  the trader receives an expected payoff of at least $\frac{1}{a} D_{T}(l\theta',\htheta)$.
+When the cost function is adjusted to $C_{\lambda}$ with an inverse liquidity parameter $\lambda$ Section \ref{sec:exputil},  the trader receives an expected payoff of at least $\frac{1}{a} D_{T}(\lambda\theta,\htheta)$.
 
@@
 The cost function has the same form as the log partition function $T$ of the exponential family and the payoff is determined by the sufficient statistics of the data $\phi(x)$.
+
 \subsection{Budget-limited Aggregation}

### 2014-02-12 WriteUp/maxentpredmkts.tex
more edits and space opts

--- 2014-02-12 WriteUp/maxentpredmkts.tex
@@
 We consider the design of prediction market mechanisms known as automated market makers.
-We show that we can design these mechanisms via the mold of \emph{exponential family distributions}, a common distribution template used in statistics.
+We show that we can design these mechanisms via the mold of \emph{exponential family distributions}, a popular and well-studied probability distribution template used in statistics.
 We give a full development of this relationship and explore a range of benefits.
@@
 We develop a very natural analysis of the market behavior as well as the price equilibrium under the assumption that the traders elicit risk aversion according to exponential utility.
-We also consider similar phenomena in other scenarios, e.g.
-when the traders are budget limited.
+We also consider similar phenomena under alternative models, such as when traders are budget limited.
 %
@@
 \begin{itemize}
-	\item the \emph{payoff function} of the market corresponds to the \emph{sufficient statistics} of the probability distribution;
+	\item the market's \emph{payoff function} corresponds to the \emph{sufficient statistics} of the distribution;
 	\item the vector of \emph{outstanding shares} in the market corresponds to the \emph{natural parameter vector} of the distribution;
@@
 In Section~\ref{sec:bayesians} we consider traders that use a conjugate prior to update their beliefs, and we study how their trades would affect the market state.
-\jake{Do we show the following?} We show that although these two kinds of traders pick the eventual market state as a convex combination of their private belief and current market state, they do so in dual spaces.
 	\item In Section~\ref{sec:exputil} we consider when our agents are risk-averse in an interesting special case, that is under the assumption they utilize \emph{exponential utility} to optimize their bets.
@@
 In this case we can characterize precisely how a single trader interacts with the market as well as the equilibrium reached given multiple traders; this result is achieved via using a \emph{potential game} argument.
-The eventual market state turns out to be a weighted combination of trader beliefs and initial market state; the weights depend on the risk aversion parameter of the individual traders.
+The eventual market state is a weighted combination of traders' beliefs and the initial state; the weights are set according to risk aversion parameters.
 	\item In Section~\ref{sec:budgets} we consider \emph{budget-limited traders} who are constrained in how much they influence the market.
@@
 
-\jake{I want to add more to this list. It's ok but doesn't totally reflect what's in the last few sections.}
-
-\subsection*{Related Work}
-
+% \jake{I want to add more to this list. It's ok but doesn't totally reflect what's in the last few sections.}
+
+\textbf{Related Work.}
 The notion of an exponential family distribution is fundamental to this paper.
@@
 We let $p(x;\mu)$ be the maximum entropy distribution with expected statistic $\mu$.
-Specifically, it is the solution to the following mathematical program:\footnote{We assume that the minimum is finite and achieved for all $\mu \in \mM$.
+Specifically, it is the solution to the following program:\footnote{We assume that the minimum is finite and achieved for all $\mu \in \mM$.
 Some care is needed to ensure this holds for specific statistics and outcome spaces.
@@
 Let $\tp \in \mP$ be a density (not necessarily from an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\hmu \in \mM$ be an alternative report.
-Then note from~(\ref{exp-fam}) that
+Then from~(\ref{exp-fam}) note
 %
@@
 
-There is a well-known duality between scoring rules and cost-function based markets~\citep{acv13,H03}. To see this in our context, recall from~(\ref{equalizer-rule}) that
+There is a well-known duality between scoring rules and cost-function based markets~\citep{acv13,H03}. To see this in our context, recall from~(\ref{equalizer-rule}):
 %
@@
 Rather than working directly with the utilities of every trader, we will work with the log of their utility values.\footnote{It is important to note that the potential function analysis still applies for any monotonically increasing transformation of the traders' utility functions.} Now the log-utility of trader $i$ is $$U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i}).$$
-Thus, Theorem \ref{thm:potl} applies and we can find the equilibrium market state by maximizing $\Phi(\vec{\delta})$ for each $\delta_{i}$:
+We can now apply Theorem \ref{thm:potl}, hence the equilibrium state is obtained by jointly maximizing $\Phi(\vec{\delta})$ for each $\delta_{i}$:
 \begin{eqnarray*}
 \nabla_{\delta_{i}}\Phi(\vec{\delta})&=&\nabla T\left(\theta+\sum_{j=1}^{n}\delta_{i}\right)-\nabla T(\htheta_{i}-a_{i}\delta_{i}) = 0.
-\end{eqnarray*}¥
+\end{eqnarray*}
 This leads to the following expression for the final market state.
@@
 \frac{\theta+\sum_{i=1}^{n}\left(\frac{\htheta_{i}}{a_{i}}\right)}{1+\sum_{i=1}^{n}\frac{1}{a_{i}}}$$
-We see that the equilibrium state is a convex combination of all beliefs and the initial market state, where beliefs are weighted by the agents' risk tolerances (inverse risk aversion).
+We see that the equilibrium state is a convex combination of the initial market state and all agent beliefs, with the latter weighted according to risk tolerance.
 

### 2014-02-12 WriteUp/maxentpredmkts.tex

--- 2014-02-12 WriteUp/maxentpredmkts.tex
@@
 \section{Budget-limited Aggregation} \label{sec:budgets}
-In this section, we consider the evolution of the market state when traders are budget-limited and their budgets are known to the market maker.
-We consider a cost-function based market with cost function $C$ and payoff $\phi(x)$ with non-negative payoff.
+
+In this section, we consider the evolution of the market state when traders are budget-limited.
+% and their budgets are known to the market maker. We consider a cost-function based market with cost function $C$ and payoff $\phi(x)$ with non-negative payoff. 
 We assume that the traders trade in multiple instances of the market.
-As before, the market price is interpreted as a probability density over the outcome space by viewing the share vector as the natural parameters of an exponential family distribution.
-We measure the error in prediction using the standard log loss.
-
-We show that traders with faulty information can only induce a limited amount of additional loss to the market's prediction.
+As before, the market price is interpreted as a probability density over the outcome space and the share vector as the natural parameter of an exponential family distribution.
+Consistent with the connections drawn in Section~\ref{sec:scoring} and throughout, we measure the error in prediction using the standard log loss.
+
+We show that traders with faulty information can only impose a limited amount of additional loss to the market's prediction.
 Further, since informative traders experience an expected increase in budget,  they will eventually be unconstrained and allowed to carry out  unrestricted trades.
@@
 \paragraph{Budget-limited trades}
+
 Let $\alpha$ be the budget of a trader in the market.
-Suppose with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
-Now suppose that $\alpha<C(\htheta)-C(\theta)$; that is, the trader's budget precludes purchasing enough shares to move the market state to his belief.
+Suppose that with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
+Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is, the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
 In this case, we want to budget-limit the trader's influence on the market state.
 
-\begin{definition}
-Let the current market state be given by $\theta$.
-Let the final market state $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
+Let the current market state be given by $\theta$ and let the final market state be $\theta'=\lambda\htheta+(1-\lambda)\theta$ where $\lambda=\min\left(1, \frac{\alpha}{C(\htheta) - C(\theta)}\right)$.
 The cost to the trader to move the market state from $\theta$ to $\theta'$ is at most his budget $\alpha$ and is called his \emph{budget-limited trade}.
-\end{definition}
 

### 2014-02-12 WriteUp/maxentpredmkts.tex
tiny edit

@@
 In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{T}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
-When the cost function is adjusted to $C_{\lambda}$ with an inverse liquidity parameter $\lambda$ Section \ref{sec:exputil},  the trader receives an expected payoff of at least $\frac{1}{a} D_{T}(\lambda\theta,\htheta)$.
+When the cost function is adjusted to $C_{\lambda}$ with an inverse liquidity parameter $\lambda$ as in Section \ref{sec:exputil},  the trader receives an expected payoff of at least $\frac{1}{a} D_{T}(\lambda\theta,\htheta)$.
 

### 2014-02-12 WriteUp/maxentpredmkts.tex
we are at 18...

@@
 Indeed,~\citet{gneiting2007strictly} observe that specific instances of scoring rules for continuous outcomes are lacking, and survey various possibilities.
-In an early work,~\citet{matheson1976scoring} designed scoring rules for real-valued outcomes from those for binary outcomes.
+% In an early work,~\citet{matheson1976scoring} designed scoring rules for real-valued outcomes from those for binary outcomes.
 
@@
 
-Prediction markets have been known to perform well in practice~\citep{AGT07,Wol04,pennock2001real}.
+Prediction markets are known to perform well in practice~\citep{AGT07,pennock2001real}.
 However, a sound theory for interpreting trader behavior and market prices is an ongoing area of study~\citep{wolfers2006interpreting}.
-At one extreme, agents are assumed myopic and risk-neutral, which implies that they move the market state to their belief~\citep{chen2010new}.
+At one extreme, agents are assumed myopic and risk-neutral, implying they move the market state to their belief~\citep{chen2010new}.
 At the other extreme, agents are strategic and the market fully incorporates all information~\citep{ostrovsky2012}.
-Strategic behavior in prediction markets is also addressed in~\citep{chen2010gaming,dimitrov08}.
+Strategic behavior in prediction markets is addressed in~\citep{chen2010gaming,dimitrov08}.
 
@@
 \begin{definition} \label{def:proper-score}
-A scoring rule $S$ is  \emph{proper} for statistic $\phi$ if for each $\mu \in \mM$ and $p \in \mP$ with expected statistic $\mu$, we have
+A scoring rule $S$ is  \emph{proper} for statistic $\phi$ if for each $\mu \in \mM$ and $p \in \mP$ with expected statistic $\mu$, we have for all $\hmu \neq \mu$
 %
 \begin{equation} \label{eq-proper}
-\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)]
-\end{equation}
-%
-for all alternative $\hmu \neq \mu$.% If the domain is $\mD = \mP$, the set of all possible densities, then we simply say the scoring rule is \emph{proper}.
+\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)].
+\end{equation}
+%
+% If the domain is $\mD = \mP$, the set of all possible densities, then we simply say the scoring rule is \emph{proper}.
 \end{definition}
@@
 
-When the outcome space is finite we can recover classical scoring rules from the definition by using the statistic $\phi : \mX \rightarrow \{0,1\}^{\mX}$ that maps an outcome $x$ to a unit vector with a 1 in the component corresponding to $x$.
+When the outcome space is finite we recover classical scoring rules by using the statistic $\phi : \mX \rightarrow \{0,1\}^{\mX}$ that maps an outcome $x$ to a unit vector with a 1 in the component corresponding to $x$.
 The expectation of $\phi$ is then exactly the probability mass function.
@@
 \begin{eqnarray}
-& & \Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\hmu)] \nonumber\\
+\Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\hmu)]
 & = & \la \theta, \mu \ra - T(\theta) - \la \htheta, \mu \ra + T(\htheta) \nonumber\\
@@
 \log p(x;\mu) & = & G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra \\
-& = & \la \theta, \mu \ra - T(\theta) - \la \theta, \mu - \phi(x) \ra \\
-& = & \la \theta, \phi(x) \ra - T(\theta),
+& = & \la \theta, \mu \ra - T(\theta) - \la \theta, \mu - \phi(x) \ra = \la \theta, \phi(x) \ra - T(\theta),
 \end{eqnarray*}
@@
 We are interested in eliciting the mean $\mu$ and variance $\sigma^2$, so as a statistic we take $\phi(x) = (x, x^2)$ for which $\Exp_p[\phi(x)] = (\mu, \mu^2+\sigma^2)$.
-The maximum entropy distribution for a given mean and variance is the normal distribution, and taking its log density gives the scoring rule
+The max entropy distribution for a given mean and variance is the Gaussian, whose log density gives the scoring rule
 %
@@
 As we saw in Section~\ref{sec:pred-market}, such a trader moves the prices to its own expectation $\mu$.
-However, this means that the market does not perform any meaningful aggregation of the agent's belief, as the final prices are simply the final agent's expectation.
+However, this means that the market does not perform meaningful aggregation of agents' beliefs, as the final prices are simply the final agent's expectation.
 
@@
 In this section we relax the standard assumption that agents in the market are risk-neutral.
-We show that, with sufficient extra structure to the agents' beliefs and utilities, the market performs a clean aggregation of the agents' expectations in the form of a simple weighted average.
+We show that with sufficient extra structure to the agents' beliefs and utilities, the market performs a clean aggregation of the agents' beliefs via a simple weighted average.
 Assume that the agent has an exponential utility function for wealth $w$:
@@
 %
-Here $a$ controls the level of risk aversion: the agent is more risk averse as $a$ increases, and as $a$ tends to 0 we approach linear utility (risk-neutrality).
+Here $a$ controls the risk aversion: the agent's aversion grows as $a$ increases, and as $a$ tends to 0 we approach linear utility (risk-neutrality).
 Specifically, $a$ is the Arrow-Pratt coefficient of absolute risk aversion, and exponential utilities of the form~(\ref{exp-util}) are the unique utilities that exhibit constant absolute risk aversion~\citep[Chap.\ 11]{VarianMicro}.
@@
 \begin{proof}
-The agent's optimal trade maximizes its expected utility, or equivalently the certainty equivalent. From Lemma~\ref{lem:exp-util} and the fact that $T = C$, the agent therefore maximizes
+The agent's optimal trade maximizes its expected utility, or equivalently the certainty equivalent. From Lemma~\ref{lem:exp-util} and that $T = C$, the agent maximizes
 %
@@
 Therefore, define the \emph{target shares} as $\ttheta = \grad T^{-1}_{\lambda}(\hmu)$.
-The target shares $\htheta$ and natural parameter $\ttheta$ are related by
-%
-$$
-\grad T(\htheta) = \grad T_{\lambda}(\ttheta) = \hmu.
-$$
-%
-It is straightforward to check that $\grad T_{\lambda}(\ttheta) = \grad T(\lambda \ttheta)$ so we have
+The target shares $\htheta$ and natural parameter $\ttheta$ are related by $\grad T(\htheta) = \grad T_{\lambda}(\ttheta) = \hmu$.
+In addition it is straightforward to check that $\grad T_{\lambda}(\ttheta) = \grad T(\lambda \ttheta)$ so we have
 %
@@
 In the exponential family market, the cost function $C$ is identical to the log partition function $T$ defined in (\ref{log-part}).
-Let $\vec{\delta}$ be the matrix of share vectors purchased by every trader in the market when the market has reached equilibrium.
+Let $\vec{\delta}$ be the matrix of share vectors purchased by every trader in the market at equilibrium.
 Let $\theta$ be the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his risk aversion parameter.
@@
 
-Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
-$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
+Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is $\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$.
 \end{proof}
@@
 \begin{eqnarray*}
-D_{T}(\theta',\htheta)&=&D_{T}\left(\lambda\htheta+(1-\lambda)\theta,\htheta\right)\\
-&\leq&\lambda D_{T}( \htheta ,\htheta)+ (1-\lambda) D_{T}( \theta ,\htheta) 
+D_{T}(\theta',\htheta)&=&D_{T}\left(\lambda\htheta+(1-\lambda)\theta,\htheta\right) \leq \lambda D_{T}( \htheta ,\htheta)+ (1-\lambda) D_{T}( \theta ,\htheta) 
 = (1-\lambda) D_{T}( \theta ,\htheta)
@@
 \end{eqnarray*}¥
-A trader who adjust the market state may expect positive profit $\geq \lambda D_{T}(\theta,\htheta)$.
+A trader who adjusts the market state may expect positive profit $\geq \lambda D_{T}(\theta,\htheta)$.
  \end{proof}

### 2014-02-12 WriteUp/maxentpredmkts.tex
returned the storkey reference
revised abstract and intro

--- 2014-02-12 WriteUp/maxentpredmkts.tex
@@
 We give a full development of this relationship and explore a range of benefits.
-We draw connections between the information aggregation of market prices and the belief aggregation of learning agents that relying on exponential family distributions.
-We develop a very natural analysis of the market behavior as well as the price equilibrium under the assumption that the traders elicit risk aversion according to exponential utility.
-We also consider similar phenomena under alternative models, such as when traders are budget limited.
+We draw connections between the information aggregation of market prices and the belief aggregation of learning agents that rely on exponential family distributions.
+We develop a very natural analysis of the market behavior as well as the price equilibrium under the assumption that the traders exhibit risk aversion according to exponential utility.
+We also consider similar aspects under alternative models, such as when traders are budget constrained.
 %
@@
 Further, since the trades are done sequentially, the trader is allowed to observe all past trades in the market and update his private information based on this information.
-In some sense the market prices, which are in effect the prices at which the marginal trader is willing to buy or sell the available securities, can be interpreted as an aggregate ``consensus probability forecast'' of the event in question.
+In this sense the market prices, which are in effect the prices at which the marginal trader is willing to buy or sell the available securities, can be interpreted as an aggregate ``consensus probability forecast'' of the event in question.
 % For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins.
@@
 
-We give an brief overview of the work on prediction market design below.
-Much of this work has focused heavily on structural properties of the mechanism: incentive compatibility, the market maker loss, the available liquidity, the fluctuations of the prices as a function of the trading volume, to name a few.
+%We give an brief overview of the work on prediction market design below. 
+Much of the work on prediction market design has focused heavily on structural properties of the mechanism: incentive compatibility, the market maker loss, the available liquidity, the fluctuations of the prices as a function of the trading volume, to name a few.
 Absent from much of the literature is a corresponding \emph{semantics} of the market behavior or the observed prices.
@@
 In the present work we show that a number of classical statistical tools can be leveraged to design a prediction market framework in the mold of \emph{exponential family distributions}; we show that this statistical framework leads to a number of attractive properties and interpretations.
-Common concepts in statistics -- e.g.
-\emph{entropy maximization}, \emph{log loss}, and \emph{bayesian inference} -- relate to natural aspects of our class of mechanisms.
-In particular, the central objects in our market framework can be interpreted via conceptual objects used to define exponential families:
+Common concepts in statistics---including \emph{entropy maximization}, \emph{log loss}, and \emph{bayesian inference}---relate to natural aspects of our class of mechanisms.
+In particular, the central objects in our market framework can be interpreted via concepts used to define exponential families:
 \begin{itemize}
@@
 	\item the market prices correspond to \emph{mean parameters};
-	\item the profit potential for a trader corresponds to a \emph{Kullback-Leibler divergence} between the trader's belief and that of the market.
+	\item the market's cost function corresponds to the distribution's \emph{log-partition} function.
+%	\item the profit potential for a trader corresponds to a \emph{Kullback-Leibler divergence} between the trader's belief and that of the market.
 \end{itemize}
@@
 
-We begin with a discussion of scoring rules in Section~\ref{sec:scoring} for exponential family distributions, and in particular we show that the logarithmic scoring rule is the only score that is \emph{proper} for the class of exponential families.
+We begin in Section~\ref{sec:scoring} with a discussion of scoring rules based on exponential family distributions, and we show how the framework leads to a variety of scoring rules for continuous outcome spaces.
 We turn our attention to market design in Section~\ref{sec:maxentmarkets} and give a full description of our proposed mechanisms.
@@
 In Section~\ref{sec:bayesians} we consider traders that use a conjugate prior to update their beliefs, and we study how their trades would affect the market state.
-	\item In Section~\ref{sec:exputil} we consider when our agents are risk-averse in an interesting special case, that is under the assumption they utilize \emph{exponential utility} to optimize their bets.
-We show how this dovetails nicely with the notion of a \emph{certainty equivalent} which is often used in risk-aversion settings.
-In this case we can characterize precisely how a single trader interacts with the market as well as the equilibrium reached given multiple traders; this result is achieved via using a \emph{potential game} argument.
-The eventual market state is a weighted combination of traders' beliefs and the initial state; the weights are set according to risk aversion parameters.
+	\item In Section~\ref{sec:exputil} we consider \emph{risk-averse} agents that optimize their bets according to exponential utility. 
+	%We show how this dovetails nicely with the notion of a \emph{certainty equivalent} which is often used in risk-aversion settings. 
+	In this case we can characterize precisely how a single trader interacts with the market, as well as the equilibrium reached given multiple traders; this result is achieved via a potential game argument.
+The eventual market state is a weighted combination of traders' beliefs and the initial state; the weights are proportional to risk aversion parameters.
 	\item In Section~\ref{sec:budgets} we consider \emph{budget-limited traders} who are constrained in how much they influence the market.
@@
 The notion of an exponential family distribution is fundamental to this paper.
-For a comprehensive introductions to these distributions, see~\citep{Barndorff78,WainJordan08}.
+For comprehensive introductions to these distributions, see~\citep{Barndorff78,WainJordan08}.
 Exponential families are intimately tied to the notions of log loss and entropy, but can be generalized to other types of convex losses and information, as shown by~\citet{GrunwaldDawid}, who also make a connection to scoring rules.

### 2014-02-12 WriteUp/maxentpredmkts.tex

--- 2014-02-12 WriteUp/maxentpredmkts.tex
@@
 At the other extreme, agents are strategic and the market fully incorporates all information~\citep{ostrovsky2012}.
-Strategic behavior in prediction markets is addressed in~\citep{chen2010gaming,dimitrov08}.
+% Strategic behavior in prediction markets is addressed in~\citep{chen2010gaming,dimitrov08}.
 
@@
 We see the potential to draw more on the mathematical finance literature to take into account risk aversion, as prediction markets are simply single-period financial markets~\citep[Part I]{follmerFinance}.
+We also note that connections between Machine Learning and market mechanisms have been explored in \cite{storkey11}.
 

### 2014-04-08 WriteUp/Original/maxentpredmkts.tex
inadvertent changes

--- 2014-02-12 WriteUp/maxentpredmkts.tex
@@
-% Adjusted to add \usepackage[numbers]{natbib}
-% \bibliographystyle{acmsmall} and \documentclass[prodmode,acmec]{ec-acmsmall}
-% Jan 5, 2013 - David Parkes
-%
-% v2-acmsmall-sample.tex, dated March 6 2012
-% This is a sample file for ACM small trim journals
-%
-% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
-% (c) 2010 Association for Computing Machinery (ACM)
-%
-% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
-% Users can also go through the FAQs available on the journal's submission webpage.
-%
-% Steps to compile: latex, bibtex, latex latex
-%
-% For tracking purposes => this is v1.3 - March 2012
-
-\documentclass[prodmode,acmec]{ec-acmsmall} % Aptara syntax
-
-% Package to generate and customize Algorithm as per ACM style
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+
+\documentclass[prodmode,acmec]{ec-acmsmall} %
+
+%
 \usepackage[ruled]{algorithm2e}
@@
 
-% Metadata Information
+%
 \acmVolume{X}
@@
 \usepackage{amssymb}
-%\usepackage{amsthm}
+%
 \usepackage{parskip}
@@
 
-%new commands
+%
 \newcommand{\la}{\langle}
@@
 \newcommand{\ttheta}{\tilde{\theta}}
-%---------------------------------------------------------------------------
+%
 \newtheorem{thm}{Theorem}[section]
@@
 \newcommand{\unrmk}{\bigbreak}
-%\newcommand{\proof}{\noindent{\bf Proof. }}
+%
 \newcommand{\unproof}{\hfill$\Box$\bigbreak}
@@
 \newcommand{\qvec}{\mathbf{q}}
-\newcommand{\lpf}{\mathbf{\psi}} %logpartition function
+\newcommand{\lpf}{\mathbf{\psi}} %
 \newcommand{\family}{\mathcal F} 
-\newcommand{\suff}{\pmb{\phi}} % suff stats
+\newcommand{\suff}{\pmb{\phi}} %
 \newcommand{\muvec}{\pmb{\mu}}
@@
 
-% Document starts
+%
 
 \begin{document}
-% Page heads
+%
 \markboth{Abernethy et al.}{Information Aggregation in Exponential Family Markets}
-% Title portion
+%
 \title{Information Aggregation in Exponential Family Markets}
@@
 \affil{Google India}}
-%%\author{GANG ZHOU
-%%\affil{College of William and Mary}
-%%YAFENG WU
-%%\affil{University of Virginia}
-%%TING YAN
-%%\affil{Eaton Innovation Center}
-%%TIAN HE
-%%\affil{University of Minnesota}
-%%CHENGDU HUANG
-%%\affil{Google}
-%%JOHN A. STANKOVIC
-%%\affil{University of Virginia}
-%%TAREK F. ABDELZAHER
-%%\affil{University of Illinois at Urbana-Champaign}}
-% NOTE! Affiliations placed here should be for the institution where the
-%       BULK of the research was done. If the author has gone to a new
-%       institution, before publication, the (above) affiliation should NOT be changed.
-%       The authors 'current' address may be given in the "Author's addresses:" block (below).
-%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
-%       currently affiliated with NASA.
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
 
@@
 %
-% In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets.
-We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data.
-Similarly,  the market state can be interpreted as a distribution over the outcome space.
-In particular, we show that a proper scoring rule can be derived from maximum entropy distributions.
-This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces.
-In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility.
-We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function.
-These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders.
-In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market.
-Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
+%
 \end{abstract}
@@
 
-%\acmformat{Jacob Abernethy, Sindhu Kutty, S\'{e}bastien Lahaie, and Rahul Sami, 2014. Information Aggregation in Exponential Family Markets}
-%%Gang Zhou, Yafeng Wu, Ting Yan, Tian He, Chengdu Huang, John A. Stankovic, and Tarek F. Abdelzaher, 2010. A multifrequency MAC specially designed for  wireless sensor network applications.
-
-% At a minimum you need to supply the author names, year and a title.
-% IMPORTANT:
-% Full first names whenever they are known, surname last, followed by a period.
-% In the case of two authors, 'and' is placed between them.
-% In the case of three or more authors, the serial comma is used, that is, all author names
-% except the last one but including the penultimate author's name are followed by a comma,
-% and then 'and' is placed before the final author's name.
-% If only first and middle initials are known, then each initial
-% is followed by a period and they are separated by a space.
-% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.
+%
+%
+
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
 
 \begin{bottomstuff}
-%%This work is supported by the National Science Foundation, under
-%%grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.
-%%
-%%Author's addresses: G. Zhou, Computer Science Department,
-%%College of William and Mary; Y. Wu  {and} J. A. Stankovic,
-%%Computer Science Department, University of Virginia; T. Yan,
-%%Eaton Innovation Center; T. He, Computer Science Department,
-%%University of Minnesota; C. Huang, Google; T. F. Abdelzaher,
-%%(Current address) NASA Ames Research Center, Moffett Field, California 94035.
+%
+%
+%
+%
+%
+%
+%
+%
+%
 \end{bottomstuff}
@@
 \section{Introduction}
-%------------------------------------------------------------
-%Outline
-%------------------------------------------------------------
-%-introduction [Sindhu; Thursday]
-%-literature review
-%- maxent scoring rule [Sebastien]
-%	- generalized lmsr
-%	- exp family review
-%	[- does modularity (a la Hanson) still hold?]
-%- prediction market setup [Sebastien]
-%	- MLE with risk neutrality
-% 	need a definition of term 'exponential family prediction market'
-%- conjugate prior market [Sindhu; Wednesday] done
-%	- also on hyperparam?
-%- exp-util [Sebastien]
-%- equilibrium for repeated trades [Sindhu; Tuesday] done
-% 	[- rate of convergence?]
-%- budgets and adversarial markets [Sindhu; Monday] done
-% [any liquidity results?] [Sindhu: weekend]
-% conclusion
-% bibliography
-
-%------------------------------------------------------------
-% Naming conventions:
-%------------------------------------------------------------
-%htheta belief
-%theta current market state
-%thetaÕ final state
-%T is the log partition function
-
-%- prediction market construction
-%	- exponential failies
-%	- scoring rules
-%	- information market
-%- Traders
-%	- aggregation process
-%		- bayesian traders
-%		- frequentist
-%	- utility
-%		- linear
-%		- exponential
-%	- budgets
-%- multiple trades
-%- multiple traders
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+
+%
+%
+%
+%
+%
+%
+%
+
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
 
@@
 In this sense the market prices, which are in effect the prices at which the marginal trader is willing to buy or sell the available securities, can be interpreted as an aggregate ``consensus probability forecast'' of the event in question.
-% For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins.
-
-% JAKE: I think most people reading this will have enough background on markets.
-% Currently, most deployed prediction markets have binary or finite discrete outcome spaces.
-This means that the outcome of every event can be characterized by a finite set of mutually exclusive exhaustive outcomes.
-The securities in these market are usually tied to each outcome so that buying (respectively, selling) a security related to an outcome, indicates a belief in an increased (respectively, decreased) probability of occurrence of the outcome.
-
-% JAKE: removing this paragraph, we don't want to harp too much on the efficiency question
-%       as this has already been adressed in a number of papers 
-% Directly applying techniques used for binary outcome spaces to large outcome spaces is computationally infeasible.
-In fact, it has been shown that computing and updating prices for combinatorial outcomes spaces is $\#$P-hard \cite{Chen08}.
-Infinite outcome spaces pose a particular challenge in prediction markets since it is not clear how to define securities in such spaces.
-The most direct way of discretizing the space may lead to loss of information from the traders.
-Extending prediction market techniques to this space has been a subject of recent interest.
-Various techniques to bound market maker loss in both a combinatorial and infinite outcome space have been studied (see, for instance, \cite{Abernethy13,OthmanS11}).
-
-% \jake{I want to use this paragraph for a general overview of mm frameworks} One popular form of prediction markets is the market scoring rule~\cite{H03}.
-A market scoring rule considers all trades as a single chronological sequence.
-Traders earn rewards proportional to the incremental reduction in prediction loss caused by their trades in comparison to the previous trade.
-In other words, their rewards depend on the change in market probabilities caused by their trade, as well as on the eventual outcome.
-Thus, each trader has an incentive to minimize the prediction loss.
-In this format, the {\em market maker} who runs the market can suffer an overall loss, but Hanson~\cite{H03} showed that, for market scoring rules on finite outcome spaces, the loss of the market maker can be bounded.
-
-%We give an brief overview of the work on prediction market design below. 
+%
+
+%
+%
+
+%
+%
+%
+
+%
+
+%
 Much of the work on prediction market design has focused heavily on structural properties of the mechanism: incentive compatibility, the market maker loss, the available liquidity, the fluctuations of the prices as a function of the trading volume, to name a few.
@@
 	\item the market's cost function corresponds to the distribution's \emph{log-partition} function.
-%	\item the profit potential for a trader corresponds to a \emph{Kullback-Leibler divergence} between the trader's belief and that of the market.
+%
 \end{itemize}
-% \sindhu{I don't think we have the KL divergence result here -- is this the bregman divergence in the budget section?}
+%
 
@@
 	\item In Section~\ref{sec:exputil} we consider \emph{risk-averse} agents that optimize their bets according to exponential utility. 
-	%We show how this dovetails nicely with the notion of a \emph{certainty equivalent} which is often used in risk-aversion settings. 
+	%
 	In this case we can characterize precisely how a single trader interacts with the market, as well as the equilibrium reached given multiple traders; this result is achieved via a potential game argument.
@@
 
-% \jake{I want to add more to this list. It's ok but doesn't totally reflect what's in the last few sections.}
+%
 
@@
 Indeed,~\citet{gneiting2007strictly} observe that specific instances of scoring rules for continuous outcomes are lacking, and survey various possibilities.
-% In an early work,~\citet{matheson1976scoring} designed scoring rules for real-valued outcomes from those for binary outcomes.
+%
 
@@
 At the other extreme, agents are strategic and the market fully incorporates all information~\citep{ostrovsky2012}.
-% Strategic behavior in prediction markets is addressed in~\citep{chen2010gaming,dimitrov08}.
+%
 
@@
 %
-% If the domain is $\mD = \mP$, the set of all possible densities, then we simply say the scoring rule is \emph{proper}.
+%
 \end{definition}
@@
 %
-%Typically the exponential distribution is parametrized by its rate $\lambda = 1/\mu$.
-In that case the rule takes the form $S(\mu, x) = -\lambda x + \log \lambda$, and we see that $\lambda$ corresponds to the natural parameter.
+%
 %
@@
 
-%\subsection{Exponential Smoothing}
-%One way to model the exponential family prediction market might be through exponential smoothing.
-%\begin{itemize}
-%\item A \emph{time series} is a set of data points sampled at periodic instances.
-%\item \emph{Smoothing} is the process of creating an approximation function based on time series that while capturing  patterns in the data is relatively insensitive to noise.
-%\item Natural first step is to use moving average. But average on how many samples (say $k$)? What to do until the first $k$ samples have been received?
-%\item
-%Next attempt might be to use weighted moving average. Advantage is that you can give higher weight to more recent terms. But still same disadvantage as simple moving average technique.
-%\item
-%\emph{Exponential Smoothing} attempts to remove this disadvantage.
-%\end{itemize}
-%
-%Exponential smoothing is defined iteratively for $\alpha\in(0,1)$ as
-%\begin{eqnarray*}
-%s_{1} &=& x_{0}\\
-%s_{t}&=&\alpha x_{t-1}+(1-\alpha)s_{t-1}
-%\end{eqnarray*}
-%
-%Smaller values of $\alpha$ means greater smoothing e.g., $\alpha=0$ gives a constant function ($=x_{0}$).
-%Note that the selection of $x_{0}$ is important -- and increases in importance as $\alpha$.
-Some advantages over moving average: all data points matter (decreasing importance with time) and computationally only last data point need be stored.
-%\pagebreak
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
 \section{Exponential Family Markets}
@@
 In his seminal paper~\citet{H03} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both in order to perform well in thin or thick markets.
-In this section, we adapt the generalized log scoring rule to a market scoring rule which leads to markets with simple closed-form cost functions for many statistics of interest.%The exponential families framework also allows us to move beyond risk-neutral agents to understand the aggregation properties of prediction markets under alternative behaviors, such as Bayesian updating and risk aversion.
+In this section, we adapt the generalized log scoring rule to a market scoring rule which leads to markets with simple closed-form cost functions for many statistics of interest.%
 
@@
 
-%* DIFFERENTIAL ENTROPY
-%* UNBOUNDED LOSS?
-%* EXAMPLE OF LOSS FOR EXPONENTIAL/GAUSSIAN
+%
+%
+%
 
@@
 
-%As before, we are interested in eliciting the sufficient statistics of the data.
-We assume that the outcome is drawn from an exponential family distribution; the prediction market is setup as before with the cost function corresponding to the log partition function and the payoff function corresponding to the sufficient statistics.
-Thus, the market state provides an estimate on the natural parameter of the distribution from which the outcome is drawn.
-Additionally, we assume that the market also makes public the total number of traders that have traded in the market.
-
-%The goal is to aggregate information from risk neutral agents who have a belief distribution over the natural parameters.
-This prior distribution is updated by the agents based on the current market state, They also each have to access to the empirical mean of sufficient statistics based on a fixed number  $m$ of data points.
-Assuming a conjugate prior, both the prior and posterior belief distributions on the natural parameters are also an exponential family distributions.
+%
+
+%
 
@@
 %
-Thus the posterior mean is a convex combination of the prior and empirical means, and their relative weights depend on the phantom and empirical sample sizes.% need a reference here
+Thus the posterior mean is a convex combination of the prior and empirical means, and their relative weights depend on the phantom and empirical sample sizes.%
 
@@
 
-%Suppose the current market state is $\theta$ and $i$ traders have traded in the market when trader $i+1$ with prior belief distribution $p(\theta;b^{i}_{0})$ enters the market.
-Here $b^{i}_{0} = (n\nu_{i}, n)$.
-This trader also has access to private information in the form of empirical sufficient statistics $\hmu_{i}$ from $m$ data points.
-Recall from Proposition \ref{prop:exp} that natural parameter $\theta$ corresponds to expected statistics $\nabla T(\theta)$.
-Thus, he updates his belief as $p(\theta;b_{i})$ where $b=(m\hmu+mi \nabla T(\theta) + n\nu_{i}, n+mi+m)$.
-
-%Suppose the trader wishes to maximize his expected payoff.
-Then the number of shares $\delta_{i}$ that he purchases when the current market state is $\theta$ is given by $$\arg\max_{\delta_{i}}\E_{\theta\sim b_{i}}\E_{x\sim \theta}\left[ \delta_{i}\phi(x) - T(\delta_{i} + \theta) + T(\theta)\right]$$
-%But, from Equation \ref{eq:prior} we have $\E_{\theta\sim b}\E_{x\sim \theta}[\phi(x)]=\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}$.
-To obtain the maximum, we set the gradient of the above expression with respect to $\delta_{i}$ to $0$.
-Thus, we have for the optimal number of shares $\delta_{i}^{*}$ $$\nabla T(\delta_{i}^{*} + \theta)=\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}$$ Thus, from Proposition \ref{prop:exp} we have that for an exponential family prediction market, the final market state is given by $$\nabla G\left(\frac{n\nu+m\hmu+mi \nabla T(\theta)}{ n+m(i+1)}\right)$$ where $G$ is the convex conjugate of $T$.
-Thus, the final market state is a convex combination of prior, posterior and market means.
-
-%Let $p(x;\theta)$ denote a probability density drawn from an exponential family with sufficient statistic $\phi : \mX \rightarrow \bR^d$, where $\theta$ is the natural parameter:
-%$$
-%p(x;\theta) = \exp \left[ \la \theta, \phi(x) \ra \right],
-%$$
-%and
-%$$
-%g(\theta) = \log \int_{\mX} \exp\la\phi(x),\theta\ra d\!x.
-%$$
-%Recall that $\grad g(\theta) = \Exp[\phi(x)]$ and $\grad^2 g(\theta) = \Var[\phi(x)]$. The family of conjugate priors is also an exponential family and takes the form
-%$$
-%p(\theta; n,\nu) = \exp\left[ \la n\nu,\theta \ra - ng(\theta) - h(\nu,n) \right].
-%$$
-%Here the feature map is $\psi(\theta) = (\theta, -g(\theta))$, the natural parameter is $(n\nu, n)$ where $n \in \bR$ and $\nu \in \bR^d$.
-The normalizer $h(\nu, n)$ is convex in $(n\nu,n)$.
-It is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$.
-The justification for this is that
-%$$
-%\Exp_{\theta} \left[ \Exp_x \left[ \phi(x) | \theta \right]\right] 
-%= \Exp_{\theta} \left[\grad g(\theta)\right]
-%= \nu.
-%$$
-%(The proof is straightforward but not obvious. I have a reference for this but it's impossible to read---it's a mathematical statistics paper.)
-
-%Suppose we draw a sample $X = (x_1,\ldots,x_m)$ of size $m$, and denote the empirical mean by $\mu[X] = \sum_{i=1}^m \phi(x_i)$. The posterior distribution is then
-%$$
-%p(\theta|X) \propto p(X|\theta)p(\theta|n,\nu) \propto \exp\left[ \la\mu[X]+n\nu,\theta\ra - (m+n)g(\theta) \right],
-%$$
-%and so the posterior mean is 
-%\begin{equation} \label{posterior-mean}
-%\frac{m\mu[X] + n\nu}{m+n}.
-%\end{equation}
-
-%In the context of prediction markets, one could imagine an agent who uses the current market estimate as its prior, and draws a empirical sample of size $m$.
-Its belief then takes the form~(\ref{posterior-mean}), where $n$ depends on the importance the agent places on the market estimate (perhaps based on how long the market has been running).
-However, note that the \emph{mean parameter} becomes a convex combination of market state and empirical belief, and this does not translate to the \emph{natural parameter}, which is what we would have liked.
-The new natural parameter is
-%$$
-%\grad g^{-1} \left(\frac{m\mu[X] + n\nu}{m+n}\right) = \grad g^* \left(\frac{m\mu[X] + n\nu}{m+n}\right).
-%$$
-%where $g^*$ is the convex conjugate of $g$.
-
-%\section{Risk Aversion, Exponential Utility, and Market Equilibrium} \label{sec:exputil}
-%\subsection{Behavior of Traders endowed with Exponential Utility}
+%
+
+%
+%
+
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+
+%
+%
+%
+%
+%
+%
+%
+%
+
+%
+%
+%
+%
+%
+
+%
+%
 
@@
 
-%(Discussion, you can skip this paragraph.) This is because for higher $\lambda$ we are making the cost function ``more convex''.
-Specifically, for $\lambda_1 > \lambda_2$ it's the case that $C_{\lambda_1}$ is an increasing convex transformation of $C(\lambda_2)$.
-This holds because the ``coefficient of risk aversion'' $A(\theta) = C''(\theta)/C'(\theta)$ increases with $\lambda$ at every point.
-I've written it assuming single-dimensional $\theta$ but there are more complicated looking generalizations for vector $\theta$.
-In fact, this coefficient might be the ``right'' notion of liquidity at a point, not the second derivative or largest eigenvalue of the Hessian.
-We should discuss.
+%
 
@@
 We now pose the question: how will a trader reason about a possible future investment when the trader holds an existing portfolio?
-%When the agent is exposed to the market given a prior investment this should clearly affect the agent's preference for additional risk.
+%
 In the context of a trader possessing an exponential family belief together with exponential utility, we show that we can explicitly analyze how an agent incorporates an existing portfolio.
@@
 
-% As before, we consider the exponential family prediction market where traders have exponential belief. Let us also suppose that traders in this market have exponential utility \[
-% U(w) = -\frac{1}{a} \exp(-aw).
-% \]
-% %
-% Here $a$ is the coefficient of risk aversion (higher means more risk averse, and the utility function is more concave).
+%
+%
+%
+%
+%
 
@@
 Based on this belief, let $\delta_1$ be the vector of shares the agent has purchased on first entry in the market.
-%Thus, his belief distribution is given by the density $p(x;\htheta)=\exp(\la\hat{\theta}, \phi(x)\ra - T(\hat{\theta})),$ where $T(\hat{\theta}) =\int_{\mathcal{X}}\exp\{\la\hat{\theta}, \phi(x)\ra\}\ dx$ is the log partition function and $\phi(x)$ are the sufficient statistics of the outcome $x\in \mathcal{X}$.
+%
 On a subsequent entry into this market with market state $\theta'$, his optimal purchase $\delta_{2}^{*}$ is given by the solution of
-%Here $\theta' = \theta+\delta_1+\delta'$ where $\delta'$ captures the movement of the share vector by other traders in the market.
+%
 %
@@
 Let $\theta$ be the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his risk aversion parameter.
-% \jake{I need to fix a small technical detail with the fact that we are using the log utilities to use the potential argument.}
+%
 
@@
 In this section, we consider the evolution of the market state when traders are budget-limited.
-% and their budgets are known to the market maker. We consider a cost-function based market with cost function $C$ and payoff $\phi(x)$ with non-negative payoff. 
+%
 We assume that the traders trade in multiple instances of the market.
@@
 \paragraph{Budget of Informative Traders}
-%Given the  information-theoretic interpretation of the cost-function based prediction market, we now show that the informative trader in the prediction market defined above increases his budget in a round in expectation {\em under his own belief distribution}.
-We now characterize the expected change in budget for an informative trader. %The following result holds for any round $t$; for simplicity, we have therefore dropped the superscript from the notation.
+%
+We now characterize the expected change in budget for an informative trader. %
 \begin{lemma}\label{thm:growth}
@@
 Then, the expectation (over the trader's belief) of the trader's profit  is strictly positive whenever his budget is positive and his belief differs from the previous market position $\theta$.
-% Suppose that each informative trader gets a random sample of data, 
-%resulting in a sequence of trader beliefs $\theta_i$, and hence a sequence of budget-limited market positions
-%$\hat{\theta}_i$, before a final outcome $x$.
-Then, the expectation, over trader $i$'s belief distribution $\theta_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\hat{\theta}_{i-1}$.
+%
+%
+%
 \end {lemma}
@@
 &=& T(\theta)-\theta\nabla T(\htheta)-(T(\theta')-\theta' \nabla T(\htheta))\\
-%&=& T(\theta)-T(\htheta)-\nabla T(\htheta)(\theta-\htheta)-(T(\theta')-T(\htheta)- \nabla T(\htheta)(\theta'-\htheta))\\
+%
 &=& D_{T}(\theta,\htheta)-D_{T}(\theta',\htheta) \geq \lambda D_{T}(\theta,\htheta)\geq0
@@
 = (1-\lambda) D_{T}( \theta ,\htheta)
-%\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
+%
 \end{eqnarray*}¥
@@
 
-%\section{Conclusions}
-%\sindhu{Do we need this section?}
-%
-%----------------------------------------------------------------------------------------
-%	BIBLIOGRAPHY
-%----------------------------------------------------------------------------------------
+%
+%
+%
+%
+%
+%
 \bibliographystyle{acmsmall}
@@
 
-%----------------------------------------------------------------------------------------
+%
 
@@
 
-%{\cor The exponential family market maker has unbounded worst case loss.}
-%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
-%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
-%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
+%
+%
+%
+%
 \paragraph{Loss of the market maker}
@@
 &=& -\log \frac{\sigma_{1}}{\sigma_{2}}-\frac{\mu_{2}(\mu_{1}-\mu_{2})-\frac{1}{2}(\mu_{1}^{2}-\mu_{2}^{2}+\sigma_{1}^{2}-\sigma_{2}^{2})}{\sigma_{2}^{2}}\\
-%&=& \log \frac{\sigma_{2}}{\sigma_{1}}-\frac{\sigma_{1}^{2}-\sigma_{2}^{2}-(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}\\
+%
 &=& \log \frac{\sigma_{2}}{\sigma_{1}}+\frac{\sigma_{2}^{2}-\sigma_{1}^{2}}{2\sigma_{2}^{2}}+\frac{(\mu_{1}-\mu_{2})^{2}}{2\sigma_{2}^{2}}
@@
 So the Gaussian market maker loss is 
-%parallels to quadratic market maker loss? 
+%
 \begin{eqnarray*}
@@
 \rmk In statistics, the standard score (aka z-score), $(x-\mu)/\sigma$, is the (signed) number of standard deviations an observation is above the mean. \unrmk
-%\pagebreak
+%
 
@@
 $$L(\theta,x)=-\log(P_{\theta}(x))=\log\int \exp\{\theta^{T}\phi(x)\}\ dx-\theta^{T}\phi(x)=C(\theta)-\theta^{T}\phi(x)$$
-%with the correspondence between $\beta$ and $\theta$ as defined earlier. %We extend this loss function to give zero loss in the absence of input.
+%
 
@@
 
-This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders. %imp
+This means that if it can be made prohibitively expensive for an attacker to generate clones, we can set up the prediction market with mostly informative traders. %
 
 In Section \ref{sec:inf} we show that for an informative trader in every round, his budget increases in expectation. 
-%We provide an information-theoretic justification for this claim in Section \ref{sec:infth}. 
+%
 The intuition behind this  is that a trader's prediction moves the input moves the market probability closer to the true probability distribution resulting in net expected profit.
@@
 Then, the expectation (over the trader's belief) of the trader's profit  is greater than zero whenever his budget is positive and his belief differs from the previous market position $\theta$.
-% Suppose that each informative trader gets a random sample of data, 
-%resulting in a sequence of trader beliefs $\theta_i$, and hence a sequence of budget-limited market positions
-%$\hat{\theta}_i$, before a final outcome $x$.
-Then, the expectation, over trader $i$'s belief distribution $\theta_i$, of trader $i$'s realized profit is greater than zero whenever her budget is positive and her belief differs from the previous market position $\hat{\theta}_{i-1}$.
+%
+%
+%
 \end {theorem}
@@
 &=&(1-\lambda) D_{T}( \theta ,\htheta)
-%\implies D_{\psi}(\theta',\beta)+aD_{\psi}(\theta',\beta)&\leq& aD_{\psi}( \theta ,\beta)\\
+%
 \end{eqnarray*}¥

### 2014-04-08 WriteUp/response.txt
edits needed for camera-ready version

- Reduce complications in notation: Please only keep necessary notation in this paper. 
- intro revision: we meant this comment for the narrow case of MSR-style markets.
  - Ostrovsky and of Wolfers and Zitzewitz are exactly about these issues. Pennock, in his Ph.D. thesis (1999), asks the question of what happens in equilibrium when traders have different priors and budget constraints, and shows that in equilibrium, the price can be interpreted as a budget-weighted average of trader beliefs. Beyond this, most of the empirical work on prediction markets is asking exactly the question of what prediction market prices actually mean. The recent working paper of Rothschild and Sethi ("Trading Strategies and Market Microstructure: Evidence from a Prediction Market") discusses how to interpret prices in the context of a continuous market rather than equilibrium prices -- the references in that paper are also worth looking at. Much of the work on manipulation is also implicitly focused on exactly this question (including the paper "Gaming Prediction Markets" with one of the authors of this paper as a co-author!)
- If accepted, the authors should spend some effort in discussing the connection of this work to that of Abernethy-Frongilllo (COLT 2012)
  - The reviewer makes a really good point regarding our omission of a reference to [Abernethy-Frongillo'12, Thm 11]; this work is certainly related although perhaps not to the extent suggested. It is correct that [AF'12, Thm 11] is the right result to use in the 'only if' part of Thm 2.2 (we will definitely include this), which follows up on prior work of [Savage'71] and [Gneiting-Raftery'07]. However, Thm 2.2 is not an instance of or equivalent to any result in [AF'12]; the construction in Thm 2.2 is to apply a scoring rule to a parametrized family of densities. One finds a "family of densities" constructed in the proof of [AF'12, Thm 11], but this does not fall within the exponential family setting.
    - That paper shows how to construct a proper scoring rule for linear properties from any convex function using Bregman divergence. This paper seems to basically present a way of choosing that convex function; this is not uninteresting in itself, but I'm concerned as to why this connection is not made clearer in the paper.
- Highlight key contributions: the authors could have just focused on presenting characterization of the class of new proper scoring rules and behavior and analysis of agents with exponential utility function. The connection between exponential family and the proposed prediction market setting is interesting and helpful if it is briefly mentioned and discussed in the end of the paper. The paper in its current form reads like suggesting that using exponential families is justified at the modeling level, which sounds a little weak (and in fact not necessary). So I dont think they should be mentioned in the beginning. For example, I got confused by "outstanding shares". Also not all exponential family distributions have a mean parameter (which is fine if you choose to model the market price distribution in this way, but this may limit the generality of the framework).
- Please give a formal definition of exponential family distributions (with citation) and examples of commonly studied distributions that can be modeled by the exponential family studied in this paper. 
- The importance of the underlying measure v is understated. If v is always the Lebesgue measure then many distributions cannot be modeled by the exponential family as presented in the paper, e.g. normal distributions with fixed variance. Many discussions and proofs are made without reference to v. My question is: is it true that all the results and comments can be applied to arbitrary v? (I just checked a few but not all since many calculations are already oversimplified.)
  -  Please give an example of non-Lebesgue v.
  - Review 3: The reviewer makes a good point, we did not really address alternative base measure, and we will need to add more detail here. However, all results hold for any arbitrary base measure v. Often v does not appear because we work directly with the expectation mu and do not write out the implicit integral. All integrals are taken with respect to dv, e.g. (5).
- For example, I found the reference to Bergman divergence a little annoying. I also think that maximum entropy principle is overemphasized---proving that the proposed scoring rule is proper is enough to me. I also got lost at the last paragraph of Section 1.
- P3: Section 2, after first paragraph. Please clarify that P contains ALL distributions on the outcome space.
- P4: please clarify the setup. Are u always using a single outcome x?, same in the example section at Page 7
- I found the "underlying density" and "full density" discussion on P4 very confusing (what are these two densities?).
- P4, last sentence. Please quote what exactly was proven by [Grunwald and Dawid]
- P5: sufficient statistics - please clarify which distributions are not like this. Can it be that the exponential distribution is not like this?
  - Review 1: The concept of a sufficient statistic also arises for distributions other than exponential families. An example of a non-exponential family with sufficient statistic is the uniform distribution over [0,r], where r is the parameter. The sufficient statistic for r is the maximum of the observations.
- P6, THM 2.2. Where the sufficient statistics condition fits in. Can it be that the entropy maximizing exponential distribution is not like that. Please clarify.
- P7: U are using mu for both the report and expectation. Please use a different notation (e.g. hat(mu)
- Can you model classical prediction markets with finite outcome space under the proposed framework? If so then what distribution in the exponential family its correspond to?
  - Classical prediction markets (specifically, LMSR) can be recovered from our framework by using the multinomial distribution, which is an exponential family. It was originally covered in the examples of Sections 2.2 and 3.3 but removed due to space constraints.
- Section 5: Please clarify the setup up-front. What is known to the principal on the traders? Does it know $a$ or it can be estimated
  - In Section 5, the principal (market maker) does not need to know the risk aversion parameter 'a' for Thms 5.2 and 5.3 to hold. The aggregation arises directly from the agents' choice of trades.
- What kind of equilibria are studied in Section 5.2? I.e. what is the game (especially what are the strategies)? What is the dynamics under the game? How equilibria are reached from the initial state of the market?
  - we should add a few more details regarding the results of 5.2, but here are some short answers: The utility functions are defined in 5.2, the strategies are the vectors of share purchases of each agent, we are interested in pure-strategy Nash equilibria, and the dynamics are not required to establish the existence of an equilibrium. We will expand upon these in the revision.
- Section 6: What exactly does the statement "the traders trade in multiple instances of the market" mean? What are the formal definitions of an "uninformative" and an "informative" trader? The semantics of the time-varying budget alpha_i^t are not clear.
- Regarding budgets and play money: indeed the play-money markets are a case where budget constraints are key. We have experience with this in other work where a play-money market was released for the US Elections. Here the budget constraints help alleviate the problem that traders can easily manipulate prices with unreasonable trades -- the goal of Section 6 was to give formal results to this effect. Risk aversion is also plausible with play-money; for instance, a trader high on the leaderboard may not want to risk her standing with large late trades.
  - the example for risk-aversion of someone not wanting to lose their place on the leaderboard doesn't really hold water: that's a type of tournament incentive: very different from traditional risk-aversion.

### 2014-04-08 WriteUp/response.txt
edited for readability

@@
 - Reduce complications in notation: Please only keep necessary notation in this paper. 
+	- For example, I found the reference to Bergman divergence a little annoying.
+I also think that maximum entropy principle is overemphasized---proving that the proposed scoring rule is proper is enough to me.
+I also got lost at the last paragraph of Section 1.
+
 - intro revision: we meant this comment for the narrow case of MSR-style markets.
@@
 Much of the work on manipulation is also implicitly focused on exactly this question (including the paper "Gaming Prediction Markets" with one of the authors of this paper as a co-author!)
+
 - If accepted, the authors should spend some effort in discussing the connection of this work to that of Abernethy-Frongilllo (COLT 2012)
@@
 This paper seems to basically present a way of choosing that convex function; this is not uninteresting in itself, but I'm concerned as to why this connection is not made clearer in the paper.
+
 - Highlight key contributions: the authors could have just focused on presenting characterization of the class of new proper scoring rules and behavior and analysis of agents with exponential utility function.
@@
 Also not all exponential family distributions have a mean parameter (which is fine if you choose to model the market price distribution in this way, but this may limit the generality of the framework).
+
 - Please give a formal definition of exponential family distributions (with citation) and examples of commonly studied distributions that can be modeled by the exponential family studied in this paper.
+
 - The importance of the underlying measure v is understated.
@@
 My question is: is it true that all the results and comments can be applied to arbitrary v?
-(I just checked a few but not all since many calculations are already oversimplified.)
-  -  Please give an example of non-Lebesgue v.
+(I just checked a few but not all since many calculations are already oversimplified.) Please give an example of non-Lebesgue v.
 	- Review 3: The reviewer makes a good point, we did not really address alternative base measure, and we will need to add more detail here.
@@
 (5).
-- For example, I found the reference to Bergman divergence a little annoying.
-I also think that maximum entropy principle is overemphasized---proving that the proposed scoring rule is proper is enough to me.
-I also got lost at the last paragraph of Section 1.
+
 - P3: Section 2, after first paragraph. Please clarify that P contains ALL distributions on the outcome space.
+
 - P4: please clarify the setup. Are u always using a single outcome x?, same in the example section at Page 7
+
 - I found the "underlying density" and "full density" discussion on P4 very confusing (what are these two densities?).
+
 - P4, last sentence. Please quote what exactly was proven by [Grunwald and Dawid]
+
 - P5: sufficient statistics - please clarify which distributions are not like this. Can it be that the exponential distribution is not like this?
@@
 The sufficient statistic for r is the maximum of the observations.
+
 - P6, THM 2.2. Where the sufficient statistics condition fits in. Can it be that the entropy maximizing exponential distribution is not like that. Please clarify.
+
 - P7: U are using mu for both the report and expectation. Please use a different notation (e.g. hat(mu)
+
 - Can you model classical prediction markets with finite outcome space under the proposed framework? If so then what distribution in the exponential family its correspond to?
@@
 It was originally covered in the examples of Sections 2.2 and 3.3 but removed due to space constraints.
+
 - Section 5: Please clarify the setup up-front. What is known to the principal on the traders? Does it know $a$ or it can be estimated
 	- In Section 5, the principal (market maker) does not need to know the risk aversion parameter 'a' for Thms 5.2 and 5.3 to hold. The aggregation arises directly from the agents' choice of trades.
+
 - What kind of equilibria are studied in Section 5.2?
@@
 We will expand upon these in the revision.
+
 - Section 6: What exactly does the statement "the traders trade in multiple instances of the market" mean?
@@
 The semantics of the time-varying budget alpha_i^t are not clear.
+
 - Regarding budgets and play money: indeed the play-money markets are a case where budget constraints are key.

### 2014-04-08 WriteUp/response.txt
tagged

--- 2014-04-08 WriteUp/response.txt
@@
-- Reduce complications in notation: Please only keep necessary notation in this paper. 
-- intro revision: we meant this comment for the narrow case of MSR-style markets.
+- Reduce complications in notation: Please only keep necessary notation in this paper. @sebastien
+  - For example, I found the reference to Bergman divergence a little annoying.
+I also think that maximum entropy principle is overemphasized---proving that the proposed scoring rule is proper is enough to me.
+I also got lost at the last paragraph of Section 1.
+
+- intro revision: we meant this comment for the narrow case of MSR-style markets. @jake
   - Ostrovsky and of Wolfers and Zitzewitz are exactly about these issues.
@@
 Much of the work on manipulation is also implicitly focused on exactly this question (including the paper "Gaming Prediction Markets" with one of the authors of this paper as a co-author!)
-- If accepted, the authors should spend some effort in discussing the connection of this work to that of Abernethy-Frongilllo (COLT 2012)
+
+- If accepted, the authors should spend some effort in discussing the connection of this work to that of Abernethy-Frongilllo (COLT 2012) @sebastien 
   - The reviewer makes a really good point regarding our omission of a reference to [Abernethy-Frongillo'12, Thm 11]; this work is certainly related although perhaps not to the extent suggested.
@@
 Also not all exponential family distributions have a mean parameter (which is fine if you choose to model the market price distribution in this way, but this may limit the generality of the framework).
+@sebastien
+
 - Please give a formal definition of exponential family distributions (with citation) and examples of commonly studied distributions that can be modeled by the exponential family studied in this paper.
+@sindhu (normal, exponential, beta)
+
 - The importance of the underlying measure v is understated.
@@
 My question is: is it true that all the results and comments can be applied to arbitrary v?
-(I just checked a few but not all since many calculations are already oversimplified.)
+(I just checked a few but not all since many calculations are already oversimplified.) @jake
   -  Please give an example of non-Lebesgue v.
@@
 (5).
-- For example, I found the reference to Bergman divergence a little annoying.
-I also think that maximum entropy principle is overemphasized---proving that the proposed scoring rule is proper is enough to me.
-I also got lost at the last paragraph of Section 1.
-- P3: Section 2, after first paragraph. Please clarify that P contains ALL distributions on the outcome space.
-- P4: please clarify the setup. Are u always using a single outcome x?, same in the example section at Page 7
-- I found the "underlying density" and "full density" discussion on P4 very confusing (what are these two densities?).
-- P4, last sentence. Please quote what exactly was proven by [Grunwald and Dawid]
-- P5: sufficient statistics - please clarify which distributions are not like this. Can it be that the exponential distribution is not like this?
+
+- P3: Section 2, after first paragraph. Please clarify that P contains ALL distributions on the outcome space. @sindhu
+
+- P4: please clarify the setup. Are u always using a single outcome x?, same in the example section at Page 7 @sindhu
+
+- I found the "underlying density" and "full density" discussion on P4 very confusing (what are these two densities?). @sindhu
+
+- P4, last sentence. Please quote what exactly was proven by [Grunwald and Dawid] @sindhu
+
+- P5: sufficient statistics - please clarify which distributions are not like this. Can it be that the exponential distribution is not like this? @sindhu 
   - Review 1: The concept of a sufficient statistic also arises for distributions other than exponential families.
@@
 The sufficient statistic for r is the maximum of the observations.
-- P6, THM 2.2. Where the sufficient statistics condition fits in. Can it be that the entropy maximizing exponential distribution is not like that. Please clarify.
-- P7: U are using mu for both the report and expectation. Please use a different notation (e.g. hat(mu)
-- Can you model classical prediction markets with finite outcome space under the proposed framework? If so then what distribution in the exponential family its correspond to?
+
+- P6, THM 2.2. Where the sufficient statistics condition fits in. Can it be that the entropy maximizing exponential distribution is not like that. Please clarify. @sindhu
+
+- P7: U are using mu for both the report and expectation. Please use a different notation (e.g. hat(mu) @sindhu
+
+- Can you model classical prediction markets with finite outcome space under the proposed framework? If so then what distribution in the exponential family its correspond to? @sebastien 
   - Classical prediction markets (specifically, LMSR) can be recovered from our framework by using the multinomial distribution, which is an exponential family.
 It was originally covered in the examples of Sections 2.2 and 3.3 but removed due to space constraints.
-- Section 5: Please clarify the setup up-front. What is known to the principal on the traders? Does it know $a$ or it can be estimated
+
+- Section 5: Please clarify the setup up-front. What is known to the principal on the traders? Does it know $a$ or it can be estimated @sebastien 
   - In Section 5, the principal (market maker) does not need to know the risk aversion parameter 'a' for Thms 5.2 and 5.3 to hold. The aggregation arises directly from the agents' choice of trades.
+
 - What kind of equilibria are studied in Section 5.2?
@@
 How equilibria are reached from the initial state of the market?
+@jake
   - we should add a few more details regarding the results of 5.2, but here are some short answers: The utility functions are defined in 5.2, the strategies are the vectors of share purchases of each agent, we are interested in pure-strategy Nash equilibria, and the dynamics are not required to establish the existence of an equilibrium.
 We will expand upon these in the revision.
+
 - Section 6: What exactly does the statement "the traders trade in multiple instances of the market" mean?
@@
 The semantics of the time-varying budget alpha_i^t are not clear.
+@sindhu
+
 - Regarding budgets and play money: indeed the play-money markets are a case where budget constraints are key.
@@
 Risk aversion is also plausible with play-money; for instance, a trader high on the leaderboard may not want to risk her standing with large late trades.
+@sindhu
   - the example for risk-aversion of someone not wanting to lose their place on the leaderboard doesn't really hold water: that's a type of tournament incentive: very different from traditional risk-aversion.

### 2014-04-10 WriteUp/response.txt
edits to the tex file; reordered response file

@@
 
-- Please give a formal definition of exponential family distributions (with citation) and examples of commonly studied distributions that can be modeled by the exponential family studied in this paper.
-@sindhu (normal, exponential, beta)
-
 - The importance of the underlying measure v is understated.
@@
 (5).
-
-- P3: Section 2, after first paragraph. Please clarify that P contains ALL distributions on the outcome space. @sindhu
-
-- P4: please clarify the setup. Are u always using a single outcome x?, same in the example section at Page 7 @sindhu
-
-- I found the "underlying density" and "full density" discussion on P4 very confusing (what are these two densities?). @sindhu
-
-- P4, last sentence. Please quote what exactly was proven by [Grunwald and Dawid] @sindhu
-
-- P5: sufficient statistics - please clarify which distributions are not like this. Can it be that the exponential distribution is not like this? @sindhu 
-  - Review 1: The concept of a sufficient statistic also arises for distributions other than exponential families.
-An example of a non-exponential family with sufficient statistic is the uniform distribution over [0,r], where r is the parameter.
-The sufficient statistic for r is the maximum of the observations.
-
-- P6, THM 2.2. Where the sufficient statistics condition fits in. Can it be that the entropy maximizing exponential distribution is not like that. Please clarify. @sindhu
-
-- P7: U are using mu for both the report and expectation. Please use a different notation (e.g. hat(mu) @sindhu
 
@@
 
-- Section 6: What exactly does the statement "the traders trade in multiple instances of the market" mean?
+- [COMPLETE] change rahul sami's affiliation @sindhu
+- [COMPLETE] remove parskip @sindhu
+- P3: Section 2, after first paragraph. Please clarify that P contains ALL distributions on the outcome space. @sindhu
+  “i think we already say that?"
+- [COMPLETE] P4: please clarify the setup. Are u always using a single outcome x? same in the example section at Page 7 @sindhu 
+  "changed sentence to: We consider a measurable space consisting of a set of (mutually exclusive, exhaustive) outcomes $\mX$ together with a $\sigma$-algebra $\mF$.
+i feel this change takes care of both these points"
+- [COMPLETE] Please give a formal definition of exponential family distributions (with citation) @sindhu 
+  "we didn't have a reference for this; i've added wainwright jordan."
+- [COMPLETE] Section 6: What exactly does the statement "the traders trade in multiple instances of the market" mean?
 What are the formal definitions of an "uninformative" and an "informative" trader?
@@
 @sindhu
-
+changed to:  "We assume that the traders trade in multiple instances of the market.
+That is, after the initial trades, the outcome is revealed and the traders are paid off.
+Then, the market runs again with the same set of traders and the outcome drawn from a (possibly different) exponential family distribution.
+The goal of this section is to analyze how the budget of traders evolve under these circumstances."
+- P7: U are using mu for both the report and expectation. Please use a different notation (e.g. hat(mu) @sindhu 
+- examples of commonly studied distributions that can be modeled by the exponential family studied in this paper. @sindhu (normal, exponential, beta)
+- P5: sufficient statistics - please clarify which distributions are not like this. Can it be that the exponential distribution is not like this? @sindhu
+  - Review 1: The concept of a sufficient statistic also arises for distributions other than exponential families.
+An example of a non-exponential family with sufficient statistic is the uniform distribution over [0,r], where r is the parameter.
+The sufficient statistic for r is the maximum of the observations.
 - Regarding budgets and play money: indeed the play-money markets are a case where budget constraints are key.
@@
   - the example for risk-aversion of someone not wanting to lose their place on the leaderboard doesn't really hold water: that's a type of tournament incentive: very different from traditional risk-aversion.
+- I found the "underlying density" and "full density" discussion on P4 very confusing (what are these two densities?). @sindhu
+- P4, last sentence. Please quote what exactly was proven by [Grunwald and Dawid] @sindhu
+- P6, THM 2.2. Where the sufficient statistics condition fits in. Can it be that the entropy maximizing exponential distribution is not like that. Please clarify. @sindhu

### 2014-04-10 WriteUp/maxentpredmkts.tex
edits to the tex file; reordered response file

@@
 %\usepackage{amsthm}
-\usepackage{parskip}
+%\usepackage{parskip}
 \usepackage{booktabs}
@@
 RAHUL SAMI
-\affil{Google India}}
+\affil{University of Michigan, Ann Arbor}\footnote{Current affiliation: Google, Inc. This work was done while the author was at the University of Michigan.}}
 %%\author{GANG ZHOU
@@
 
-We consider a measurable space consisting of a set of outcomes $\mX$ together with a $\sigma$-algebra $\mF$.
+We consider a measurable space consisting of a set of (mutually exclusive, exhaustive) outcomes $\mX$ together with a $\sigma$-algebra $\mF$.
 An agent or expert has a \emph{belief} over potential outcomes taking the form of a probability measure absolutely continuous with respect to a base measure~$\nu$.\footnote{Recall that a measure $P$ is absolutely continuous with respect to~$\nu$ if $P(A) = 0$ for every $A \in \mF$ for which $\nu(A) = 0$.
@@
 
-It is well-known that solutions to~(\ref{maxent-prog}) are \emph{exponential family} distributions, whose densities with respect to $\nu$ take the form
+It is well-known that solutions to~(\ref{maxent-prog}) are \emph{exponential family} distributions (see, for instance, \cite{WainJordan08}).
+These are distributions whose densities with respect to $\nu$ take the form
 %
@@
 We assume that the traders trade in multiple instances of the market.
+That is, after the initial trades, the outcome is revealed and the traders are paid off.
+Then, the market runs again with the same set of traders and the outcome drawn from a (possibly different) exponential family distribution.
+The goal of this section is to analyze how the budget of traders evolve under these circumstances.
 As before, the market price is interpreted as a probability density over the outcome space and the share vector as the natural parameter of an exponential family distribution.

### 2014-04-11 WriteUp/maxentpredmkts.tex
more changes including comments in response file

@@
 \newcommand{\hmu}{\ensuremath{\hat{\mu}}}
+\newcommand{\hsigma}{\ensuremath{\hat{\sigma}}}
+\newcommand{\hSigma}{\ensuremath{\hat{\Sigma}}}
 \newcommand{\mP}{\ensuremath{\mathcal{P}}}
@@
 
+ A note about play-money markets: Even in prediction markets set up with play-money, risk aversion is plausible since it captures the fact that a trader high on the leaderboard may not want to risk her standing with large late trades.
+Further, in such markets especially, budget constraints are key since they help alleviate the problem that traders can easily manipulate prices with unreasonable trades.
+Our results are therefore particularly relevant in markets with play-money.
 % \jake{I want to add more to this list. It's ok but doesn't totally reflect what's in the last few sections.}
@@
 The family is \emph{minimal} if there is no $\alpha \in \Theta$ such that $\la \alpha, \phi(x) \ra$ is a constant over $\mX$ ($\nu$-almost everywhere); minimality is a property of the associated statistic $\phi$, usually called the \emph{sufficient statistic} in the literature.
+
+Some commonly studied distributions that are in fact exponential families include the binomial, beta and normal distributions.
+To see for instance, that the normal distribution is an exponential family note that instead of the usual parametrization using mean  $\mu$ and  variance $\sigma$, it can be parametrized in terms of $\vec{\theta}=(\theta_{1},\theta_{2})$ where $\theta_{1}=\frac{\mu}{\sigma^{2}}$ and $\theta_{2}=\frac{-1}{2\sigma^{2}}$,  with corresponding log partition function $-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$.
 
@@
 where $\htheta = \grad G(\hmu)$ is the natural parameter for the exponential family with statistic $\hmu$.
-We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, which is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
+Note that the expected log score does not depend  on the entire maximum entropy distribution but just its expected statistic $\mu$.
+This is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
+%We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, which is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
+%
 %
@@
 Consider the logarithmic scoring rule $S(\mu,x) = \log p(x;\mu)$ defined over a set of densities $\mD$ parametrized by $\mM$.
-The scoring rule is proper if and only if $\mD$ is the exponential family with statistic $\phi$.
+The scoring rule is proper if and only if $\mD$ is the exponential family with statistic $\phi$, where $\mu=\E_{p(x;\mu)}[\phi(x)]$.
 \end{theorem}
@@
 The following examples illustrate the construction.
+Here, we have used $\mu$ and $\sigma$ for the true mean and variance and $\hmu$ and $\hsigma$ for the elicited mean and variance respectively (and $\hSigma$ for the variance of the multivariate normal).
 %
@@
 \begin{equation} \label{exp-scoring}
-S(\mu, x) = -\frac{x}{\mu} - \log \mu.
+S(\hmu, x) = -\frac{x}{\hmu} - \log \hmu.
 \end{equation}
@@
 \begin{equation} \label{weibull-scoring}
-S(\mu, x) = (k-1) \log x - k \log \mu - \Gamma\left(1+\frac{1}{k}\right)^k \left(\frac{x}{\mu}\right)^k,
+S(\hmu, x) = (k-1) \log x - k \log \hmu - \Gamma\left(1+\frac{1}{k}\right)^k \left(\frac{x}{\hmu}\right)^k,
 \end{equation}
@@
 \begin{equation}
-S((\mu, \sigma^2), x) = -\frac{(x-\mu)^2}{\sigma^2} - \log \sigma^2.
+S((\hmu, \hsigma^2), x) = -\frac{(x-\hmu)^2}{\hsigma^2} - \log \hsigma^2.
 \end{equation}
@@
 \begin{equation}
-S((\mu, \Sigma), x) = -(x - \mu)'\Sigma^{-1}(x - \mu) - \log |\Sigma|.
+S((\hmu, \hSigma), x) = -(x - \hmu)'\hSigma^{-1}(x - \hmu) - \log |\hSigma|.
 \end{equation}
@@
 We assume that the traders trade in multiple instances of the market.
-That is, after the initial trades, the outcome is revealed and the traders are paid off.
+That is, after the initial setup and trades, the outcome is revealed and the traders are paid off.
 Then, the market runs again with the same set of traders and the outcome drawn from a (possibly different) exponential family distribution.
@@
 
+Note that while an uninformative trader is unconstrained in his belief, the informative trader's belief corresponds to the parameter of the true distribution.
 \paragraph{Limited Damage}

### 2014-04-11 WriteUp/response.txt

--- 2014-04-08 WriteUp/response.txt
@@
 
-- Section 6: What exactly does the statement "the traders trade in multiple instances of the market" mean?
-What are the formal definitions of an "uninformative" and an "informative" trader?
-The semantics of the time-varying budget alpha_i^t are not clear.
-@sindhu
-
-- Regarding budgets and play money: indeed the play-money markets are a case where budget constraints are key.
+- [COMPLETE] change rahul sami's affiliation
+- [COMPLETE] remove parskip
+- [COMPLETE] P4: please clarify the setup. Are u always using a single outcome x? same in the example section at Page 7 @sindhu 
+  "changed sentence to: We consider a measurable space consisting of a set of (mutually exclusive, exhaustive) outcomes $\mX$ together with a $\sigma$-algebra $\mF$.
+i feel this change takes care of both these points"
+- [COMPLETE] Please give a formal definition of exponential family distributions (with citation) @sindhu 
+  "we didn't have a reference for this; i've added wainwright jordan."
+- [COMPLETE] P7: U are using mu for both the report and expectation. Please use a different notation (e.g. hat(mu) @sindhu 
+  "i think the reviewer means that in the definition of the scoring rule, we should indicate that the first argument is the agent report and not the true mean. so i've changed that."
+- [COMPLETE] Section 6: 
+  "Changed first paragraph of this section and added a sentence to the "Budget-limited trades" paragraph"
+  - [COMPLETE] What exactly does the statement "the traders trade in multiple instances of the market" mean? 
+  - [COMPLETE] The semantics of the time-varying budget alpha_i^t are not clear. @sindhu
+  - [COMPLETE] What are the formal definitions of an "uninformative" and an "informative" trader? 
+- [COMPLETE] I found the "underlying density" and "full density" discussion on P4 very confusing (what are these two densities?). @sindhu
+  "I think he means p.6 just before thm 2.2; i've reworded this sentence."
+- [COMPLETE] Regarding budgets and play money: indeed the play-money markets are a case where budget constraints are key.
 We have experience with this in other work where a play-money market was released for the US Elections.
@@
 @sindhu
+  "added a note to the introduction."
   - the example for risk-aversion of someone not wanting to lose their place on the leaderboard doesn't really hold water: that's a type of tournament incentive: very different from traditional risk-aversion.
+- [COMPLETE] P6, THM 2.2. Where the sufficient statistics condition fits in. Can it be that the entropy maximizing exponential distribution is not like that. Please clarify. @sindhu
+  "changed statement of thm 2.2 to state the relationship between \mu and \phi explicitly"
+- [COMPLETE] examples of commonly studied distributions that can be modeled by the exponential family studied in this paper. @sindhu 
+  " i added examples before prop 2.1 (normal, binomial and beta) and added a detailed mapping for the normal distribution"
+- P3: Section 2, after first paragraph. Please clarify that P contains ALL distributions on the outcome space. @sindhu
+  "already says that?"
+- P4, last sentence. Please quote what exactly was proven by [Grunwald and Dawid] @sindhu
+  "i'm not sure about this one"

### 2014-04-11 WriteUp/maxentpredmkts.tex

@@
 %
-%
 \begin{theorem} \label{maxent-score}
@@
 %
-
 As further intuition for the result, note that~(\ref{breg-div}) is the definition of the `Bregman divergence' with respect to strictly convex function $T$, written $D_T$. Therefore we have 
@@
 %
-
 In the above, Example~\ref{ex:gaussian} illustrates an important point about parametrizations of the elicited expectations.
@@
 
-Let $q_{0}$ be the initial and $q_{f}$ the final market state.
-Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
-\begin{eqnarray*}
-\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
-&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
-&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
-&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
-&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
-&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
-&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
-&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
-&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
-\end{eqnarray*}
-
 \subsection{Example: Gaussian Markets}
@@
 
+\subsection{An Information-Theoretic Interpretation}
+\label{sec:infth}
+
+ We observe a useful alternative view of the market 
+scoring rule prediction market for exponential family learning. We connect the cost, payoff and profit function to information-theoretic 
+quantities associated with the exponential family.
+
+The following result has been previously pointed out by Amari \cite{Amari-KL}. 
+\begin{theorem}\label{lem:profit_decomposition}{\bf (Profit Decomposition)}: Consider an exponential 
+family $\family$ of distributions over some set of statistics $\suff(x)$, with natural parameters $\beta$. Let $\pi, \rho \in \family$ be any two probability distributions in the family. We use 
+$\beta_\pi$ to denote the natural parameters of $\pi$ and $\mu_{\pi}$ to denote the expected value of the sufficient statistics under $\pi$. Likewise,
+we can define $\beta_\rho$ and  $\mu_\rho$. Let $\lpf(\beta_{\rho})$ indicate the log partition function of $\rho$.
+Let $H(\pi)$ denote the entropy of the distribution $\pi$, and 
+ $K(\pi||\rho)$ denote the KL-divergence of $\rho$ relative to $\pi$.
+Then, the following equality holds:  
+\begin{equation}
+ K(\pi||\rho) + H(\pi) = \lpf(\beta_{\rho}) - \beta_\rho \cdot \mu_\pi
+\label{eq:profit_decomposition}
+\end{equation}
+\end{theorem}
+\begin{proof}
+\begin{eqnarray*}
+ K(\pi||\rho) + H(\pi) &=& \int_x \pi(x)\log{\frac{\pi(x)}{\rho(x)}}dx\\
+ &&\quad - \int_x \pi(x)\log{\pi(x)}dx \\
+&=& - \int_x \pi(x)\log{\rho(x)}dx \\
+&=& - \int_x \pi(x)\left[ \beta_\rho \cdot \suff(x) - \lpf(\beta_{\rho})\right]dx\\
+&=& - \beta_\rho \cdot \int_x \pi(x) \suff(x) dx\\
+&&\quad + \lpf(\beta_{\rho}) \int_x \pi(x) dx \\
+&=&  \lpf(\beta_{\rho}) - \beta_\rho \cdot \mu_\pi
+\end{eqnarray*}
+\end{proof}
+
+Equation~\ref{eq:profit_decomposition} gives us an alternative view of the
+market scoring rule construction. Assume that $\pi$ is the true distribution, and consider two arbitrary
+distributions $\rho_1, \rho_2 \in \family$.
+Note that $\lpf(\rho)$  is independent of $\pi$, and $\betavec_\rho \cdot \muvec_\pi$ is linear in the probabilities $\pi(x)$.
+If we want to measure loss by the KL-divergence, we can do so (in expectation) by setting a cost function that captures the first term, and defining security quantities ($\betavec$) and payoffs ($\suff$) to capture the second term.
+In particular, in a market with cost function $\lpf$, if the
+market price initially implies a distribution $\rho_1$, and a trader moves
+the market to price than implies a distribution $\rho_2$, then the cost she
+incurs is $\lpf(\rho_2) - \lpf(\rho_1)$.
+The number of securities bought to make this trade is given by the vector ($\betavec_{\rho_2} - \betavec_{\rho_1}$), and the expected payoff of the securities are given by $\muvec_\pi$.
+Thus, by equation \ref{eq:profit_decomposition}, the {\em net profit} of the trader is equal to $K(\pi||\rho_1) - K(\pi||\rho_2)$, {\it i.e.}, the reduction in KL-divergence with respect to the true distribution.
+We note one useful property of this construction: For a fixed vector $\betavec$ of purchased securities, the cost is independent of the outcome
+(and outcome distribution $\pi$), while the payoff is independent of the initial market state in which these securities were purchased.
+
+\subsection{Properties of the Maximum Entropy Market}
+
+{\thm Negative differential entropy $-H(p)\defeq \int_{x}p(x) \log p(x) dx$ is unbounded from above for an exponential family distribution.}
+\proof
+Note that, for an exponential family distribution
+\begin{eqnarray*}
+H(p) &=& -\int_{x}p(x)\log p(x)dx\\
+&=& -\int_{x}p(x)[\beta\cdot \phi(x)-\psi(\beta)] dx\\
+&=& -\beta\cdot\E[ \phi(x)]+\psi(\beta)
+\end{eqnarray*}
+If the range of $\beta$ is unbounded, the negative differential entropy is unbounded as well.
+\unproof
+
+%{\cor The exponential family market maker has unbounded worst case loss.}
+%\proof From \cite{ACV13} we know that the worst case market maker loss can be expressed in terms of the conjugate of the cost function as 
+%$$\sup_{\mu\in\phi(\mathcal{X})}C^{*}(\mu) - \min_{\mu\in\conv(\phi(\mathcal{X}))}C^{*}(\mu)$$ 
+%Unbounded differential entropy for an exponential family distribution thus implies unbounded market maker loss.\unproof
+\paragraph{Loss of the market maker}
+First we derive an expression for the conjugate dual $C^{*}$ in terms of the primal variables.
+
+Recall the definition of the conjugate dual:
+$$C^{*}(\mu) = \sup_{q}q\cdot\mu - C(q)$$
+The supremum is achieved at $q'$ such that $\nabla C(q')=\mu$. So we may rewrite:
+$$C^{*}(\nabla C(q')) = \sup_{q}q\cdot\nabla C(q') - C(q)$$
+This supremum is achieved at $q$ such that $\nabla C(q')=\nabla C(q)$. One such value of $q$ is $q'$.
+So we have 
+$$C^{*}(\nabla C(q')) = q'\cdot\nabla C(q') - C(q')$$
+Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
+
+Let $q_{0}$ be the initial and $q_{f}$ the final market state.
+Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
+\begin{eqnarray*}
+\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
+&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
+&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
+&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
+&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
+&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
+&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
+&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
+&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
+\end{eqnarray*}
+
 Thus, the expected payoff is at least $\frac{1}{a} D_{T}(\theta'\lambda,\beta)$.

### 2014-04-17 WriteUp/maxentpredmkts.tex
fixed a small typo

@@
 RAHUL SAMI
-\affil{University of Michigan, Ann Arbor}\footnote{Current affiliation: Google, Inc. This work was done while the author was at the University of Michigan.}}
+\affil{University of Michigan, Ann Arbor\footnote{Current affiliation: Google, Inc. This work was done while the author was at the University of Michigan.}}}
 %%\author{GANG ZHOU

### 2014-04-21 WriteUp/response.txt
addressed my reviewer comments

--- 2014-04-08 WriteUp/response.txt
@@
 
-- If accepted, the authors should spend some effort in discussing the connection of this work to that of Abernethy-Frongilllo (COLT 2012) @sebastien 
-  - The reviewer makes a really good point regarding our omission of a reference to [Abernethy-Frongillo'12, Thm 11]; this work is certainly related although perhaps not to the extent suggested.
-It is correct that [AF'12, Thm 11] is the right result to use in the 'only if' part of Thm 2.2 (we will definitely include this), which follows up on prior work of [Savage'71] and [Gneiting-Raftery'07].
-However, Thm 2.2 is not an instance of or equivalent to any result in [AF'12]; the construction in Thm 2.2 is to apply a scoring rule to a parametrized family of densities.
-One finds a "family of densities" constructed in the proof of [AF'12, Thm 11], but this does not fall within the exponential family setting.
-    - That paper shows how to construct a proper scoring rule for linear properties from any convex function using Bregman divergence.
-This paper seems to basically present a way of choosing that convex function; this is not uninteresting in itself, but I'm concerned as to why this connection is not made clearer in the paper.
-
-- Highlight key contributions: the authors could have just focused on presenting characterization of the class of new proper scoring rules and behavior and analysis of agents with exponential utility function.
-The connection between exponential family and the proposed prediction market setting is interesting and helpful if it is briefly mentioned and discussed in the end of the paper.
-The paper in its current form reads like suggesting that using exponential families is justified at the modeling level, which sounds a little weak (and in fact not necessary).
-So I dont think they should be mentioned in the beginning.
-For example, I got confused by "outstanding shares".
-Also not all exponential family distributions have a mean parameter (which is fine if you choose to model the market price distribution in this way, but this may limit the generality of the framework).
-@sebastien
-
-- Please give a formal definition of exponential family distributions (with citation) and examples of commonly studied distributions that can be modeled by the exponential family studied in this paper.
-@sindhu (normal, exponential, beta)
-
-- The importance of the underlying measure v is understated.
-If v is always the Lebesgue measure then many distributions cannot be modeled by the exponential family as presented in the paper, e.g.
-normal distributions with fixed variance.
-Many discussions and proofs are made without reference to v.
-My question is: is it true that all the results and comments can be applied to arbitrary v?
-(I just checked a few but not all since many calculations are already oversimplified.) @jake
-  -  Please give an example of non-Lebesgue v.
-  - Review 3: The reviewer makes a good point, we did not really address alternative base measure, and we will need to add more detail here.
-However, all results hold for any arbitrary base measure v.
-Often v does not appear because we work directly with the expectation mu and do not write out the implicit integral.
-All integrals are taken with respect to dv, e.g.
-(5).
-
-- P3: Section 2, after first paragraph. Please clarify that P contains ALL distributions on the outcome space. @sindhu
-
-- P4: please clarify the setup. Are u always using a single outcome x?, same in the example section at Page 7 @sindhu
-
-- I found the "underlying density" and "full density" discussion on P4 very confusing (what are these two densities?). @sindhu
-
-- P4, last sentence. Please quote what exactly was proven by [Grunwald and Dawid] @sindhu
-
-- P5: sufficient statistics - please clarify which distributions are not like this. Can it be that the exponential distribution is not like this? @sindhu 
-  - Review 1: The concept of a sufficient statistic also arises for distributions other than exponential families.
-An example of a non-exponential family with sufficient statistic is the uniform distribution over [0,r], where r is the parameter.
-The sufficient statistic for r is the maximum of the observations.
-
-- P6, THM 2.2. Where the sufficient statistics condition fits in. Can it be that the entropy maximizing exponential distribution is not like that. Please clarify. @sindhu
-
-- P7: U are using mu for both the report and expectation. Please use a different notation (e.g. hat(mu) @sindhu
-
-- Can you model classical prediction markets with finite outcome space under the proposed framework? If so then what distribution in the exponential family its correspond to? @sebastien 
-  - Classical prediction markets (specifically, LMSR) can be recovered from our framework by using the multinomial distribution, which is an exponential family.
-It was originally covered in the examples of Sections 2.2 and 3.3 but removed due to space constraints.
-
-- Section 5: Please clarify the setup up-front. What is known to the principal on the traders? Does it know $a$ or it can be estimated @sebastien 
-  - In Section 5, the principal (market maker) does not need to know the risk aversion parameter 'a' for Thms 5.2 and 5.3 to hold. The aggregation arises directly from the agents' choice of trades.
-
 - What kind of equilibria are studied in Section 5.2?
@@
 @sindhu
+  "added a note to the introduction."
   - the example for risk-aversion of someone not wanting to lose their place on the leaderboard doesn't really hold water: that's a type of tournament incentive: very different from traditional risk-aversion.
+- [COMPLETE] P6, THM 2.2. Where the sufficient statistics condition fits in. Can it be that the entropy maximizing exponential distribution is not like that. Please clarify. @sindhu
+  "changed statement of thm 2.2 to state the relationship between \mu and \phi explicitly"
+- [COMPLETE] examples of commonly studied distributions that can be modeled by the exponential family studied in this paper. @sindhu 
+  " i added examples before prop 2.1 (normal, binomial and beta) and added a detailed mapping for the normal distribution"
+- P3: Section 2, after first paragraph. Please clarify that P contains ALL distributions on the outcome space. @sindhu
+  "already says that?"
+- P4, last sentence. Please quote what exactly was proven by [Grunwald and Dawid] @sindhu
+  "i'm not sure about this one"
+[COMPLETE]- The importance of the underlying measure v is understated.
+If v is always the Lebesgue measure then many distributions cannot be modeled by the exponential family as presented in the paper, e.g.
+normal distributions with fixed variance.
+Many discussions and proofs are made without reference to v.
+My question is: is it true that all the results and comments can be applied to arbitrary v?
+(I just checked a few but not all since many calculations are already oversimplified.) @sebastien: clarified that base measure is implicit in integrals.
+[COMPLETE] -  Please give an example of non-Lebesgue v: explained the case of fixed variance, unknown mean in a new footnote.
+[COMPLETE]- Can you model classical prediction markets with finite outcome space under the proposed framework?
+If so then what distribution in the exponential family its correspond to?
+@sebastien: added examples of scoring rules and prediction markets with finite outcome space, i.e., the usual log score and LMSR.
+[COMPLETE] - Highlight key contributions: the authors could have just focused on presenting characterization of the class of new proper scoring rules and behavior and analysis of agents with exponential utility function.
+The connection between exponential family and the proposed prediction market setting is interesting and helpful if it is briefly mentioned and discussed in the end of the paper.
+The paper in its current form reads like suggesting that using exponential families is justified at the modeling level, which sounds a little weak (and in fact not necessary).
+So I dont think they should be mentioned in the beginning.
+For example, I got confused by "outstanding shares".
+Also not all exponential family distributions have a mean parameter (which is fine if you choose to model the market price distribution in this way, but this may limit the generality of the framework).
+@sebastien: added sentence explaining that we aren't actually modeling agent beliefs as exponential families in the scoring rule section.
+[COMPLETE] - Section 5: Please clarify the setup up-front.
+What is known to the principal on the traders?
+Does it know $a$ or it can be estimated @sebastien: added sentence in Section 5 explaining market maker does not need to know a for results to hold.
+[COMPLETE] - If accepted, the authors should spend some effort in discussing the connection of this work to that of Abernethy-Frongilllo (COLT 2012) 
+- That paper shows how to construct a proper scoring rule for linear properties from any convex function using Bregman divergence.
+This paper seems to basically present a way of choosing that convex function; this is not uninteresting in itself, but I'm concerned as to why this connection is not made clearer in the paper.
+@sebastien: added discussion of all this right after Thm 2.2, and also used the right result from Aberneth-Frongillo'12 in the proof of Thm 2.2.

### 2014-04-21 WriteUp/maxentpredmkts.tex
addressed my reviewer comments

--- 2014-04-11 WriteUp/maxentpredmkts.tex
@@
 We draw connections between the information aggregation of market prices and the belief aggregation of learning agents that rely on exponential family distributions.
-We develop a very natural analysis of the market behavior as well as the price equilibrium under the assumption that the traders exhibit risk aversion according to exponential utility.
-We also consider similar aspects under alternative models, such as when traders are budget constrained.
+We develop a natural analysis of the market behavior as well as the price equilibrium under the assumption that the traders exhibit risk aversion according to exponential utility.
+We also consider similar aspects under alternative models, such as budget-constrained traders.
 %
@@
 
-\terms{Algorithms, Economics}
-
-\keywords{logarithmic score, exponential family, maximum entropy, risk aversion, budget constraints}
+%\terms{Algorithms, Economics}
+
+\keywords{logarithmic score; exponential family; maximum entropy; risk aversion; budget constraints}
 
@@
 This will affect the state of the market, thus updating the predictive probabilities for the event.
-Further, since the trades are done sequentially, the trader is allowed to observe all past trades in the market and update his private information based on this information.
+Further, since the trades are done sequentially, the trader may observe all past trades in the market and update his private beliefs based on this information.
 In this sense the market prices, which are in effect the prices at which the marginal trader is willing to buy or sell the available securities, can be interpreted as an aggregate ``consensus probability forecast'' of the event in question.
@@
 In the present work we show that a number of classical statistical tools can be leveraged to design a prediction market framework in the mold of \emph{exponential family distributions}; we show that this statistical framework leads to a number of attractive properties and interpretations.
-Common concepts in statistics---including \emph{entropy maximization}, \emph{log loss}, and \emph{bayesian inference}---relate to natural aspects of our class of mechanisms.
+Common concepts in statistics---including \emph{entropy maximization}, \emph{log loss}, and \emph{Bayesian inference}---relate to natural aspects of our class of mechanisms.
 In particular, the central objects in our market framework can be interpreted via concepts used to define exponential families:
@@
 	\item the market's \emph{payoff function} corresponds to the \emph{sufficient statistics} of the distribution;
-	\item the vector of \emph{outstanding shares} in the market corresponds to the \emph{natural parameter vector} of the distribution;
+	\item the vector of \emph{outstanding shares} in the market corresponds to the \emph{natural parameter} of the distribution;
 	\item the market prices correspond to \emph{mean parameters};
@@
 	%We show how this dovetails nicely with the notion of a \emph{certainty equivalent} which is often used in risk-aversion settings. 
-	In this case we can characterize precisely how a single trader interacts with the market, as well as the equilibrium reached given multiple traders; this result is achieved via a potential game argument.
+	In this case we can characterize precisely how a single trader interacts with the market, as well as the equilibrium reached given multiple traders; the latter result is achieved via a potential game argument.
 The eventual market state is a weighted combination of traders' beliefs and the initial state; the weights are proportional to risk aversion parameters.
@@
 Scoring rules are a measure of prediction accuracy, and we are concerned here with scoring rules for statistic expectations, typically over infinite outcome spaces.
-Such rules have been characterized by~\citet{Savage71}; see also~\citep{frongillo2013elicitingmeans,lambert2008eliciting}.
+Such rules have been characterized by~\citet{Savage71}; see also~\citep{AbernethyFr12,gneiting2007strictly,lambert2008eliciting}.
 Our rules are of course special cases of this characterization, but it appears the range of elegant scoring rules that arise from exponential families has not been appreciated.
@@
 Let $$\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}$$ be the set of realizable statistic expectations.
+Unless otherwise mentioned, all integrals including expectations are taken with respect to base measure $\nu$, which is therefore implicit in many of the formulas.
 A scoring rule $S : \mM \times \mX \rightarrow \bR \cup \{-\infty\}$ pays the agent $S(\hmu,x)$ according to how well its report $\hmu \in \mM$ agrees with the eventual outcome $x \in \mX$.
@@
 The function $T$ is known as the \emph{log-partition} or \emph{cumulant} function corresponding to the exponential family.
-Its domain is $\Theta = \{ \theta \in \bR^d : T(\theta) < +\infty \}$, called the natural parameter space.
+Its domain is $\Theta = \{ \theta \in \bR^d : T(\theta) < +\infty \}$, called the natural parameter space, and we assume throughout that it is nonempty.\footnote{This may impose additional conditions on the choice of base measure $\nu$ in certain special cases.
+For instance, if the density over $\bR$ is known to have unit variance and statistic $\phi(x) = x$, then one uses the base measure $\nu(x) = e^{-x^2/2}$ leading to a normal distribution with fixed variance~\citep{Barndorff78}.}
+%
 The exponential family is \emph{regular} if $\Theta$ is open---almost all exponential families of interest, and all those we consider in this work, are regular.
@@
 For the converse, assume the defined log score is proper.
-By the Savage characterization of proper scoring rules for expectations (see~\citet[Thm.\ 1]{gneiting2007strictly} and~\citet{Savage71}), we must have
+By the Savage characterization of proper scoring rules for expectations (see~\citet[Thm.\ 11]{AbernethyFr12}), we must have
 %
@@
 %
-
-As further intuition for the result, note that~(\ref{breg-div}) is the definition of the `Bregman divergence' with respect to strictly convex function $T$, written $D_T$. Therefore we have 
+It has been understood since~\citep{mccarthy1956measures,Savage71} that scoring rules for densities and expectations can be characterized in terms of an underlying convex function.
+The generalized log scoring rule of Theorem~\ref{maxent-score} obtains by applying the modern characterization of~\citet{AbernethyFr12} to the optimal value function $G$ (negative maximum entropy).
+\citet[Thm.\ 6]{BanerjeeDhGh05} prove a similar result by showing a bijection between exponential families and Bregman divergences.
+
+The notion of a Bregman divergence in fact provides further intuition for the result.
+Note that~(\ref{breg-div}) is the definition of the Bregman divergence with respect to strictly convex function $T$, written $D_T$.
+Therefore we have
 %
@@
 The scoring rule compensates the agent according to the log likelihood of the eventual outcome according to this distribution.
-The interpretation is that the agent is only providing partial information about the underlying density, so the principal first infers a full density according to the principle of maximum entropy, and then scores the agent using the usual log score.
+The interpretation is that the agent is only providing partial information about the underlying density, so the principal first infers a full density according to the principle of maximum entropy, and then scores the agent using the standard log score.
+We stress that the exponential families framework is a tool in the construction of scoring rules; we are not actually modeling any agent's belief via an exponential family.
 
@@
 Here, we have used $\mu$ and $\sigma$ for the true mean and variance and $\hmu$ and $\hsigma$ for the elicited mean and variance respectively (and $\hSigma$ for the variance of the multivariate normal).
+
+%
+\begin{example}
+As a base measure we take the counting measure over the finite set $\{1,2,\ldots,d\}$, and we consider the statistic $\phi(i) \in \{0,1\}^d$ that maps outcome $i$ to the unit vector with component $i$ equal to 1.
+The expectation of $\phi(i)$ is simply the entire probability distribution $p$, which is a multinomial distribution (an instance of an exponential family).
+We recover the standard log scoring rule $S(\hat{p}, i) = \log \hat{p}(i)$.
+\end{example}
+%
+
 %
@@
 The relevant entities now are the payoff function, the effective domain of shares, and the cost function.
+%
+\begin{example}
+We consider a finite outcome space $\mX = \{1,2,\ldots,d\}$ and set up a market with an Arrow-Debreu security $\phi_i$ for every outcome $i$: the security pays off 1 if $i$ occurs and 0 otherwise.
+The log-partition function recovers the classical LMSR cost function~\citep{H07}:
+$C(\theta) = \log \sum_{i \in \mX} e^{\theta_i}.$
+The effective domain is $\Theta = \bR^d$, meaning that a security can always be bought or sold short.
+\end{example}
+%
+
 %
@@
 %
-
+We stress that the aggregation in Theorem~\ref{thm:exp-util} arises directly from the agents' choice of trades.
+The market maker does not need to know the risk aversion parameters.
 Note that as, $a$ tends to 0, we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$.
@@
 Also, recall that $\nabla C^{*}(\nabla C(q))=q$ for the exponential family market. (I think there are some restrictions on $q$.)
-
-Let $q_{0}$ be the initial and $q_{f}$ the final market state.
-Then if $\mu$ is the expected value of the outcome sufficient statistics (\ie  the mean parameter) under the true distribution, the loss of the exponential family market maker can be written as:
-\begin{eqnarray*}
-\phi(x)(q_{f}-q_{0})-C(q_{f})+C(q_{0})&=& \phi(x) (q_{f}-q_{0})-q_{f}\nabla C(q_{f})+ C^{*}(\nabla C(q_{f}))\\
-&&+q_{0}\nabla C(q_{0})- C^{*}(\nabla C(q_{0}))\\
-&=&q_{f}(\phi(x) -\nabla C(q_{f}))+ C^{*}(\nabla C(q_{f}))\\
-&&-q_{0}(\phi(x) -\nabla C(q_{0}))- C^{*}(\nabla C(q_{0}))\\
-&=&-C^{*}(\nabla C(q_{0}))+ C^{*}(\phi(x))-q_{0}(\phi(x) -\nabla C(q_{0}))\\
-&&+C^{*}(\nabla C(q_{f}))- C^{*}(\phi(x))+q_{f}(\phi(x) -\nabla C(q_{f}))\\
-&=&  C^{*}(\phi(x))-C^{*}(\nabla C(q_{0}))-\nabla C^{*}(\nabla C(q_{0}))(\phi(x) -\nabla C(q_{0}))\\
-&&-\left[C^{*}(\phi(x))-C^{*}(\nabla C(q_{f}))-\nabla C^{*}(\nabla C(q_{f}))(\phi(x) -\nabla C(q_{f}))\right]\\
-&=&D_{C^{*}}[\phi(x),\nabla C(q_{0}))]-D_{C^{*}}[\phi(x),\nabla C(q_{f}))]
-\end{eqnarray*}
 

### 2014-04-21 WriteUp/EC14-template/EC14-sample.tex
added folder with EC templates

%\documentclass[prodmode]{acmsmall-ec14}
%\documentclass[prodmode,license]{acmsmall-ec14}
\documentclass[prodmode,permissions]{acmsmall-ec14}

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

%\conferenceinfo{EC'14,}{June 8--12, 2014, Stanford, CA, USA.}
%\CopyrightYear{2014}
%\crdata{978-1-4503-2565-3/12/06}
%\copyrighttext{Copyright is held by the owner/author(s). Publication rights licensed to ACM.}

\doi{2600057.2600058}% TeXSupport

% Document starts
\begin{document}

% Page heads
%\markboth{G. Zhou et al.}{A Multifrequency MAC Specially Designed for WSN Applications}

% Title portion
\title{An EC’14 Sample File Only, \underline{N}ote \underline{I}nitial \underline{C}aps to be Used  for the Titles} 
\author{GANG ZHOU
\affil{College of William and Mary}
YAFENG WU
\affil{University of Virginia}
TING YAN
\affil{Eaton Innovation Center}
TIAN HE
\affil{University of Minnesota}
CHENGDU HUANG
\affil{Google}
JOHN A. STANKOVIC
\affil{University of Virginia}
TAREK F. ABDELZAHER
\affil{University of Illinois at Urbana-Champaign}}
% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{abstract}
Multifrequency media access control has been well understood in
general wireless ad hoc networks, while in wireless sensor networks,
researchers still focus on single frequency solutions. In wireless
sensor networks, each device is typically equipped with a single
radio transceiver and applications adopt much smaller packet sizes
compared to those in general wireless ad hoc networks. Hence, the
multifrequency MAC protocols proposed for general wireless ad hoc
networks are not suitable for wireless sensor network applications,
which we further demonstrate through our simulation experiments. In
this article, we propose MMSN, which takes advantage of
multifrequency availability while, at the same time, takes into
consideration the restrictions of wireless sensor networks. Through
extensive experiments, MMSN exhibits the prominent ability to utilize
parallel transmissions among neighboring nodes. When multiple physical
frequencies are available, it also achieves increased energy
efficiency, demonstrating the ability to work against radio
interference and the tolerance to a wide range of measured time
synchronization errors.
\end{abstract}

\category{C.2.2}{Computer-Communication Networks}{Network Protocols}

%\terms{Design, Algorithms, Performance}

\keywords{Wireless sensor networks, media access control, multi-channel, radio interference, time synchronization}

\begin{bottomstuff}
This work is supported by the National Science Foundation, undergrant
CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609. See the
Acknowledgements section before REFERENCES.

Author’s addresses: G. Zhou, Computer Science Department, College of
William \& Mary; email: \url{g.zhou@sample.com}; Y. Wu and J.A. Stankovic, CS  Dept, Univ. of
Virginia; email: \{y.wu,ja.stankovic\}@sample.com; T. Yan, Eaton Innovation Center; email: t.yam@sample.com; T. He, Computer Science
Dept., Univ. of Minnesota; email: t.he@sample.com; C. Huang, Google; email: c.huang@google.com; T.F. Abdelzaher, CS
Dept, Univ. of Illinois at Urbana-Champaign; email: tf.abdelzaher@sample.com.
\end{bottomstuff}

\maketitle

\section{Introduction}

As a new technology, Wireless Sensor Networks (WSNs) has a wide
range of applications [Culler 2001,Bahl 2002,Akyildiz 2001], including
environment monitoring, smart buildings, medical care, industrial and
military applications. Among them, a recent trend is to develop
commercial sensor networks that require pervasive sensing of both
environment and human beings, for example, assisted living
[Akyildiz 2002,Harvard 2001,CROSSBOW] and smart homes
[Harvard 2001,Adya 2001,CROSSBOW].
% quote
\begin{quote}
``For these applications, sensor devices are incorporated into human
cloths [Natarajan 2001,Zhou 2006,Bahl 2002,Adya 2001] for monitoring
health related information like EKG readings, fall detection, and voice recognition".
\end{quote}
While collecting all these multimedia information
[Akyildiz 2002] requires a high network throughput, off-the-shelf
sensor devices only provide very limited bandwidth in a single
channel: 19.2Kbps in MICA2 [Bahl 2002] and 250Kbps in MICAz.

In this article, we propose MMSN, abbreviation for Multifrequency
Media access control for wireless Sensor Networks. The main
contributions of this work can be summarized as follows.
% itemize
\begin{itemize}
\item To the best of our knowledge, the MMSN protocol is the first
multifrequency MAC protocol especially designed for WSNs, in which
each device is equipped with a single radio transceiver and
the MAC layer packet size is very small.
\item Instead of using pairwise RTS/CTS frequency negotiation
[Adya 2001,Culler 2001; Tzamaloukas 2001; Zhou 2006],
we propose lightweight frequency assignments, which are good choices
for many deployed comparatively static WSNs.
\item We develop new toggle transmission and snooping techniques to
enable a single radio transceiver in a sensor device to achieve
scalable performance, avoiding the nonscalable ``one
control channel + multiple data channels'' design [Natarajan 2001].
\end{itemize}

% Head 1
\section{MMSN Protocol}

% Head 2
\subsection{Frequency Assignment}

We propose a suboptimal distribution to be used by each node, which is
easy to compute and does not depend on the number of competing
nodes. A natural candidate is an increasing geometric sequence, in
which
% Numbered Equation
\begin{equation}
\label{eqn:01}
P(t)=\frac{b^{\frac{t+1}{T+1}}-b^{\frac{t}{T+1}}}{b-1},
\end{equation}
where $t=0,{\ldots}\,,T$, and $b$ is a number greater than $1$.

In our algorithm, we use the suboptimal approach for simplicity and
generality. We need to make the distribution of the selected back-off
time slice at each node conform to what is shown in Equation
(\ref{eqn:01}). It is implemented as follows: First, a random
variable $\alpha$ with a uniform distribution within the interval
$(0, 1)$ is generated on each node, then time slice $i$ is selected
according to the following equation:
% Unnumbered Equation
\[
i=\lfloor(T+1)\log_b[\alpha(b-1)+1]\rfloor.
\]
It can be easily proven that the distribution of $i$ conforms to Equation
(\ref{eqn:01}).

So protocols [Bahl 2002,Culler 2001,Zhou 2006,Adya 2001,Culler 2001;
Tzamaloukas-01; Akyildiz-01] that use RTS/CTS
controls\footnote{RTS/CTS controls are required to be implemented by
802.11-compliant devices. They can be used as an optional mechanism
to avoid Hidden Terminal Problems in the 802.11 standard and
protocols based on those similar to [Akyildiz 2001] and
[Adya 2001].} for frequency negotiation and reservation are not
suitable for WSN applications, even though they exhibit good
performance in general wireless ad hoc
networks.

% Head 3
\subsubsection{Exclusive Frequency Assignment}

In exclusive frequency assignment, nodes first exchange their IDs
among two communication hops so that each node knows its two-hop
neighbors' IDs. In the second broadcast, each node beacons all
neighbors' IDs it has collected during the first broadcast period.

% Head 4
\paragraph{Eavesdropping}

Even though the even selection scheme leads to even sharing of
available frequencies among any two-hop neighborhood, it involves a
number of two-hop broadcasts. To reduce the communication cost, we
propose a lightweight eavesdropping scheme.

\subsection{Basic Notations}

As Algorithm~\ref{alg:one} states, for each frequency
number, each node calculates a random number (${\textit{Rnd}}_{\alpha}$) for
itself and a random number (${\textit{Rnd}}_{\beta}$) for each of its two-hop
neighbors with the same pseudorandom number generator.
% Algorithm
\begin{algorithm}[t]
\SetAlgoNoLine
\KwIn{Node $\alpha$'s ID ($ID_{\alpha}$), and node $\alpha$'s
neighbors' IDs within two communication hops.}
\KwOut{The frequency number ($FreNum_{\alpha}$) node $\alpha$ gets assigned.}
$index$ = 0; $FreNum_{\alpha}$ = -1\;
\Repeat{$FreNum_{\alpha} > -1$}{
        $Rnd_{\alpha}$ = Random($ID_{\alpha}$, $index$)\;
        $Found$ = $TRUE$\;
        \For{each node $\beta$ in $\alpha$'s two communication hops
    }{
      $Rnd_{\beta}$ = Random($ID_{\beta}$, $index$)\;
      \If{($Rnd_{\alpha} < Rnd_{\beta}$) \text{or} ($Rnd_{\alpha}$ ==
          $Rnd_{\beta}$ \text{and} $ID_{\alpha} < ID_{\beta}$)\;
      }{
        $Found$ = $FALSE$; break\;
      }
        }
     \eIf{$Found$}{
           $FreNum_{\alpha}$ = $index$\;
         }{
           $index$ ++\;
     }
      }
\caption{Frequency Number Computation}
\label{alg:one}
\end{algorithm}

Bus masters are divided into two disjoint sets, $\mathcal{M}_{RT}$
and $\mathcal{M}_{NRT}$.
% description
\begin{description}
\item[RT Masters]
$\mathcal{M}_{RT}=\{ \vec{m}_{1},\dots,\vec{m}_{n}\}$ denotes the
$n$ RT masters issuing real-time constrained requests. To model the
current request issued by an $\vec{m}_{i}$ in $\mathcal{M}_{RT}$,
three parameters---the recurrence time $(r_i)$, the service cycle
$(c_i)$, and the relative deadline $(d_i)$---are used, with their
relationships.
\item[NRT Masters]
$\mathcal{M}_{NRT}=\{ \vec{m}_{n+1},\dots,\vec{m}_{n+m}\}$ is a set
of $m$ masters issuing nonreal-time constrained requests. In our
model, each $\vec{m}_{j}$ in $\mathcal{M}_{NRT}$ needs only one
parameter, the service cycle, to model the current request it
issues.
\end{description}

Here, a question may arise, since each node has a global ID. Why
don't we just map nodes' IDs within two hops into a group of
frequency numbers and assign those numbers to all nodes within two
hops?

\section{Simulator}
\label{sec:sim}

If the model checker requests successors of a state which are not
created yet, the state space uses the simulator to create the
successors on-the-fly. To create successor states the simulator
conducts the following steps.
% enumerate
\begin{enumerate}
\item Load state into microcontroller model.
\item Determine assignments needed for resolving nondeterminism.
\item For each assignment.
      \begin{enumerate}
      \item either call interrupt handler or simulate effect of next instruction, or
      \item evaluate truth values of atomic propositions.
      \end{enumerate}
\item Return resulting states.
\end{enumerate}
Figure~\ref{fig:one} shows a typical microcontroller C program that
controls an automotive power window lift. The program is one of the
programs used in the case study described in Section~\ref{sec:sim}.
At first sight, the programs looks like an ANSI~C program. It
contains function calls, assignments, if clauses, and while loops.
% Figure
\begin{figure}
\centerline{\includegraphics{acmsmall-mouse}}
\caption{Code before preprocessing.}
\label{fig:one}
\end{figure}

\subsection{Problem Formulation}

The objective of variable coalescence-based offset assignment is to find
both the coalescence scheme and the MWPC on the coalesced graph. We start
with a few definitions and lemmas for variable coalescence.

% Enunciations
\begin{definition}[Coalesced Node (C-Node)]A C-node is a set of
live ranges (webs) in the AG or IG that are coalesced. Nodes within the same
C-node cannot interfere with each other on the IG. Before any coalescing is
done, each live range is a C-node by itself.
\end{definition}

\begin{definition}[C-AG (Coalesced Access Graph)]The C-AG is the access
graph after node coalescence, which is composed of all C-nodes and C-edges.
\end{definition}

\begin{lemma}
The C-MWPC problem is NP-complete.
\end{lemma}
\begin{proof} C-MWPC can be easily reduced to the MWPC problem assuming a
coalescence graph without any edge or a fully connected interference graph.
Therefore, each C-node is an uncoalesced live range after value separation
and C-PC is equivalent to PC. A fully connected interference graph is made
possible when all live ranges interfere with each other. Thus, the C-MWPC
problem is NP-complete.
\end{proof}

\begin{lemma}[Lemma Subhead]The solution to the C-MWPC problem is no
worse than the solution to the MWPC.
\end{lemma}
\begin{proof}
Simply, any solution to the MWPC is also a solution to the
C-MWPC. But some solutions to C-MWPC may not apply to the MWPC (if any
coalescing were made).
\end{proof}

\section{Performance Evaluation}

During all the experiments, the Geographic Forwarding (GF)
[Akyildiz 2001] routing protocol is used. GF exploits geographic
information of nodes and conducts local data-forwarding to achieve
end-to-end routing. Our simulation is
configured according to the settings in
Table~\ref{tab:one}. Each run lasts for 2 minutes and
repeated 100 times. For each data value we present in the results,
we also give its 90\% confidence interval.
% Table
\begin{table}%
\tbl{Simulation Configuration\label{tab:one}}{%
\begin{tabular}{|l|l|}
\hline
TERRAIN{$^a$}   & (200m$\times$200m) Square\\\hline
Node Number     & 289\\\hline
Node Placement  & Uniform\\\hline
Application     & Many-to-Many/Gossip CBR Streams\\\hline
Payload Size    & 32 bytes\\\hline
Routing Layer   & GF\\\hline
MAC Layer       & CSMA/MMSN\\\hline
Radio Layer     & RADIO-ACCNOISE\\\hline
Radio Bandwidth & 250Kbps\\\hline
Radio Range     & 20m--45m\\\hline
\end{tabular}}
\begin{tabnote}%
\Note{Source:}{This is a table
sourcenote. This is a table sourcenote. This is a table
sourcenote.}
\vskip2pt
\Note{Note:}{This is a table footnote.}
\tabnoteentry{$^a$}{This is a table footnote. This is a
table footnote. This is a table footnote.}
\end{tabnote}%
\end{table}%

\section{Conclusions}

In this article, we develop the first multifrequency MAC protocol for
WSN applications in which each device adopts a
single radio transceiver. The different MAC design requirements for
WSNs and general wireless ad-hoc networks are
compared, and a complete WSN multifrequency MAC design (MMSN) is
put forth. During the MMSN design, we analyze and evaluate different
choices for frequency assignments and also discuss the nonuniform
back-off algorithms for the slotted media access design.

% Start of "Sample References" section

\section{Typical references in new ACM Reference Format}
A paginated journal article \cite{Abril07}, an enumerated
journal article \cite{Cohen07}, a reference to an entire issue \cite{JCohen96},
a monograph (whole book) \cite{Kosiur01}, a monograph/whole book in a series (see 2a in spec. document)
\cite{Harel79}, a divisible-book such as an anthology or compilation \cite{Editor00}
followed by the same example, however we only output the series if the volume number is given
\cite{Editor00a} (so Editor00a's series should NOT be present since it has no vol. no.),
a chapter in a divisible book \cite{Spector90}, a chapter in a divisible book
in a series \cite{Douglass98}, a multi-volume work as book \cite{Knuth97},
an article in a proceedings (of a conference, symposium, workshop for example)
(paginated proceedings article) \cite{Andler79}, a proceedings article
with all possible elements \cite{Smith10}, an example of an enumerated
proceedings article \cite{VanGundy07},
an informally published work \cite{Harel78}, a doctoral dissertation \cite{Clarkson85},
a master's thesis: \cite{anisi03}, an online document / world wide web resource \cite{Thornburg01}, \cite{Ablamowicz07},
\cite{Poker06}, a video game (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03}
and \cite{Lee05} and (Case 3) a patent \cite{JoeScientist001},
work accepted for publication \cite{rous08}, 'YYYYb'-test for prolific author
\cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might contain
'duplicate' DOI and URLs (some SIAM articles) \cite{Kirschmer:2010:AEI:1958016.1958018}.
Boris / Barbara Beeton: multi-volume works as books
\cite{MR781536} and \cite{MR781537}.

% Appendix
\appendix
\section*{APPENDIX}
\setcounter{section}{1}
In this appendix, we measure
the channel switching time of Micaz [CROSSBOW] sensor devices.
In our experiments, one mote alternatingly switches between Channels
11 and 12. Every time after the node switches to a channel, it sends
out a packet immediately and then changes to a new channel as soon
as the transmission is finished. We measure the
number of packets the test mote can send in 10 seconds, denoted as
$N_{1}$. In contrast, we also measure the same value of the test
mote without switching channels, denoted as $N_{2}$. We calculate
the channel-switching time $s$ as
\begin{eqnarray}%
s=\frac{10}{N_{1}}-\frac{10}{N_{2}}. \nonumber
\end{eqnarray}%
By repeating the experiments 100 times, we get the average
channel-switching time of Micaz motes: 24.3$\mu$s.

\appendixhead{ZHOU}

% Acknowledgments
\begin{acks}
The authors would like to thank Dr. Maura Turolla of Telecom
Italia for providing specifications about the application scenario.
\end{acks}

% Bibliography
\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{acmsmall-sample-bibfile}

\end{document}

### 2014-04-21 WriteUp/maxentpredmkts.tex
switched to camera-ready style file

@@
-% Adjusted to add \usepackage[numbers]{natbib}
-% \bibliographystyle{acmsmall} and \documentclass[prodmode,acmec]{ec-acmsmall}
-% Jan 5, 2013 - David Parkes
-%
-% v2-acmsmall-sample.tex, dated March 6 2012
-% This is a sample file for ACM small trim journals
-%
-% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
-% (c) 2010 Association for Computing Machinery (ACM)
-%
-% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
-% Users can also go through the FAQs available on the journal's submission webpage.
-%
-% Steps to compile: latex, bibtex, latex latex
-%
-% For tracking purposes => this is v1.3 - March 2012
-
-\documentclass[prodmode,acmec]{ec-acmsmall} % Aptara syntax
+%\documentclass[prodmode]{acmsmall-ec14}
+\documentclass[prodmode,license]{acmsmall-ec14}
+%\documentclass[prodmode,permissions]{acmsmall-ec14}
 
@@
 % Metadata Information
-\acmVolume{X}
-\acmNumber{X}
-\acmArticle{X}
-\acmYear{2014}
-\acmMonth{2}
+%\acmVolume{X}
+%\acmNumber{X}
+%\acmArticle{X}
+%\acmYear{2014}
+%\acmMonth{2}
+
+\doi{2600057.2602896}
+
+\clubpenalty=10000 
+\widowpenalty = 10000
 
@@
 % Page heads
-\markboth{Abernethy et al.}{Information Aggregation in Exponential Family Markets}
+%\markboth{Abernethy et al.}{Information Aggregation in Exponential Family Markets}
 % Title portion
@@
 RAHUL SAMI
-\affil{University of Michigan, Ann Arbor\footnote{Current affiliation: Google, Inc. This work was done while the author was at the University of Michigan.}}}
+\affil{University of Michigan, Ann Arbor}}
 %%\author{GANG ZHOU
@@
 \begin{bottomstuff}
-%%This work is supported by the National Science Foundation, under
-%%grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.
-%%
-%%Author's addresses: G. Zhou, Computer Science Department,
-%%College of William and Mary; Y. Wu  {and} J. A. Stankovic,
-%%Computer Science Department, University of Virginia; T. Yan,
-%%Eaton Innovation Center; T. He, Computer Science Department,
-%%University of Minnesota; C. Huang, Google; T. F. Abdelzaher,
-%%(Current address) NASA Ames Research Center, Moffett Field, California 94035.
+Rahul Sami's current affiliation: Google, Inc. This work was done while the author was at the University of Michigan.
+Author's addresses: J.
+Abernethy and S.
+Kutty, Department of Electrical Engineering and Computer Science, University of Michigan, \url{{skutty, jabernet}@umich.edu}.
+S.
+Lahaie, Microsoft Research, New York City, \url{slahaie@microsoft.com}.
+Rahul Sami, Google Inc., \url{rahulsami@gmail.com}.
 \end{bottomstuff}
@@
 %----------------------------------------------------------------------------------------
-\bibliographystyle{acmsmall}
+%\bibliographystyle{acmsmall}
+
+\bibliographystyle{ACM-Reference-Format-Journals}
 \bibliography{maxentpredmkts}
+
 \end{document}
-
 %----------------------------------------------------------------------------------------

### 2014-04-21 WriteUp/maxentpredmkts.tex
took a pass and made small edits and typo fixes

@@
 For comprehensive introductions to these distributions, see~\citep{Barndorff78,WainJordan08}.
-Exponential families are intimately tied to the notions of log loss and entropy, but can be generalized to other types of convex losses and information, as shown by~\citet{GrunwaldDawid}, who also make a connection to scoring rules.
+Exponential families are intimately tied to the notions of log loss and entropy, but can be generalized to other types of convex losses and information as shown by~\citet{GrunwaldDawid}, who also make a connection to scoring rules.
 
@@
 % Strategic behavior in prediction markets is addressed in~\citep{chen2010gaming,dimitrov08}.
+The aggregation properties of market mechanisms have also been explored from a machine learning perspective in~\cite{storkey11}.
 
@@
 We see the potential to draw more on the mathematical finance literature to take into account risk aversion, as prediction markets are simply single-period financial markets~\citep[Part I]{follmerFinance}.
-We also note that connections between Machine Learning and market mechanisms have been explored in \cite{storkey11}.
 
@@
 
-When the outcome space is finite we recover classical scoring rules by using the statistic $\phi : \mX \rightarrow \{0,1\}^{\mX}$ that maps an outcome $x$ to a unit vector with a 1 in the component corresponding to $x$.
+%When the outcome space is finite we recover classical scoring rules by using the statistic $\phi : \mX \rightarrow \{0,1\}^{\mX}$ that maps an outcome $x$ to a unit vector with a 1 in the component corresponding to $x$.
 The expectation of $\phi$ is then exactly the probability mass function.
@@
 
-It is well-known that solutions to~(\ref{maxent-prog}) are \emph{exponential family} distributions (see, for instance, \cite{WainJordan08}).
+It is well-known that solutions to~(\ref{maxent-prog}) are \emph{exponential family} distributions; see, for instance \cite{WainJordan08}.
 These are distributions whose densities with respect to $\nu$ take the form
@@
 
-Some commonly studied distributions that are in fact exponential families include the binomial, beta and normal distributions.
-To see for instance, that the normal distribution is an exponential family note that instead of the usual parametrization using mean  $\mu$ and  variance $\sigma$, it can be parametrized in terms of $\vec{\theta}=(\theta_{1},\theta_{2})$ where $\theta_{1}=\frac{\mu}{\sigma^{2}}$ and $\theta_{2}=\frac{-1}{2\sigma^{2}}$,  with corresponding log partition function $-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$.
-
+Some commonly studied distributions that are in fact exponential families include the binomial, beta, Poisson, exponential, and normal distributions; the sufficient statistics and parametrizations for the exponential and normal will be covered in Section~\ref{sec:examples}.
+%To see for instance, that the normal distribution is an exponential family note that instead of the usual parametrization using mean  $\mu$ and  variance $\sigma$, it can be parametrized in terms of $\vec{\theta}=(\theta_{1},\theta_{2})$ where $\theta_{1}=\frac{\mu}{\sigma^{2}}$ and $\theta_{2}=\frac{-1}{2\sigma^{2}}$,  with corresponding log partition function $-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$.
+[SL: I think this is too much so removed it.]
+%
 The following proposition collects the relevant results on regular exponential families; proofs may be found in~\citet[Prop.\ 3.1--3.2, Thm.\ 3.3--3.4]{WainJordan08} and see also~\citet[Lem.\ 1, Thm.\ 2]{BanerjeeDhGh05b}.
@@
 Let $\tp \in \mP$ be a density (not necessarily from an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\hmu \in \mM$ be an alternative report.
-Then from~(\ref{exp-fam}) note
+Then from~(\ref{exp-fam}) we have
 %
@@
 It has been understood since~\citep{mccarthy1956measures,Savage71} that scoring rules for densities and expectations can be characterized in terms of an underlying convex function.
-The generalized log scoring rule of Theorem~\ref{maxent-score} obtains by applying the modern characterization of~\citet{AbernethyFr12} to the optimal value function $G$ (negative maximum entropy).
+The generalized log scoring rule of Theorem~\ref{maxent-score} obtains by applying the modern characterization of~\citet[Thm.\ 11]{AbernethyFr12} to the optimal value function $G$ (negative maximum entropy).
 \citet[Thm.\ 6]{BanerjeeDhGh05} prove a similar result by showing a bijection between exponential families and Bregman divergences.
@@
 \subsection{Examples: Moments over the Real Line}
+\label{sec:examples}
 
@@
 The interpretation is that the agent is only providing partial information about the underlying density, so the principal first infers a full density according to the principle of maximum entropy, and then scores the agent using the standard log score.
-We stress that the exponential families framework is a tool in the construction of scoring rules; we are not actually modeling any agent's belief via an exponential family.
+We stress that the exponential families framework is a tool in the construction of scoring rules at this point; we are not actually modeling any agent's belief via an exponential family.
 
@@
 The following examples illustrate the construction.
-Here, we have used $\mu$ and $\sigma$ for the true mean and variance and $\hmu$ and $\hsigma$ for the elicited mean and variance respectively (and $\hSigma$ for the variance of the multivariate normal).
-
-%
-\begin{example}
+Here, we have used $\mu$ and $\sigma$ for the true mean and variance and $\hmu$ and $\hsigma$ for the elicited mean and variance respectively (and $\hSigma$ for the covariance matrix of the multivariate normal).
+
+%
+\begin{example} \label{ex:multinomial}
 As a base measure we take the counting measure over the finite set $\{1,2,\ldots,d\}$, and we consider the statistic $\phi(i) \in \{0,1\}^d$ that maps outcome $i$ to the unit vector with component $i$ equal to 1.
@@
 
-As a concrete example, recall that in the classic finite-outcome case the statistic has a component for each outcome $x$ such that $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise.
+To be concrete, recall from Example~\ref{ex:multinomial} that in the classic finite-outcome case the statistic has a component for each outcome $x$ such that $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise.
 Therefore the corresponding security pays 1 dollar if outcome $x$ occurs.
@@
 
-There is a well-known duality between scoring rules and cost-function based markets~\citep{acv13,H03}. To see this in our context, recall from~(\ref{equalizer-rule}):
+There is a well-known duality between scoring rules and cost-function based markets~\citep{acv13,H03}. To see this in our context, recall from~(\ref{equalizer-rule}) that
 %
@@
 We consider a finite outcome space $\mX = \{1,2,\ldots,d\}$ and set up a market with an Arrow-Debreu security $\phi_i$ for every outcome $i$: the security pays off 1 if $i$ occurs and 0 otherwise.
-The log-partition function recovers the classical LMSR cost function~\citep{H07}:
+The log-partition function recovers the classical LMSR cost function:
 $C(\theta) = \log \sum_{i \in \mX} e^{\theta_i}.$
@@
 As before let the data distribution be given by $p(x;\theta)=\exp(\la\theta,\phi(x)\ra-T(\theta))$ where $T$ is the log partition function and $\phi$ are the sufficient statistics.
-Instead of direct beliefs about the data distribution the agent maintains a conjugate prior over the parameters $\theta$.
+Instead of direct beliefs about the data distribution, the agent maintains a conjugate prior over the parameters $\theta$.
 Every exponential family admits a conjugate prior of the form
 %
-$$p(\theta;b_{0})=\exp(\la n\nu,\theta\ra + nT(\theta) - \psi(\nu,n)).$$
-%
-Note that this is also an exponential family with natural parameter $b_0 = (n\nu,n)$ where $\nu \in \bR^d$ and $n$ is a positive integer.
+$$p(\theta;n\nu, n)=\exp(\la n\nu,\theta\ra + nT(\theta) - \psi(\nu,n)).$$
+%
+Note that this is also an exponential family with natural parameter $(n\nu,n)$ where $\nu \in \bR^d$ and $n$ is a positive integer.
 The sufficient statistic maps $\theta$ to $(\theta, T(\theta))$, and the log partition function $\psi$ is defined as the normalizer as usual.
-For a complete treatment of exponential families conjugate priors, see for instance~\citet{Barndorff78}.
+For a complete treatment of exponential families conjugate priors, see for instance~\citep{Barndorff78}.
 Now~\citet[Thm.\ 2]{diaconis79} and~\citet{jewell1974credible} have shown that
@@
 \begin{equation}\label{eq:prior}
-\E_{\theta\sim b_{0}}\E_{x\sim \theta}[\phi(x)]=\nu,
+\E_{\theta\sim (n\nu,n)}\E_{x\sim \theta}[\phi(x)]=\nu,
 \end{equation}
@@
 The market maker does not need to know the risk aversion parameters.
-Note that as, $a$ tends to 0, we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$.
+Note that as $a$ tends to 0 we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$.
 As $a$ grows larger (the agent grows more risk averse) the agent makes smaller trades to reduce it exposure, and the final state stays closer to the current state $\theta$.
@@
 
-We draw a well-known result from game theory regarding the class of \emph{potential games}.
+We draw on a well-known result from game theory regarding the class of \emph{potential games}.
 We say a function $f(\vec{x})$ is at a \emph{local optimum} if changing any coordinate of $\vec{x}$ does not increase the value of $f$.
@@
 Let $\vec{\delta}$ be the matrix of share vectors purchased by every trader in the market at equilibrium.
-Let $\theta$ be the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his risk aversion parameter.
+Let $\theta$ be the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution, and $a_{i}$ his risk aversion parameter.
 % \jake{I need to fix a small technical detail with the fact that we are using the log utilities to use the potential argument.}
@@
 Define a potential function as  $\Phi(\vec{\delta}) =  T\left(\theta+\sum_{i}\delta_{i}\right)+\sum_{i}\frac{1}{a_{i}}T(\htheta_{i}-a_{i}\delta_{i})$.
-
-Rather than working directly with the utilities of every trader, we will work with the log of their utility values.\footnote{It is important to note that the potential function analysis still applies for any monotonically increasing transformation of the traders' utility functions.} Now the log-utility of trader $i$ is $$U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i}).$$
+%
+Rather than working directly with the utilities of every trader, we will work with the log of their utility values.\footnote{It is important to note that the potential function analysis still applies for any monotonically increasing transformation of the traders' utility functions.} Now the log-utility of trader $i$ is
+%
+$$
+\textstyle
+U_{i}(\vec{\delta})=- T(\theta+\sum_{j}\delta_{j}) + T(\theta+\sum_{j\neq i}\delta_{j}) - \frac{1}{a_{i}} T(\htheta_{i} - a_{i}\delta_{i}) + \frac{1}{a_{i}} T(\htheta_{i}).
+$$
+%
 We can now apply Theorem \ref{thm:potl}, hence the equilibrium state is obtained by jointly maximizing $\Phi(\vec{\delta})$ for each $\delta_{i}$:
-\begin{eqnarray*}
-\nabla_{\delta_{i}}\Phi(\vec{\delta})&=&\nabla T\left(\theta+\sum_{j=1}^{n}\delta_{i}\right)-\nabla T(\htheta_{i}-a_{i}\delta_{i}) = 0.
-\end{eqnarray*}
+%
+$$
+\textstyle
+\nabla_{\delta_{i}}\Phi(\vec{\delta})= \nabla T\left(\theta+\sum_{j=1}^{n}\delta_{j}\right)-\nabla T(\htheta_{i}-a_{i}\delta_{i}) = 0.
+$$
+%
 This leads to the following expression for the final market state.
@@
 Then, the market runs again with the same set of traders and the outcome drawn from a (possibly different) exponential family distribution.
-The goal of this section is to analyze how the budget of traders evolve under these circumstances.
+The goal of this section is to analyze how the budgets of traders evolve under these circumstances.
 As before, the market price is interpreted as a probability density over the outcome space and the share vector as the natural parameter of an exponential family distribution.
@@
 We assume that these entities trade in multiple instances of the market; thus the exposure of the market maker is over several rounds.
-The log loss function for $\theta$ shares held is defined as $L(\theta,x)=-\log p(x;\theta)=C(\theta)-\la \theta\phi(x)\ra$.
+The log loss function for $\theta$ shares held is defined as $L(\theta,x)=-\log p(x;\theta)=C(\theta)-\la \theta,\phi(x)\ra$.
 
@@
 \end{eqnarray*}
-
-Where $\Delta_{i}^{t}$ is called the myopic impact of a trader $i$ in round $t$.
+%
+Here $\Delta_{i}^{t}$ is called the myopic impact of a trader $i$ in round $t$.
 Thus, the myopic impact captures incremental gain in prediction due to the trader in a round and is equal to the change in his budget in that round.
@@
 \end{proof}
-
+%
 An interesting aspect of Lemma~\ref{lem:damage} is that the log loss can be quantified in the same units as the traders' budgets.
@@
 \end{eqnarray*}¥
-A trader who adjusts the market state may expect positive profit $\geq \lambda D_{T}(\theta,\htheta)$.
+A trader who adjusts the market state may expect positive profit of at least $\lambda D_{T}(\theta,\htheta)$.
  \end{proof}
@@
 
-%\section{Conclusions}
-%\sindhu{Do we need this section?}
-%
+% Acknowledgments
+\begin{acks}
+We thank Rafael Frongillo and the anonymous referees for valuable feedback.
+\end{acks}
+
 %----------------------------------------------------------------------------------------

### 2014-04-21 WriteUp/maxentpredmkts.tex
ran a spelling check

@@
 recalling~(\ref{breg-div}).
-Now~\citet{BanerjeeDhGh05} have observed (among others) that the Kullback-Leibler divergence between two exponential family distributions is the Bregman divergence, with repect to the log-partition function, between their natural parameters.
+Now~\citet{BanerjeeDhGh05} have observed (among others) that the Kullback-Leibler divergence between two exponential family distributions is the Bregman divergence, with respect to the log-partition function, between their natural parameters.
 The agent's expected profit is therefore the KL divergence between the market's implied expectation and the exponential family corresponding to the agent's expectation, a well-known property from the classical LMSR~\citep{H07}.

### 2014-04-22 WriteUp/maxentpredmkts.tex
added references to Raf's work (from his email)

@@
 The aggregation properties of market mechanisms have also been explored from a machine learning perspective in~\cite{storkey11}.
-
-We are not aware of any works that consider risk-averse agents within cost-function based markets.
+More closely related to our work, \citet{FrongilloRe13} also touch upon the statistical interpretations of prediction market costs and payoffs from the viewpoint of exponential families.
+
+We are not aware of any works that consider risk-averse agents within cost-function based markets, with the exception of~\citep[Sec.\ 3.1]{FrongilloPeRe12} who assume agents are price takers.
 However, risk aversion is a fundamental component of mathematical finance and portfolio optimization, and there are close connections between the notion of a cost function and that of a convex risk measure~\citep{follmerRisk,follmer2011entropic}.

### 2014-04-24 WriteUp/maxentpredmkts.tex
down to 18 pages

@@
 Absent from much of the literature is a corresponding \emph{semantics} of the market behavior or the observed prices.
-That is, how can we interpret the equilibrium market state when we have a number of traders with diverse beliefs on the underlying state of the world?
-In what sense is the market an aggregation mechanism?
+That is, how can we interpret the equilibrium market state when we have a number of traders with diverse beliefs on the state of the world?
+How is the market an aggregation mechanism?
 Do price changes relate to our usual Bayesian notion of information incorporation via posterior updating?
@@
 In addition to showing the syntactic relationship between exponential families and prediction markets, we explore a number of rich semantic implications as well.
-In particular, we show that our formulation allows us to analyze the evolution of the market under various models of trader behavior:
+In particular, we show that our formulation allows us to analyze the evolution of the market under various trader models:
 \begin{itemize}
@@
 To generalize the rule to expected statistics rather than full densities, we consider a subset of densities $\mD \subseteq \mP$.
-If there is a bijection between the sets $\mD$ and $\mM$, then we say that $\mM$ parametrizes $\mD$ and write $p(\cdot\,;\mu)$ for the density mapping to $\mu$.
+If $\mD$ and $\mM$ are bijective, then we say that $\mM$ parametrizes $\mD$ and write $p(\cdot\,;\mu)$ for the density mapping to $\mu$.
 Given such a family parametrized by the relevant statistics, the generalized log scoring rule is then
@@
 The following observation is crucial.
-Let $\tp \in \mP$ be a density (not necessarily from an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\hmu \in \mM$ be an alternative report.
+Let $\tp \in \mP$ be any arbitrary density with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\hmu \in \mM$ be an alternative report.
 Then from~(\ref{exp-fam}) we have
@@
 %
-$$
+$
 S(\mu,x) = G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra
-$$
+$
 for some strictly convex function $G$. Let $T = G^*$, so that $\grad G = {\grad T}^{-1}$, and let $\theta = \grad G(\mu)$. Then the above can be written as
@@
 %
-Again, we stress that this scoring rule elicits the mean and variance of any density over the real numbers, not just those of a normal distribution.
+Again, we stress that this scoring rule elicits the mean and variance of any density over $\bR$, not just those of a normal distribution.
 The construction easily generalizes to a multi-dimensional outcome space by taking the log density of the multivariate normal:
@@
 Under this approach, each component~$i$ of the statistic $\phi$ is interpreted as the payoff function of a security; that is, a single share of security $i$ pays off $\phi_i(x)$ when $x \in \mX$ occurs.
-Thus if the portfolio of shares held by the agent is $\delta \in \bR^d$, where entry $\delta_i$ corresponds to the number of shares of security $i$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \delta, \phi(x) \ra$.
+Thus if the portfolio of shares held by the agent is $\delta \in \bR^d$, where entry $\delta_i$ corresponds to the number of shares of security $i$, then the payoff to the agent when $x$ occurs is the inner product $\la \delta, \phi(x) \ra$.
 
@@
 
-There is a well-known duality between scoring rules and cost-function based markets~\citep{acv13,H03}. To see this in our context, recall from~(\ref{equalizer-rule}) that
+There is a well-known duality between scoring rules and cost-function based markets~\citep{acv13,H03}. In our context, recall from~(\ref{equalizer-rule}) that
 %
@@
 %
-$$
+$
 \grad C(\theta + \delta) = \frac{n\nu + \hmu}{n+1}.
-$$
+$
 %
@@
 $$
-Suppose as before that the agent's belief over outcomes takes the form of a density $p$ with respect to base measure $\nu$.
+Suppose that the agent's belief over outcomes is a density $p$ with respect to base measure $\nu$.
 There is a close relationship between the log-partition function and the certainty equivalent under exponential utility~\citep[see][]{ben2007old}.
@@
 %
-Rather than working directly with the utilities of every trader, we will work with the log of their utility values.\footnote{It is important to note that the potential function analysis still applies for any monotonically increasing transformation of the traders' utility functions.} Now the log-utility of trader $i$ is
+Rather than working directly with the utilities of every trader, we will work with the log of their utility values.\footnote{Note: the potential function analysis still applies for any monotonically increasing transformation of the traders' utility functions.} Now the log-utility of trader $i$ is
 %
@@
 Suppose that with infinite budget, the trader would have moved the market state from $\theta$ to $\htheta$, where $\htheta$ represents his true belief.
-Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is, the trader's budget does not allow for purchasing enough shares to move the market state to his belief.
+Now suppose further that $\alpha<C(\htheta)-C(\theta)$; that is, the trader's budget disallows purchasing enough shares to move the market state to his belief.
 In this case, we want to budget-limit the trader's influence on the market state.
@@
 \begin{eqnarray*}
-\alpha_{i}^{t}-\alpha_{i}^{t-1}&=&C(\theta)-C(\theta')-(\theta-\theta')^{T}\phi(x^{t})\\
-&=&L(\theta,x^{t})-L(\theta',x^{t})=\Delta_{i}^{t}
+\alpha_{i}^{t}-\alpha_{i}^{t-1}&=&C(\theta)-C(\theta')-(\theta-\theta')^{T}\phi(x^{t})= L(\theta,x^{t})-L(\theta',x^{t})=\Delta_{i}^{t}
 \end{eqnarray*}
@@
 
-Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is $\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$.
+Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is 
+\[\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}\]
 \end{proof}
@@
 \begin{proof}
-Let the cost function $C$ be equal to the log partition function $T$ of the belief distribution.
-The payoff is given by the sufficient statistics $\phi(x)$.
-Then, the trader's expected net payoff is given by
+Let $C$ be the log partition function $T$ of the belief distribution. The payoff is given by the sufficient statistics $\phi(x)$. Then the trader's expected net payoff is given by
 \begin{eqnarray*}
@@
 \end{eqnarray*}¥
-A trader who adjusts the market state may expect positive profit of at least $\lambda D_{T}(\theta,\htheta)$.
+A trader who adjusts the market state may expect profit of at least $\lambda D_{T}(\theta,\htheta)$.
  \end{proof}
@@
 In this case, the trader who moves the market state can expect his profit to be positive and at least $\frac{1}{a}D_{T}(\theta,\htheta)$ where $a$ is the exponential utility parameter.
-When the cost function is adjusted to $C_{\lambda}$ with an inverse liquidity parameter $\lambda$ as in Section \ref{sec:exputil},  the trader receives an expected payoff of at least $\frac{1}{a} D_{T}(\lambda\theta,\htheta)$.
+When the cost function is adjusted to $C_{\lambda}$ with an inverse liquidity parameter $\lambda$ (Section \ref{sec:exputil}), the trader receives expected payoff of at least $\frac{1}{a} D_{T}(\lambda\theta,\htheta)$.
 
@@
 \begin{acks}
-We thank Rafael Frongillo and the anonymous referees for valuable feedback.
+We would like to sincerely thank Rafael Frongillo and the anonymous referees for valuable feedback.
 \end{acks}

### 2014-04-24 WriteUp/SubmittedVersion/final_submit.tex
final version

--- 2014-04-24 WriteUp/maxentpredmkts.tex
@@
 
-We are not aware of any works that consider risk-averse agents within cost-function based markets, with the exception of~\citep[Sec.\ 3.1]{FrongilloPeRe12} who assume agents are price takers.
+One of the aims of the present work is to understand the behavior of cost-function prediction markets with risk-averse agents, a subject which has received only limited attention; the only work we are aware of is \citep[Sec.\ 3.1]{FrongilloPeRe12}.
 However, risk aversion is a fundamental component of mathematical finance and portfolio optimization, and there are close connections between the notion of a cost function and that of a convex risk measure~\citep{follmerRisk,follmer2011entropic}.
 Indeed, they arise from the same axioms as noted by~\citet{othman2011liqu}.
-We see the potential to draw more on the mathematical finance literature to take into account risk aversion, as prediction markets are simply single-period financial markets~\citep[Part I]{follmerFinance}.
+We see the potential to draw more on the mathematical finance literature to take into account risk aversion, as prediction markets can be viewed simply as single-period financial markets~\citep[Part I]{follmerFinance}.
 
@@
 \begin{theorem}[\citet{monderer1996potential}]\label{thm:potl}
-Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}= (\delta_{1},\ldots,\delta_{i},\ldots,\delta_{n})$.
-If there exists a potential function $\Phi(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=\Phi(\vec{\delta})-\Phi(\vec{\delta}_{-i},\delta_{i}')$$
-then $\vec{\delta}$ is a Nash equilibrium if and only if $\Phi(\vec{\delta})$ is at a local optimum.
+Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given the strategy profile $\vec{\delta}= (\delta_{1},\ldots,\delta_{i},\ldots,\delta_{n})$.
+Assume there exists a potential function $\Phi(\vec{\delta})$ such that, for any $\vec \delta$ and $\delta_i'$, $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=\Phi(\vec{\delta})-\Phi(\vec{\delta}_{-i},\delta_{i}').$$
+Then $\vec{\delta}$ is a pure-strategy Nash equilibrium if and only if $\vec \delta$ is a local optimum for $\Phi(\cdot)$.
 \end{theorem}
@@
 We see that the equilibrium state is a convex combination of the initial market state and all agent beliefs, with the latter weighted according to risk tolerance.
+
+We finally note that we have not considered the dynamics by which such an equilibrium would be reached, nor the rate of convergence etc., yet we think such questions provide fruitful directions for future research.
 
